### 2020-12-03 computational model/TE_computation.Rmd
computational model

---
title: "TE_computation"
date: "Last update: Dec 2nd 2020"
output: html_document
---

# Preparations
## load libraries
```{r setup, include=FALSE}

library(here)
library(tidyverse)
library(tidylog)
library(regclass)

`%notin%` <- Negate(`%in%`)
```

# Generating dataframe for computational model
```{r}
#### Set 1: Calculate other variables based on V, W, and B ####

## Creating variables: V, W, and B
# Items on vocabulary measure (e.g. words on the CDI)
V <- 600 # This is constant

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 50) %>% # a sequence of 0 to 2V
  rep(., 6) # generate 6 repetitinos of the sequence

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.1) %>% # a sequence of 0.5 to 1
  rep(., 25) # generate 25 repetitions of the sequence

## Creating the dataframe
TE_fromVWB <- data.frame(V, W) %>%
  arrange(W) %>%
  mutate(B = B) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= 600))

#### Set 2: Calculate other variables based on V, D, and N ####

## Creating variables: D and N
D <- seq(0, V, by = 50) %>% # Words produced in the dominant language
  rep(., 13) # generate 13 repetitions of the sequence

N <- seq(0, V, by =50) %>% # Words produced in non-dominant language
  rep(., 13) # generate 13 repetitions of the sequence

## Creating the dataframe
TE_fromVDN <- data.frame(V, D) %>%
  arrange(D) %>%
  mutate(N = N) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create other variables
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>%
  # Remove cases where B = NA
  filter(!is.na(B))

#### Combine the 2 sets ####
TE_computation <- TE_fromVWB %>%
  bind_rows(TE_fromVDN)

```

# Correlation
```{r}
Correlation <- TE_computation %>%
  select(-V) %>%
  cor_matrix()

#write.table(Correlation, "TEcomputation_correlations.tsv", col.names=NA)
```

# Plotting relations between variables
## Stacked barchart showing the vocabulary composition across different balance proportions
```{r}

#plot_composition <- TE_computation %>%
#  # Convert to long data
#  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
#  filter(vocab_type %notin% c("D", "N", "C")) %>%
#  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
#  mutate(bin_W = cut(W, breaks = 3, 
#                     labels = c("0-400", "401-800", "801-1200"))) %>%
#  group_by(bin_W, vocab_type, B) %>%
#  summarise(n_words = mean(number)) %>%
#  ggplot(aes(x=B, y = n_words, fill = vocab_type)) + 
#  geom_bar(position="stack", stat="identity") +
#  facet_grid(. ~ bin_W) +
#  #scale_y_continuous(limits = c(0, 1200)) +
#  theme_minimal() +
#  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
#  labs(x = "Balance", 
#       y = "Number of words")

plot_composition <- TE_computation %>%
  # Convert to long data
  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("D", "N", "C")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  group_by(bin_B, vocab_type, W) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x=W, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ bin_B) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  #scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
  labs(x = "Total vocabulary size", 
       y = "Number of words", 
       title = "Vocabulary composition across different balance proportions")

ggsave("plot_composition.png", plot_composition)
```

## Plotting relations between TEs and different variables
##### Relation between TE & W, D, & N #####
```{r}
### Plot: Y = TE, X = D / N
plot_TEandWDN <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = TE)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_TEandWDN.png", plot_TEandWDN)

### Plot: Y = TE, X = D / N, color by B
plot_TEandWDN_byB <- TE_computation %>%
  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = TE, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total number of TEs",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_TEandWDN_byB.png", plot_TEandWDN_byB)
```

##### Relation between W & TE, Du, & Nu #####
```{r}
### Plot: Y = W, X = TE/ Du / Nu
plot_WandTEDuNu <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal()

ggsave("plot_WandTEDuNu.png", plot_WandTEDuNu)

### Plot: Y = W, X = TE/ Du / Nu, color by B
plot_WandTEDuNu_byB <- TE_computation %>%
  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = W, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Total vocabulary size",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 1200)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_WandTEDuNu_byB.png", plot_WandTEDuNu_byB)
```

##### Relation between C & W, D, & N #####
```{r}
### Plot: Y = C, X = W / D / N
plot_CandWDN <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  #mutate(bin_B = cut(B, breaks = 6, 
  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal()

ggsave("plot_CandWDN.png", plot_CandWDN)

### Plot: Y = C, X = W / D / N, color by B
plot_CandWDN_byB <- TE_computation %>%
  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %in% c("W", "D", "N")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
  mutate(bin_B = cut(B, breaks = 6, 
                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
  ggplot(aes(x = number, y = C, colour = bin_B)) +
  facet_grid(. ~vocab_type) +
  labs(x = "Words produced", y = "Number of concepts",
       colour = "Bins of balance") +
  geom_point(shape = 23) +
  geom_smooth(method="lm", se=F) +
  scale_x_continuous(limits = c(0, 1200)) +
  scale_y_continuous(limits = c(0, 600)) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave("plot_CandWDN_byB.png", plot_CandWDN_byB)
```

### 2020-12-07 Analyze_TEs.Rmd
Data (n = 230; unique n = 201)

---
title: "acquisition of TEs"
date: "Last update: Nov 27th 2020"
output: 
  html_document: 
    code_folding: show
    collapsed: no
    df_print: kable
    highlight: espresso
    number_sections: yes
    theme: spacelab
    toc: yes
    toc_float: yes
editor_options: 
  chunk_output_type: console
---
# Preparations
## load libraries
```{r}

# install.packages("devtools")
#devtools::install_github("mikabr/ggpirate")

library(ggpirate)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)

library(lme4)
library(lmerTest)
library(scales)
```

# load data
```{r}
cdi_ws = readRDS(here::here("Data","output_analyzeSent.Rdata"))
```

# Participant Exclusions
exclude babies out of 18-33m age range, monolinguals and those with >10% 3rd language exposure
```{r}
# exclude kids that are bilingual at one visit, monolingual at another (regardless of direction)
# switch from other to bilingual, or vice versa, or from other to monolingual and vice versa are ok
exclude.lang_group_change  = cdi_ws %>%
  filter(lang_group !="other") %>%
  group_by(baby_id) %>% 
  distinct(ID_testdate, .keep_all = T) %>%
  mutate(visit_num = 1:n()) %>%  
  mutate(changed_language_group_status = if_else(visit_num == 1, NA, 
                                                 lang_group != lag(lang_group))) %>%
  # uncomment see all visits for babies that have one visit excluded
  #filter(any(changed_language_group_status == TRUE) ) %>% View()
  filter(changed_language_group_status == TRUE) %>% 
  pull(ID_testdate)

# Create exclusion criteria variables
cdi_ws_TE <- cdi_ws %>%
  mutate(
    # create an exclusion criteria for language
    exclude.language = case_when(
      # exclude babies with more than 10% exposure to a 3rd language
      lang_exp_other > 10 ~ 1,
      # exclude language group status changed from bilingual to mono or vice versa
      ID_testdate %in% exclude.lang_group_change ~ 1, 
      lang_group == "bilingual" ~ 0, #bilingual
      lang_group == "monolingual" ~ 0, # monolingual
      TRUE ~ 1), # all other cases exlude

    # create an exclusion criteria for age
    exclude.age = case_when(!between(age_months_binned, 18, 33) ~ 1, 
                            TRUE ~ 0) # all other cases keep
  )

# Exclude participants
keepers_ws_TE <- cdi_ws_TE %>%
  filter(
    # exclude participants who have exposure to a third language/ are neither bilingual nor monolingual
    exclude.language == 0,
    # exclude participants who don't fit age range
    exclude.age == 0,
    # exclude participants who don't have both CDI filled
    both_cdi_filled == "Y",
    # keep only babies with at least 15% of E and 15% of F (i.e., exclude monolinguals)
    lang_exp_eng >= 15 & lang_exp_eng <= 85,
    lang_exp_fre >= 15 & lang_exp_fre <= 85)

### Check how many distinct participants and data point we have
keepers_ws_TE %>% summarize(N_visits = n_distinct(ID_testdate),
                            N_babies = n_distinct(baby_id),
                            N_admin = sum(n_administrations, na.rm=T))

### Check if there is any NAs in the vocabulary measures (should all return 0)
keepers_ws_TE %>% filter(is.na(vocab_score)) %>% 
  select(ID_testdate, vocab_type, vocab_score) %>%
  nrow()

keepers_ws_TE %>% filter(is.na(eng_unique_words)) %>% 
  select(ID_testdate) %>% 
  nrow()

keepers_ws_TE %>% filter(is.na(fre_unique_words)) %>% 
  select(ID_testdate) %>%
  nrow()

keepers_ws_TE %>% filter(is.na(number_of_te)) %>% 
  select(ID_testdate) %>%
  nrow()
```

# Prepare the dataset
```{r}

keepers_ws_TE = keepers_ws_TE %>%
  # keep only necessary variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% #n_administrations
  spread(vocab_type, vocab_score) %>%

  ######
  rename(baby_id = Baby_ID, age_months_binned = age.months,
         word_vocab = Totalvocabulary,
         total_words_eng = TotalNW_Eng, total_words_fre = TotalNW_Fr,
         eng_unique_words = Eng_unique_words, fre_unique_words = Fr_unique_words,
         number_of_te = TotalnumberTEs,
         lang_exp_eng = LangExpEng, lang_exp_fre = LangExpFr, lang_exp_other = LangExpOther) %>%

  # calculate age in months (with decimal)
  mutate(age_month_decimal = age_days/30) %>%
  # create balance groups (by exposure)
  mutate(BalanceGroup_byExposure = cut(balance, breaks = 4, 
                                       labels = c("80_20", "70_30", "60_40", "50_50"))) %>%
  mutate(TotalNW_Dom_exposure = case_when(lang_dom == "English" ~ total_words_eng,
                                          lang_dom == "French" ~ total_words_fre),
         TotalNW_NonDom_exposure = case_when(lang_dom == "English" ~ total_words_fre,
                                             lang_dom == "French" ~ total_words_eng),
         Dom_unique_words_exposure = case_when(lang_dom == "French" ~ fre_unique_words, 
                                               lang_dom == "English" ~ eng_unique_words),
         NonDom_unique_words_exposure = case_when(lang_dom == "French" ~ eng_unique_words, 
                                         lang_dom == "English" ~ fre_unique_words)) %>%

  #defining dominance based on vocabulary size
  mutate(LangDom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  mutate(LangDom_vocab = as.factor(LangDom_vocab)) %>%
  mutate(TotalNW_Dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                                 TRUE ~ total_words_fre),
         TotalNW_NonDom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                                    TRUE ~ total_words_eng),
         Dom_unique_words = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                      TRUE ~ fre_unique_words),
         NonDom_unique_words = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                         TRUE ~ eng_unique_words)) %>%
  mutate(TotalUnique = Dom_unique_words + NonDom_unique_words) %>%

  #Defining vocabulary-based balance score
  mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
  #Defining proportion of TE (with reference to child's own total vocab size)
  mutate(propTE = number_of_te*2/word_vocab) %>% 
  #Defining proportion of TE (with reference to CDI)
  mutate(propTE_CDI = number_of_te/672)

# check balancedness cutoffs
keepers_ws_TE %>%
  ggplot(aes(x = lang_exp_eng, y = BalanceGroup_byExposure)) + 
  geom_point()

keepers_ws_TE %>%
  ggplot(aes(x = lang_exp_fre, y = BalanceGroup_byExposure)) + 
  geom_point()

```

# Descriptives
## number of CDIs & Babys
```{r}
# how many unique babies
keepers_ws_TE %>%
  distinct(baby_id) %>%
  count()

keepers_ws_TE %>%
  distinct(participant_id) %>%
  count()

# some babies made multiple visits to the lab
keepers_ws_TE %>% 
  distinct(ID_testdate) %>% 
  count()

N_duplicatedBabyID <- keepers_ws_TE %>%
  filter(duplicated(baby_id))

# Create dataframe of demographics
d_participants <- keepers_ws_TE %>%
  filter(vocab_type == "word_vocab") %>%
  select(baby_id, age_days, gender, 
         lang_exp_eng, lang_exp_fre, lang_exp_other, years_education) %>%
  mutate(Female = ifelse(gender == "F", 1, 0)) %>%
  summarise(n = n(),
            age.days = mean(age_days, na.rm = TRUE),
            lang_exp_eng = mean(lang_exp_eng, na.rm = TRUE),
            lang_exp_fre = mean(lang_exp_fre, na.rm = TRUE), 
            lang_exp_other = mean(lang_exp_other, na.rm = TRUE),
            years_education = mean(years_education, na.rm = TRUE),
            number_female = sum(Female))
```

## Gender 
```{r}
keepers_ws_TE %>%
  group_by(Gender) %>% 
  distinct(baby_id, .keep.all=T) %>%
  count()
```

## English dominant
```{r}
# dominance defined by language exposure
keepers_ws_TE %>% group_by(lang_dom) %>% distinct(baby_id) %>%
  count()

# dominance defined by vocab size
keepers_ws_TE %>% group_by(LangDom_vocab) %>% distinct(baby_id) %>%
  count()
```

## years of maternal education
```{r}
keepers_ws_TE %>% distinct(baby_id, .keep_all = T) %>%
  summarize(mean = mean(years_education, na.rm=T), 
            sd = sd(years_education, na.rm=T))
```

## age range, mean
```{r}
keepers_ws_TE %>%
  distinct(baby_id, .keep_all = T) %>% 
  summarize(sd = sd(age_days, na.rm=T))

do.call(cbind, lapply(keepers_ws_TE %>%
                        distinct(baby_ID, .keep_all = T) %>% 
                        select(age_days,age_continuous),
                      summary))

```

## average # of TEs
```{r}
keepers_ws_TE %>%
  filter(bothCDI_filled=="Y") %>%
  summarise(mean = mean(propTE, na.rm=T), 
            min= min(propTE, na.rm=T),
            max= max(propTE, na.rm=T))

```

## Percentile information from WordbankR
```{r}

## English Word&Sentence: Vocabulary size at 90% percentile
english_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(EngWS_90percentile = production)

## Canadian French Word&Sentence: Vocabulary size at 90% percentile
french_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(FrWS_90percentile = production)

## Merge the two percentile dataframe
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Total_90percentile = EngWS_90percentile + FrWS_90percentile) %>%
  # keep only necessary columns
  select(age, EngWS_90percentile, FrWS_90percentile, Total_90percentile) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)
```

# Adding percentile info to keepers_ws_TE dataframe & 
## Creating a new propTE based on the percentile
```{r}
# merge the dataframes
keepers_ws_TE <- keepers_ws_TE %>%
  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
  # this means the percentile should be the same for all >30m
  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
  # merge with the percentile data frame
  left_join(ws_percentile, by = "age_months_percentile") %>%
  # calculate TE in relation to CDI percentile data
  mutate(propTE_percentile = number_of_te*2/Total_90percentile)
  # %>% select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
  #       propTE, propTE_CDI, propTE_percentile)

# check if any baby learn more words than the percentile
# (there are 2 babies know more than the percentile in their dominant language)
MoreThanPercentile <- keepers_ws_TE %>%
  select(ID_testdate, age_months_binned, lang_dom, LangDom_vocab, 
         word_vocab, total_words_eng, EngWS_90percentile, 
         total_words_fre, FrWS_90percentile, word_vocab, Total_90percentile) %>%
  filter(total_words_eng > EngWS_90percentile | 
           total_words_fre > FrWS_90percentile |
           word_vocab > Total_90percentile)
```

## Create predicted TEs for simulation
```{r}
keepers_ws_TE_simulation <- keepers_ws_TE %>%
  # Predicted propTE (Simulation)
  mutate(propDomVocab = TotalNW_Dom / 672, # Proportion of Dom/NonDom vocab
         propNonDomVocab = TotalNW_NonDom / 672, 
         Predicted_propTE = propDomVocab * propNonDomVocab, # Predicted proportion of TE
         Diff = propTE - Predicted_propTE) %>% # Difference between Predicted and Observed propTE
  # Predicted raw TE (with reference to CDI)
  mutate(Predicted_TotalnumberTEs_CDI = (TotalNW_Dom * TotalNW_NonDom)/672) %>%
  # Predicted raw TE (with reference to CDI percentile)
  mutate(Predicted_TotalnumberTEs_percentile = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)) %>%
  # Create two groups based on observed TE
  mutate(TEgroup = ifelse(number_of_te < 100, "Observed TE less than 100", "Observed TE more than 100"))

  #select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
  #       propTE, propTE_CDI, propTE_percentile, Predicted_propTE, Diff, 
  #       Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile)

```

###### Linear regressions of predicted TE models ###### 
```{r}

### Model 1: Predicted_TotalnumberTEs_percentile
Prediction_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te*(Total_90percentile/2) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

### Model 2: Predicted_TotalnumberTEs_CDI
CDI_model <- keepers_ws_TE_simulation %>%
  lm(number_of_te * 672 ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
     data = .)

anova(Percentile_model, CDI_model)
```

###### Theoretical distribution of TEs (Simulation) ###### 
```{r}
###Prediction 1: Predicted raw TE = (TotalNW_Dom*TotalNW_NonDom)/Total CDI
# Ploting Y = observed rawTE  & X = Predicted raw TE

keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_CDI, y = number_of_te)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs (based on CDI)", y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430)

# Kolmogorov-Smirnov test to compare the observed and predicted TE
stats::ks.test(keepers_ws_TE_simulation$number_of_te,
               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_CDI)

###Prediction 2: Predicted raw TE = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)
# Ploting Y = observed rawTE  & X = Predicted raw TE

TEprediction_percentile <- keepers_ws_TE_simulation %>%
  ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs (based on percentile)", y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,430) +
  ylim(-10,430)

ggsave(here("TEprediction_percentile.png"), TEprediction_percentile)

# Kolmogorov-Smirnov test to compare the observed and predicted TE
stats::ks.test(keepers_ws_TE_simulation$number_of_te,
               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_percentile)

keepers_ws_TE_simulation_LessThan10 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE less than 100")

stats::ks.test(keepers_ws_TE_simulation_LessThan10$number_of_te,
               keepers_ws_TE_simulation_LessThan10$Predicted_TotalnumberTEs_percentile)

keepers_ws_TE_simulation_MoreThan10 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE more than 100")

stats::ks.test(keepers_ws_TE_simulation_MoreThan10$number_of_te,
               keepers_ws_TE_simulation_MoreThan10$Predicted_TotalnumberTEs_percentile)

### Plotting the 2 predictions together

TEprediction_percentile_CDI <- keepers_ws_TE_simulation %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile, TEgroup) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te, TEgroup), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  facet_grid(. ~TEgroup) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
  guides(colour = guide_legend(nrow = 1))

ggsave(here("TEprediction_percentile&CDI.png"), TEprediction_percentile_CDI)

#Less Than 100 observed TEs
TEprediction_percentile_CDI_LessThan100 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE less than 100") %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs \n (Less than 100 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,100) +
  ylim(-10,100) +
  theme(plot.title = element_text(size = 12, face = "bold"), 
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.text=element_text(size = 10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("TEprediction_percentile&CDI_LessThan100.png"), TEprediction_percentile_CDI_LessThan100)

#Less Than 50 observed TEs
TEprediction_percentile_CDI_LessThan50 <- keepers_ws_TE_simulation %>%
  filter(number_of_te < 50) %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = 
       "Observed vs. Predicted total number of TEs \n (Less than 50 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,50) +
  ylim(-10,50) +
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.box="vertical", legend.text=element_text(size = 10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("TEprediction_percentile&CDI_LessThan50.png"), TEprediction_percentile_CDI_LessThan50)

# More than 100 Observed TEs
TEprediction_percentile_CDI_MoreThan100 <- keepers_ws_TE_simulation %>%
  filter(TEgroup == "Observed TE more than 100") %>%
  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
  pivot_longer(-c(baby_id, number_of_te), 
               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
  labs(x = "Predicted total number of TEs", 
       y = "Observed total number of TEs",
       title = "Observed vs. Predicted total number of TEs \n (More than 100 observed TEs)") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_smooth(method = "lm", se = F) +
  geom_point(shape = 23) +
  geom_abline(linetype = 2, alpha = 0.5) +
  xlim(-10,500) +
  ylim(-10,500) +
  theme(plot.title = element_text(size = 12, face = "bold"),
        axis.text.x = element_text(angle = 90), 
        legend.position = "bottom", legend.text=element_text(size=10)) +
  guides(colour = guide_legend(nrow = 2))

ggsave(here("TEprediction_percentile&CDI_MoreThan100.png"), TEprediction_percentile_CDI_MoreThan100)

###Prediction 2: Predicted raw TE = TotalNW_NonDom
# Ploting Y = observed rawTE  & X = Predicted raw TE

#keepers_ws_TE_simulation%>%
#  ggplot(aes(x = TotalNW_NonDom, y = number_of_te)) +
#  labs(x = "Predicted total number of TEs", 
#       y = "Observed total number of TEs",
#       title = 
#       "Observed vs. Predicted total number of TEs 
#       (based on total number of NonDom vocab)") +
#  theme(plot.title = element_text(hjust = 0.5)) +
#  geom_smooth(method = "lm", se = F) +
#  geom_point(shape = 23) +
#  geom_abline(linetype = 2, alpha = 0.5) +
#  xlim(-10,500) +
#  ylim(-10,500)

## Kolmogorov-Smirnov test to compare the observed and predicted TE
#stats::ks.test(keepers_ws_TE_simulation$TotalnumberTEs, keepers_ws_TE_simulation$TotalNW_NonDom)

```

# plots
## proportion of TEs
```{r}
keepers_ws_TE %>%
  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
  ggplot(aes(x = age.days/30, 
             y = propTE)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2))+
  ylab("Percentage of Translation Equivalents") +
  xlab("Age in Months")+
  ggtitle("Percentage of Translation Equivalents by Age") +  
  theme_classic(base_size = 22) +
  scale_x_continuous(breaks= seq(18,32,by = 2))+
  geom_point(size=3) + 
  geom_smooth(method="lm" )

ggsave(here("Output","PlotbyAge.png"), height=10,width=10)
```

## Bar graphs with individual scores (percentage TEs by dominance)
```{r}
keepers_ws_summarize = keepers_ws_TE %>% 
  group_by(balance_group15) %>%
  summarize(mean = mean(propTE, na.rm = T),
         N = n(),
         sd = sd(propTE, na.rm=T),
         se = sd(propTE, na.rm = T)/ sqrt(N))

ggplot(data=keepers_ws_summarize,
       aes(x=balance_group15,
           y=mean)) +
    geom_bar(aes(fill = balance_group15),
             show.legend =F,
             colour="black",
             stat="identity",
             position="dodge", width=.7) +

   geom_point(data = keepers_ws_TE,aes( y = propTE, fill = balance_group15), alpha = 0.9, show.legend = F, color = "black", pch = 21,size=1.75, 
   stat = "identity",position=position_jitterdodge()) +
  geom_errorbar(aes(ymin = mean - se, 
                    ymax = mean + se), 
                width=0.2,
                position=position_dodge(.9)) +
    ylab("Percentage of Translation Equivalents\n") +
    xlab("Balance") +
scale_x_discrete(breaks= c("80_20", "70_30", "60_40", "50_50"),
                        labels =c("80:20","70:30","60:40","50:50"))+
    ggtitle("Percentage of Translation Equivalents by Balance") +
    theme_classic(base_size = 22) +
  scale_y_continuous(labels = scales::percent_format(accuracy=2))+
    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
          plot.title = element_text(hjust = .5))

ggsave(here("output","PlotTE_byBalance.png"), width=9.5, height= 7)
```
## age by balancedness
```{r}
keepers_ws_TE %>%
  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
  group_by(balance_group15) %>%
  summarize(median.age = median(age.days))

keepers_ws_TE %>%
  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
  ggplot(aes(x=balance_group15, y=age.days/30, colour=balance_group15))+
  geom_pirate(bars=F) +ylab("Age (Months)") + coord_flip()

```

## TE: age plot by balancedness
```{r}
keepers_ws_TE %>%
  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
  ggplot(aes(x = age.days/30, 
             y = propTE)) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 2))+
  ylab("Percentage of Translation Equivalents\n") +
  xlab("Age in Months")+
  ggtitle("Percentage of Translation Equivalents by Age and Balance") +  guides(colour = guide_legend(reverse=TRUE))+

  theme_classic(base_size = 22) +
 # scale_x_continuous(breaks= seq(18,32,by = 2))+
  geom_point(size=3, colour= "darkgrey") + 
  geom_smooth(size =2, aes(colour=balance_group15), method="lm", se=F)+
  scale_colour_discrete(name="Balance", 
                        breaks= c("80_20", "70_30", "60_40", "50_50"),
                        labels =c("least balanced (80:20)","less balanced (70:30)","more balanced (60:40)","most balanced (50:50)"))
ggsave(here("Output","PlotTE_BalanceByAge.png"), width=13, height= 7)

```

# Statistical tests
## ANOVA for balance
```{r}
res.aov <- aov(propTE ~ balance_group15*cent.age.months, data=keepers_ws_TE) 
summary(res.aov)
```

## Age (old code, not used for poster)
### Running linear model
```{r}
  linear.model<-lmer(propTE ~ cent.age.months+ (1|Baby_ID), data = keepers_ws_TE)
  summary(linear.model)
```
### Running quadratic model
```{r}
 quadratic.model <-lmer(propTE ~ poly(cent.age.months,2) + (1|Baby_ID), data = keepers_ws_TE)
  summary(quadratic.model)
```
### Cubic model
```{r}
cubic.model<-lm(propTE ~ poly(cent.age.months,3), data = keepers_ws_TE)
summary(cubic.model)
```

### 2020-12-07 Analyze_TEs.Rmd
fixed typo and created an output folder

@@
   spread(vocab_type, vocab_score) %>%
-
-  ######
-  rename(baby_id = Baby_ID, age_months_binned = age.months,
-         word_vocab = Totalvocabulary,
-         total_words_eng = TotalNW_Eng, total_words_fre = TotalNW_Fr,
-         eng_unique_words = Eng_unique_words, fre_unique_words = Fr_unique_words,
-         number_of_te = TotalnumberTEs,
-         lang_exp_eng = LangExpEng, lang_exp_fre = LangExpFr, lang_exp_other = LangExpOther) %>%
-
   # calculate age in months (with decimal)
@@
 d_participants <- keepers_ws_TE %>%
-  filter(vocab_type == "word_vocab") %>%
   select(baby_id, age_days, gender, 
@@
 keepers_ws_TE %>%
-  group_by(Gender) %>% 
+  group_by(gender) %>% 
   distinct(baby_id, .keep.all=T) %>%
@@
 do.call(cbind, lapply(keepers_ws_TE %>%
-                        distinct(baby_ID, .keep_all = T) %>% 
+                        distinct(baby_id, .keep_all = T) %>% 
                         select(age_days,age_continuous),
@@
 keepers_ws_TE %>%
-  filter(bothCDI_filled=="Y") %>%
+  filter(both_cdi_filled=="Y") %>%
   summarise(mean = mean(propTE, na.rm=T), 
@@
 ### Model 1: Predicted_TotalnumberTEs_percentile
-Prediction_model <- keepers_ws_TE_simulation %>%
+Percentile_model <- keepers_ws_TE_simulation %>%
   lm(number_of_te*(Total_90percentile/2) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
@@
 
-ggsave(here("TEprediction_percentile.png"), TEprediction_percentile)
+ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile)
 
@@
 
-ggsave(here("TEprediction_percentile&CDI.png"), TEprediction_percentile_CDI)
+ggsave(here("Output/TEprediction_percentile&CDI.png"), TEprediction_percentile_CDI)
 
@@
 
-ggsave(here("TEprediction_percentile&CDI_LessThan100.png"), TEprediction_percentile_CDI_LessThan100)
+ggsave(here("Output/TEprediction_percentile&CDI_LessThan100.png"), TEprediction_percentile_CDI_LessThan100)
 
@@
 
-ggsave(here("TEprediction_percentile&CDI_LessThan50.png"), TEprediction_percentile_CDI_LessThan50)
+ggsave(here("Output/TEprediction_percentile&CDI_LessThan50.png"), TEprediction_percentile_CDI_LessThan50)
 
@@
 
-ggsave(here("TEprediction_percentile&CDI_MoreThan100.png"), TEprediction_percentile_CDI_MoreThan100)
+ggsave(here("Output/TEprediction_percentile&CDI_MoreThan100.png"), TEprediction_percentile_CDI_MoreThan100)
 

### 2020-12-18 computational model/TE_computation.Rmd
Updated computational model

@@
 library(tidylog)
-library(regclass)
+library(ggpubr)
+library(regclass) # for correlation
+library(effects)
+# library(jtools)
+library(FRACTION) # this allows us to use the is.wholenumber function
 
 `%notin%` <- Negate(`%in%`)
+
 ```
@@
 # Total word vocabulary of a child
-W <- seq(0, 2*V, by = 50) %>% # a sequence of 0 to 2V
-  rep(., 6) # generate 6 repetitinos of the sequence
+W <- seq(0, 2*V, by = 25) %>% # a sequence of 0 to 2V
+  rep(., 21) # generate 21 repetitinos of the sequence
 
 # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.1) %>% # a sequence of 0.5 to 1
-  rep(., 25) # generate 25 repetitions of the sequence
+B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
+  rep(., 49) # generate 25 repetitions of the sequence
 
@@
   # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600))
+  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  # filter out cases where D or N aren't whole numbers
+  filter_at(vars(D, N), any_vars(is.wholenumber(.)))
 
@@
 ## Creating variables: D and N
-D <- seq(0, V, by = 50) %>% # Words produced in the dominant language
-  rep(., 13) # generate 13 repetitions of the sequence
+D <- seq(0, V, by = 25) %>% # Words produced in the dominant language
+  rep(., 25) # generate 13 repetitions of the sequence
 
-N <- seq(0, V, by =50) %>% # Words produced in non-dominant language
-  rep(., 13) # generate 13 repetitions of the sequence
+N <- seq(0, V, by = 25) %>% # Words produced in non-dominant language
+  rep(., 25) # generate 25 repetitions of the sequence
 
@@
 
-#### Combine the 2 sets ####
+#### Set 3: Calculate other variables based on V, D, and B ####
+
+## Creating variables: D and N
+D <- seq(0, V, by = 25) %>% # Words produced in the dominant language
+  rep(., 21) # generate 13 repetitions of the sequence
+
+B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
+  rep(., 25) # generate 25 repetitions of the sequence
+
+## Creating the dataframe
+TE_fromVDB <- data.frame(V, D) %>%
+  arrange(D) %>%
+  mutate(B = B) %>%
+  # Create other variables
+  mutate(W = D/B, # Total word vocabulary of a child
+         N = W-D, # Words produced in non-dominant language
+         TE = (D*N)/V, # If D and N are independent, TE = DN/V
+         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
+         Du = D-TE, # Unique words in dominant language
+         Nu = N-TE) %>% # Unique words in non-dominant language
+  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
+  filter(Nu >= 0) %>%
+  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
+  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  # filter out cases where N isn't whole number
+  filter(is.wholenumber(N))
+
+#### Set 4: Calculate other variables based on V, N, and B ####
+
+## Creating variables: D and N
+N <- seq(0, V, by = 25) %>% # Words produced in the dominant language
+  rep(., 21) # generate 13 repetitions of the sequence
+
+B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
+  rep(., 25) # generate 25 repetitions of the sequence
+
+## Creating the dataframe
+TE_fromVNB <- data.frame(V, N) %>%
+  arrange(N) %>%
+  mutate(B = B) %>%
+  # Create other variables
+  mutate(W = D/(1-B), # Total word vocabulary of a child
+         D = W-N, # Words produced in dominant language
+         TE = (D*N)/V, # If D and N are independent, TE = DN/V
+         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
+         Du = D-TE, # Unique words in dominant language
+         Nu = N-TE) %>% # Unique words in non-dominant language
+  # It is not possible for N > D, so keep only cases where D >= N
+  filter(D >= N) %>%
+  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
+  filter(Nu >= 0) %>%
+  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
+  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  # filter out cases where D isn't whole number
+  filter(is.wholenumber(D)) %>%
+  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
+
+#### Combine the 4 sets ####
 TE_computation <- TE_fromVWB %>%
-  bind_rows(TE_fromVDN)
+  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
+  # constrain all variables but B (balance) to be integers 
+  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
 
@@
 
+mycor <- rcorr(as.matrix(TE_computation), type="pearson")
+mycor$r
+mycor$P
+
+library(corrplot)
+mycor <- cor(TE_computation)
+corrplot(mycor)
 #write.table(Correlation, "TEcomputation_correlations.tsv", col.names=NA)
@@
 
-plot_composition <- TE_computation %>%
+plot_composition_byB <- TE_computation %>%
   # Convert to long data
@@
                      labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  group_by(bin_B, vocab_type, W) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  mutate(bin_W = cut_interval(W, 8,
+                     labels = c(150, 300, 450, 600, 750, 900, 1050, 1200))) %>%
+  group_by(bin_B, vocab_type, bin_W) %>%
   summarise(n_words = mean(number)) %>%
-  ggplot(aes(x=W, y = n_words, fill = vocab_type)) + 
+  ggplot(aes(x = bin_W, y = n_words, fill = vocab_type)) + 
   geom_bar(position="stack", stat="identity") +
@@
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) +
   labs(x = "Total vocabulary size", 
-       y = "Number of words", 
+       y = "Number of concepts", 
        title = "Vocabulary composition across different balance proportions")
 
-ggsave("plot_composition.png", plot_composition)
+ggsave("plot_composition_byB.png", plot_composition_byB)
+
+plot_composition_byW <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N", "C")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  mutate(bin_W = cut_interval(W, 6,
+                     labels = c(200, 400, 600, 800, 1000, 1200))) %>%
+  group_by(bin_B, vocab_type, bin_W) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ bin_W) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
+  labs(x = "Balance", 
+       y = "Number of concepts", 
+       title = "Vocabulary composition across different vocabulary size")
+
+```
+
+## Compare different vocabulary measures across balance
+```{r}
+plot_words <- TE_computation %>%
+  # Convert to long data
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = as.factor(vocab_type)) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  group_by(bin_B, vocab_type, W) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = W, y = n_words, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Total vocabulary Size (W)", 
+       y = "Number of words", 
+       color = "Balance",
+       title = "Word vocabulary size by balance under different vocabulary measures") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_words.png", plot_words,
+       width = 10, height = 5)
 ```
@@
 ```
+
+# Plotting relations between variables (while holding some variables constant)
+## TE ~ predictors (input)
+```{r}
+#TE_Constant <- lm(TE ~ ., data = TE_computation) 
+#summary(TE_Constant)
+
+#### run a regression with TE ~ B, D, N
+TE_InputConstant <- lm(TE ~ 0 + B + D + N, data = TE_computation) 
+summary(TE_InputConstant)
+
+### plot the relation between TE & B / D / N
+#effect_plot(TE_InputConstant, pred = B, plot.points = TRUE)
+effect_B <- effect(term = "B", mod = TE_InputConstant) %>%
+  data.frame()
+
+plotConstant_TE_B <- ggplot(effect_B, aes(x = B, y = fit)) +
+  geom_line() +
+  ylim(-25, 600) +
+  labs(x = 'Balance', y = 'Number of TEs') +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_B.png", plotConstant_TE_B)
+
+### plot the relation between TE & D
+#effect_plot(TE_InputConstant, pred = D, plot.points = TRUE)
+effect_D <- effect(term = "D", mod = TE_InputConstant) %>%
+  data.frame()
+
+plotConstant_TE_D <- ggplot(effect_D, aes(x = D, y = fit)) +
+  geom_line() +
+  ylim(-25, 600) +
+  labs(x = 'Dominant vocabulary', y = 'Number of TEs') +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_D.png", plotConstant_TE_D)
+
+### plot the relation between TE & N
+#effect_plot(TE_InputConstant, pred = N, plot.points = TRUE)
+effect_N <- effect(term = "N", mod = TE_InputConstant) %>%
+  data.frame()
+
+plotConstant_TE_N <- ggplot(effect_N, aes(x = N, y = fit)) +
+  geom_line() +
+  ylim(-25, 600) +
+  labs(x = 'Non-dominant vocabulary', y = 'Number of TEs') +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_N.png", plotConstant_TE_D)
+
+## put the three plots together
+plotConstant_TE_BDN <- ggarrange(plotConstant_TE_B, plotConstant_TE_D, plotConstant_TE_N, 
+                                 ncol = 3, nrow = 1,
+                                 widths = 3, heights = 1) %>%
+  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + B + D + N", face = "bold", size = 14))
+
+ggsave("plotConstant_TE_BDN.png", plotConstant_TE_BDN, 
+       width = 10, height = 5)
+```
+
+## TE ~ predictors (output)
+```{r}
+#### run a regression with TE ~ C, Du, Nu
+TE_OutputConstant <- lm(TE ~ 0 + C + Du + Nu, data = TE_computation) 
+summary(TE_OutputConstant)
+
+## plot the relation between TE & C
+#effect_plot(TE_OutputConstant, pred = C, plot.points = TRUE)
+effect_C <- effect(term = "C", mod = TE_OutputConstant) %>%
+  data.frame()
+
+plotConstant_TE_C <- ggplot(effect_C, aes(x = C, y = fit)) +
+  geom_line() +
+  labs(x = 'Concept vocabulary', y = 'Number of TEs') +
+  ylim(-300, 600) +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_C.png", plotConstant_TE_C)
+
+## plot the relation between TE & Du
+#effect_plot(TE_OutputConstant, pred = Du, plot.points = TRUE)
+effect_Du <- effect(term = "Du", mod = TE_OutputConstant) %>%
+  data.frame()
+
+plotConstant_TE_Du <- ggplot(effect_Du, aes(x = Du, y = fit)) +
+  geom_line() +
+  labs(x = 'Unique dominant vocabulary', y = 'Number of TEs') +
+  ylim(-300, 600) +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)
+
+## plot the relation between TE & Nu
+#effect_plot(TE_OutputConstant, pred = Nu, plot.points = TRUE)
+effect_Nu <- effect(term = "Nu", mod = TE_OutputConstant) %>%
+  data.frame()
+
+plotConstant_TE_Nu <- ggplot(effect_Nu, aes(x = Nu, y = fit)) +
+  geom_line() +
+  labs(x = 'Unique non-dominant vocabulary', y = 'Number of TEs') +
+  ylim(-300, 600) +
+  xlim(0,600) +
+  theme_minimal()
+
+#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)
+
+## put the three plots together
+plotConstant_TE_CDuNu <- ggarrange(plotConstant_TE_C, plotConstant_TE_Du, plotConstant_TE_Nu, 
+                                 ncol = 3, nrow = 1,
+                                 widths = 3, heights = 1) %>%
+  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + C + Du + Nu", face = "bold", size = 14))
+
+ggsave("plotConstant_TE_CDuNu.png", plotConstant_TE_CDuNu, 
+       width = 10, height = 5)
+```
+
+# Estimating  if the models is biased towards ot against learning TEs
+```{r}
+## plot
+plot_learnVSpredict <- TE_computation %>%
+  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
+         learned_TE = (D*N)/(W/2)) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  ggplot(aes(x = learned_TE, y = predicted_TE, color = bin_B)) +
+  geom_point() +
+  geom_abline(intercept = 0, slope = 1) +
+  labs(x = "Learned translation equivalents (= (D*N)/(W/2))",
+       y = "Predicted translation equivalents under the model (= (D*N)/V)",
+       color = "Bins of balance") +                                                    
+  theme_minimal()
+
+ggsave("plot_learnVSpredict.png", plot_learnVSpredict)
+
+## Kolmogorov-Smirnov test to compare the observed and predicted TE
+TE_computation <- TE_computation %>%
+  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
+         learned_TE = (D*N)/(W/2))
+
+stats::ks.test(TE_computation$predicted_TE, TE_computation$learned_TE)
+
+```

### 2020-12-24 computational model/TE_computation.Rmd
Updated computational codes

@@
 #### Set 1: Calculate other variables based on V, W, and B ####
-
 ## Creating variables: V, W, and B
 # Items on vocabulary measure (e.g. words on the CDI)
-V <- 600 # This is constant
+V <- 600
 
 # Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25) %>% # a sequence of 0 to 2V
-  rep(., 21) # generate 21 repetitinos of the sequence
+W <- seq(0, 2*V, by = 25)
 
 # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
-  rep(., 49) # generate 25 repetitions of the sequence
+B <- seq(0.5, 1, by = 0.025)
+
+# Generating sequence of W and B
+W_seq <- W %>% # a sequence of 0 to 2V
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>%  # a sequence of 0.5 to 1
+  rep(., length(W)) # generate N repetitions of the sequence based on length of W
 
 ## Creating the dataframe
-TE_fromVWB <- data.frame(V, W) %>%
-  arrange(W) %>%
-  mutate(B = B) %>%
+TE_fromVWB <- data.frame(V, W_seq) %>%
+  arrange(W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B_seq) %>%
   # Create other variables
@@
 #### Set 2: Calculate other variables based on V, D, and N ####
+## Creating variables: D and N
+D <- seq(0, V, by = 25)
 
-## Creating variables: D and N
-D <- seq(0, V, by = 25) %>% # Words produced in the dominant language
-  rep(., 25) # generate 13 repetitions of the sequence
+N <- seq(0, V, by = 25)
 
-N <- seq(0, V, by = 25) %>% # Words produced in non-dominant language
-  rep(., 25) # generate 25 repetitions of the sequence
+# Generating sequence of D and N
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+N_seq <- N %>% # Words produced in non-dominant language
+  rep(., length(D)) # generate N repetitions of the sequence based on length of D
 
 ## Creating the dataframe
-TE_fromVDN <- data.frame(V, D) %>%
-  arrange(D) %>%
-  mutate(N = N) %>%
+TE_fromVDN <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(N = N_seq) %>%
   # It is not possible for N > D, so keep only cases where D >= N
@@
 #### Set 3: Calculate other variables based on V, D, and B ####
+## Creating variables: D and N
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
 
-## Creating variables: D and N
-D <- seq(0, V, by = 25) %>% # Words produced in the dominant language
-  rep(., 21) # generate 13 repetitions of the sequence
-
-B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
-  rep(., 25) # generate 25 repetitions of the sequence
+B_seq <- B %>% # balance
+  rep(., length(D))  # generate N repetitions of the sequence based on length of D
 
 ## Creating the dataframe
-TE_fromVDB <- data.frame(V, D) %>%
-  arrange(D) %>%
-  mutate(B = B) %>%
+TE_fromVDB <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(B = B_seq) %>%
   # Create other variables
@@
 #### Set 4: Calculate other variables based on V, N, and B ####
+## Creating variables: D and N
+N_seq <- N %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
 
-## Creating variables: D and N
-N <- seq(0, V, by = 25) %>% # Words produced in the dominant language
-  rep(., 21) # generate 13 repetitions of the sequence
-
-B <- seq(0.5, 1, by = 0.025) %>% # a sequence of 0.5 to 1
-  rep(., 25) # generate 25 repetitions of the sequence
+B_seq <- B %>% # a sequence of 0.5 to 1
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
 
 ## Creating the dataframe
-TE_fromVNB <- data.frame(V, N) %>%
-  arrange(N) %>%
-  mutate(B = B) %>%
+TE_fromVNB <- data.frame(V, N_seq) %>%
+  arrange(N_seq) %>%
+  rename(N = N_seq) %>%
+  mutate(B = B_seq) %>%
   # Create other variables
@@
   filter(is.wholenumber(D)) %>%
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
+  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
+  filter(is.wholenumber(Nu))
 

### 2021-01-06 computational model/TE_computation.Rmd
Amending the simulation of the computational data set
Adding more levels of V

@@
 # Generating dataframe for computational model
+#### Set 1: Calculate other variables based on V, W, and B ####
 ```{r}
-#### Set 1: Calculate other variables based on V, W, and B ####
-## Creating variables: V, W, and B
+### Generating variables: V, W, and B
+## V = 600 ##
 # Items on vocabulary measure (e.g. words on the CDI)
@@
 
-## Creating the dataframe
-TE_fromVWB <- data.frame(V, W_seq) %>%
+# Creating the dataframe
+TE_V600 <- data.frame(V, W_seq) %>%
   arrange(W_seq) %>%
   rename(W = W_seq) %>%
-  mutate(B = B_seq) %>%
+  mutate(B = B_seq)
+
+## V = 450 ##
+# Items on vocabulary measure (e.g. words on the CDI)
+V <- 450
+
+# Total word vocabulary of a child
+W <- seq(0, 2*V, by = 25)
+
+# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
+B <- seq(0.5, 1, by = 0.025)
+
+# Generating sequence of W and B
+W_seq <- W %>% # a sequence of 0 to 2V
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>%  # a sequence of 0.5 to 1
+  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+
+# Creating the dataframe
+TE_V450 <- data.frame(V, W_seq) %>%
+  arrange(W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B_seq)
+
+## V = 300 ##
+# Items on vocabulary measure (e.g. words on the CDI)
+V <- 300
+
+# Total word vocabulary of a child
+W <- seq(0, 2*V, by = 25)
+
+# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
+B <- seq(0.5, 1, by = 0.025)
+
+# Generating sequence of W and B
+W_seq <- W %>% # a sequence of 0 to 2V
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>%  # a sequence of 0.5 to 1
+  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+
+# Creating the dataframe
+TE_V300 <- data.frame(V, W_seq) %>%
+  arrange(W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B_seq)
+
+## V = 150 ##
+# Items on vocabulary measure (e.g. words on the CDI)
+V <- 150
+
+# Total word vocabulary of a child
+W <- seq(0, 2*V, by = 25)
+
+# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
+B <- seq(0.5, 1, by = 0.025)
+
+# Generating sequence of W and B
+W_seq <- W %>% # a sequence of 0 to 2V
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>%  # a sequence of 0.5 to 1
+  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+
+# Creating the dataframe
+TE_V150 <- data.frame(V, W_seq) %>%
+  arrange(W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B_seq)
+
+### Combining all dataframes and Calculate other variables
+TE_fromVWB <- TE_fromVWB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # Create other variables
@@
   # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  filter_at(vars(D, N), all_vars(. <= V)) %>%
   # filter out cases where D or N aren't whole numbers
   filter_at(vars(D, N), any_vars(is.wholenumber(.))) 
+```
 
 #### Set 2: Calculate other variables based on V, D, and N ####
-## Creating variables: D and N
+```{r}
+### Generating variables: D and N
+## V = 600 ##
+V <- 600
 D <- seq(0, V, by = 25) 
-
 N <- seq(0, V, by = 25)
@@
 
-## Creating the dataframe
-TE_fromVDN <- data.frame(V, D_seq) %>%
+# Creating the dataframe
+TE_V600 <- data.frame(V, D_seq) %>%
   arrange(D_seq) %>%
   rename(D = D_seq) %>%
-  mutate(N = N_seq) %>%
+  mutate(N = N_seq)
+
+## V = 450 ##
+V <- 450
+D <- seq(0, V, by = 25) 
+N <- seq(0, V, by = 25)
+
+# Generating sequence of D and N
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+N_seq <- N %>% # Words produced in non-dominant language
+  rep(., length(D)) # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V450 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(N = N_seq)
+
+## V = 300 ##
+V <- 300
+D <- seq(0, V, by = 25) 
+N <- seq(0, V, by = 25)
+
+# Generating sequence of D and N
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+N_seq <- N %>% # Words produced in non-dominant language
+  rep(., length(D)) # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V300 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(N = N_seq)
+
+## V = 150 ##
+V <- 150
+D <- seq(0, V, by = 25) 
+N <- seq(0, V, by = 25)
+
+# Generating sequence of D and N
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+N_seq <- N %>% # Words produced in non-dominant language
+  rep(., length(D)) # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V150 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(N = N_seq)
+
+### Combining all dataframes and Calculate other variables
+TE_fromVDN <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # It is not possible for N > D, so keep only cases where D >= N
@@
   filter(!is.na(B)) 
+```
 
 #### Set 3: Calculate other variables based on V, D, and B ####
-## Creating variables: D and N
+```{r}
+### Creating variables: D and B
+## V = 600 #
+V <- 600
+D <- seq(0, V, by = 25)
+
 D_seq <- D %>% # Words produced in the dominant language
@@
 
-## Creating the dataframe
-TE_fromVDB <- data.frame(V, D_seq) %>%
+# Creating the dataframe
+TE_V600 <- data.frame(V, D_seq) %>%
   arrange(D_seq) %>%
   rename(D = D_seq) %>%
-  mutate(B = B_seq) %>%
+  mutate(B = B_seq)
+
+## V = 450 #
+V <- 450
+D <- seq(0, V, by = 25)
+
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # balance
+  rep(., length(D))  # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V450 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(B = B_seq)
+
+## V = 300 #
+V <- 300
+D <- seq(0, V, by = 25)
+
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # balance
+  rep(., length(D))  # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V300 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(B = B_seq)
+
+## V = 150 #
+V <- 150
+D <- seq(0, V, by = 25)
+
+D_seq <- D %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # balance
+  rep(., length(D))  # generate N repetitions of the sequence based on length of D
+
+# Creating the dataframe
+TE_V150 <- data.frame(V, D_seq) %>%
+  arrange(D_seq) %>%
+  rename(D = D_seq) %>%
+  mutate(B = B_seq)
+
+### Combining all dataframes and Calculate other variables
+TE_fromVDB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # Create other variables
@@
   # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  filter_at(vars(D, N), all_vars(. <= V)) %>%
   # filter out cases where N isn't whole number
   filter(is.wholenumber(N))
+```
 
 #### Set 4: Calculate other variables based on V, N, and B ####
-## Creating variables: D and N
+```{r}
+### Generating variables: N and B
+## V = 600 #
+V <- 600
+N <- seq(0, V, by = 25)
+
 N_seq <- N %>% # Words produced in the dominant language
@@
 
-## Creating the dataframe
-TE_fromVNB <- data.frame(V, N_seq) %>%
+# Creating the dataframe
+TE_V600 <- data.frame(V, N_seq) %>%
   arrange(N_seq) %>%
   rename(N = N_seq) %>%
-  mutate(B = B_seq) %>%
+  mutate(B = B_seq)
+
+## V = 450 #
+V <- 450
+N <- seq(0, V, by = 25)
+
+N_seq <- N %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # a sequence of 0.5 to 1
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+# Creating the dataframe
+TE_V450 <- data.frame(V, N_seq) %>%
+  arrange(N_seq) %>%
+  rename(N = N_seq) %>%
+  mutate(B = B_seq)
+
+## V = 300 #
+V <- 300
+N <- seq(0, V, by = 25)
+
+N_seq <- N %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # a sequence of 0.5 to 1
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+# Creating the dataframe
+TE_V300 <- data.frame(V, N_seq) %>%
+  arrange(N_seq) %>%
+  rename(N = N_seq) %>%
+  mutate(B = B_seq)
+
+## V = 150 #
+V <- 150
+N <- seq(0, V, by = 25)
+
+N_seq <- N %>% # Words produced in the dominant language
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+B_seq <- B %>% # a sequence of 0.5 to 1
+  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+
+# Creating the dataframe
+TE_V150 <- data.frame(V, N_seq) %>%
+  arrange(N_seq) %>%
+  rename(N = N_seq) %>%
+  mutate(B = B_seq)
+
+### Combining all dataframes and Calculate other variables
+TE_fromVNB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # Create other variables
@@
   # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
+  filter_at(vars(D, N), all_vars(. <= V)) %>%
   # filter out cases where D isn't whole number
@@
   filter(is.wholenumber(Nu)) 
+```
 
 #### Combine the 4 sets ####
+```{r}
 TE_computation <- TE_fromVWB %>%
@@
   filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
-
 ```
@@
 Correlation <- TE_computation %>%
-  select(-V) %>%
   cor_matrix()
 
+library(Hmisc)
 mycor <- rcorr(as.matrix(TE_computation), type="pearson")
@@
        title = "Vocabulary composition across different vocabulary size") 
-
 ```
@@
 ```{r}
+vocab_label <- c("Concept Vocabulary (C)", "Unique Dominant Words (Du)", "Unique Non-dominant Words (Nu)")
+
 plot_words <- TE_computation %>%
@@
   pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N")) %>%
   mutate(vocab_type = as.factor(vocab_type)) %>%
+  mutate(vocab_type = recode(vocab_type, C = "Concept vocabulary (C)", Du = "Unique dominant words (Du)",
+         Nu = "Unique non-dominant words (Nu)", TE = "Translation equivalents")) %>%
   mutate(bin_B = cut(B, breaks = 6, 

### 2021-01-07 Analyze_TEs.Rmd
New stacked barchart

@@
 library(wordbankr)
+library("ggpubr")
 
@@
   #Defining vocabulary-based balance score
-  mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
+  #mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
+  mutate(balance_vocab = TotalNW_Dom/(TotalNW_Dom+TotalNW_NonDom)) %>%
   #Defining proportion of TE (with reference to child's own total vocab size)
@@
 d_participants <- keepers_ws_TE %>%
-  select(baby_id, age_days, gender, 
+  select(baby_id, age_continuous, gender, 
          lang_exp_eng, lang_exp_fre, lang_exp_other, years_education) %>%
@@
   summarise(n = n(),
-            age.days = mean(age_days, na.rm = TRUE),
+            age_Month = mean(age_continuous, na.rm = TRUE),
             lang_exp_eng = mean(lang_exp_eng, na.rm = TRUE),
@@
             years_education = mean(years_education, na.rm = TRUE),
-            number_female = sum(Female))
+            number_female = sum(Female)) %>%
+  mutate(PercentageFemale = number_female/n)
+
 ```
@@
 
+# Correlation
+```{r}
+Correlation <- keepers_ws_TE %>%
+  select(age_continuous, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
+         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
+  rename(Age_month = age_continuous,
+         B = balance_vocab, 
+         W = word_vocab,
+         D = TotalNW_Dom, 
+         N = TotalNW_NonDom, 
+         TE = number_of_te, 
+         Du = Dom_unique_words, 
+         Nu = NonDom_unique_words, 
+         C = concept_vocab)
+
+mycor <- rcorr(as.matrix(Correlation), type="pearson")
+mycor$r
+mycor$P
+
+library(corrplot)
+mycor <- cor(Correlation) 
+corrplot(mycor)
+write.csv(mycor, "ObservedTE_correlations.csv", col.names=NA)
+```
+
+###### Stacked barchat: Y=Total vocab, X= individual baby (stacked: TE, unique dom, unique nondom) ######
+```{r}
+
+##Creating dataset for stacked barchart
+
+stacked <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  arrange(desc(number_of_te)) %>%
+  mutate(TE = number_of_te) %>%
+  pivot_longer(-c(baby_id, balance_vocab, word_vocab, TE), 
+               names_to = "Type", values_to = "NumberOfWords") %>%
+  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te")))
+
+##Stacked barchart (sort by total number of words)
+
+stacked %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+    labs(x = "Individual infant", 
+       y = "Number of words produced") + 
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))
+
+##Stacked barchart (sort by total number of TEs)
+
+stacked %>%
+  ggplot(aes(x = reorder(baby_id, -TE, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+    labs(x = "Individual infant", 
+       y = "Number of words produced") + 
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))
+
+#stacked %>%
+#  mutate(bin_balance = cut_width(balance_vocab, 0.333333333333333, boundary = 0)) %>%
+#  ggplot(aes(x = reorder(Baby_ID, -NumberOfWords, sum), y = NumberOfWords)) +
+#  geom_col(aes(fill = Type)) +
+#  facet_grid(. ~ bin_balance, labeller=labeller(bin_balance = balance_facet)) + 
+#  labs(x = "Individual infant", 
+#       y = "Number of words produced") + 
+#  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c"))
+
+Bal_StackedPlot <- stacked %>%
+  mutate(bin_B = cut(balance_vocab, breaks = 2, 
+                     labels = c("More balanced", "Less balanced"))) %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  facet_grid(. ~ bin_B) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Least balanced: 0 - .333") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
+       width = 10, height = 5)
+
+```
+
+```{r}
+stacked <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  arrange(desc(number_of_te)) %>%
+  mutate(TE = number_of_te) %>%
+  pivot_longer(-c(baby_id, balance_vocab, word_vocab, TE), 
+               names_to = "Type", values_to = "NumberOfWords") %>%
+  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
+  #Subset balance_vocab by 3 levels 
+  mutate(BalSubset = cut(balance_vocab, breaks = 3, 
+                         labels=c("HighBalance", "MidRangeBalance", "LeastBalance")))
+
+LeastBal_plot <- stacked %>%
+  filter(BalSubset == "LeastBalance") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Least balanced") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+MidRangeBal_plot <- stacked %>%
+  filter(BalSubset == "MidRangeBalance") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = " ") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+HighBal_plot <- stacked %>%
+  filter(BalSubset == "HighBalance") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "High balance") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+Bal_StackedPlot <- ggarrange(LeastBal_plot, 
+          MidRangeBal_plot + rremove("ylab"), 
+          HighBal_plot + rremove("ylab"),
+          ncol = 3, nrow = 1,
+          common.legend = TRUE, legend="bottom")
+
+ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
+       width = 10, height = 5)
+
+```
+
+## Linear regression: TE ~ B + D + N
+```{r}
+TE_input_lm <- lmer(number_of_te ~ 0 + balance_vocab + TotalNW_Dom + TotalNW_NonDom,
+                  data = keepers_ws_TE)
+
+summary(TE_input_lm)
+```
+
+## Linear regression: TE ~ Du * Nu
+```{r}
+TE_output_lm <- lm(number_of_te ~ 0 + Dom_unique_words * NonDom_unique_words,
+                  data = keepers_ws_TE)
+
+summary(TE_output_lm)
+
+### plot the Du*Nu interaction
+#effect_plot(TE_InputConstant, pred = B, plot.points = TRUE)
+interaction_DuNu <- effect(term = "Dom_unique_words:NonDom_unique_words", mod = TE_output_lm) %>%
+  data.frame()
+
+plot_interactionDuNu <- ggplot(interaction_DuNu, 
+                               aes(x = Dom_unique_words, y = fit, group = NonDom_unique_words)) +
+  geom_line(aes(color = NonDom_unique_words)) +
+  labs(x = 'Unique dominant words',
+       y = 'Number of TEs',
+       color = 'Unique non-dominant words')
+
+  ylim(-25, 600) +
+  labs(x = 'Balance', y = 'Number of TEs') +
+  theme_minimal()
+
+```
+
 ## Create predicted TEs for simulation
@@
   ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
-  facet_grid(. ~TEgroup) +
-  labs(x = "Predicted total number of TEs (based on percentile)", y = "Observed total number of TEs",
-       title = "Observed vs. Predicted total number of TEs") +
+#  facet_grid(. ~TEgroup) +
+  labs(x = "Simulated total number of TEs (based on percentile)", y = "Observed total number of TEs") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   xlim(-10,430) +
-  ylim(-10,430)
+  ylim(-10,430) +
+  theme_minimal()
 
@@
   ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
-  facet_grid(. ~TEgroup) +
+#  facet_grid(. ~TEgroup) +
   labs(x = "Predicted total number of TEs", 

### 2021-01-12 computational model/TE_computation.Rmd
new plots

@@
        title = "Vocabulary composition across different vocabulary size")
+
+plot_composition_byV <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N", "C")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
+  labs(x = "Balance", 
+       y = "Number of concepts", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
+
+ggsave("plot_composition_byV.png", plot_composition_byV,
+        width = 10, height = 5)
+
+plot_composition_byV <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N", "C")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(bin_B, vocab_type, V) %>%
+  summarize(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) +
+  geom_bar(position = "stack", stat = "identity", reverse = TRUE) +
+  facet_grid(. ~ V) +
+  theme_minimal() +
+  ggtitle("Vocabulary composition across different developmental timepoints") +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom")
+
 ```

### 2021-02-01 Analyze_TEs.Rmd
20210201

@@
 
+library(Hmisc)
 mycor <- rcorr(as.matrix(Correlation), type="pearson")
@@
 library(corrplot)
-mycor <- cor(Correlation) 
+mycor <- round(cor(Correlation), 2)
 corrplot(mycor)
 write.csv(mycor, "ObservedTE_correlations.csv", col.names=NA)
+
+cor.test(Correlation$TE, Correlation$B, method = "pearson")
+cor.test(Correlation$TE, Correlation$W, method = "pearson")
+cor.test(Correlation$TE, Correlation$D, method = "pearson")
+cor.test(Correlation$TE, Correlation$N, method = "pearson")
+
 ```
@@
 
+##### Facet by W (< 100 vs. > 100)
+
 ###Prediction 2: Predicted raw TE = TotalNW_NonDom

### 2021-02-01 computational model/TE_computation_Data.Rmd
20210201

---
title: "TE_computation"
date: "Last update: Dec 2nd 2020"
output: html_document
---

# Preparations
## load libraries
```{r setup, include=FALSE}

library(here)
library(tidyverse)
library(tidylog)
library(ggpubr)
library(regclass) # for correlation
library(effects)
# library(jtools)
#library(FRACTION) # this allows us to use the is.wholenumber function

`%notin%` <- Negate(`%in%`)

```

# Generating dataframe for computational model
#### Set 1: Calculate other variables based on V, W, and B ####
```{r}
### Generating variables: V, W, and B
## V = 600 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 600

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V600 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 550 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 550

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V550 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 500 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 500

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V500 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 450 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 450

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V450 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 400 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 400

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V400 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 350 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 350

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V350 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 300 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 300

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V300 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 250 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 250

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V250 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 200 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 200

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V200 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 150 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 150

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V150 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 100 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 100

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V100 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

## V = 50 ##
# Items on vocabulary measure (e.g. words on the CDI)
V <- 50

# Total word vocabulary of a child
W <- seq(0, 2*V, by = 25)

# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
B <- seq(0.5, 1, by = 0.025)

# Generating sequence of W and B
W_seq <- W %>% # a sequence of 0 to 2V
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>%  # a sequence of 0.5 to 1
  rep(., length(W)) # generate N repetitions of the sequence based on length of W

# Creating the dataframe
TE_V50 <- data.frame(V, W_seq) %>%
  arrange(W_seq) %>%
  rename(W = W_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVWB <- TE_fromVWB <- rbind(TE_V50, TE_V100, TE_V150, TE_V200, TE_V250, TE_V300, 
                                  TE_V350, TE_V400, TE_V450, TE_V500, TE_V550, TE_V600) %>%
  # Create other variables
  mutate(D = B*W, # Words produced in the dominant language
         N = (1-B)*W, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

  ## filter out cases where D or N aren't whole numbers
  #filter_at(vars(D, N), any_vars(is.wholenumber(.))) 
```

#### Set 2: Calculate other variables based on V, D, and N ####
```{r}
### Generating variables: D and N
## V = 600 ##
V <- 600
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V600 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 550 ##
V <- 550
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V550 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 500 ##
V <- 500
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V500 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 450 ##
V <- 450
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V450 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 400 ##
V <- 400
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V400 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 350 ##
V <- 350
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V350 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 300 ##
V <- 300
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V300 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 250 ##
V <- 250
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V250 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 200 ##
V <- 200
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V200 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 150 ##
V <- 150
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V150 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 100 ##
V <- 100
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V100 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

## V = 50 ##
V <- 50
D <- seq(0, V, by = 25) 
N <- seq(0, V, by = 25)

# Generating sequence of D and N
D_seq <- D %>% # Words produced in the dominant language
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

N_seq <- N %>% # Words produced in non-dominant language
  rep(., length(D)) # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V50 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(N = N_seq)

### Combining all dataframes and Calculate other variables
TE_fromVDN <- rbind(TE_V50, TE_V100, TE_V150, TE_V200, TE_V250, TE_V300, 
                    TE_V350, TE_V400, TE_V450, TE_V500, TE_V550, TE_V600) %>%
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # Create other variables
  mutate(W = D+N, # Total word vocabulary of a child
         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>%
  # Remove cases where B = NA
  filter(!is.na(B)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)
```

#### Set 3: Calculate other variables based on V, D, and B ####
```{r}
### Creating variables: D and B
## V = 600 #
V <- 600
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V600 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 550 #
V <- 550
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V550 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 500 #
V <- 500
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V500 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 450 #
V <- 450
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V450 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 400 #
V <- 400
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V400 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 350 #
V <- 350
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V350 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 300 #
V <- 300
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V300 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 250 #
V <- 250
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V250 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 200 #
V <- 200
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V200 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 150 #
V <- 150
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V150 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 100 #
V <- 100
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V100 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

## V = 50 #
V <- 50
D <- seq(0, V, by = 25)

D_seq <- D %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # balance
  rep(., length(D))  # generate N repetitions of the sequence based on length of D

# Creating the dataframe
TE_V50 <- data.frame(V, D_seq) %>%
  arrange(D_seq) %>%
  rename(D = D_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVDB <- rbind(TE_V50, TE_V100, TE_V150, TE_V200, TE_V250, TE_V300, 
                    TE_V350, TE_V400, TE_V450, TE_V500, TE_V550, TE_V600) %>%
  # Create other variables
  mutate(W = D/B, # Total word vocabulary of a child
         N = W-D, # Words produced in non-dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

  ## filter out cases where N isn't whole number
  #filter(is.wholenumber(N))
```

#### Set 4: Calculate other variables based on V, N, and B ####
```{r}
### Generating variables: N and B
## V = 600 #
V <- 600
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V600 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 550 #
V <- 550
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V550 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 500 #
V <- 500
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V500 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 450 #
V <- 450
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V450 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 400 #
V <- 400
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V400 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 350 #
V <- 350
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V350 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 300 #
V <- 300
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V300 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 250 #
V <- 250
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V250 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 200 #
V <- 200
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V200 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 150 #
V <- 150
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V150 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 100 #
V <- 100
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V100 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

## V = 50 #
V <- 50
N <- seq(0, V, by = 25)

N_seq <- N %>% # Words produced in the dominant language
  rep(., length(B)) # generate N repetitions of the sequence based on length of B

B_seq <- B %>% # a sequence of 0.5 to 1
  rep(., length(N)) # generate N repetitions of the sequence based on length of N

# Creating the dataframe
TE_V50 <- data.frame(V, N_seq) %>%
  arrange(N_seq) %>%
  rename(N = N_seq) %>%
  mutate(B = B_seq)

### Combining all dataframes and Calculate other variables
TE_fromVNB <- rbind(TE_V50, TE_V100, TE_V150, TE_V200, TE_V250, TE_V300, 
                    TE_V350, TE_V400, TE_V450, TE_V500, TE_V550, TE_V600) %>%
  # Create other variables
  mutate(W = N/(1-B), # Total word vocabulary of a child
         D = W-N, # Words produced in dominant language
         TE = (D*N)/V, # If D and N are independent, TE = DN/V
         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
         Du = D-TE, # Unique words in dominant language
         Nu = N-TE) %>% # Unique words in non-dominant language
  # It is not possible for N > D, so keep only cases where D >= N
  filter(D >= N) %>%
  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
  filter(Nu >= 0) %>%
  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
  filter_at(vars(D, N), all_vars(. <= V)) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

  ## filter out cases where D isn't whole number
  #filter(is.wholenumber(D)) %>%
  #filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
  #filter(is.wholenumber(Nu)) 
```

#### Combine the 4 sets ####
```{r}
TE_computation <- TE_fromVWB %>%
  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
  # Round to keep whole numbers only
  mutate_at(vars(-B), round, 0)

write.csv(TE_computation, 'TE_computation.csv')

  ## constrain all variables but B (balance) to be integers 
  #filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
```

### 2021-02-01 computational model/TE_computation_Simulation.Rmd
20210201

--- 2020-12-24 computational model/TE_computation.Rmd
@@
 # library(jtools)
-library(FRACTION) # this allows us to use the is.wholenumber function
+#library(FRACTION) # this allows us to use the is.wholenumber function
 
@@
 
-# Generating dataframe for computational model
+# Read in data
 ```{r}
-#### Set 1: Calculate other variables based on V, W, and B ####
-## Creating variables: V, W, and B
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 600
-
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
-
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
-
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
-
-## Creating the dataframe
-TE_fromVWB <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(D = B*W, # Words produced in the dominant language
-         N = (1-B)*W, # Words produced in non-dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where D or N aren't whole numbers
-  filter_at(vars(D, N), any_vars(is.wholenumber(.)))
-
-#### Set 2: Calculate other variables based on V, D, and N ####
-## Creating variables: D and N
-D <- seq(0, V, by = 25)
-
-N <- seq(0, V, by = 25)
-
-# Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_fromVDN <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq) %>%
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # Create other variables
-  mutate(W = D+N, # Total word vocabulary of a child
-         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>%
-  # Remove cases where B = NA
-  filter(!is.na(B))
-
-#### Set 3: Calculate other variables based on V, D, and B ####
-## Creating variables: D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_fromVDB <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(W = D/B, # Total word vocabulary of a child
-         N = W-D, # Words produced in non-dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where N isn't whole number
-  filter(is.wholenumber(N))
-
-#### Set 4: Calculate other variables based on V, N, and B ####
-## Creating variables: D and N
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-## Creating the dataframe
-TE_fromVNB <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(W = D/(1-B), # Total word vocabulary of a child
-         D = W-N, # Words produced in dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where D isn't whole number
-  filter(is.wholenumber(D)) %>%
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
-  filter(is.wholenumber(Nu))
-
-#### Combine the 4 sets ####
-TE_computation <- TE_fromVWB %>%
-  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
-  # constrain all variables but B (balance) to be integers 
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
-
+TE_computation <- read.csv(here::here("computational model/TE_computation.csv"))
 ```
@@
 
+plot_composition_byV <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N", "C")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) + 
+  labs(x = "Balance", 
+       y = "Number of concepts", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
+
+ggsave("plot_composition_byV.png", plot_composition_byV,
+        width = 10, height = 5)
+
 ```
@@
        width = 10, height = 5)
+
+plot_inputs <- TE_computation %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("W", "D", "N", "TE")) %>%
+  mutate(vocab_type = as.factor(vocab_type)) %>%
+  mutate(vocab_type = fct_relevel(vocab_type, "W")) %>%
+  mutate(vocab_type = recode(vocab_type, W = "Total vocabulary (W)", D = "Dominant vocabulary (D)", 
+                             N = "Non-dominant vocabulary (N)", TE = "Translation equivalents (TE)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = V, y = n_words, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Different timepoints = Learnable vocaobulary (V)", 
+       y = "Number of words", 
+       color = "Balance",
+       title = "Word vocabulary size by balance under different input measures and translation equivalents") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs.png", plot_inputs,
+       width = 10, height = 5)
+
+plot_inputs_TE <- TE_computation %>%
+  # V = 600
+  filter(V == 600) %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("W", "D", "N")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             W = "Panel A: Total vocabulary (W)", 
+                             D = "Panel B: Dominant vocabulary (D)", 
+                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  ggplot(aes(x = number, y = TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Number of words", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)"
+       #, title = "Number of translation equivalent by balance in relation to different input measures "
+       ) +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs_TE.png", plot_inputs_TE,
+       width = 10, height = 5)
+
+plot_inputs_TE_V <- TE_computation %>%
+  filter(W == 250) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  ggplot(aes(x = V, y = TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  #facet_grid(. ~ bin_W) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Developmental timepoint (V)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+plot_inputs_TE_V <- TE_computation %>%
+  #filter(W == 100) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  mutate(bin_W = cut_interval(W, 4,
+                     labels = c("W = 0 - 300" , "W = 301 - 600", "W = 601 - 900", "W = 901 - 1200"))) %>%
+  ggplot(aes(x = V, y = TE, color = bin_W)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ bin_B) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Developmental timepoint (V)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs_TE_V.png", plot_inputs_TE_V,
+       width = 10, height = 5)
+
+```
+
+## Plotting different scenarios
+```{r}
+plot_scenarios <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("D", "N", "TE")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("N", "D", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
+  labs(x = "Balance", 
+       y = "Number of words (W)", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
+
 ```
@@
   mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
-  #mutate(bin_B = cut(B, breaks = 6, 
-  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
   ggplot(aes(x = number, y = TE)) +
-  facet_grid(. ~vocab_type) +
+  facet_grid(. ~ vocab_type) +
   labs(x = "Words produced", y = "Total number of TEs") +
-  geom_point(shape = 23) +
+  #geom_point(shape = 23) +
   geom_smooth(method="lm", se=F) +
@@
        colour = "Bins of balance") +
-  geom_point(shape = 23) +
+  #geom_point(shape = 23) +
   geom_smooth(method="lm", se=F) +
@@
   labs(x = "Words produced", y = "Total vocabulary size") +
-  geom_point(shape = 23) +
+  #geom_point(shape = 23) +
   geom_smooth(method="lm", se=F) +
@@
   labs(x = "Words produced", y = "Number of concepts") +
-  geom_point(shape = 23) +
+  #geom_point(shape = 23) +
   geom_smooth(method="lm", se=F) +
@@
 ```
+
+#Simulation3: Non-independence of D and N
+```{r}
+# I * (D*N/V)
+I1 = 1.5
+I2 = 0.5
+
+TE_computation <- TE_computation %>%
+  ## Calculate where I > 1
+  mutate(TE_easier = I1*(D*N/V)) %>%
+  ## Calculate where I < 1
+  mutate(TE_harder = I2*(D*N/V))
+
+Simulation3 <- TE_computation %>%
+  ## Calculate where I > 1
+  mutate(TE_easier = I1*(D*N/V)) %>%
+  ## Calculate where I < 1
+  mutate(TE_harder = I2*(D*N/V)) %>%
+  pivot_longer(c(TE, TE_harder, TE_easier), 
+               names_to = "Model", values_to = "number_TE") %>%
+  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
+  mutate(Model = recode(Model, 
+                        TE = "I = 1 (assuming independence of D and N)",
+                        TE_easier = "I = 1.5 (TEs are more easily learned)",
+                        TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
+  # Convert to long data
+  #pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  #filter(vocab_type == c("W", "D", "N")) %>%
+  #mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
+  #mutate(vocab_type = recode(vocab_type, 
+  #                           W = "Panel A: Total vocabulary (W)", 
+  #                           D = "Panel B: Dominant vocabulary (D)", 
+  #                           N = "Panel C: Non-dominant vocabulary (N)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ Model) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Total word vocabulary (W)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)"
+       #, title = "Number of translation equivalent by balance in relation to different input measures "
+       ) +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_Simulation3.png", Simulation3, 
+       width = 10, height = 5)
+
+Simulation3 <- TE_computation %>%
+  pivot_longer(c(TE_harder, TE_easier), 
+               names_to = "Nonindependence_type", values_to = "Simulated_TEs") %>%
+  mutate(Nonindependence_type = recode(Nonindependence_type, 
+                                       TE_easier = "I = 1.5 (TEs are more easily learned)",
+                                       TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
+  ggplot(aes(y = Simulated_TEs, x = TE, color = Nonindependence_type)) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  labs(y = "Simulated number of TEs assuming non-independence of D and N", 
+       x = "Expected number of TEs assuming independence of D and N",
+       color = "Scenarios"
+       #, title = "Non-independence vs. Independence of dominant and non-dominant words"
+       ) +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  theme_minimal()
+
+ggsave("plot_Simulation3.png", Simulation3, 
+       width = 8, height = 6)
+
+```

### 2021-02-26 Analyze_TEs_old.Rmd
Organized Simulations in order

--- 2021-02-01 Analyze_TEs.Rmd
@@
 
+write.csv(keepers_ws_TE, "Data/keepers_ws_TE.csv", row.names = F)
+
 # check balancedness cutoffs
@@
 
+```
+
+# Descriptives: Vocabulary measures
+```{r}
+## Mean word vocab size
+keepers_ws_TE %>% 
+  summarize(mean_W = mean(word_vocab, na.rm=T), 
+            sd_W = sd(word_vocab, na.rm=T),
+            min_W = min(word_vocab, na.rm=T),
+            max_W = max(word_vocab, na.rm=T))
+
+## Mean TE size
+keepers_ws_TE %>% 
+  summarize(mean_TE = mean(number_of_te, na.rm=T), 
+            sd_TE = sd(number_of_te, na.rm=T),
+            min_TE = min(number_of_te, na.rm=T),
+            max_TE = max(number_of_te, na.rm=T))
+
+## Mean unique word size
+keepers_ws_TE %>% 
+  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
+            sd_Du = sd(Dom_unique_words, na.rm=T),
+            min_Du = min(Dom_unique_words, na.rm=T),
+            max_Du = max(Dom_unique_words, na.rm=T))
+
+keepers_ws_TE %>% 
+  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
+            sd_Nu = sd(NonDom_unique_words, na.rm=T),
+            min_Nu = min(NonDom_unique_words, na.rm=T),
+            max_Nu = max(NonDom_unique_words, na.rm=T))
+
+t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
+       paired = TRUE) # paired-samples t-test to compare Du and Nu
+
+## Stacked barchart to plot TE, Du, and Nu of individual infants (facet by balance)
+stacked <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset balance_vocab by 3 levels 
+  mutate(BalSubset = cut(balance_vocab, breaks = 3, 
+                         labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
+  arrange(desc(number_of_te)) %>%
+  mutate(TE = number_of_te) %>%
+  pivot_longer(-c(baby_id, balance_vocab, BalSubset, word_vocab, TE), 
+               names_to = "Type", values_to = "NumberOfWords") %>%
+  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te")))
+
+LeastBal_plot <- stacked %>%.
+  filter(BalSubset == "LeastBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Least balanced") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+MediumBal_plot <- stacked %>%
+  filter(BalSubset == "MediumBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Medium balanced") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+MostBal_plot <- stacked %>%
+  filter(BalSubset == "MostBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Most balanced") + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+Bal_StackedPlot <- ggarrange(LeastBal_plot, 
+          MediumBal_plot + rremove("ylab"), 
+          MostBal_plot + rremove("ylab"),
+          ncol = 3, nrow = 1,
+          common.legend = TRUE, legend="bottom")
+
+ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
+       width = 10, height = 5)
+
+## Anova to compare the total vocab size among the 3 balance groups in the stacked barchart
+stacked_wide <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, concept_vocab, propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset balance_vocab by 3 levels 
+  mutate(BalSubset = cut(balance_vocab, breaks = 3, 
+                         labels=c("MostBalanced", "MediumBalanced", "LeastBalanced")))
+
+W_byB_anova <- aov(word_vocab ~ BalSubset, data = stacked_wide)
+summary(W_byB_anova)
+
+## Mean proportion of TE
+stacked_wide %>% 
+  filter(BalSubset == "LeastBalanced") %>%
+  summarize(mean_propTE = mean(propTE, na.rm=T), 
+            sd_propTE = sd(propTE, na.rm=T),
+            min_propTE = min(propTE, na.rm=T),
+            max_propTE = max(propTE, na.rm=T))
+
+stacked_wide %>% 
+  filter(BalSubset == "MediumBalanced") %>%
+  summarize(mean_propTE = mean(propTE, na.rm=T), 
+            sd_propTE = sd(propTE, na.rm=T),
+            min_propTE = min(propTE, na.rm=T),
+            max_propTE = max(propTE, na.rm=T))
+
+stacked_wide %>% 
+  filter(BalSubset == "MostBalanced") %>%
+  summarize(mean_propTE = mean(propTE, na.rm=T), 
+            sd_propTE = sd(propTE, na.rm=T),
+            min_propTE = min(propTE, na.rm=T),
+            max_propTE = max(propTE, na.rm=T))
+
+## Correlation between concept vocab and TE across different balance group
+stacked_wide_most <- stacked_wide %>%
+  filter(BalSubset == "MostBalanced")
+
+cor.test(stacked_wide_most$number_of_te, stacked_wide_most$concept_vocab, method = "pearson")
+
+stacked_wide_med <- stacked_wide %>%
+  filter(BalSubset == "MediumBalanced")
+
+cor.test(stacked_wide_med$number_of_te, stacked_wide_med$concept_vocab, method = "pearson")
+
+stacked_wide_least <- stacked_wide %>%
+  filter(BalSubset == "LeastBalanced")
+
+cor.test(stacked_wide_least$number_of_te, stacked_wide_least$concept_vocab, method = "pearson")
+
+## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 balance groups in the stacked barchart
+stacked_wide %>% 
+  group_by(BalSubset) %>%
+  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
+            sd_Du = sd(Dom_unique_words, na.rm=T),
+            min_Du = min(Dom_unique_words, na.rm=T),
+            max_Du = max(Dom_unique_words, na.rm=T))
+
+Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
+summary(Du_byB_anova)
+
+pairwise.t.test(stacked_wide$Dom_unique_words, stacked_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+stacked_wide %>% 
+  group_by(BalSubset) %>%
+  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
+            sd_Nu = sd(NonDom_unique_words, na.rm=T),
+            min_Nu = min(NonDom_unique_words, na.rm=T),
+            max_Nu = max(NonDom_unique_words, na.rm=T))
+
+Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
+summary(Nu_byB_anova)
+
+pairwise.t.test(stacked_wide$NonDom_unique_words, stacked_wide$BalSubset,
+                 p.adjust.method = "BH")
 ```
@@
   theme(plot.title = element_text(hjust = 0.5))
-
-ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
-       width = 10, height = 5)
-
-```
-
-```{r}
-stacked <- keepers_ws_TE %>%
-  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
-  arrange(desc(number_of_te)) %>%
-  mutate(TE = number_of_te) %>%
-  pivot_longer(-c(baby_id, balance_vocab, word_vocab, TE), 
-               names_to = "Type", values_to = "NumberOfWords") %>%
-  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
-  #Subset balance_vocab by 3 levels 
-  mutate(BalSubset = cut(balance_vocab, breaks = 3, 
-                         labels=c("HighBalance", "MidRangeBalance", "LeastBalance")))
-
-LeastBal_plot <- stacked %>%
-  filter(BalSubset == "LeastBalance") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = "Least balanced") + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-MidRangeBal_plot <- stacked %>%
-  filter(BalSubset == "MidRangeBalance") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = " ") + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-HighBal_plot <- stacked %>%
-  filter(BalSubset == "HighBalance") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = "High balance") + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-Bal_StackedPlot <- ggarrange(LeastBal_plot, 
-          MidRangeBal_plot + rremove("ylab"), 
-          HighBal_plot + rremove("ylab"),
-          ncol = 3, nrow = 1,
-          common.legend = TRUE, legend="bottom")
 

### 2021-02-27 computational model/TE_computation_Simulation.Rmd
revised simulation models

--- 2021-01-06 computational model/TE_computation.Rmd
@@
 library(ggpubr)
-library(regclass) # for correlation
-library(effects)
+#library(regclass) # for correlation
+#library(effects)
 # library(jtools)
-library(FRACTION) # this allows us to use the is.wholenumber function
+#library(FRACTION) # this allows us to use the is.wholenumber function
 
@@
 
-# Generating dataframe for computational model
-#### Set 1: Calculate other variables based on V, W, and B ####
+# Read in data
 ```{r}
-### Generating variables: V, W, and B
-## V = 600 ##
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 600
+TE_computation <- read.csv(here::here("computational model/TE_computation.csv"))
+```
 
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
+# Simulation 1
+## Infants of the same developmental level with different word vocabularies and balances of exposure
+### V = 600
+```{r}
+# Generating the parameters
+V <- 600 # setting V = 600 for the Simulation (keep V constant)
 
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
+W <- seq(0, 2*V, by = 25) # a range of total word vocabulary of a child
+
+B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
 
@@
 # Creating the dataframe
-TE_V600 <- data.frame(V, W_seq) %>%
+data_Simulation1 <- data.frame(V, W_seq) %>%
   arrange(W_seq) %>%
   rename(W = W_seq) %>%
-  mutate(B = B_seq)
+  mutate(B = B_seq) %>%
+  # Create other variables
+  mutate(D = B*W, # Words produced in the dominant language
+         N = (1-B)*W, # Words produced in non-dominant language
+         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
+  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
+  filter_at(vars(D, N), all_vars(. <= V)) %>%
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0)
 
-## V = 450 ##
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 450
+# Visualizing Simulation 1
+Simulation1 <- data_Simulation1 %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             W = "Panel A: Total vocabulary (W)", 
+                             D = "Panel B: Dominant vocabulary (D)", 
+                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
+  mutate(B = as.factor(B)) %>%
+  ggplot(aes(x = number, y = TE, color = B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=12)) +
+  labs(x = "Number of words", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)"
+       #, title = "Number of translation equivalent by balance in relation to different input measures "
+       ) +
+  guides(colour = guide_legend(nrow = 1))
 
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
+ggsave("plot_Simulation1.png", Simulation1,
+       width = 10, height = 5)
 
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
+```
 
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
+# Simulation 2
+## Patterns of translation equivalent learning as a function of different developmental timepoints
+### V = 150, 300, 450, or 600
+```{r}
+# V = 600
+## Generating the parameters
+V <- 600
+W <- V # holding W constant
+B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
+
+## Generating sequence of W and B
+W_seq <- W %>% 
   rep(., length(B)) # generate N repetitions of the sequence based on length of B
 
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+## Creating the dataframe
+TE_V600 <- data.frame(V, W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B)
 
-# Creating the dataframe
-TE_V450 <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B_seq)
+# V = 450
+## Generating the parameters
+V <- 450
+W <- V # holding W constant
+B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
 
-## V = 300 ##
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 300
-
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
-
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
-
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
+## Generating sequence of W and B
+W_seq <- W %>% 
   rep(., length(B)) # generate N repetitions of the sequence based on length of B
 
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+## Creating the dataframe
+TE_V450 <- data.frame(V, W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B)
 
-# Creating the dataframe
-TE_V300 <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B_seq)
+# V = 300
+## Generating the parameters
+V <- 300
+W <- V # holding W constant
+B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
 
-## V = 150 ##
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 150
-
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
-
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
-
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
+## Generating sequence of W and B
+W_seq <- W %>% 
   rep(., length(B)) # generate N repetitions of the sequence based on length of B
 
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+## Creating the dataframe
+TE_V300 <- data.frame(V, W_seq) %>%
+  rename(W = W_seq) %>%
+  mutate(B = B)
 
-# Creating the dataframe
+# V = 150
+## Generating the parameters
+V <- 150
+W <- V # holding W constant
+B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
+
+## Generating sequence of W and B
+W_seq <- W %>% 
+  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+
+## Creating the dataframe
 TE_V150 <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
   rename(W = W_seq) %>%
-  mutate(B = B_seq)
+  mutate(B = B)
 
-### Combining all dataframes and Calculate other variables
-TE_fromVWB <- TE_fromVWB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
+# Combining all the dataframes
+data_Simulation2 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # Create other variables
@@
          TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
          Du = D-TE, # Unique words in dominant language
          Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= V)) %>%
-  # filter out cases where D or N aren't whole numbers
-  filter_at(vars(D, N), any_vars(is.wholenumber(.))) 
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0)
+
+# Visualizing Simulation 2
+Simulation2 <- data_Simulation2 %>%
+  pivot_longer(-c(V, W, B), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  group_by(V, B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), 
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5),
+        text = element_text(size=12)) + 
+  labs(x = "Balance", 
+       y = "Number of concepts", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
+
+ggsave("plot_Simulation2.png", Simulation2,
+        width = 10, height = 5)
+
 ```
 
-#### Set 2: Calculate other variables based on V, D, and N ####
+# Simulation 3
+## Non-independence of dominant-language (D) and non-dominant language words (N)
 ```{r}
-### Generating variables: D and N
-## V = 600 ##
+# V = 600
+## Generating the parameters
 V <- 600
@@
 
-# Generating sequence of D and N
+## Generating sequence of D and N
 D_seq <- D %>% # Words produced in the dominant language
@@
 
-# Creating the dataframe
+## Creating the dataframe
 TE_V600 <- data.frame(V, D_seq) %>%
@@
 
-## V = 450 ##
+# V = 450
+## Generating the parameters
 V <- 450
@@
 
-# Creating the dataframe
+## Creating the dataframe
 TE_V450 <- data.frame(V, D_seq) %>%
@@
 
-## V = 300 ##
+# V = 300
+## Generating the parameters
 V <- 300
@@
 
-# Creating the dataframe
+## Creating the dataframe
 TE_V300 <- data.frame(V, D_seq) %>%
@@
 
-## V = 150 ##
+# V = 150
+## Generating the parameters
 V <- 150
@@
 
-# Creating the dataframe
+## Creating the dataframe
 TE_V150 <- data.frame(V, D_seq) %>%
@@
 
-### Combining all dataframes and Calculate other variables
-TE_fromVDN <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
+# Combining all the dataframes
+data_Simulation3 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
   # It is not possible for N > D, so keep only cases where D >= N
   filter(D >= N) %>%
-  # Create other variables
+  # Create TE 
   mutate(W = D+N, # Total word vocabulary of a child
          B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>%
-  # Remove cases where B = NA
-  filter(!is.na(B)) 
-```
+         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
+  # Remove cases where B = NaN (those were all cases where both D and N are 0)
+  filter(!is.na(B)) %>%
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0)
 
-#### Set 3: Calculate other variables based on V, D, and B ####
-```{r}
-### Creating variables: D and B
-## V = 600 #
-V <- 600
-D <- seq(0, V, by = 25)
+# Simulation3: TE = I * (D*N/V)
+## Defining the independence parameter I
+I1 = 1.5 # I > 1, TEs are MORE easily learned than unique words
+I2 = 0.5 # I < 1, TEs are LESS easily learned than unique words
 
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-# Creating the dataframe
-TE_V600 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq)
-
-## V = 450 #
-V <- 450
-D <- seq(0, V, by = 25)
-
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-# Creating the dataframe
-TE_V450 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq)
-
-## V = 300 #
-V <- 300
-D <- seq(0, V, by = 25)
-
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-# Creating the dataframe
-TE_V300 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq)
-
-## V = 150 #
-V <- 150
-D <- seq(0, V, by = 25)
-
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-# Creating the dataframe
-TE_V150 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq)
-
-### Combining all dataframes and Calculate other variables
-TE_fromVDB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
-  # Create other variables
-  mutate(W = D/B, # Total word vocabulary of a child
-         N = W-D, # Words produced in non-dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= V)) %>%
-  # filter out cases where N isn't whole number
-  filter(is.wholenumber(N))
-```
-
-#### Set 4: Calculate other variables based on V, N, and B ####
-```{r}
-### Generating variables: N and B
-## V = 600 #
-V <- 600
-N <- seq(0, V, by = 25)
-
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-# Creating the dataframe
-TE_V600 <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq)
-
-## V = 450 #
-V <- 450
-N <- seq(0, V, by = 25)
-
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-# Creating the dataframe
-TE_V450 <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq)
-
-## V = 300 #
-V <- 300
-N <- seq(0, V, by = 25)
-
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-# Creating the dataframe
-TE_V300 <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq)
-
-## V = 150 #
-V <- 150
-N <- seq(0, V, by = 25)
-
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-# Creating the dataframe
-TE_V150 <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq)
-
-### Combining all dataframes and Calculate other variables
-TE_fromVNB <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
-  # Create other variables
-  mutate(W = D/(1-B), # Total word vocabulary of a child
-         D = W-N, # Words produced in dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= V)) %>%
-  # filter out cases where D isn't whole number
-  filter(is.wholenumber(D)) %>%
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
-  filter(is.wholenumber(Nu)) 
-```
-
-#### Combine the 4 sets ####
-```{r}
-TE_computation <- TE_fromVWB %>%
-  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
-  # constrain all variables but B (balance) to be integers 
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
-```
-
-# Correlation
-```{r}
-Correlation <- TE_computation %>%
-  cor_matrix()
-
-library(Hmisc)
-mycor <- rcorr(as.matrix(TE_computation), type="pearson")
-mycor$r
-mycor$P
-
-library(corrplot)
-mycor <- cor(TE_computation)
-corrplot(mycor)
-#write.table(Correlation, "TEcomputation_correlations.tsv", col.names=NA)
-```
-
-# Plotting relations between variables
-## Stacked barchart showing the vocabulary composition across different balance proportions
-```{r}
-
-#plot_composition <- TE_computation %>%
-#  # Convert to long data
-#  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-#  filter(vocab_type %notin% c("D", "N", "C")) %>%
-#  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
-#  mutate(bin_W = cut(W, breaks = 3, 
-#                     labels = c("0-400", "401-800", "801-1200"))) %>%
-#  group_by(bin_W, vocab_type, B) %>%
-#  summarise(n_words = mean(number)) %>%
-#  ggplot(aes(x=B, y = n_words, fill = vocab_type)) + 
-#  geom_bar(position="stack", stat="identity") +
-#  facet_grid(. ~ bin_W) +
-#  #scale_y_continuous(limits = c(0, 1200)) +
-#  theme_minimal() +
-#  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
-#  labs(x = "Balance", 
-#       y = "Number of words")
-
-plot_composition_byB <- TE_computation %>%
-  # Convert to long data
-  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %notin% c("D", "N", "C")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+## Visualizing the simulation
+Simulation3 <- data_Simulation3 %>%
+  ## Calculate where I > 1
+  mutate(TE_easier = I1*(D*N/V)) %>%
+  ## Calculate where I < 1
+  mutate(TE_harder = I2*(D*N/V)) %>%
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0) %>%
+  # Prepare dataset to plot
+  pivot_longer(c(TE, TE_harder, TE_easier), 
+               names_to = "Model", values_to = "number_TE") %>%
+  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
+  mutate(Model = recode(Model, 
+                        TE = "I = 1 (assuming independence of D and N)",
+                        TE_easier = "I = 1.5 (TEs are more easily learned)",
+                        TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
   mutate(bin_B = cut(B, breaks = 6, 
                      labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  #mutate(bin_W = cut_interval(W, 12,
-  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
-  mutate(bin_W = cut_interval(W, 8,
-                     labels = c(150, 300, 450, 600, 750, 900, 1050, 1200))) %>%
-  group_by(bin_B, vocab_type, bin_W) %>%
-  summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = bin_W, y = n_words, fill = vocab_type)) + 
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ bin_B) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  #scale_y_continuous(limits = c(0, 1200)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
-        plot.title = element_text(hjust = 0.5)) +
-  labs(x = "Total vocabulary size", 
-       y = "Number of concepts", 
-       title = "Vocabulary composition across different balance proportions")
-
-ggsave("plot_composition_byB.png", plot_composition_byB)
-
-plot_composition_byW <- TE_computation %>%
-  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %notin% c("D", "N", "C")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  mutate(bin_W = cut_interval(W, 6,
-                     labels = c(200, 400, 600, 800, 1000, 1200))) %>%
-  group_by(bin_B, vocab_type, bin_W) %>%
-  summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ bin_W) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  #scale_y_continuous(limits = c(0, 1200)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
-  labs(x = "Balance", 
-       y = "Number of concepts", 
-       title = "Vocabulary composition across different vocabulary size") 
-```
-
-## Compare different vocabulary measures across balance
-```{r}
-vocab_label <- c("Concept Vocabulary (C)", "Unique Dominant Words (Du)", "Unique Non-dominant Words (Nu)")
-
-plot_words <- TE_computation %>%
-  # Convert to long data
-  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %notin% c("D", "N")) %>%
-  mutate(vocab_type = as.factor(vocab_type)) %>%
-  mutate(vocab_type = recode(vocab_type, C = "Concept vocabulary (C)", Du = "Unique dominant words (Du)",
-         Nu = "Unique non-dominant words (Nu)", TE = "Translation equivalents")) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  #mutate(bin_W = cut_interval(W, 12,
-  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
-  group_by(bin_B, vocab_type, W) %>%
-  summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = W, y = n_words, color = bin_B)) +
-  #geom_point() +
+  # ggplot
+  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
   stat_smooth(method = lm, se = F) +
   theme_minimal() + 
-  facet_grid(. ~ vocab_type) +
+  facet_grid(. ~ Model) +
   ylim(0, 600) +
@@
         legend.position = "bottom") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  labs(x = "Total vocabulary Size (W)", 
-       y = "Number of words", 
-       color = "Balance",
-       title = "Word vocabulary size by balance under different vocabulary measures") +
+  theme(plot.title = element_text(hjust = 0.5),
+        text = element_text(size=12)) +
+  labs(x = "Total word vocabulary (W)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
   guides(colour = guide_legend(nrow = 1))
 
-ggsave("plot_words.png", plot_words,
+ggsave("plot_Simulation3.png", Simulation3, 
        width = 10, height = 5)
-```
-
-## Plotting relations between TEs and different variables
-##### Relation between TE & W, D, & N #####
-```{r}
-### Plot: Y = TE, X = D / N
-plot_TEandWDN <- TE_computation %>%
-  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("W", "D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
-  #mutate(bin_B = cut(B, breaks = 6, 
-  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = TE)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Total number of TEs") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_y_continuous(limits = c(0, 600)) +
-  theme_minimal()
-
-ggsave("plot_TEandWDN.png", plot_TEandWDN)
-
-### Plot: Y = TE, X = D / N, color by B
-plot_TEandWDN_byB <- TE_computation %>%
-  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("W", "D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = TE, colour = bin_B)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Total number of TEs",
-       colour = "Bins of balance") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_x_continuous(limits = c(0, 1200)) +
-  scale_y_continuous(limits = c(0, 600)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
-  guides(colour = guide_legend(nrow = 1))
-
-ggsave("plot_TEandWDN_byB.png", plot_TEandWDN_byB)
-```
-
-##### Relation between W & TE, Du, & Nu #####
-```{r}
-### Plot: Y = W, X = TE/ Du / Nu
-plot_WandTEDuNu <- TE_computation %>%
-  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
-  #mutate(bin_B = cut(B, breaks = 6, 
-  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = W)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Total vocabulary size") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_y_continuous(limits = c(0, 1200)) +
-  theme_minimal()
-
-ggsave("plot_WandTEDuNu.png", plot_WandTEDuNu)
-
-### Plot: Y = W, X = TE/ Du / Nu, color by B
-plot_WandTEDuNu_byB <- TE_computation %>%
-  pivot_longer(c(TE, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("TE", "Du", "Nu")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("TE", "Du", "Nu"))) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = W, colour = bin_B)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Total vocabulary size",
-       colour = "Bins of balance") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_y_continuous(limits = c(0, 1200)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
-  guides(colour = guide_legend(nrow = 1))
-
-ggsave("plot_WandTEDuNu_byB.png", plot_WandTEDuNu_byB)
-```
-
-##### Relation between C & W, D, & N #####
-```{r}
-### Plot: Y = C, X = W / D / N
-plot_CandWDN <- TE_computation %>%
-  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("W", "D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
-  #mutate(bin_B = cut(B, breaks = 6, 
-  #                   labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = C)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Number of concepts") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_y_continuous(limits = c(0, 600)) +
-  theme_minimal()
-
-ggsave("plot_CandWDN.png", plot_CandWDN)
-
-### Plot: Y = C, X = W / D / N, color by B
-plot_CandWDN_byB <- TE_computation %>%
-  pivot_longer(c(TE, W, D, N, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %in% c("W", "D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = number, y = C, colour = bin_B)) +
-  facet_grid(. ~vocab_type) +
-  labs(x = "Words produced", y = "Number of concepts",
-       colour = "Bins of balance") +
-  geom_point(shape = 23) +
-  geom_smooth(method="lm", se=F) +
-  scale_x_continuous(limits = c(0, 1200)) +
-  scale_y_continuous(limits = c(0, 600)) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
-  guides(colour = guide_legend(nrow = 1))
-
-ggsave("plot_CandWDN_byB.png", plot_CandWDN_byB)
-```
-
-# Plotting relations between variables (while holding some variables constant)
-## TE ~ predictors (input)
-```{r}
-#TE_Constant <- lm(TE ~ ., data = TE_computation) 
-#summary(TE_Constant)
-
-#### run a regression with TE ~ B, D, N
-TE_InputConstant <- lm(TE ~ 0 + B + D + N, data = TE_computation) 
-summary(TE_InputConstant)
-
-### plot the relation between TE & B / D / N
-#effect_plot(TE_InputConstant, pred = B, plot.points = TRUE)
-effect_B <- effect(term = "B", mod = TE_InputConstant) %>%
-  data.frame()
-
-plotConstant_TE_B <- ggplot(effect_B, aes(x = B, y = fit)) +
-  geom_line() +
-  ylim(-25, 600) +
-  labs(x = 'Balance', y = 'Number of TEs') +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_B.png", plotConstant_TE_B)
-
-### plot the relation between TE & D
-#effect_plot(TE_InputConstant, pred = D, plot.points = TRUE)
-effect_D <- effect(term = "D", mod = TE_InputConstant) %>%
-  data.frame()
-
-plotConstant_TE_D <- ggplot(effect_D, aes(x = D, y = fit)) +
-  geom_line() +
-  ylim(-25, 600) +
-  labs(x = 'Dominant vocabulary', y = 'Number of TEs') +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_D.png", plotConstant_TE_D)
-
-### plot the relation between TE & N
-#effect_plot(TE_InputConstant, pred = N, plot.points = TRUE)
-effect_N <- effect(term = "N", mod = TE_InputConstant) %>%
-  data.frame()
-
-plotConstant_TE_N <- ggplot(effect_N, aes(x = N, y = fit)) +
-  geom_line() +
-  ylim(-25, 600) +
-  labs(x = 'Non-dominant vocabulary', y = 'Number of TEs') +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_N.png", plotConstant_TE_D)
-
-## put the three plots together
-plotConstant_TE_BDN <- ggarrange(plotConstant_TE_B, plotConstant_TE_D, plotConstant_TE_N, 
-                                 ncol = 3, nrow = 1,
-                                 widths = 3, heights = 1) %>%
-  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + B + D + N", face = "bold", size = 14))
-
-ggsave("plotConstant_TE_BDN.png", plotConstant_TE_BDN, 
-       width = 10, height = 5)
-```
-
-## TE ~ predictors (output)
-```{r}
-#### run a regression with TE ~ C, Du, Nu
-TE_OutputConstant <- lm(TE ~ 0 + C + Du + Nu, data = TE_computation) 
-summary(TE_OutputConstant)
-
-## plot the relation between TE & C
-#effect_plot(TE_OutputConstant, pred = C, plot.points = TRUE)
-effect_C <- effect(term = "C", mod = TE_OutputConstant) %>%
-  data.frame()
-
-plotConstant_TE_C <- ggplot(effect_C, aes(x = C, y = fit)) +
-  geom_line() +
-  labs(x = 'Concept vocabulary', y = 'Number of TEs') +
-  ylim(-300, 600) +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_C.png", plotConstant_TE_C)
-
-## plot the relation between TE & Du
-#effect_plot(TE_OutputConstant, pred = Du, plot.points = TRUE)
-effect_Du <- effect(term = "Du", mod = TE_OutputConstant) %>%
-  data.frame()
-
-plotConstant_TE_Du <- ggplot(effect_Du, aes(x = Du, y = fit)) +
-  geom_line() +
-  labs(x = 'Unique dominant vocabulary', y = 'Number of TEs') +
-  ylim(-300, 600) +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)
-
-## plot the relation between TE & Nu
-#effect_plot(TE_OutputConstant, pred = Nu, plot.points = TRUE)
-effect_Nu <- effect(term = "Nu", mod = TE_OutputConstant) %>%
-  data.frame()
-
-plotConstant_TE_Nu <- ggplot(effect_Nu, aes(x = Nu, y = fit)) +
-  geom_line() +
-  labs(x = 'Unique non-dominant vocabulary', y = 'Number of TEs') +
-  ylim(-300, 600) +
-  xlim(0,600) +
-  theme_minimal()
-
-#ggsave("plotConstant_TE_Du.png", plotConstant_TE_Du)
-
-## put the three plots together
-plotConstant_TE_CDuNu <- ggarrange(plotConstant_TE_C, plotConstant_TE_Du, plotConstant_TE_Nu, 
-                                 ncol = 3, nrow = 1,
-                                 widths = 3, heights = 1) %>%
-  annotate_figure(top = text_grob("linear regression model: TE ~ 0 + C + Du + Nu", face = "bold", size = 14))
-
-ggsave("plotConstant_TE_CDuNu.png", plotConstant_TE_CDuNu, 
-       width = 10, height = 5)
-```
-
-# Estimating  if the models is biased towards ot against learning TEs
-```{r}
-## plot
-plot_learnVSpredict <- TE_computation %>%
-  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
-         learned_TE = (D*N)/(W/2)) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
-  ggplot(aes(x = learned_TE, y = predicted_TE, color = bin_B)) +
-  geom_point() +
-  geom_abline(intercept = 0, slope = 1) +
-  labs(x = "Learned translation equivalents (= (D*N)/(W/2))",
-       y = "Predicted translation equivalents under the model (= (D*N)/V)",
-       color = "Bins of balance") +                                                    
-  theme_minimal()
-
-ggsave("plot_learnVSpredict.png", plot_learnVSpredict)
-
-## Kolmogorov-Smirnov test to compare the observed and predicted TE
-TE_computation <- TE_computation %>%
-  mutate(predicted_TE = TE, #predicted TE is based on D*N/V (where V is the max. possible words)
-         learned_TE = (D*N)/(W/2))
-
-stats::ks.test(TE_computation$predicted_TE, TE_computation$learned_TE)
 

### 2021-03-22 computational model/TE_computation_Simulation.Rmd
updated simulation analysis

@@
 
-```
-
-# Read in data
-```{r}
-TE_computation <- read.csv(here::here("computational model/TE_computation.csv"))
 ```
@@
 W <- V # holding W constant
-B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
+B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
 
@@
          TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
+         DU = D-TE, # Unique words in dominant language
+         NU = N-TE) %>% # Unique words in non-dominant language
   # Round to keep whole numbers only
@@
 Simulation2 <- data_Simulation2 %>%
-  pivot_longer(-c(V, W, B), names_to = "vocab_type", values_to = "number") %>%
+  mutate(bin_B = case_when(B < 0.6 ~ "0.5",
+                           B >= 0.6 & B < 0.7 ~ "0.6",
+                           B >= 0.7 & B < 0.8 ~ "0.7",
+                           B >= 0.8 & B < 0.9 ~ "0.8",
+                           TRUE ~ "0.9")) %>%
+  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
   filter(vocab_type %notin% c("D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
-  group_by(V, B, vocab_type) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("NU", "DU", "TE"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
   summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = B, y = n_words, fill = vocab_type)) + 
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
   geom_bar(position="stack", stat="identity") +

### 2021-03-22 Analyze_TEs.Rmd
updated observed data analysis

--- 2020-12-07 Analyze_TEs.Rmd
@@
   mutate(propTE_CDI = number_of_te/672)
+
+write.csv(keepers_ws_TE, "Data/keepers_ws_TE.csv", row.names = F)
 
@@
 
+```
+
+# Figure 6
+```{r}
+Observed_plot_inputs_TE <- keepers_ws_TE %>% 
+  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, TotalNW_Dom, TotalNW_NonDom, balance_vocab) %>%
+  pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             word_vocab = "Panel A: Total vocabulary", 
+                             TotalNW_Dom = "Panel B: Dominant vocabulary", 
+                             TotalNW_NonDom = "Panel C: Non-dominant vocabulary")) %>%
+  mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
+                           balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
+                           balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
+                           balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
+                           TRUE ~ "0.9")) %>%
+  ggplot(aes(x = number, y = number_of_te, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  ylim(0, 400) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Number of words", 
+       y = "Total observed number of translation equivalents", 
+       color = "Balance"
+       #, title = "Number of translation equivalent by balance in relation to different input measures "
+       ) +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
+       width = 10, height = 5)
+
+```
+
+# Figure 7
+```{r}
+plot_stacked_byV <- keepers_ws_TE %>%
+  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
+  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
+                              TRUE ~ "28-33 months")) %>%
+  #mutate(V_subset = case_when(age_months_percentile <= 21 ~ "18-21",
+  #                            age_months_percentile > 21 & age_months_percentile <= 25 ~ "22-25",
+  #                            age_months_percentile > 25 & age_months_percentile <= 29 ~ "26-29",
+  #                            TRUE ~ "30-33")) %>%
+  #mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
+  #                         balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
+  #                         balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
+  #                         balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
+  #                         TRUE ~ "0.9")) %>%
+  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
+                               balance_vocab >= 0.80 ~ "Least Balanced",
+                               TRUE ~ "Medium Balanced")) %>%
+  mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
+  arrange(desc(number_of_te)) %>%
+  mutate(TE = number_of_te) %>%
+  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, V_subset, BalSubset, word_vocab, TE), 
+               names_to = "Type", values_to = "NumberOfWords") %>%
+  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
+  mutate(Type = recode(Type, 
+                       NonDom_unique_words = "Unique non-dominant word", 
+                       Dom_unique_words = "Unique dominant word", 
+                       number_of_te = "Translation equivalent")) %>%
+  group_by(V_subset, BalSubset, Type) %>%
+  summarise(n_words = mean(NumberOfWords)) %>%
+  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V_subset) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 0) , 
+        legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) + 
+  labs(x = "Balance", 
+       y = "Observed number of concepts", 
+       title = "Vocabulary composition across different subset of developmental timepoints")
+
+ggsave("Output/Observed_plot_stacked_byV.png", plot_stacked_byV,
+       width = 10, height = 5)
+
+## Mean proportion of TE
+stacked_byV_wide <- keepers_ws_TE %>%
+  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
+         propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
+  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
+                              TRUE ~ "28-33 months")) %>%
+  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
+                               balance_vocab >= 0.80 ~ "Least Balanced",
+                               TRUE ~ "Medium Balanced"))
+
+stacked_byV_wide %>%
+  group_by(V_subset) %>%
+  summarise(mean_propTE = mean(propTE, na.rm=T), 
+            sd_propTE = sd(propTE, na.rm=T),
+            min_propTE = min(propTE, na.rm=T),
+            max_propTE = max(propTE, na.rm=T))
+
+# Anova to compare the number of TEs among the 3 developmental groups in the stacked barchart
+stacked_byV_wide %>%  
+  group_by(V_subset) %>%
+  summarise(mean_TE = mean(number_of_te, na.rm=T), 
+            sd_TE = sd(number_of_te, na.rm=T),
+            min_TE = min(number_of_te, na.rm=T),
+            max_TE = max(number_of_te, na.rm=T))
+
+TE_byV_anova <- aov(propTE ~ V_subset, data = stacked_byV_wide)
+summary(TE_byV_anova)
+
+pairwise.t.test(stacked_byV_wide$number_of_te, stacked_byV_wide$V_subset,
+                 p.adjust.method = "BH")
+
+## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
+stacked_byV_wide %>%  
+  group_by(BalSubset) %>%
+  summarise(mean_Du = mean(Dom_unique_words, na.rm=T), 
+            sd_Du = sd(Dom_unique_words, na.rm=T),
+            min_Du = min(Dom_unique_words, na.rm=T),
+            max_Du = max(Dom_unique_words, na.rm=T))
+
+Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
+summary(Du_byB_anova)
+
+pairwise.t.test(stacked_byV_wide$Dom_unique_words, stacked_byV_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+stacked_byV_wide %>% 
+  group_by(BalSubset) %>%
+  summarise(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
+            sd_Nu = sd(NonDom_unique_words, na.rm=T),
+            min_Nu = min(NonDom_unique_words, na.rm=T),
+            max_Nu = max(NonDom_unique_words, na.rm=T))
+
+Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
+summary(Nu_byB_anova)
+
+pairwise.t.test(stacked_byV_wide$NonDom_unique_words, stacked_byV_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+```
+
+# Descriptives: Vocabulary measures
+```{r}
+## Mean word vocab size
+keepers_ws_TE %>% 
+  summarize(mean_W = mean(word_vocab, na.rm=T), 
+            sd_W = sd(word_vocab, na.rm=T),
+            min_W = min(word_vocab, na.rm=T),
+            max_W = max(word_vocab, na.rm=T))
+
+## Mean TE size
+keepers_ws_TE %>% 
+  summarize(mean_TE = mean(number_of_te, na.rm=T), 
+            sd_TE = sd(number_of_te, na.rm=T),
+            min_TE = min(number_of_te, na.rm=T),
+            max_TE = max(number_of_te, na.rm=T))
+
+## Mean unique word size
+keepers_ws_TE %>% 
+  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
+            sd_Du = sd(Dom_unique_words, na.rm=T),
+            min_Du = min(Dom_unique_words, na.rm=T),
+            max_Du = max(Dom_unique_words, na.rm=T))
+
+keepers_ws_TE %>% 
+  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
+            sd_Nu = sd(NonDom_unique_words, na.rm=T),
+            min_Nu = min(NonDom_unique_words, na.rm=T),
+            max_Nu = max(NonDom_unique_words, na.rm=T))
+
+t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
+       paired = TRUE) # paired-samples t-test to compare Du and Nu
+
+## Stacked barchart to plot TE, Du, and Nu of individual infants (facet by balance)
+stacked <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset balance_vocab by 3 levels 
+  #mutate(BalSubset = cut(balance_vocab, breaks = 3, 
+  #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
+  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
+                               balance_vocab >= 0.80 ~ "LeastBalanced",
+                               TRUE ~ "MediumBalanced")) %>%
+  arrange(desc(number_of_te)) %>%
+  mutate(TE = number_of_te) %>%
+  pivot_longer(-c(baby_id, balance_vocab, BalSubset, word_vocab, TE), 
+               names_to = "Type", values_to = "NumberOfWords") %>%
+  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
+  mutate(Type = recode(Type, 
+                       NonDom_unique_words = "Unique non-dominant word", 
+                       Dom_unique_words = "Unique dominant word", 
+                       number_of_te = "Translation equivalent"))
+
+LeastBal_plot <- stacked %>%
+  filter(BalSubset == "LeastBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Least balanced") + 
+  theme_minimal() + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+MediumBal_plot <- stacked %>%
+  filter(BalSubset == "MediumBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Medium balanced") + 
+  theme_minimal() + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+MostBal_plot <- stacked %>%
+  filter(BalSubset == "MostBalanced") %>%
+  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
+  geom_col(aes(fill = Type)) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_y_continuous(limits = c(0, 1050)) +
+  labs(x = "Individual infant", 
+       y = "Number of concepts produced",
+       title = "Most balanced") + 
+  theme_minimal() + 
+  theme(plot.title = element_text(hjust = 0.5))
+
+Bal_StackedPlot <- ggarrange(LeastBal_plot, 
+          MediumBal_plot + rremove("ylab"), 
+          MostBal_plot + rremove("ylab"),
+          ncol = 3, nrow = 1,
+          common.legend = TRUE, legend="bottom")
+
+ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
+       width = 10, height = 5)
+
+## Anova to compare the total vocab size among the 3 balance groups in the stacked barchart
+stacked_wide <- keepers_ws_TE %>%
+  select(baby_id, balance_vocab, word_vocab, concept_vocab, propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  # Subset balance_vocab by 3 levels 
+  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
+                               balance_vocab >= 0.80 ~ "LeastBalanced",
+                               TRUE ~ "MediumBalanced"))
+
+W_byB_anova <- aov(word_vocab ~ BalSubset, data = stacked_wide)
+summary(W_byB_anova)
+
+pairwise.t.test(stacked_wide$word_vocab, stacked_wide$BalSubset, p.adj = "BH")
+
+## Mean proportion of TE
+stacked_wide %>%
+  group_by(BalSubset) %>%
+  summarize(mean_propTE = mean(propTE, na.rm=T), 
+            sd_propTE = sd(propTE, na.rm=T),
+            min_propTE = min(propTE, na.rm=T),
+            max_propTE = max(propTE, na.rm=T))
+
+## Correlation between concept vocab and TE across different balance group
+stacked_wide_most <- stacked_wide %>%
+  filter(BalSubset == "MostBalanced")
+
+cor.test(stacked_wide_most$number_of_te, stacked_wide_most$concept_vocab, method = "pearson")
+
+stacked_wide_med <- stacked_wide %>%
+  filter(BalSubset == "MediumBalanced")
+
+cor.test(stacked_wide_med$number_of_te, stacked_wide_med$concept_vocab, method = "pearson")
+
+stacked_wide_least <- stacked_wide %>%
+  filter(BalSubset == "LeastBalanced")
+
+cor.test(stacked_wide_least$number_of_te, stacked_wide_least$concept_vocab, method = "pearson")
+
+## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 balance groups in the stacked barchart
+stacked_wide %>% 
+  group_by(BalSubset) %>%
+  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
+            sd_Du = sd(Dom_unique_words, na.rm=T),
+            min_Du = min(Dom_unique_words, na.rm=T),
+            max_Du = max(Dom_unique_words, na.rm=T))
+
+Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
+summary(Du_byB_anova)
+
+pairwise.t.test(stacked_wide$Dom_unique_words, stacked_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+stacked_wide %>% 
+  group_by(BalSubset) %>%
+  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
+            sd_Nu = sd(NonDom_unique_words, na.rm=T),
+            min_Nu = min(NonDom_unique_words, na.rm=T),
+            max_Nu = max(NonDom_unique_words, na.rm=T))
+
+Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
+summary(Nu_byB_anova)
+
+pairwise.t.test(stacked_wide$NonDom_unique_words, stacked_wide$BalSubset,
+                 p.adjust.method = "BH")
 ```
@@
   # compute estimate total vocab by adding the two languages
-  mutate(Total_90percentile = EngWS_90percentile + FrWS_90percentile) %>%
+  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
   # keep only necessary columns
-  select(age, EngWS_90percentile, FrWS_90percentile, Total_90percentile) %>%
+  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
   # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
   rename(age_months_percentile = age)
+
+write.csv(ws_percentile, "Output/ws_percentile.csv", row.names = F)
 ```
@@
   # merge with the percentile data frame
-  left_join(ws_percentile, by = "age_months_percentile") %>%
+  left_join(ws_percentile, by = "age_months_percentile")
+
+%>%
   # calculate TE in relation to CDI percentile data
@@
 
-## Create predicted TEs for simulation
+# Correlation
 ```{r}
-keepers_ws_TE_simulation <- keepers_ws_TE %>%
-  # Predicted propTE (Simulation)
-  mutate(propDomVocab = TotalNW_Dom / 672, # Proportion of Dom/NonDom vocab
-         propNonDomVocab = TotalNW_NonDom / 672, 
-         Predicted_propTE = propDomVocab * propNonDomVocab, # Predicted proportion of TE
-         Diff = propTE - Predicted_propTE) %>% # Difference between Predicted and Observed propTE
-  # Predicted raw TE (with reference to CDI)
-  mutate(Predicted_TotalnumberTEs_CDI = (TotalNW_Dom * TotalNW_NonDom)/672) %>%
-  # Predicted raw TE (with reference to CDI percentile)
-  mutate(Predicted_TotalnumberTEs_percentile = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)) %>%
-  # Create two groups based on observed TE
-  mutate(TEgroup = ifelse(number_of_te < 100, "Observed TE less than 100", "Observed TE more than 100"))
+Correlation <- keepers_ws_TE %>%
+  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
+         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
+  # recode the balance_vocab so that it would be in descending order, so the value of the most balanced would be the greatest
+  mutate(balance_vocab = balance_vocab*-1) %>%
+  # rename variables
+  rename(Age_month = age_continuous,
+         V = Average_90percentile,
+         B = balance_vocab, 
+         W = word_vocab,
+         D = TotalNW_Dom, 
+         N = TotalNW_NonDom, 
+         TE = number_of_te, 
+         Du = Dom_unique_words, 
+         Nu = NonDom_unique_words, 
+         C = concept_vocab)
 
-  #select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
-  #       propTE, propTE_CDI, propTE_percentile, Predicted_propTE, Diff, 
-  #       Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile)
+library(Hmisc)
+mycor <- rcorr(as.matrix(Correlation), type="pearson")
+mycor$r
+round(mycor$P, 2)
+
+library(corrplot)
+mycor <- round(cor(Correlation), 2)
+corrplot(mycor)
+write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
+
+cor.test(Correlation$TE, Correlation$B, method = "pearson")
+cor.test(Correlation$TE, Correlation$W, method = "pearson")
+cor.test(Correlation$TE, Correlation$D, method = "pearson")
+cor.test(Correlation$TE, Correlation$N, method = "pearson")
+cor.test(Correlation$TE, Correlation$V, method = "pearson")
+cor.test(Correlation$Du, Correlation$B, method = "pearson")
+cor.test(Correlation$Nu, Correlation$B, method = "pearson")
 
@@
 
-###### Linear regressions of predicted TE models ###### 
+###### Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) ###### 
 ```{r}
+## Create predicted TEs for simulation
+keepers_ws_TE_simulation <- keepers_ws_TE %>%
+  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile)
 
-### Model 1: Predicted_TotalnumberTEs_percentile
+## Linear regressions of predicted TE models
 Percentile_model <- keepers_ws_TE_simulation %>%
-  lm(number_of_te*(Total_90percentile/2) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
      data = .)
 
-### Model 2: Predicted_TotalnumberTEs_CDI
-CDI_model <- keepers_ws_TE_simulation %>%
-  lm(number_of_te * 672 ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
+summary(Percentile_model)
 
-anova(Percentile_model, CDI_model)
-```
-
-###### Theoretical distribution of TEs (Simulation) ###### 
-```{r}
-###Prediction 1: Predicted raw TE = (TotalNW_Dom*TotalNW_NonDom)/Total CDI
-# Ploting Y = observed rawTE  & X = Predicted raw TE
-
-keepers_ws_TE_simulation %>%
-  ggplot(aes(x = Predicted_TotalnumberTEs_CDI, y = number_of_te)) +
-  facet_grid(. ~TEgroup) +
-  labs(x = "Predicted total number of TEs (based on CDI)", y = "Observed total number of TEs",
-       title = "Observed vs. Predicted total number of TEs") +
+## Ploting Y = observed rawTE  & X = Predicted rawTE 
+TEprediction_percentile <- keepers_ws_TE_simulation %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+#  facet_grid(. ~TEgroup) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   xlim(-10,430) +
-  ylim(-10,430)
+  ylim(-10,430) +
+  theme_minimal()
+
+ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
+       height=5, width=5)
 
 # Kolmogorov-Smirnov test to compare the observed and predicted TE
-stats::ks.test(keepers_ws_TE_simulation$number_of_te,
-               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_CDI)
+#stats::ks.test(keepers_ws_TE_simulation$number_of_te,
+#               keepers_ws_TE_simulation$Predicted_TE_percentile)
 
-###Prediction 2: Predicted raw TE = TotalNW_Dom * TotalNW_NonDom/(Total_90percentile/2)
-# Ploting Y = observed rawTE  & X = Predicted raw TE
+## Looking closer at the infants with total vocabulary < 150
 
-TEprediction_percentile <- keepers_ws_TE_simulation %>%
-  ggplot(aes(x = Predicted_TotalnumberTEs_percentile, y = number_of_te)) +
-  facet_grid(. ~TEgroup) +
-  labs(x = "Predicted total number of TEs (based on percentile)", y = "Observed total number of TEs",
-       title = "Observed vs. Predicted total number of TEs") +
+keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300)
+
+stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
+               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)
+
+keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300)
+
+stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
+               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)
+
+# Plot
+TEprediction_percentile_facetW <- keepers_ws_TE_simulation %>%
+  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  facet_wrap(~ W_group, scales = "free") +
+  #facet_grid(. ~ W_group, scales = "free") +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
   theme(plot.title = element_text(hjust = 0.5)) +
@@
   geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,430) +
-  ylim(-10,430)
+  theme_minimal()
 
-ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile)
+ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetW,
+       height=5,width=10)
 
-# Kolmogorov-Smirnov test to compare the observed and predicted TE
-stats::ks.test(keepers_ws_TE_simulation$number_of_te,
-               keepers_ws_TE_simulation$Predicted_TotalnumberTEs_percentile)
+Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
 
-keepers_ws_TE_simulation_LessThan10 <- keepers_ws_TE_simulation %>%
-  filter(TEgroup == "Observed TE less than 100")
+summary(Percentile_model_LessThan300W)
 
-stats::ks.test(keepers_ws_TE_simulation_LessThan10$number_of_te,
-               keepers_ws_TE_simulation_LessThan10$Predicted_TotalnumberTEs_percentile)
+Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
 
-keepers_ws_TE_simulation_MoreThan10 <- keepers_ws_TE_simulation %>%
-  filter(TEgroup == "Observed TE more than 100")
-
-stats::ks.test(keepers_ws_TE_simulation_MoreThan10$number_of_te,
-               keepers_ws_TE_simulation_MoreThan10$Predicted_TotalnumberTEs_percentile)
-
-### Plotting the 2 predictions together
-
-TEprediction_percentile_CDI <- keepers_ws_TE_simulation %>%
-  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile, TEgroup) %>%
-  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
-                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
-  pivot_longer(-c(baby_id, number_of_te, TEgroup), 
-               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
-  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
-  facet_grid(. ~TEgroup) +
-  labs(x = "Predicted total number of TEs", 
-       y = "Observed total number of TEs",
-       title = 
-       "Observed vs. Predicted total number of TEs") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,500) +
-  ylim(-10,500) +
-  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") +
-  guides(colour = guide_legend(nrow = 1))
-
-ggsave(here("Output/TEprediction_percentile&CDI.png"), TEprediction_percentile_CDI)
-
-#Less Than 100 observed TEs
-TEprediction_percentile_CDI_LessThan100 <- keepers_ws_TE_simulation %>%
-  filter(TEgroup == "Observed TE less than 100") %>%
-  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
-  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
-                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
-  pivot_longer(-c(baby_id, number_of_te), 
-               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
-  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
-  labs(x = "Predicted total number of TEs", 
-       y = "Observed total number of TEs",
-       title = 
-       "Observed vs. Predicted total number of TEs \n (Less than 100 observed TEs)") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,100) +
-  ylim(-10,100) +
-  theme(plot.title = element_text(size = 12, face = "bold"), 
-        axis.text.x = element_text(angle = 90), 
-        legend.position = "bottom", legend.text=element_text(size = 10)) +
-  guides(colour = guide_legend(nrow = 2))
-
-ggsave(here("Output/TEprediction_percentile&CDI_LessThan100.png"), TEprediction_percentile_CDI_LessThan100)
-
-#Less Than 50 observed TEs
-TEprediction_percentile_CDI_LessThan50 <- keepers_ws_TE_simulation %>%
-  filter(number_of_te < 50) %>%
-  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
-  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
-                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
-  pivot_longer(-c(baby_id, number_of_te), 
-               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
-  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
-  labs(x = "Predicted total number of TEs", 
-       y = "Observed total number of TEs",
-       title = 
-       "Observed vs. Predicted total number of TEs \n (Less than 50 observed TEs)") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,50) +
-  ylim(-10,50) +
-  theme(plot.title = element_text(size = 12, face = "bold"),
-        axis.text.x = element_text(angle = 90), 
-        legend.position = "bottom", legend.box="vertical", legend.text=element_text(size = 10)) +
-  guides(colour = guide_legend(nrow = 2))
-
-ggsave(here("Output/TEprediction_percentile&CDI_LessThan50.png"), TEprediction_percentile_CDI_LessThan50)
-
-# More than 100 Observed TEs
-TEprediction_percentile_CDI_MoreThan100 <- keepers_ws_TE_simulation %>%
-  filter(TEgroup == "Observed TE more than 100") %>%
-  select(baby_id, number_of_te, Predicted_TotalnumberTEs_CDI, Predicted_TotalnumberTEs_percentile) %>%
-  dplyr::rename("=Dom*NonDom/TotalCDI" = Predicted_TotalnumberTEs_CDI,
-                "=Dom*NonDom/Percentile" = Predicted_TotalnumberTEs_percentile) %>%
-  pivot_longer(-c(baby_id, number_of_te), 
-               names_to = "Predicted_TotalnumberTEs", values_to = "PredictedNumber") %>%
-  ggplot(aes(x = PredictedNumber, y = number_of_te, color = Predicted_TotalnumberTEs)) +
-  labs(x = "Predicted total number of TEs", 
-       y = "Observed total number of TEs",
-       title = "Observed vs. Predicted total number of TEs \n (More than 100 observed TEs)") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,500) +
-  ylim(-10,500) +
-  theme(plot.title = element_text(size = 12, face = "bold"),
-        axis.text.x = element_text(angle = 90), 
-        legend.position = "bottom", legend.text=element_text(size=10)) +
-  guides(colour = guide_legend(nrow = 2))
-
-ggsave(here("Output/TEprediction_percentile&CDI_MoreThan100.png"), TEprediction_percentile_CDI_MoreThan100)
-
-###Prediction 2: Predicted raw TE = TotalNW_NonDom
-# Ploting Y = observed rawTE  & X = Predicted raw TE
-
-#keepers_ws_TE_simulation%>%
-#  ggplot(aes(x = TotalNW_NonDom, y = number_of_te)) +
-#  labs(x = "Predicted total number of TEs", 
-#       y = "Observed total number of TEs",
-#       title = 
-#       "Observed vs. Predicted total number of TEs 
-#       (based on total number of NonDom vocab)") +
-#  theme(plot.title = element_text(hjust = 0.5)) +
-#  geom_smooth(method = "lm", se = F) +
-#  geom_point(shape = 23) +
-#  geom_abline(linetype = 2, alpha = 0.5) +
-#  xlim(-10,500) +
-#  ylim(-10,500)
-
-## Kolmogorov-Smirnov test to compare the observed and predicted TE
-#stats::ks.test(keepers_ws_TE_simulation$TotalnumberTEs, keepers_ws_TE_simulation$TotalNW_NonDom)
+summary(Percentile_model_MoreThan300W)
 

### 2021-03-22 computational model/TE_computation_Simulation.Rmd
adding spaces between ggplot facets

@@
         plot.title = element_text(hjust = 0.5),
-        text = element_text(size=12)) +
+        text = element_text(size=12),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
   labs(x = "Number of words", 
@@
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  #scale_y_continuous(limits = c(0, 1200)) +
   theme_minimal() +
@@
         plot.title = element_text(hjust = 0.5),
-        text = element_text(size=12)) + 
+        text = element_text(size=12),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) + 
   labs(x = "Balance", 
@@
   theme(plot.title = element_text(hjust = 0.5),
-        text = element_text(size=12)) +
+        text = element_text(size=12),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
   labs(x = "Total word vocabulary (W)", 

### 2021-03-22 computational model/TE_computation_Simulation.Rmd
changing I to IND

@@
 
-# Simulation3: TE = I * (D*N/V)
+# Simulation3: TE = IND * (D*N/V)
 ## Defining the independence parameter I
-I1 = 1.5 # I > 1, TEs are MORE easily learned than unique words
-I2 = 0.5 # I < 1, TEs are LESS easily learned than unique words
+IND1 = 1.5 # IND > 1, TEs are MORE easily learned than unique words
+IND2 = 0.5 # IND < 1, TEs are LESS easily learned than unique words
 
@@
 Simulation3 <- data_Simulation3 %>%
-  ## Calculate where I > 1
-  mutate(TE_easier = I1*(D*N/V)) %>%
-  ## Calculate where I < 1
-  mutate(TE_harder = I2*(D*N/V)) %>%
+  ## Calculate where IND > 1
+  mutate(TE_easier = IND1*(D*N/V)) %>%
+  ## Calculate where IND < 1
+  mutate(TE_harder = IND2*(D*N/V)) %>%
   # Round to keep whole numbers only
@@
   mutate(Model = recode(Model, 
-                        TE = "I = 1 (assuming independence of D and N)",
-                        TE_easier = "I = 1.5 (TEs are more easily learned)",
-                        TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
+                        TE = "IND = 1 (assuming independence of D and N)",
+                        TE_easier = "IND = 1.5 (TEs are more easily learned)",
+                        TE_harder = "IND = 0.5 (TE are less easily learned)")) %>%
   mutate(bin_B = cut(B, breaks = 6, 

### 2021-03-26 computational model/TE_computation_Simulation.Rmd
updated simulation (generating from D & N)

@@
 ```{r}
+V = 600
+
 # Generating the parameters
-V <- 600 # setting V = 600 for the Simulation (keep V constant)
+W <- V # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W1 <- data.frame(V, W, N)
 
-W <- seq(0, 2*V, by = 25) # a range of total word vocabulary of a child
+W <- V-100 # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W2 <- data.frame(V, W, N)
 
-B <- seq(0.5, 1, by = 0.1) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
+W <- V-100*2 # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W3 <- data.frame(V, W, N)
 
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+W <- V-100*3 # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W4 <- data.frame(V, W, N)
 
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
+W <- V-100*4 # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W5 <- data.frame(V, W, N)
+
+W <- V-100*5 # a range of total word vocabulary of a child
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_W6 <- data.frame(V, W, N)
 
 # Creating the dataframe
-data_Simulation1 <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B_seq) %>%
+data_Simulation1 <- rbind(TE_W1, TE_W2, TE_W3, TE_W4, TE_W5, TE_W6) %>%
+  # Get dominant words D
+  mutate(D = W - N) %>%
   # Create other variables
-  mutate(D = B*W, # Words produced in the dominant language
-         N = (1-B)*W, # Words produced in non-dominant language
+  mutate(B = N/(D+N), # balance,
          TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= V)) %>%
   # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0)
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 6 groups
+  mutate(bin_B = case_when(B == 0.0 ~ "0",
+                           B > 0 & B <= 0.1 ~ "0.1",
+                           B > 0.1 & B <= 0.2 ~ "0.2",
+                           B > 0.2 & B <= 0.3 ~ "0.3", 
+                           B > 0.3 & B <= 0.4 ~ "0.4", 
+                           TRUE ~ "0.5")) %>%
+  arrange(bin_B)
 
@@
   mutate(vocab_type = recode(vocab_type, 
-                             W = "Panel A: Total vocabulary (W)", 
-                             D = "Panel B: Dominant vocabulary (D)", 
-                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
-  mutate(B = as.factor(B)) %>%
-  ggplot(aes(x = number, y = TE, color = B)) +
-  #geom_point() +
+                             W = "Panel 1A: Total vocabulary (W)", 
+                             D = "Panel 1B: Dominant vocabulary (D)", 
+                             N = "Panel 1C: Non-dominant vocabulary (N)")) %>%
+  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
   stat_smooth(method = lm, se = F) +
+  scale_color_manual(values=c("#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotted", "dotdash", "dashed", "twodash", "solid")) +
+  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
+  #scale_color_discrete(direction = -1) +
   theme_minimal() + 
   facet_grid(. ~ vocab_type) +
-  ylim(0, 600) +
   theme(axis.text.x = element_text(angle = 90),
         legend.position = "bottom",
-        plot.title = element_text(hjust = 0.5),
-        text = element_text(size=12),
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
         panel.spacing.x = unit(2, "lines"),
         panel.background = element_rect(fill = NA, color = "black")) +
+  ylim(0, 200) +
+  xlim(0, 700) +
   labs(x = "Number of words", 
        y = "Total number of translation equivalents (TE)", 
-       color = "Balance (B)"
-       #, title = "Number of translation equivalent by balance in relation to different input measures "
-       ) +
-  guides(colour = guide_legend(nrow = 1))
+       color = "Balance (B)",
+       linetype = "Balance (B)",
+       title = "1) Simulated data") +
+  guides(color = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
 
@@
        width = 10, height = 5)
-
 ```
@@
 ## Patterns of translation equivalent learning as a function of different developmental timepoints
-### V = 150, 300, 450, or 600
+### V = 200, 400, or 600
 ```{r}
@@
 W <- V # holding W constant
-B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_V600 <- data.frame(V, W, N)
 
-## Generating sequence of W and B
-W_seq <- W %>% 
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
+V <- 400
+W <- V # holding W constant
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_V400 <- data.frame(V, W, N)
 
-## Creating the dataframe
-TE_V600 <- data.frame(V, W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B)
+V <- 200
+W <- V # holding W constant
+N <- seq(0, W/2, by = 10) # a range of non-dominant words
+TE_V200 <- data.frame(V, W, N)
 
-# V = 450
-## Generating the parameters
-V <- 450
-W <- V # holding W constant
-B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
-
-## Generating sequence of W and B
-W_seq <- W %>% 
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-## Creating the dataframe
-TE_V450 <- data.frame(V, W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B)
-
-# V = 300
-## Generating the parameters
-V <- 300
-W <- V # holding W constant
-B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
-
-## Generating sequence of W and B
-W_seq <- W %>% 
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-## Creating the dataframe
-TE_V300 <- data.frame(V, W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B)
-
-# V = 150
-## Generating the parameters
-V <- 150
-W <- V # holding W constant
-B <- seq(0.5, 1, by = 0.01) # a range of Balance, i.e. proportion use of dominant language (B ≤ .50)
-
-## Generating sequence of W and B
-W_seq <- W %>% 
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-## Creating the dataframe
-TE_V150 <- data.frame(V, W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B)
-
-# Combining all the dataframes
-data_Simulation2 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
+# Creating the dataframe
+data_Simulation2 <- rbind(TE_V200, TE_V400, TE_V600) %>%
+  # Get dominant words D
+  mutate(D = W - N) %>%
   # Create other variables
-  mutate(D = B*W, # Words produced in the dominant language
-         N = (1-B)*W, # Words produced in non-dominant language
+  mutate(B = N/(D+N), # balance,
          TE = (D*N)/V, # If D and N are independent, TE = DN/V
@@
   # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0)
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 3 groups
+  mutate(bin_B = case_when(B >= 0.35 ~ "Most Balanced",
+                           B <= 0.20 ~ "Least Balanced",
+                           TRUE ~ "Medium Balanced")) %>%
+  mutate(bin_B  = fct_rev(bin_B)) %>%
+  arrange(bin_B)
 
@@
 Simulation2 <- data_Simulation2 %>%
-  mutate(bin_B = case_when(B < 0.6 ~ "0.5",
-                           B >= 0.6 & B < 0.7 ~ "0.6",
-                           B >= 0.7 & B < 0.8 ~ "0.7",
-                           B >= 0.8 & B < 0.9 ~ "0.8",
-                           TRUE ~ "0.9")) %>%
   pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
@@
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
+  ylim(0, 600) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 90), 
-        legend.position = "bottom",
-        plot.title = element_text(hjust = 0.5),
-        text = element_text(size=12),
-        panel.spacing.x = unit(2, "lines"),
-        panel.background = element_rect(fill = NA, color = "black")) + 
-  labs(x = "Balance", 
-       y = "Number of concepts", 
-       title = "Vocabulary composition across different developmental timepoints (V)")
+  theme(legend.position = "bottom",
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        panel.spacing.x = unit(2, "lines")) + 
+  labs(x = "\nBalance (B)", 
+       y = "Number of concepts (C)", 
+       title = "Panel A: Simulated data")
 
@@
         width = 10, height = 5)
-
 ```
@@
 V <- 600
-D <- seq(0, V, by = 25) 
-N <- seq(0, V, by = 25)
+W <- V
+N <- seq(0, W/2, by = 10)
+TE_V600 <- data.frame(V, W, N)
 
-## Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
+V <- 450
+W <- V
+N <- seq(0, W/2, by = 10)
+TE_V450 <- data.frame(V, W, N)
 
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
+V <- 300
+W <- V
+N <- seq(0, W/2, by = 10)
+TE_V300 <- data.frame(V, W, N)
 
-## Creating the dataframe
-TE_V600 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq)
+V <- 150
+W <- V
+N <- seq(0, W/2, by = 10)
+TE_V150 <- data.frame(V, W, N)
 
-# V = 450
-## Generating the parameters
-V <- 450
-D <- seq(0, V, by = 25) 
-N <- seq(0, V, by = 25)
+# Creating the dataframe
+data_Simulation3 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
+  # Get dominant words D
+  mutate(D = W - N) %>%
+  # Create other variables
+  mutate(B = N/(D+N), # balance,
+         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 6 groups
+  mutate(bin_B = case_when(B == 0.0 ~ "0",
+                           B > 0 & B <= 0.1 ~ "0.1",
+                           B > 0.1 & B <= 0.2 ~ "0.2",
+                           B > 0.2 & B <= 0.3 ~ "0.3", 
+                           B > 0.3 & B <= 0.4 ~ "0.4", 
+                           TRUE ~ "0.5")) %>%
+  arrange(bin_B)
 
-## Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_V450 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq)
-
-# V = 300
-## Generating the parameters
-V <- 300
-D <- seq(0, V, by = 25) 
-N <- seq(0, V, by = 25)
-
-## Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_V300 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq)
-
-# V = 150
-## Generating the parameters
-V <- 150
-D <- seq(0, V, by = 25) 
-N <- seq(0, V, by = 25)
-
-## Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_V150 <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq)
-
-# Combining all the dataframes
-data_Simulation3 <- rbind(TE_V150, TE_V300, TE_V450, TE_V600) %>%
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # Create TE 
-  mutate(W = D+N, # Total word vocabulary of a child
-         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
-  # Remove cases where B = NaN (those were all cases where both D and N are 0)
-  filter(!is.na(B)) %>%
-  # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0)
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0", "0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
 
@@
   # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0) %>%
+  mutate_at(vars(-c(B, bin_B)), round, 0) %>%
   # Prepare dataset to plot
@@
                         TE_harder = "IND = 0.5 (TE are less easily learned)")) %>%
-  mutate(bin_B = cut(B, breaks = 6, 
-                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
   # ggplot
-  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
+  ggplot(aes(x = W, y = number_TE, color = bin_B, linetype = bin_B)) +
   stat_smooth(method = lm, se = F) +
+  scale_color_manual(values=c("#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotted", "dotdash", "dashed", "twodash", "solid")) +
+  #scale_color_discrete(direction = -1) +
   theme_minimal() + 
   facet_grid(. ~ Model) +
-  ylim(0, 600) +
   theme(axis.text.x = element_text(angle = 90),
@@
         text = element_text(size=12),
-        panel.spacing.x = unit(2, "lines"),
+        panel.spacing.x = unit(2.5, "lines"),
         panel.background = element_rect(fill = NA, color = "black")) +
@@
        y = "Total number of translation equivalents (TE)", 
-       color = "Balance (B)") +
-  guides(colour = guide_legend(nrow = 1))
+       color = "Balance (B)",
+       linetype = "Balance (B)") +
+  guides(colour = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
 

### 2021-05-17 Analyze_TEs.Rmd
updated simulation generation and analysis

@@
 
-library(ggpirate)
 library(here)
@@
 library(wordbankr)
+library(ggplot2)
+library(patchwork)
+library(gridExtra)
 library("ggpubr")
@@
 library(scales)
+
+`%notin%` <- Negate(`%in%`)
+
 ```
@@
 ```{r}
-
-keepers_ws_TE = keepers_ws_TE %>%
+keepers_ws_TE <- keepers_ws_TE %>%
   # keep only necessary variables
@@
   mutate(age_month_decimal = age_days/30) %>%
-  # create balance groups (by exposure)
-  mutate(BalanceGroup_byExposure = cut(balance, breaks = 4, 
-                                       labels = c("80_20", "70_30", "60_40", "50_50"))) %>%
-  mutate(TotalNW_Dom_exposure = case_when(lang_dom == "English" ~ total_words_eng,
-                                          lang_dom == "French" ~ total_words_fre),
-         TotalNW_NonDom_exposure = case_when(lang_dom == "English" ~ total_words_fre,
-                                             lang_dom == "French" ~ total_words_eng),
-         Dom_unique_words_exposure = case_when(lang_dom == "French" ~ fre_unique_words, 
-                                               lang_dom == "English" ~ eng_unique_words),
-         NonDom_unique_words_exposure = case_when(lang_dom == "French" ~ eng_unique_words, 
-                                         lang_dom == "English" ~ fre_unique_words)) %>%
-
+  #defining dominance based on input
+  #mutate(TotalNW_Dom_input = case_when(lang_dom == "English" ~ total_words_eng,
+  #                                        lang_dom == "French" ~ total_words_fre),
+  #       TotalNW_NonDom_input = case_when(lang_dom == "English" ~ total_words_fre,
+  #                                           lang_dom == "French" ~ total_words_eng),
+  #       Dom_unique_words_input = case_when(lang_dom == "French" ~ fre_unique_words, 
+  #                                             lang_dom == "English" ~ eng_unique_words),
+  #       NonDom_unique_words_input = case_when(lang_dom == "French" ~ eng_unique_words, 
+  #                                       lang_dom == "English" ~ fre_unique_words)) %>%
+  #mutate(balance_input = balance/100) %>%
   #defining dominance based on vocabulary size
@@
   mutate(TotalUnique = Dom_unique_words + NonDom_unique_words) %>%
-
   #Defining vocabulary-based balance score
-  #mutate(balance_vocab = TotalNW_NonDom/TotalNW_Dom) %>%
-  mutate(balance_vocab = TotalNW_Dom/(TotalNW_Dom+TotalNW_NonDom)) %>%
-  #Defining proportion of TE (with reference to child's own total vocab size)
-  mutate(propTE = number_of_te*2/word_vocab) %>% 
-  #Defining proportion of TE (with reference to CDI)
-  mutate(propTE_CDI = number_of_te/672)
+  mutate(balance_vocab = TotalNW_NonDom/(TotalNW_Dom+TotalNW_NonDom)) %>%
+  #defining dominance based on input (with reference to the lang_dom defined by vocab size)
+  mutate(LangDom_input = case_when(LangDom_vocab == "English" ~ lang_exp_eng,
+                                    TRUE ~ lang_exp_fre),
+         LangNonDom_input = case_when(LangDom_vocab == "English" ~ lang_exp_fre,
+                                    TRUE ~ lang_exp_eng)) %>%
+  mutate(balance_input = LangNonDom_input/(LangDom_input+LangNonDom_input))
 
@@
 
-# check balancedness cutoffs
-keepers_ws_TE %>%
-  ggplot(aes(x = lang_exp_eng, y = BalanceGroup_byExposure)) + 
-  geom_point()
-
-keepers_ws_TE %>%
-  ggplot(aes(x = lang_exp_fre, y = BalanceGroup_byExposure)) + 
-  geom_point()
-
-```
-
-# Figure 6
+```
+
+# Percentile information from WordbankR
+```{r}
+
+## English Word&Sentence: Vocabulary size at 90% percentile
+english_ws_percentile <-  
+  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
+                      production,
+                      quantile = "standard") %>%
+  # keep only the 90% percentile
+  filter(quantile == 0.90) %>%
+  rename(EngWS_90percentile = production)
+
+## Canadian French Word&Sentence: Vocabulary size at 90% percentile
+french_ws_percentile <-  
+  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
+                      production,
+                      quantile = "standard") %>%
+  # keep only the 90% percentile
+  filter(quantile == 0.90) %>%
+  rename(FrWS_90percentile = production)
+
+## Merge the two percentile dataframe
+ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
+                       by = "age") %>%
+  # compute estimate total vocab by adding the two languages
+  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
+  # keep only necessary columns
+  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
+  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
+  rename(age_months_percentile = age)
+
+write.csv(ws_percentile, "Output/ws_percentile.csv", row.names = F)
+```
+
+# CDI percentile for the hypothetical child Jamie
+```{r}
+english_ws_percentile_18 <-  
+  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
+                      production,
+                      quantile = "standard") %>%
+  filter(age == 18) %>%
+  rename(English = production)
+
+french_ws_percentile_18 <-  
+  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
+                      production,
+                      quantile = "standard") %>%
+  filter(age == 18) %>%
+  rename(French = production)
+
+ws_percentile_18 <- merge(english_ws_percentile_18, french_ws_percentile_18, 
+                       by = c("form", "age","quantile")) %>%
+  # compute estimate total vocab by adding the two languages
+  mutate(Average_90percentile = (English + French)/2) %>%
+  select(-c(language.x, language.y))
+
+```
+
+# Adding percentile info to keepers_ws_TE dataframe & 
+## Creating a new propTE based on the percentile
+```{r}
+# merge the dataframes
+keepers_ws_TE <- keepers_ws_TE %>%
+  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
+  # this means the percentile should be the same for all >30m
+  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
+  # merge with the percentile data frame
+  left_join(ws_percentile, by = "age_months_percentile") 
+```
+
+# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (V = 600)
+```{r}
+## generate the simulation data
+simulation1 <- function (V) {
+  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
+  N_seq <- seq(0, V, by = 10) # sequence of N from 0 to V at an interval of 10
+  D <- rep (D_seq, length(N_seq))
+  N <- rep(N_seq, length(D_seq))
+  data_simulation1 <<- data.frame(V, D, N)
+}
+
+simulation1(600) # setting V = 600
+
+## clean the simulation data
+data_simulation1 <- data_simulation1 %>%
+  arrange(D, N) %>%
+  filter(N <= D) %>%
+  # Get word vocabulary W
+  mutate(W = D + N) %>%
+  # Create other variables
+  mutate(B = N/(D+N), # balance,
+         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 6 groups
+  mutate(bin_B = cut(B, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  arrange(D, N, bin_B)
+
+## visualizing Simulation 1
+simulation1 <- data_simulation1 %>%
+  # convert to long data
+  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("D", "N", "W"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             D = "Panel 1A: Dominant vocabulary (D)", 
+                             N = "Panel 1B: Non-dominant vocabulary (N)",
+                             W = "Panel 1C: Total vocabulary (W)")) %>%
+  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
+  stat_smooth(method = lm, se = F) +
+  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
+  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
+  #scale_color_discrete(direction = -1) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom",
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
+  ylim(0, 200) +
+  labs(x = "Number of words", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)",
+       linetype = "Balance (B)",
+       title = "1) Simulated data") +
+  guides(color = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
+
+ggsave("Output/plot_Simulation1.png", simulation1,
+       width = 10, height = 5)
+```
+
+## Observed data: Testing Simulation 1
 ```{r}
@@
   pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             word_vocab = "Panel A: Total vocabulary", 
-                             TotalNW_Dom = "Panel B: Dominant vocabulary", 
-                             TotalNW_NonDom = "Panel C: Non-dominant vocabulary")) %>%
-  mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
-                           balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
-                           balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
-                           balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
-                           TRUE ~ "0.9")) %>%
-  ggplot(aes(x = number, y = number_of_te, color = bin_B)) +
+                             TotalNW_Dom = "Panel 2A: Dominant vocabulary (D)", 
+                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary (N)",
+                             word_vocab = "Panel 2C: Total vocabulary (W)")) %>%
+  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  #mutate(bin_B = case_when(balance_vocab < 0.2 ~ "0.1",
+  #                         balance_vocab >= 0.2 & balance_vocab < 0.3 ~ "0.2",
+  #                         balance_vocab >= 0.3 & balance_vocab < 0.4 ~ "0.3",
+  #                         balance_vocab >= 0.4 & balance_vocab < 0.5 ~ "0.5",
+  #                         TRUE ~ "0.5")) %>%
+  ggplot(aes(x = number, y = number_of_te, color = bin_B, linetype = bin_B)) +
   #geom_point() +
   stat_smooth(method = lm, se = F) +
+  #scale_color_manual(values=c("#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
+  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
   theme_minimal() + 
   facet_grid(. ~ vocab_type) +
-  ylim(0, 400) +
   theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom") +
-  theme(plot.title = element_text(hjust = 0.5)) +
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        legend.position = "bottom",
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
   labs(x = "Number of words", 
-       y = "Total observed number of translation equivalents", 
-       color = "Balance"
-       #, title = "Number of translation equivalent by balance in relation to different input measures "
-       ) +
-  guides(colour = guide_legend(nrow = 1))
-
-ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)",
+       linetype = "Balance (B)",
+       title = "2) Observed data") +
+  guides(colour = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
+
+plot_Simulation1plusData <- ggarrange(simulation1, 
+                                      Observed_plot_inputs_TE,
+                                      nrow = 2,
+                                      common.legend = TRUE, legend = "bottom")
+
+ggsave("Output/plot_Simulation1plusData.png", plot_Simulation1plusData,
+       width = 10, height = 10)
+
+#ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
+#       width = 10, height = 5)
+
+### one-way ANOVA to compare the 5 balance groups
+aov_TE_binB <- keepers_ws_TE %>% 
+  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  aov(number_of_te ~ bin_B, data = .)
+
+summary(aov_TE_binB)
+
+mean_TE_binB <- keepers_ws_TE %>% 
+  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  group_by(bin_B) %>%
+  summarize(n = n(),
+            mean_TE = mean(number_of_te, na.rm = T),
+            sd_TE = sd(number_of_te, na.rm = T),
+            min_TE = min(number_of_te, na.rm = T),
+            max_TE = max(number_of_te, na.rm = T),)
+
+```
+
+# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (V = 300, 450, or 600)
+```{r}
+## generate the simulation data
+V <- seq(300, 600, by = 150) # set V to 300, 450, and 600
+
+simulation2 <- function (V) {
+  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
+  N_seq <- seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
+  D <- rep(D_seq, length(N_seq))
+  N <- rep(N_seq, length(D_seq))
+  data_simulation2 <<- data.frame(V, D, N)
+}
+
+data_simulation2 <- do.call(rbind, lapply(V, simulation2)) %>%
+  arrange(V, D, N) %>%
+  filter(N <= D) %>%
+  # Create other variables
+  mutate(W = D + N, # word vocabulary W
+         B = N/(D+N), # balance,
+         TE = (D*N)/V, # If D and N are independent, TE = DN/V
+         DI = D-TE, # Unique words in dominant language
+         NI = N-TE) %>% # Unique words in non-dominant language
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 3 groups
+  mutate(bin_B = case_when(B >= 0.36 ~ "Most Balanced",
+                           B <= 0.20 ~ "Least Balanced",
+                           TRUE ~ "Medium Balanced")) %>%
+  mutate(bin_B  = fct_rev(bin_B)) %>%
+  arrange(D, N, bin_B)
+
+## visualizing Simulation 2
+simulation2 <- data_simulation2 %>%
+  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("NI", "DI", "TE"))) %>%
+  mutate(V_label = case_when(V == 600 ~ "V = 600",
+                             V == 450 ~ "V = 450",
+                             V == 300 ~ "V = 300")) %>%
+  group_by(V, V_label, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V_label) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
+  ylim(0, 600) +
+  theme_minimal() +
+  theme(legend.position = "bottom",
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        panel.spacing.x = unit(2, "lines")) + 
+  labs(x = "\nBalance (B)", 
+       y = "Number of concepts (C)", 
+       title = "Panel A: Simulated data")
+
+ggsave("Output/plot_Simulation2.png", simulation2,
         width = 10, height = 5)
-
-```
-
-# Figure 7
+```
+
+## Observed data: Testing Simulation 2
 ```{r}
@@
   # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
-                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
-                              TRUE ~ "28-33 months")) %>%
-  #mutate(V_subset = case_when(age_months_percentile <= 21 ~ "18-21",
-  #                            age_months_percentile > 21 & age_months_percentile <= 25 ~ "22-25",
-  #                            age_months_percentile > 25 & age_months_percentile <= 29 ~ "26-29",
-  #                            TRUE ~ "30-33")) %>%
-  #mutate(bin_B = case_when(balance_vocab < 0.6 ~ "0.5",
-  #                         balance_vocab >= 0.6 & balance_vocab < 0.7 ~ "0.6",
-  #                         balance_vocab >= 0.7 & balance_vocab < 0.8 ~ "0.7",
-  #                         balance_vocab >= 0.8 & balance_vocab < 0.9 ~ "0.8",
-  #                         TRUE ~ "0.9")) %>%
-  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
-                               balance_vocab >= 0.80 ~ "Least Balanced",
+  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
+                              TRUE ~ "28-33 months (V = 620.4 - 638.9)")) %>%
+  mutate(BalSubset = case_when(balance_vocab >= 0.36 ~ "Most Balanced",
+                               balance_vocab <= 0.20 ~ "Least Balanced",
                                TRUE ~ "Medium Balanced")) %>%
-  mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
+  mutate(BalSubset = fct_rev(BalSubset)) %>%
+ # mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
   arrange(desc(number_of_te)) %>%
@@
   mutate(Type = recode(Type, 
-                       NonDom_unique_words = "Unique non-dominant word", 
-                       Dom_unique_words = "Unique dominant word", 
-                       number_of_te = "Translation equivalent")) %>%
-  group_by(V_subset, BalSubset, Type) %>%
+                       NonDom_unique_words = "Initial non-dominant word (NI)", 
+                       Dom_unique_words = "Initial dominant word (DI)", 
+                       number_of_te = "Translation equivalent (TE)")) %>%
+  group_by(age_months_percentile, BalSubset, Type) %>%
   summarise(n_words = mean(NumberOfWords)) %>%
@@
   geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ V_subset) +
+#  facet_grid(. ~ age_months_percentile) +
+  facet_wrap(. ~ age_months_percentile, ncol = 4) +
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  #scale_y_continuous(limits = c(0, 1200)) +
+  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
+  ylim(0, 600) +
   theme_minimal() +
@@
         legend.position = "bottom",
-        plot.title = element_text(hjust = 0.5)) + 
-  labs(x = "Balance", 
-       y = "Observed number of concepts", 
-       title = "Vocabulary composition across different subset of developmental timepoints")
-
-ggsave("Output/Observed_plot_stacked_byV.png", plot_stacked_byV,
-       width = 10, height = 5)
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        panel.spacing.x = unit(2, "lines")) + 
+  labs(x = "\nBalance (B)", 
+       y = "Number of concepts (C)",
+       title = "Panel B: Observed data")
+
+plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
+                                      plot_stacked_byV,
+                                      nrow = 2,
+                                      heights=c(0.9,1))
+
+ggsave("Output/plot_Simulation2plusData.png", plot_Simulation2plusData,
+       width = 10, height = 10)
+
+# One-way Anova to compare the 3 V groups
+keepers_ws_TE_3Vsubset <- keepers_ws_TE %>%
+  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
+  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
+                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))
+
+summary(aov(concept_vocab ~ V_subset, data = keepers_ws_TE_3Vsubset))
+
+mean_conceptV_Vsubset <- keepers_ws_TE_3Vsubset %>%
+  group_by(V_subset) %>%
+  summarise(mean_conceptV = mean(concept_vocab, na.rm = T),
+            sd_conceptV = sd(concept_vocab, na.rm = T),
+            min_conceptV = min(concept_vocab, na.rm = T),
+            max_conceptV = max(concept_vocab, na.rm = T))
+
+pairwise.t.test(keepers_ws_TE_3Vsubset$concept_vocab, keepers_ws_TE_3Vsubset$V_subset,
+                 p.adjust.method = "bonferroni")
 
@@
   group_by(BalSubset) %>%
-  summarise(mean_Du = mean(Dom_unique_words, na.rm=T), 
-            sd_Du = sd(Dom_unique_words, na.rm=T),
-            min_Du = min(Dom_unique_words, na.rm=T),
-            max_Du = max(Dom_unique_words, na.rm=T))
-
-Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
-summary(Du_byB_anova)
+  summarise(mean_DI = mean(Dom_unique_words, na.rm=T), 
+            sd_DI = sd(Dom_unique_words, na.rm=T),
+            min_DI = min(Dom_unique_words, na.rm=T),
+            max_DI = max(Dom_unique_words, na.rm=T))
+
+DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
+summary(DI_byB_anova)
 
@@
   group_by(BalSubset) %>%
-  summarise(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
-            sd_Nu = sd(NonDom_unique_words, na.rm=T),
-            min_Nu = min(NonDom_unique_words, na.rm=T),
-            max_Nu = max(NonDom_unique_words, na.rm=T))
-
-Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
-summary(Nu_byB_anova)
+  summarise(mean_NI = mean(NonDom_unique_words, na.rm=T), 
+            sd_NI = sd(NonDom_unique_words, na.rm=T),
+            min_NI = min(NonDom_unique_words, na.rm=T),
+            max_NI = max(NonDom_unique_words, na.rm=T))
+
+NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
+summary(NI_byB_anova)
 
@@
 
-# Descriptives: Vocabulary measures
-```{r}
-## Mean word vocab size
+# Simulation 3: Non-independence of dominant-language (D) and non-dominant language words (N)
+```{r}
+## generate the simulation data
+V <- seq(150, 600, by = 150) # set V to 300, 450, and 600
+
+simulation3 <- function (V) {
+  D_seq = seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 50
+  N_seq = seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
+  D <- rep(D_seq, length(N_seq))
+  N <- rep(N_seq, length(D_seq))
+  data_simulation3 <<- data.frame(V, D, N)
+}
+
+data_simulation3 <- do.call(rbind, lapply(V, simulation3)) %>%
+  arrange(V, D, N) %>%
+  filter(N <= D) %>% # eliminate cases when N > D
+  mutate(W = D+N, # word vocabulary W
+         B = N/(D+N), # balance B
+         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
+  mutate_at(vars(-B), round, 0) %>%
+  mutate(bin_B = cut(B, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  arrange(bin_B)
+
+## Simulation3: TE = IND * (D*N/V)
+### defining the independence parameter I
+IND1 = 1.5 # IND > 1, TEs are MORE easily learned than unique words
+IND2 = 0.5 # IND < 1, TEs are LESS easily learned than unique words
+
+### visualizing the simulation
+simulation3 <- data_simulation3 %>%
+  ## calculate where IND > 1
+  mutate(TE_easier = IND1*(D*N/V)) %>%
+  ## calculate where IND < 1
+  mutate(TE_harder = IND2*(D*N/V)) %>%
+  # round to keep whole numbers only
+  mutate_at(vars(-c(B, bin_B)), round, 0) %>%
+  # prepare dataset to plot
+  pivot_longer(c(TE, TE_harder, TE_easier), 
+               names_to = "Model", values_to = "number_TE") %>%
+  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
+  mutate(Model = recode(Model, 
+                        TE = "IND = 1 (Independence Account)",
+                        TE_easier = "IND = 1.5 (Enhancement Account)",
+                        TE_harder = "IND = 0.5 (Avoidance Account)")) %>%
+  # ggplot
+  ggplot(aes(x = W, y = number_TE, color = bin_B, linetype = bin_B)) +
+  stat_smooth(method = lm, se = F) +
+  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
+  # scale_color_discrete(direction = -1) +
+  theme_minimal() + 
+  facet_grid(. ~ Model) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5),
+        text = element_text(size=12),
+        panel.spacing.x = unit(2.5, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
+  labs(x = "Total word vocabulary (W)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)",
+       linetype = "Balance (B)") +
+  guides(colour = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
+
+ggsave("Output/plot_Simulation3.png", simulation3, 
+       width = 10, height = 5)
+```
+
+# Testing Simulation 3: Figures 7 & 8
+## Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) 
+```{r}
+## Create predicted TEs for simulation
+keepers_ws_TE_simulation <- keepers_ws_TE %>%
+  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile) %>%
+  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, TotalNW_Dom, TotalNW_NonDom, word_vocab, age_months_binned)
+
+## LM model
+Percentile_model <- keepers_ws_TE_simulation %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model)
+
+## Ploting Y = observed rawTE  & X = Predicted rawTE 
+TEprediction_percentile <- keepers_ws_TE_simulation %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+#  facet_grid(. ~TEgroup) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+#  geom_text(aes(label=ID_testdate)) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  xlim(-10,450) +
+  ylim(-10,450) +
+  theme_minimal()
+
+ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
+       height=5, width=5)
+
+## Identifying outliers
+CheckOutliers_model <- keepers_ws_TE_simulation %>%
+  lm(number_of_te ~ 0 + Predicted_TE_percentile, 
+#  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom,
+     data = .)
+summary(CheckOutliers_model)
+
+keepers_ws_TE_simulation$CD <- cooks.distance(CheckOutliers_model)
+
+TEprediction_percentile_CD <- keepers_ws_TE_simulation %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_smooth(method = "lm", se = T) +
+  geom_point(size=3, aes(color=CD)) +
+  geom_text(aes(label=ID_testdate)) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal()
+
+qplot_CD <- qplot(keepers_ws_TE_simulation$CD)
+
+grid.arrange(TEprediction_percentile_CD, qplot_CD, ncol = 2)
+
+keepers_ws_TE_simulation_removeCD <- filter(keepers_ws_TE_simulation, CD < 0.4)
+
+TEprediction_percentile_removeCD <- keepers_ws_TE_simulation_removeCD %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_smooth(method = "lm", se = T) +
+  geom_point(size=3) +
+  geom_text(aes(label=ID_testdate)) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal()
+
+grid.arrange(TEprediction_percentile_removeCD, TEprediction_percentile_CD, ncol = 2)
+
+Percentile_model_removeCD <- keepers_ws_TE_simulation_removeCD %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_removeCD)
+
+## Ploting Y = observed rawTE  & X = Predicted rawTE 
+TEprediction_percentile_NoOutliers <- keepers_ws_TE_simulation_removeCD %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+#  facet_grid(. ~TEgroup) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  xlim(-10,430) +
+  ylim(-10,430) +
+  theme_minimal()
+
+Percentile_model_NoOutliers_LessThan300W <- keepers_ws_TE_simulation_removeCD %>%
+  filter(word_vocab < 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_NoOutliers_LessThan300W)
+
+Percentile_model_NoOutliers_MoreThan300W <- keepers_ws_TE_simulation_removeCD %>%
+  filter(word_vocab > 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_NoOutliers_MoreThan300W)
+
+## Looking closer at the infants with total vocabulary < 300
+
+keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300)
+
+stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
+               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)
+
+keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300)
+
+stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
+               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)
+
+# Plot
+TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
+  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
+  filter(W_group == "Less than 300 total vocabulary") %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
+       title = "Less than 300 total vocabulary") +
+  xlim(0,80) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+TEprediction_percentile_facet_moreW300 <- keepers_ws_TE_simulation %>%
+  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
+  filter(W_group == "More than 300 total vocabulary") %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
+       title = "More than 300 total vocabulary") +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal() +
+  theme(plot.title = element_text(hjust = 0.5))
+
+TEprediction_percentile_facetByW300 <- 
+  ggarrange(TEprediction_percentile_facet_lessW300 + rremove("x.title"), 
+            TEprediction_percentile_facet_moreW300 + rremove("xy.title"), 
+            ncol = 2, nrow = 1) %>%
+  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model"))
+
+ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetByW300,
+       height=5,width=10)
+
+Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_LessThan300W)
+
+Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_MoreThan300W)
+
+age_info <- keepers_ws_TE %>%
+  mutate(group = if_else(word_vocab < 300, "LessThan300W", "MoreThan300W")) %>%
+  group_by(group) %>%
+  summarize(n = n(),
+             mean_age = mean(age_continuous, na.rm = T),
+             sd_age = sd(age_continuous, na.rm = T),
+             min_age = min(age_continuous, na.rm = T),
+             max_age = max(age_continuous, na.rm = T))
+
+Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
+  filter(number_of_te < 130) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_LessThan300W)
+
+Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
+  filter(number_of_te > 130) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+
+summary(Percentile_model_MoreThan300W)
+
+```
+
+# Descriptives
+## Gender 
+```{r}
 keepers_ws_TE %>%
-  summarize(mean_W = mean(word_vocab, na.rm=T), 
-            sd_W = sd(word_vocab, na.rm=T),
-            min_W = min(word_vocab, na.rm=T),
-            max_W = max(word_vocab, na.rm=T))
-
-## Mean TE size
+  group_by(gender) %>% 
+  distinct(baby_id, .keep.all=T) %>%
+  count()
+```
+
+## English dominant
+```{r}
+# dominance defined by language exposure
+keepers_ws_TE %>% group_by(lang_dom) %>% distinct(baby_id) %>%
+  count()
+
+# dominance defined by vocab size
 keepers_ws_TE %>% 
-  summarize(mean_TE = mean(number_of_te, na.rm=T), 
-            sd_TE = sd(number_of_te, na.rm=T),
-            min_TE = min(number_of_te, na.rm=T),
-            max_TE = max(number_of_te, na.rm=T))
-
-## Mean unique word size
+  distinct(ID_testdate, .keep_all = TRUE) %>%
+  mutate(Eng_Dom = ifelse(LangDom_vocab == "English", 1, 0)) %>%
+  summarise(n = n(),
+            n_EngDom = sum(Eng_Dom),
+            percent_EngDom = n_EngDom/n * 100,
+            percent_FrDom = 100 - percent_EngDom)
+
+```
+
+## demographics
+```{r}
+# How many data points including some babies who made multiple visits to the lab (N)
 keepers_ws_TE %>% 
-  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
-            sd_Du = sd(Dom_unique_words, na.rm=T),
-            min_Du = min(Dom_unique_words, na.rm=T),
-            max_Du = max(Dom_unique_words, na.rm=T))
-
+  distinct(ID_testdate) %>% 
+  count()
+
+N_duplicatedBabyID <- keepers_ws_TE %>%
+  filter(duplicated(baby_id))
+
+# how many unique babies
 keepers_ws_TE %>%
-  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
-            sd_Nu = sd(NonDom_unique_words, na.rm=T),
-            min_Nu = min(NonDom_unique_words, na.rm=T),
-            max_Nu = max(NonDom_unique_words, na.rm=T))
-
+  distinct(baby_id) %>%
+  count()
+
+keepers_ws_TE %>%
+  distinct(participant_id) %>%
+  count()
+
+# Create dataframe of demographics
+d_participants <- keepers_ws_TE %>%
+  select(ID_testdate, gender,age_continuous, years_education, lang_exp_eng, lang_exp_fre, lang_exp_other) %>%
+  mutate(female = ifelse(gender == "F", 1, 0)) %>%
+  pivot_longer(-c(ID_testdate), names_to = "info", values_to = "value") %>%
+  group_by(info) %>%
+  summarise(n_subject = n_distinct(ID_testdate),
+            mean = mean(value, na.rm=T),
+            sd = sd(value, na.rm=T),
+            min = min(value, na.rm=T),
+            max = max(value, na.rm=T))
+```
+
+## average # of vocabulary
+```{r}
+vocabulary_average <- keepers_ws_TE %>%
+  select(ID_testdate, word_vocab, TotalNW_Dom, TotalNW_NonDom, concept_vocab,
+         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  pivot_longer(-ID_testdate, names_to = "vocabulary", values_to = "number") %>%
+  group_by(vocabulary) %>%
+  summarise(mean = mean(number, na.rm = T), 
+            sd = sd(number, na.rm = T),
+            min = min(number, na.rm = T),
+            max = max(number, na.rm = T)) %>%
+  mutate_if(is.numeric, round, digits = 2)
+
+## paired-sample t-test to compare D and N
+t.test(keepers_ws_TE$TotalNW_Dom, keepers_ws_TE$TotalNW_NonDom, 
+       paired = TRUE)
+
+## paired-sample t-test to compare DI and NI
 t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
-       paired = TRUE) # paired-samples t-test to compare Du and Nu
-
-## Stacked barchart to plot TE, Du, and Nu of individual infants (facet by balance)
+       paired = TRUE) 
+```
+
+## average balance and comparison with balance by input
+```{r}
+### Average balance by vocab
+keepers_ws_TE %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+### grouped by language
+keepers_ws_TE %>%
+  group_by(LangDom_vocab) %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+
+t.test(balance_vocab ~ LangDom_vocab, keepers_ws_TE) # t-test to compare the two language groups
+
+### Comparsion with balance by input (which is defined with reference to the lang_dom defined by vocab size)
+keepers_ws_TE %>%
+  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
+  summarise(n = n(),
+            n_consistent = sum(consistent_LangDom),
+            percentage_consistent = n_consistent/n*100,
+            n_inconsistent = n - n_consistent,
+            percentage_inconsistent = n_inconsistent/n*100)
+
+keepers_ws_TE_withoutInconsistent <- keepers_ws_TE %>%
+  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
+  filter(consistent_LangDom == 1)
+
+### Correlation between balance by vocabulary and balance by input
+cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$balance_input, method = "pearson")
+
+### Correlation between balance by vocabulary and raw language exposure to the non-dominant language
+cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input, method = "pearson")
+
+plot(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input,
+     xlab = "Vocabulary balance (= N/W)",
+     ylab = "raw exposure to the non-dominant language") +
+  lines(lowess(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input), col="blue") # lowess line (x,y)
+
+  abline(lm(keepers_ws_TE$LangNonDom_input~keepers_ws_TE$balance_vocab), col="red") # regression line (y~x)
+lines(lowess(wt,mpg), col="blue") # lowess line (x,y)
+
+```
+
+# Correlation
+```{r}
+Correlation <- keepers_ws_TE %>%
+  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
+         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
+  # rename variables
+  rename(Age_month = age_continuous,
+         V = Average_90percentile,
+         B = balance_vocab, 
+         W = word_vocab,
+         D = TotalNW_Dom, 
+         N = TotalNW_NonDom, 
+         TE = number_of_te, 
+         DI = Dom_unique_words, 
+         NI = NonDom_unique_words, 
+         C = concept_vocab)
+
+# Correlation with corrections for multiple comparisons
+library(psych)
+TE_corr <- corr.test(Correlation, use = "pairwise", method = "pearson", adjust = "BY")
+
+write.csv(round(TE_corr$r, 3), file="Output/ObservedTE_correlations_r.csv")  ## coefficients
+write.csv(round(TE_corr$p, 3), file="Output/ObservedTE_correlations_p.csv")  ## p-values
+
+library(Hmisc)
+mycor <- rcorr(as.matrix(Correlation), type="pearson")
+mycor$r
+round(mycor$P, 3)
+round(p.adjust(mycor$P, "BY"))
+
+library(corrplot)
+corrplot(mycor)
+write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
+
+cor.test(Correlation$TE, Correlation$B, method = "pearson")
+cor.test(Correlation$TE, Correlation$W, method = "pearson")
+cor.test(Correlation$TE, Correlation$D, method = "pearson")
+cor.test(Correlation$TE, Correlation$N, method = "pearson")
+cor.test(Correlation$TE, Correlation$V, method = "pearson")
+cor.test(Correlation$DI, Correlation$B, method = "pearson")
+cor.test(Correlation$NI, Correlation$B, method = "pearson")
+
+```
+
+# Figure 5: Stacked barchart to plot TE, DI, and NI of individual infants (facet by balance)
+```{r}
 stacked <- keepers_ws_TE %>%
@@
   #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
-  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
-                               balance_vocab >= 0.80 ~ "LeastBalanced",
+  mutate(BalSubset = case_when(balance_vocab >= 0.35 ~ "MostBalanced",
+                               balance_vocab <= 0.20 ~ "LeastBalanced",
                                TRUE ~ "MediumBalanced")) %>%
@@
   mutate(Type = recode(Type, 
-                       NonDom_unique_words = "Unique non-dominant word", 
-                       Dom_unique_words = "Unique dominant word", 
+                       NonDom_unique_words = "Initial non-dominant word", 
+                       Dom_unique_words = "Initial dominant word", 
                        number_of_te = "Translation equivalent"))
@@
   # Subset balance_vocab by 3 levels 
-  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "MostBalanced",
-                               balance_vocab >= 0.80 ~ "LeastBalanced",
+  mutate(BalSubset = case_when(balance_vocab >= 0.35 ~ "MostBalanced",
+                               balance_vocab <= 0.20 ~ "LeastBalanced",
                                TRUE ~ "MediumBalanced"))
@@
   group_by(BalSubset) %>%
-  summarize(mean_propTE = mean(propTE, na.rm=T), 
+  summarise(mean_propTE = mean(propTE, na.rm=T), 
             sd_propTE = sd(propTE, na.rm=T),
@@
   group_by(BalSubset) %>%
-  summarize(mean_Du = mean(Dom_unique_words, na.rm=T), 
-            sd_Du = sd(Dom_unique_words, na.rm=T),
-            min_Du = min(Dom_unique_words, na.rm=T),
-            max_Du = max(Dom_unique_words, na.rm=T))
-
-Du_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
-summary(Du_byB_anova)
+  summarize(mean_DI = mean(Dom_unique_words, na.rm=T), 
+            sd_DI = sd(Dom_unique_words, na.rm=T),
+            min_DI = min(Dom_unique_words, na.rm=T),
+            max_DI = max(Dom_unique_words, na.rm=T))
+
+DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
+summary(DI_byB_anova)
 
@@
   group_by(BalSubset) %>%
-  summarize(mean_Nu = mean(NonDom_unique_words, na.rm=T), 
-            sd_Nu = sd(NonDom_unique_words, na.rm=T),
-            min_Nu = min(NonDom_unique_words, na.rm=T),
-            max_Nu = max(NonDom_unique_words, na.rm=T))
-
-Nu_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
-summary(Nu_byB_anova)
+  summarize(mean_NI = mean(NonDom_unique_words, na.rm=T), 
+            sd_NI = sd(NonDom_unique_words, na.rm=T),
+            min_NI = min(NonDom_unique_words, na.rm=T),
+            max_NI = max(NonDom_unique_words, na.rm=T))
+
+NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
+summary(NI_byB_anova)
 
@@
                  p.adjust.method = "BH")
-```
-
-# Descriptives
-## number of CDIs & Babys
-```{r}
-# how many unique babies
-keepers_ws_TE %>%
-  distinct(baby_id) %>%
-  count()
-
-keepers_ws_TE %>%
-  distinct(participant_id) %>%
-  count()
-
-# some babies made multiple visits to the lab
-keepers_ws_TE %>% 
-  distinct(ID_testdate) %>% 
-  count()
-
-N_duplicatedBabyID <- keepers_ws_TE %>%
-  filter(duplicated(baby_id))
-
-# Create dataframe of demographics
-d_participants <- keepers_ws_TE %>%
-  select(baby_id, age_continuous, gender, 
-         lang_exp_eng, lang_exp_fre, lang_exp_other, years_education) %>%
-  mutate(Female = ifelse(gender == "F", 1, 0)) %>%
-  summarise(n = n(),
-            age_Month = mean(age_continuous, na.rm = TRUE),
-            lang_exp_eng = mean(lang_exp_eng, na.rm = TRUE),
-            lang_exp_fre = mean(lang_exp_fre, na.rm = TRUE), 
-            lang_exp_other = mean(lang_exp_other, na.rm = TRUE),
-            years_education = mean(years_education, na.rm = TRUE),
-            number_female = sum(Female)) %>%
-  mutate(PercentageFemale = number_female/n)
-
-```
-
-## Gender 
-```{r}
-keepers_ws_TE %>%
-  group_by(gender) %>% 
-  distinct(baby_id, .keep.all=T) %>%
-  count()
-```
-
-## English dominant
-```{r}
-# dominance defined by language exposure
-keepers_ws_TE %>% group_by(lang_dom) %>% distinct(baby_id) %>%
-  count()
-
-# dominance defined by vocab size
-keepers_ws_TE %>% group_by(LangDom_vocab) %>% distinct(baby_id) %>%
-  count()
-```
-
-## years of maternal education
-```{r}
-keepers_ws_TE %>% distinct(baby_id, .keep_all = T) %>%
-  summarize(mean = mean(years_education, na.rm=T), 
-            sd = sd(years_education, na.rm=T))
-```
-
-## age range, mean
-```{r}
-keepers_ws_TE %>%
-  distinct(baby_id, .keep_all = T) %>% 
-  summarize(sd = sd(age_days, na.rm=T))
-
-do.call(cbind, lapply(keepers_ws_TE %>%
-                        distinct(baby_id, .keep_all = T) %>% 
-                        select(age_days,age_continuous),
-                      summary))
-
-```
-
-## average # of TEs
-```{r}
-keepers_ws_TE %>%
-  filter(both_cdi_filled=="Y") %>%
-  summarise(mean = mean(propTE, na.rm=T), 
-            min= min(propTE, na.rm=T),
-            max= max(propTE, na.rm=T))
-
-```
-
-## Percentile information from WordbankR
-```{r}
-
-## English Word&Sentence: Vocabulary size at 90% percentile
-english_ws_percentile <-  
-  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
-                      production,
-                      quantile = "standard") %>%
-  # keep only the 90% percentile
-  filter(quantile == 0.90) %>%
-  rename(EngWS_90percentile = production)
-
-## Canadian French Word&Sentence: Vocabulary size at 90% percentile
-french_ws_percentile <-  
-  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
-                      production,
-                      quantile = "standard") %>%
-  # keep only the 90% percentile
-  filter(quantile == 0.90) %>%
-  rename(FrWS_90percentile = production)
-
-## Merge the two percentile dataframe
-ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
-                       by = "age") %>%
-  # compute estimate total vocab by adding the two languages
-  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
-  # keep only necessary columns
-  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
-  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
-  rename(age_months_percentile = age)
-
-write.csv(ws_percentile, "Output/ws_percentile.csv", row.names = F)
-```
-
-# Adding percentile info to keepers_ws_TE dataframe & 
-## Creating a new propTE based on the percentile
-```{r}
-# merge the dataframes
-keepers_ws_TE <- keepers_ws_TE %>%
-  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe
-  # this means the percentile should be the same for all >30m
-  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
-  # merge with the percentile data frame
-  left_join(ws_percentile, by = "age_months_percentile")
-
-%>%
-  # calculate TE in relation to CDI percentile data
-  mutate(propTE_percentile = number_of_te*2/Total_90percentile)
-  # %>% select(baby_id, age_months_binned, age_months_percentile, word_vocab, Total_90percentile,
-  #       propTE, propTE_CDI, propTE_percentile)
-
-# check if any baby learn more words than the percentile
-# (there are 2 babies know more than the percentile in their dominant language)
-MoreThanPercentile <- keepers_ws_TE %>%
-  select(ID_testdate, age_months_binned, lang_dom, LangDom_vocab, 
-         word_vocab, total_words_eng, EngWS_90percentile, 
-         total_words_fre, FrWS_90percentile, word_vocab, Total_90percentile) %>%
-  filter(total_words_eng > EngWS_90percentile | 
-           total_words_fre > FrWS_90percentile |
-           word_vocab > Total_90percentile)
-```
-
-# Correlation
-```{r}
-Correlation <- keepers_ws_TE %>%
-  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
-         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
-  # recode the balance_vocab so that it would be in descending order, so the value of the most balanced would be the greatest
-  mutate(balance_vocab = balance_vocab*-1) %>%
-  # rename variables
-  rename(Age_month = age_continuous,
-         V = Average_90percentile,
-         B = balance_vocab, 
-         W = word_vocab,
-         D = TotalNW_Dom, 
-         N = TotalNW_NonDom, 
-         TE = number_of_te, 
-         Du = Dom_unique_words, 
-         Nu = NonDom_unique_words, 
-         C = concept_vocab)
-
-library(Hmisc)
-mycor <- rcorr(as.matrix(Correlation), type="pearson")
-mycor$r
-round(mycor$P, 2)
-
-library(corrplot)
-mycor <- round(cor(Correlation), 2)
-corrplot(mycor)
-write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
-
-cor.test(Correlation$TE, Correlation$B, method = "pearson")
-cor.test(Correlation$TE, Correlation$W, method = "pearson")
-cor.test(Correlation$TE, Correlation$D, method = "pearson")
-cor.test(Correlation$TE, Correlation$N, method = "pearson")
-cor.test(Correlation$TE, Correlation$V, method = "pearson")
-cor.test(Correlation$Du, Correlation$B, method = "pearson")
-cor.test(Correlation$Nu, Correlation$B, method = "pearson")
-
-```
-
-###### Theoretical distribution of TEs (Simulation): D * N / V (TotalNW_Dom * TotalNW_NonDom / Average_90percentile) ###### 
-```{r}
-## Create predicted TEs for simulation
-keepers_ws_TE_simulation <- keepers_ws_TE %>%
-  mutate(Predicted_TE_percentile = TotalNW_Dom * TotalNW_NonDom/Average_90percentile)
-
-## Linear regressions of predicted TE models
-Percentile_model <- keepers_ws_TE_simulation %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model)
-
-## Ploting Y = observed rawTE  & X = Predicted rawTE 
-TEprediction_percentile <- keepers_ws_TE_simulation %>%
-  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
-#  facet_grid(. ~TEgroup) +
-  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,430) +
-  ylim(-10,430) +
-  theme_minimal()
-
-ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
-       height=5, width=5)
-
-# Kolmogorov-Smirnov test to compare the observed and predicted TE
-#stats::ks.test(keepers_ws_TE_simulation$number_of_te,
-#               keepers_ws_TE_simulation$Predicted_TE_percentile)
-
-## Looking closer at the infants with total vocabulary < 150
-
-keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
-  filter(word_vocab < 300)
-
-stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
-               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)
-
-keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
-  filter(word_vocab > 300)
-
-stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
-               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)
-
-# Plot
-TEprediction_percentile_facetW <- keepers_ws_TE_simulation %>%
-  mutate(W_group = if_else(word_vocab < 300, "Less than 300 total vocabulary", "More than 300 total vocabulary")) %>%
-  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
-  facet_wrap(~ W_group, scales = "free") +
-  #facet_grid(. ~ W_group, scales = "free") +
-  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  theme_minimal()
-
-ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetW,
-       height=5,width=10)
-
-Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
-  filter(word_vocab < 300) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_LessThan300W)
-
-Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
-  filter(word_vocab > 300) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_MoreThan300W)
-
 ```

### 2021-05-17 Analyze_TEs.Rmd
fixed simulation2 figure

@@
   mutate(BalSubset = fct_rev(BalSubset)) %>%
- # mutate(BalSubset = factor(BalSubset, levels=c("Most Balanced", "Medium Balanced", "Least Balanced"))) %>%
   arrange(desc(number_of_te)) %>%
@@
                        number_of_te = "Translation equivalent (TE)")) %>%
-  group_by(age_months_percentile, BalSubset, Type) %>%
+  group_by(V_subset, BalSubset, Type) %>%
   summarise(n_words = mean(NumberOfWords)) %>%
@@
   geom_bar(position="stack", stat="identity") +
-#  facet_grid(. ~ age_months_percentile) +
-  facet_wrap(. ~ age_months_percentile, ncol = 4) +
+  facet_grid(. ~ V_subset) +
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 

### 2021-05-18 Analyze_TEs.Rmd
fixed typos

@@
 simulation3 <- function (V) {
-  D_seq = seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 50
+  D_seq = seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
   N_seq = seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
@@
 
-plot(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input,
      xlab = "Vocabulary balance (= N/W)",
-     ylab = "raw exposure to the non-dominant language") +
-  lines(lowess(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input), col="blue") # lowess line (x,y)
-
+balance_correlation <- plot(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input,
+                            xlab = "Vocabulary balance (= N/W)",
+                            ylab = "Raw exposure to the non-dominant language",
+                            cex.lab=1.0) +
+  #lines(lowess(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input), col="blue") # lowess line (x,y)
   abline(lm(keepers_ws_TE$LangNonDom_input~keepers_ws_TE$balance_vocab), col="red") # regression line (y~x)
-lines(lowess(wt,mpg), col="blue") # lowess line (x,y)
 

### 2021-06-03 Analyze_TEs.Rmd
updated analysis

--- 2021-05-17 Analyze_TEs.Rmd
@@
 # exclude kids that are bilingual at one visit, monolingual at another (regardless of direction)
-# switch from other to bilingual, or vice versa, or from other to monolingual and vice versa are ok
-exclude.lang_group_change  = cdi_ws %>%
+exclude.lang_group_change <- cdi_ws %>%
   filter(lang_group !="other") %>%
@@
   filter(changed_language_group_status == TRUE) %>% 
-  pull(ID_testdate)
+  pull(baby_id)
 
@@
       # exclude language group status changed from bilingual to mono or vice versa
-      ID_testdate %in% exclude.lang_group_change ~ 1, 
+      baby_id %in% exclude.lang_group_change ~ 1, 
       lang_group == "bilingual" ~ 0, #bilingual
@@
     both_cdi_filled == "Y",
-    # keep only babies with at least 15% of E and 15% of F (i.e., exclude monolinguals)
-    lang_exp_eng >= 15 & lang_exp_eng <= 85,
-    lang_exp_fre >= 15 & lang_exp_fre <= 85)
+    # keep only babies with at least 25% of E and 25% of F (i.e., exclude monolinguals)
+    lang_exp_eng >= 25 & lang_exp_eng <= 75,
+    lang_exp_fre >= 25 & lang_exp_fre <= 75)
 
@@
 
+# descriptives
+## how many distinct participants and data point we have
+```{r}
+keepers_ws_TE %>% summarize(N_babies = n_distinct(baby_id),
+                            N_admin = n_distinct(ID_testdate))
+
+# number of participants who contributed multiple data points
+keepers_ws_TE %>% 
+  count(baby_id) %>% 
+  filter(n>1) %>%
+  nrow()
+
+N_duplicatedBabyID <- keepers_ws_TE %>%
+  filter(duplicated(baby_id)) # checking which baby_id has multiple data
+```
+
+## percentage of english-dominant & french-dominant children
+```{r}
+# dominance defined by language exposure
+keepers_ws_TE %>% group_by(lang_dom) %>% distinct(baby_id) %>%
+  count()
+
+# dominance defined by vocab size
+keepers_ws_TE %>% 
+  distinct(ID_testdate, .keep_all = TRUE) %>%
+  mutate(Eng_Dom = ifelse(LangDom_vocab == "English", 1, 0)) %>%
+  summarise(n = n(),
+            n_EngDom = sum(Eng_Dom),
+            percent_EngDom = n_EngDom/n * 100,
+            percent_FrDom = 100 - percent_EngDom)
+```
+
+## gender: percentage of girls
+```{r}
+keepers_ws_TE %>%
+  group_by(gender) %>% 
+  distinct(baby_id, .keep.all=T) %>%
+  count() 
+```
+
+## demographics table: age (in months), maternal edu in years, lang_exp_eng, lang_exp_fre, lang_exp_other
+```{r}
+# Create dataframe of demographics
+d_participants <- keepers_ws_TE %>%
+  select(ID_testdate, age_continuous, years_education, lang_exp_eng, lang_exp_fre, lang_exp_other) %>%
+  pivot_longer(-c(ID_testdate), names_to = "info", values_to = "value") %>%
+  group_by(info) %>%
+  summarise(n_subject = n_distinct(ID_testdate),
+            mean = mean(value, na.rm=T),
+            sd = sd(value, na.rm=T),
+            min = min(value, na.rm=T),
+            max = max(value, na.rm=T))
+```
+
+## who filled out the CDIs
+```{r}
+keepers_ws_TE %>% 
+  # combine grandmother & other family member
+  mutate(cdi_filled_by = replace(cdi_filled_by, cdi_filled_by == "Grandmother", "Other family member")) %>% 
+  # count number of respondents
+  count(cdi_filled_by) %>%
+  mutate(percentage = n/sum(n)*100) # calculate percentages
+```
+
+## average # of vocabulary
+```{r}
+vocabulary_average <- keepers_ws_TE %>%
+  select(ID_testdate, word_vocab, TotalNW_Dom, TotalNW_NonDom, concept_vocab,
+         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+  pivot_longer(-ID_testdate, names_to = "vocabulary", values_to = "number") %>%
+  group_by(vocabulary) %>%
+  summarise(mean = mean(number, na.rm = T), 
+            sd = sd(number, na.rm = T),
+            min = min(number, na.rm = T),
+            max = max(number, na.rm = T)) %>%
+  mutate_if(is.numeric, round, digits = 2)
+
+## paired-sample t-test to compare DOM and NONDOM
+t.test(keepers_ws_TE$TotalNW_Dom, keepers_ws_TE$TotalNW_NonDom, 
+       paired = TRUE)
+
+## paired-sample t-test to compare DI and NI
+t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
+       paired = TRUE) 
+```
+
+## average balance and comparison with balance by input
+```{r}
+# Average BALANCE score (defined by vocabulary balance)
+keepers_ws_TE %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+
+# grouped by language-dominance
+keepers_ws_TE %>%
+  group_by(LangDom_vocab) %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+
+t.test(balance_vocab ~ LangDom_vocab, keepers_ws_TE) # t-test to compare the two language groups
+
+# Comparsion with balance by input (which is defined with reference to the lang_dom defined by vocab size)
+keepers_ws_TE %>%
+  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
+  summarise(n = n(),
+            n_consistent = sum(consistent_LangDom),
+            percentage_consistent = n_consistent/n*100,
+            n_inconsistent = n - n_consistent,
+            percentage_inconsistent = n_inconsistent/n*100)
+
+keepers_ws_TE_withoutInconsistent <- keepers_ws_TE %>%
+  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
+  filter(consistent_LangDom == 1)
+
+## Correlation between balance by vocabulary and balance by input
+#cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$balance_input, method = "pearson")
+
+### Correlation between balance by vocabulary and raw language exposure to the non-dominant language
+cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input, method = "pearson")
+
+balance_correlation <- plot(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input,
+                            xlab = "Vocabulary balance (= N/W)",
+                            ylab = "Raw exposure to the non-dominant language",
+                            cex.lab=1.0) +
+  #lines(lowess(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input), col="blue") # lowess line (x,y)
+  abline(lm(keepers_ws_TE$LangNonDom_input~keepers_ws_TE$balance_vocab), col="red") # regression line (y~x)
+
+balance_correlation <- keepers_ws_TE %>%
+  ggplot(aes(x = balance_vocab, y = LangNonDom_input)) +
+  stat_smooth(method = lm, se = F, color = "black") +
+  geom_point(shape = 1) + 
+  theme_light() + 
+  labs(x = "Vocabulary BALANCE (= NONDOM/WORD)", 
+       y = "Raw exposure to the non-dominant language (%)") +
+  theme(text = element_text(size=12))
+
+ggsave("Output/plot_BALANCE_correlation.png", balance_correlation,
+       width = 6, height = 5)
+
+```
+
 # Percentile information from WordbankR
@@
 
-# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (V = 600)
+# Correlation among variables in the observed data
+```{r}
+Correlation <- keepers_ws_TE %>%
+  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
+         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
+  # rename variables
+  rename(Age_month = age_continuous,
+         V = Average_90percentile,
+         B = balance_vocab, 
+         W = word_vocab,
+         D = TotalNW_Dom, 
+         N = TotalNW_NonDom, 
+         TE = number_of_te, 
+         DI = Dom_unique_words, 
+         NI = NonDom_unique_words, 
+         C = concept_vocab)
+
+# Correlation with corrections for multiple comparisons
+library(psych)
+TE_corr <- corr.test(Correlation, use = "pairwise", method = "pearson", adjust = "BY")
+
+write.csv(round(TE_corr$r, 3), file="Output/ObservedTE_correlations_r.csv")  ## coefficients
+write.csv(round(TE_corr$p, 3), file="Output/ObservedTE_correlations_p.csv")  ## p-values
+
+library(Hmisc)
+mycor <- rcorr(as.matrix(Correlation), type="pearson")
+mycor$r
+round(mycor$P, 3)
+round(p.adjust(mycor$P, "BY"))
+
+library(corrplot)
+corrplot(mycor)
+write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
+
+cor.test(Correlation$TE, Correlation$B, method = "pearson")
+cor.test(Correlation$TE, Correlation$W, method = "pearson")
+cor.test(Correlation$TE, Correlation$D, method = "pearson")
+cor.test(Correlation$TE, Correlation$N, method = "pearson")
+cor.test(Correlation$TE, Correlation$V, method = "pearson")
+cor.test(Correlation$DI, Correlation$B, method = "pearson")
+cor.test(Correlation$NI, Correlation$B, method = "pearson")
+
+```
+
+# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (LEARNABLE = 600)
 ```{r}
 ## generate the simulation data
-simulation1 <- function (V) {
-  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
-  N_seq <- seq(0, V, by = 10) # sequence of N from 0 to V at an interval of 10
-  D <- rep (D_seq, length(N_seq))
-  N <- rep(N_seq, length(D_seq))
-  data_simulation1 <<- data.frame(V, D, N)
+simulation1 <- function (LEARNABLE) {
+  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
+  NONDOM_seq <- seq(0, LEARNABLE, by = 10) # sequence of N from 0 to V at an interval of 10
+  DOM <- rep (DOM_seq, length(NONDOM_seq))
+  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
+  data_simulation1 <<- data.frame(LEARNABLE, DOM, NONDOM)
 }
@@
 data_simulation1 <- data_simulation1 %>%
-  arrange(D, N) %>%
-  filter(N <= D) %>%
+  arrange(DOM, NONDOM) %>%
+  filter(NONDOM <= DOM) %>%
   # Get word vocabulary W
-  mutate(W = D + N) %>%
+  mutate(WORD = DOM + NONDOM) %>%
   # Create other variables
-  mutate(B = N/(D+N), # balance,
-         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
+  mutate(BALANCE = NONDOM/(DOM+NONDOM), 
+         TE = (DOM*NONDOM)/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
   # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0) %>%
+  mutate_at(vars(-BALANCE), round, 0) %>%
   # bin B into 6 groups
-  mutate(bin_B = cut(B, breaks = 5, 
+  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  arrange(D, N, bin_B)
+  arrange(DOM, NONDOM, bin_BALANCE)
 
@@
   # convert to long data
-  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("D", "N", "W"))) %>%
+  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             D = "Panel 1A: Dominant vocabulary (D)", 
-                             N = "Panel 1B: Non-dominant vocabulary (N)",
-                             W = "Panel 1C: Total vocabulary (W)")) %>%
-  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
+                             DOM = "Panel 1A: Dominant vocabulary \n (DOM)", 
+                             NONDOM = "Panel 1B: Non-dominant vocabulary \n (NONDOM)",
+                             WORD = "Panel 1C: Total vocabulary \n (WORD)")) %>%
+  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
   stat_smooth(method = lm, se = F) +
@@
         panel.background = element_rect(fill = NA, color = "black")) +
-  ylim(0, 200) +
+  ylim(0, 600) +
+  xlim(0, 1000) +
   labs(x = "Number of words", 
        y = "Total number of translation equivalents (TE)", 
-       color = "Balance (B)",
-       linetype = "Balance (B)",
+       color = "Vocabulary balance (BALANCE)",
+       linetype = "Vocabulary balance (BALANCE)",
        title = "1) Simulated data") +
@@
   mutate(vocab_type = recode(vocab_type, 
-                             TotalNW_Dom = "Panel 2A: Dominant vocabulary (D)", 
-                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary (N)",
-                             word_vocab = "Panel 2C: Total vocabulary (W)")) %>%
-  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+                             TotalNW_Dom = "Panel 2A: Dominant vocabulary \n (DOM)", 
+                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary \n (NONDOM)",
+                             word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
@@
   #                         TRUE ~ "0.5")) %>%
-  ggplot(aes(x = number, y = number_of_te, color = bin_B, linetype = bin_B)) +
+  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
   #geom_point() +
   stat_smooth(method = lm, se = F) +
-  #scale_color_manual(values=c("#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
   scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
@@
         panel.background = element_rect(fill = NA, color = "black")) +
+  ylim(0, 600) +
+  xlim(0, 1000) +
   labs(x = "Number of words", 
        y = "Total number of translation equivalents (TE)", 
-       color = "Balance (B)",
-       linetype = "Balance (B)",
+       color = "Vocabulary balance (BALANCE)",
+       linetype = "Vocabulary balance (BALANCE)",
        title = "2) Observed data") +
@@
 
-#ggsave("Output/Observed_plot_inputs_TE.png", Observed_plot_inputs_TE,
-#       width = 10, height = 5)
-
 ### one-way ANOVA to compare the 5 balance groups
-aov_TE_binB <- keepers_ws_TE %>% 
-  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+aov_TE_binBALANCE <- keepers_ws_TE %>% 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  aov(number_of_te ~ bin_B, data = .)
-
-summary(aov_TE_binB)
-
-mean_TE_binB <- keepers_ws_TE %>% 
-  mutate(bin_B = cut(balance_vocab, breaks = 5, 
+  aov(number_of_te ~ bin_BALANCE, data = .)
+
+summary(aov_TE_binBALANCE)
+
+check <- keepers_ws_TE %>% 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))
+
+pairwise.t.test(check$number_of_te, check$bin_BALANCE)
+
+mean_TE_binBALANCE <- keepers_ws_TE %>% 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  group_by(bin_B) %>%
+  group_by(bin_BALANCE) %>%
   summarize(n = n(),
@@
 
-# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (V = 300, 450, or 600)
+# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (LEARNABLE = 300, 450, or 600)
 ```{r}
 ## generate the simulation data
-V <- seq(300, 600, by = 150) # set V to 300, 450, and 600
-
-simulation2 <- function (V) {
-  D_seq <- seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 100
-  N_seq <- seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
-  D <- rep(D_seq, length(N_seq))
-  N <- rep(N_seq, length(D_seq))
-  data_simulation2 <<- data.frame(V, D, N)
+LEARNABLE <- seq(300, 600, by = 150) # set LEARNABLE to 300, 450, and 600
+
+simulation2 <- function (LEARNABLE) {
+  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
+  NONDOM_seq <- seq(0, LEARNABLE, by = 25) # sequence of N from 0 to V at an interval of 25
+  DOM <- rep(DOM_seq, length(NONDOM_seq))
+  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
+  data_simulation2 <<- data.frame(LEARNABLE, DOM, NONDOM)
 }
 
-data_simulation2 <- do.call(rbind, lapply(V, simulation2)) %>%
-  arrange(V, D, N) %>%
-  filter(N <= D) %>%
+data_simulation2 <- do.call(rbind, lapply(LEARNABLE, simulation2)) %>%
+  arrange(LEARNABLE, DOM, NONDOM) %>%
+  filter(NONDOM <= DOM) %>%
   # Create other variables
-  mutate(W = D + N, # word vocabulary W
-         B = N/(D+N), # balance,
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         DI = D-TE, # Unique words in dominant language
-         NI = N-TE) %>% # Unique words in non-dominant language
+  mutate(WORD = DOM + NONDOM, # word vocabulary 
+         BALANCE = NONDOM/(DOM+NONDOM), # balance,
+         TE = (DOM*NONDOM)/LEARNABLE, # If D and N are independent, TE = DN/V
+         DOM_SINGLET = DOM-TE, # Unique words in dominant language
+         NONDOM_SINGLET = NONDOM-TE) %>% # Unique words in non-dominant language
   # Round to keep whole numbers only
-  mutate_at(vars(-B), round, 0) %>%
+  mutate_at(vars(-BALANCE), round, 0) %>%
   # bin B into 3 groups
-  mutate(bin_B = case_when(B >= 0.36 ~ "Most Balanced",
-                           B <= 0.20 ~ "Least Balanced",
+  mutate(bin_BALANCE = case_when(BALANCE >= 0.36 ~ "Most Balanced",
+                                 BALANCE <= 0.20 ~ "Least Balanced",
                                  TRUE ~ "Medium Balanced")) %>%
-  mutate(bin_B  = fct_rev(bin_B)) %>%
-  arrange(D, N, bin_B)
+  mutate(bin_BALANCE  = fct_rev(bin_BALANCE)) %>%
+  arrange(DOM, NONDOM, bin_BALANCE)
 
@@
 simulation2 <- data_simulation2 %>%
-  pivot_longer(-c(V, W, B, bin_B), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %notin% c("D", "N")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("NI", "DI", "TE"))) %>%
-  mutate(V_label = case_when(V == 600 ~ "V = 600",
-                             V == 450 ~ "V = 450",
-                             V == 300 ~ "V = 300")) %>%
-  group_by(V, V_label, bin_B, vocab_type) %>%
+  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
+  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
+                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
+                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
+  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
   summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
   geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ V_label) +
+  facet_grid(. ~ LEARNABLE_label) +
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
+  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
   ylim(0, 600) +
@@
         panel.spacing.x = unit(2, "lines")) + 
-  labs(x = "\nBalance (B)", 
-       y = "Number of concepts (C)", 
+  labs(x = "\nVocabulary balance (BALANCE)", 
+       y = "Number of concepts (CONCEPT)", 
        title = "Panel A: Simulated data")
@@
 ```{r}
-plot_stacked_byV <- keepers_ws_TE %>%
+plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
   select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
   # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
-                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
-                              TRUE ~ "28-33 months (V = 620.4 - 638.9)")) %>%
-  mutate(BalSubset = case_when(balance_vocab >= 0.36 ~ "Most Balanced",
+  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
+                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
+                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
+  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                                balance_vocab <= 0.20 ~ "Least Balanced",
@@
   mutate(TE = number_of_te) %>%
-  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, V_subset, BalSubset, word_vocab, TE), 
+  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
                names_to = "Type", values_to = "NumberOfWords") %>%
@@
   mutate(Type = recode(Type, 
-                       NonDom_unique_words = "Initial non-dominant word (NI)", 
-                       Dom_unique_words = "Initial dominant word (DI)", 
+                       NonDom_unique_words = "Singlet in non-dominant language (NONDOM-SINGLET)", 
+                       Dom_unique_words = "Singlet in dominant language (DOM-SINGLET)", 
                        number_of_te = "Translation equivalent (TE)")) %>%
-  group_by(V_subset, BalSubset, Type) %>%
+  group_by(LEARNABLE_subset, BalSubset, Type) %>%
   summarise(n_words = mean(NumberOfWords)) %>%
@@
   geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ V_subset) +
+  facet_grid(. ~ LEARNABLE_subset) +
   scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
@@
         panel.spacing.x = unit(2, "lines")) + 
-  labs(x = "\nBalance (B)", 
-       y = "Number of concepts (C)",
+  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
+  labs(x = "\nVocabulary balance (BALANCE)", 
+       y = "Number of concepts (CONCEPT)",
        title = "Panel B: Observed data")
@@
 plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
-                                      plot_stacked_byV,
+                                      plot_stacked_byLEARNABLE,
                                       nrow = 2,
-                                      heights=c(0.9,1))
+                                      heights=c(0.8,1))
 
@@
 pairwise.t.test(keepers_ws_TE_3Vsubset$concept_vocab, keepers_ws_TE_3Vsubset$V_subset,
-                 p.adjust.method = "bonferroni")
+                 p.adjust.method = "BH")
 
@@
   select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
-         propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
+         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
   # Subset developmental level (i.e., Average_90percentile) by 3 levels 
@@
                               TRUE ~ "28-33 months")) %>%
-  mutate(BalSubset = case_when(balance_vocab <= 0.65 ~ "Most Balanced",
-                               balance_vocab >= 0.80 ~ "Least Balanced",
+  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
+                               balance_vocab <= 0.20 ~ "Least Balanced",
                                TRUE ~ "Medium Balanced"))
@@
 
-TE_byV_anova <- aov(propTE ~ V_subset, data = stacked_byV_wide)
+TE_byV_anova <- aov(number_of_te ~ V_subset, data = stacked_byV_wide)
 summary(TE_byV_anova)
@@
 
-# Simulation 3: Non-independence of dominant-language (D) and non-dominant language words (N)
+# Simulation 3: Non-independence of dominant-language (DOM) and non-dominant language words (NONDOM)
 ```{r}
 ## generate the simulation data
-V <- seq(150, 600, by = 150) # set V to 300, 450, and 600
-
-simulation3 <- function (V) {
-  D_seq = seq(100, V, by = 100) # sequence of D from 100 to V at an interval of 50
-  N_seq = seq(0, V, by = 25) # sequence of N from 0 to V at an interval of 25
-  D <- rep(D_seq, length(N_seq))
-  N <- rep(N_seq, length(D_seq))
-  data_simulation3 <<- data.frame(V, D, N)
+LEARNABLE <- seq(150, 600, by = 150) # set LEARNABLE to 300, 450, and 600
+
+simulation3 <- function (LEARNABLE) {
+  DOM_seq = seq(100, LEARNABLE, by = 100) # sequence of DOM from 100 to LEARNABLE at an interval of 100
+  NONDOM_seq = seq(0, LEARNABLE, by = 25) # sequence of NONDOM from 0 to LEARNABLE at an interval of 25
+  DOM <- rep(DOM_seq, length(NONDOM_seq))
+  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
+  data_simulation3 <<- data.frame(LEARNABLE, DOM, NONDOM)
 }
 
-data_simulation3 <- do.call(rbind, lapply(V, simulation3)) %>%
-  arrange(V, D, N) %>%
-  filter(N <= D) %>% # eliminate cases when N > D
-  mutate(W = D+N, # word vocabulary W
-         B = N/(D+N), # balance B
-         TE = D*N/V) %>% # If D and N are independent, TE = DN/V
-  mutate_at(vars(-B), round, 0) %>%
-  mutate(bin_B = cut(B, breaks = 5, 
+data_simulation3 <- do.call(rbind, lapply(LEARNABLE, simulation3)) %>%
+  arrange(LEARNABLE, DOM, NONDOM) %>%
+  filter(NONDOM <= DOM) %>% # eliminate cases when NONDOM > DOM
+  mutate(WORD = DOM+NONDOM, # word vocabulary
+         BALANCE = NONDOM/(DOM+NONDOM), # balance 
+         TE = DOM*NONDOM/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
+  mutate_at(vars(-BALANCE), round, 0) %>%
+  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  arrange(bin_B)
-
-## Simulation3: TE = IND * (D*N/V)
-### defining the independence parameter I
-IND1 = 1.5 # IND > 1, TEs are MORE easily learned than unique words
-IND2 = 0.5 # IND < 1, TEs are LESS easily learned than unique words
+  arrange(bin_BALANCE)
+
+## Simulation3: TE = BIAS * (DOM*NONDOM/LEARNABLE)
+### defining the bias parameter I
+BIAS1 = 1.5 # BIAS > 1, TEs are MORE easily learned than unique words
+BIAS2 = 0.5 # BIAS < 1, TEs are LESS easily learned than unique words
 
@@
 simulation3 <- data_simulation3 %>%
-  ## calculate where IND > 1
-  mutate(TE_easier = IND1*(D*N/V)) %>%
-  ## calculate where IND < 1
-  mutate(TE_harder = IND2*(D*N/V)) %>%
+  mutate(TE_easier = BIAS1*(DOM*NONDOM/LEARNABLE), # calculate where BIAS > 1
+         TE_harder = BIAS2*(DOM*NONDOM/LEARNABLE)) %>% # calculate where BIAS < 1
   # round to keep whole numbers only
-  mutate_at(vars(-c(B, bin_B)), round, 0) %>%
+  mutate_at(vars(-c(BALANCE, bin_BALANCE)), round, 0) %>%
   # prepare dataset to plot
@@
   mutate(Model = recode(Model, 
-                        TE = "IND = 1 (Independence Account)",
-                        TE_easier = "IND = 1.5 (Enhancement Account)",
-                        TE_harder = "IND = 0.5 (Avoidance Account)")) %>%
+                        TE = "BIAS = 1 (Neutral Account)",
+                        TE_easier = "BIAS = 1.5 (Preference Account)",
+                        TE_harder = "BIAS = 0.5 (Avoidance Account)")) %>%
   # ggplot
-  ggplot(aes(x = W, y = number_TE, color = bin_B, linetype = bin_B)) +
+  ggplot(aes(x = WORD, y = number_TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
   stat_smooth(method = lm, se = F) +
@@
   scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
-  # scale_color_discrete(direction = -1) +
   theme_minimal() + 
@@
         panel.background = element_rect(fill = NA, color = "black")) +
-  labs(x = "Total word vocabulary (W)", 
+  labs(x = "Total word vocabulary (WORD)", 
        y = "Total number of translation equivalents (TE)", 
-       color = "Balance (B)",
-       linetype = "Balance (B)") +
+       color = "Vocabulary balance (BALANCE)",
+       linetype = "Vocabulary balance (BALANCE)") +
   guides(colour = guide_legend(reverse = T, nrow = 1),
@@
 
-# Descriptives
-## Gender 
-```{r}
-keepers_ws_TE %>%
-  group_by(gender) %>% 
-  distinct(baby_id, .keep.all=T) %>%
-  count()
-```
-
-## English dominant
-```{r}
-# dominance defined by language exposure
-keepers_ws_TE %>% group_by(lang_dom) %>% distinct(baby_id) %>%
-  count()
-
-# dominance defined by vocab size
-keepers_ws_TE %>% 
-  distinct(ID_testdate, .keep_all = TRUE) %>%
-  mutate(Eng_Dom = ifelse(LangDom_vocab == "English", 1, 0)) %>%
-  summarise(n = n(),
-            n_EngDom = sum(Eng_Dom),
-            percent_EngDom = n_EngDom/n * 100,
-            percent_FrDom = 100 - percent_EngDom)
-
-```
-
-## demographics
-```{r}
-# How many data points including some babies who made multiple visits to the lab (N)
-keepers_ws_TE %>% 
-  distinct(ID_testdate) %>% 
-  count()
-
-N_duplicatedBabyID <- keepers_ws_TE %>%
-  filter(duplicated(baby_id))
-
-# how many unique babies
-keepers_ws_TE %>%
-  distinct(baby_id) %>%
-  count()
-
-keepers_ws_TE %>%
-  distinct(participant_id) %>%
-  count()
-
-# Create dataframe of demographics
-d_participants <- keepers_ws_TE %>%
-  select(ID_testdate, gender,age_continuous, years_education, lang_exp_eng, lang_exp_fre, lang_exp_other) %>%
-  mutate(female = ifelse(gender == "F", 1, 0)) %>%
-  pivot_longer(-c(ID_testdate), names_to = "info", values_to = "value") %>%
-  group_by(info) %>%
-  summarise(n_subject = n_distinct(ID_testdate),
-            mean = mean(value, na.rm=T),
-            sd = sd(value, na.rm=T),
-            min = min(value, na.rm=T),
-            max = max(value, na.rm=T))
-```
-
-## average # of vocabulary
-```{r}
-vocabulary_average <- keepers_ws_TE %>%
-  select(ID_testdate, word_vocab, TotalNW_Dom, TotalNW_NonDom, concept_vocab,
-         number_of_te, Dom_unique_words, NonDom_unique_words) %>%
-  pivot_longer(-ID_testdate, names_to = "vocabulary", values_to = "number") %>%
-  group_by(vocabulary) %>%
-  summarise(mean = mean(number, na.rm = T), 
-            sd = sd(number, na.rm = T),
-            min = min(number, na.rm = T),
-            max = max(number, na.rm = T)) %>%
-  mutate_if(is.numeric, round, digits = 2)
-
-## paired-sample t-test to compare D and N
-t.test(keepers_ws_TE$TotalNW_Dom, keepers_ws_TE$TotalNW_NonDom, 
-       paired = TRUE)
-
-## paired-sample t-test to compare DI and NI
-t.test(keepers_ws_TE$Dom_unique_words, keepers_ws_TE$NonDom_unique_words, 
-       paired = TRUE) 
-```
-
-## average balance and comparison with balance by input
-```{r}
-### Average balance by vocab
-keepers_ws_TE %>%
-  summarise(mean = mean(balance_vocab, na.rm = T),
-            sd = sd(balance_vocab, na.rm = T),
-            min = min(balance_vocab, na.rm = T),
-            max = max(balance_vocab, na.rm = T))
-### grouped by language
-keepers_ws_TE %>%
-  group_by(LangDom_vocab) %>%
-  summarise(mean = mean(balance_vocab, na.rm = T),
-            sd = sd(balance_vocab, na.rm = T),
-            min = min(balance_vocab, na.rm = T),
-            max = max(balance_vocab, na.rm = T))
-
-t.test(balance_vocab ~ LangDom_vocab, keepers_ws_TE) # t-test to compare the two language groups
-
-### Comparsion with balance by input (which is defined with reference to the lang_dom defined by vocab size)
-keepers_ws_TE %>%
-  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
-  summarise(n = n(),
-            n_consistent = sum(consistent_LangDom),
-            percentage_consistent = n_consistent/n*100,
-            n_inconsistent = n - n_consistent,
-            percentage_inconsistent = n_inconsistent/n*100)
-
-keepers_ws_TE_withoutInconsistent <- keepers_ws_TE %>%
-  mutate(consistent_LangDom = if_else(lang_dom != LangDom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
-  filter(consistent_LangDom == 1)
-
-### Correlation between balance by vocabulary and balance by input
-cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$balance_input, method = "pearson")
-
-### Correlation between balance by vocabulary and raw language exposure to the non-dominant language
-cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input, method = "pearson")
-
-plot(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input,
-     xlab = "Vocabulary balance (= N/W)",
-     ylab = "raw exposure to the non-dominant language") +
-  lines(lowess(keepers_ws_TE$balance_vocab, keepers_ws_TE$LangNonDom_input), col="blue") # lowess line (x,y)
-
-  abline(lm(keepers_ws_TE$LangNonDom_input~keepers_ws_TE$balance_vocab), col="red") # regression line (y~x)
-lines(lowess(wt,mpg), col="blue") # lowess line (x,y)
-
-```
-
-# Correlation
-```{r}
-Correlation <- keepers_ws_TE %>%
-  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, TotalNW_Dom, TotalNW_NonDom,
-         number_of_te, Dom_unique_words, NonDom_unique_words, concept_vocab) %>%
-  # rename variables
-  rename(Age_month = age_continuous,
-         V = Average_90percentile,
-         B = balance_vocab, 
-         W = word_vocab,
-         D = TotalNW_Dom, 
-         N = TotalNW_NonDom, 
-         TE = number_of_te, 
-         DI = Dom_unique_words, 
-         NI = NonDom_unique_words, 
-         C = concept_vocab)
-
-# Correlation with corrections for multiple comparisons
-library(psych)
-TE_corr <- corr.test(Correlation, use = "pairwise", method = "pearson", adjust = "BY")
-
-write.csv(round(TE_corr$r, 3), file="Output/ObservedTE_correlations_r.csv")  ## coefficients
-write.csv(round(TE_corr$p, 3), file="Output/ObservedTE_correlations_p.csv")  ## p-values
-
-library(Hmisc)
-mycor <- rcorr(as.matrix(Correlation), type="pearson")
-mycor$r
-round(mycor$P, 3)
-round(p.adjust(mycor$P, "BY"))
-
-library(corrplot)
-corrplot(mycor)
-write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
-
-cor.test(Correlation$TE, Correlation$B, method = "pearson")
-cor.test(Correlation$TE, Correlation$W, method = "pearson")
-cor.test(Correlation$TE, Correlation$D, method = "pearson")
-cor.test(Correlation$TE, Correlation$N, method = "pearson")
-cor.test(Correlation$TE, Correlation$V, method = "pearson")
-cor.test(Correlation$DI, Correlation$B, method = "pearson")
-cor.test(Correlation$NI, Correlation$B, method = "pearson")
-
-```
-
 # Figure 5: Stacked barchart to plot TE, DI, and NI of individual infants (facet by balance)
@@
   #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
-  mutate(BalSubset = case_when(balance_vocab >= 0.35 ~ "MostBalanced",
+  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "MostBalanced",
                                balance_vocab <= 0.20 ~ "LeastBalanced",
@@
   # Subset balance_vocab by 3 levels 
-  mutate(BalSubset = case_when(balance_vocab >= 0.35 ~ "MostBalanced",
+  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "MostBalanced",
                                balance_vocab <= 0.20 ~ "LeastBalanced",
@@
 ```
-
-# Statistical tests
-## ANOVA for balance
-```{r}
-res.aov <- aov(propTE ~ balance_group15*cent.age.months, data=keepers_ws_TE) 
-summary(res.aov)
-```
-
-## Age (old code, not used for poster)
-### Running linear model
-```{r}
-  linear.model<-lmer(propTE ~ cent.age.months+ (1|Baby_ID), data = keepers_ws_TE)
-  summary(linear.model)
-```
-### Running quadratic model
-```{r}
- quadratic.model <-lmer(propTE ~ poly(cent.age.months,2) + (1|Baby_ID), data = keepers_ws_TE)
-  summary(quadratic.model)
-```
-### Cubic model
-```{r}
-cubic.model<-lm(propTE ~ poly(cent.age.months,3), data = keepers_ws_TE)
-summary(cubic.model)
-```

### 2021-06-03 Analyze_TEs.Rmd
updated analysis

@@
   rename(Age_month = age_continuous,
-         V = Average_90percentile,
-         B = balance_vocab, 
-         W = word_vocab,
-         D = TotalNW_Dom, 
-         N = TotalNW_NonDom, 
+         LEARNABLE = Average_90percentile,
+         BALANCE = balance_vocab, 
+         WORD = word_vocab,
+         DOM = TotalNW_Dom, 
+         NONDOM = TotalNW_NonDom, 
          TE = number_of_te, 
-         DI = Dom_unique_words, 
-         NI = NonDom_unique_words, 
-         C = concept_vocab)
+         "DOM-SINGLET" = Dom_unique_words, 
+         "NONDOM-SINGLET" = NonDom_unique_words, 
+         CONCEPT = concept_vocab)
 

### 2021-06-09 simulations_old_versions/computational model/TE_computation_Simulation.Rmd
organizing folders

--- 2021-03-26 computational model/TE_computation_Simulation.Rmd
@@
 
+```
+
+# Simulation 1
+## Infants of the same developmental level with different word vocabularies and balances of exposure
+### V = 600
+```{r}
+V = 600
+
+D_seq = seq(50, 600, by = 25)
+N_seq = seq(0, 600, by = 25)
+
+D = rep(D_seq, length(N_seq))
+N = rep(N_seq, length(D_seq))
+
+data_Simulation1 <- data.frame(V, D, N) %>%
+  arrange(D, N) %>%
+  filter(N <= D) %>%
+  # Get word vocabulary W
+  mutate(W = D + N) %>%
+  # Create other variables
+  mutate(B = N/(D+N), # balance,
+         TE = (D*N)/V) %>% # If D and N are independent, TE = DN/V
+  # Round to keep whole numbers only
+  mutate_at(vars(-B), round, 0) %>%
+  # bin B into 6 groups
+  mutate(bin_B = cut(B, breaks = 5, 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  arrange(W, bin_B)
+
+# Visualizing Simulation 1
+Simulation1 <- data_Simulation1 %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N), names_to = "vocab_type", values_to = "number") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             W = "Panel 1A: Total vocabulary (W)", 
+                             D = "Panel 1B: Dominant vocabulary (D)", 
+                             N = "Panel 1C: Non-dominant vocabulary (N)")) %>%
+  ggplot(aes(x = number, y = TE, color = bin_B, linetype = bin_B)) +
+  stat_smooth(method = lm, se = F) +
+  scale_color_manual(values=c("#bdbdbd", "#969696", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotted", "dotdash", "dashed", "twodash", "solid")) +
+  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
+  #scale_color_discrete(direction = -1) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom",
+        plot.title = element_text(face = 'bold', size = 18, hjust = 0, vjust = 2),
+        plot.title.position = "plot",
+        text = element_text(size=14),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) +
+  ylim(0, 200) +
+  xlim(0, 700) +
+  labs(x = "Number of words", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)",
+       linetype = "Balance (B)",
+       title = "1) Simulated data") +
+  guides(color = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
+
+ggsave("plot_Simulation1.png", Simulation1,
+       width = 10, height = 5)
 ```

### 2021-06-09 simulations_old_versions/computational model/old/old_TE_computation_Simulation.Rmd
organizing folders

--- 2020-12-24 computational model/TE_computation.Rmd
@@
 
-# Generating dataframe for computational model
+# Read in data
 ```{r}
-#### Set 1: Calculate other variables based on V, W, and B ####
-## Creating variables: V, W, and B
-# Items on vocabulary measure (e.g. words on the CDI)
-V <- 600
+TE_computation <- read.csv(here::here("computational model/TE_computation.csv"))
+```
 
-# Total word vocabulary of a child
-W <- seq(0, 2*V, by = 25)
+# Simulation 1
+## Infants of the same developmental level with different word vocabularies and balances of exposure
+### V = 600
+```{r}
+Simulation1 <- TE_computation %>%
+  # V = 600
+  filter(V == 600) %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("W", "D", "N")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("W", "D", "N"))) %>%
+  mutate(vocab_type = recode(vocab_type, 
+                             W = "Panel A: Total vocabulary (W)", 
+                             D = "Panel B: Dominant vocabulary (D)", 
+                             N = "Panel C: Non-dominant vocabulary (N)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  ggplot(aes(x = number, y = TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Number of words", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)"
+       #, title = "Number of translation equivalent by balance in relation to different input measures "
+       ) +
+  guides(colour = guide_legend(nrow = 1))
 
-# Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-B <- seq(0.5, 1, by = 0.025)
-
-# Generating sequence of W and B
-W_seq <- W %>% # a sequence of 0 to 2V
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>%  # a sequence of 0.5 to 1
-  rep(., length(W)) # generate N repetitions of the sequence based on length of W
-
-## Creating the dataframe
-TE_fromVWB <- data.frame(V, W_seq) %>%
-  arrange(W_seq) %>%
-  rename(W = W_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(D = B*W, # Words produced in the dominant language
-         N = (1-B)*W, # Words produced in non-dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where D or N aren't whole numbers
-  filter_at(vars(D, N), any_vars(is.wholenumber(.)))
-
-#### Set 2: Calculate other variables based on V, D, and N ####
-## Creating variables: D and N
-D <- seq(0, V, by = 25)
-
-N <- seq(0, V, by = 25)
-
-# Generating sequence of D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-N_seq <- N %>% # Words produced in non-dominant language
-  rep(., length(D)) # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_fromVDN <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(N = N_seq) %>%
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # Create other variables
-  mutate(W = D+N, # Total word vocabulary of a child
-         B = D/(D+N), # Balance, i.e. proportion use of non-dominant language (B ≤ .50)
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>%
-  # Remove cases where B = NA
-  filter(!is.na(B))
-
-#### Set 3: Calculate other variables based on V, D, and B ####
-## Creating variables: D and N
-D_seq <- D %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # balance
-  rep(., length(D))  # generate N repetitions of the sequence based on length of D
-
-## Creating the dataframe
-TE_fromVDB <- data.frame(V, D_seq) %>%
-  arrange(D_seq) %>%
-  rename(D = D_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(W = D/B, # Total word vocabulary of a child
-         N = W-D, # Words produced in non-dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where N isn't whole number
-  filter(is.wholenumber(N))
-
-#### Set 4: Calculate other variables based on V, N, and B ####
-## Creating variables: D and N
-N_seq <- N %>% # Words produced in the dominant language
-  rep(., length(B)) # generate N repetitions of the sequence based on length of B
-
-B_seq <- B %>% # a sequence of 0.5 to 1
-  rep(., length(N)) # generate N repetitions of the sequence based on length of N
-
-## Creating the dataframe
-TE_fromVNB <- data.frame(V, N_seq) %>%
-  arrange(N_seq) %>%
-  rename(N = N_seq) %>%
-  mutate(B = B_seq) %>%
-  # Create other variables
-  mutate(W = D/(1-B), # Total word vocabulary of a child
-         D = W-N, # Words produced in dominant language
-         TE = (D*N)/V, # If D and N are independent, TE = DN/V
-         C = W-TE, # Concept vocabulary (or total conceptual vocabulary size)
-         Du = D-TE, # Unique words in dominant language
-         Nu = N-TE) %>% # Unique words in non-dominant language
-  # It is not possible for N > D, so keep only cases where D >= N
-  filter(D >= N) %>%
-  # In reality, it would be impossible to get negative numbers for any vocabulary variables, so filter out all cases with negative numbers
-  filter(Nu >= 0) %>%
-  # It is also not possible for D or N to exceed V, so filter out cases where D or N > V
-  filter_at(vars(D, N), all_vars(. <= 600)) %>%
-  # filter out cases where D isn't whole number
-  filter(is.wholenumber(D)) %>%
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.))) %>%
-  filter(is.wholenumber(Nu))
-
-#### Combine the 4 sets ####
-TE_computation <- TE_fromVWB %>%
-  bind_rows(TE_fromVDN, TE_fromVDB, TE_fromVNB) %>%
-  # constrain all variables but B (balance) to be integers 
-  filter_at(vars(TE, C, Du, Nu), any_vars(is.wholenumber(.)))
+ggsave("plot_Simulation1.png", plot_inputs_TE,
+       width = 10, height = 5)
 
@@
 
-# Correlation
+# Simulation 2
+## Patterns of translation equivalent learning as a function of different developmental timepoints
 ```{r}
-Correlation <- TE_computation %>%
-  select(-V) %>%
-  cor_matrix()
+timepoints <- c(150, 300, 450, 600)
 
-mycor <- rcorr(as.matrix(TE_computation), type="pearson")
-mycor$r
-mycor$P
+Simulation2 <- TE_computation %>%
+  filter(V %in% timepoints) %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type %notin% c("D", "N", "C")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("Nu", "Du", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom",
+        plot.title = element_text(hjust = 0.5)) + 
+  labs(x = "Balance", 
+       y = "Number of concepts", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
 
-library(corrplot)
-mycor <- cor(TE_computation)
-corrplot(mycor)
-#write.table(Correlation, "TEcomputation_correlations.tsv", col.names=NA)
+ggsave("plot_Simulation2.png", plot_composition_byV,
+        width = 10, height = 5)
+
+```
+
+# Simulation 3
+## Non-independence of dominant-language (D) and non-dominant language words (N)
+```{r}
+
+# I * (D*N/V)
+I1 = 1.5
+I2 = 0.5
+
+Simulation3 <- TE_computation %>%
+  filter(V==600) %>%
+  ## Calculate where I > 1
+  mutate(TE_easier = I1*(D*N/V)) %>%
+  ## Calculate where I < 1
+  mutate(TE_harder = I2*(D*N/V)) %>%
+  pivot_longer(c(TE, TE_harder, TE_easier), 
+               names_to = "Model", values_to = "number_TE") %>%
+  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
+  mutate(Model = recode(Model, 
+                        TE = "I = 1 (assuming independence of D and N)",
+                        TE_easier = "I = 1.5 (TEs are more easily learned)",
+                        TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  ggplot(aes(x = W, y = number_TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ Model) +
+  ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Total word vocabulary (W)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_Simulation3.png", Simulation3, 
+       width = 10, height = 5)
+
+#Simulation3 <- TE_computation %>%
+#  pivot_longer(c(TE_harder, TE_easier), 
+#               names_to = "Nonindependence_type", values_to = "Simulated_TEs") %>%
+#  mutate(Nonindependence_type = recode(Nonindependence_type, 
+#                                       TE_easier = "I = 1.5 (TEs are more easily learned)",
+#                                       TE_harder = "I = 0.5 (TE are less easily learned)")) %>%
+#  ggplot(aes(y = Simulated_TEs, x = TE, color = Nonindependence_type)) +
+#  geom_smooth(method = "lm", se = F) +
+#  geom_point(shape = 23) +
+#  geom_abline(linetype = 2, alpha = 0.5) +
+#  labs(y = "Simulated number of TEs assuming non-independence of D and N", 
+#       x = "Expected number of TEs assuming independence of D and N",
+#       color = "Scenarios"
+#       #, title = "Non-independence vs. Independence of dominant and non-dominant words"
+#       ) +
+#  theme(plot.title = element_text(hjust = 0.5)) +
+#  theme_minimal()
+
+##ggsave("plot_Simulation3.png", Simulation3, 
+##       width = 8, height = 6)
+
 ```
@@
        width = 10, height = 5)
+
+plot_inputs <- TE_computation %>%
+  # Convert to long data
+  pivot_longer(c(W, D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("W", "D", "N", "TE")) %>%
+  mutate(vocab_type = as.factor(vocab_type)) %>%
+  mutate(vocab_type = fct_relevel(vocab_type, "W")) %>%
+  mutate(vocab_type = recode(vocab_type, W = "Total vocabulary (W)", D = "Dominant vocabulary (D)", 
+                             N = "Non-dominant vocabulary (N)", TE = "Translation equivalents (TE)")) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 12,
+  #                   labels = c(100,200, 300, 400, 500, 600, 700, 800, 900, 1000, 1100, 1200))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = V, y = n_words, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ vocab_type) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Different timepoints = Learnable vocaobulary (V)", 
+       y = "Number of words", 
+       color = "Balance",
+       title = "Word vocabulary size by balance under different input measures and translation equivalents") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs.png", plot_inputs,
+       width = 10, height = 5)
+
+plot_inputs_TE_V <- TE_computation %>%
+  filter(W == 250) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  ggplot(aes(x = V, y = TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  #facet_grid(. ~ bin_W) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Developmental timepoint (V)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+plot_inputs_TE_V <- TE_computation %>%
+  filter(W == 400) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  #mutate(bin_W = cut_interval(W, 3,
+  #                   labels = c("W = 0 - 400" , "W = 401 - 800", "W = 801 - 1200"))) %>%
+  ggplot(aes(x = V, y = TE, color = bin_B)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  #facet_grid(. ~ bin_B) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Developmental timepoint (V)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs_TE_V.png", plot_inputs_TE_V,
+       width = 6, height = 6)
+
+plot_inputs_TE_V <- TE_computation %>%
+  #filter(W == 100) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  mutate(bin_W = cut_interval(W, 4,
+                     labels = c("W = 0 - 300" , "W = 301 - 600", "W = 601 - 900", "W = 901 - 1200"))) %>%
+  ggplot(aes(x = V, y = TE, color = bin_W)) +
+  #geom_point() +
+  stat_smooth(method = lm, se = F) +
+  theme_minimal() + 
+  facet_grid(. ~ bin_B) +
+  #ylim(0, 600) +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  labs(x = "Developmental timepoint (V)", 
+       y = "Total number of translation equivalents (TE)", 
+       color = "Balance (B)") +
+  guides(colour = guide_legend(nrow = 1))
+
+ggsave("plot_inputs_TE_V.png", plot_inputs_TE_V,
+       width = 10, height = 5)
+
+```
+
+## Plotting different scenarios
+```{r}
+plot_scenarios <- TE_computation %>%
+  pivot_longer(c(D, N, TE, C, Du, Nu), names_to = "vocab_type", values_to = "number") %>%
+  filter(vocab_type == c("D", "N", "TE")) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("N", "D", "TE"))) %>%
+  mutate(bin_B = cut(B, breaks = 6, 
+                     labels = c("0.5", "0.6", "0.7", "0.8", "0.9", "1"))) %>%
+  group_by(V, bin_B, vocab_type) %>%
+  summarise(n_words = mean(number)) %>%
+  ggplot(aes(x = bin_B, y = n_words, fill = vocab_type)) + 
+  geom_bar(position="stack", stat="identity") +
+  facet_grid(. ~ V) +
+  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  #scale_y_continuous(limits = c(0, 1200)) +
+  theme_minimal() +
+  theme(axis.text.x = element_text(angle = 90), legend.position = "bottom") + 
+  labs(x = "Balance", 
+       y = "Number of words (W)", 
+       title = "Vocabulary composition across different developmental timepoints (V)")
+
 ```

### 2021-06-09 simulations_old_versions/simulation/vocab_simulation.Rmd
organizing folders

---
title: "Simulation of TE learning"
output: html_notebook
---

This is a first attempt to simulate TE learning in bilinguals.

Hypothetical baby learing 70% English and 30% French

```{r}
library(tidyverse)
library(purrr)
library(corrr)

set.seed(6)

```

Function for simple vocabulary simulation. All words of equal difficulty.

```{r vocab-sim}
## Function that simulates vocab for one baby

# prop_dom is proportion exposure to dominant language
# learn_cutoff is the number of experiences (encounters) needed to learn a word
# experience is the number of times a word could be encountered
# te_cutoff is multiplier for threshold when a TE is lexicalized. te_cutoff = 1 means same experienc needed to learn TE as unique word. Higher values --> harder to learn TEs, i.e. te_cutoff = .5 implies half the experience is needed to learn the TE

VOCAB_SIZE = 100 # Vocab size set here for all simulations.

vocab_sim <- function(prop_dom, learn_cutoff = 30, experience = 50, te= 1) {

  learn_cutoff_first = learn_cutoff # cutoff if first word learned
  learn_cutoff_second = learn_cutoff*te # cutoff if second word learned

  prop_dom = prop_dom # proportion exposure to dominant language
  prop_nondom = 1- prop_dom # proportion exposure to non-dominant language

# Samples how often individual words will be encountered, proportional to exposure
# sample() takes a sample of the specified size from the elements of x using either with or without replacement.

#experience_dom <- prop_dom * sample(1:experience, experience, replace = TRUE)
#experience_nondom <- prop_nondom * sample(1:experience, experience, replace = TRUE)

experience_dom <- prop_dom * sample(1:experience, VOCAB_SIZE, replace = TRUE)
experience_nondom <- prop_nondom * sample(1:experience, VOCAB_SIZE, replace = TRUE)

# Based on above vectors, determine whether each concept is labeled uniquely, or has a TE
vocab <- data.frame(prop_dom, experience_dom, experience_nondom) %>%
  mutate(prop_dom = prop_dom) %>%
  mutate(learn_cutoff = learn_cutoff) %>%
  mutate(experience = experience) %>%
  mutate(te = te) %>%
  mutate(learned_dom = case_when(
    experience_dom >= experience_nondom & experience_dom >= learn_cutoff_first ~ T,
    experience_dom >= experience_nondom & experience_dom < learn_cutoff_first ~ F,
    experience_dom < experience_nondom & experience_dom >= learn_cutoff_second ~ T,
    experience_dom < experience_nondom & experience_dom < learn_cutoff_second ~ F

  )) %>%
  mutate(learned_nondom = case_when(
    experience_nondom >= experience_dom & experience_nondom >= learn_cutoff_first ~ T,
    experience_nondom >= experience_dom & experience_nondom < learn_cutoff_first ~ F,
    experience_nondom < experience_dom & experience_nondom >= learn_cutoff_second ~ T,
    experience_nondom < experience_dom & experience_nondom < learn_cutoff_second ~ F
  )) %>%
#  mutate(learned_dom = if_else(experience_dom >= learn_cutoff, T, F)) %>%
#  mutate(learned_nondom = if_else(experience_nondom >= learn_cutoff, T, F)) %>%
  mutate(learned_unique_dom = if_else(learned_dom & !learned_nondom, T, F)) %>%
  mutate(learned_unique_nondom = if_else(learned_nondom & !learned_dom, T, F)) %>%
  mutate(learned_unique_all = as.logical(learned_unique_dom + learned_unique_nondom)) %>%
  mutate(learned_te = if_else(learned_dom & learned_nondom, T, F)) %>%
  mutate(learned_concepts = if_else(learned_dom | learned_nondom, T, F)) %>%
  select(-experience_dom, -experience_nondom)

return(vocab)
}

# Check that the above function works

vocab_sim(.5, 20, 100, 1.0)

```

Creates summary of simulation

```{r vocab-sim-summary}
## Function that summarizes vocab

vocab_sim_summary <- function(prop_dom, learn_cutoff=15, experience=100, te = 1) {

  vocab <- vocab_sim(prop_dom, learn_cutoff, experience, te) %>%
    group_by(prop_dom, learn_cutoff, experience, te) %>%
  summarize_all(sum)

  return(vocab)

}

## Check that the above function words

vocab_sim_summary(.5, 20, 100)

```

```{r parameter-setup}

# This chunk creates the dataframes/vectors of the space of parameters for the simulation to model

# Create vector with proportions from .1 to .5
seq_prop <-seq(from = .5, to = 1, by = .1)
props <- rep(seq_prop, 25) # 20 repetitions of the different proportion exposures

# For now keep cutoff constant.  Can use this code in the future if necessary

# seq_cutoff <-seq(from = 30, to = 50, by = 10)
# cutoffs <- rep(seq_cutoff, length(props)/length(seq_cutoff))

cutoffs <- rep(50, length(props))

# Create vector with different levels of experience

seq_experience <- seq(from = 50, to = 450, by = 100)
experience <- rep(seq_experience, length(props)/length(seq_experience))

# Create vector for different TE weights.  For now, just use .5, 1, 1.5

seq_te <- c(.5, 1, 1.5)
te <- rep(seq_te, length(props)/length(seq_te))

# Check that the lengths for each vector are equal. These should all return "TRUE"

length(cutoffs) == length(props)
length(experience) == length(props)
length(te) == length(props)

# Get all combinations, then extract columns as lists.  Couldn't find more elegant solution
combo1 <- crossing(props, cutoffs, experience, te) # Full factorial combination of each vector

# Repeat each combo 20 times

combo <- map_dfr(seq_len(20), ~combo1)

# Now need to extract each of the columns separately.  Could be coded more elegantly, but this works for now.

props1 <- combo$props
cutoffs1 <- combo$cutoffs
experience1 <- combo$experience
te1 <- combo$te

# Check how many infants will be simulated

length(props1)

```
Run simulation across all the parameter combinations created in the previous step

```{r}
# Run simulation. This step might take a while

vocab_simulations_full <- 
  pmap(list(props1, cutoffs1, experience1, te1), vocab_sim_summary) %>%
  bind_rows()

# Cleanup and recode some of the columns for easier visualization/interpretation

vocab_simulations <- vocab_simulations_full%>%
  mutate(te = as.factor(as.character(te))) %>%
  mutate(te = recode(te, "0.5" = "te_easier" , "1" = "te_equal", "1.5" = "te_harder")) %>%
  mutate(learn_cutoff = as.factor(as.character(learn_cutoff))) %>%
  mutate(learn_cutoff = recode(learn_cutoff, "30" = "words_easy", "40" = "words_medium", "50" = "words_hard")) %>%
  mutate(learned_words = learned_dom + learned_nondom) %>%
  mutate(prop_dom = as.factor(as.character(prop_dom)))

# Pivot the data into long format

vocab_simulations_long <- vocab_simulations %>%
  pivot_longer(cols = starts_with("learned"), names_to = "word_type", values_to = "n_words")

```

```{r correlations}

# Calculating correlations between different vocabulary measures

all_correl <- vocab_simulations %>%
  ungroup() %>%
    mutate(prop_dom = as.numeric(as.character(prop_dom))) %>%
  filter(prop_dom <= .75) %>%
  mutate(experience = as.numeric(as.character(experience))) %>%
  mutate(learned_propte = learned_te/learned_words) %>%
  relocate(learned_te, learned_propte, learned_dom, learned_nondom, learned_words, learned_concepts) %>%
  group_by(te) %>%
  select(starts_with("learned")| starts_with("prop")|starts_with("experience")) %>%
  group_map(~correlate(.x))

te_neutral <- all_correl[[2]] %>%
  mutate(across(is.numeric, round, 2))

# write_tsv(te_neutral, "correlations.tsv")

```

```{r}

# Plot to compare different vocabulary measures under different TE learning strategies, across experience

plot_words <- vocab_simulations_long %>%
  mutate(prop_dom = as.factor(prop_dom)) %>%
  filter(word_type %in% c("learned_words", "learned_concepts", "learned_dom", "learned_nondom", "learned_te")) %>%
  ggplot(aes(x = experience, y = n_words, color = prop_dom)) +
  geom_point() +
  stat_smooth(method = lm, se = F) +
  theme_minimal() + 
  facet_grid(te ~ word_type) +
  ggtitle("Word vocabulary size by dominance under different measures") +
  theme(axis.text.x = element_text(angle = 90))

ggsave("plot_plot_words.png", plot_words)
plot_words

```

Vocabulary composition across different TE scenarios
```{r}
# Plot of simulations

plot_composition <- vocab_simulations_long %>%
 # filter(te == "te_equal") %>%
 # filter(experience == 350) %>% # Fix a level of experience
  filter(word_type %in% c("learned_te", "learned_unique_dom", "learned_unique_nondom")) %>%
  mutate(word_type = fct_relevel(word_type, "learned_unique_nondom", "learned_unique_dom", "learned_te")) %>%
 # group_by(learn_cutoff) %>%
  group_by(prop_dom, word_type, learn_cutoff, experience, te) %>%
  summarize(n_words = mean(n_words)) %>%
 # mutate(word_type = fct_rev(word_type)) %>%
  ggplot(aes(x = prop_dom, y = n_words, fill = word_type)) +
   geom_bar(position = "stack", stat = "identity", reverse = TRUE) +
  facet_grid(te ~ experience) +
  theme_minimal() +
    ggtitle("Vocabulary composition across different TE scenarios, with different proportions of dominant language exposure") +
     theme(axis.text.x = element_text(angle = 90), legend.position = "bottom")

ggsave("plot_composition.png", plot_composition)
plot_composition

```

```{r}

# non-dominant vocab predicts TEs, but stronger prediction for less balanced infants

vocab_simulations %>%
  filter(te == "te_equal") %>%
  mutate(experience = as.factor(as.character(experience))) %>%
  ggplot(aes(x = learned_nondom, y = learned_te, color = prop_dom)) +
 # facet_grid(. ~ experience) +
  geom_point() +
  theme_minimal() +
 # stat_smooth(method = "lm") +
    ggtitle("Predicting TEs from non-dominant vocabulary")

```

```{r}
# Basically shows that simulations converge on the same thing as computations that are now in the paper

vocab_simulations %>%
filter(te == "te_equal") %>%
  mutate(predict_te = learned_dom * learned_nondom / VOCAB_SIZE) %>%
  ggplot(aes(x = learned_te, y = predict_te, color = prop_dom)) +
    geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  theme_minimal()

```

### 2021-06-09 02_prepare_data.Rmd
added data_cleaning scripts

---
title: "TE_bilingual_vocabulary_model: Prepare the keepers_data for analysis"
author: "Rachel K.Y. Tsui, Ana Maria Gonzalez-Barrero, Esther Schott, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
```

# Intro
This script implements and documents how we prepared the keepers_data for analysis.

First, we have to read in the keepers_data who holds the bilingual data relevant for the current study.

```{r load_data_raw}
keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE.csv"))
```

# Pivot data from long to wide
```{r}
keepers_ws_TE_final <- keepers_ws_TE %>%
  # remove irrelevant variables
  select(-c(vocab_type_dom, vocab_cross_lang, n_administrations)) %>% 
  # change data from long to wide format by spreading "vocab_type"
  pivot_wider(names_from = vocab_type, values_from = vocab_score)
```

# Define language of dominance 
The dominant language was deemed to be the language in which the child produced a greater number of words (i.e., that language with a greater vocabulary size).

* If the child produced a greater number of words in English than French, then English is the dominant language.
* If the child produced a greater numnber of words in French than English, then French is the dominant language.
```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # define lang_dom_vocab
  mutate(lang_dom_vocab = case_when(total_words_eng > total_words_fre ~ "English",
                                 TRUE ~ "French")) %>%
  # make lang_dom_vocab a factor with 2 levels
  mutate(lang_dom_vocab = as.factor(lang_dom_vocab)) 
```

# Define vocabulary measures based on language dominance
Here, we re-defined vocabulary measures with reference to the language of dominance defined in the previous step. The vocabulary measures include:

* total number of words in the dominant language
* total number of words in the non-dominant language
* total number of singlets in the dominant language
* total number of singlets in the non-dominant language
* total number of singlets (i.e., across both languages)
```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  mutate(
    # define total number of words in the DOMINANT language
    total_words_dom = case_when(total_words_eng > total_words_fre ~ total_words_eng,
                            TRUE ~ total_words_fre),
    # define total number of words in the NON-DOMINANT language
    total_words_nondom = case_when(total_words_eng > total_words_fre ~ total_words_fre,
                               TRUE ~ total_words_eng),
    # define total number of singlets in the DOMINANT language
    total_singlet_dom = case_when(total_words_eng > total_words_fre ~ eng_unique_words,
                                  TRUE ~ fre_unique_words),
    # define total number of singlets in the NON-DOMINANT language
    total_singlet_nondom = case_when(total_words_eng > total_words_fre ~ fre_unique_words,
                                     TRUE ~ eng_unique_words),
    # compute total number of singlets across both languages
    singlet_vocab = total_singlet_dom + total_singlet_nondom)
```

# Define language balance
A balance score is calculated for each child.

* **balance_vocab**: **The main balance measure** we used in the current study is a *Vocabulary Balance*, which is determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages. The formula for calculating the vocabulary balance score is NONDOM/(DOM+NONDOM).
* **balance_input**: Balance can also be considered in terms of input in each language. To make balance_vocab and balance_input comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance (lang_dom_vocab), rather than the language that children heard most and least often. The formula for calculating the input balance score is also NONDOM/(DOM+NONDOM).

```{r}
keepers_ws_TE_final <- keepers_ws_TE_final %>%
  # calculate Vocabulary Balance balance_vocab
  mutate(balance_vocab = total_words_nondom/(total_words_dom+total_words_nondom)) %>%
  # calculate Input Balance balance_input
  mutate(
    # define input dominance (with reference to the lang_dom_vocab defined by vocab size)
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_eng,
                              TRUE ~ lang_exp_fre),
    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                 TRUE ~ lang_exp_eng),
    # calculate the balance score
    balance_input = lang_dom_input/(lang_dom_input+lang_dom_input))
```

# Save the final keepers_data
The final keepers_data is saved in the folder "data_keepers"
```{r}
write.csv(keepers_ws_TE_final, "data_keepers/keepers_ws_TE_final.csv", row.names = F)
```

### 2021-06-11 01_data_wordbank.Rmd
codes to prepare data

---
title: "TE_bilingual_vocabulary_model: Retrieve percentile information from wordbankr"
author: "Rachel K.Y. Tsui, Ana Maria Gonzalez-Barrero, Esther Schott, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
```

# Intro
This script implements and documents how we retrieve the percentile information from wordbankr.

In our study, we define developmental level of a child as the number of potentially learnable words. The number of learnable words is determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages. So we need to retrieve the percentile information from Wordbank (Frank et al., 2016).
```{r load_wordbankr}
library(wordbankr)
```

# Obtain the 90th percentile information from WordbankR
Here, we extract the information about the number of words child can produce at the 90th percentile across ages.

## English CDI-Words and sentences: Vocabulary size at 90th percentile
```{r eng_percentile_info}
english_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("English (American)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(EngWS_90percentile = production)
```

## French CDI-Mots et énoncés: Vocabulary size at 90th percentile
```{r fre_percentile_info}
french_ws_percentile <-  
  fit_vocab_quantiles(get_administration_data("French (Quebecois)", "WS"), 
                      production,
                      quantile = "standard") %>%
  # keep only the 90% percentile
  filter(quantile == 0.90) %>%
  rename(FrWS_90percentile = production)
```

# Merge the two percentile dataframe
Then we merge the two percentile dataframe by age.
```{r merge_percentile}
ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
                       by = "age") %>%
  # compute estimate total vocab by adding the two languages
  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
  # keep only necessary columns
  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
  # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
  rename(age_months_percentile = age)
```

# Save the percentile data for subsequent analysis
The percentile data is saved in the folder "data_keepers"
```{r}
write.csv(ws_percentile, "data_keepers/ws_percentile.csv", row.names = F)
```

### 2021-06-11 02_clean_data.Rmd
codes to prepare data

---
title: "TE_bilingual_vocabulary_model: Clean raw data to keep only bilinguals"
author: "Rachel K.Y. Tsui, Ana Maria Gonzalez-Barrero, Esther Schott, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output: 
  html_document:
    toc: true
    toc_float: true
    number_sections: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(tidyverse)
```

# Intro
This script implements and documents the data cleaning process.

The raw data from Gonzalez-Barrero et al. (2020) comprises archival expressive vocabulary data from English and French monolingual children, as well as English-French bilingual children. Our current study is interested in bilingual vocabulary development, so we are only keeping data from bilingual children.

```{r load_data_raw}
# raw archival data from Gonzalez-Barrero, Schott, & Byers-Heinlein’s (2020)
cdi_ws = readRDS(here::here("data_raw","output_analyzeSent.Rdata"))
```

# Generate keepers_data
We only keep data from bilinguals, who have at least 25% exposure to both English and French, with <10% exposure to a third language. Here, we are removing data from infants who:

* are out of the 18-33month age range
* do not have at least 25% of language exposure to both English and French 
* have an exposure of >10% to a third language
* changed the language group across multiple visits (i.e., those are bilingual at one visit but monolingual at another, regardless of direction)
* do not have both CDIs filled

```{r}
# change which infants changed lang_group & create a "exclude.language_group_change" criterion for subsequent step
exclude.lang_group_change <- cdi_ws %>%
  filter(lang_group !="other") %>%
  group_by(baby_id) %>% 
  distinct(ID_testdate, .keep_all = T) %>%
  mutate(visit_num = 1:n()) %>%  
  mutate(changed_language_group_status = if_else(visit_num == 1, NA, 
                                                 lang_group != lag(lang_group))) %>%
  # uncomment the next line to see all visits for babies that have one visit excluded
  # filter(any(changed_language_group_status == TRUE) ) %>% View()
  filter(changed_language_group_status == TRUE) %>% 
  pull(baby_id)

# create exclusion criteria which identifies all infants who don't fit the required age/language criteria listed above
cdi_ws_TE <- cdi_ws %>%
  mutate(
    # create an exclusion variable for age
    exclude.age = if_else(!between(age_months_binned, 18, 33), 1, 0), # give 1 to those who are not within 18 to 33 months
    # create an exclusion variable for language
    exclude.language = case_when( 
      lang_exp_eng < 25 ~ 1, # give 1 to those who do not have at least 25% exposure to English
      lang_exp_fre < 25 ~ 1, # give 1 to those who do not have at least 25% exposure to French
      lang_exp_other > 10 ~ 1, # give 1 to those with more than 10% exposure to a third language
      baby_id %in% exclude.lang_group_change ~ 1, # give 1 to those who changed language group from bilingual to mono or vice versa
      TRUE ~ 0) # give 0 to all other cases
    )

# generate a keepers data which exclude infants who don't fite the criteria listed
keepers_ws_TE <- cdi_ws_TE %>%
  filter(
    # keep only those with a 0 in exclude.age
    exclude.age == 0,
    # keep only those with a 0 in exclude.language
    exclude.language == 0,
    # keep only those who have both CDIs filled
    both_cdi_filled == "Y")
```

# Check the keepers_data
## How many distinct infants and total number of data points are there?
```{r}
keepers_ws_TE %>% summarize(N_visits = n_distinct(ID_testdate),
                            N_babies = n_distinct(baby_id))
```

## Are there any NAs in the vocabulary measures? (should all return 0)
### Vocabulary score in total vocabulary in English/French, total word vocabulary, and total concept vocabulary
```{r}
keepers_ws_TE %>% filter(is.na(vocab_score)) %>% 
  select(ID_testdate, vocab_type, vocab_score) %>%
  nrow()
```

### Unique words in English
```{r}
keepers_ws_TE %>% filter(is.na(eng_unique_words)) %>% 
  select(ID_testdate) %>% 
  nrow()
```

### Unique words in French
```{r}
keepers_ws_TE %>% filter(is.na(fre_unique_words)) %>% 
  select(ID_testdate) %>%
  nrow()
```

### Number of translation equivalents
```{r}
keepers_ws_TE %>% filter(is.na(number_of_te)) %>% 
  select(ID_testdate) %>%
  nrow()
```

# Save the keepers_data
The keepers_data is saved in the folder "data_keepers"
```{r}
write.csv(keepers_ws_TE, "data_keepers/keepers_ws_TE.csv", row.names = F)
```

### 2021-06-11 03_prepare_data.Rmd
codes to prepare data

--- 2021-06-09 02_prepare_data.Rmd
@@
 * If the child produced a greater numnber of words in French than English, then French is the dominant language.
+
 ```{r}
@@
 * total number of singlets (i.e., across both languages)
+
 ```{r}
@@
                               TRUE ~ lang_exp_fre),
-    lang_dom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
+    lang_nondom_input = case_when(lang_dom_vocab == "English" ~ lang_exp_fre,
                                  TRUE ~ lang_exp_eng),
     # calculate the balance score
-    balance_input = lang_dom_input/(lang_dom_input+lang_dom_input))
+    balance_input = lang_nondom_input/(lang_dom_input+lang_nondom_input))
+```
+
+# Append percentile data to the keepers_data
+Finally, we add the percentile data obtained from wordbankr to the keepers_data. This percentile data indicates the number of potentially learnable words for a child at the relevant age.
+
+```{r}
+# read in the percentile data
+ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv"))
+
+# merge the percentile data to the keepers_ws_TE_final data
+keepers_ws_TE_final <- keepers_ws_TE_final %>%
+  # as the upper age limit of the CDI-WS is 30m, create a new age variable which replaces all >30m as 30m for merging with the percentile dataframe (this means the percentile should be the same for all >30m)
+  mutate(age_months_percentile = ifelse(age_months_binned > 30, 30, age_months_binned)) %>%
+  # merge with the percentile data frame
+  left_join(ws_percentile, by = "age_months_percentile") 
 ```

### 2021-06-11 paper/paper_TE_bilingual_vocabulary_model.Rmd
rmarkdown (not finished yet)

---
title: "Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants"

shorttitle: "Are translation equivalents special for bilingual infants?"

header-includes:
  - \usepackage{amsmath}
  - \renewcommand{\topfraction}{1}
  - \renewcommand{\bottomfraction}{1}
  - \renewcommand{\textfraction}{.1}
  - \renewcommand{\floatpagefraction}{1}
  - \setcounter{topnumber}{9}
  - \setcounter{bottomnumber}{9}
  - \setcounter{totalnumber}{20}
  - \setcounter{dbltopnumber}{9}
  - \DeclareUnicodeCharacter{2265}{\ensuremath{\ge}}
  - \DeclareUnicodeCharacter{2264}{\ensuremath{\le}}

author: 
  - name          : "Rachel Ka-Ying Tsui"
    affiliation   : "1, 2"
    corresponding : yes    
    address       : "Department of Psychology, 7141 Sherbrooke St. West, Montreal, QC, Canada, H2T1V2, or Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
    email         : "rachelkytsui@gmail.com"
  - name          : "Ana Maria Gonzalez-Barrero"
    affiliation   : "1" 
  - name          : "Esther Schott"
    affiliation   : "1" 
  - name          : "Krista Byers-Heinlein"
    affiliation   : "1"

affiliation:
  - id            : "1"
    institution   : "Department of Psychology, Concordia University"
  - id            : "2"
    institution   : "Laboratory for Language Development, RIKEN Center for Brain Science"

abstract: |
  The acquisition of translation equivalents is often considered a special component of bilingual children’s vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages. This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are neither harder nor easier to learn. To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language. Study 2 tested model-derived predictions against vocabulary data from 200 French–English bilingual children aged 18–33 months. Results showed a close match between the model predictions and bilingual children’s patterns of translation equivalent learning. At smaller vocabulary sizes, data matched the Preference Account, while at larger vocabulary sizes they matched the Neutral Account. Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.

keywords          : "bilingualism, infants, translation equivalents, vocabulary development, word learning, computational modeling"

wordcount         :

floatsintext      : yes
figurelist        : no
tablelist         : no
footnotelist      : no
lineno            : yes
mask              : no # "yes" to omit identifying information from the title page

#bibliography      : ["gaze-following_references.bib"]
#replace_ampersands: yes

class             : "man"
output            : papaja::apa6_pdf
---

```{r load_packages_settings, include = FALSE}
library(rmarkdown)
library(knitr)
library(papaja)
library(kableExtra)
library(english)
library(here)
library(tidyverse)
library(tidylog)
library(wordbankr)
library(ggplot2)
library(patchwork)
library(gridExtra)
library("ggpubr")
library(lme4)
library(lmerTest)
library(scales)

`%notin%` <- Negate(`%in%`)

knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)

```

```{r load_data}
# read in the final cleaned keepers_data
keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE_final.csv"))
```

Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings. For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French. These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer, Bornstein, & De Coster, 2006; Pearson et al., 1995). Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics. For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak & Byers-Heinlein, 2019). That is, upon hearing the English word “apple”, the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds. In vocabulary acquisition, bilingual children must learn a first label for a referent ( a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent. Is translation equivalent learning different from singlet learning? The current paper contrasts three competing accounts: 1) translation equivalents are hard to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are neither harder nor easier to learn than singlets (Neutral Account). To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model as a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level. In Study 1, we use the Bilingual Vocabulary Model to derive a set of predictions, which we then test against vocabulary data from 200 18- to 33-month-old bilingual children in Study 2.

## Accounts of translation equivalent learning

### Avoidance Account: Translation equivalents are harder to learn than singlets
Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies (e.g., Imedadze, 1978; Swain & Wesche 1975; Volterra & Taeschner, 1978). The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent. This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman & Wachtel, 1988; Markman, 1992; 1994). For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug”, they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.

Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein & Werker, 2009; Davidson & Tell, 2005; Houston-Price, Caloghiris, & Raviglione, 2010). When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word. However, a strong form of mutual exclusivity might prevent bilinguals from associating a translation equivalent word with its referent, given that in this case the referent is already associated with another word (albeit in the other language). Thus, mutual exclusivity could prevent bilinguals from acquiring translation equivalents, leading to an abundance of singlets in their vocabularies.

Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David & Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017). Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy. For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010). A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al, 2020).

Overall, converging evidence refutes the position that a strong form of mutual exclusivity operates on the developing bilingual lexicon to prevent bilinguals from acquiring translation equivalents. Nonetheless, it leaves open the possibility that translation equivalents may be less likely acquired in favour of learning singlets even if translation equivalents are not completely avoided. If bilingual children avoid lexical overlap across languages even to a small degree, then under the Avoidance Account translation equivalents would be harder to learn than singlets.

### Preference Account: Translation equivalents are easier to learn than singlets
Contrary to the Avoidance Account, the Preference Account posits that translation equivalents are easier to learn than singlets. At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two. When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017). Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak & Byers-Heinlein, 2018; Singh, 2013). In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020). The Preference Account predicts that translation equivalents would be more easily learned than singlets.

There are several lines of empirical evidence to support the Preference Account. For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages (Lanvers, 1999; Pearson & Fernández, 1994). The main explanation that has been given for this finding is that additional exposure to their non-dominant language — which became their new dominant language — enabled fast mapping of words to already-lexicalized concepts.

Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015). The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like “cat” and “dog” are strongly semantically related). Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in the lexicons. This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English “dog” and French _“chien”_) and words that had other cross-language relations (e.g., “cat” and _“chien”_). The authors then simulated bilingual vocabulaires by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals. Comparison with actual bilingual children’s vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation. The authors therefore concluded that bilingual children show a preference for learning translation equivalents over singlets. Note that, in their study, translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word “cat”). However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning including vocabulary size in each language and the developmental level of a child — an issue that we will return to later in the introduction.

Overall, there is some evidence that bilingual children more readily learn translation equivalents than singlets. If the strong semantic overlap between translation equivalents facilitates their learning, then under the Preference Account translation equivalents would be more easily learned than singlets.

### Neutral Account: Translation equivalents are similar to learn than singlets
The previous accounts rely on the idea that bilingual vocabulary development unfolds differently than monolingual development, as monolinguals encounter only singlets but bilinguals encounter both singlets and translation equivalents. There is an underlying assumption that translation equivalent learning is somehow special relative to singlet learning — the Avoidance Account proposes that translation equivalents are harder to learn than singlets, whereas the Preference Account proposes that translation equivalents are easier to learn than singlets. However, it is also possible that translation equivalents are neither harder nor easier to learn than singlets. We call this the Neutral Account.

The Neutral Account implies that bilingual children’s two languages develop relatively independently. Indeed, language and processing measures in bilinguals tend to be tightly correlated within a particular language, with only weak cross-language correlations. For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, & Hurtado, 2010). Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013). Bilingual children who heard a greater amount of a particular language tended to process that language faster, and in turn learned more words in that language.

In a study whose results directly support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children. It was revealed that the percentage of translation equivalents observed in English–Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children. This evidence suggested that the acquisition of translation equivalents is not that different from the expected overlap between two independent lexicons, implying that singlets and translation equivalents are equally learnable. In sum, the Neutral Account predicts that translation equivalents are neither harder nor easier for bilingual children to learn than singlets.

## Contributors to translation equivalent knowledge
The previous section discussed three theoretical accounts concerning the relative learnability of translation equivalents. However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than as an integrated system. In this section, we consider two proximal variables that we expect to predict the number of translation equivalents children know: vocabulary size in each language, and word learnability as a function of children’s developmental level.

### Vocabulary size in bilinguals’ two languages
Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know. For example, a child with a less balanced vocabulary across the two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language. Conversely, it seems reasonable to expect that if a child knows a similar number of words in each language and thus has a more balanced vocabulary across the two languages, there would be more potential for some of those words to be translation equivalents.

Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which appears to be tightly linked to their exposure to each language. In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, & Hurtado, 2010; Place & Hoff, 2011; Pearson et al., 1997). Bilingual children tend to know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011). This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.

One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts. This is known as the Complementarity Principle (Grosjean, 2016). For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010). If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words. However, the Complementarity Principle may not necessarily impact bilingual word acquisition in infancy. Most words in children’s early vocabularies could be considered “home words”, which include words for social and daily routines (“hello”, “more”, “diaper”), common nouns (“doggie”), and everyday verbs (“walk”). Such words are likely to be encountered across contexts where children spend the majority of their time, such as at home and at childcare. Thus, growing up in a bilingual context from birth, bilingual children presumably encounter “home words” in both languages. Accordingly, we assume that for the most part, bilingual children’s opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle. We further consider implications of this assumption in the discussion section.

Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, & Weisleder, 2013). Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).

### Word learnability as a function of developmental level
An often overlooked factor that could contribute to bilingual children’s learning of translation equivalents is related to the changes in the learnability of different words over time based on children’s developmental level. Evidence from monolingual children shows that some types of words are characteristically learned before others. For example, across many languages, including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996). Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007). This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky et al., 2019). If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development. That is, potential singlets might be “too hard” to be learned at a particular age. Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.

## The Bilingual Vocabulary Model
Taking into account the contributions of language exposure and developmental level to bilingual children’s vocabulary acquisition, we put forward the Bilingual Vocabulary Model. This model proposes that the number of translation equivalents that children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children’s developmental level. We formalize learning a translation equivalent pair  as the joint probability of learning each of the words in the pair. This provides a clear empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent. The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B | A) where P(B | A) = P(B) if A and B are independent (see Box 1 for a detailed example). The full model is shown in Figure 1. In the next paragraphs, we define each of the model parameters in detail, and these are also summarized in Table 1.

&nbsp;

```{r box1, results="asis", fig.cap = " "}
box1 <- data.frame(title = "To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed. To determine independence, we must know the probability of being a psychology major, the probability of being left handed, and the probability of being both a psychology major and left-handed. The numerator of these probabilities will be the number of students who are left-handed, psychology majors, and both (respectively) and the denominator will be the total population of students we are observing. Imagine a college of 1000 students. If 100/1000 (or 1/10) students are left-handed, and 200/1000 (or 1/5) students are psychology majors, then if these variables are independent we expect a proportion of 1/10 × 1/5 = 1/50 students to be left-handed psychology majors. To determine the number we expect to observe in the college, we multiple 1/50 × 1000 = 20 students. When we compare this expected number to the actual number of left-handed students, there are three possible outcomes. First, we may observe many more than 20 left-handed psychology students at the college (say 100 students), which suggests that being left-handed increases the probability of majoring in psychology. Or, second, we may observe many fewer than 20 left-handed psychology students at the college (say 5 students), this suggests that being left-handed decreases the probability of majoring in psychology (in this example by a factor of 1/4). Finally, if left-handedness and majoring in psychology are independent, we can predict the number of left-handed psychology students by multiplying the observed number of left-handed students (100) by the observed number of psychology students (200), and dividing by the total population of the college (1000), so for example 100×200/1000 = 20. Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.") %>%
  rename("Box 1. Example of test for independence." = title)

kable(box1,
      format="latex", booktabs=TRUE, longtable = TRUE) %>%
  column_spec(1, width = "450px") %>%
  kable_styling(font_size = 10)

```

```{r fig1, fig.cap="Illustration of the Bilingual Vocabulary Model.", fig.align='center', out.width='100%'}

knitr::include_graphics("./Figure1_illustration.jpg")
```

The model takes four main parameters: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS > 1) or against (BIAS < 1) learning translation equivalents. The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language. 
Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words). If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI. For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation). Very young children would not be expected to know many of the “harder” words on the CDI, such as “lawn mower”, “sidewalk”, or “vitamins'', due to children’s immature cognitive machinery and conceptual development. A more reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age. For example, Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words. Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016). Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable — if imperfect — estimate of the number of learnable words (LEARNABLE) that a child of Jamie’s age could potentially acquire in each language. Thus, we might expect that Jamie could potentially have learned up to 245 English words and 245 French words, although he has thus far only learned 50 in English and 20 in French.

Using the mathematical concept of independence, we can then quantify the number of translation equivalents (TE) expected given children’s vocabulary sizes in the dominant (DOM) and non-dominant (NONDOM) languages, as well as the number of potentially learnable words (LEARNABLE). If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) — which is the possible number of words that could overlap across both languages — to predict the number of translation equivalents. 
We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral account. Adding this parameter, translation equivalents can be derived from TE = BIAS × (DOM × NONDOM)/LEARNABLE. For the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets; for the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets; for the the Neutral Account, BIAS is exactly 1 (i.e., the model is unbiased with respect to whether translation equivalents are more difficult or easier to acquire than singlets). Going back to the example of 18-month-old Jamie, we would set the denominator at 245 which is the number of potentially learnable words at 18 months. If translation equivalents are half as easy to learn as singlets (following the Avoidance Account), we would expect Jamie to produce .5×(50×20/245) = 2.0 translation equivalents. Conversely, if translation equivalents are twice as easy to learn as singlets (following the Preference Account), we would expect Jamie to produce 2×(50×20/245) = 8.2 translation equivalents. Under the Neutral Account, we would expect Jamie to learn 1×(50×20/245) = 4.1 translation equivalents.

Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.

Balance of vocabulary (BALANCE) is the proportion of total words that children produce in each language. For convenience, balance is defined in reference to the non-dominant language with the formula NONDOM/(DOM+NONDOM), such that scores can range from 0.0 (completely unbalanced) to 0.5 (completely balanced). For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29. Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many words he produces are translation equivalents).

Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM). 
Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child — that is, the total number of concepts that are lexicalized in either language. This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD). Finally, we can also calculate singlets that are produced in each language, that is words for which the child does not yet produce a translation equivalent. Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM). It is also possible to decompose children’s word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.

```{r table1, results = "asis"}

table1 <- data.frame(
  Variable = c("LEARNABLE", "DOM", "NONDOM", "BIAS", "BALANCE", "WORD", "TE", "CONCEPT", "DOM-SINGLET", "NONDOM-SINGLET"),
  Definition = c("Number of learnable words in each language, given the child’s developmental level",
                   "Words produced in the dominant language",
                   "Words produced in the non-dominant language",
                   "Bias parameter",
                   "Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages)",
                   "Word vocabulary (or total vocabulary size)",
                   "Translation equivalents produced",
                   "Concept vocabulary (or total conceptual vocabulary size)",
                   "Singlets in dominant language",
                   "Singlets in non-dominant language"),
  Constraints = c("Varies by age. No greater than the number of words on CDI.",
                  "DOM \u2265 NONDOM (i.e., always produce more words in dominant than non-dominant language); DOM \u2264 LEARNABLE",
                  "NONDOM \u2264 DOM (always produce fewer words in non-dominant than dominant language); NONDOM \u2264 LEARNABLE",
                  "BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account",
                  ".00 \u2264 BALANCE \u2264 .50 (greater values indicate children producing a more similar number of words in their two languages)",
                  "W \u2264 2LEARNABLE (maximum word vocabulary is knowing each word in both languages)",
                  "", "", "", ""),
  "Relationship to other parameters" = c("Maximum number that could be learned in DOM or NONDOM",
                                         "DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM",
                                         "NONDOM = BALANCE×WORD; NONDOM = WORD - DOM",
                                         "Account = BIAS×(DOM×NONDOM/LEARNABLE)",
                                         "BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)",
                                         "WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)",
                                         "If DOM and NONDOM are independent, TE = DOM×NONDOM/LEARNABLE",
                                         "CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET",
                                         "DOM-SINGLET = DOM - TE",
                                         "NONDOM_SINGLET = NONCOM - TE")
)

landscape(
  kable(table1,
        format = "latex", booktabs = TRUE,
        caption = "Summary of the parameters in the Bilingual Vocabulary Model.") %>%
    kableExtra::kable_styling(font_size = 7) %>%
    pack_rows("Main Parameters", 1, 4) %>%
    pack_rows("Derived Parameters", 5, 10) %>%
    column_spec(2:4, width = "175px")
)
```

## Current research
The current research aimed to better understand the nature of translation equivalent learning in bilingual children. Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels. We also compared predicted learning outcomes for when translation equivalents are harder to learn, or easier to learn, or similar to learn than singlets.

In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997). Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).

# Study 1: Simulations
Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction (see also Figure 1), which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning. Note that usually only three values are necessary to calculate all the other variables (see Table 1). Most commonly, we can calculate other variables based on the total number of learnable words (LEARNABLE) together with either the words known in each language (DOM and NONDOM) or word vocabulary plus balance (WORD and BALANCE) which allow us to compute DOM and NONDOM. It is also possible to calculate other variables based on the total number of learnable words (LEARNABLE) with balance and words known in either language (BALANCE and DOM or BALANCE and NONDOM).

Three simulations were generated to explore expected patterns of translation equivalent learning under the Bilingual Vocabulary Model. In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NON-DOM), and word vocabulary (WORD). In the second simulation, we explored relationships between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE). In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the bias parameter (BIAS) to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts. A summary of the parameter values used in each simulation is provided in Table 2.

```{r table2, results = "asis"}

table2 <- data.frame(
  Simulation = c("1", "2", "3"),
  LEARNABLE = c("Constant at 600", "Varied at 300, 450, and 600", "Varied at 150, 300, 450, and 600"),
  DOM = c("Varied, ranging from 100 to LEARNABLE at an interval of 100", 
          "Varied, ranging from 100 to LEARNABLE at an interval of 100",
          "Varied, ranging from 100 to LEARNABLE at an interval of 100 "),
  NONDOM = c("Varied, ranging from 0 to DOM at an interval of 10",
             "Varied, ranging from 0 to DOM at an interval of 25",
             "Varied, ranging from 0 to DOM at an interval of 25"),
  WORD = c("Calculated as WORD = DOM + NONDOM",
           "Calculated as WORD = DOM + NONDOM",
           "Calculated as WORD = DOM + NONDOM"),
  BALANCE = c("Calculated as BALANCE = NONDOM / (DOM+NONDOM)",
              "Calculated as BALANCE = NONDOM / (DOM+NONDOM)",
              "Calculated as BALANCE = NONDOM / (DOM+NONDOM)"),
  BIAS = c("Constant at 1", "Constant at 1", 
           "Varied at 0.5 (Avoidance Account), \n 1 (Neutral Account), \n and 1.5 (Preference Account)"),
  n_data = c(216, 161, 166)
)

landscape(
  kable(table2,
        col.names = c("Simulation",
                      "Learnable words (LEARNABLE)",
                      "Words in dominant Language (DOM)",
                      "Words in non-dominant language (NONDOM)",
                      "Word vocabulary (WORD)",
                      "Balance of vocabulary (BALANCE)",
                      "Bias parameter (BIAS)",
                      "Total number of data points generated"),
        align = "clllllll",
        format="latex", booktabs = TRUE,
        caption = "Summary of the parameters used in each simulation.") %>%
    column_spec(1:2, width = "45px") %>%
    column_spec(3:7, width = "75px") %>%
    column_spec(8, width = "45px") %>%
    kable_styling(font_size = 8)
)

```

## Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
In Simulation 1, we first illustrate the mathematical relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE. For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016). We set BIAS to 1, meaning that in these examples translation equivalents are similarly easy to learn as singlets.

We first illustrate with three hypothetical children. Infant Annie (small vocabulary, unbalanced exposure) produces 270 words in the dominant language and 30 words in her non-dominant language. She has a word vocabulary of 300, and a balance score of .10 (10% of her words are in the non-dominant language). Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents. Infant Bernie (small vocabulary, balanced exposure) produces 180 dominant-language words, and 120 non-dominant language words. Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40% of his words are in the non-dominant language). Based on our formula, we expect Bernie to produce 36 translation equivalents. Comparing Annie and Bernie, two children who produce the same word vocabulary (i.e., WORD is held constant), the child with more balanced language vocabulary (Bernie) is expected to produce more translation equivalents. Like Bernie, infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM) for a total of 900 words (WORD), and thus BALANCE = .40. Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE). Infants Bernie and Charlie illustrate that for two children equal in BALANCE, the child with larger word vocabulary (WORD) is expected to produce more translation equivalents (TE). Other vocabulary metrics are calculated for each hypothetical child as described in Table 3.

```{r table3, results="asis"}

# create columns that contain information to be added to table3 in the next step
## Definition column: the definition of each variable in the table
Definition <- c("Bias parameter", "Learnable words", "Words produced in the dominant language", "Words produced in the non-dominant language",
                "Word vocabulary (or total vocabulary size)", "Vocabulary balance", "Translation equivalents produced", 
                "Concept vocabulary (or total conceptual vocabulary size)", "Singlets in dominant language", "Singlets in non-dominant language")
## Calculation column: the formula used to calculate the derived parameters
Calculation <- c("", "", "", "", 
                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM * NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")

table3 <- data.frame(
    # define the main parameters
    hypotheical_child = c("Infant Anne (small vocabulary, unbalanced)",
                        "Infant Bernie (small vocabulary, balanced)",
                        "Infant Charlie (large vocabulary, balanced)"),
    BIAS = 1, 
    LEARNABLE = 600, 
    DOM = c(270, 180, 540),
    NONDOM = c(30, 120, 360)
  ) %>%
  # calculate the derived parameters from the main parameters
  mutate(WORD = DOM + NONDOM,
         BALANCE = NONDOM/(DOM+NONDOM),
         TE = DOM*NONDOM/LEARNABLE,
         CONCEPT = WORD - TE,
         "DOM-SINGLET" = DOM - TE, 
         "NONDOM-SINGLET" = NONDOM - TE) %>%
  # transpose the data set for presentation
  pivot_longer(-("hypotheical_child"), names_to = "Variable", values_to = "value") %>%
  pivot_wider(names_from = "hypotheical_child", values_from = "value") %>%
  # append other information for presentation
  mutate(Definition = Definition, 
         Calculation = Calculation) %>%
  # reorder columns
  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)")

landscape(
  kable(table3,
      align = "llllll",
      format="latex", booktabs = TRUE,
      caption = "Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.") %>%
  pack_rows("Main Parameters", 1, 4) %>%
  pack_rows("Derived Parameters", 5, 10) %>%
  column_spec(c(1), width = "100px") %>%
  column_spec(c(2:3), width = "140px") %>%
  column_spec(c(4:6), width = "60px") %>%
  kable_styling(font_size = 8) 
)

```

We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied. Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation). Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)). In Figure 2, we plotted TE knowledge as predicted by DOM, NONDOM, and WORD at different levels of BALANCE. Across all three Panels (1A, 1B, and 1C), children with the most balanced vocabulary consistently produce more translation equivalents than other children. Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increase, the number of TE also increases regardless of BALANCE. Interestingly, Panel 1B shows that NONDOM and TE are extremely tightly coupled. In sum, we observe three important patterns, which will serve as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:

* Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE). 
* Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE). 
* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge(see panel 1B of Figure 2).

```{r simulation1, echo = FALSE}
# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (LEARNABLE = 600)

## generate the simulation data
simulation1 <- function (LEARNABLE) {
  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
  NONDOM_seq <- seq(0, LEARNABLE, by = 10) # sequence of N from 0 to V at an interval of 10
  DOM <- rep (DOM_seq, length(NONDOM_seq))
  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
  data_simulation1 <<- data.frame(LEARNABLE, DOM, NONDOM)
}

simulation1(600) # setting V = 600

## clean the simulation data
data_simulation1 <- data_simulation1 %>%
  arrange(DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>%
  # Get word vocabulary W
  mutate(WORD = DOM + NONDOM) %>%
  # Create other variables
  mutate(BALANCE = NONDOM/(DOM+NONDOM), 
         TE = (DOM*NONDOM)/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  # Round to keep whole numbers only
  mutate_at(vars(-BALANCE), round, 0) %>%
  # bin B into 5 groups
  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(DOM, NONDOM, bin_BALANCE)

## visualizing Simulation 1
simulation1 <- data_simulation1 %>%
  # convert to long data
  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             DOM = "Panel 1A: \n Dominant vocabulary \n (DOM)", 
                             NONDOM = "Panel 1B: \n Non-dominant vocabulary \n (NONDOM)",
                             WORD = "Panel 1C: \n Total vocabulary \n (WORD)")) %>%
  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=12),
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "1) Simulated data") +
  guides(color = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1)) 
```

```{r simulation1-plus_observed_data, echo = FALSE}
Observed_plot_inputs_TE <- keepers_ws_TE %>% 
  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, total_words_dom, total_words_nondom, balance_vocab) %>%
  pivot_longer(c(word_vocab, total_words_dom, total_words_nondom), names_to = "vocab_type", values_to = "number") %>%
  mutate(vocab_type = factor(vocab_type, levels=c("total_words_dom", "total_words_nondom", "word_vocab"))) %>%
  mutate(vocab_type = recode(vocab_type, 
                             total_words_dom = "Panel 2A: \n Dominant vocabulary \n (DOM)", 
                             total_words_nondom = "Panel 2B:\n Non-dominant vocabulary \n (NONDOM)",
                             word_vocab = "Panel 2C: \n Total vocabulary \n (WORD)")) %>%
  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
  #geom_point() +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ vocab_type) +
  theme(axis.text.x = element_text(angle = 90),
        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=12),
        legend.position = "bottom",
        panel.spacing.x = unit(2, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  ylim(0, 600) +
  xlim(0, 1000) +
  labs(x = "Number of words", 
       y = "Number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)",
       title = "2) Observed data") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))

# combine Simulation1 visualization + observed data visualization
plot_Simulation1plusData <- ggarrange(simulation1, 
                                      Observed_plot_inputs_TE,
                                      nrow = 2,
                                      common.legend = TRUE, legend = "bottom")
```

```{r fig2, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to total dominant vocabulary size (DOM; Panel A), total non-dominant vocabulary size (NONDOM; Panel B), and total word vocabulary (WORD; Panel C). Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600. Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%"}

plot_Simulation1plusData
```

## Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn the word for 600 different concepts (i.e., LEARNABLE held constant at 600). As laid out in the introduction, in the Bilingual Vocabulary Model the learnability of different words changes over time, where the possible number of words that can be learned increases as a child grows older. Therefore, in Simulation 2, we looked at the expected patterns of translation equivalent learning across varying levels of LEARNABLE (i.e., the number of learnable words as developmental level changes). Additionally, we further examined vocabulary composition by computing the number of singlets in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) language. BIAS was once again kept constant at 1.

Translation equivalent knowledge was simulated across children at three developmental levels (the number of LEARNABLE words = 300, 450, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM). In total, data from 161 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation). Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM with the formula NONDOM/(DOM+NONDOM). We also counted the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children’s concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET. Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), medium balanced (BALANCE = .20 - .35), and least balanced (BALANCE = .00 - .02).

In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT) as shown by the largest bars for the children at LEARNABLE = 600 compared to the shortest bars for children at LEARNABLE = 300. Moreover, we continue to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE), as evidenced by the green bars that are the highest for the most balanced children (BALANCE = .36 - .50) compared to the medium balanced (BALANCE = .21 - .35) and the least balanced children (BALANCE = .00 - .20). Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents. Across all three developmental levels of LEARNABLE, the light blue bars representing singlets in the dominant language (DOM-SINGLET) grow with less balanced exposure, whereas the dark blue bars representing singlets in the non-dominant language (NONDOM-SINGLET) grow with more balanced exposure. Overall, we observe 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model. Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will

* Prediction 2a: Have larger concept vocabularies (CONCEPT) 
* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with a less balanced vocabulary (BALANCE) produce the most DOM-SINGLET.
* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with a more balanced vocabulary (BALANCE) produce the most NONDOM-SINGLET.

```{r simulation2, echo = FALSE}
# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (LEARNABLE = 300, 450, or 600)

## generate the simulation data
LEARNABLE <- seq(300, 600, by = 150) # set LEARNABLE to 300, 450, and 600

simulation2 <- function (LEARNABLE) {
  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
  NONDOM_seq <- seq(0, LEARNABLE, by = 25) # sequence of N from 0 to V at an interval of 25
  DOM <- rep(DOM_seq, length(NONDOM_seq))
  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
  data_simulation2 <<- data.frame(LEARNABLE, DOM, NONDOM)
}

data_simulation2 <- do.call(rbind, lapply(LEARNABLE, simulation2)) %>%
  arrange(LEARNABLE, DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>%
  # Create other variables
  mutate(WORD = DOM + NONDOM, # word vocabulary 
         BALANCE = NONDOM/(DOM+NONDOM), # balance,
         TE = (DOM*NONDOM)/LEARNABLE, # If D and N are independent, TE = DN/V
         DOM_SINGLET = DOM-TE, # Unique words in dominant language
         NONDOM_SINGLET = NONDOM-TE) %>% # Unique words in non-dominant language
  # Round to keep whole numbers only
  mutate_at(vars(-BALANCE), round, 0) %>%
  # bin B into 3 groups
  mutate(bin_BALANCE = case_when(BALANCE >= 0.36 ~ "Most Balanced",
                                 BALANCE <= 0.20 ~ "Least Balanced",
                                 TRUE ~ "Medium Balanced")) %>%
  mutate(bin_BALANCE  = fct_rev(bin_BALANCE)) %>%
  arrange(DOM, NONDOM, bin_BALANCE)

## visualizing Simulation 2
simulation2 <- data_simulation2 %>%
  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
  filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
  summarise(n_words = mean(number)) %>%
  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_label) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=10),
        panel.spacing.x = unit(2, "lines")) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)", 
       title = "Panel A: Simulated data")
```

```{r simulation2-plus_observed_data, echo = FALSE}
plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
                               balance_vocab <= 0.20 ~ "Least Balanced",
                               TRUE ~ "Medium Balanced")) %>%
  mutate(BalSubset = fct_rev(BalSubset)) %>%
  arrange(desc(number_of_te)) %>%
  mutate(TE = number_of_te) %>%
  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
               names_to = "Type", values_to = "NumberOfWords") %>%
  mutate(Type = factor(Type, levels=c("total_singlet_nondom", "total_singlet_dom", "number_of_te"))) %>%
  mutate(Type = recode(Type, 
                       total_singlet_nondom = "Singlet in non-dominant language (NONDOM-SINGLET)", 
                       total_singlet_dom = "Singlet in dominant language (DOM-SINGLET)", 
                       number_of_te = "Translation equivalent (TE)")) %>%
  group_by(LEARNABLE_subset, BalSubset, Type) %>%
  summarise(n_words = mean(NumberOfWords)) %>%
  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
  geom_bar(position="stack", stat="identity") +
  facet_grid(. ~ LEARNABLE_subset) +
  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
  ylim(0, 600) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 0) , 
        legend.position = "bottom",
        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
        plot.title.position = "plot",
        text = element_text(size=10),
        panel.spacing.x = unit(2, "lines")) + 
  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
  labs(x = "\nVocabulary balance (BALANCE)", 
       y = "Number of concepts (CONCEPT)",
       title = "Panel B: Observed data")

# combine Simulation2 visualization + observed data visualization
plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
                                      plot_stacked_byLEARNABLE,
                                      nrow = 2,
                                      heights=c(0.8,1))
```

```{r fig3, fig.cap="Number of translation equivalents (TE) and singlets in dominant (DOM-SINGLET) and non-dominant language (NONDOM-SINGLET) across different developmental levels/ages, which sets the number of LEARNABLE words. Panel A represents the model simulation in Study 1, where developmental levels of simulated children are set at three values: LEARNABLE = 300, 450, and 600. Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right). Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="80%"}

plot_Simulation2plusData
```

## Simulation 3: Bias towards or against translation equivalent learning compared to singlets
In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE. In our final simulation, we examined models where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account. Mathematically, this requires varying the BIAS parameter. For the Preference Account, BIAS will be greater than 1, meaning that TEs are more easily learned than singlets. On the other hand, for the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets.

Translation equivalent knowledge was first simulated across different developmental levels (as indicated by number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM. Again, BALANCE and word vocabulary (WORD) were calculated based on the values of DOM and NONDOM. The final simulated data set contained 166 data points (see Table 2 for a summary of the parameters used). Three scenarios of translation equivalent learning (TE) were then generated using the formula TE = BIAS × DOM×NONDOM/LEARNABLE. To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50% less likely to be learned than singlets). To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are neither easier nor harder to learn than singlets). Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets). In Figure 4, we illustrate the three different scenarios of simulated translation equivalent knowledge. Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies produced more translation equivalents. Thus, overall relationships between vocabulary balance and translation equivalent knowledge remained similar across the Avoidance, Preference, and Neutral Accounts. What changed is the slope of translation equivalent learning: the slopes were the shallowest under the Avoidance Account where BIAS = 0.5, whereas the slopes are steepest under the Preference Account where BIAS = 1.5. With this, we further outline Prediction Set 3:

* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning, with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).

```{r simulation3, echo = FALSE}
# Simulation 3: Non-independence of dominant-language (DOM) and non-dominant language words (NONDOM)

## generate the simulation data
LEARNABLE <- seq(150, 600, by = 150) # set LEARNABLE to 300, 450, and 600

simulation3 <- function (LEARNABLE) {
  DOM_seq = seq(100, LEARNABLE, by = 100) # sequence of DOM from 100 to LEARNABLE at an interval of 100
  NONDOM_seq = seq(0, LEARNABLE, by = 25) # sequence of NONDOM from 0 to LEARNABLE at an interval of 25
  DOM <- rep(DOM_seq, length(NONDOM_seq))
  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
  data_simulation3 <<- data.frame(LEARNABLE, DOM, NONDOM)
}

data_simulation3 <- do.call(rbind, lapply(LEARNABLE, simulation3)) %>%
  arrange(LEARNABLE, DOM, NONDOM) %>%
  filter(NONDOM <= DOM) %>% # eliminate cases when NONDOM > DOM
  mutate(WORD = DOM+NONDOM, # word vocabulary
         BALANCE = NONDOM/(DOM+NONDOM), # balance 
         TE = DOM*NONDOM/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
  mutate_at(vars(-BALANCE), round, 0) %>%
  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
  arrange(bin_BALANCE)

## Simulation3: TE = BIAS * (DOM*NONDOM/LEARNABLE)
### defining the bias parameter I
BIAS1 = 1.5 # BIAS > 1, TEs are MORE easily learned than unique words
BIAS2 = 0.5 # BIAS < 1, TEs are LESS easily learned than unique words

### visualizing the simulation
plot_simulation3 <- data_simulation3 %>%
  mutate(TE_easier = BIAS1*(DOM*NONDOM/LEARNABLE), # calculate where BIAS > 1
         TE_harder = BIAS2*(DOM*NONDOM/LEARNABLE)) %>% # calculate where BIAS < 1
  # round to keep whole numbers only
  mutate_at(vars(-c(BALANCE, bin_BALANCE)), round, 0) %>%
  # prepare dataset to plot
  pivot_longer(c(TE, TE_harder, TE_easier), 
               names_to = "Model", values_to = "number_TE") %>%
  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
  mutate(Model = recode(Model, 
                        TE = "BIAS = 1 \n (Neutral Account)",
                        TE_easier = "BIAS = 1.5 \n (Preference Account)",
                        TE_harder = "BIAS = 0.5 \n (Avoidance Account)")) %>%
  # ggplot
  ggplot(aes(x = WORD, y = number_TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
  stat_smooth(method = lm, se = F) +
  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
  theme_minimal() + 
  facet_grid(. ~ Model) +
  theme(axis.text.x = element_text(angle = 90),
        legend.position = "bottom") +
  theme(plot.title = element_text(hjust = 0.5),
        text = element_text(size=11),
        panel.spacing.x = unit(2.5, "lines"),
        panel.background = element_rect(fill = NA, color = "black")) +
  labs(x = "Word vocabulary (WORD)", 
       y = "Number of translation equivalents (TE)", 
       color = "Vocabulary balance (BALANCE)",
       linetype = "Vocabulary balance (BALANCE)") +
  guides(colour = guide_legend(reverse = T, nrow = 1),
         linetype = guide_legend(reverse = T, nrow = 1))
```

```{r fig4, fig.cap="Different scenarios of expected translation equivalents learning (TE) while assuming bias towards (BIAS < 1) or against learning translation equivalents (BIAS > 1), or no bias (BIAS = 1).", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%"}

plot_simulation3
```

# Study 2: Empirical data
In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables. In Study 2, we tested these predictions using archival data obtained from 200 French–English bilingual children aged 18 to 33 months.

## Method
Ethics approval was obtained by the Human Research Ethics Board of XYZ University (ethics protocol number 0000) and informed consent was obtained from the participants’ parents.

### Participants
```{r participants, include = FALSE}
# how many distinct participants and data point we have
keepers_ws_TE %>% 
  summarize(N_babies = n_distinct(baby_id),
            N_admin = n_distinct(ID_testdate))

# how many girls and boys we have
d_gender <- keepers_ws_TE %>%
  group_by(gender) %>% 
  distinct(baby_id, .keep.all=T) 
  # %>% count()

# number of participants who contributed multiple data points
d_multiple <- keepers_ws_TE %>% 
  count(baby_id) %>% 
  filter(n>1)

# checking which baby_id has multiple data
n_duplicatedBabyID <- keepers_ws_TE %>%
  filter(duplicated(baby_id)) %>%
  arrange(baby_id)

## percentage of english-dominant & french-dominant children
# dominance defined by vocab size
d_lang_dom <- keepers_ws_TE %>% 
  distinct(ID_testdate, .keep_all = TRUE) %>%
  mutate(Eng_Dom = ifelse(lang_dom_vocab == "English", 1, 0)) %>%
  summarise(n = n(),
            n_EngDom = sum(Eng_Dom),
            percent_EngDom = n_EngDom/n * 100,
            percent_FrDom = 100 - percent_EngDom)
```
Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al. (2020). Some children took part in more than one in-lab study (n = `r length(d_multiple$baby_id)`); thus, they contributed data at more than one time point. This resulted in a larger number of datapoints relative to the number of unique participants. The total number of data points included in the analyses was `r length(keepers_ws_TE$ID_testdate)` (i.e., `r length(keepers_ws_TE$eng_cdi_filled[keepers_ws_TE$eng_cdi_filled == "Y"])` English and `r length(keepers_ws_TE$fre_cdi_filled[keepers_ws_TE$fre_cdi_filled == "Y"])` French CDI questionnaires). Participants were recruited through government birth lists, online ads, daycares, and infant-parent group activities (e.g., children’s library activities). Inclusion criteria were the following: full term-pregnancy ( i.e., > 36 weeks of gestation), normal birth weight (> 2500 grams), and absence of major medical conditions (i.e., meningitis). Only children who had complete data in both CDI forms (i.e., English and French) were retained for analysis. Bilingual children were defined as those exposed at least 25% of the time over the course of their lives globally to both English and French and with less than 10% of exposure to a third language. For children who participated more than once, their language exposure followed such criteria between visits. Following the approach in Study 1, children’s dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM). Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 1)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant. Data collection was conducted in Montréal, Québec, Canada. Montréal is a multicultural city where both English and French are widely used in society. Children’s demographic characteristics are presented in Table 4.

```{r table4, results="asis", echo = FALSE}
# Create dataframe of demographics
d_participants <- keepers_ws_TE %>%
  # keep only relevant variables
  select(ID_testdate, age_continuous, years_education, lang_exp_eng, lang_exp_fre, lang_exp_other) %>%
  # rename variables
  rename("Age in months" = age_continuous,
         "Maternal education in years" = years_education,
         "% Global exposure to English"  = lang_exp_eng,
         "% Global exposure to French"  = lang_exp_fre, 
         "% Global exposure to Other" = lang_exp_other) %>%
  # summarize the data
  pivot_longer(-c(ID_testdate), names_to = "info", values_to = "value") %>%
  group_by(info) %>%
  summarise(mean = mean(value, na.rm=T),
            sd = sd(value, na.rm=T),
            min = min(value, na.rm=T),
            max = max(value, na.rm=T)) %>%
  mutate(across(where(is.numeric),round, 1)) %>%
  # combine min and max to form range
  unite("range", min:max, sep = " - ", remove = TRUE, na.rm = FALSE) %>%
  # arrange rows in custom order
  mutate(info = factor(info, levels = c("Age in months", "Maternal education in years", 
                                        "% Global exposure to English", "% Global exposure to French",
                                        "% Global exposure to Other"))) %>%
  arrange(info)

# generate the apa table
papaja::apa_table(d_participants, format.args = list(digits = 1),
                  caption = "Demographic characteristics of participants (data points = 229).",
                  col.names =c("","Mean","$SD$","Range"),
                  align=c("l","c","c","c"))
```

### Measures

```{r cdi_filled_by, include = FALSE}
# who filled out the CDIs
d_cdi_filled_by <- keepers_ws_TE %>% 
  # combine grandmother & other family member
  mutate(cdi_filled_by = replace(cdi_filled_by, cdi_filled_by == "Grandmother", "Other family member")) %>% 
  # count number of respondents
  count(cdi_filled_by) %>%
  mutate(percentage = n/sum(n)*100) # calculate percentages
```

* **MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI).** Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI. Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997). We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64%), fathers (7%), both parents (4%), others (< 1%; e.g., grandmother), and respondent not indicated (25%). In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms. Our analyses focused on the vocabulary checklist of this questionnaire, which includes different nouns, verbs, adjectives, and other words used by young children. There are 680 words in the English CDI version and 664 in the Québec French version.

* **Language Exposure Questionnaire using the MAPLE approach.**

### Procedure

## Results
### Testing Prediction Set 1: 
### Testing Prediction Set 2:
### Testing Prediction Set 3:

# Discussion

### 2021-06-11 paper/paper_TE_bilingual_vocabulary_model.tex
rmarkdown (not finished yet)

--- 2021-06-11 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
----
-title: "Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants"
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+%
+\documentclass[
+  english,
+  ,man,floatsintext]{apa6}
+\usepackage{lmodern}
+\usepackage{amsmath}
+\usepackage{ifxetex,ifluatex}
+\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+  \usepackage{amssymb}
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\hypersetup{
+  pdftitle={Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants},
+  pdfauthor={Rachel Ka-Ying Tsui1, 2, Ana Maria Gonzalez-Barrero1, Esther Schott1, \& Krista Byers-Heinlein1},
+  pdflang={en-EN},
+  pdfkeywords={bilingualism, infants, translation equivalents, vocabulary development, word learning, computational modeling},
+  hidelinks,
+  pdfcreator={LaTeX via pandoc}}
+\urlstyle{same} % disable monospaced font for URLs
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
+% Manuscript styling
+\usepackage{upgreek}
+\captionsetup{font=singlespacing,justification=justified}
 
-shorttitle: "Are translation equivalents special for bilingual infants?"
+% Table formatting
+\usepackage{longtable}
+\usepackage{lscape}
+% \usepackage[counterclockwise]{rotating}   % Landscape page setup for large tables
+\usepackage{multirow}		% Table styling
+\usepackage{tabularx}		% Control Column width
+\usepackage[flushleft]{threeparttable}	% Allows for three part tables with a specified notes section
+\usepackage{threeparttablex}            % Lets threeparttable work with longtable
 
-header-includes:
-  - \usepackage{amsmath}
-  - \renewcommand{\topfraction}{1}
-  - \renewcommand{\bottomfraction}{1}
-  - \renewcommand{\textfraction}{.1}
-  - \renewcommand{\floatpagefraction}{1}
-  - \setcounter{topnumber}{9}
-  - \setcounter{bottomnumber}{9}
-  - \setcounter{totalnumber}{20}
-  - \setcounter{dbltopnumber}{9}
-  - \DeclareUnicodeCharacter{2265}{\ensuremath{\ge}}
-  - \DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
+% Create new environments so endfloat can handle them
+% \newenvironment{ltable}
+%   {\begin{landscape}\centering\begin{threeparttable}}
+%   {\end{threeparttable}\end{landscape}}
+\newenvironment{lltable}{\begin{landscape}\centering\begin{ThreePartTable}}{\end{ThreePartTable}\end{landscape}}
 
-author: 
-  - name          : "Rachel Ka-Ying Tsui"
-    affiliation   : "1, 2"
-    corresponding : yes    
-    address       : "Department of Psychology, 7141 Sherbrooke St.
-West, Montreal, QC, Canada, H2T1V2, or Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
-    email         : "rachelkytsui@gmail.com"
-  - name          : "Ana Maria Gonzalez-Barrero"
-    affiliation   : "1" 
-  - name          : "Esther Schott"
-    affiliation   : "1" 
-  - name          : "Krista Byers-Heinlein"
-    affiliation   : "1"
+% Enables adjusting longtable caption width to table width
+% Solution found at http://golatex.de/longtable-mit-caption-so-breit-wie-die-tabelle-t15767.html
+\makeatletter
+\newcommand\LastLTentrywidth{1em}
+\newlength\longtablewidth
+\setlength{\longtablewidth}{1in}
+\newcommand{\getlongtablewidth}{\begingroup \ifcsname LT@\roman{LT@tables}\endcsname \global\longtablewidth=0pt \renewcommand{\LT@entry}[2]{\global\advance\longtablewidth by ##2\relax\gdef\LastLTentrywidth{##2}}\@nameuse{LT@\roman{LT@tables}} \fi \endgroup}
 
-affiliation:
-  - id            : "1"
-    institution   : "Department of Psychology, Concordia University"
-  - id            : "2"
-    institution   : "Laboratory for Language Development, RIKEN Center for Brain Science"
+% \setlength{\parindent}{0.5in}
+% \setlength{\parskip}{0pt plus 0pt minus 0pt}
 
-abstract: |
-  The acquisition of translation equivalents is often considered a special component of bilingual children’s vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
-This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are neither harder nor easier to learn.
-To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
-Study 2 tested model-derived predictions against vocabulary data from 200 French–English bilingual children aged 18–33 months.
-Results showed a close match between the model predictions and bilingual children’s patterns of translation equivalent learning.
+% \usepackage{etoolbox}
+\makeatletter
+\patchcmd{\HyOrg@maketitle}
+  {\section{\normalfont\normalsize\abstractname}}
+  {\section*{\normalfont\normalsize\abstractname}}
+  {}{\typeout{Failed to patch abstract.}}
+\patchcmd{\HyOrg@maketitle}
+  {\section{\protect\normalfont{\@title}}}
+  {\section*{\protect\normalfont{\@title}}}
+  {}{\typeout{Failed to patch title.}}
+\makeatother
+\shorttitle{Are translation equivalents special for bilingual infants?}
+\keywords{bilingualism, infants, translation equivalents, vocabulary development, word learning, computational modeling}
+\usepackage{lineno}
+
+\linenumbers
+\usepackage{csquotes}
+\usepackage{amsmath}
+\renewcommand{\topfraction}{1}
+\renewcommand{\bottomfraction}{1}
+\renewcommand{\textfraction}{.1}
+\renewcommand{\floatpagefraction}{1}
+\setcounter{topnumber}{9}
+\setcounter{bottomnumber}{9}
+\setcounter{totalnumber}{20}
+\setcounter{dbltopnumber}{9}
+\DeclareUnicodeCharacter{2265}{\ensuremath{\ge}}
+\DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
+\ifxetex
+  % Load polyglossia as late as possible: uses bidi with RTL langages (e.g. Hebrew, Arabic)
+  \usepackage{polyglossia}
+  \setmainlanguage[]{english}
+\else
+  \usepackage[shorthands=off,main=english]{babel}
+\fi
+\ifluatex
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+
+\title{Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants}
+\author{Rachel Ka-Ying Tsui\textsuperscript{1, 2}, Ana Maria Gonzalez-Barrero\textsuperscript{1}, Esther Schott\textsuperscript{1}, \& Krista Byers-Heinlein\textsuperscript{1}}
+\date{}
+
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Psychology, Concordia University\\\textsuperscript{2} Laboratory for Language Development, RIKEN Center for Brain Science}
+
+\abstract{
+The acquisition of translation equivalents is often considered a special component of bilingual children's vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
+This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are neither harder nor easier to learn.
+To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model --- the Bilingual Vocabulary Model --- which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
+Study 2 tested model-derived predictions against vocabulary data from 200 French--English bilingual children aged 18--33 months.
+Results showed a close match between the model predictions and bilingual children's patterns of translation equivalent learning.
 At smaller vocabulary sizes, data matched the Preference Account, while at larger vocabulary sizes they matched the Neutral Account.
 Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.
+}
 
-keywords          : "bilingualism, infants, translation equivalents, vocabulary development, word learning, computational modeling"
-
-wordcount         :
-
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
-footnotelist      : no
-lineno            : yes
-mask              : no # "yes" to omit identifying information from the title page
-
-#bibliography      : ["gaze-following_references.bib"]
-#replace_ampersands: yes
-
-class             : "man"
-output            : papaja::apa6_pdf
----
-
-```{r load_packages_settings, include = FALSE}
-library(rmarkdown)
-library(knitr)
-library(papaja)
-library(kableExtra)
-library(english)
-library(here)
-library(tidyverse)
-library(tidylog)
-library(wordbankr)
-library(ggplot2)
-library(patchwork)
-library(gridExtra)
-library("ggpubr")
-library(lme4)
-library(lmerTest)
-library(scales)
-
-`%notin%` <- Negate(`%in%`)
-
-knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE, cache = TRUE)
-
-```
-
-```{r load_data}
-# read in the final cleaned keepers_data
-keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE_final.csv"))
-```
+\begin{document}
+\maketitle
 
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
-For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer, Bornstein, & De Coster, 2006; Pearson et al., 1995).
-Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics.
-For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak & Byers-Heinlein, 2019).
-That is, upon hearing the English word “apple”, the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent ( a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent.
+For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
+These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Pearson et al., 1995).
+Translation equivalents are thought to hold a special status in a bilingual's developing lexicon due to the strong overlap in their semantics.
+For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
+That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
+In vocabulary acquisition, bilingual children must learn a first label for a referent ( a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent.
 Is translation equivalent learning different from singlet learning?
 The current paper contrasts three competing accounts: 1) translation equivalents are hard to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are neither harder nor easier to learn than singlets (Neutral Account).
-To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model as a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level.
+To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model as a computational account of vocabulary learning, with parameters including bilinguals' vocabulary in each language and their developmental level.
 In Study 1, we use the Bilingual Vocabulary Model to derive a set of predictions, which we then test against vocabulary data from 200 18- to 33-month-old bilingual children in Study 2.
 
-## Accounts of translation equivalent learning
+\hypertarget{accounts-of-translation-equivalent-learning}{%
+\subsection{Accounts of translation equivalent learning}\label{accounts-of-translation-equivalent-learning}}
 
-### Avoidance Account: Translation equivalents are harder to learn than singlets
-Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies (e.g., Imedadze, 1978; Swain & Wesche 1975; Volterra & Taeschner, 1978).
+\hypertarget{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}{%
+\subsubsection{Avoidance Account: Translation equivalents are harder to learn than singlets}\label{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}}
+
+Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children's early vocabularies (e.g., Imedadze, 1978; Swain \& Wesche 1975; Volterra \& Taeschner, 1978).
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman & Wachtel, 1988; Markman, 1992; 1994).
-For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug”, they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.
+This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman \& Wachtel, 1988; Markman, 1992; 1994).
+For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like ``wug,'' they assume that ``wug'' refers to the garlic press --- the object unknown to them --- rather than to the cup, the object for which they already know the word.
 
-Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein & Werker, 2009; Davidson & Tell, 2005; Houston-Price, Caloghiris, & Raviglione, 2010).
+Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein \& Werker, 2009; Davidson \& Tell, 2005; Houston-Price, Caloghiris, \& Raviglione, 2010).
 When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word.
@@
 
-Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David & Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017).
+Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David \& Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017).
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
-For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010).
+For example, when hearing a novel word like ``nil,'' monolingual children aged 17--22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein \& Werker, 2009; 2013; Houston-Price et al., 2010).
 A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al, 2020).
@@
 
-### Preference Account: Translation equivalents are easier to learn than singlets
+\hypertarget{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}{%
+\subsubsection{Preference Account: Translation equivalents are easier to learn than singlets}\label{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}}
+
 Contrary to the Avoidance Account, the Preference Account posits that translation equivalents are easier to learn than singlets.
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
-Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak & Byers-Heinlein, 2018; Singh, 2013).
+When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
+Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak \& Byers-Heinlein, 2018; Singh, 2013).
 In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
@@
 There are several lines of empirical evidence to support the Preference Account.
-For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages (Lanvers, 1999; Pearson & Fernández, 1994).
-The main explanation that has been given for this finding is that additional exposure to their non-dominant language — which became their new dominant language — enabled fast mapping of words to already-lexicalized concepts.
+For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages (Lanvers, 1999; Pearson \& Fernández, 1994).
+The main explanation that has been given for this finding is that additional exposure to their non-dominant language --- which became their new dominant language --- enabled fast mapping of words to already-lexicalized concepts.
 
 Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015).
-The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like “cat” and “dog” are strongly semantically related).
+The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related).
 Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in the lexicons.
-This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English “dog” and French _“chien”_) and words that had other cross-language relations (e.g., “cat” and _“chien”_).
+This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English ``dog'' and French \emph{``chien''}) and words that had other cross-language relations (e.g., ``cat'' and \emph{``chien''}).
 The authors then simulated bilingual vocabulaires by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals.
-Comparison with actual bilingual children’s vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
+Comparison with actual bilingual children's vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
 The authors therefore concluded that bilingual children show a preference for learning translation equivalents over singlets.
-Note that, in their study, translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word “cat”).
-However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning including vocabulary size in each language and the developmental level of a child — an issue that we will return to later in the introduction.
+Note that, in their study, translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat'').
+However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning including vocabulary size in each language and the developmental level of a child --- an issue that we will return to later in the introduction.
 
@@
 
-### Neutral Account: Translation equivalents are similar to learn than singlets
+\hypertarget{neutral-account-translation-equivalents-are-similar-to-learn-than-singlets}{%
+\subsubsection{Neutral Account: Translation equivalents are similar to learn than singlets}\label{neutral-account-translation-equivalents-are-similar-to-learn-than-singlets}}
+
 The previous accounts rely on the idea that bilingual vocabulary development unfolds differently than monolingual development, as monolinguals encounter only singlets but bilinguals encounter both singlets and translation equivalents.
-There is an underlying assumption that translation equivalent learning is somehow special relative to singlet learning — the Avoidance Account proposes that translation equivalents are harder to learn than singlets, whereas the Preference Account proposes that translation equivalents are easier to learn than singlets.
+There is an underlying assumption that translation equivalent learning is somehow special relative to singlet learning --- the Avoidance Account proposes that translation equivalents are harder to learn than singlets, whereas the Preference Account proposes that translation equivalents are easier to learn than singlets.
 However, it is also possible that translation equivalents are neither harder nor easier to learn than singlets.
@@
 
-The Neutral Account implies that bilingual children’s two languages develop relatively independently.
+The Neutral Account implies that bilingual children's two languages develop relatively independently.
 Indeed, language and processing measures in bilinguals tend to be tightly correlated within a particular language, with only weak cross-language correlations.
-For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, & Hurtado, 2010).
-Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
+For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, \& Hurtado, 2010).
+Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
 Bilingual children who heard a greater amount of a particular language tended to process that language faster, and in turn learned more words in that language.
@@
 In a study whose results directly support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children.
-It was revealed that the percentage of translation equivalents observed in English–Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
+It was revealed that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
 This evidence suggested that the acquisition of translation equivalents is not that different from the expected overlap between two independent lexicons, implying that singlets and translation equivalents are equally learnable.
@@
 
-## Contributors to translation equivalent knowledge
+\hypertarget{contributors-to-translation-equivalent-knowledge}{%
+\subsection{Contributors to translation equivalent knowledge}\label{contributors-to-translation-equivalent-knowledge}}
+
 The previous section discussed three theoretical accounts concerning the relative learnability of translation equivalents.
 However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than as an integrated system.
-In this section, we consider two proximal variables that we expect to predict the number of translation equivalents children know: vocabulary size in each language, and word learnability as a function of children’s developmental level.
+In this section, we consider two proximal variables that we expect to predict the number of translation equivalents children know: vocabulary size in each language, and word learnability as a function of children's developmental level.
 
-### Vocabulary size in bilinguals’ two languages
+\hypertarget{vocabulary-size-in-bilinguals-two-languages}{%
+\subsubsection{Vocabulary size in bilinguals' two languages}\label{vocabulary-size-in-bilinguals-two-languages}}
+
 Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know.
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which appears to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, & Hurtado, 2010; Place & Hoff, 2011; Pearson et al., 1997).
-Bilingual children tend to know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
+In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, \& Hurtado, 2010; Place \& Hoff, 2011; Pearson et al., 1997).
+Bilingual children tend to know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
 
-One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts.
+One important consideration in thinking about bilingual children's experience is whether they encounter and use their languages within the same or different contexts.
 This is known as the Complementarity Principle (Grosjean, 2016).
@@
 However, the Complementarity Principle may not necessarily impact bilingual word acquisition in infancy.
-Most words in children’s early vocabularies could be considered “home words”, which include words for social and daily routines (“hello”, “more”, “diaper”), common nouns (“doggie”), and everyday verbs (“walk”).
+Most words in children's early vocabularies could be considered ``home words,'' which include words for social and daily routines (``hello,'' ``more,'' ``diaper''), common nouns (``doggie''), and everyday verbs (``walk'').
 Such words are likely to be encountered across contexts where children spend the majority of their time, such as at home and at childcare.
-Thus, growing up in a bilingual context from birth, bilingual children presumably encounter “home words” in both languages.
-Accordingly, we assume that for the most part, bilingual children’s opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle.
+Thus, growing up in a bilingual context from birth, bilingual children presumably encounter ``home words'' in both languages.
+Accordingly, we assume that for the most part, bilingual children's opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle.
 We further consider implications of this assumption in the discussion section.
 
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, & Weisleder, 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).
+Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013).
+Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman et al., 2008; Swingley \& Humphrey, 2018).
 
-### Word learnability as a function of developmental level
-An often overlooked factor that could contribute to bilingual children’s learning of translation equivalents is related to the changes in the learnability of different words over time based on children’s developmental level.
+\hypertarget{word-learnability-as-a-function-of-developmental-level}{%
+\subsubsection{Word learnability as a function of developmental level}\label{word-learnability-as-a-function-of-developmental-level}}
+
+An often overlooked factor that could contribute to bilingual children's learning of translation equivalents is related to the changes in the learnability of different words over time based on children's developmental level.
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
@@
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
-That is, potential singlets might be “too hard” to be learned at a particular age.
+That is, potential singlets might be ``too hard'' to be learned at a particular age.
 Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
 
-## The Bilingual Vocabulary Model
-Taking into account the contributions of language exposure and developmental level to bilingual children’s vocabulary acquisition, we put forward the Bilingual Vocabulary Model.
-This model proposes that the number of translation equivalents that children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children’s developmental level.
+\hypertarget{the-bilingual-vocabulary-model}{%
+\subsection{The Bilingual Vocabulary Model}\label{the-bilingual-vocabulary-model}}
+
+Taking into account the contributions of language exposure and developmental level to bilingual children's vocabulary acquisition, we put forward the Bilingual Vocabulary Model.
+This model proposes that the number of translation equivalents that children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children's developmental level.
 We formalize learning a translation equivalent pair as the joint probability of learning each of the words in the pair.
 This provides a clear empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
-The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B | A) where P(B | A) = P(B) if A and B are independent (see Box 1 for a detailed example).
+The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B \textbar{} A) where P(B \textbar{} A) = P(B) if A and B are independent (see Box 1 for a detailed example).
 The full model is shown in Figure 1.
@@
 
-&nbsp;
+~
 
-```{r box1, results="asis", fig.cap = " "}
-box1 <- data.frame(title = "To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
+\begingroup\fontsize{10}{12}\selectfont
+
+\begin{longtable}{>{\raggedright\arraybackslash}p{450px}}
+\toprule
+Box 1. Example of test for independence.\\
+\midrule
+To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
 To determine independence, we must know the probability of being a psychology major, the probability of being left handed, and the probability of being both a psychology major and left-handed.
@@
 Finally, if left-handedness and majoring in psychology are independent, we can predict the number of left-handed psychology students by multiplying the observed number of left-handed students (100) by the observed number of psychology students (200), and dividing by the total population of the college (1000), so for example 100×200/1000 = 20.
-Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.") %>%
-  rename("Box 1. Example of test for independence." = title)
+Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.\\
+\bottomrule
+\end{longtable}
+\endgroup{}
 
-kable(box1,
-      format="latex", booktabs=TRUE, longtable = TRUE) %>%
-  column_spec(1, width = "450px") %>%
-  kable_styling(font_size = 10)
+\begin{figure}
 
-```
+{\centering \includegraphics[width=1\linewidth]{./Figure1_illustration}
 
-```{r fig1, fig.cap="Illustration of the Bilingual Vocabulary Model.", fig.align='center', out.width='100%'}
+}
 
-knitr::include_graphics("./Figure1_illustration.jpg")
-```
+\caption{Illustration of the Bilingual Vocabulary Model.}\label{fig:fig1}
+\end{figure}
 
-The model takes four main parameters: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS > 1) or against (BIAS < 1) learning translation equivalents.
+The model takes four main parameters: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS \textgreater{} 1) or against (BIAS \textless{} 1) learning translation equivalents.
 The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language.
@@
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-Very young children would not be expected to know many of the “harder” words on the CDI, such as “lawn mower”, “sidewalk”, or “vitamins'', due to children’s immature cognitive machinery and conceptual development.
-A more reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
+Very young children would not be expected to know many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins'\,', due to children's immature cognitive machinery and conceptual development.
+A more reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
 For example, Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
 Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
-Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable — if imperfect — estimate of the number of learnable words (LEARNABLE) that a child of Jamie’s age could potentially acquire in each language.
+Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a child of Jamie's age could potentially acquire in each language.
 Thus, we might expect that Jamie could potentially have learned up to 245 English words and 245 French words, although he has thus far only learned 50 in English and 20 in French.
 
-Using the mathematical concept of independence, we can then quantify the number of translation equivalents (TE) expected given children’s vocabulary sizes in the dominant (DOM) and non-dominant (NONDOM) languages, as well as the number of potentially learnable words (LEARNABLE).
-If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) — which is the possible number of words that could overlap across both languages — to predict the number of translation equivalents.
+Using the mathematical concept of independence, we can then quantify the number of translation equivalents (TE) expected given children's vocabulary sizes in the dominant (DOM) and non-dominant (NONDOM) languages, as well as the number of potentially learnable words (LEARNABLE).
+If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) --- which is the possible number of words that could overlap across both languages --- to predict the number of translation equivalents.
 We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral account.
@@
 Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
-Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child — that is, the total number of concepts that are lexicalized in either language.
+Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child --- that is, the total number of concepts that are lexicalized in either language.
 This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD).
@@
 Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM).
-It is also possible to decompose children’s word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
+It is also possible to decompose children's word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
 
-```{r table1, results = "asis"}
+\begin{landscape}\begin{table}
 
-table1 <- data.frame(
-  Variable = c("LEARNABLE", "DOM", "NONDOM", "BIAS", "BALANCE", "WORD", "TE", "CONCEPT", "DOM-SINGLET", "NONDOM-SINGLET"),
-  Definition = c("Number of learnable words in each language, given the child’s developmental level",
-                   "Words produced in the dominant language",
-                   "Words produced in the non-dominant language",
-                   "Bias parameter",
-                   "Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages)",
-                   "Word vocabulary (or total vocabulary size)",
-                   "Translation equivalents produced",
-                   "Concept vocabulary (or total conceptual vocabulary size)",
-                   "Singlets in dominant language",
-                   "Singlets in non-dominant language"),
-  Constraints = c("Varies by age. No greater than the number of words on CDI.",
-                  "DOM \u2265 NONDOM (i.e., always produce more words in dominant than non-dominant language); DOM \u2264 LEARNABLE",
-                  "NONDOM \u2264 DOM (always produce fewer words in non-dominant than dominant language); NONDOM \u2264 LEARNABLE",
-                  "BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account",
-                  ".00 \u2264 BALANCE \u2264 .50 (greater values indicate children producing a more similar number of words in their two languages)",
-                  "W \u2264 2LEARNABLE (maximum word vocabulary is knowing each word in both languages)",
-                  "", "", "", ""),
-  "Relationship to other parameters" = c("Maximum number that could be learned in DOM or NONDOM",
-                                         "DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM",
-                                         "NONDOM = BALANCE×WORD; NONDOM = WORD - DOM",
-                                         "Account = BIAS×(DOM×NONDOM/LEARNABLE)",
-                                         "BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)",
-                                         "WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)",
-                                         "If DOM and NONDOM are independent, TE = DOM×NONDOM/LEARNABLE",
-                                         "CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET",
-                                         "DOM-SINGLET = DOM - TE",
-                                         "NONDOM_SINGLET = NONCOM - TE")
-)
+\caption{\label{tab:table1}Summary of the parameters in the Bilingual Vocabulary Model.}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{l>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}}
+\toprule
+Variable & Definition & Constraints & Relationship.to.other.parameters\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Main Parameters}}\\
+\hspace{1em}LEARNABLE & Number of learnable words in each language, given the child’s developmental level & Varies by age.
+No greater than the number of words on CDI.
+& Maximum number that could be learned in DOM or NONDOM\\
+\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (i.e., always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD - DOM\\
+\hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & Account = BIAS×(DOM×NONDOM/LEARNABLE)\\
+\addlinespace[0.3em]
+\multicolumn{4}{l}{\textbf{Derived Parameters}}\\
+\hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & .00 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)\\
+\hspace{1em}TE & Translation equivalents produced &  & If DOM and NONDOM are independent, TE = DOM×NONDOM/LEARNABLE\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language &  & DOM-SINGLET = DOM - TE\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language &  & NONDOM\_SINGLET = NONCOM - TE\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
 
-landscape(
-  kable(table1,
-        format = "latex", booktabs = TRUE,
-        caption = "Summary of the parameters in the Bilingual Vocabulary Model.") %>%
-    kableExtra::kable_styling(font_size = 7) %>%
-    pack_rows("Main Parameters", 1, 4) %>%
-    pack_rows("Derived Parameters", 5, 10) %>%
-    column_spec(2:4, width = "175px")
-)
-```
+\hypertarget{current-research}{%
+\subsection{Current research}\label{current-research}}
 
-## Current research
 The current research aimed to better understand the nature of translation equivalent learning in bilingual children.
@@
 
-In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997).
+In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997).
 Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
 
-# Study 1: Simulations
+\hypertarget{study-1-simulations}{%
+\section{Study 1: Simulations}\label{study-1-simulations}}
+
 Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction (see also Figure 1), which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
@@
 
-```{r table2, results = "asis"}
+\begin{landscape}\begin{table}
 
-table2 <- data.frame(
-  Simulation = c("1", "2", "3"),
-  LEARNABLE = c("Constant at 600", "Varied at 300, 450, and 600", "Varied at 150, 300, 450, and 600"),
-  DOM = c("Varied, ranging from 100 to LEARNABLE at an interval of 100", 
-          "Varied, ranging from 100 to LEARNABLE at an interval of 100",
-          "Varied, ranging from 100 to LEARNABLE at an interval of 100 "),
-  NONDOM = c("Varied, ranging from 0 to DOM at an interval of 10",
-             "Varied, ranging from 0 to DOM at an interval of 25",
-             "Varied, ranging from 0 to DOM at an interval of 25"),
-  WORD = c("Calculated as WORD = DOM + NONDOM",
-           "Calculated as WORD = DOM + NONDOM",
-           "Calculated as WORD = DOM + NONDOM"),
-  BALANCE = c("Calculated as BALANCE = NONDOM / (DOM+NONDOM)",
-              "Calculated as BALANCE = NONDOM / (DOM+NONDOM)",
-              "Calculated as BALANCE = NONDOM / (DOM+NONDOM)"),
-  BIAS = c("Constant at 1", "Constant at 1", 
-           "Varied at 0.5 (Avoidance Account), \n 1 (Neutral Account), \n and 1.5 (Preference Account)"),
-  n_data = c(216, 161, 166)
-)
+\caption{\label{tab:table2}Summary of the parameters used in each simulation.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{>{\centering\arraybackslash}p{45px}>{\raggedright\arraybackslash}p{45px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{45px}}
+\toprule
+Simulation & Learnable words (LEARNABLE) & Words in dominant Language (DOM) & Words in non-dominant language (NONDOM) & Word vocabulary (WORD) & Balance of vocabulary (BALANCE) & Bias parameter (BIAS) & Total number of data points generated\\
+\midrule
+1 & Constant at 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 10 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 216\\
+2 & Varied at 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 161\\
+3 & Varied at 150, 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Varied at 0.5 (Avoidance Account),
+ 1 (Neutral Account), 
+ and 1.5 (Preference Account) & 166\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
 
-landscape(
-  kable(table2,
-        col.names = c("Simulation",
-                      "Learnable words (LEARNABLE)",
-                      "Words in dominant Language (DOM)",
-                      "Words in non-dominant language (NONDOM)",
-                      "Word vocabulary (WORD)",
-                      "Balance of vocabulary (BALANCE)",
-                      "Bias parameter (BIAS)",
-                      "Total number of data points generated"),
-        align = "clllllll",
-        format="latex", booktabs = TRUE,
-        caption = "Summary of the parameters used in each simulation.") %>%
-    column_spec(1:2, width = "45px") %>%
-    column_spec(3:7, width = "75px") %>%
-    column_spec(8, width = "45px") %>%
-    kable_styling(font_size = 8)
-)
+\hypertarget{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}{%
+\subsection{Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}
 
-```
-
-## Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
 In Simulation 1, we first illustrate the mathematical relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
@@
 Infant Annie (small vocabulary, unbalanced exposure) produces 270 words in the dominant language and 30 words in her non-dominant language.
-She has a word vocabulary of 300, and a balance score of .10 (10% of her words are in the non-dominant language).
+She has a word vocabulary of 300, and a balance score of .10 (10\% of her words are in the non-dominant language).
 Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents.
 Infant Bernie (small vocabulary, balanced exposure) produces 180 dominant-language words, and 120 non-dominant language words.
-Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40% of his words are in the non-dominant language).
+Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40\% of his words are in the non-dominant language).
 Based on our formula, we expect Bernie to produce 36 translation equivalents.
@@
 
-```{r table3, results="asis"}
+\begin{landscape}\begin{table}
 
-# create columns that contain information to be added to table3 in the next step
-## Definition column: the definition of each variable in the table
-Definition <- c("Bias parameter", "Learnable words", "Words produced in the dominant language", "Words produced in the non-dominant language",
-                "Word vocabulary (or total vocabulary size)", "Vocabulary balance", "Translation equivalents produced", 
-                "Concept vocabulary (or total conceptual vocabulary size)", "Singlets in dominant language", "Singlets in non-dominant language")
-## Calculation column: the formula used to calculate the derived parameters
-Calculation <- c("", "", "", "", 
-                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM * NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
-
-table3 <- data.frame(
-    # define the main parameters
-    hypotheical_child = c("Infant Anne (small vocabulary, unbalanced)",
-                        "Infant Bernie (small vocabulary, balanced)",
-                        "Infant Charlie (large vocabulary, balanced)"),
-    BIAS = 1, 
-    LEARNABLE = 600, 
-    DOM = c(270, 180, 540),
-    NONDOM = c(30, 120, 360)
-  ) %>%
-  # calculate the derived parameters from the main parameters
-  mutate(WORD = DOM + NONDOM,
-         BALANCE = NONDOM/(DOM+NONDOM),
-         TE = DOM*NONDOM/LEARNABLE,
-         CONCEPT = WORD - TE,
-         "DOM-SINGLET" = DOM - TE, 
-         "NONDOM-SINGLET" = NONDOM - TE) %>%
-  # transpose the data set for presentation
-  pivot_longer(-("hypotheical_child"), names_to = "Variable", values_to = "value") %>%
-  pivot_wider(names_from = "hypotheical_child", values_from = "value") %>%
-  # append other information for presentation
-  mutate(Definition = Definition, 
-         Calculation = Calculation) %>%
-  # reorder columns
-  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)")
-
-landscape(
-  kable(table3,
-      align = "llllll",
-      format="latex", booktabs = TRUE,
-      caption = "Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.") %>%
-  pack_rows("Main Parameters", 1, 4) %>%
-  pack_rows("Derived Parameters", 5, 10) %>%
-  column_spec(c(1), width = "100px") %>%
-  column_spec(c(2:3), width = "140px") %>%
-  column_spec(c(4:6), width = "60px") %>%
-  kable_styling(font_size = 8) 
-)
-
-```
+\caption{\label{tab:table3}Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{100px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}}
+\toprule
+Variable & Definition & Calculation & Infant Anne (small vocabulary, unbalanced) & Infant Bernie (small vocabulary, balanced) & Infant Charlie (large vocabulary, balanced)\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Main Parameters}}\\
+\hspace{1em}BIAS & Bias parameter &  & 1.0 & 1.0 & 1.0\\
+\hspace{1em}LEARNABLE & Learnable words &  & 600.0 & 600.0 & 600.0\\
+\hspace{1em}DOM & Words produced in the dominant language &  & 270.0 & 180.0 & 540.0\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language &  & 30.0 & 120.0 & 360.0\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Derived Parameters}}\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & DOM + NONDOM & 300.0 & 300.0 & 900.0\\
+\hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0.1 & 0.4 & 0.4\\
+\hspace{1em}TE & Translation equivalents produced & DOM * NONDOM / LEARNABLE & 13.5 & 36.0 & 324.0\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286.5 & 264.0 & 576.0\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256.5 & 144.0 & 216.0\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16.5 & 84.0 & 36.0\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
 
@@
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
-Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
+Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
 In Figure 2, we plotted TE knowledge as predicted by DOM, NONDOM, and WORD at different levels of BALANCE.
@@
 
-* Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE). 
-* Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE). 
-* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge(see panel 1B of Figure 2).
+\begin{itemize}
+\tightlist
+\item
+  Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE).
+\item
+  Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE).
+\item
+  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge(see panel 1B of Figure 2).
+\end{itemize}
 
-```{r simulation1, echo = FALSE}
-# Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (LEARNABLE = 600)
+\begin{figure}
 
-## generate the simulation data
-simulation1 <- function (LEARNABLE) {
-  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
-  NONDOM_seq <- seq(0, LEARNABLE, by = 10) # sequence of N from 0 to V at an interval of 10
-  DOM <- rep (DOM_seq, length(NONDOM_seq))
-  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
-  data_simulation1 <<- data.frame(LEARNABLE, DOM, NONDOM)
+{\centering \includegraphics[width=1.1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig2-1}
+
 }
 
-simulation1(600) # setting V = 600
+\caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to total dominant vocabulary size (DOM; Panel A), total non-dominant vocabulary size (NONDOM; Panel B), and total word vocabulary (WORD; Panel C).
+Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600.
+Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig2}
+\end{figure}
 
-## clean the simulation data
-data_simulation1 <- data_simulation1 %>%
-  arrange(DOM, NONDOM) %>%
-  filter(NONDOM <= DOM) %>%
-  # Get word vocabulary W
-  mutate(WORD = DOM + NONDOM) %>%
-  # Create other variables
-  mutate(BALANCE = NONDOM/(DOM+NONDOM), 
-         TE = (DOM*NONDOM)/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
-  # Round to keep whole numbers only
-  mutate_at(vars(-BALANCE), round, 0) %>%
-  # bin B into 5 groups
-  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  arrange(DOM, NONDOM, bin_BALANCE)
+\hypertarget{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}{%
+\subsection{Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}
 
-## visualizing Simulation 1
-simulation1 <- data_simulation1 %>%
-  # convert to long data
-  pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
-  mutate(vocab_type = recode(vocab_type, 
-                             DOM = "Panel 1A: \n Dominant vocabulary \n (DOM)", 
-                             NONDOM = "Panel 1B: \n Non-dominant vocabulary \n (NONDOM)",
-                             WORD = "Panel 1C: \n Total vocabulary \n (WORD)")) %>%
-  ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
-  stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
-  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
-  theme_minimal() + 
-  facet_grid(. ~ vocab_type) +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom",
-        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
-        plot.title.position = "plot",
-        text = element_text(size=12),
-        panel.spacing.x = unit(2, "lines"),
-        panel.background = element_rect(fill = NA, color = "black")) +
-  ylim(0, 600) +
-  xlim(0, 1000) +
-  labs(x = "Number of words", 
-       y = "Number of translation equivalents (TE)", 
-       color = "Vocabulary balance (BALANCE)",
-       linetype = "Vocabulary balance (BALANCE)",
-       title = "1) Simulated data") +
-  guides(color = guide_legend(reverse = T, nrow = 1),
-         linetype = guide_legend(reverse = T, nrow = 1)) 
-```
-
-```{r simulation1-plus_observed_data, echo = FALSE}
-Observed_plot_inputs_TE <- keepers_ws_TE %>% 
-  select(baby_id, ID_testdate, age_days, number_of_te, word_vocab, total_words_dom, total_words_nondom, balance_vocab) %>%
-  pivot_longer(c(word_vocab, total_words_dom, total_words_nondom), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("total_words_dom", "total_words_nondom", "word_vocab"))) %>%
-  mutate(vocab_type = recode(vocab_type, 
-                             total_words_dom = "Panel 2A: \n Dominant vocabulary \n (DOM)", 
-                             total_words_nondom = "Panel 2B:\n Non-dominant vocabulary \n (NONDOM)",
-                             word_vocab = "Panel 2C: \n Total vocabulary \n (WORD)")) %>%
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
-  #geom_point() +
-  stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
-  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
-  theme_minimal() + 
-  facet_grid(. ~ vocab_type) +
-  theme(axis.text.x = element_text(angle = 90),
-        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
-        plot.title.position = "plot",
-        text = element_text(size=12),
-        legend.position = "bottom",
-        panel.spacing.x = unit(2, "lines"),
-        panel.background = element_rect(fill = NA, color = "black")) +
-  ylim(0, 600) +
-  xlim(0, 1000) +
-  labs(x = "Number of words", 
-       y = "Number of translation equivalents (TE)", 
-       color = "Vocabulary balance (BALANCE)",
-       linetype = "Vocabulary balance (BALANCE)",
-       title = "2) Observed data") +
-  guides(colour = guide_legend(reverse = T, nrow = 1),
-         linetype = guide_legend(reverse = T, nrow = 1))
-
-# combine Simulation1 visualization + observed data visualization
-plot_Simulation1plusData <- ggarrange(simulation1, 
-                                      Observed_plot_inputs_TE,
-                                      nrow = 2,
-                                      common.legend = TRUE, legend = "bottom")
-```
-
-```{r fig2, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to total dominant vocabulary size (DOM; Panel A), total non-dominant vocabulary size (NONDOM; Panel B), and total word vocabulary (WORD; Panel C).
-Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600.
-Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%"}
-
-plot_Simulation1plusData
-```
-
-## Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
 In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn the word for 600 different concepts (i.e., LEARNABLE held constant at 600).
@@
 Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM with the formula NONDOM/(DOM+NONDOM).
-We also counted the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children’s concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
+We also counted the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children's concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
 Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), medium balanced (BALANCE = .20 - .35), and least balanced (BALANCE = .00 - .02).
@@
 
-* Prediction 2a: Have larger concept vocabularies (CONCEPT) 
-* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
-* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with a less balanced vocabulary (BALANCE) produce the most DOM-SINGLET.
-* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with a more balanced vocabulary (BALANCE) produce the most NONDOM-SINGLET.
+\begin{itemize}
+\tightlist
+\item
+  Prediction 2a: Have larger concept vocabularies (CONCEPT)
+\item
+  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+\item
+  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with a less balanced vocabulary (BALANCE) produce the most DOM-SINGLET.
+\item
+  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with a more balanced vocabulary (BALANCE) produce the most NONDOM-SINGLET.
+\end{itemize}
 
-```{r simulation2, echo = FALSE}
-# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (LEARNABLE = 300, 450, or 600)
+\begin{figure}
 
-## generate the simulation data
-LEARNABLE <- seq(300, 600, by = 150) # set LEARNABLE to 300, 450, and 600
+{\centering \includegraphics[width=0.8\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig3-1}
 
-simulation2 <- function (LEARNABLE) {
-  DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
-  NONDOM_seq <- seq(0, LEARNABLE, by = 25) # sequence of N from 0 to V at an interval of 25
-  DOM <- rep(DOM_seq, length(NONDOM_seq))
-  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
-  data_simulation2 <<- data.frame(LEARNABLE, DOM, NONDOM)
 }
 
-data_simulation2 <- do.call(rbind, lapply(LEARNABLE, simulation2)) %>%
-  arrange(LEARNABLE, DOM, NONDOM) %>%
-  filter(NONDOM <= DOM) %>%
-  # Create other variables
-  mutate(WORD = DOM + NONDOM, # word vocabulary 
-         BALANCE = NONDOM/(DOM+NONDOM), # balance,
-         TE = (DOM*NONDOM)/LEARNABLE, # If D and N are independent, TE = DN/V
-         DOM_SINGLET = DOM-TE, # Unique words in dominant language
-         NONDOM_SINGLET = NONDOM-TE) %>% # Unique words in non-dominant language
-  # Round to keep whole numbers only
-  mutate_at(vars(-BALANCE), round, 0) %>%
-  # bin B into 3 groups
-  mutate(bin_BALANCE = case_when(BALANCE >= 0.36 ~ "Most Balanced",
-                                 BALANCE <= 0.20 ~ "Least Balanced",
-                                 TRUE ~ "Medium Balanced")) %>%
-  mutate(bin_BALANCE  = fct_rev(bin_BALANCE)) %>%
-  arrange(DOM, NONDOM, bin_BALANCE)
-
-## visualizing Simulation 2
-simulation2 <- data_simulation2 %>%
-  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
-  filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
-  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
-                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
-                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
-  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
-  summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ LEARNABLE_label) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
-  ylim(0, 600) +
-  theme_minimal() +
-  theme(legend.position = "bottom",
-        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
-        plot.title.position = "plot",
-        text = element_text(size=10),
-        panel.spacing.x = unit(2, "lines")) + 
-  labs(x = "\nVocabulary balance (BALANCE)", 
-       y = "Number of concepts (CONCEPT)", 
-       title = "Panel A: Simulated data")
-```
-
-```{r simulation2-plus_observed_data, echo = FALSE}
-plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
-  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
-  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
-                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
-                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
-  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
-                               balance_vocab <= 0.20 ~ "Least Balanced",
-                               TRUE ~ "Medium Balanced")) %>%
-  mutate(BalSubset = fct_rev(BalSubset)) %>%
-  arrange(desc(number_of_te)) %>%
-  mutate(TE = number_of_te) %>%
-  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
-               names_to = "Type", values_to = "NumberOfWords") %>%
-  mutate(Type = factor(Type, levels=c("total_singlet_nondom", "total_singlet_dom", "number_of_te"))) %>%
-  mutate(Type = recode(Type, 
-                       total_singlet_nondom = "Singlet in non-dominant language (NONDOM-SINGLET)", 
-                       total_singlet_dom = "Singlet in dominant language (DOM-SINGLET)", 
-                       number_of_te = "Translation equivalent (TE)")) %>%
-  group_by(LEARNABLE_subset, BalSubset, Type) %>%
-  summarise(n_words = mean(NumberOfWords)) %>%
-  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ LEARNABLE_subset) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
-  ylim(0, 600) +
-  theme_minimal() +
-  theme(axis.text.x = element_text(angle = 0) , 
-        legend.position = "bottom",
-        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
-        plot.title.position = "plot",
-        text = element_text(size=10),
-        panel.spacing.x = unit(2, "lines")) + 
-  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
-  labs(x = "\nVocabulary balance (BALANCE)", 
-       y = "Number of concepts (CONCEPT)",
-       title = "Panel B: Observed data")
-
-# combine Simulation2 visualization + observed data visualization
-plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
-                                      plot_stacked_byLEARNABLE,
-                                      nrow = 2,
-                                      heights=c(0.8,1))
-```
-
-```{r fig3, fig.cap="Number of translation equivalents (TE) and singlets in dominant (DOM-SINGLET) and non-dominant language (NONDOM-SINGLET) across different developmental levels/ages, which sets the number of LEARNABLE words.
+\caption{Number of translation equivalents (TE) and singlets in dominant (DOM-SINGLET) and non-dominant language (NONDOM-SINGLET) across different developmental levels/ages, which sets the number of LEARNABLE words.
 Panel A represents the model simulation in Study 1, where developmental levels of simulated children are set at three values: LEARNABLE = 300, 450, and 600.
 Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right).
-Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="80%"}
+Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.}\label{fig:fig3}
+\end{figure}
 
-plot_Simulation2plusData
-```
+\hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}{%
+\subsection{Simulation 3: Bias towards or against translation equivalent learning compared to singlets}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}}
 
-## Simulation 3: Bias towards or against translation equivalent learning compared to singlets
 In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
@@
 Three scenarios of translation equivalent learning (TE) were then generated using the formula TE = BIAS × DOM×NONDOM/LEARNABLE.
-To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50% less likely to be learned than singlets).
+To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50\% less likely to be learned than singlets).
 To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are neither easier nor harder to learn than singlets).
-Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets).
+Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50\% more likely to be learned than singlets).
 In Figure 4, we illustrate the three different scenarios of simulated translation equivalent knowledge.
@@
 
-* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning, with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+\begin{itemize}
+\tightlist
+\item
+  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning, with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+\end{itemize}
 
-```{r simulation3, echo = FALSE}
-# Simulation 3: Non-independence of dominant-language (DOM) and non-dominant language words (NONDOM)
+\begin{figure}
 
-## generate the simulation data
-LEARNABLE <- seq(150, 600, by = 150) # set LEARNABLE to 300, 450, and 600
+{\centering \includegraphics[width=1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig4-1}
 
-simulation3 <- function (LEARNABLE) {
-  DOM_seq = seq(100, LEARNABLE, by = 100) # sequence of DOM from 100 to LEARNABLE at an interval of 100
-  NONDOM_seq = seq(0, LEARNABLE, by = 25) # sequence of NONDOM from 0 to LEARNABLE at an interval of 25
-  DOM <- rep(DOM_seq, length(NONDOM_seq))
-  NONDOM <- rep(NONDOM_seq, length(DOM_seq))
-  data_simulation3 <<- data.frame(LEARNABLE, DOM, NONDOM)
 }
 
-data_simulation3 <- do.call(rbind, lapply(LEARNABLE, simulation3)) %>%
-  arrange(LEARNABLE, DOM, NONDOM) %>%
-  filter(NONDOM <= DOM) %>% # eliminate cases when NONDOM > DOM
-  mutate(WORD = DOM+NONDOM, # word vocabulary
-         BALANCE = NONDOM/(DOM+NONDOM), # balance 
-         TE = DOM*NONDOM/LEARNABLE) %>% # If DOM and NONDOM are independent, TE = DOM*NONDOM/LEARNABLE
-  mutate_at(vars(-BALANCE), round, 0) %>%
-  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  arrange(bin_BALANCE)
+\caption{Different scenarios of expected translation equivalents learning (TE) while assuming bias towards (BIAS < 1) or against learning translation equivalents (BIAS > 1), or no bias (BIAS = 1).}\label{fig:fig4}
+\end{figure}
 
-## Simulation3: TE = BIAS * (DOM*NONDOM/LEARNABLE)
-### defining the bias parameter I
-BIAS1 = 1.5 # BIAS > 1, TEs are MORE easily learned than unique words
-BIAS2 = 0.5 # BIAS < 1, TEs are LESS easily learned than unique words
+\hypertarget{study-2-empirical-data}{%
+\section{Study 2: Empirical data}\label{study-2-empirical-data}}
 
-### visualizing the simulation
-plot_simulation3 <- data_simulation3 %>%
-  mutate(TE_easier = BIAS1*(DOM*NONDOM/LEARNABLE), # calculate where BIAS > 1
-         TE_harder = BIAS2*(DOM*NONDOM/LEARNABLE)) %>% # calculate where BIAS < 1
-  # round to keep whole numbers only
-  mutate_at(vars(-c(BALANCE, bin_BALANCE)), round, 0) %>%
-  # prepare dataset to plot
-  pivot_longer(c(TE, TE_harder, TE_easier), 
-               names_to = "Model", values_to = "number_TE") %>%
-  mutate(Model = factor(Model, levels=c("TE_harder", "TE", "TE_easier"))) %>%
-  mutate(Model = recode(Model, 
-                        TE = "BIAS = 1 \n (Neutral Account)",
-                        TE_easier = "BIAS = 1.5 \n (Preference Account)",
-                        TE_harder = "BIAS = 0.5 \n (Avoidance Account)")) %>%
-  # ggplot
-  ggplot(aes(x = WORD, y = number_TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
-  stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
-  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
-  theme_minimal() + 
-  facet_grid(. ~ Model) +
-  theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom") +
-  theme(plot.title = element_text(hjust = 0.5),
-        text = element_text(size=11),
-        panel.spacing.x = unit(2.5, "lines"),
-        panel.background = element_rect(fill = NA, color = "black")) +
-  labs(x = "Word vocabulary (WORD)", 
-       y = "Number of translation equivalents (TE)", 
-       color = "Vocabulary balance (BALANCE)",
-       linetype = "Vocabulary balance (BALANCE)") +
-  guides(colour = guide_legend(reverse = T, nrow = 1),
-         linetype = guide_legend(reverse = T, nrow = 1))
-```
+In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
+In Study 2, we tested these predictions using archival data obtained from 200 French--English bilingual children aged 18 to 33 months.
 
-```{r fig4, fig.cap="Different scenarios of expected translation equivalents learning (TE) while assuming bias towards (BIAS < 1) or against learning translation equivalents (BIAS > 1), or no bias (BIAS = 1).", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%"}
+\hypertarget{method}{%
+\subsection{Method}\label{method}}
 
-plot_simulation3
-```
+Ethics approval was obtained by the Human Research Ethics Board of XYZ University (ethics protocol number 0000) and informed consent was obtained from the participants' parents.
 
-# Study 2: Empirical data
-In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
-In Study 2, we tested these predictions using archival data obtained from 200 French–English bilingual children aged 18 to 33 months.
+\hypertarget{participants}{%
+\subsubsection{Participants}\label{participants}}
 
-## Method
-Ethics approval was obtained by the Human Research Ethics Board of XYZ University (ethics protocol number 0000) and informed consent was obtained from the participants’ parents.
-
-### Participants
-```{r participants, include = FALSE}
-# how many distinct participants and data point we have
-keepers_ws_TE %>% 
-  summarize(N_babies = n_distinct(baby_id),
-            N_admin = n_distinct(ID_testdate))
-
-# how many girls and boys we have
-d_gender <- keepers_ws_TE %>%
-  group_by(gender) %>% 
-  distinct(baby_id, .keep.all=T) 
-  # %>% count()
-
-# number of participants who contributed multiple data points
-d_multiple <- keepers_ws_TE %>% 
-  count(baby_id) %>% 
-  filter(n>1)
-
-# checking which baby_id has multiple data
-n_duplicatedBabyID <- keepers_ws_TE %>%
-  filter(duplicated(baby_id)) %>%
-  arrange(baby_id)
-
-## percentage of english-dominant & french-dominant children
-# dominance defined by vocab size
-d_lang_dom <- keepers_ws_TE %>% 
-  distinct(ID_testdate, .keep_all = TRUE) %>%
-  mutate(Eng_Dom = ifelse(lang_dom_vocab == "English", 1, 0)) %>%
-  summarise(n = n(),
-            n_EngDom = sum(Eng_Dom),
-            percent_EngDom = n_EngDom/n * 100,
-            percent_FrDom = 100 - percent_EngDom)
-```
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
-(2020).
-Some children took part in more than one in-lab study (n = `r length(d_multiple$baby_id)`); thus, they contributed data at more than one time point.
+Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Some children took part in more than one in-lab study (n = 28); thus, they contributed data at more than one time point.
 This resulted in a larger number of datapoints relative to the number of unique participants.
-The total number of data points included in the analyses was `r length(keepers_ws_TE$ID_testdate)` (i.e., `r length(keepers_ws_TE$eng_cdi_filled[keepers_ws_TE$eng_cdi_filled == "Y"])` English and `r length(keepers_ws_TE$fre_cdi_filled[keepers_ws_TE$fre_cdi_filled == "Y"])` French CDI questionnaires).
-Participants were recruited through government birth lists, online ads, daycares, and infant-parent group activities (e.g., children’s library activities).
-Inclusion criteria were the following: full term-pregnancy ( i.e., > 36 weeks of gestation), normal birth weight (> 2500 grams), and absence of major medical conditions (i.e., meningitis).
+The total number of data points included in the analyses was 229 (i.e., 229 English and 229 French CDI questionnaires).
+Participants were recruited through government birth lists, online ads, daycares, and infant-parent group activities (e.g., children's library activities).
+Inclusion criteria were the following: full term-pregnancy ( i.e., \textgreater{} 36 weeks of gestation), normal birth weight (\textgreater{} 2500 grams), and absence of major medical conditions (i.e., meningitis).
 Only children who had complete data in both CDI forms (i.e., English and French) were retained for analysis.
-Bilingual children were defined as those exposed at least 25% of the time over the course of their lives globally to both English and French and with less than 10% of exposure to a third language.
+Bilingual children were defined as those exposed at least 25\% of the time over the course of their lives globally to both English and French and with less than 10\% of exposure to a third language.
 For children who participated more than once, their language exposure followed such criteria between visits.
-Following the approach in Study 1, children’s dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
-Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 1)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant.
+Following the approach in Study 1, children's dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
+Within the 229 data points, 59.80\% of children were English-dominant and 40.17\% were French-dominant.
 Data collection was conducted in Montréal, Québec, Canada.
 Montréal is a multicultural city where both English and French are widely used in society.
-Children’s demographic characteristics are presented in Table 4.
+Children's demographic characteristics are presented in Table 4.
 
-```{r table4, results="asis", echo = FALSE}
-# Create dataframe of demographics
-d_participants <- keepers_ws_TE %>%
-  # keep only relevant variables
-  select(ID_testdate, age_continuous, years_education, lang_exp_eng, lang_exp_fre, lang_exp_other) %>%
-  # rename variables
-  rename("Age in months" = age_continuous,
-         "Maternal education in years" = years_education,
-         "% Global exposure to English"  = lang_exp_eng,
-         "% Global exposure to French"  = lang_exp_fre, 
-         "% Global exposure to Other" = lang_exp_other) %>%
-  # summarize the data
-  pivot_longer(-c(ID_testdate), names_to = "info", values_to = "value") %>%
-  group_by(info) %>%
-  summarise(mean = mean(value, na.rm=T),
-            sd = sd(value, na.rm=T),
-            min = min(value, na.rm=T),
-            max = max(value, na.rm=T)) %>%
-  mutate(across(where(is.numeric),round, 1)) %>%
-  # combine min and max to form range
-  unite("range", min:max, sep = " - ", remove = TRUE, na.rm = FALSE) %>%
-  # arrange rows in custom order
-  mutate(info = factor(info, levels = c("Age in months", "Maternal education in years", 
-                                        "% Global exposure to English", "% Global exposure to French",
-                                        "% Global exposure to Other"))) %>%
-  arrange(info)
+\begin{table}[tbp]
 
-# generate the apa table
-papaja::apa_table(d_participants, format.args = list(digits = 1),
-                  caption = "Demographic characteristics of participants (data points = 229).",
-                  col.names =c("","Mean","$SD$","Range"),
-                  align=c("l","c","c","c"))
-```
+\begin{center}
+\begin{threeparttable}
 
-### Measures
+\caption{\label{tab:table4}Demographic characteristics of participants (data points = 229).}
 
-```{r cdi_filled_by, include = FALSE}
-# who filled out the CDIs
-d_cdi_filled_by <- keepers_ws_TE %>% 
-  # combine grandmother & other family member
-  mutate(cdi_filled_by = replace(cdi_filled_by, cdi_filled_by == "Grandmother", "Other family member")) %>% 
-  # count number of respondents
-  count(cdi_filled_by) %>%
-  mutate(percentage = n/sum(n)*100) # calculate percentages
-```
+\begin{tabular}{lccc}
+\toprule
+ & Mean & $SD$ & Range\\
+\midrule
+Age in months & 24.4 & 4.7 & 18.4 - 33.5\\
+Maternal education in years & 16.6 & 2.1 & 10 - 21\\
+\% Global exposure to English & 51.7 & 14.8 & 25 - 75\\
+\% Global exposure to French & 47.8 & 15.0 & 25 - 75\\
+\% Global exposure to Other & 0.6 & 1.8 & 0 - 10\\
+\bottomrule
+\end{tabular}
 
-* **MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI).** Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
+\end{threeparttable}
+\end{center}
+
+\end{table}
+
+\hypertarget{measures}{%
+\subsubsection{Measures}\label{measures}}
+
+\begin{itemize}
+\item
+  \textbf{MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI).} Bilingual children's expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
 Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64%), fathers (7%), both parents (4%), others (< 1%; e.g., grandmother), and respondent not indicated (25%).
+We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (25\%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
@@
 There are 680 words in the English CDI version and 664 in the Québec French version.
+\item
+  \textbf{Language Exposure Questionnaire using the MAPLE approach.}
+\end{itemize}
 
-* **Language Exposure Questionnaire using the MAPLE approach.**
+\hypertarget{procedure}{%
+\subsubsection{Procedure}\label{procedure}}
 
-### Procedure
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-## Results
-### Testing Prediction Set 1: 
-### Testing Prediction Set 2:
-### Testing Prediction Set 3:
+\hypertarget{testing-prediction-set-1}{%
+\subsubsection{Testing Prediction Set 1:}\label{testing-prediction-set-1}}
 
-# Discussion
+\hypertarget{testing-prediction-set-2}{%
+\subsubsection{Testing Prediction Set 2:}\label{testing-prediction-set-2}}
+
+\hypertarget{testing-prediction-set-3}{%
+\subsubsection{Testing Prediction Set 3:}\label{testing-prediction-set-3}}
+
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+\end{document}

### 2021-06-14 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated rmarkdown file

@@
   The acquisition of translation equivalents is often considered a special component of bilingual children’s vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
-This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are neither harder nor easier to learn.
+This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
 To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
@@
 library("ggpubr")
+library(broom)
+library(lsr)
+library(Hmisc)
 library(lme4)
@@
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are hard to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are neither harder nor easier to learn than singlets (Neutral Account).
-To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model as a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level.
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn than singlets (Neutral Account).
+To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level.
 In Study 1, we use the Bilingual Vocabulary Model to derive a set of predictions, which we then test against vocabulary data from 200 18- to 33-month-old bilingual children in Study 2.
@@
 
-Overall, converging evidence refutes the position that a strong form of mutual exclusivity operates on the developing bilingual lexicon to prevent bilinguals from acquiring translation equivalents.
+Overall, converging evidence refutes the position that a strong form of mutual exclusivity prevents bilinguals from acquiring translation equivalents.
 Nonetheless, it leaves open the possibility that translation equivalents may be less likely acquired in favour of learning singlets even if translation equivalents are not completely avoided.
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
+When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
 Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak & Byers-Heinlein, 2018; Singh, 2013).
 In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
-The Preference Account predicts that translation equivalents would be more easily learned than singlets.
+The Preference Account predicts that translation equivalents are more easily learned than singlets.
 
@@
 The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like “cat” and “dog” are strongly semantically related).
-Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in the lexicons.
+Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in children’s lexicons.
 This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English “dog” and French _“chien”_) and words that had other cross-language relations (e.g., “cat” and _“chien”_).
@@
 Comparison with actual bilingual children’s vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
-The authors therefore concluded that bilingual children show a preference for learning translation equivalents over singlets.
-Note that, in their study, translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word “cat”).
-However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning including vocabulary size in each language and the developmental level of a child — an issue that we will return to later in the introduction.
+The authors therefore concluded that bilingual children learn translation equivalents more easily than singlets.
+Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word “cat”).
+However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning including vocabulary size in each language and the developmental level of a child — a point that we will return to later in the introduction.
 
 Overall, there is some evidence that bilingual children more readily learn translation equivalents than singlets.
-If the strong semantic overlap between translation equivalents facilitates their learning, then under the Preference Account translation equivalents would be more easily learned than singlets.
+If the strong semantic overlap between translation equivalents facilitates their learning, then under the Preference Account translation equivalents will be more easily learned than singlets.
 
@@
 There is an underlying assumption that translation equivalent learning is somehow special relative to singlet learning — the Avoidance Account proposes that translation equivalents are harder to learn than singlets, whereas the Preference Account proposes that translation equivalents are easier to learn than singlets.
-However, it is also possible that translation equivalents are neither harder nor easier to learn than singlets.
+However, it is also possible that translation equivalents are neither harder nor easier for bilingual children to learn than singlets.
 We call this the Neutral Account.
@@
 The Neutral Account implies that bilingual children’s two languages develop relatively independently.
-Indeed, language and processing measures in bilinguals tend to be tightly correlated within a particular language, with only weak cross-language correlations.
+Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
 For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, & Hurtado, 2010).
 Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
-Bilingual children who heard a greater amount of a particular language tended to process that language faster, and in turn learned more words in that language.
-
-In a study whose results directly support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children.
-It was revealed that the percentage of translation equivalents observed in English–Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence suggested that the acquisition of translation equivalents is not that different from the expected overlap between two independent lexicons, implying that singlets and translation equivalents are equally learnable.
-In sum, the Neutral Account predicts that translation equivalents are neither harder nor easier for bilingual children to learn than singlets.
+Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
+
+In a study whose results support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children.
+The researchers found that the percentage of translation equivalents observed in English–Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
+This evidence implied that singlets and translation equivalents are equally learnable.
+In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
 
@@
 The previous section discussed three theoretical accounts concerning the relative learnability of translation equivalents.
-However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than as an integrated system.
-In this section, we consider two proximal variables that we expect to predict the number of translation equivalents children know: vocabulary size in each language, and word learnability as a function of children’s developmental level.
+However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than integrated within the larger context of bilingual lexical development.
+In this section, we consider two proximal variables that we expect to predict the number of translation equivalents bilingual children know: vocabulary size in each language, and word learnability as a function of children’s developmental level.
 
@@
 
-Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which appears to be tightly linked to their exposure to each language.
+Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
 In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, & Hurtado, 2010; Place & Hoff, 2011; Pearson et al., 1997).
-Bilingual children tend to know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
+Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
-However, the Complementarity Principle may not necessarily impact bilingual word acquisition in infancy.
+However, we argue that the Complementarity Principle is unlikely to strongly impact bilingual word acquisition in infancy.
 Most words in children’s early vocabularies could be considered “home words”, which include words for social and daily routines (“hello”, “more”, “diaper”), common nouns (“doggie”), and everyday verbs (“walk”).
@@
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
-For example, across many languages, including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
+For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
 Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007).
@@
 Taking into account the contributions of language exposure and developmental level to bilingual children’s vocabulary acquisition, we put forward the Bilingual Vocabulary Model.
-This model proposes that the number of translation equivalents that children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children’s developmental level.
+This model proposes that the number of translation equivalents that bilingual children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children’s developmental level.
 We formalize learning a translation equivalent pair as the joint probability of learning each of the words in the pair.
-This provides a clear empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
+This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
 The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B | A) where P(B | A) = P(B) if A and B are independent (see Box 1 for a detailed example).
@@
 
-&nbsp;
-
+\newpage
 ```{r box1, results="asis", fig.cap = " "}
@@
   column_spec(1, width = "450px") %>%
-  kable_styling(font_size = 10)
-
-```
-
-```{r fig1, fig.cap="Illustration of the Bilingual Vocabulary Model.", fig.align='center', out.width='100%'}
+  kable_styling(font_size = 9)
+
+```
+
+```{r fig1, fig.cap="Illustration of the Bilingual Vocabulary Model.", fig.align='center', out.width='90%'}
 
@@
 A more reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
-For example, Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
+For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
 Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
@@
 For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29.
-Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many words he produces are translation equivalents).
+Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many of the words he produces are translation equivalents).
 
@@
   Constraints = c("Varies by age. No greater than the number of words on CDI.",
-                  "DOM \u2265 NONDOM (i.e., always produce more words in dominant than non-dominant language); DOM \u2264 LEARNABLE",
-                  "NONDOM \u2264 DOM (always produce fewer words in non-dominant than dominant language); NONDOM \u2264 LEARNABLE",
+                  "DOM \u2265 NONDOM (children always produce more words in dominant than non-dominant language); DOM \u2264 LEARNABLE",
+                  "NONDOM \u2264 DOM (children always produce fewer words in non-dominant than dominant language); NONDOM \u2264 LEARNABLE",
                   "BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account",
-                  ".00 \u2264 BALANCE \u2264 .50 (greater values indicate children producing a more similar number of words in their two languages)",
-                  "W \u2264 2LEARNABLE (maximum word vocabulary is knowing each word in both languages)",
+                  "0 \u2264 BALANCE \u2264 .50 (greater values indicate children producing a more similar number of words in their two languages)",
+                  "W \u2264 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages)",
                   "", "", "", ""),
@@
                                          "NONDOM = BALANCE×WORD; NONDOM = WORD - DOM",
-                                         "Account = BIAS×(DOM×NONDOM/LEARNABLE)",
+                                         "",
                                          "BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)",
                                          "WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)",
-                                         "If DOM and NONDOM are independent, TE = DOM×NONDOM/LEARNABLE",
+                                         "TE = BIAS × DOM×NONDOM/LEARNABLE",
                                          "CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET",
@@
     pack_rows("Derived Parameters", 5, 10) %>%
-    column_spec(2:4, width = "175px")
+    column_spec(2:4, width = "175px") %>%
+    footnote(general = "All vocabulary measures are constrained to be integers.")
 )
@@
   BIAS = c("Constant at 1", "Constant at 1", 
-           "Varied at 0.5 (Avoidance Account), \n 1 (Neutral Account), \n and 1.5 (Preference Account)"),
+           "Varied at 0.5 (Avoidance Account), 1 (Neutral Account), and 1.5 (Preference Account)"),
   n_data = c(216, 161, 166)
@@
 
-## Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
-In Simulation 1, we first illustrate the mathematical relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
+## 1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
+In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
 For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
@@
 ## Definition column: the definition of each variable in the table
-Definition <- c("Bias parameter", "Learnable words", "Words produced in the dominant language", "Words produced in the non-dominant language",
+Definition <- c("Bias parameter", "Learnable words in each language", "Words produced in the dominant language", "Words produced in the non-dominant language",
                 "Word vocabulary (or total vocabulary size)", "Vocabulary balance", "Translation equivalents produced", 
@@
 Calculation <- c("", "", "", "", 
-                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM * NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
+                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM × NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
 
@@
 )
-
 ```
@@
 We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
+BIAS was once again constant at 1.
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
 Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 2, we plotted TE knowledge as predicted by DOM, NONDOM, and WORD at different levels of BALANCE.
-Across all three Panels (1A, 1B, and 1C), children with the most balanced vocabulary consistently produce more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increase, the number of TE also increases regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE are extremely tightly coupled.
-In sum, we observe three important patterns, which will serve as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
+In Figure 2, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
+Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
+Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
 
@@
 * Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE).
-* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge(see panel 1B of Figure 2).
+* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 2).
 
@@
 
-```{r fig2, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to total dominant vocabulary size (DOM; Panel A), total non-dominant vocabulary size (NONDOM; Panel B), and total word vocabulary (WORD; Panel C).
-Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600.
-Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%"}
+```{r fig2, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
+Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
+Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%", fig.pos = "H"}
 
@@
 
-## Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
-In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn the word for 600 different concepts (i.e., LEARNABLE held constant at 600).
-As laid out in the introduction, in the Bilingual Vocabulary Model the learnability of different words changes over time, where the possible number of words that can be learned increases as a child grows older.
-Therefore, in Simulation 2, we looked at the expected patterns of translation equivalent learning across varying levels of LEARNABLE (i.e., the number of learnable words as developmental level changes).
+## 1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
+In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE held constant at 600).
+As laid out in the introduction, under the Bilingual Vocabulary Model, the learnability of different words changes with a child’s developmental level, where LEARNABLE increases as a child grows older.
+Therefore, in Simulation 2, we looked at the expected patterns of translation equivalent learning across varying levels of LEARNABLE (i.e., the number of learnable words in each language as developmental level changes).
 Additionally, we further examined vocabulary composition by computing the number of singlets in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) language.
@@
 In total, data from 161 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
-Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM with the formula NONDOM/(DOM+NONDOM).
-We also counted the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children’s concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
-Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), medium balanced (BALANCE = .20 - .35), and least balanced (BALANCE = .00 - .02).
-
-In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT) as shown by the largest bars for the children at LEARNABLE = 600 compared to the shortest bars for children at LEARNABLE = 300.
-Moreover, we continue to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE), as evidenced by the green bars that are the highest for the most balanced children (BALANCE = .36 - .50) compared to the medium balanced (BALANCE = .21 - .35) and the least balanced children (BALANCE = .00 - .20).
-Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents.
-Across all three developmental levels of LEARNABLE, the light blue bars representing singlets in the dominant language (DOM-SINGLET) grow with less balanced exposure, whereas the dark blue bars representing singlets in the non-dominant language (NONDOM-SINGLET) grow with more balanced exposure.
-Overall, we observe 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
-Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
-
-* Prediction 2a: Have larger concept vocabularies (CONCEPT) 
-* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
-* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with a less balanced vocabulary (BALANCE) produce the most DOM-SINGLET.
-* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with a more balanced vocabulary (BALANCE) produce the most NONDOM-SINGLET.
+Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children’s concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
+Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), to medium balanced (BALANCE = .20 - .35), to least balanced (BALANCE = .00 - .02).
 
@@
 Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right).
-Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="80%"}
+Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="80%", fig.pos = "H"}
 
@@
 
-## Simulation 3: Bias towards or against translation equivalent learning compared to singlets
+In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+Moreover, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
+Overall, we generated 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
+Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
+
+* Prediction 2a: Have larger concept vocabularies (CONCEPT). 
+* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET.
+* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+
+## 1.3 Simulation 3: Bias towards or against translation equivalent learning compared to singlets
 In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
-In our final simulation, we examined models where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
+In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
 Mathematically, this requires varying the BIAS parameter.
@@
 
-Translation equivalent knowledge was first simulated across different developmental levels (as indicated by number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
+Translation equivalent (TE) knowledge was first simulated across different developmental levels (as indicated by number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
 Again, BALANCE and word vocabulary (WORD) were calculated based on the values of DOM and NONDOM.
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are neither easier nor harder to learn than singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equal to learn as singlets).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets).
-In Figure 4, we illustrate the three different scenarios of simulated translation equivalent knowledge.
-Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies produced more translation equivalents.
-Thus, overall relationships between vocabulary balance and translation equivalent knowledge remained similar across the Avoidance, Preference, and Neutral Accounts.
-What changed is the slope of translation equivalent learning: the slopes were the shallowest under the Avoidance Account where BIAS = 0.5, whereas the slopes are steepest under the Preference Account where BIAS = 1.5.
+In Figure 4, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
+Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
+Thus, overall relationships between BALANCE and TE remained similar across the Avoidance, Preference, and Neutral Accounts.
+What changed was the slope of translation equivalent learning: the slopes were the shallowest under the Avoidance Account where BIAS = 0.5, whereas the slopes were steepest under the Preference Account where BIAS = 1.5.
 With this, we further outline Prediction Set 3:
 
-* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning, with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn than singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
 
@@
 
-```{r fig4, fig.cap="Different scenarios of expected translation equivalents learning (TE) while assuming bias towards (BIAS < 1) or against learning translation equivalents (BIAS > 1), or no bias (BIAS = 1).", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%"}
+```{r fig4, fig.cap="Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%", fig.pos = "H"}
 
@@
 In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
-In Study 2, we tested these predictions using archival data obtained from 200 French–English bilingual children aged 18 to 33 months.
-
-## Method
-Ethics approval was obtained by the Human Research Ethics Board of XYZ University (ethics protocol number 0000) and informed consent was obtained from the participants’ parents.
-
-### Participants
+In Study 2, we tested these predictions using archival vocabulary data from 200 French–English bilingual children aged 18 to 33 months.
+
+## 2.1 Method
+Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certification Number 10000439) and informed consent was obtained from the children’s parents.
+
+## 2.1.1 Participants
 ```{r participants, include = FALSE}
@@
 ```
+
 Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
@@
 Bilingual children were defined as those exposed at least 25% of the time over the course of their lives globally to both English and French and with less than 10% of exposure to a third language.
-For children who participated more than once, their language exposure followed such criteria between visits.
+For children who participated more than once, their language exposure followed such criteria for all visits.
 Following the approach in Study 1, children’s dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
-Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 1)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant.
+Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 2)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant.
 Data collection was conducted in Montréal, Québec, Canada.
 Montréal is a multicultural city where both English and French are widely used in society.
-Children’s demographic characteristics are presented in Table 4.
-
-```{r table4, results="asis", echo = FALSE}
+Children’s demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
+
+```{r table4, results="asis"}
 # Create dataframe of demographics
@@
 
-### Measures
+## 2.1.2 Measures
 
@@
 
-* **MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI).** Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
+### MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)
+Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
 Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64%), fathers (7%), both parents (4%), others (< 1%; e.g., grandmother), and respondent not indicated (25%).
+We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), and respondent not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
@@
 
-* **Language Exposure Questionnaire using the MAPLE approach.**
-
-### Procedure
-
-## Results
-### Testing Prediction Set 1: 
-### Testing Prediction Set 2:
-### Testing Prediction Set 3:
+Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero et al.
+(2020) by three proficient bilingual French–English adults who carefully examined each language version of the CDI.
+Word pairs that made reference to the same concept (e.g., English “apple” and French _“pomme”_) were considered to be translation equivalents.
+In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
+Words that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were also considered translation equivalents.
+Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
+A full list of translation equivalents is available at [https://osf.io/7fz6c/].
+
+After determining the dominant language of a child based on the vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
+Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).
+
+### Language Exposure Questionnaire using the MAPLE approach
+Children’s language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE, Byers-Heinlein et al., 2018).
+The LEQ is a structured interview that lasts approximately 15 minutes.
+It includes targeted questions that quantify the child’s language exposure from birth until their current age.
+The LEQ and MAPLE provide a global language exposure estimate based on the number of hours the child is exposed to each language within all contexts (e.g., home, daycare, etc.).
+Children’s average global exposure to each language is described in Table 4.
+
+## 2.1.3 Procedure
+Caregivers were asked to fill out the CDI questionnaires as part of their child’s participation in experimental studies on language development, speech perception, and word learning.
+Caregivers were instructed to check off the words produced by their child using either a CDI paper questionnaire or the same questionnaire administered on a tablet.
+Data from paper based questionnaires was double entered and checked by trained research assistants.
+
+## 2.2 Results
+Data analyses were conducted using R (Version 4.0.2, 2020).
+Analysis scripts and the data set used in the present study are available at [https://osf.io/2t5kw/].
+We first present descriptive measures of vocabulary, and then tests of the three sets of predictions generated in Study 1.
+
+## 2.2.1 Descriptive measures of vocabulary
+
+```{r vocabulary_descriptives, include = FALSE}
+# average vocabulary scores
+vocabulary_average <- keepers_ws_TE %>%
+  select(ID_testdate, word_vocab, total_words_dom, total_words_nondom, concept_vocab,
+         number_of_te, total_singlet_dom, total_singlet_nondom) %>%
+  pivot_longer(-ID_testdate, names_to = "vocabulary", values_to = "number") %>%
+  group_by(vocabulary) %>%
+  summarise(mean = mean(number, na.rm = T), 
+            sd = sd(number, na.rm = T),
+            min = min(number, na.rm = T),
+            max = max(number, na.rm = T)) %>%
+  mutate(across(where(is.numeric), round, 1))
+
+## paired-sample t-test to compare DOM and NONDOM
+t_test_DOM_NONDOM <- tidy(t.test(keepers_ws_TE$total_words_dom, keepers_ws_TE$total_words_nondom, paired = TRUE))
+
+d_DOM_NONDOM <- cohensD(keepers_ws_TE$total_words_dom, keepers_ws_TE$total_words_nondom, method = "paired") # cohen's d effect size
+
+## paired-sample t-test to compare DOM-SINGLET and NONDOM-SINGLET
+t_test_SINGLET <- tidy(t.test(keepers_ws_TE$total_singlet_dom, keepers_ws_TE$total_singlet_nondom, paired = TRUE))
+
+d_SINGLET <- cohensD(keepers_ws_TE$total_singlet_dom, keepers_ws_TE$total_singlet_nondom, method = "paired")  # cohen's d effect size
+```
+
+On average, bilinguals in the sample had a mean word vocabulary size (WORD) of `r vocabulary_average$mean[7]` (SD = `r vocabulary_average$sd[7]`), with a wide range of `r vocabulary_average$min[7]` - `r vocabulary_average$max[7]` words.
+As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = `r vocabulary_average$mean[5]`, SD = `r vocabulary_average$sd[5]`, range = `r vocabulary_average$min[5]` - `r vocabulary_average$max[5]`) than in their non-dominant language (NONDOM; M = `r vocabulary_average$mean[6]`, SD = `r vocabulary_average$sd[6]`, range = `r vocabulary_average$min[6]` - `r vocabulary_average$max[6]`), _t_(`r t_test_DOM_NONDOM$parameter[1]`) = `r t_test_DOM_NONDOM$statistic[1]`, $p `r papaja::printp(t_test_DOM_NONDOM$p.value[1])`$, $d = `r round(d_DOM_NONDOM, 2)`$.
+
+Children produced an average of `r vocabulary_average$mean[2]` translation equivalents (TE; SD = `r vocabulary_average$sd[2]`, range = `r vocabulary_average$min[2]` - `r vocabulary_average$max[2]`).
+The remainder of words were singlets: Children produced many more singlets in their dominant language (DOM-SINGLET; M = `r vocabulary_average$mean[3]`, SD = `r vocabulary_average$sd[3]`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r vocabulary_average$mean[4]`, SD = `r vocabulary_average$sd[4]`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
+On average, children’s concept vocabulary size was `r vocabulary_average$mean[1]` (CONCEPT; SD = `r vocabulary_average$sd[1]`, range = `r vocabulary_average$min[1]` - `r vocabulary_average$max[1]`).
+
+```{r balance_descriptives, include = FALSE}
+# Average BALANCE score (defined by vocabulary balance)
+balance_average <- keepers_ws_TE %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+
+# Average BALANCE score grouped by language-dominance
+balance_average_group <- keepers_ws_TE %>%
+  group_by(lang_dom_vocab) %>%
+  summarise(mean = mean(balance_vocab, na.rm = T),
+            sd = sd(balance_vocab, na.rm = T),
+            min = min(balance_vocab, na.rm = T),
+            max = max(balance_vocab, na.rm = T))
+
+## paired-sample t-test to compare balance score between English-dominant vs. French-dominant children 
+t_test_balance <- tidy(t.test(balance_vocab ~ lang_dom_vocab, keepers_ws_TE))
+
+d_balance <- cohensD(balance_vocab ~ lang_dom_vocab, data = keepers_ws_TE, method = "pooled") # cohen's d effect size
+
+# vocabulary balance vs. input balance
+
+## check to see how many children have consistent/inconsistent dominant language between vocabulary-defined and input-defined dominance
+dominance_consistency <- keepers_ws_TE %>%
+  mutate(consistent_lang_dom = if_else(lang_dom != lang_dom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
+  summarise(n = n(),
+            n_consistent = sum(consistent_lang_dom),
+            percentage_consistent = n_consistent/n*100,
+            n_inconsistent = n - n_consistent,
+            percentage_inconsistent = n_inconsistent/n*100)
+
+## correlation between vocabulary balance and input balance (raw language exposure to the non-dominant language)
+correlation_balance <- tidy(cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$lang_nondom_input, method = "pearson"))
+
+```
+
+Vocabulary balance (BALANCE) was then determined based on the proportion of total words produced in the non-dominant language following the formula BALANCE = NONDOM/WORD as in Study 1.
+On average, bilingual children in our sample had a balance score BALANCE of `r balance_average$mean[1]` (SD = `r balance_average$sd[1]`), ranging from `r balance_average$min[1]` to `r balance_average$max[1]`.
+Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
+The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 2)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
+
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, but for young bilinguals balance can also be considered in terms of input in each language.
+We therefore compare the vocabulary balance with the proportion of exposure bilingual children received in their non-dominant language.
+To make values comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance, rather than the language that children heard most and least often.
+For most children, the language in which they produced the most words was also the language that they heard most often (`r dominance_consistency$n_consistent` children, `r round(dominance_consistency$percentage_consistent,2)`%), although this was not the case for some children (`r dominance_consistency$n_inconsistent` children, `r round(dominance_consistency$percentage_inconsistent,2)`%).
+The correlation between vocabulary-defined BALANCE and the raw percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 5).
+Thus, these two constructs were related, although not identical.
+
+&nbsp;
+```{r fig5, fig.cap="Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.", echo=FALSE, dpi=600, fig.align='center', out.width="75%", fig.pos = "H"}
+
+# plotting the correlation between balance by vocabulary and balance by input (raw language exposure to the non-dominant language)
+keepers_ws_TE %>%
+  ggplot(aes(x = balance_vocab, y = lang_nondom_input)) +
+  stat_smooth(method = lm, se = F, color = "black") +
+  geom_point(shape = 1) + 
+  theme_light() + 
+  labs(x = "Balance based on vocabulary (BALANCE)", 
+       y = "Balance based on exposure") +
+  theme(text = element_text(size=16))
+
+```
+
+## 2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures
+Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson’s correlations.
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations).
+
+```{r testing_prediction1, include = FALSE}
+
+# correlation between TE and BALANCE
+cor_TE_balance <- tidy(cor.test(keepers_ws_TE$number_of_te, keepers_ws_TE$balance_vocab, method = "pearson"))
+
+# one-way ANOVA to compare TE among the 5 BALANCE groups
+
+keepers_ws_TE <- keepers_ws_TE %>% # divide children into 5 balance subset groups (0 < BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5)
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))
+
+aov_TE_binBALANCE <- tidy(keepers_ws_TE %>% # one-way ANOVA
+                            aov(number_of_te ~ bin_BALANCE, data = .))
+
+# correlation between TE and WORD
+cor_TE_word <- tidy(cor.test(keepers_ws_TE$number_of_te, keepers_ws_TE$word_vocab, method = "pearson"))
+
+# correlation between TE and DOM
+cor_TE_dom <- tidy(cor.test(keepers_ws_TE$number_of_te, keepers_ws_TE$total_words_dom, method = "pearson"))
+
+# correlation between TE and NONDOM
+cor_TE_nondom <- tidy(cor.test(keepers_ws_TE$number_of_te, keepers_ws_TE$total_words_nondom, method = "pearson"))
+```
+
+Prediction 1a was that children with more balanced vocabularies would produce more translation equivalents.
+As shown in Figure 2 Row 2, our vocabulary data confirmed the prediction, _r_(`r cor_TE_balance$parameter[1]`) = `r round(cor_TE_balance$estimate[1],2)`, $p `r papaja::printp(cor_TE_balance$p.value[1])`$, where children with the most balanced vocabulary produced the most translation equivalents.
+We further tested this prediction by dividing children into 5 balance subset groups (0 < BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, _F_(`r aov_TE_binBALANCE$df[1]`, `r aov_TE_binBALANCE$df[2]`) = `r aov_TE_binBALANCE$statistic[1]`, $p = `r papaja::printp(aov_TE_binBALANCE$p.value[1])`$.
+The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the least translation equivalents.
+Detailed descriptive statistics are reported in Table 6.
+
+Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): _r_(`r cor_TE_word$parameter[1]`) = `r round(cor_TE_word$estimate[1],2)`, $p `r papaja::printp(cor_TE_word$p.value[1])`$, and dominant-language vocabulary (DOM): _r_(`r cor_TE_dom$parameter[1]`) = `r round(cor_TE_dom$estimate[1],2)`, $p `r papaja::printp(cor_TE_dom$p.value[1])`$.
+Figure 2 Row 2 further illustrates these relationships observed in our dataset.
+
+Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
+As shown in Figure 2 Row 2, we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
+
+```{r table5, results = "asis"}
+# correlation among variables in the observed data
+
+## keep only relevant variables for the correlation analysis
+correlation_data <- keepers_ws_TE %>%
+  select(age_continuous, Average_90percentile, balance_vocab, word_vocab, total_words_dom, total_words_nondom,
+         number_of_te, total_singlet_dom, total_singlet_nondom, concept_vocab) %>%
+  # rename variables
+  rename("Age (in month)" = age_continuous,
+         LEARNABLE = Average_90percentile,
+         BALANCE = balance_vocab, 
+         WORD = word_vocab,
+         DOM = total_words_dom, 
+         NONDOM = total_words_nondom, 
+         TE = number_of_te, 
+         "DOM-SINGLET" = total_singlet_dom, 
+         "NONDOM-SINGLET" = total_singlet_nondom, 
+         CONCEPT = concept_vocab)
+
+## correlation analysis
+### corstars function to add significant * to the correlation matrix
+corstars <-function(x, method=c("pearson", "spearman"), removeTriangle=c("upper", "lower")){ # x = data
+    #Compute correlation matrix
+    require(Hmisc)
+    x <- as.matrix(x)
+    correlation_matrix<-rcorr(x, type=method[1])
+    R <- correlation_matrix$r # Matrix of correlation coeficients
+    p <- correlation_matrix$P # Matrix of p-value
+
+    ## Define notions for significance levels; spacing is important.
+    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
+
+    ## trunctuate the correlation matrix to two decimal
+    R <- format(round(cbind(rep(-1.11, ncol(x)), R), 2))[,-1]
+
+    ## build a new matrix that includes the correlations with their apropriate stars
+    Rnew <- matrix(paste(R, mystars, sep=""), ncol=ncol(x))
+    diag(Rnew) <- paste(diag(R), " ", sep="")
+    rownames(Rnew) <- colnames(x)
+    colnames(Rnew) <- paste(colnames(x), "", sep="")
+
+    ## remove upper triangle of correlation matrix
+    if(removeTriangle[1]=="upper"){
+      Rnew <- as.matrix(Rnew)
+      Rnew[upper.tri(Rnew, diag = TRUE)] <- ""
+      Rnew <- as.data.frame(Rnew)
+    }
+
+    ## remove lower triangle of correlation matrix
+    else if(removeTriangle[1]=="lower"){
+      Rnew <- as.matrix(Rnew)
+      Rnew[lower.tri(Rnew, diag = TRUE)] <- ""
+      Rnew <- as.data.frame(Rnew)
+    }
+
+    ## remove last column and return the correlation matrix
+    Rnew <- cbind(Rnew[1:length(Rnew)-1])
+    return(Rnew)
+}
+
+correlation_table <- corstars(correlation_data, "pearson", "lower") # generate correlation matrix with our data using corstars function
+
+# print table
+papaja::apa_table(correlation_table,
+                  caption = "Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).",
+                  note = "*** p < .001, ** p < .01, * p < .05.",
+                  align="l",
+                  landscape = TRUE,
+                  font_size = "scriptsize")
+```
+
+```{r table6, results="asis"}
+
+# TE descriptive statistics of the 5 balance group
+mean_TE_binBALANCE <- keepers_ws_TE %>% 
+  group_by(bin_BALANCE) %>%
+  summarise(n = n(),
+            mean_TE = mean(number_of_te, na.rm = T),
+            sd_TE = sd(number_of_te, na.rm = T),
+            min_TE = min(number_of_te, na.rm = T),
+            max_TE = max(number_of_te, na.rm = T)) %>%
+  mutate(across(!bin_BALANCE & !sd_TE, ~ round(.,0))) %>%
+  mutate_at(vars(sd_TE), ~ round(.,2)) %>%
+  # combine min and max to form range
+  unite("range", min_TE:max_TE, sep = " - ", remove = TRUE, na.rm = FALSE) %>%
+  # sort bin_BALANCE with descending order
+  arrange(desc(bin_BALANCE)) %>%
+  # rename each level of bin_BALANCE
+  mutate(bin_BALANCE = recode(bin_BALANCE,
+                              "0.5" = "0.4 < BALANCE \u2264 0.5",
+                              "0.4" = "0.3 < BALANCE \u2264 0.4",
+                              "0.3" = "0.2 < BALANCE \u2264 0.3",
+                              "0.2" = "0.1 < BALANCE \u2264 0.2",
+                              "0.1" = "0 < BALANCE \u2264 0.1")) %>%
+  # rename column
+  rename(" " = bin_BALANCE, "N" = n, "Mean" = mean_TE, "SD" = sd_TE)
+
+# print table
+#papaja::apa_table(mean_TE_binBALANCE,
+#                  caption = "Average number of translation equivalents (TE) produced by each balance group.",
+#                  col.names =c("","N","Mean","$SD$","range"),
+#                  align="c")
+
+kable(mean_TE_binBALANCE,
+      format="latex", booktabs=TRUE, align = "c",
+      caption = "Average number of translation equivalents (TE) produced by each balance group.") %>%
+  column_spec(1, width = "150px") %>%
+  column_spec(2:4, width = "50px") %>%
+  column_spec(5, width = "75px") %>%
+  kable_styling(position = "center")
+```
+
+## 2.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels
+Prediction Set 2 pertained to expected patterns of acquisition of translation equivalents and singlets for children of different developmental levels.
+In our data set, developmental level was approximated by children’s age.
+Figure 3 Panel B shows the concept vocabulary (CONCEPT) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+To illustrate the acquisition of translation equivalents and singlets at different developmental levels, we divided children into three age groups: younger children of 18–22 months, middle children of 23–27 months, and older children of 28–33 months.
+
+```{r testing_prediction2, include = FALSE}
+
+# testing prediction 2a: To compare CONCEPT among the 3 LEARNABLE groups in the stacked bar chart
+
+## correlation between age (in months) and CONCEPT
+cor_age_CONCEPT <- tidy(cor.test(keepers_ws_TE$age_continuous, keepers_ws_TE$concept_vocab, method = "pearson"))
+
+## one-way ANOVE to compare CONCEPT among the 3 LEARNABLE groups
+keepers_ws_TE_3LEARNABLEsubset <- keepers_ws_TE %>%
+  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
+  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
+                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))
+
+aov_age_CONCEPT <- tidy(aov(concept_vocab ~ LEARNABLE_subset, data = keepers_ws_TE_3LEARNABLEsubset))
+
+pairwise.t.test(keepers_ws_TE_3LEARNABLEsubset$concept_vocab, keepers_ws_TE_3LEARNABLEsubset$LEARNABLE_subset,
+                p.adjust.method = "BH")
+
+## mean CONCEPT by LEARNABLE subset
+mean_concept_LEARNABLEsubset <- keepers_ws_TE_3LEARNABLEsubset %>%
+  group_by(LEARNABLE_subset) %>%
+  summarise(mean_concept = mean(concept_vocab, na.rm = T),
+            sd_concept = sd(concept_vocab, na.rm = T),
+            min_concept = min(concept_vocab, na.rm = T),
+            max_concept = max(concept_vocab, na.rm = T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+# testing prediction 2b: To compare TE among the 3 LEARNABLE groups in the stacked bar chart
+
+## correlation between age (in months) and TE
+cor_age_TE <- tidy(cor.test(keepers_ws_TE$age_continuous, keepers_ws_TE$number_of_te, method = "pearson"))
+
+## one-way ANOVA to compare TE among LEARNABLE
+stacked_byLEARNABLE_wide <- keepers_ws_TE %>%
+  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
+         number_of_te, total_singlet_dom, total_singlet_nondom) %>%
+  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
+  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
+                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
+                              TRUE ~ "28-33 months")) %>%
+  # Subset balance level by 3 groups
+  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
+                               balance_vocab <= 0.20 ~ "Least Balanced",
+                               TRUE ~ "Medium Balanced"))
+
+aov_age_TE <- tidy(aov(number_of_te ~ LEARNABLE_subset, data = stacked_byLEARNABLE_wide))
+
+pairwise.t.test(stacked_byLEARNABLE_wide$number_of_te, stacked_byLEARNABLE_wide$LEARNABLE_subset,
+                 p.adjust.method = "BH")
+
+## mean TE by LEARNABLE subset
+mean_TE_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
+  group_by(LEARNABLE_subset) %>%
+  summarise(mean_TE = mean(number_of_te, na.rm=T), 
+            sd_TE = sd(number_of_te, na.rm=T),
+            min_TE = min(number_of_te, na.rm=T),
+            max_TE = max(number_of_te, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+# testing prediction 2c: To compare DOM-SINGLET among the 3 LEARNABLE groups in the stacked bar chart
+
+## correlation between DOM-SINGLET and age (in month)
+cor_age_DOMSING <- tidy(cor.test(keepers_ws_TE$age_continuous, keepers_ws_TE$total_singlet_dom, method = "pearson"))
+
+## correlation between DOM-SINGLET and BALANCE
+cor_BALANCE_DOMSING <- tidy(cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$total_singlet_dom, method = "pearson"))
+
+## one-way ANOVA to compare DOM-SINGLET among LEARNABLE
+aov_balance_DOMSING <- tidy(aov(total_singlet_dom ~ BalSubset, data = stacked_byLEARNABLE_wide))
+
+pairwise.t.test(stacked_byLEARNABLE_wide$total_singlet_dom, stacked_byLEARNABLE_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+## mean DOM-SINGLET by LEARNABLE subset
+mean_DOMSINGLET_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
+  group_by(BalSubset) %>%
+  summarise(mean_DOM_SINGLET = mean(total_singlet_dom, na.rm=T), 
+            sd_DOM_SINGLET = sd(total_singlet_dom, na.rm=T),
+            min_DOM_SINGLET = min(total_singlet_dom, na.rm=T),
+            max_DOM_SINGLET = max(total_singlet_dom, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+# testing Prediction 2d: To compare NONDOM-SINGLET among the 3 LEARNABLE groups in the stacked barchart
+
+## correlation between NONDOM-SINGLET and age (in month)
+cor_age_NONDOMSING <- tidy(cor.test(keepers_ws_TE$age_continuous, keepers_ws_TE$total_singlet_nondom, method = "pearson"))
+
+## correlation between NONDOM-SINGLET and BALANCE
+cor_BALANCE_NONDOMSING <- tidy(cor.test(keepers_ws_TE$balance_vocab, keepers_ws_TE$total_singlet_nondom, method = "pearson"))
+
+## one-way ANOVA to compare NONDOM-SINGLET among LEARNABLE
+aov_balance_NONDOMSING <- tidy(aov(total_singlet_nondom ~ BalSubset, data = stacked_byLEARNABLE_wide))
+
+pairwise.t.test(stacked_byLEARNABLE_wide$total_singlet_nondom, stacked_byLEARNABLE_wide$BalSubset,
+                 p.adjust.method = "BH")
+
+## mean NONDOM-SINGLET by LEARNABLE subset
+mean_NONDOMSINGLET_LEARNABLEsubset <-stacked_byLEARNABLE_wide %>% 
+  group_by(BalSubset) %>%
+  summarise(mean_NONDOM_SINGLET = mean(total_singlet_nondom, na.rm=T), 
+            sd_NONDOM_SINGLET = sd(total_singlet_nondom, na.rm=T),
+            min_NONDOM_SINGLET = min(total_singlet_nondom, na.rm=T),
+            max_NONDOM_SINGLET = max(total_singlet_nondom, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+```
+
+Prediction 2a was that older children (i.e., those at a later developmental level) would have larger concept vocabularies than younger children (i.e., those at an earlier developmental level).
+We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, _r_(`r cor_age_CONCEPT$parameter[1]`) = `r round(cor_age_CONCEPT$estimate[1],2)`, $p `r papaja::printp(cor_age_CONCEPT$p.value[1])`$, and therefore confirmed the prediction.
+This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, ($F$(`r aov_age_CONCEPT$df[1]`, `r aov_age_CONCEPT$df[2]`) = `r round(aov_age_CONCEPT$statistic[1],2)`, $p `r papaja::printp(aov_age_CONCEPT$p.value[1])`$).
+Older children of 28–33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[3],2)` ($p$s < .001), those middle children of 23–27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[2],2)`, and those younger children of 18–22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[1],2)` ($p$s < .001).
+
+Prediction 2b was that older children would produce more translation equivalents than younger children.
+First, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, _r_(`r cor_age_TE$parameter[1]`) = `r round(cor_age_TE$estimate[1],2)`, $p `r papaja::printp(cor_age_TE$p.value[1])`$, and therefore confirmed the prediction.
+In a one-way ANOVA with age group as factor, we further found that groups differed in how many translation equivalents they produced ($F$(`r aov_age_TE$df[1]`, `r aov_age_TE$df[2]`) = `r round(aov_age_TE$statistic[1],2)`, $p `r papaja::printp(aov_age_TE$p.value[1])`$).
+Younger children of 18–22 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[1],2)` translation equivalents, middle children of 23–27 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[2],2)` translation equivalents, and older children of 28–33 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[3],2)` translation equivalents ($p$s < .01).
+
+Prediction 2c was that both older children and those with the least balanced vocabularies (BALANCE) would produce more dominant-language singlets (DOM-SINGLET).
+This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_DOMSING$parameter[1]`) = `r round(cor_age_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_age_DOMSING$p.value[1])`$, and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), _r_(`r cor_BALANCE_DOMSING$parameter[1]`) = `r round(cor_BALANCE_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_DOMSING$p.value[1])`$.
+As shown in Figure 3 Panel B, children were divided into least balanced (range of balance: .00 - .20), medium balanced (range of balance: .20 - .35) and most balanced (range of balance: .35 - .50) groups (i.e., the same criteria as in Figure 5).
+In a one-way ANOVA with balance group as a between-subjects factor, we observed that the least balanced children produced the most singlets in their dominant language ($p$s < .001), with the least balanced, medium balanced, most balanced children producing respectively: `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[1],2)`, `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[2],2)`, and `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[3],2)` words in their dominant language ($F$(`r aov_balance_DOMSING$df[1]`, `r aov_balance_DOMSING$df[2]`) = `r round(aov_balance_DOMSING$statistic[1],2)`, $p `r papaja::printp(aov_balance_DOMSING$p.value[1])`$).
+
+Prediction 2d was that older children and those with the most balanced vocabularies (BALANCE) would produce more singlets in their non-dominant language.
+This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_NONDOMSING$parameter[1]`) = `r round(cor_age_NONDOMSING$estimate[1],2)`, $p = `r papaja::printp(cor_age_NONDOMSING$p.value[1])`$, and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), _r_(`r cor_BALANCE_NONDOMSING$parameter[1]`) = `r round(cor_BALANCE_NONDOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_NONDOMSING$p.value[1])`$.
+In a one-way ANOVA with balance group as a between-subjects factor, we confirmed that children who differed in how balanced their vocabulary knowledge was also differed in how many singlets they produced in their non-dominant language ($F$(`r aov_balance_NONDOMSING$df[1]`, `r aov_balance_NONDOMSING$df[2]`) = `r round(aov_balance_NONDOMSING$statistic[1],2)`, $p `r papaja::printp(aov_balance_NONDOMSING$p.value[1])`$).
+As shown in Figure 3 Panel B, we observed that children produced very few singlets in their non-dominant language, although the most balanced children produced the most singlets in their non-dominant language (mean of the most balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[3],2)` > mean of the medium balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[2],2)` > mean of the most balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[1],2)`; $p$s < .001).
+
+## 2.2.4 Testing Prediction Set 3: Rate of translation equivalent learning
+Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
+To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate BIAS parameter.
+
+First, we will walk through the parameters in this model.
+The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
+As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
+For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
+For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since the 90th percentile information was available only up to 30 months.
+
+```{r table7, results = "asis"}
+# read in the percentile file
+ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv")) %>%
+  # as the upper age limit of the CDI-WS is 30m, create a new row for age 30-33m
+  mutate(age_months_percentile = case_when(age_months_percentile == 30 ~ "30 - 33", 
+                                           TRUE ~ as.character(age_months_percentile))) %>%
+  # rename variables for presentation
+  rename("Age (months)" = age_months_percentile,
+         "Number of English words produced at 90th percentile" = EngWS_90percentile,
+         "Number of French words produced at 90th percentile" = FrWS_90percentile,
+         "Average (LEARNABLE)" = Average_90percentile) %>%
+  # round to 1 decimal place
+  mutate(across(where(is.numeric),round,1))
+
+kable(ws_percentile,
+        format = "latex", booktabs = TRUE, align = "c",
+        caption = "The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.") %>%
+  kableExtra::kable_styling(font_size = 10) %>%
+  column_spec(1, width = "75px") %>%
+  column_spec(2:4, width = "125px")
+
+```
+
+Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
+To reproduce the Bilingual Vocabulary Model’s formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included in the model (denoted in R by using a colon rather than an asterisk between the interacting predictors).
+Therefore, our final linear regression model equation was:
+
+_ObservedTE ~ 0 + Dominant vocabulary:Non-dominant vocabulary/90 percentile of CDI items._
+
+\noindent (In R language, the model was entered as:
+
+_ObservedTE * 90 percentile of CDI items ~ 0 + Dominant vocabulary:Non-dominant vocabulary_)
+
+\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was consistent with the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
+Otherwise, a coefficient less than 1 represents a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 represents a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
+
+```{r testing_prediction3, include = FALSE}
+
+# testing Prediction 3: linear regression model comparing simulated and observed TE
+
+## creating predicted TEs for simulation
+keepers_ws_TE_simulation <- keepers_ws_TE %>%
+  mutate(Predicted_TE_percentile = total_words_dom * total_words_nondom/Average_90percentile) %>%
+  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, total_words_dom, total_words_nondom, 
+         word_vocab, age_months_binned)
+
+## linear regression model
+test_prediction3_model <- keepers_ws_TE_simulation %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom, data = .)
+
+# checking outliers
+## calculating cook's distance
+check_outliers_model <- keepers_ws_TE_simulation %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom,
+     data = .)
+
+keepers_ws_TE_simulation$CD <- cooks.distance(check_outliers_model)
+
+number_of_outliers <- filter(keepers_ws_TE_simulation, CD > 0.4) # check number of outliers with cook's distance > 0.4
+
+## linear regression model without the outliers
+model_remove_outliers <- keepers_ws_TE_simulation %>%
+  filter(CD < 0.4) %>% 
+  lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom, 
+     data = .)
+
+# splitting the linear regression model for children with WORD < 300 and those with WORD > 300
+
+## linear regression model for those with WORD > 300
+percentile_model_more300w <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom, 
+     data = .)
+
+## linear regression model for those with WORD < 300
+percentile_model_less300w <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom, 
+     data = .)
+
+```
+
+Our model showed an excellent model fit of $R^2$ = `r round(summary(test_prediction3_model)$r.squared,2)`, indicating that our model explained `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of the variance in bilinguals’ translation equivalent knowledge.
+The linear regression model estimated a BIAS coefficient of `r round(summary(test_prediction3_model)$coefficient[1],2)`, $p `r papaja::printp(summary(test_prediction3_model)$coefficient[4])`$.
+This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words.
+
+To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child’s expected translation equivalent knowledge (setting BIAS = 1), which is plotted against our observed data in Figure 6.
+Expected and observed translation equivalents were closely aligned with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1), suggesting that the Neutral Account provides a parsimonious explanation for bilinguals’ translation equivalent knowledge.
+This provides evidence for the notion that translation equivalents are neither harder nor easier to learn than singlets in bilingual vocabulary learning.
+Note that visual inspection suggested that there could be some possible outliers.
+Cook’s distance was estimated for our linear regression model listed above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a cook’s distance over 0.4.
+After removing those `r words(length(number_of_outliers$ID_testdate))` data points, the linear regression model returned a coefficient of `r round(summary(model_remove_outliers)$coefficient[1],2)`, $p `r papaja::printp(summary(model_remove_outliers)$coefficient[4])`$, with $R^2$ = `r round(summary(model_remove_outliers)$r.squared,2)`.
+As the model fit was similar to the model without eliminating the `r words(length(number_of_outliers$ID_testdate))` outlier data points, we proceeded with the full data set keeping the `r words(length(number_of_outliers$ID_testdate))` potential outlier data points.
+
+```{r fig6, fig.cap="The number of simulated and observed translation equivalents plotted against each other.
+The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
+The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="70%", out.width="70%"}
+
+# ploting Y = observed rawTE  & X = Predicted rawTE 
+keepers_ws_TE_simulation %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  xlim(-10,450) +
+  ylim(-10,450) +
+  theme_minimal() + 
+  theme(text = element_text(size=16))
+
+```
+
+Despite the good overall fit to the data, a close examination of Figure 6 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
+Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1) for children with more than 300 total vocabulary.
+To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
+The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_more300w)$coefficient[1],2)`, $p `r papaja::printp(summary(percentile_model_more300w)$coefficient[4])`$, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_less300w)$coefficient[1],2)`, $p `r papaja::printp(summary(percentile_model_less300w)$coefficient[4])`$.
+Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size ($R^2$ = `r round(summary(percentile_model_more300w)$r.squared,2)`) than children with smaller vocabulary size ($R^2$ = `r round(summary(percentile_model_less300w)$r.squared,2)`).
+Overall, this analysis suggests that  translation equivalent learning for children with larger vocabularies corresponds best to the Neutral Account, but translation equivalent learning for children with smaller vocabularies corresponds best to the Preference Account.
+
+```{r fig7, fig.cap="The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
+The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%"}
+
+# plotting the two separate models (those with WORD < 300 + those with WORD > 300)
+TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
+  mutate(W_group = if_else(word_vocab < 300, "Less than 300 word vocabulary (WORD)", "More than 300 word vocabulary (WORD)")) %>%
+  filter(W_group == "Less than 300 word vocabulary (WORD)") %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
+       title = "Less than 300 word vocabulary (WORD)") +
+  xlim(0,80) +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 11, hjust = 0.5),
+  text = element_text(size=11))
+
+TEprediction_percentile_facet_moreW300 <- keepers_ws_TE_simulation %>%
+  mutate(W_group = if_else(word_vocab < 300, "Less than 300 word vocabulary (WORD)", "More than 300 word vocabulary (WORD)")) %>%
+  filter(W_group == "More than 300 word vocabulary (WORD)") %>%
+  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
+  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs",
+       title = "More than 300 word vocabulary (WORD)") +
+  geom_smooth(method = "lm", se = F) +
+  geom_point(shape = 23) +
+  geom_abline(linetype = 2, alpha = 0.5) +
+  theme_minimal() +
+  theme(plot.title = element_text(size = 11, hjust = 0.5),
+        text = element_text(size=11))
+
+TEprediction_percentile_facetByW300 <- 
+  ggarrange(TEprediction_percentile_facet_lessW300 + rremove("x.title"), 
+            TEprediction_percentile_facet_moreW300 + rremove("xy.title"), 
+            ncol = 2, nrow = 1) %>%
+  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model"))
+
+TEprediction_percentile_facetByW300
+```
 
 # Discussion
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn than singlet words (i.e., the first label for a particular referent).
+To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were available to be learned given the child’s age.
+The relative difficulty of learning translation equivalents relative to singlets was modeled via the bias parameter (BIAS), which indicated whether translation equivalent learning is consistent with the Avoidance (BIAS < 1), Preference (BIAS > 1), or Neutral Account (BIAS = 1).
+
+## Confirmation of model predictions
+In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18-33 months.
+Three sets of model predictions were confirmed in our empirical dataset.
+
+Prediction Set 1 pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
+In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
+This pattern is consistent with reports from previous research (Davidson & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
+This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Poulin-Dubois et al., 2013; Legacy et al., 2015).
+Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
+A similar pattern has been reported by Legacy and colleagues (2015), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2015).
+
+Prediction Set 2 pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+We operationalized developmental level in terms of children’s age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
+Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
+Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
+This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David & Wei, 2008; Legacy et al., 2016).
+Predictions 2c and 2d were also confirmed by our vocabulary data.
+While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Place & Hoff, 2011; Pearson et al., 1997).
+Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
+
+Overall, we observed a strong correspondence between the data simulated under the Bilingual Vocabulary Model and our observed data.
+Moreover, our model predicted numerous disparate patterns that have been previously reported in the literature.
+
+Having validated our overall approach, Prediction 3 motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents learning are similar to learn than singlets), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
+However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
+Modeling their data separately, we found evidence for the Preference account: younger children at around `r round(keepers_ws_TE %>% filter(word_vocab < 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months appeared to learn translation equivalents more easily than singlets, whereas older children at around `r round(keepers_ws_TE %>% filter(word_vocab > 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months learned translation equivalents similarly to singlets.
+This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
+This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
+The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
+Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
+The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children’s own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
+
+## Developmental change in translation equivalent learning
+The developmental change of bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman & Bowman, 1989).
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman & Bowman, 1989).
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au & Glusman, 1990, David & Tell, 2005).
+At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
+Yet, this explanation would not predict that younger bilinguals’ data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance account, before perhaps returning to the Neutral account once children realize that each referent should have a label in each language.
+Thus, changes in one-to-one mapping biases do not provide a complete explanation for our results.
+
+Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
+For example, a caregiver may choose to say to their English–French bilingual child “Can you grab the livre?” if they know their child understands the French word “livre” but not the English equivalent “book”.
+This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
+Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
+Overall, more research will be needed to understand why translation equivalents appear to be over-represented in younger bilinguals’ vocabularies.
+
+## Assumptions, limitations, and future directions
+Our Bilingual Vocabulary Model presented an integrated computational account of translation equivalent learning, focusing on the joint probability of learning the word for a concept in each language.
+To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
+However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Hoff, 2003; Fernald, Marchman, & Weisleder, 2013), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020, Rowe, 2012).
+It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+Such a model may better characterize and predict bilingual vocabulary development as a function of experience.
+Moreover, it would be important to extend our Bilingual Vocabulary Model to longitudinal data or data of a different bilingual population to investigate if it is possible to replicate the qualitative shift where bilingual children’s ability to learn translation equivalents appears to change across development.
+
+Another limitation of our model is that it takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
+However, the reality of bilingual experience might be more complex.
+First, some concepts expressed as a single word in one language may be lexicalized by two words in another language (e.g., English has a single word for “sister” but Mandarin has separate words for _“jiějie”_ [older sister] and _“mèimei”_ [younger sister]).
+As another example, some words may not have a translation equivalent in the other language (e.g., the Japanese word _“sushi”_ is borrowed into other languages).
+Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a “bowl” in English but an _“assiette”_ [plate] in French).
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
+Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) (Bosch & Ramon-Casas, 2014).
+Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
+While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
+Moreover, if they do prove to be important, such factors could be added to future iterations of the Bilingual Vocabulary Model.
+
+Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
+However, following the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., “soap”, “bath”, “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2016; Tardif et al., 2008).
+For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and simple nouns (“ball”, “dog”) are among the first words children across languages and cultures.
+It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
+Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
+
+Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the bias parameter (BIAS) and the parameter that accounts for how many words are potentially learnable at a particular age (LEARNABLE).
+Under the Bilingual Vocabulary Model, the learnability parameter and the bias parameter jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter changes by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also change by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
+Our model estimated the number of learnable words to be the number that children at the 90th percentile at a particular age produce.
+Small changes to this approach (e.g., taking the number of words children at the 95th percentile produce) would likely not drastically alter our results, nor change the qualitative shift that we observed in our data.
+Nonetheless, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+
+## Conclusions
+In sum, the acquisition of translation equivalents has been considered a special component in bilingual children’ vocabulary development.
+Previous research has put forward three diverging accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+We proposed the Bilingual Vocabulary Model, which provides a quantitative way to test these accounts, by modeling translation equivalent learning in relation to vocabulary size in each language and the number of potentially learnable words, which is constrained by children’s developmental level.
+Results using archival data from a large number of young French–English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
+Moreover, our model parsimoniously explained previously disparate observations about bilingual children’s translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.

### 2021-06-14 paper/paper_TE_bilingual_vocabulary_model.tex
updated rmarkdown file

@@
 The acquisition of translation equivalents is often considered a special component of bilingual children's vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
-This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are neither harder nor easier to learn.
+This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
 To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model --- the Bilingual Vocabulary Model --- which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
@@
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are hard to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are neither harder nor easier to learn than singlets (Neutral Account).
-To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model as a computational account of vocabulary learning, with parameters including bilinguals' vocabulary in each language and their developmental level.
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn than singlets (Neutral Account).
+To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals' vocabulary in each language and their developmental level.
 In Study 1, we use the Bilingual Vocabulary Model to derive a set of predictions, which we then test against vocabulary data from 200 18- to 33-month-old bilingual children in Study 2.
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
+When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
 Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak \& Byers-Heinlein, 2018; Singh, 2013).
 In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
-The Preference Account predicts that translation equivalents would be more easily learned than singlets.
+The Preference Account predicts that translation equivalents are more easily learned than singlets.
 
@@
 The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related).
-Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in the lexicons.
+Using a statistical model that allowed free semantic relations among vocabulary data from monolingual and bilingual children, the results suggested that words were learned faster when they were semantically connected to more known words in children's lexicons.
 This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English ``dog'' and French \emph{``chien''}) and words that had other cross-language relations (e.g., ``cat'' and \emph{``chien''}).
@@
 Comparison with actual bilingual children's vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
-The authors therefore concluded that bilingual children show a preference for learning translation equivalents over singlets.
-Note that, in their study, translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat'').
-However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning including vocabulary size in each language and the developmental level of a child --- an issue that we will return to later in the introduction.
+The authors therefore concluded that bilingual children learn translation equivalents more easily than singlets.
+Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat'').
+However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning including vocabulary size in each language and the developmental level of a child --- a point that we will return to later in the introduction.
 
 Overall, there is some evidence that bilingual children more readily learn translation equivalents than singlets.
-If the strong semantic overlap between translation equivalents facilitates their learning, then under the Preference Account translation equivalents would be more easily learned than singlets.
+If the strong semantic overlap between translation equivalents facilitates their learning, then under the Preference Account translation equivalents will be more easily learned than singlets.
 
@@
 There is an underlying assumption that translation equivalent learning is somehow special relative to singlet learning --- the Avoidance Account proposes that translation equivalents are harder to learn than singlets, whereas the Preference Account proposes that translation equivalents are easier to learn than singlets.
-However, it is also possible that translation equivalents are neither harder nor easier to learn than singlets.
+However, it is also possible that translation equivalents are neither harder nor easier for bilingual children to learn than singlets.
 We call this the Neutral Account.
@@
 The Neutral Account implies that bilingual children's two languages develop relatively independently.
-Indeed, language and processing measures in bilinguals tend to be tightly correlated within a particular language, with only weak cross-language correlations.
+Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
 For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, \& Hurtado, 2010).
 Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
-Bilingual children who heard a greater amount of a particular language tended to process that language faster, and in turn learned more words in that language.
+Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
 
-In a study whose results directly support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children.
-It was revealed that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence suggested that the acquisition of translation equivalents is not that different from the expected overlap between two independent lexicons, implying that singlets and translation equivalents are equally learnable.
-In sum, the Neutral Account predicts that translation equivalents are neither harder nor easier for bilingual children to learn than singlets.
+In a study whose results support the Neutral Account, Pearson and colleagues (1995) randomly paired the single-language English lexicons from a subset of bilingual children to the single-language Spanish lexicons from another subset of bilingual children to derive a percentage of by-chance lexical overlaps shared between monolingual lexicons of two randomly paired children.
+The researchers found that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
+This evidence implied that singlets and translation equivalents are equally learnable.
+In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
 
@@
 The previous section discussed three theoretical accounts concerning the relative learnability of translation equivalents.
-However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than as an integrated system.
-In this section, we consider two proximal variables that we expect to predict the number of translation equivalents children know: vocabulary size in each language, and word learnability as a function of children's developmental level.
+However, to date, aspects of translation equivalent learning have mostly been examined in isolation, rather than integrated within the larger context of bilingual lexical development.
+In this section, we consider two proximal variables that we expect to predict the number of translation equivalents bilingual children know: vocabulary size in each language, and word learnability as a function of children's developmental level.
 
@@
 
-Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which appears to be tightly linked to their exposure to each language.
+Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
 In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, \& Hurtado, 2010; Place \& Hoff, 2011; Pearson et al., 1997).
-Bilingual children tend to know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
+Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
-However, the Complementarity Principle may not necessarily impact bilingual word acquisition in infancy.
+However, we argue that the Complementarity Principle is unlikely to strongly impact bilingual word acquisition in infancy.
 Most words in children's early vocabularies could be considered ``home words,'' which include words for social and daily routines (``hello,'' ``more,'' ``diaper''), common nouns (``doggie''), and everyday verbs (``walk'').
@@
 Taking into account the contributions of language exposure and developmental level to bilingual children's vocabulary acquisition, we put forward the Bilingual Vocabulary Model.
-This model proposes that the number of translation equivalents that children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children's developmental level.
+This model proposes that the number of translation equivalents that bilingual children produce is a function of vocabulary learning in each language, in the context of the number of potentially learnable words given the children's developmental level.
 We formalize learning a translation equivalent pair as the joint probability of learning each of the words in the pair.
-This provides a clear empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
+This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
 The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B \textbar{} A) where P(B \textbar{} A) = P(B) if A and B are independent (see Box 1 for a detailed example).
@@
 
-~
-
-\begingroup\fontsize{10}{12}\selectfont
+\newpage
+\begingroup\fontsize{9}{11}\selectfont
 
@@
 
-{\centering \includegraphics[width=1\linewidth]{./Figure1_illustration}
+{\centering \includegraphics[width=0.9\linewidth]{./Figure1_illustration}
 
@@
 A more reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
-For example, Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
+For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
 Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
@@
 & Maximum number that could be learned in DOM or NONDOM\\
-\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (i.e., always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM\\
-\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD - DOM\\
-\hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & Account = BIAS×(DOM×NONDOM/LEARNABLE)\\
+\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (children always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (children always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD - DOM\\
+\hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & \\
 \addlinespace[0.3em]
 \multicolumn{4}{l}{\textbf{Derived Parameters}}\\
-\hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & .00 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
-\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)\\
-\hspace{1em}TE & Translation equivalents produced &  & If DOM and NONDOM are independent, TE = DOM×NONDOM/LEARNABLE\\
+\hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & 0 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)\\
+\hspace{1em}TE & Translation equivalents produced &  & TE = BIAS × DOM×NONDOM/LEARNABLE\\
 \hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
@@
 \bottomrule
+\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
+\multicolumn{4}{l}{\rule{0pt}{1em}All vocabulary measures are constrained to be integers.}\\
 \end{tabular}
@@
 2 & Varied at 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 161\\
-3 & Varied at 150, 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Varied at 0.5 (Avoidance Account),
- 1 (Neutral Account), 
- and 1.5 (Preference Account) & 166\\
+3 & Varied at 150, 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Varied at 0.5 (Avoidance Account), 1 (Neutral Account), and 1.5 (Preference Account) & 166\\
 \bottomrule
@@
 \hypertarget{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}{%
-\subsection{Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}
+\subsection{1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}
 
-In Simulation 1, we first illustrate the mathematical relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
+In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
 For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
@@
 \hspace{1em}BIAS & Bias parameter &  & 1.0 & 1.0 & 1.0\\
-\hspace{1em}LEARNABLE & Learnable words &  & 600.0 & 600.0 & 600.0\\
+\hspace{1em}LEARNABLE & Learnable words in each language &  & 600.0 & 600.0 & 600.0\\
 \hspace{1em}DOM & Words produced in the dominant language &  & 270.0 & 180.0 & 540.0\\
@@
 \hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0.1 & 0.4 & 0.4\\
-\hspace{1em}TE & Translation equivalents produced & DOM * NONDOM / LEARNABLE & 13.5 & 36.0 & 324.0\\
+\hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 13.5 & 36.0 & 324.0\\
 \hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286.5 & 264.0 & 576.0\\
@@
 We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
+BIAS was once again constant at 1.
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
 Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 2, we plotted TE knowledge as predicted by DOM, NONDOM, and WORD at different levels of BALANCE.
-Across all three Panels (1A, 1B, and 1C), children with the most balanced vocabulary consistently produce more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increase, the number of TE also increases regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE are extremely tightly coupled.
-In sum, we observe three important patterns, which will serve as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
+In Figure 2, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
+Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
+Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
 
@@
 \item
-  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge(see panel 1B of Figure 2).
+  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 2).
 \end{itemize}
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to total dominant vocabulary size (DOM; Panel A), total non-dominant vocabulary size (NONDOM; Panel B), and total word vocabulary (WORD; Panel C).
-Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600.
+\caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
+Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
 Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig2}
@@
 \hypertarget{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}{%
-\subsection{Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}
+\subsection{1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}
 
-In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn the word for 600 different concepts (i.e., LEARNABLE held constant at 600).
-As laid out in the introduction, in the Bilingual Vocabulary Model the learnability of different words changes over time, where the possible number of words that can be learned increases as a child grows older.
-Therefore, in Simulation 2, we looked at the expected patterns of translation equivalent learning across varying levels of LEARNABLE (i.e., the number of learnable words as developmental level changes).
+In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE held constant at 600).
+As laid out in the introduction, under the Bilingual Vocabulary Model, the learnability of different words changes with a child's developmental level, where LEARNABLE increases as a child grows older.
+Therefore, in Simulation 2, we looked at the expected patterns of translation equivalent learning across varying levels of LEARNABLE (i.e., the number of learnable words in each language as developmental level changes).
 Additionally, we further examined vocabulary composition by computing the number of singlets in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) language.
@@
 In total, data from 161 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
-Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM with the formula NONDOM/(DOM+NONDOM).
-We also counted the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children's concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
-Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), medium balanced (BALANCE = .20 - .35), and least balanced (BALANCE = .00 - .02).
+Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children's concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
+Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), to medium balanced (BALANCE = .20 - .35), to least balanced (BALANCE = .00 - .02).
 
-In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT) as shown by the largest bars for the children at LEARNABLE = 600 compared to the shortest bars for children at LEARNABLE = 300.
-Moreover, we continue to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE), as evidenced by the green bars that are the highest for the most balanced children (BALANCE = .36 - .50) compared to the medium balanced (BALANCE = .21 - .35) and the least balanced children (BALANCE = .00 - .20).
-Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents.
-Across all three developmental levels of LEARNABLE, the light blue bars representing singlets in the dominant language (DOM-SINGLET) grow with less balanced exposure, whereas the dark blue bars representing singlets in the non-dominant language (NONDOM-SINGLET) grow with more balanced exposure.
-Overall, we observe 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
-Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
-
-\begin{itemize}
-\tightlist
-\item
-  Prediction 2a: Have larger concept vocabularies (CONCEPT)
-\item
-  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
-\item
-  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with a less balanced vocabulary (BALANCE) produce the most DOM-SINGLET.
-\item
-  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with a more balanced vocabulary (BALANCE) produce the most NONDOM-SINGLET.
-\end{itemize}
-
-\begin{figure}
+\begin{figure}[H]
 
@@
 
+In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+Moreover, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
+Overall, we generated 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
+Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
+
+\begin{itemize}
+\tightlist
+\item
+  Prediction 2a: Have larger concept vocabularies (CONCEPT).
+\item
+  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+\item
+  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least \emph{balanced} vocabulary (BALANCE) will produce the most DOM-SINGLET.
+\item
+  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most \emph{balanced} vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+\end{itemize}
+
 \hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}{%
-\subsection{Simulation 3: Bias towards or against translation equivalent learning compared to singlets}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}}
+\subsection{1.3 Simulation 3: Bias towards or against translation equivalent learning compared to singlets}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}}
 
 In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
-In our final simulation, we examined models where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
+In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
 Mathematically, this requires varying the BIAS parameter.
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50\% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are neither easier nor harder to learn than singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equal to learn as singlets).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50\% more likely to be learned than singlets).
-In Figure 4, we illustrate the three different scenarios of simulated translation equivalent knowledge.
-Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies produced more translation equivalents.
-Thus, overall relationships between vocabulary balance and translation equivalent knowledge remained similar across the Avoidance, Preference, and Neutral Accounts.
-What changed is the slope of translation equivalent learning: the slopes were the shallowest under the Avoidance Account where BIAS = 0.5, whereas the slopes are steepest under the Preference Account where BIAS = 1.5.
+In Figure 4, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
+Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
+Thus, overall relationships between BALANCE and TE remained similar across the Avoidance, Preference, and Neutral Accounts.
+What changed was the slope of translation equivalent learning: the slopes were the shallowest under the Avoidance Account where BIAS = 0.5, whereas the slopes were steepest under the Preference Account where BIAS = 1.5.
 With this, we further outline Prediction Set 3:
@@
 \item
-  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning, with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn than singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
 \end{itemize}
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\caption{Different scenarios of expected translation equivalents learning (TE) while assuming bias towards (BIAS < 1) or against learning translation equivalents (BIAS > 1), or no bias (BIAS = 1).}\label{fig:fig4}
+\caption{Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.}\label{fig:fig4}
 \end{figure}
@@
 In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
-In Study 2, we tested these predictions using archival data obtained from 200 French--English bilingual children aged 18 to 33 months.
+In Study 2, we tested these predictions using archival vocabulary data from 200 French--English bilingual children aged 18 to 33 months.
 
 \hypertarget{method}{%
-\subsection{Method}\label{method}}
+\subsection{2.1 Method}\label{method}}
 
-Ethics approval was obtained by the Human Research Ethics Board of XYZ University (ethics protocol number 0000) and informed consent was obtained from the participants' parents.
+Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certification Number 10000439) and informed consent was obtained from the children's parents.
 
 \hypertarget{participants}{%
-\subsubsection{Participants}\label{participants}}
+\subsection{2.1.1 Participants}\label{participants}}
 
@@
 Bilingual children were defined as those exposed at least 25\% of the time over the course of their lives globally to both English and French and with less than 10\% of exposure to a third language.
-For children who participated more than once, their language exposure followed such criteria between visits.
+For children who participated more than once, their language exposure followed such criteria for all visits.
 Following the approach in Study 1, children's dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
-Within the 229 data points, 59.80\% of children were English-dominant and 40.17\% were French-dominant.
+Within the 229 data points, 59.83\% of children were English-dominant and 40.17\% were French-dominant.
 Data collection was conducted in Montréal, Québec, Canada.
 Montréal is a multicultural city where both English and French are widely used in society.
-Children's demographic characteristics are presented in Table 4.
+Children's demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
 
@@
 \hypertarget{measures}{%
-\subsubsection{Measures}\label{measures}}
+\subsection{2.1.2 Measures}\label{measures}}
 
-\begin{itemize}
-\item
-  \textbf{MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI).} Bilingual children's expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
+\hypertarget{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}{%
+\subsubsection{MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)}\label{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}}
+
+Bilingual children's expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
 Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (25\%).
+We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (24\%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
@@
 There are 680 words in the English CDI version and 664 in the Québec French version.
-\item
-  \textbf{Language Exposure Questionnaire using the MAPLE approach.}
-\end{itemize}
+
+Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero et al.~(2020) by three proficient bilingual French--English adults who carefully examined each language version of the CDI.
+Word pairs that made reference to the same concept (e.g., English ``apple'' and French \emph{``pomme''}) were considered to be translation equivalents.
+In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
+Words that had similar phonetic realizations (e.g., English ``alligator'' and French \emph{``alligator''}) were also considered translation equivalents.
+Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
+A full list of translation equivalents is available at {[}\url{https://osf.io/7fz6c/}{]}.
+
+After determining the dominant language of a child based on the vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
+Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).
+
+\hypertarget{language-exposure-questionnaire-using-the-maple-approach}{%
+\subsubsection{Language Exposure Questionnaire using the MAPLE approach}\label{language-exposure-questionnaire-using-the-maple-approach}}
+
+Children's language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch \& Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE, Byers-Heinlein et al., 2018).
+The LEQ is a structured interview that lasts approximately 15 minutes.
+It includes targeted questions that quantify the child's language exposure from birth until their current age.
+The LEQ and MAPLE provide a global language exposure estimate based on the number of hours the child is exposed to each language within all contexts (e.g., home, daycare, etc.).
+Children's average global exposure to each language is described in Table 4.
 
 \hypertarget{procedure}{%
-\subsubsection{Procedure}\label{procedure}}
+\subsection{2.1.3 Procedure}\label{procedure}}
+
+Caregivers were asked to fill out the CDI questionnaires as part of their child's participation in experimental studies on language development, speech perception, and word learning.
+Caregivers were instructed to check off the words produced by their child using either a CDI paper questionnaire or the same questionnaire administered on a tablet.
+Data from paper based questionnaires was double entered and checked by trained research assistants.
 
 \hypertarget{results}{%
-\subsection{Results}\label{results}}
+\subsection{2.2 Results}\label{results}}
 
-\hypertarget{testing-prediction-set-1}{%
-\subsubsection{Testing Prediction Set 1:}\label{testing-prediction-set-1}}
+Data analyses were conducted using R (Version 4.0.2, 2020).
+Analysis scripts and the data set used in the present study are available at {[}\url{https://osf.io/2t5kw/}{]}.
+We first present descriptive measures of vocabulary, and then tests of the three sets of predictions generated in Study 1.
 
-\hypertarget{testing-prediction-set-2}{%
-\subsubsection{Testing Prediction Set 2:}\label{testing-prediction-set-2}}
+\hypertarget{descriptive-measures-of-vocabulary}{%
+\subsection{2.2.1 Descriptive measures of vocabulary}\label{descriptive-measures-of-vocabulary}}
 
-\hypertarget{testing-prediction-set-3}{%
-\subsubsection{Testing Prediction Set 3:}\label{testing-prediction-set-3}}
+On average, bilinguals in the sample had a mean word vocabulary size (WORD) of 295 (SD = 254.60), with a wide range of 6 - 1071 words.
+As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = 206.10, SD = 175.60, range = 4 - 657) than in their non-dominant language (NONDOM; M = 88.90, SD = 98.50, range = 2 - 469), \emph{t}(228) = 13.89, \(p < .001\), \(d = 0.92\).
+
+Children produced an average of 67.70 translation equivalents (TE; SD = 85.10, range = 1 - 409).
+The remainder of words were singlets: Children produced many more singlets in their dominant language (DOM-SINGLET; M = 138.40, SD = 124.40, range = 2 - 523) than in their non-dominant language (NONDOM-SINGLET; M = 21.20, SD = 20.10, range = 0 - 94), \emph{t}(228) = 13.89, \(p < .001\), \(d = 0.92\).
+On average, children's concept vocabulary size was 227.30 (CONCEPT; SD = 181.30, range = 4 - 695).
+
+Vocabulary balance (BALANCE) was then determined based on the proportion of total words produced in the non-dominant language following the formula BALANCE = NONDOM/WORD as in Study 1.
+On average, bilingual children in our sample had a balance score BALANCE of 0.31 (SD = 0.13), ranging from 0.02 to 0.50.
+Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
+The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.17\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).
+
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, but for young bilinguals balance can also be considered in terms of input in each language.
+We therefore compare the vocabulary balance with the proportion of exposure bilingual children received in their non-dominant language.
+To make values comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance, rather than the language that children heard most and least often.
+For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79.04\%), although this was not the case for some children (48 children, 20.96\%).
+The correlation between vocabulary-defined BALANCE and the raw percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 5).
+Thus, these two constructs were related, although not identical.
+
+~
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.75\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig5-1}
+
+}
+
+\caption{Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.}\label{fig:fig5}
+\end{figure}
+
+\hypertarget{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}{%
+\subsection{2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures}\label{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}}
+
+Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson's correlations.
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations).
+
+Prediction 1a was that children with more balanced vocabularies would produce more translation equivalents.
+As shown in Figure 2 Row 2, our vocabulary data confirmed the prediction, \emph{r}(227) = 0.25, \(p < .001\), where children with the most balanced vocabulary produced the most translation equivalents.
+We further tested this prediction by dividing children into 5 balance subset groups (0 \textless{} BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, \emph{F}(4, 224) = 3.61, \(p = .007\).
+The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the least translation equivalents.
+Detailed descriptive statistics are reported in Table 6.
+
+Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): \emph{r}(227) = 0.90, \(p < .001\), and dominant-language vocabulary (DOM): \emph{r}(227) = 0.76, \(p < .001\).
+Figure 2 Row 2 further illustrates these relationships observed in our dataset.
+
+Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
+As shown in Figure 2 Row 2, we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
+
+\begin{lltable}
+
+\begin{TableNotes}[para]
+\normalsize{\textit{Note.} *** p < .001, ** p < .01, * p < .05.}
+\end{TableNotes}
+
+\scriptsize{
+
+\begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
+\caption{\label{tab:table5}Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).}\\
+\toprule
+ & \multicolumn{1}{c}{Age (in month)} & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET}\\
+\midrule
+\endfirsthead
+\caption*{\normalfont{Table \ref{tab:table5} continued}}\\
+\toprule
+ & \multicolumn{1}{c}{Age (in month)} & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET}\\
+\midrule
+\endhead
+Age (in month) &  & 0.95**** & -0.24*** & 0.65**** & 0.69**** & 0.45**** & 0.48**** & 0.65**** & 0.19**\ \\\
+LEARNABLE &  &  & -0.23*** & 0.62**** & 0.65**** & 0.43**** & 0.44**** & 0.62**** & 0.21**\ \\\
+BALANCE &  &  &  & -0.07\ \ \ \ & -0.29**** & 0.35**** & 0.25**** & -0.58**** & 0.63****\\
+WORD &  &  &  &  & 0.96**** & 0.87**** & 0.90**** & 0.74**** & 0.44****\\
+DOM &  &  &  &  &  & 0.70**** & 0.76**** & 0.89**** & 0.23***\\
+NONDOM &  &  &  &  &  &  & 0.99**** & 0.31**** & 0.72****\\
+TE &  &  &  &  &  &  &  & 0.38**** & 0.60****\\
+DOM-SINGLET &  &  &  &  &  &  &  &  & -0.09\ \ \ \\\
+NONDOM-SINGLET &  &  &  &  &  &  &  &  & \\
+CONCEPT &  &  &  &  &  &  &  &  & \\
+\bottomrule
+\addlinespace
+\insertTableNotes
+\end{longtable}
+
+}
+
+\end{lltable}
+
+\begin{table}
+
+\caption{\label{tab:table6}Average number of translation equivalents (TE) produced by each balance group.}
+\centering
+\begin{tabular}[t]{>{\centering\arraybackslash}p{150px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{75px}}
+\toprule
+  & N & Mean & SD & range\\
+\midrule
+0.4 < BALANCE ≤ 0.5 & 69 & 92 & 114.44 & 4 - 409\\
+0.3 < BALANCE ≤ 0.4 & 58 & 68 & 89.30 & 1 - 376\\
+0.2 < BALANCE ≤ 0.3 & 52 & 64 & 57.13 & 5 - 221\\
+0.1 < BALANCE ≤ 0.2 & 32 & 47 & 35.17 & 7 - 137\\
+0 < BALANCE ≤ 0.1 & 18 & 18 & 11.57 & 2 - 52\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+\hypertarget{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}{%
+\subsection{2.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels}\label{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}}
+
+Prediction Set 2 pertained to expected patterns of acquisition of translation equivalents and singlets for children of different developmental levels.
+In our dataset, developmental level was approximated by children's age.
+Figure 3 Panel B shows the concept vocabulary (CONCEPT) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+To illustrate the acquisition of translation equivalents and singlets at different developmental levels, we divided children into three age groups: younger children of 18--22 months, middle children of 23--27 months, and older children of 28--33 months.
+
+Prediction 2a was that older children (i.e., those at a later developmental level) would have larger concept vocabularies than younger children (i.e., those at an earlier developmental level).
+We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, \emph{r}(227) = 0.69, \(p < .001\), and therefore confirmed the prediction.
+This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, (\(F\)(2, 226) = 90.86, \(p < .001\)).
+Older children of 28--33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of 414.60 (\(p\)s \textless{} .001), those middle children of 23--27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of 252.10, and those younger children of 18--22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of 119.90 (\(p\)s \textless{} .001).
+
+Prediction 2b was that older children would produce more translation equivalents than younger children.
+First, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, \emph{r}(227) = 0.48, \(p < .001\), and therefore confirmed the prediction.
+In a one-way ANOVA with age group as factor, we further found that groups differed in how many translation equivalents they produced (\(F\)(2, 226) = 31.74, \(p < .001\)).
+Younger children of 18--22 months produced an average of 33.70 translation equivalents, middle children of 23--27 months produced an average of 71.10 translation equivalents, and older children of 28--33 months produced an average of 131.40 translation equivalents (\(p\)s \textless{} .01).
+
+Prediction 2c was that both older children and those with the least balanced vocabularies (BALANCE) would produce more dominant-language singlets (DOM-SINGLET).
+This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.65, \(p < .001\), and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), \emph{r}(227) = -0.58, \(p < .001\).
+As shown in Figure 3 Panel B, children were divided into least balanced (range of balance: .00 - .20), medium balanced (range of balance: .20 - .35) and most balanced (range of balance: .35 - .50) groups (i.e., the same criteria as in Figure 5).
+In a one-way ANOVA with balance group as a between-subjects factor, we observed that the least balanced children produced the most singlets in their dominant language (\(p\)s \textless{} .001), with the least balanced, medium balanced, most balanced children producing respectively: 255.50, 141.50, and 71.90 words in their dominant language (\(F\)(2, 226) = 50.77, \(p < .001\)).
+
+Prediction 2d was that older children and those with the most balanced vocabularies (BALANCE) would produce more singlets in their non-dominant language.
+This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.19, \(p = .005\), and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), \emph{r}(227) = 0.63, \(p < .001\).
+In a one-way ANOVA with balance group as a between-subjects factor, we confirmed that children who differed in how balanced their vocabulary knowledge was also differed in how many singlets they produced in their non-dominant language (\(F\)(2, 226) = 61.89, \(p < .001\)).
+As shown in Figure 3 Panel B, we observed that children produced very few singlets in their non-dominant language, although the most balanced children produced the most singlets in their non-dominant language (mean of the most balanced children = 35.10 \textgreater{} mean of the medium balanced children = 15.50 \textgreater{} mean of the most balanced children = 5.70; \(p\)s \textless{} .001).
+
+\hypertarget{testing-prediction-set-3-rate-of-translation-equivalent-learning}{%
+\subsection{2.2.4 Testing Prediction Set 3: Rate of translation equivalent learning}\label{testing-prediction-set-3-rate-of-translation-equivalent-learning}}
+
+Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
+To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate BIAS parameter.
+
+First, we will walk through the parameters in this model.
+The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
+As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
+For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
+For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since the 90th percentile information was available only up to 30 months.
+
+\begin{table}
+
+\caption{\label{tab:table7}The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.}
+\centering
+\fontsize{10}{12}\selectfont
+\begin{tabular}[t]{>{\centering\arraybackslash}p{75px}>{\centering\arraybackslash}p{125px}>{\centering\arraybackslash}p{125px}>{\centering\arraybackslash}p{125px}}
+\toprule
+Age (months) & Number of English words produced at 90th percentile & Number of French words produced at 90th percentile & Average (LEARNABLE)\\
+\midrule
+16 & 129.0 & 97.0 & 113.0\\
+17 & 201.5 & 161.1 & 181.3\\
+18 & 268.7 & 221.1 & 244.9\\
+19 & 330.5 & 277.0 & 303.8\\
+20 & 387.0 & 328.8 & 357.9\\
+\addlinespace
+21 & 438.1 & 376.5 & 407.3\\
+22 & 483.8 & 420.0 & 451.9\\
+23 & 524.2 & 459.4 & 491.8\\
+24 & 559.3 & 494.7 & 527.0\\
+25 & 589.0 & 525.9 & 557.4\\
+\addlinespace
+26 & 613.4 & 553.0 & 583.2\\
+27 & 632.4 & 575.9 & 604.1\\
+28 & 646.0 & 594.7 & 620.4\\
+29 & 654.3 & 609.4 & 631.9\\
+30 - 33 & 657.9 & 620.0 & 638.9\\
+\bottomrule
+\end{tabular}
+\end{table}
+
+Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
+To reproduce the Bilingual Vocabulary Model's formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included in the model (denoted in R by using a colon rather than an asterisk between the interacting predictors).
+Therefore, our final linear regression model equation was:
+
+\emph{ObservedTE \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary/90 percentile of CDI items.}
+
+\noindent (In R language, the model was entered as:
+
+\emph{ObservedTE * 90 percentile of CDI items \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary})
+
+\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was consistent with the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
+Otherwise, a coefficient less than 1 represents a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 represents a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
+
+Our model showed an excellent model fit of \(R^2\) = 0.96, indicating that our model explained 96\% of the variance in bilinguals' translation equivalent knowledge.
+The linear regression model estimated a BIAS coefficient of 1.02, \(p < .001\).
+This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words.
+
+To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child's expected translation equivalent knowledge (setting BIAS = 1), which is plotted against our observed data in Figure 6.
+Expected and observed translation equivalents were closely aligned with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1), suggesting that the Neutral Account provides a parsimonious explanation for bilinguals' translation equivalent knowledge.
+This provides evidence for the notion that translation equivalents are neither harder nor easier to learn than singlets in bilingual vocabulary learning.
+Note that visual inspection suggested that there could be some possible outliers.
+Cook's distance was estimated for our linear regression model listed above and identified two data points with a cook's distance over 0.4.
+After removing those two data points, the linear regression model returned a coefficient of 1.05, \(p < .001\), with \(R^2\) = 0.96.
+As the model fit was similar to the model without eliminating the two outlier data points, we proceeded with the full data set keeping the two potential outlier data points.
+
+\begin{figure}
+
+{\centering \includegraphics[width=0.7\linewidth,height=0.7\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig6-1}
+
+}
+
+\caption{The number of simulated and observed translation equivalents plotted against each other.
+The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
+The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.}\label{fig:fig6}
+\end{figure}
+
+Despite the good overall fit to the data, a close examination of Figure 6 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
+Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1) for children with more than 300 total vocabulary.
+To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
+The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.02, \(p < .001\), whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.22, \(p < .001\).
+Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size (\(R^2\) = 0.97) than children with smaller vocabulary size (\(R^2\) = 0.88).
+Overall, this analysis suggests that translation equivalent learning for children with larger vocabularies corresponds best to the Neutral Account, but translation equivalent learning for children with smaller vocabularies corresponds best to the Preference Account.
+
+\begin{figure}
+
+{\centering \includegraphics[width=1.2\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig7-1}
+
+}
+
+\caption{The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
+The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.}\label{fig:fig7}
+\end{figure}
 
@@
 
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn than singlet words (i.e., the first label for a particular referent).
+To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were available to be learned given the child's age.
+The relative difficulty of learning translation equivalents relative to singlets was modeled via the bias parameter (BIAS), which indicated whether translation equivalent learning is consistent with the Avoidance (BIAS \textless{} 1), Preference (BIAS \textgreater{} 1), or Neutral Account (BIAS = 1).
+
+\hypertarget{confirmation-of-model-predictions}{%
+\subsection{Confirmation of model predictions}\label{confirmation-of-model-predictions}}
+
+In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18-33 months.
+Three sets of model predictions were confirmed in our empirical dataset.
+
+Prediction Set 1 pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
+In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
+This pattern is consistent with reports from previous research (Davidson \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
+This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Poulin-Dubois et al., 2013; Legacy et al., 2015).
+Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
+A similar pattern has been reported by Legacy and colleagues (2015), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2015).
+
+Prediction Set 2 pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
+Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
+Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
+This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy et al., 2016).
+Predictions 2c and 2d were also confirmed by our vocabulary data.
+While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Place \& Hoff, 2011; Pearson et al., 1997).
+Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
+
+Overall, we observed a strong correspondence between the data simulated under the Bilingual Vocabulary Model and our observed data.
+Moreover, our model predicted numerous disparate patterns that have been previously reported in the literature.
+
+Having validated our overall approach, Prediction 3 motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents learning are similar to learn than singlets), with this model explaining 96\% of variance in the data.
+However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
+Modeling their data separately, we found evidence for the Preference account: younger children at around 22 months appeared to learn translation equivalents more easily than singlets, whereas older children at around 28 months learned translation equivalents similarly to singlets.
+This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
+This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
+The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
+Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
+The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children's own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
+
+\hypertarget{developmental-change-in-translation-equivalent-learning}{%
+\subsection{Developmental change in translation equivalent learning}\label{developmental-change-in-translation-equivalent-learning}}
+
+The developmental change of bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman \& Bowman, 1989).
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman \& Bowman, 1989).
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990, David \& Tell, 2005).
+At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
+Yet, this explanation would not predict that younger bilinguals' data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance account, before perhaps returning to the Neutral account once children realize that each referent should have a label in each language.
+Thus, changes in one-to-one mapping biases do not provide a complete explanation for our results.
+
+Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
+For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the livre?'' if they know their child understands the French word ``livre'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
+Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
+Overall, more research will be needed to understand why translation equivalents appear to be over-represented in younger bilinguals' vocabularies.
+
+\hypertarget{assumptions-limitations-and-future-directions}{%
+\subsection{Assumptions, limitations, and future directions}\label{assumptions-limitations-and-future-directions}}
+
+Our Bilingual Vocabulary Model presented an integrated computational account of translation equivalent learning, focusing on the joint probability of learning the word for a concept in each language.
+To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
+However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Hoff, 2003; Fernald, Marchman, \& Weisleder, 2013), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020, Rowe, 2012).
+It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+Such a model may better characterize and predict bilingual vocabulary development as a function of experience.
+Moreover, it would be important to extend our Bilingual Vocabulary Model to longitudinal data or data of a different bilingual population to investigate if it is possible to replicate the qualitative shift where bilingual children's ability to learn translation equivalents appears to change across development.
+
+Another limitation of our model is that it takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
+However, the reality of bilingual experience might be more complex.
+First, some concepts expressed as a single word in one language may be lexicalized by two words in another language (e.g., English has a single word for ``sister'' but Mandarin has separate words for \emph{``jiějie''} {[}older sister{]} and \emph{``mèimei''} {[}younger sister{]}).
+As another example, some words may not have a translation equivalent in the other language (e.g., the Japanese word \emph{``sushi''} is borrowed into other languages).
+Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a ``bowl'' in English but an \emph{``assiette''} {[}plate{]} in French).
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
+Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English--French pair ``banana'' -- \emph{``banane''}) more easily than those that do not share similar phonological form (e.g., the English--French pair ``dog'' -- \emph{``chien''}) (Bosch \& Ramon-Casas, 2014).
+Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
+While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
+Moreover, if they do prove to be important, such factors could be added to future iterations of the Bilingual Vocabulary Model.
+
+Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
+However, following the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2016; Tardif et al., 2008).
+For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and simple nouns (``ball,'' ``dog'') are among the first words children across languages and cultures.
+It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
+Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
+
+Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the bias parameter (BIAS) and the parameter that accounts for how many words are potentially learnable at a particular age (LEARNABLE).
+Under the Bilingual Vocabulary Model, the learnability parameter and the bias parameter jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter changes by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also change by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
+Our model estimated the number of learnable words to be the number that children at the 90th percentile at a particular age produce.
+Small changes to this approach (e.g., taking the number of words children at the 95th percentile produce) would likely not drastically alter our results, nor change the qualitative shift that we observed in our data.
+Nonetheless, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+
+\hypertarget{conclusions}{%
+\subsection{Conclusions}\label{conclusions}}
+
+In sum, the acquisition of translation equivalents has been considered a special component in bilingual children' vocabulary development.
+Previous research has put forward three diverging accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+We proposed the Bilingual Vocabulary Model, which provides a quantitative way to test these accounts, by modeling translation equivalent learning in relation to vocabulary size in each language and the number of potentially learnable words, which is constrained by children's developmental level.
+Results using archival data from a large number of young French--English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
+Moreover, our model parsimoniously explained previously disparate observations about bilingual children's translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
+
 \end{document}

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
suppressed automatic table numbering

@@
   - \usepackage{amsmath}
+  - \usepackage[labelformat=empty]{caption}
+  - \usepackage{caption}
   - \renewcommand{\topfraction}{1}
@@
 ---
+
+\captionsetup[table]{labelformat=empty}
 
@@
 \newpage
-```{r box1, results="asis", fig.cap = " "}
+```{r box1, results="asis", tab.cap = NULL}
 box1 <- data.frame(title = "To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
@@
         format = "latex", booktabs = TRUE,
-        caption = "Summary of the parameters in the Bilingual Vocabulary Model.") %>%
+        caption = "Table 1. Summary of the parameters in the Bilingual Vocabulary Model.") %>%
     kableExtra::kable_styling(font_size = 7) %>%
@@
         format="latex", booktabs = TRUE,
-        caption = "Summary of the parameters used in each simulation.") %>%
+        caption = "Table 2. Summary of the parameters used in each simulation.") %>%
     column_spec(1:2, width = "45px") %>%
@@
       format="latex", booktabs = TRUE,
-      caption = "Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.") %>%
+      caption = "Table 3.
+Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.") %>%
   pack_rows("Main Parameters", 1, 4) %>%
@@
 keepers_ws_TE %>% 
-  summarize(N_babies = n_distinct(baby_id),
+  summarise(N_babies = n_distinct(baby_id),
             N_admin = n_distinct(ID_testdate))
@@
 
+&nbsp;
 ```{r table4, results="asis"}
@@
 papaja::apa_table(d_participants, format.args = list(digits = 1),
-                  caption = "Demographic characteristics of participants (data points = 229).",
+                  caption = "Table 4. Demographic characteristics of participants (data points = 229).",
                   col.names =c("","Mean","$SD$","Range"),
-                  align=c("l","c","c","c"))
+                  align=c("l","c","c","c"),
+                  placement = "h")
 ```
@@
 papaja::apa_table(correlation_table,
-                  caption = "Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).",
+                  caption = "Table 5. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).",
                   note = "*** p < .001, ** p < .01, * p < .05.",
@@
       format="latex", booktabs=TRUE, align = "c",
-      caption = "Average number of translation equivalents (TE) produced by each balance group.") %>%
+      caption = "Table 6. Average number of translation equivalents (TE) produced by each balance group.") %>%
   column_spec(1, width = "150px") %>%
@@
         format = "latex", booktabs = TRUE, align = "c",
-        caption = "The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.") %>%
+        caption = "Table 7.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.") %>%
   kableExtra::kable_styling(font_size = 10) %>%

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.tex
suppressed automatic table numbering

@@
 \usepackage{amsmath}
+\usepackage[labelformat=empty]{caption}
+\usepackage{caption}
 \renewcommand{\topfraction}{1}
@@
 
+\captionsetup[table]{labelformat=empty}
+
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
@@
 
-\caption{\label{tab:table1}Summary of the parameters in the Bilingual Vocabulary Model.}
+\caption{\label{tab:table1}Table 1. Summary of the parameters in the Bilingual Vocabulary Model.}
 \centering
@@
 
-\caption{\label{tab:table2}Summary of the parameters used in each simulation.}
+\caption{\label{tab:table2}Table 2. Summary of the parameters used in each simulation.}
 \centering
@@
 
-\caption{\label{tab:table3}Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.}
+\caption{\label{tab:table3}Table 3.
+Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.}
 \centering
@@
 
-\begin{table}[tbp]
+~
+
+\begin{table}[h]
 
@@
 
-\caption{\label{tab:table4}Demographic characteristics of participants (data points = 229).}
+\caption{\label{tab:table4}Table 4. Demographic characteristics of participants (data points = 229).}
 
@@
 \begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table5}Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).}\\
+\caption{\label{tab:table5}Table 5. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).}\\
 \toprule
@@
 
-\caption{\label{tab:table6}Average number of translation equivalents (TE) produced by each balance group.}
+\caption{\label{tab:table6}Table 6. Average number of translation equivalents (TE) produced by each balance group.}
 \centering
@@
 Prediction Set 2 pertained to expected patterns of acquisition of translation equivalents and singlets for children of different developmental levels.
-In our dataset, developmental level was approximated by children's age.
+In our data set, developmental level was approximated by children's age.
 Figure 3 Panel B shows the concept vocabulary (CONCEPT) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
@@
 
-\caption{\label{tab:table7}The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.}
+\caption{\label{tab:table7}Table 7.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.}
 \centering

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
added references

@@
 
-#bibliography      : ["gaze-following_references.bib"]
-#replace_ampersands: yes
+bibliography      : ["paper_TE_bilingual_vocabulary_model_references.bib"]
+replace_ampersands: yes
 
@@
 For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer, Bornstein, & De Coster, 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents [also called doublets; @Umbel_etal_1992], and are observed amongst bilingual children’s first words [e.g., @David_Wei_2008; @DeHouwer_etal_2006; @Pearson_etal_1995].
 Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics.
-For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak & Byers-Heinlein, 2019).
+For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent [e.g., @Bilson_etal_2015; @Floccia_etal_2020; @Jardak_Byers-Heinlein_2019].
 That is, upon hearing the English word “apple”, the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent ( a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent.
+In vocabulary acquisition, bilingual children must learn a first label for a referent [a “singlet”; @Umbel_etal_1992] before they can learn its translation equivalent.
 Is translation equivalent learning different from singlet learning?
@@
 ### Avoidance Account: Translation equivalents are harder to learn than singlets
-Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies (e.g., Imedadze, 1978; Swain & Wesche 1975; Volterra & Taeschner, 1978).
+Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies [e.g., @Imedadze_1967; @Swain_Wesche_1975; @Volterra_Taeschner_1978].
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman & Wachtel, 1988; Markman, 1992; 1994).
+This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level [@Markman_Wachtel_1988; @Markman_1992; @Markman_1994].
 For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug”, they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.
 
-Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein & Werker, 2009; Davidson & Tell, 2005; Houston-Price, Caloghiris, & Raviglione, 2010).
+Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition [@Byers-Heinlein_Werker_2009; @Davidson_Tell_2005; @Houston-Price_etal_2010].
 When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word.
@@
 
-Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David & Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017).
+Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development [@David_Wei_2008; @DeHouwer_etal_2006; @Holowka_etal_2002; @Pearson_etal_1995; @Legacy_etal_2017].
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
-For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010).
-A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al, 2020).
+For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects [@Byers-Heinlein_Werker_2009; @Byers-Heinlein_Werker_2013; @Houston-Price_etal_2010].
+A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals [@Lewis_etal_2020].
 
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
-Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak & Byers-Heinlein, 2018; Singh, 2013).
-In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
+When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets [e.g., @Montanari_2010; @Poulin-Dubois_etal_2013; @Poulin-Dubois_etal_2018].
+Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related [e.g., @Floccia_etal_2020; @Jardak_Byers-Heinlein_2019; @Singh_2014].
+In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition [e.g., @Bilson_etal_2015; @Floccia_etal_2020].
 The Preference Account predicts that translation equivalents are more easily learned than singlets.
@@
 There are several lines of empirical evidence to support the Preference Account.
-For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages (Lanvers, 1999; Pearson & Fernández, 1994).
+For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages [@Lanvers_1999; @Pearson_Fernandez_1994].
 The main explanation that has been given for this finding is that additional exposure to their non-dominant language — which became their new dominant language — enabled fast mapping of words to already-lexicalized concepts.
 
-Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015).
+Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years [@Bilson_etal_2015].
 The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like “cat” and “dog” are strongly semantically related).
@@
 Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
-For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, & Hurtado, 2010).
-Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
+For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language [@Marchman_etal_2010].
+Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages [e.g., @Pearson_Fernandez_1994; @Pearson_etal_1997], and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size [@Hurtado_etal_2014].
 Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, & Hurtado, 2010; Place & Hoff, 2011; Pearson et al., 1997).
-Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
+In general, more language exposure leads to larger vocabulary size [e.g., @Barnes_Garcia_2013; @Boyce_etal_2013; @Hurtado_etal_2014; @Marchman_etal_2010; @Place_Hoff_2011; @Pearson_etal_1997].
+Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure [i.e., non-dominant language; @Pearson_etal_1997; @Place_Hoff_2011].
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts.
-This is known as the Complementarity Principle (Grosjean, 2016).
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
+This is known as the Complementarity Principle [@Grosjean_2016].
+For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language [@Bialystok_etal_2010].
 If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
@@
 
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, & Weisleder, 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).
+Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech [e.g., @Brent_Siskind_2001; @Swingley_Humphrey_2018], children’s efficiency of processing words they hear [e.g., @Hurtado_etal_2014; @Weisleder_Fernald_2013], cognitive development and perceptual bias [e.g., @Benedict_1979; @Goodman_etal_2008], and family socioeconomic status [@Fernald_etal_2013].
+Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently [@Brent_Siskind_2001; @Goodman_etal_2008; @Swingley_Humphrey_2018].
 
@@
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
-For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
-Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007).
-This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky et al., 2019).
+For example, across many languages including English, children show a noun bias in their early lexicons [@Braginsky_etal_2019; @Goodman_etal_2008], although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired [@Tardif_1996].
+Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time [@Fenson_etal_2007].
+This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types [@Bergelson_2020; @Braginsky_etal_2019].
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
@@
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories [CDI; @Fenson_etal_2007], one option would be to set LEARNABLE to be the total number of items on the CDI.
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
@@
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
-Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
+Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words [retrieved from the Wordbank database; @Frank_etal_2016].
 Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable — if imperfect — estimate of the number of learnable words (LEARNABLE) that a child of Jamie’s age could potentially acquire in each language.
@@
 
-In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997).
+In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English [@Fenson_etal_2007] and Québec French [@Trudeau_etal_1999].
 Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old [i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; @Frank_etal_2016].
 We set BIAS to 1, meaning that in these examples translation equivalents are similarly easy to learn as singlets.
@@
 
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
-(2020).
+Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as @Gonzalez-Barrero_2020.
 Some children took part in more than one in-lab study (n = `r length(d_multiple$baby_id)`); thus, they contributed data at more than one time point.
@@
 Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
-Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
+Caregivers completed the original CDI English version [@Fenson_etal_2007] and its Québec French adaptation [@Trudeau_etal_1999].
 We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), and respondent not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
@@
 
-Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero et al.
-(2020) by three proficient bilingual French–English adults who carefully examined each language version of the CDI.
+Translation equivalents (TE) were determined in the same manner as @Gonzalez-Barrero_2020 by three proficient bilingual French–English adults who carefully examined each language version of the CDI.
 Word pairs that made reference to the same concept (e.g., English “apple” and French _“pomme”_) were considered to be translation equivalents.
@@
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
-A full list of translation equivalents is available at [https://osf.io/7fz6c/].
+A full list of translation equivalents is available at [https://osf.io/2t5kw/].
 
@@
 ### Language Exposure Questionnaire using the MAPLE approach
-Children’s language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE, Byers-Heinlein et al., 2018).
+Children’s language exposure was measured using the Language Exposure Questionnaire [LEQ; @Bosch_Sebastian-Galles_2001] and the Multilingual Approach to Parent Language Estimates [MAPLE; @Byers-Heinlein_etal_2020].
 The LEQ is a structured interview that lasts approximately 15 minutes.
@@
 The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
+As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank [@Frank_etal_2016], and Table 7 lists the denominator at different ages.
 For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
@@
 
+&nbsp;
 ```{r fig7, fig.cap="The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
@@
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (Davidson & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+This pattern is consistent with reports from previous research [@David_Wei_2008; @Legacy_etal_2016; @Montanari_2010; @Pearson_etal_1995; @Pearson_etal_1997].
+Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size [@Legacy_etal_2016; @Montanari_2010].
 Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Poulin-Dubois et al., 2013; Legacy et al., 2015).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents [@Poulin-Dubois_etal_2013; @Legacy_etal_2016].
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2015), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2015).
+A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child [@Legacy_etal_2016].
 
@@
 We operationalized developmental level in terms of children’s age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
-Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
+Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature [@Pearson_etal_1993].
 Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
-This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David & Wei, 2008; Legacy et al., 2016).
+This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older [@David_Wei_2008; @Legacy_etal_2016].
 Predictions 2c and 2d were also confirmed by our vocabulary data.
 While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Place & Hoff, 2011; Pearson et al., 1997).
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language [e.g., @Boyce_etal_2013; @Hoff_etal_2012; @Marchman_etal_2010; @Place_Hoff_2011; @Pearson_etal_1997].
 Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
@@
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
-This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
+This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets [@Bilson_etal_2015].
 The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
-Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
+Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account [e.g., @Bilson_etal_2015; @Pearson_etal_1995].
 The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children’s own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
@@
 The developmental change of bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman & Bowman, 1989).
-In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman & Bowman, 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au & Glusman, 1990, David & Tell, 2005).
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents [@Halberda_2003; @Lewis_etal_2020; @Merriman_etal_1989].
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent [@Halberda_2003; @Merriman_etal_1989].
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words [@Lewis_etal_2020], while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
 At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
-Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary [@Kremin_etal_2021; @Nicoladis_Secco_2000].
 For example, a caregiver may choose to say to their English–French bilingual child “Can you grab the livre?” if they know their child understands the French word “livre” but not the English equivalent “book”.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Hoff, 2003; Fernald, Marchman, & Weisleder, 2013), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020, Rowe, 2012).
+However, our model does not consider other qualitative factors including family socioeconomic status [e.g., @Hoff_2003; @Fernald_etal_2013], parents’ interaction with their children [e.g., @Blewitt_etal_2009; @Yu_Smith_2012], and the quality of parental language input over time [e.g., @Raneri_etal_2020; @Rowe_2012].
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a “bowl” in English but an _“assiette”_ [plate] in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
-Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) (Bosch & Ramon-Casas, 2014).
-Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
+There is mixed evidence for whether bilingual adults maintain separate [@Jared_etal_2012] versus integrated [@Ameel_etal_2009] conceptual representations across their two languages, and little to no data from bilingual children.
+Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) [@Bosch_Ramon-Casas_2014].
+Likewise, some bilingual children learn language pairs that share more cognates than others [e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; @Schepens_etal_2013].
 While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
@@
 Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
-However, following the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+However, following the Complementarity Principle [@Grosjean_2016], bilinguals may have different experiences in each of their languages.
 For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., “soap”, “bath”, “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2016; Tardif et al., 2008).
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced [e.g., @Braginsky_etal_2019; @Tardif_etal_2008].
 For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and simple nouns (“ball”, “dog”) are among the first words children across languages and cultures.
@@
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
+
+\newpage
+
+# References
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+
+<div id = "refs"></div>
+\endgroup

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.tex
added references

@@
 \fi
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
+  % set entry spacing
+  \ifnum #2 > 0
+  \setlength{\parskip}{#2\baselineskip}
+  \fi
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
@@
 For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents {[}also called doublets; Umbel, Pearson, Fernández, and Oller (1992){]}, and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Pearson, Fernández, \& Oller, 1995).
 Translation equivalents are thought to hold a special status in a bilingual's developing lexicon due to the strong overlap in their semantics.
-For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
+For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent ( a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent.
+In vocabulary acquisition, bilingual children must learn a first label for a referent {[}a ``singlet''; Umbel, Pearson, Fernández, and Oller (1992){]} before they can learn its translation equivalent.
 Is translation equivalent learning different from singlet learning?
@@
 
-Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children's early vocabularies (e.g., Imedadze, 1978; Swain \& Wesche 1975; Volterra \& Taeschner, 1978).
+Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children's early vocabularies (e.g., Imedadze, 1967; Swain \& Wesche, 1975; Volterra \& Taeschner, 1978).
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman \& Wachtel, 1988; Markman, 1992; 1994).
+This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman, 1992, 1994; Markman \& Wachtel, 1988).
 For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like ``wug,'' they assume that ``wug'' refers to the garlic press --- the object unknown to them --- rather than to the cup, the object for which they already know the word.
@@
 
-Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David \& Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017).
+Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Holowka, Brosseau-Lapré, \& Petitto, 2002; Legacy et al., 2017; Pearson, Fernández, \& Oller, 1995).
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
-For example, when hearing a novel word like ``nil,'' monolingual children aged 17--22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein \& Werker, 2009; 2013; Houston-Price et al., 2010).
-A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al, 2020).
+For example, when hearing a novel word like ``nil,'' monolingual children aged 17--22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein \& Werker, 2009, 2013; Houston-Price, Caloghiris, \& Raviglione, 2010).
+A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis, Cristiano, Lake, Kwan, \& Frank, 2020).
 
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
-Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak \& Byers-Heinlein, 2018; Singh, 2013).
-In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
+When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; Poulin-Dubois, Kuzyk, Legacy, Zesiger, \& Friend, 2018).
+Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019; Singh, 2014).
+In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Floccia et al., 2020).
 The Preference Account predicts that translation equivalents are more easily learned than singlets.
@@
 
-Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015).
+Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson, Yoshida, Tran, Woods, \& Hills, 2015).
 The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related).
@@
 For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, \& Hurtado, 2010).
-Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
+Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernández, 1994; Pearson, Fernández, Lewedeg, \& Oller, 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado, Grüter, Marchman, \& Fernald, 2014).
 Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, \& Hurtado, 2010; Place \& Hoff, 2011; Pearson et al., 1997).
-Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
+In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2013; Boyce, Gillam, Innocenti, Cook, \& Ortiz, 2013; Hurtado, Grüter, Marchman, \& Fernald, 2014; Marchman, Fernald, \& Hurtado, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Place \& Hoff, 2011).
+Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure {[}i.e., non-dominant language; Pearson, Fernández, Lewedeg, and Oller (1997); Place and Hoff (2011){]}.
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 This is known as the Complementarity Principle (Grosjean, 2016).
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
+For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok, Luk, Peets, \& Yang, 2010).
 If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
@@
 
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman et al., 2008; Swingley \& Humphrey, 2018).
+Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado, Grüter, Marchman, \& Fernald, 2014; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman, Dale, \& Li, 2008), and family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013).
+Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman, Dale, \& Li, 2008; Swingley \& Humphrey, 2018).
 
@@
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
-For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
+For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky, Yurovsky, Marchman, \& Frank, 2019; Goodman, Dale, \& Li, 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
 Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007).
-This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky et al., 2019).
+This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky, Yurovsky, Marchman, \& Frank, 2019).
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
@@
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories {[}CDI; Fenson et al.
+(2007){]}, one option would be to set LEARNABLE to be the total number of items on the CDI.
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
@@
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
-Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
+Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words {[}retrieved from the Wordbank database; Frank, Braginsky, Yurovsky, and Marchman (2016){]}.
 Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a child of Jamie's age could potentially acquire in each language.
@@
 
-In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997).
+In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau, Frank, \& Poulin-Dubois, 1999).
 Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old {[}i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank, Braginsky, Yurovsky, and Marchman (2016){]}.
 We set BIAS to 1, meaning that in these examples translation equivalents are similarly easy to learn as singlets.
@@
 
-Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero, Schott, and Byers-Heinlein (2020).
 Some children took part in more than one in-lab study (n = 28); thus, they contributed data at more than one time point.
@@
 Bilingual children's expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
-Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
+Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau, Frank, \& Poulin-Dubois, 1999).
 We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (24\%).
@@
 
-Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero et al.~(2020) by three proficient bilingual French--English adults who carefully examined each language version of the CDI.
+Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero, Schott, and Byers-Heinlein (2020) by three proficient bilingual French--English adults who carefully examined each language version of the CDI.
 Word pairs that made reference to the same concept (e.g., English ``apple'' and French \emph{``pomme''}) were considered to be translation equivalents.
@@
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
-A full list of translation equivalents is available at {[}\url{https://osf.io/7fz6c/}{]}.
+A full list of translation equivalents is available at {[}\url{https://osf.io/2t5kw/}{]}.
 
@@
 
-Children's language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch \& Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE, Byers-Heinlein et al., 2018).
+Children's language exposure was measured using the Language Exposure Questionnaire {[}LEQ; Bosch and Sebastián-Gallés (2001){]} and the Multilingual Approach to Parent Language Estimates {[}MAPLE; Byers-Heinlein et al.
+(2020){]}.
 The LEQ is a structured interview that lasts approximately 15 minutes.
@@
 The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
+As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman, 2016), and Table 7 lists the denominator at different ages.
 For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
@@
 
+~
+
 \begin{figure}
@@
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (Davidson \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+This pattern is consistent with reports from previous research (David \& Wei, 2008; Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Montanari, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Pearson, Fernández, \& Oller, 1995).
+Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Montanari, 2010).
 Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Poulin-Dubois et al., 2013; Legacy et al., 2015).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013).
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2015), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2015).
+A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016).
 
@@
 We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
-Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
+Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson, Fernández, \& Oller, 1993).
 Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
-This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy et al., 2016).
+This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016).
 Predictions 2c and 2d were also confirmed by our vocabulary data.
 While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Place \& Hoff, 2011; Pearson et al., 1997).
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce, Gillam, Innocenti, Cook, \& Ortiz, 2013; Hoff et al., 2012; Marchman, Fernald, \& Hurtado, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Place \& Hoff, 2011).
 Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
@@
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
-This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
+This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson, Yoshida, Tran, Woods, \& Hills, 2015).
 The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
-Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
+Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Pearson, Fernández, \& Oller, 1995).
 The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children's own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
@@
 The developmental change of bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman \& Bowman, 1989).
-In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman \& Bowman, 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990, David \& Tell, 2005).
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis, Cristiano, Lake, Kwan, \& Frank, 2020; Merriman, Bowman, \& MacWhinney, 1989).
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman, Bowman, \& MacWhinney, 1989).
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis, Cristiano, Lake, Kwan, \& Frank, 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
 At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
-Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin, Alves, Orena, Polka, \& Byers-Heinlein, 2021; Nicoladis \& Secco, 2000).
 For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the livre?'' if they know their child understands the French word ``livre'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Hoff, 2003; Fernald, Marchman, \& Weisleder, 2013), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020, Rowe, 2012).
+However, our model does not consider other qualitative factors including family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013; e.g., Hoff, 2003), parents' interaction with their children (e.g., Blewitt, Rump, Shealy, \& Cook, 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri, Holzen, Newman, \& Bernstein Ratner, 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a ``bowl'' in English but an \emph{``assiette''} {[}plate{]} in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
+There is mixed evidence for whether bilingual adults maintain separate (Jared, Cormier, Levy, \& Wade-Woolley, 2012) versus integrated (Ameel, Malt, Storms, \& Van Assche, 2009) conceptual representations across their two languages, and little to no data from bilingual children.
 Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English--French pair ``banana'' -- \emph{``banane''}) more easily than those that do not share similar phonological form (e.g., the English--French pair ``dog'' -- \emph{``chien''}) (Bosch \& Ramon-Casas, 2014).
-Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
+Likewise, some bilingual children learn language pairs that share more cognates than others {[}e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens, Dijkstra, Grootjen, and Heuven (2013){]}.
 While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
@@
 For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2016; Tardif et al., 2008).
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky, Yurovsky, Marchman, \& Frank, 2019; Tardif et al., 2008).
 For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and simple nouns (``ball,'' ``dog'') are among the first words children across languages and cultures.
@@
 
+\newpage
+
+\hypertarget{references}{%
+\section{References}\label{references}}
+
+\begingroup
+\setlength{\parindent}{-0.5in}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\hypertarget{ref-Ameel_etal_2009}{}%
+Ameel, E., Malt, B.
+C., Storms, G., \& Van Assche, F.
+(2009).
+Semantic convergence in the bilingual lexicon.
+\emph{Journal of Memory and Language}, \emph{60}(2), 270--290.
+\url{https://doi.org/10.1016/j.jml.2008.10.001}
+
+\leavevmode\hypertarget{ref-Au_Glusman_1990}{}%
+Au, T.
+K.-F., \& Glusman, M.
+(1990).
+The principle of mutual exclusivity in word learning: To honor or not to honor?
+\emph{Child Development}, \emph{61}(5), 1474--1490.
+\url{https://doi.org/10.1111/j.1467-8624.1990.tb02876.x}
+
+\leavevmode\hypertarget{ref-Barnes_Garcia_2013}{}%
+Barnes, J., \& Garcia, I.
+(2013).
+Vocabulary growth and composition in monolingual and bilingual basque infants and toddlers.
+\emph{International Journal of Bilingualism}, \emph{17}(3), 357--374.
+\url{https://doi.org/10.1177/1367006912438992}
+
+\leavevmode\hypertarget{ref-Benedict_1979}{}%
+Benedict, H. (1979). Early lexical development: Comprehension and production. \emph{Journal of Child Language}, \emph{6}(2), 183--200. \url{https://doi.org/10.1017/S0305000900002245}
+
+\leavevmode\hypertarget{ref-Bergelson_2020}{}%
+Bergelson, E.
+(2020).
+The comprehension boost in early word learning: Older infants are better learners.
+\emph{Child Development Perspectives}, \emph{14}(3), 142--149.
+\url{https://doi.org/10.1111/cdep.12373}
+
+\leavevmode\hypertarget{ref-Bialystok_etal_2010}{}%
+Bialystok, E., Luk, G., Peets, K.
+F., \& Yang, S.
+(2010).
+Receptive vocabulary differences in monolingual and bilingual children.
+\emph{Bilingualism (Cambridge, England)}, \emph{13}(4), 525--531.
+\url{https://doi.org/10.1017/S1366728909990423}
+
+\leavevmode\hypertarget{ref-Bilson_etal_2015}{}%
+Bilson, S., Yoshida, H., Tran, C.
+D., Woods, E.
+A., \& Hills, T.
+T.
+(2015).
+Semantic facilitation in bilingual first language acquisition.
+\emph{Cognition}, \emph{140}, 122--134.
+\url{https://doi.org/10.1016/j.cognition.2015.03.013}
+
+\leavevmode\hypertarget{ref-Blewitt_etal_2009}{}%
+Blewitt, P., Rump, K.
+M., Shealy, S.
+E., \& Cook, S.
+A.
+(2009).
+Shared book reading: When and how questions affect young children's word learning.
+\emph{Journal of Educational Psychology}, \emph{101}(2), 294-\/-304.
+\url{https://doi.org/10.1037/a0013844}
+
+\leavevmode\hypertarget{ref-Bosch_Ramon-Casas_2014}{}%
+Bosch, L., \& Ramon-Casas, M.
+(2014).
+First translation equivalents in bilingual toddlers' expressive vocabulary: Does form similarity matter?
+\emph{International Journal of Behavioral Development}, \emph{38}(4), 317--322.
+\url{https://doi.org/10.1177/0165025414532559}
+
+\leavevmode\hypertarget{ref-Bosch_Sebastian-Galles_2001}{}%
+Bosch, L., \& Sebastián-Gallés, N.
+(2001).
+Evidence of early language discrimination abilities in infants from bilingual environments.
+\emph{Infancy}, \emph{21}(1), 29--49.
+\url{https://doi.org/10.1207/S15327078IN0201_3}
+
+\leavevmode\hypertarget{ref-Boyce_etal_2013}{}%
+Boyce, L.
+K., Gillam, S.
+L., Innocenti, M.
+S., Cook, G.
+A., \& Ortiz, E.
+(2013).
+An examination of language input and vocabulary development of young latino dual language learners living in poverty.
+\emph{First Language}, \emph{33}(6), 572--593.
+\url{https://doi.org/10.1177/0142723713503145}
+
+\leavevmode\hypertarget{ref-Braginsky_etal_2019}{}%
+Braginsky, M., Yurovsky, D., Marchman, V.
+A., \& Frank, M.
+C.
+(2019).
+Consistency and variability in children's word learning across languages.
+\emph{Open Mind}, \emph{2019}(3), 52--67.
+\url{https://doi.org/10.1162/opmi_a_00026}
+
+\leavevmode\hypertarget{ref-Brent_Siskind_2001}{}%
+Brent, M.
+R., \& Siskind, J.
+M.
+(2001).
+The role of exposure to isolated words in early vocabulary development.
+\emph{Cognition}, \emph{81}(2), B33--B44.
+\url{https://doi.org/10.1016/S0010-0277(01)00122-6}
+
+\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2020}{}%
+Byers-Heinlein, K., Schott, E., Gonzalez-Barrero, A.
+M., Brouillard, M., Dubé, D., Jardak, A., \ldots{} Tamayo, M.
+P.
+(2020).
+MAPLE: A multilingual approach to parent language estimates.
+\emph{Bilingualism: Language and Cognition}, \emph{23}(5), 951--957.
+\url{https://doi.org/10.1017/S1366728919000282}
+
+\leavevmode\hypertarget{ref-Byers-Heinlein_Werker_2009}{}%
+Byers-Heinlein, K., \& Werker, J.
+F.
+(2009).
+Monolingual, bilingual, trilingual: Infants' language experience influences the development of a word-learning heuristic.
+\emph{Developmental Science}, \emph{12}(5), 815--823.
+\url{https://doi.org/10.1111/j.1467-7687.2009.00902.x}
+
+\leavevmode\hypertarget{ref-Byers-Heinlein_Werker_2013}{}%
+Byers-Heinlein, K., \& Werker, J.
+F.
+(2013).
+Lexicon structure and the disambiguation of novel words: Evidence from bilingual infants.
+\emph{Cognition}, \emph{128}(3), 407--416.
+\url{https://doi.org/10.1016/j.cognition.2013.05.010}
+
+\leavevmode\hypertarget{ref-David_Wei_2008}{}%
+David, A., \& Wei, L.
+(2008).
+Individual differences in the lexical development of french--english bilingual children.
+\emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.
+\url{https://doi.org/10.1080/13670050802149200}
+
+\leavevmode\hypertarget{ref-Davidson_Tell_2005}{}%
+Davidson, D., \& Tell, D.
+(2005).
+Monolingual and bilingual children's use of mutual exclusivity in the naming of whole objects.
+\emph{Journal of Experimental Child Psychology}, \emph{92}(1), 25--45.
+\url{https://doi.org/10.1016/j.jecp.2005.03.007}
+
+\leavevmode\hypertarget{ref-DeHouwer_etal_2006}{}%
+De Houwer, A., Bornstein, M.
+H., \& De Coster, S.
+(2006).
+Early understanding of two words for the same thing: A CDI study of lexical comprehension in infant bilinguals.
+\emph{International Journal of Bilingual}, \emph{10}(3), 331--347.
+\url{https://doi.org/10.1177/13670069060100030401}
+
+\leavevmode\hypertarget{ref-Fenson_etal_2007}{}%
+Fenson, L., Marchman, V.
+A., Thal, D.
+J., Dale, P.
+S., Reznick, J.
+S., \& Bates, E.
+(2007).
+\emph{MacArthur-bates communicative development inventories (CDIs)} (2nd ed.).
+Baltimore, MD: Brookes Publishing.
+
+\leavevmode\hypertarget{ref-Fernald_etal_2013}{}%
+Fernald, A., Marchman, V.
+A., \& Weisleder, A.
+(2013).
+SES differences in language processing skill and vocabulary are evident at 18 monthss.
+\emph{Developmental Science}, \emph{16}, 234--248.
+\url{https://doi.org/10.1111/desc.12019}
+
+\leavevmode\hypertarget{ref-Floccia_etal_2020}{}%
+Floccia, C., Luche, C.
+D., Lepadatu, I., Chow, J., Ratnage, P., \& Plunkett, K.
+(2020).
+Translation equivalent and cross-language semantic priming in bilingual toddlers.
+\emph{Journal of Memory and Language}, \emph{112}, 104086.
+\url{https://doi.org/10.1016/j.jml.2019.104086}
+
+\leavevmode\hypertarget{ref-Frank_etal_2016}{}%
+Frank, M.
+C., Braginsky, M., Yurovsky, D., \& Marchman, V.
+A.
+(2016).
+Wordbank: An open repository for developmental vocabulary data.
+\emph{Journal of Child Language}.
+\url{https://doi.org/10.1017/S0305000916000209}
+
+\leavevmode\hypertarget{ref-Gonzalez-Barrero_2020}{}%
+Gonzalez-Barrero, A.
+M., Schott, E., \& Byers-Heinlein, K.
+(2020).
+\emph{Bilingual adjusted vocabulary: A developmentally-informed bilingual vocabulary measure}.
+Preprint.
+https://doi.org/\url{https://psyarxiv.com/x7s4u/}
+
+\leavevmode\hypertarget{ref-Goodman_etal_2008}{}%
+Goodman, J.
+C., Dale, P.
+S., \& Li, P.
+(2008).
+Does frequency count?
+Parental input and the acquisition of vocabulary.
+\emph{Journal of Child Language}, \emph{35}(3), 515--531.
+\url{https://doi.org/10.1017/S0305000907008641}
+
+\leavevmode\hypertarget{ref-Grosjean_2016}{}%
+Grosjean, F.
+(2016).
+The complementarity principle and its impact on processing, acquisition and dominance.
+In C.
+Silva-Corvalán \& J.
+Treffers-Daller (Eds.), \emph{Language dominance in bilinguals: Issues of measurement and operationalization} (pp.
+66--84).
+Cambridge: Cambridge University Press.
+
+\leavevmode\hypertarget{ref-Halberda_2003}{}%
+Halberda, J. (2003). The development of a word-learning strategy. \emph{Cognition}, \emph{87}(1), B23--B34. \url{https://doi.org/10.1016/S0010-0277(02)00186-5}
+
+\leavevmode\hypertarget{ref-Hoff_2003}{}%
+Hoff, E.
+(2003).
+The specificity of environmental influence: Socioeconomic status affects early vocabulary development via maternal speech.
+\emph{Child Development}, \emph{74}(5), 1368--1378.
+\url{https://doi.org/10.1111/1467-8624.00612}
+
+\leavevmode\hypertarget{ref-Hoff_etal_2012}{}%
+Hoff, E., Core, C., Place, S., Rumiche, R., Señor, M., \& Parra, M.
+(2012).
+Dual language exposure and early bilingual development.
+\emph{Journal of Child Language}, \emph{39}(1), 1--27.
+\url{https://doi.org/10.1017/S0305000910000759}
+
+\leavevmode\hypertarget{ref-Holowka_etal_2002}{}%
+Holowka, S., Brosseau-Lapré, F., \& Petitto, L.
+A.
+(2002).
+Semantic and conceptual knowledge underlying bilingual babies' first signs and words.
+\emph{Language Learning}, \emph{52}(2), 205--262.
+\url{https://doi.org/10.1111/0023-8333.00184}
+
+\leavevmode\hypertarget{ref-Houston-Price_etal_2010}{}%
+Houston-Price, C., Caloghiris, Z., \& Raviglione, E.
+(2010).
+Language experience shapes the development of the mutual exclusivity bias.
+\emph{Infancy}, \emph{15}(2), 125--150.
+\url{https://doi.org/10.1111/j.1532-7078.2009.00009.x}
+
+\leavevmode\hypertarget{ref-Hurtado_etal_2014}{}%
+Hurtado, N., Grüter, T., Marchman, V.
+A., \& Fernald, A.
+(2014).
+Relative language exposure, processing efficiency and vocabulary in spanish--english bilingual toddler.
+\emph{Bilingualism: Language and Cognition}, \emph{17}(1), 189--202.
+\url{https://doi.org/10.1017/S136672891300014X}
+
+\leavevmode\hypertarget{ref-Imedadze_1967}{}%
+Imedadze, N.
+V.
+(1967).
+On the psychological nature of child speech formation under condition of exposure to two languages.
+\emph{International Journal of Psychology}, \emph{2}(2), 129--132.
+\url{https://doi.org/10.1080/00207596708247209}
+
+\leavevmode\hypertarget{ref-Jardak_Byers-Heinlein_2019}{}%
+Jardak, A., \& Byers-Heinlein, K.
+(2019).
+Labels or concepts?
+The development of semantic networks in bilingual two-year-olds.
+\emph{Child Development}, \emph{90}(2), e212--e229.
+\url{https://doi.org/10.1111/cdev.13050}
+
+\leavevmode\hypertarget{ref-Jared_etal_2012}{}%
+Jared, D., Cormier, P., Levy, B.
+A., \& Wade-Woolley, L.
+(2012).
+Cross-language activation of phonology in young bilingual readers.
+\emph{Reading and Writing}, \emph{25}, 1327--1343.
+\url{https://doi.org/10.1007/s11145-011-9320-0}
+
+\leavevmode\hypertarget{ref-Kremin_etal_2021}{}%
+Kremin, L.
+V., Alves, J., Orena, A.
+J., Polka, L., \& Byers-Heinlein, K.
+(2021).
+Code-switching in parents' everyday speech to bilingual infants.
+\emph{Journal of Child Language}, \emph{First View}, 1--27.
+\url{https://doi.org/10.1017/S0305000921000118}
+
+\leavevmode\hypertarget{ref-Lanvers_1999}{}%
+Lanvers, U.
+(1999).
+Lexical growth patterns in a bilingual infant: The occurrence and significance of equivalents in the bilingual lexicon.
+\emph{International Journal of Bilingual Education and Bilingualism}, \emph{2}(1), 30--52.
+\url{https://doi.org/10.1080/13670059908666245}
+
+\leavevmode\hypertarget{ref-Legacy_etal_2017}{}%
+Legacy, J., Reider, J., Crivello, C., Kuzyk, O., Friend, M., Zesiger, P., \& Poulin-Dubois, D.
+(2017).
+Dog or chien?
+Translation equivalents in the receptive and expressive vocabularies of young french--english bilinguals.
+\emph{Journal of Child Language}, \emph{44}(4), 881--904.
+\url{https://doi.org/10.1017/S0305000916000295}
+
+\leavevmode\hypertarget{ref-Legacy_etal_2016}{}%
+Legacy, J., Zesiger, P., Friend, M., \& Poulin-Dubois, D.
+(2016).
+Vocabulary size, translation equivalents, and efficiency in word recognition in very young bilinguals.
+\emph{Journal of Child Language}, \emph{43}(4), 760--783.
+\url{https://doi.org/10.1017/S0305000915000252}
+
+\leavevmode\hypertarget{ref-Lewis_etal_2020}{}%
+Lewis, M., Cristiano, V., Lake, B.
+M., Kwan, T., \& Frank, M.
+C.
+(2020).
+The role of developmental change and linguistic experience in the mutual exclusivity effect.
+\emph{Cognition}, \emph{198}, 104191.
+\url{https://doi.org/10.1016/j.cognition.2020.104191}
+
+\leavevmode\hypertarget{ref-Marchman_etal_2010}{}%
+Marchman, V.
+A., Fernald, A., \& Hurtado, N.
+(2010).
+How vocabulary size in two languages relates to efficiency in spoken word recognition by young spanish-english bilinguals.
+\emph{Journal of Child Language}, \emph{37}(4), 817--840.
+\url{https://doi.org/10.1017/S0305000909990055}
+
+\leavevmode\hypertarget{ref-Markman_1992}{}%
+Markman, E.
+M.
+(1992).
+Constraints on word learning: Speculations about their nature, origins, and domain specificity.
+In M.
+R.
+Gunnar \& M.
+Maratsos (Eds.), \emph{The minnesota symposia on child psychology} (Vol.
+25, pp.
+59--101).
+Hillsdale, NJ: Lawrence Erlbaum Associates, Inc.
+
+\leavevmode\hypertarget{ref-Markman_1994}{}%
+Markman, E. M. (1994). Constraints on word meaning in early language acquisition. \emph{Lingua}, \emph{92}, 199--227. \url{https://doi.org/10.1016/0024-3841(94)90342-5}
+
+\leavevmode\hypertarget{ref-Markman_Wachtel_1988}{}%
+Markman, E.
+M., \& Wachtel, G.
+F.
+(1988).
+Children's use of mutual exclusivity to constrain the meanings of words.
+\emph{Cognitive Psychology}, \emph{20}(2), 121--157.
+\url{https://doi.org/10.1016/0010-0285(88)90017-5}
+
+\leavevmode\hypertarget{ref-Merriman_etal_1989}{}%
+Merriman, W.
+E., Bowman, L.
+L., \& MacWhinney, B.
+(1989).
+The mutual exclusivity bias in children's word learning.
+\emph{Monographs of the Society for Research in Child Development}, \emph{54}(3-4), 130.
+\url{https://doi.org/10.2307/1166130}
+
+\leavevmode\hypertarget{ref-Montanari_2010}{}%
+Montanari, S.
+(2010).
+Translation equivalents and the emergence of multiple lexicons in early trilingual development.
+\emph{First Language}, \emph{30}(1), 102--125.
+\url{https://doi.org/10.1177/0142723709350528}
+
+\leavevmode\hypertarget{ref-Nicoladis_Secco_2000}{}%
+Nicoladis, E., \& Secco, G.
+(2000).
+The role of a child's productive vocabulary in the language choice of a bilingual family.
+\emph{First Language}, \emph{20}(58), 3--28.
+\url{https://doi.org/10.1177/014272370002005801}
+
+\leavevmode\hypertarget{ref-Pearson_Fernandez_1994}{}%
+Pearson, B.
+Z., \& Fernández, S.
+C.
+(1994).
+Patterns of interaction in the lexical growth in two languages of bilingual infants and toddlers.
+\emph{Language Learning}, \emph{44}(4), 617--653.
+\url{https://doi.org/10.1111/j.1467-1770.1994.tb00633.x}
+
+\leavevmode\hypertarget{ref-Pearson_etal_1997}{}%
+Pearson, B. Z., Fernández, S. C., Lewedeg, V., \& Oller, D. K. (1997). The relation of input factors to lexical learning by bilingual infants. \emph{Applied Psycholinguistics}, \emph{18}(1), 41--58.
+
+\leavevmode\hypertarget{ref-Pearson_etal_1993}{}%
+Pearson, B.
+Z., Fernández, S.
+C., \& Oller, D.
+K.
+(1993).
+Lexical development in bilingual infants and toddlers: Comparison to monolingual norms.
+\emph{Language Learning}, \emph{43}(1), 93--120.
+\url{https://doi.org/10.1111/j.1467-1770.1993.tb00174.x}
+
+\leavevmode\hypertarget{ref-Pearson_etal_1995}{}%
+Pearson, B.
+Z., Fernández, S.
+C., \& Oller, D.
+K.
+\&.
+(1995).
+Cross-language synonyms in the lexicons of bilingual infants: One language or two?
+\emph{Journal of Child Language}, \emph{22}(2), 345--368.
+\url{https://doi.org/10.1017/s030500090000982x}
+
+\leavevmode\hypertarget{ref-Place_Hoff_2011}{}%
+Place, S., \& Hoff, E.
+(2011).
+Properties of dual language exposure that influence 2-year-olds' bilingual proficiency.
+\emph{Child Development}, \emph{82}(6), 1834--1849.
+\url{https://doi.org/10.1111/j.1467-8624.2011.01660.x}
+
+\leavevmode\hypertarget{ref-Poulin-Dubois_etal_2013}{}%
+Poulin-Dubois, D., Bialystok, E., Blaye, A., Polonia, A., \& Yott, J.
+(2013).
+Lexical access and vocabulary development in very young bilinguals.
+\emph{International Journal of Bilingualism}, \emph{17}(1), 57--70.
+\url{https://doi.org/10.1177/1367006911431198}
+
+\leavevmode\hypertarget{ref-Poulin-Dubois_etal_2018}{}%
+Poulin-Dubois, D., Kuzyk, O., Legacy, J., Zesiger, P., \& Friend, M.
+(2018).
+Translation equivalents facilitate lexical access in very young bilinguals.
+\emph{Bilingualism: Language and Cognition}, \emph{21}(4), 856--866.
+\url{https://doi.org/10.1017/S1366728917000657}
+
+\leavevmode\hypertarget{ref-Raneri_etal_2020}{}%
+Raneri, D., Holzen, K.
+von, Newman, R., \& Bernstein Ratner, N.
+(2020).
+Change in maternal speech rate to preverbal infants over the first two years of life.
+\emph{Journal of Child Language}, \emph{47}(6), 1263--1275.
+\url{https://doi.org/10.1017/S030500091900093X}
+
+\leavevmode\hypertarget{ref-Rowe_2012}{}%
+Rowe, M.
+L.
+(2012).
+A longitudinal investigation of the role of quantity and quality of child-directed speech in vocabulary development.
+\emph{Child Development}, \emph{83}(5), 1762--1774.
+\url{https://doi.org/10.1111/j.1467-8624.2012.01805.x}
+
+\leavevmode\hypertarget{ref-Schepens_etal_2013}{}%
+Schepens, J., Dijkstra, T., Grootjen, F., \& Heuven, W.
+J.
+B.
+van.
+(2013).
+Cross-language distributions of high frequency and phonetically similar cognates.
+\emph{PLOS ONE}, \emph{8}(5), e63006.
+\url{https://doi.org/10.1371/journal.pone.0063006}
+
+\leavevmode\hypertarget{ref-Singh_2014}{}%
+Singh, L. (2014). One world, two languages: Cross-language semantic priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2), 755--766. \url{https://doi.org/10.1111/cdev.12133}
+
+\leavevmode\hypertarget{ref-Swain_Wesche_1975}{}%
+Swain, M., \& Wesche, M. (1975). Linguistic interaction: Case study of a bilingual child. \emph{Language Sciences}, \emph{37}, 17--22.
+
+\leavevmode\hypertarget{ref-Swingley_Humphrey_2018}{}%
+Swingley, D., \& Humphrey, C.
+(2018).
+Quantitative linguistic predictors of infants' learning of specific english words.
+\emph{Child Development}, \emph{89}(4), 1247--1267.
+\url{https://doi.org/10.1111/cdev.12731}
+
+\leavevmode\hypertarget{ref-Tardif_1996}{}%
+Tardif, T.
+(1996).
+Nouns are not always learned before verbs: Evidence from mandarin speakers' early vocabularies.
+\emph{Developmental Psychology}, \emph{32}(3), 492--504.
+\url{https://doi.org/10.1037/0012-1649.32.3.492}
+
+\leavevmode\hypertarget{ref-Tardif_etal_2008}{}%
+Tardif, T., Fletcher, P., Liang, W., Zhang, Z., Kaciroti, N., \& Marchman, V.
+A.
+(2008).
+Baby's first 10 words.
+\emph{Developmental Psychology}, \emph{44}(4), 929--938.
+\url{https://doi.org/10.1037/0012-1649.44.4.929}
+
+\leavevmode\hypertarget{ref-Trudeau_etal_1999}{}%
+Trudeau, N., Frank, I., \& Poulin-Dubois, D.
+(1999).
+Une adaptation en français québecois du MacArthur communicative development inventory {[}a quebec french adaptation of the MacArthur communicative development inventory{]}.
+\emph{Revue d'orthophonie Et d'audiologie}, \emph{23}, 31--73.
+
+\leavevmode\hypertarget{ref-Umbel_etal_1992}{}%
+Umbel, V.
+M., Pearson, B.
+Z., Fernández, M.
+C., \& Oller, D.
+K.
+(1992).
+Measuring bilingual children's receptive vocabularies.
+\emph{Child Development}, \emph{63}(4), 1012--1020.
+\url{https://doi.org/10.1111/j.1467-8624.1992.tb01678.x}
+
+\leavevmode\hypertarget{ref-Volterra_Taeschner_1978}{}%
+Volterra, V., \& Taeschner, T. (1978). The acquisition and development of language by bilingual children. \emph{Journal of Child Language}, \emph{5}, 311--326.
+
+\leavevmode\hypertarget{ref-Weisleder_Fernald_2013}{}%
+Weisleder, A., \& Fernald, A. (2013). Talking to children matters: Early language experience strengthens processing and builds vocabulary. \emph{Psychological Science}, \emph{24}(11), 2143--2152.
+
+\leavevmode\hypertarget{ref-Yu_Smith_2012}{}%
+Yu, C., \& Smith, L. (2012). Embodied attention and word learning by toddlers. \emph{Cognition}, \emph{125}(2), 244--262.
+
+\end{CSLReferences}
+
+\endgroup
+
 \end{document}

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
fixed in-text citations (apa7)

--- 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
 
-#bibliography      : ["gaze-following_references.bib"]
-#replace_ampersands: yes
+bibliography      : ["paper_TE_bilingual_vocabulary_model_references.bib"]
+replace_ampersands: yes
 
 class             : "man"
-output            : papaja::apa6_pdf
+output            : papaja::apa6_pdf # change to apa6_docx for word document
+citation_package  : biblatex
+nocite: '@*' # include all the references in the .bib file, because in-text citations were mostly typed manually to fit the apa7 style
 ---
@@
 For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer, Bornstein, & De Coster, 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics.
@@
 That is, upon hearing the English word “apple”, the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent ( a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent.
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent.
 Is translation equivalent learning different from singlet learning?
@@
 ### Avoidance Account: Translation equivalents are harder to learn than singlets
-Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies (e.g., Imedadze, 1978; Swain & Wesche 1975; Volterra & Taeschner, 1978).
+Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies [e.g., @Imedadze_1967; @Swain_Wesche_1975; @Volterra_Taeschner_1978].
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman & Wachtel, 1988; Markman, 1992; 1994).
+This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level [@Markman_Wachtel_1988; @Markman_1992; @Markman_1994].
 For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug”, they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.
 
-Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein & Werker, 2009; Davidson & Tell, 2005; Houston-Price, Caloghiris, & Raviglione, 2010).
+Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein & Werker, 2009; Davidson & Tell, 2005; Houston-Price et al., 2010).
 When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word.
@@
 
-Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David & Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Pearson et al., 1995; Legacy et al., 2017).
+Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David & Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Legacy et al., 2017; Pearson et al., 1995).
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
 For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010).
-A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al, 2020).
+A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al., 2020).
 
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2017).
-Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak & Byers-Heinlein, 2018; Singh, 2013).
+When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
+Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related [e.g., @Floccia_etal_2020; @Jardak_Byers-Heinlein_2019; @Singh_2014].
 In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
@@
 There are several lines of empirical evidence to support the Preference Account.
-For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages (Lanvers, 1999; Pearson & Fernández, 1994).
+For example, some early case studies reported that bilinguals tended to learn more translation equivalents than singlets when experiencing a shift in their language exposure that inverted their dominant and non-dominant languages [@Lanvers_1999; @Pearson_Fernandez_1994].
 The main explanation that has been given for this finding is that additional exposure to their non-dominant language — which became their new dominant language — enabled fast mapping of words to already-lexicalized concepts.
@@
 Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
-For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, & Hurtado, 2010).
-Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernández, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2013).
+For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman et al., 2010).
+Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernandez, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2014).
 Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2012; Boyce et al., 2013; Hurtado et al., 2013; Marchman, Fernald, & Hurtado, 2010; Place & Hoff, 2011; Pearson et al., 1997).
+In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011).
 Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
@@
 One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts.
-This is known as the Complementarity Principle (Grosjean, 2016).
+This is known as the Complementarity Principle [@Grosjean_2016].
 For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
@@
 
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), and family socioeconomic status (Fernald, Marchman, & Weisleder, 2013).
+Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech [e.g., @Brent_Siskind_2001; @Swingley_Humphrey_2018], children’s efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), and family socioeconomic status (Fernald et al., 2013).
 Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).
@@
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
-For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
-Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007).
+For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired [@Tardif_1996].
+Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time [@Fenson_etal_2007].
 This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky et al., 2019).
@@
 
-In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1997).
+In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English [@Fenson_etal_2007] and Québec French (Trudeau et al., 1999).
 Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
@@
 Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right).
-Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="80%", fig.pos = "H"}
+Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.width="80%", fig.pos = "H"}
 
@@
 Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
-Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1997).
+Caregivers completed the original CDI English version [@Fenson_etal_2007] and its Québec French adaptation (Trudeau et al., 1999).
 We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), and respondent not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
@@
 ### Language Exposure Questionnaire using the MAPLE approach
-Children’s language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE, Byers-Heinlein et al., 2018).
+Children’s language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE; Byers-Heinlein et al., 2020).
 The LEQ is a structured interview that lasts approximately 15 minutes.
@@
 The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
-The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="70%", out.width="70%"}
+The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="60%", out.width="60%"}
 
@@
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (Davidson & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+This pattern is consistent with reports from previous research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
 Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
 Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Poulin-Dubois et al., 2013; Legacy et al., 2015).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2015), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2015).
+A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
 
@@
 While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Place & Hoff, 2011; Pearson et al., 1997).
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011).
 Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
@@
 The developmental change of bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman & Bowman, 1989).
-In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman & Bowman, 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au & Glusman, 1990, David & Tell, 2005).
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
 At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Hoff, 2003; Fernald, Marchman, & Weisleder, 2013), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020, Rowe, 2012).
+However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
-Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) (Bosch & Ramon-Casas, 2014).
+Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) [@Bosch_Ramon-Casas_2014].
 Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
@@
 Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
-However, following the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+However, following the Complementarity Principle [@Grosjean_2016], bilinguals may have different experiences in each of their languages.
 For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., “soap”, “bath”, “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2016; Tardif et al., 2008).
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
 For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and simple nouns (“ball”, “dog”) are among the first words children across languages and cultures.

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.tex
fixed in-text citations (apa7)

@@
 For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
-These cross-language synonyms are known as translation equivalents {[}also called doublets; Umbel, Pearson, Fernández, and Oller (1992){]}, and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Pearson, Fernández, \& Oller, 1995).
+These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual's developing lexicon due to the strong overlap in their semantics.
-For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
+For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent {[}a ``singlet''; Umbel, Pearson, Fernández, and Oller (1992){]} before they can learn its translation equivalent.
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent.
 Is translation equivalent learning different from singlet learning?
@@
 
-Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein \& Werker, 2009; Davidson \& Tell, 2005; Houston-Price, Caloghiris, \& Raviglione, 2010).
+Although mutual exclusivity is helpful for monolingual vocabulary acquisition, its use is more complex for bilingual vocabulary acquisition (Byers-Heinlein \& Werker, 2009; Davidson \& Tell, 2005; Houston-Price et al., 2010).
 When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word.
@@
 
-Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David \& Wei, 2008; De Houwer, Bornstein, \& De Coster, 2006; Holowka, Brosseau-Lapré, \& Petitto, 2002; Legacy et al., 2017; Pearson, Fernández, \& Oller, 1995).
+Contrary to earlier studies, more recent work has indicated that bilinguals do understand and produce translation equivalents from early in development (David \& Wei, 2008; De Houwer et al., 2006; Holowka et al., 2002; Legacy et al., 2017; Pearson et al., 1995).
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
-For example, when hearing a novel word like ``nil,'' monolingual children aged 17--22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein \& Werker, 2009, 2013; Houston-Price, Caloghiris, \& Raviglione, 2010).
-A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis, Cristiano, Lake, Kwan, \& Frank, 2020).
+For example, when hearing a novel word like ``nil,'' monolingual children aged 17--22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein \& Werker, 2009; 2013; Houston-Price et al., 2010).
+A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al., 2020).
 
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013; Poulin-Dubois, Kuzyk, Legacy, Zesiger, \& Friend, 2018).
+When a French--English bilingual child encounters the word \emph{``pomme''} after having learned ``apple,'' one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
 Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related (e.g., Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019; Singh, 2014).
-In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Floccia et al., 2020).
+In this context, the strong semantic overlap makes translation equivalents special, and could facilitate their acquisition (e.g., Bilson et al., 2015; Floccia et al., 2020).
 The Preference Account predicts that translation equivalents are more easily learned than singlets.
@@
 
-Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson, Yoshida, Tran, Woods, \& Hills, 2015).
+Other evidence suggesting translation equivalents might be easier to learn than singlets comes from a study that included vocabulary-checklist data from 254 monolingual and 181 bilingual children aged 6 months to 7.5 years (Bilson et al., 2015).
 The researchers used a network analysis approach to investigate how translation equivalents are learned, focusing on the semantic relationships between the words (e.g., words like ``cat'' and ``dog'' are strongly semantically related).
@@
 Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
-For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman, Fernald, \& Hurtado, 2010).
-Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernández, 1994; Pearson, Fernández, Lewedeg, \& Oller, 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado, Grüter, Marchman, \& Fernald, 2014).
+For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman et al., 2010).
+Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernandez, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2014).
 Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2013; Boyce, Gillam, Innocenti, Cook, \& Ortiz, 2013; Hurtado, Grüter, Marchman, \& Fernald, 2014; Marchman, Fernald, \& Hurtado, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Place \& Hoff, 2011).
-Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure {[}i.e., non-dominant language; Pearson, Fernández, Lewedeg, and Oller (1997); Place and Hoff (2011){]}.
+In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011).
+Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 This is known as the Complementarity Principle (Grosjean, 2016).
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok, Luk, Peets, \& Yang, 2010).
+For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
 If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
@@
 
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado, Grüter, Marchman, \& Fernald, 2014; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman, Dale, \& Li, 2008), and family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman, Dale, \& Li, 2008; Swingley \& Humphrey, 2018).
+Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), and family socioeconomic status (Fernald et al., 2013).
+Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman et al., 2008; Swingley \& Humphrey, 2018).
 
@@
 Evidence from monolingual children shows that some types of words are characteristically learned before others.
-For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky, Yurovsky, Marchman, \& Frank, 2019; Goodman, Dale, \& Li, 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
+For example, across many languages including English, children show a noun bias in their early lexicons (Braginsky et al., 2019; Goodman et al., 2008), although for other languages such as Mandarin it appears that verbs and nouns are more equally acquired (Tardif, 1996).
 Certain classes of words are rarely known at the onset of lexical development, such as prepositions and words for time (Fenson et al., 2007).
-This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky, Yurovsky, Marchman, \& Frank, 2019).
+This is thought to be due to the cognitive and linguistic machinery that must be in place in order for children to represent these concepts, a necessary prerequisite for learning certain word types (Bergelson, 2020; Braginsky et al., 2019).
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
@@
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories {[}CDI; Fenson et al.
-(2007){]}, one option would be to set LEARNABLE to be the total number of items on the CDI.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI.
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
@@
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
-Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words {[}retrieved from the Wordbank database; Frank, Braginsky, Yurovsky, and Marchman (2016){]}.
+Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
 Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a child of Jamie's age could potentially acquire in each language.
@@
 
-In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau, Frank, \& Poulin-Dubois, 1999).
+In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1999).
 Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old {[}i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank, Braginsky, Yurovsky, and Marchman (2016){]}.
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
 We set BIAS to 1, meaning that in these examples translation equivalents are similarly easy to learn as singlets.
@@
 
-{\centering \includegraphics[width=0.8\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig3-1}
+{\centering \includegraphics[width=0.8\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig3-1}
 
@@
 
-Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero, Schott, and Byers-Heinlein (2020).
+Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
 Some children took part in more than one in-lab study (n = 28); thus, they contributed data at more than one time point.
@@
 Bilingual children's expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
-Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau, Frank, \& Poulin-Dubois, 1999).
+Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1999).
 We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (24\%).
@@
 
-Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero, Schott, and Byers-Heinlein (2020) by three proficient bilingual French--English adults who carefully examined each language version of the CDI.
+Translation equivalents (TE) were determined in the same manner as Gonzalez-Barrero et al.~(2020) by three proficient bilingual French--English adults who carefully examined each language version of the CDI.
 Word pairs that made reference to the same concept (e.g., English ``apple'' and French \emph{``pomme''}) were considered to be translation equivalents.
@@
 
-Children's language exposure was measured using the Language Exposure Questionnaire {[}LEQ; Bosch and Sebastián-Gallés (2001){]} and the Multilingual Approach to Parent Language Estimates {[}MAPLE; Byers-Heinlein et al.
-(2020){]}.
+Children's language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch \& Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE; Byers-Heinlein et al., 2020).
 The LEQ is a structured interview that lasts approximately 15 minutes.
@@
 The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank, Braginsky, Yurovsky, \& Marchman, 2016), and Table 7 lists the denominator at different ages.
+As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
 For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
@@
 
-{\centering \includegraphics[width=0.7\linewidth,height=0.7\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig6-1}
+{\centering \includegraphics[width=0.6\linewidth,height=0.6\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig6-1}
 
@@
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (David \& Wei, 2008; Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Montanari, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Pearson, Fernández, \& Oller, 1995).
-Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Montanari, 2010).
+This pattern is consistent with reports from previous research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
 Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016; Poulin-Dubois, Bialystok, Blaye, Polonia, \& Yott, 2013).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016).
+A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
 
@@
 We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
-Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson, Fernández, \& Oller, 1993).
+Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
 Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
-This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy, Zesiger, Friend, \& Poulin-Dubois, 2016).
+This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy et al., 2016).
 Predictions 2c and 2d were also confirmed by our vocabulary data.
 While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce, Gillam, Innocenti, Cook, \& Ortiz, 2013; Hoff et al., 2012; Marchman, Fernald, \& Hurtado, 2010; Pearson, Fernández, Lewedeg, \& Oller, 1997; Place \& Hoff, 2011).
+These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011).
 Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
@@
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
-This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson, Yoshida, Tran, Woods, \& Hills, 2015).
+This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
 The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
-Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson, Yoshida, Tran, Woods, \& Hills, 2015; Pearson, Fernández, \& Oller, 1995).
+Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
 The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children's own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
@@
 The developmental change of bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis, Cristiano, Lake, Kwan, \& Frank, 2020; Merriman, Bowman, \& MacWhinney, 1989).
-In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman, Bowman, \& MacWhinney, 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis, Cristiano, Lake, Kwan, \& Frank, 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
+As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
+In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
+In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
 At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
-Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin, Alves, Orena, Polka, \& Byers-Heinlein, 2021; Nicoladis \& Secco, 2000).
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
 For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the livre?'' if they know their child understands the French word ``livre'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (Fernald, Marchman, \& Weisleder, 2013; e.g., Hoff, 2003), parents' interaction with their children (e.g., Blewitt, Rump, Shealy, \& Cook, 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri, Holzen, Newman, \& Bernstein Ratner, 2020; Rowe, 2012).
+However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a ``bowl'' in English but an \emph{``assiette''} {[}plate{]} in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared, Cormier, Levy, \& Wade-Woolley, 2012) versus integrated (Ameel, Malt, Storms, \& Van Assche, 2009) conceptual representations across their two languages, and little to no data from bilingual children.
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
 Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English--French pair ``banana'' -- \emph{``banane''}) more easily than those that do not share similar phonological form (e.g., the English--French pair ``dog'' -- \emph{``chien''}) (Bosch \& Ramon-Casas, 2014).
-Likewise, some bilingual children learn language pairs that share more cognates than others {[}e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens, Dijkstra, Grootjen, and Heuven (2013){]}.
+Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
 While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
@@
 For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky, Yurovsky, Marchman, \& Frank, 2019; Tardif et al., 2008).
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
 For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and simple nouns (``ball,'' ``dog'') are among the first words children across languages and cultures.
@@
 Benedict, H. (1979). Early lexical development: Comprehension and production. \emph{Journal of Child Language}, \emph{6}(2), 183--200. \url{https://doi.org/10.1017/S0305000900002245}
+
+\leavevmode\hypertarget{ref-Benjamini_Yekutieli_2001}{}%
+Benjamini, Y., \& Yekutieli, D.
+(2001).
+The control of the false discovery rate in multiple testing under dependency.
+\emph{The Annals of Statistics}, \emph{29}(4), 1165--1188.
+Retrieved from \url{http://www.jstor.org/stable/2674075}
 

### 2021-06-15 simulations_old_versions/Analyze_TEs.Rmd
organizing folders

--- 2021-06-03 Analyze_TEs.Rmd
@@
 write.csv(mycor, "Output/ObservedTE_correlations.csv", col.names=NA)
-
-cor.test(Correlation$TE, Correlation$B, method = "pearson")
-cor.test(Correlation$TE, Correlation$W, method = "pearson")
-cor.test(Correlation$TE, Correlation$D, method = "pearson")
-cor.test(Correlation$TE, Correlation$N, method = "pearson")
-cor.test(Correlation$TE, Correlation$V, method = "pearson")
-cor.test(Correlation$DI, Correlation$B, method = "pearson")
-cor.test(Correlation$NI, Correlation$B, method = "pearson")
-
 ```
@@
   mutate_at(vars(-BALANCE), round, 0) %>%
-  # bin B into 6 groups
-  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
+  # bin B into 5 groups
+  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
@@
   scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
-  #scale_color_manual(values=c("#5e4fa2", "#3288bd", "#abdda4", "#fee08b", "#fc8d59", "#d53e4f")) + 
-  #scale_color_discrete(direction = -1) +
   theme_minimal() + 
@@
                              word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  #mutate(bin_B = case_when(balance_vocab < 0.2 ~ "0.1",
-  #                         balance_vocab >= 0.2 & balance_vocab < 0.3 ~ "0.2",
-  #                         balance_vocab >= 0.3 & balance_vocab < 0.4 ~ "0.3",
-  #                         balance_vocab >= 0.4 & balance_vocab < 0.5 ~ "0.5",
-  #                         TRUE ~ "0.5")) %>%
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) 
   ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
 
-### one-way ANOVA to compare the 5 balance groups
+# testing Prediction 1a: To compare TE among the 5 BALANCE groups 
+## correlation between TE and BALANCE
+cor.test(Correlation$TE, Correlation$BALANCE, method = "pearson")
+
+## one-way ANOVA
 aov_TE_binBALANCE <- keepers_ws_TE %>% 
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
@@
 check <- keepers_ws_TE %>% 
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))
 
-pairwise.t.test(check$number_of_te, check$bin_BALANCE)
+pairwise.t.test(check$number_of_te, check$bin_BALANCE,
+                p.adjust.method = "BH")
 
 mean_TE_binBALANCE <- keepers_ws_TE %>% 
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = 5, 
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
   group_by(bin_BALANCE) %>%
-  summarize(n = n(),
+  summarise(n = n(),
             mean_TE = mean(number_of_te, na.rm = T),
@@
             min_TE = min(number_of_te, na.rm = T),
-            max_TE = max(number_of_te, na.rm = T),)
-
+            max_TE = max(number_of_te, na.rm = T)) %>%
+  mutate(across(!bin_BALANCE & !sd_TE, ~ round(.,0))) %>%
+  mutate_at(vars(sd_TE), ~ round(.,2))
+
+# testing Prediction 1b: To compare TE and WORD or DOM
+cor.test(Correlation$TE, Correlation$WORD, method = "pearson")
+cor.test(Correlation$TE, Correlation$DOM, method = "pearson")
+
+# testing Prediction 1c: To compare TE and NONDOM
+cor.test(Correlation$TE, Correlation$NONDOM, method = "pearson")
 ```
@@
 
-# One-way Anova to compare the 3 V groups
-keepers_ws_TE_3Vsubset <- keepers_ws_TE %>%
+# testing Prediction 2a: To compare CONCEPT among the 3 LEARNABLE groups in the stacked bar chart
+## correlation between age (in months) and CONCEPT
+cor.test(Correlation$CONCEPT, Correlation$Age_month, method = "pearson")
+
+## one-way ANOVE to compare CONCEPT among LEARNABLE
+keepers_ws_TE_3LEARNABLEsubset <- keepers_ws_TE %>%
   # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
+  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
                               age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
@@
 
-summary(aov(concept_vocab ~ V_subset, data = keepers_ws_TE_3Vsubset))
-
-mean_conceptV_Vsubset <- keepers_ws_TE_3Vsubset %>%
-  group_by(V_subset) %>%
-  summarise(mean_conceptV = mean(concept_vocab, na.rm = T),
-            sd_conceptV = sd(concept_vocab, na.rm = T),
-            min_conceptV = min(concept_vocab, na.rm = T),
-            max_conceptV = max(concept_vocab, na.rm = T))
-
-pairwise.t.test(keepers_ws_TE_3Vsubset$concept_vocab, keepers_ws_TE_3Vsubset$V_subset,
+summary(aov(concept_vocab ~ LEARNABLE_subset, data = keepers_ws_TE_3LEARNABLEsubset))
+
+mean_concept_LEARNABLEsubset <- keepers_ws_TE_3LEARNABLEsubset %>%
+  group_by(LEARNABLE_subset) %>%
+  summarise(mean_concept = mean(concept_vocab, na.rm = T),
+            sd_concept = sd(concept_vocab, na.rm = T),
+            min_concept = min(concept_vocab, na.rm = T),
+            max_concept = max(concept_vocab, na.rm = T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+pairwise.t.test(keepers_ws_TE_3LEARNABLEsubset$concept_vocab, keepers_ws_TE_3LEARNABLEsubset$LEARNABLE_subset,
                 p.adjust.method = "BH")
 
-## Mean proportion of TE
-stacked_byV_wide <- keepers_ws_TE %>%
+# testing Prediction 2b: To compare TE among the 3 LEARNABLE groups in the stacked bar chart
+## one-way ANOVA to compare TE among LEARNABLE
+stacked_byLEARNABLE_wide <- keepers_ws_TE %>%
   select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
@@
   # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(V_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
+  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
                               age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
@@
 
-stacked_byV_wide %>%
-  group_by(V_subset) %>%
-  summarise(mean_propTE = mean(propTE, na.rm=T), 
-            sd_propTE = sd(propTE, na.rm=T),
-            min_propTE = min(propTE, na.rm=T),
-            max_propTE = max(propTE, na.rm=T))
-
-# Anova to compare the number of TEs among the 3 developmental groups in the stacked barchart
-stacked_byV_wide %>%  
-  group_by(V_subset) %>%
+TE_byLEARNABLE_anova <- aov(number_of_te ~ LEARNABLE_subset, data = stacked_byLEARNABLE_wide)
+summary(TE_byLEARNABLE_anova)
+
+pairwise.t.test(stacked_byLEARNABLE_wide$number_of_te, stacked_byLEARNABLE_wide$LEARNABLE_subset,
+                 p.adjust.method = "BH")
+
+mean_TE_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
+  group_by(LEARNABLE_subset) %>%
   summarise(mean_TE = mean(number_of_te, na.rm=T), 
@@
             min_TE = min(number_of_te, na.rm=T),
-            max_TE = max(number_of_te, na.rm=T))
-
-TE_byV_anova <- aov(number_of_te ~ V_subset, data = stacked_byV_wide)
-summary(TE_byV_anova)
-
-pairwise.t.test(stacked_byV_wide$number_of_te, stacked_byV_wide$V_subset,
+            max_TE = max(number_of_te, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+## correlation between age (in months) and TE
+cor.test(Correlation$TE, Correlation$Age_month, method = "pearson")
+
+# testing Prediction 2c: One-way Anova to compare DOM-SINGLET among the 3 LEARNABLE groups in the stacked bar chart
+## correlation between DOM-SINGLET and age (in month)
+cor.test(Correlation$"DOM-SINGLET", Correlation$Age_month, method = "pearson")
+
+## correlation between DOM-SINGLET and BALANCE
+cor.test(Correlation$"DOM-SINGLET", Correlation$BALANCE, method = "pearson")
+
+## one-way ANOVA to compare DOM-SINGLET among LEARNABLE
+DOMSINGLET_byBALANCE_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byLEARNABLE_wide)
+summary(DOMSINGLET_byBALANCE_anova)
+
+pairwise.t.test(stacked_byLEARNABLE_wide$Dom_unique_words, stacked_byLEARNABLE_wide$BalSubset,
                  p.adjust.method = "BH")
 
-## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 developmental groups in the stacked barchart
-stacked_byV_wide %>%  
+mean_DOMSINGLET_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
   group_by(BalSubset) %>%
-  summarise(mean_DI = mean(Dom_unique_words, na.rm=T), 
-            sd_DI = sd(Dom_unique_words, na.rm=T),
-            min_DI = min(Dom_unique_words, na.rm=T),
-            max_DI = max(Dom_unique_words, na.rm=T))
-
-DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_byV_wide)
-summary(DI_byB_anova)
-
-pairwise.t.test(stacked_byV_wide$Dom_unique_words, stacked_byV_wide$BalSubset,
+  summarise(mean_DOM_SINGLET = mean(Dom_unique_words, na.rm=T), 
+            sd_DOM_SINGLET = sd(Dom_unique_words, na.rm=T),
+            min_DOM_SINGLET = min(Dom_unique_words, na.rm=T),
+            max_DOM_SINGLET = max(Dom_unique_words, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
+
+# testing Prediction 2d: One-way Anova to compare NONDOM-SINGLET among the 3 LEARNABLE groups in the stacked barchart
+## correlation between NONDOM-SINGLET and age (in month)
+cor.test(Correlation$"NONDOM-SINGLET", Correlation$Age_month, method = "pearson")
+
+## correlation between NONDOM-SINGLET and BALANCE
+cor.test(Correlation$"NONDOM-SINGLET", Correlation$BALANCE, method = "pearson")
+
+## one-way ANOVA to compare NONDOM-SINGLET among LEARNABLE
+NONDOMSINGLET_byBALANCE_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byLEARNABLE_wide)
+summary(NONDOMSINGLET_byBALANCE_anova)
+
+pairwise.t.test(stacked_byLEARNABLE_wide$NonDom_unique_words, stacked_byLEARNABLE_wide$BalSubset,
                  p.adjust.method = "BH")
 
-stacked_byV_wide %>% 
+mean_NONDOMSINGLET_LEARNABLEsubset <-stacked_byLEARNABLE_wide %>% 
   group_by(BalSubset) %>%
-  summarise(mean_NI = mean(NonDom_unique_words, na.rm=T), 
-            sd_NI = sd(NonDom_unique_words, na.rm=T),
-            min_NI = min(NonDom_unique_words, na.rm=T),
-            max_NI = max(NonDom_unique_words, na.rm=T))
-
-NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_byV_wide)
-summary(NI_byB_anova)
-
-pairwise.t.test(stacked_byV_wide$NonDom_unique_words, stacked_byV_wide$BalSubset,
-                 p.adjust.method = "BH")
-
+  summarise(mean_NONDOM_SINGLET = mean(NonDom_unique_words, na.rm=T), 
+            sd_NONDOM_SINGLET = sd(NonDom_unique_words, na.rm=T),
+            min_NONDOM_SINGLET = min(NonDom_unique_words, na.rm=T),
+            max_NONDOM_SINGLET = max(NonDom_unique_words, na.rm=T)) %>%
+  mutate(across(is.numeric, ~ round(.,1)))
 ```
@@
   mutate_at(vars(-BALANCE), round, 0) %>%
-  mutate(bin_BALANCE = cut(BALANCE, breaks = 5, 
+  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
@@
 ```{r}
-## Create predicted TEs for simulation
+# testing Prediction 3: linear regression model comparing simulated and observed TE
+## creating predicted TEs for simulation
 keepers_ws_TE_simulation <- keepers_ws_TE %>%
@@
 
-## LM model
+##linear regression model
 Percentile_model <- keepers_ws_TE_simulation %>%
@@
      data = .)
-
 summary(Percentile_model)
 
-## Ploting Y = observed rawTE  & X = Predicted rawTE 
-TEprediction_percentile <- keepers_ws_TE_simulation %>%
+## ploting Y = observed rawTE  & X = Predicted rawTE 
+plot_TEprediction_percentile <- keepers_ws_TE_simulation %>%
   ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
-#  facet_grid(. ~TEgroup) +
   labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
@@
   geom_point(shape = 23) +
-#  geom_text(aes(label=ID_testdate)) +
   geom_abline(linetype = 2, alpha = 0.5) +
@@
 
-ggsave(here("Output/TEprediction_percentile.png"), TEprediction_percentile,
+ggsave(here("Output/plot_Simulation3_data.png"), plot_TEprediction_percentile,
        height=5, width=5)
 
-## Identifying outliers
+## identifying outliers
 CheckOutliers_model <- keepers_ws_TE_simulation %>%
-  lm(number_of_te ~ 0 + Predicted_TE_percentile, 
-#  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom,
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom,
      data = .)
@@
 
-## Ploting Y = observed rawTE  & X = Predicted rawTE 
-TEprediction_percentile_NoOutliers <- keepers_ws_TE_simulation_removeCD %>%
-  ggplot(aes(x = Predicted_TE_percentile, y = number_of_te)) +
-#  facet_grid(. ~TEgroup) +
-  labs(x = "Number of TEs predicted by the Bilingual Vocabulary Model", y = "Observed total number of TEs") +
-  theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
-  xlim(-10,430) +
-  ylim(-10,430) +
-  theme_minimal()
-
-Percentile_model_NoOutliers_LessThan300W <- keepers_ws_TE_simulation_removeCD %>%
+# splitting the model for children with WORD < 300 and those with WORD > 300
+## linear regression model for those with WORD > 300
+Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300) %>%
+  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
+     data = .)
+summary(Percentile_model_NoOutliers_MoreThan300W)
+
+## linear regression model for those with WORD < 300
+Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
   filter(word_vocab < 300) %>%
@@
      data = .)
-
 summary(Percentile_model_NoOutliers_LessThan300W)
 
-Percentile_model_NoOutliers_MoreThan300W <- keepers_ws_TE_simulation_removeCD %>%
-  filter(word_vocab > 300) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_NoOutliers_MoreThan300W)
-
-## Looking closer at the infants with total vocabulary < 300
-
-keepers_ws_TE_simulation_LessThan300 <- keepers_ws_TE_simulation %>%
-  filter(word_vocab < 300)
-
-stats::ks.test(keepers_ws_TE_simulation_LessThan300$number_of_te,
-               keepers_ws_TE_simulation_LessThan300$Predicted_TE_percentile)
-
-keepers_ws_TE_simulation_MoreThan300 <- keepers_ws_TE_simulation %>%
-  filter(word_vocab > 300)
-
-stats::ks.test(keepers_ws_TE_simulation_MoreThan300$number_of_te,
-               keepers_ws_TE_simulation_MoreThan300$Predicted_TE_percentile)
-
-# Plot
+## plotting the two separate models (those with WORD < 300 + those with WORD > 300)
 TEprediction_percentile_facet_lessW300 <- keepers_ws_TE_simulation %>%
@@
 
-ggsave(here("Output/TEprediction_percentile_facetW.png"), TEprediction_percentile_facetByW300,
+ggsave(here("Output/plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
        height=5,width=10)
-
-Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
-  filter(word_vocab < 300) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_LessThan300W)
-
-Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
-  filter(word_vocab > 300) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_MoreThan300W)
-
-age_info <- keepers_ws_TE %>%
-  mutate(group = if_else(word_vocab < 300, "LessThan300W", "MoreThan300W")) %>%
-  group_by(group) %>%
-  summarize(n = n(),
-             mean_age = mean(age_continuous, na.rm = T),
-             sd_age = sd(age_continuous, na.rm = T),
-             min_age = min(age_continuous, na.rm = T),
-             max_age = max(age_continuous, na.rm = T))
-
-Percentile_model_LessThan300W <- keepers_ws_TE_simulation %>%
-  filter(number_of_te < 130) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_LessThan300W)
-
-Percentile_model_MoreThan300W <- keepers_ws_TE_simulation %>%
-  filter(number_of_te > 130) %>%
-  lm(number_of_te*(Average_90percentile) ~ 0 + TotalNW_Dom:TotalNW_NonDom, 
-     data = .)
-
-summary(Percentile_model_MoreThan300W)
-
-```
-
-# Figure 5: Stacked barchart to plot TE, DI, and NI of individual infants (facet by balance)
-```{r}
-stacked <- keepers_ws_TE %>%
-  select(baby_id, balance_vocab, word_vocab, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
-  # Subset balance_vocab by 3 levels 
-  #mutate(BalSubset = cut(balance_vocab, breaks = 3, 
-  #                       labels=c("MostBalanced", "MediumBalanced", "LeastBalanced"))) %>%
-  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "MostBalanced",
-                               balance_vocab <= 0.20 ~ "LeastBalanced",
-                               TRUE ~ "MediumBalanced")) %>%
-  arrange(desc(number_of_te)) %>%
-  mutate(TE = number_of_te) %>%
-  pivot_longer(-c(baby_id, balance_vocab, BalSubset, word_vocab, TE), 
-               names_to = "Type", values_to = "NumberOfWords") %>%
-  mutate(Type = factor(Type, levels=c("NonDom_unique_words", "Dom_unique_words", "number_of_te"))) %>%
-  mutate(Type = recode(Type, 
-                       NonDom_unique_words = "Initial non-dominant word", 
-                       Dom_unique_words = "Initial dominant word", 
-                       number_of_te = "Translation equivalent"))
-
-LeastBal_plot <- stacked %>%
-  filter(BalSubset == "LeastBalanced") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = "Least balanced") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-MediumBal_plot <- stacked %>%
-  filter(BalSubset == "MediumBalanced") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = "Medium balanced") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-MostBal_plot <- stacked %>%
-  filter(BalSubset == "MostBalanced") %>%
-  ggplot(aes(x = reorder(baby_id, -NumberOfWords, sum), y = NumberOfWords)) +
-  geom_col(aes(fill = Type)) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_y_continuous(limits = c(0, 1050)) +
-  labs(x = "Individual infant", 
-       y = "Number of concepts produced",
-       title = "Most balanced") + 
-  theme_minimal() + 
-  theme(plot.title = element_text(hjust = 0.5))
-
-Bal_StackedPlot <- ggarrange(LeastBal_plot, 
-          MediumBal_plot + rremove("ylab"), 
-          MostBal_plot + rremove("ylab"),
-          ncol = 3, nrow = 1,
-          common.legend = TRUE, legend="bottom")
-
-ggsave("Output/StackedPlot_byBalance.png", Bal_StackedPlot,
-       width = 10, height = 5)
-
-## Anova to compare the total vocab size among the 3 balance groups in the stacked barchart
-stacked_wide <- keepers_ws_TE %>%
-  select(baby_id, balance_vocab, word_vocab, concept_vocab, propTE, number_of_te, Dom_unique_words, NonDom_unique_words) %>%
-  # Subset balance_vocab by 3 levels 
-  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "MostBalanced",
-                               balance_vocab <= 0.20 ~ "LeastBalanced",
-                               TRUE ~ "MediumBalanced"))
-
-W_byB_anova <- aov(word_vocab ~ BalSubset, data = stacked_wide)
-summary(W_byB_anova)
-
-pairwise.t.test(stacked_wide$word_vocab, stacked_wide$BalSubset, p.adj = "BH")
-
-## Mean proportion of TE
-stacked_wide %>%
-  group_by(BalSubset) %>%
-  summarise(mean_propTE = mean(propTE, na.rm=T), 
-            sd_propTE = sd(propTE, na.rm=T),
-            min_propTE = min(propTE, na.rm=T),
-            max_propTE = max(propTE, na.rm=T))
-
-## Correlation between concept vocab and TE across different balance group
-stacked_wide_most <- stacked_wide %>%
-  filter(BalSubset == "MostBalanced")
-
-cor.test(stacked_wide_most$number_of_te, stacked_wide_most$concept_vocab, method = "pearson")
-
-stacked_wide_med <- stacked_wide %>%
-  filter(BalSubset == "MediumBalanced")
-
-cor.test(stacked_wide_med$number_of_te, stacked_wide_med$concept_vocab, method = "pearson")
-
-stacked_wide_least <- stacked_wide %>%
-  filter(BalSubset == "LeastBalanced")
-
-cor.test(stacked_wide_least$number_of_te, stacked_wide_least$concept_vocab, method = "pearson")
-
-## Anova to compare the unique dom vocab size and unique non-dom vocab size among the 3 balance groups in the stacked barchart
-stacked_wide %>% 
-  group_by(BalSubset) %>%
-  summarize(mean_DI = mean(Dom_unique_words, na.rm=T), 
-            sd_DI = sd(Dom_unique_words, na.rm=T),
-            min_DI = min(Dom_unique_words, na.rm=T),
-            max_DI = max(Dom_unique_words, na.rm=T))
-
-DI_byB_anova <- aov(Dom_unique_words ~ BalSubset, data = stacked_wide)
-summary(DI_byB_anova)
-
-pairwise.t.test(stacked_wide$Dom_unique_words, stacked_wide$BalSubset,
-                 p.adjust.method = "BH")
-
-stacked_wide %>% 
-  group_by(BalSubset) %>%
-  summarize(mean_NI = mean(NonDom_unique_words, na.rm=T), 
-            sd_NI = sd(NonDom_unique_words, na.rm=T),
-            min_NI = min(NonDom_unique_words, na.rm=T),
-            max_NI = max(NonDom_unique_words, na.rm=T))
-
-NI_byB_anova <- aov(NonDom_unique_words ~ BalSubset, data = stacked_wide)
-summary(NI_byB_anova)
-
-pairwise.t.test(stacked_wide$NonDom_unique_words, stacked_wide$BalSubset,
-                 p.adjust.method = "BH")
-```
-
-# plots
-## proportion of TEs
-```{r}
-keepers_ws_TE %>%
-  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
-  ggplot(aes(x = age.days/30, 
-             y = propTE)) +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 2))+
-  ylab("Percentage of Translation Equivalents") +
-  xlab("Age in Months")+
-  ggtitle("Percentage of Translation Equivalents by Age") +  
-  theme_classic(base_size = 22) +
-  scale_x_continuous(breaks= seq(18,32,by = 2))+
-  geom_point(size=3) + 
-  geom_smooth(method="lm" )
-
-ggsave(here("Output","PlotbyAge.png"), height=10,width=10)
-```
-
-## Bar graphs with individual scores (percentage TEs by dominance)
-```{r}
-keepers_ws_summarize = keepers_ws_TE %>% 
-  group_by(balance_group15) %>%
-  summarize(mean = mean(propTE, na.rm = T),
-         N = n(),
-         sd = sd(propTE, na.rm=T),
-         se = sd(propTE, na.rm = T)/ sqrt(N))
-
-ggplot(data=keepers_ws_summarize,
-       aes(x=balance_group15,
-           y=mean)) +
-    geom_bar(aes(fill = balance_group15),
-             show.legend =F,
-             colour="black",
-             stat="identity",
-             position="dodge", width=.7) +
-
-   geom_point(data = keepers_ws_TE,aes( y = propTE, fill = balance_group15), alpha = 0.9, show.legend = F, color = "black", pch = 21,size=1.75, 
-   stat = "identity",position=position_jitterdodge()) +
-  geom_errorbar(aes(ymin = mean - se, 
-                    ymax = mean + se), 
-                width=0.2,
-                position=position_dodge(.9)) +
-    ylab("Percentage of Translation Equivalents\n") +
-    xlab("Balance") +
-scale_x_discrete(breaks= c("80_20", "70_30", "60_40", "50_50"),
-                        labels =c("80:20","70:30","60:40","50:50"))+
-    ggtitle("Percentage of Translation Equivalents by Balance") +
-    theme_classic(base_size = 22) +
-  scale_y_continuous(labels = scales::percent_format(accuracy=2))+
-    theme(axis.text.x = element_text(angle = 45, hjust = 1), 
-          plot.title = element_text(hjust = .5))
-
-ggsave(here("output","PlotTE_byBalance.png"), width=9.5, height= 7)
-```
-## age by balancedness
-```{r}
-keepers_ws_TE %>%
-  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
-  group_by(balance_group15) %>%
-  summarize(median.age = median(age.days))
-
-keepers_ws_TE %>%
-  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
-  ggplot(aes(x=balance_group15, y=age.days/30, colour=balance_group15))+
-  geom_pirate(bars=F) +ylab("Age (Months)") + coord_flip()
-
-```
-
-## TE: age plot by balancedness
-```{r}
-keepers_ws_TE %>%
-  filter(vocab_type=="Totalvocabulary", bothCDI_filled=="Y") %>%
-  ggplot(aes(x = age.days/30, 
-             y = propTE)) +
-  scale_y_continuous(labels = scales::percent_format(accuracy = 2))+
-  ylab("Percentage of Translation Equivalents\n") +
-  xlab("Age in Months")+
-  ggtitle("Percentage of Translation Equivalents by Age and Balance") +  guides(colour = guide_legend(reverse=TRUE))+
-
-  theme_classic(base_size = 22) +
- # scale_x_continuous(breaks= seq(18,32,by = 2))+
-  geom_point(size=3, colour= "darkgrey") + 
-  geom_smooth(size =2, aes(colour=balance_group15), method="lm", se=F)+
-  scale_colour_discrete(name="Balance", 
-                        breaks= c("80_20", "70_30", "60_40", "50_50"),
-                        labels =c("least balanced (80:20)","less balanced (70:30)","more balanced (60:40)","most balanced (50:50)"))
-ggsave(here("Output","PlotTE_BalanceByAge.png"), width=13, height= 7)
-
-```
+```

### 2021-06-15 paper/paper_TE_bilingual_vocabulary_model_docx.Rmd
tried knitting to docx

--- 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
 
-floatsintext      : yes
-figurelist        : no
-tablelist         : no
+floatsintext      : no
+figurelist        : yes
+tablelist         : yes
 footnotelist      : no
@@
 class             : "man"
-output            : papaja::apa6_pdf # change to apa6_docx for word document
+output            : papaja::apa6_word # change to apa6_docx for word document
 citation_package  : biblatex

### 2021-06-16 paper/paper_TE_bilingual_vocabulary_model.Rmd
fixed typos/wordings

@@
   - name          : "Rachel Ka-Ying Tsui"
-    affiliation   : "1, 2"
+    affiliation   : "1"
     corresponding : yes    
     address       : "Department of Psychology, 7141 Sherbrooke St.
-West, Montreal, QC, Canada, H2T1V2, or Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
+West, Montreal, QC, Canada, H2T1V2.
+Current address: Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
     email         : "rachelkytsui@gmail.com"
@@
   - id            : "1"
-    institution   : "Department of Psychology, Concordia University"
-  - id            : "2"
-    institution   : "Laboratory for Language Development, RIKEN Center for Brain Science"
+    institution   : "Concordia University"
 
@@
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
-
-One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts.
-This is known as the Complementarity Principle [@Grosjean_2016].
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
-If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
-However, we argue that the Complementarity Principle is unlikely to strongly impact bilingual word acquisition in infancy.
-Most words in children’s early vocabularies could be considered “home words”, which include words for social and daily routines (“hello”, “more”, “diaper”), common nouns (“doggie”), and everyday verbs (“walk”).
-Such words are likely to be encountered across contexts where children spend the majority of their time, such as at home and at childcare.
-Thus, growing up in a bilingual context from birth, bilingual children presumably encounter “home words” in both languages.
-Accordingly, we assume that for the most part, bilingual children’s opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle.
-We further consider implications of this assumption in the discussion section.
-
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech [e.g., @Brent_Siskind_2001; @Swingley_Humphrey_2018], children’s efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), and family socioeconomic status (Fernald et al., 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).
 
@@
 * Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE). 
-* Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE).
+* Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE).
 * Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 2).
@@
     ## Define notions for significance levels; spacing is important.
-    mystars <- ifelse(p < .0001, "****", ifelse(p < .001, "*** ", ifelse(p < .01, "**  ", ifelse(p < .05, "*   ", "    "))))
+    mystars <- ifelse(p < .001, "***", ifelse(p < .01, "**", ifelse(p < .05, "*", " ")))
 
@@
     ## remove last column and return the correlation matrix
-    Rnew <- cbind(Rnew[1:length(Rnew)-1])
+    Rnew <- cbind(Rnew[2:length(Rnew)])
     return(Rnew)
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our model does not consider that vocabulary acquisition could also be predicted by other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), as well as other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and simple nouns (“ball”, “dog”) are among the first words children across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
+For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and common nouns (“ball”, “dog”) are among the first words children acquired across languages and cultures.
+It therefore seems reasonable to expect that bilingual children growing up in a bilingual household would be exposed to a similar set of referents and labels in each of their languages.
 Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.

### 2021-06-16 paper/paper_TE_bilingual_vocabulary_model.tex
fixed typos/wordings

@@
   pdftitle={Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants},
-  pdfauthor={Rachel Ka-Ying Tsui1, 2, Ana Maria Gonzalez-Barrero1, Esther Schott1, \& Krista Byers-Heinlein1},
+  pdfauthor={Rachel Ka-Ying Tsui1, Ana Maria Gonzalez-Barrero1, Esther Schott1, \& Krista Byers-Heinlein1},
   pdflang={en-EN},
@@
 \title{Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants}
-\author{Rachel Ka-Ying Tsui\textsuperscript{1, 2}, Ana Maria Gonzalez-Barrero\textsuperscript{1}, Esther Schott\textsuperscript{1}, \& Krista Byers-Heinlein\textsuperscript{1}}
+\author{Rachel Ka-Ying Tsui\textsuperscript{1}, Ana Maria Gonzalez-Barrero\textsuperscript{1}, Esther Schott\textsuperscript{1}, \& Krista Byers-Heinlein\textsuperscript{1}}
 \date{}
 
-\affiliation{\vspace{0.5cm}\textsuperscript{1} Department of Psychology, Concordia University\\\textsuperscript{2} Laboratory for Language Development, RIKEN Center for Brain Science}
+\affiliation{\vspace{0.5cm}\textsuperscript{1} Concordia University}
 
@@
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
-
-One important consideration in thinking about bilingual children's experience is whether they encounter and use their languages within the same or different contexts.
-This is known as the Complementarity Principle (Grosjean, 2016).
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
-If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
-However, we argue that the Complementarity Principle is unlikely to strongly impact bilingual word acquisition in infancy.
-Most words in children's early vocabularies could be considered ``home words,'' which include words for social and daily routines (``hello,'' ``more,'' ``diaper''), common nouns (``doggie''), and everyday verbs (``walk'').
-Such words are likely to be encountered across contexts where children spend the majority of their time, such as at home and at childcare.
-Thus, growing up in a bilingual context from birth, bilingual children presumably encounter ``home words'' in both languages.
-Accordingly, we assume that for the most part, bilingual children's opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle.
-We further consider implications of this assumption in the discussion section.
-
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder \& Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), and family socioeconomic status (Fernald et al., 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent \& Siskind, 2001; Goodman et al., 2008; Swingley \& Humphrey, 2018).
 
@@
 \item
-  Prediction 1b: Children who produce more dominant-language words (DOM) or more total words (WORD) will produce more translation equivalents (TE).
+  Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE).
 \item
@@
 \toprule
- & \multicolumn{1}{c}{Age (in month)} & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET}\\
+ & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET} & \multicolumn{1}{c}{CONCEPT}\\
 \midrule
@@
 \toprule
- & \multicolumn{1}{c}{Age (in month)} & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET}\\
+ & \multicolumn{1}{c}{LEARNABLE} & \multicolumn{1}{c}{BALANCE} & \multicolumn{1}{c}{WORD} & \multicolumn{1}{c}{DOM} & \multicolumn{1}{c}{NONDOM} & \multicolumn{1}{c}{TE} & \multicolumn{1}{c}{DOM-SINGLET} & \multicolumn{1}{c}{NONDOM-SINGLET} & \multicolumn{1}{c}{CONCEPT}\\
 \midrule
 \endhead
-Age (in month) &  & 0.95**** & -0.24*** & 0.65**** & 0.69**** & 0.45**** & 0.48**** & 0.65**** & 0.19**\ \\\
-LEARNABLE &  &  & -0.23*** & 0.62**** & 0.65**** & 0.43**** & 0.44**** & 0.62**** & 0.21**\ \\\
-BALANCE &  &  &  & -0.07\ \ \ \ & -0.29**** & 0.35**** & 0.25**** & -0.58**** & 0.63****\\
-WORD &  &  &  &  & 0.96**** & 0.87**** & 0.90**** & 0.74**** & 0.44****\\
-DOM &  &  &  &  &  & 0.70**** & 0.76**** & 0.89**** & 0.23***\\
-NONDOM &  &  &  &  &  &  & 0.99**** & 0.31**** & 0.72****\\
-TE &  &  &  &  &  &  &  & 0.38**** & 0.60****\\
-DOM-SINGLET &  &  &  &  &  &  &  &  & -0.09\ \ \ \\\
-NONDOM-SINGLET &  &  &  &  &  &  &  &  & \\
+Age (in month) & 0.95*** & -0.24*** & 0.65*** & 0.69*** & 0.45*** & 0.48*** & 0.65*** & 0.19** & 0.69***\\
+LEARNABLE &  & -0.23*** & 0.62*** & 0.65*** & 0.43*** & 0.44*** & 0.62*** & 0.21** & 0.66***\\
+BALANCE &  &  & -0.07 & -0.29*** & 0.35*** & 0.25*** & -0.58*** & 0.63*** & -0.21**\\
+WORD &  &  &  & 0.96*** & 0.87*** & 0.90*** & 0.74*** & 0.44*** & 0.98***\\
+DOM &  &  &  &  & 0.70*** & 0.76*** & 0.89*** & 0.23*** & 0.99***\\
+NONDOM &  &  &  &  &  & 0.99*** & 0.31*** & 0.72*** & 0.76***\\
+TE &  &  &  &  &  &  & 0.38*** & 0.60*** & 0.80***\\
+DOM-SINGLET &  &  &  &  &  &  &  & -0.09 & 0.86***\\
+NONDOM-SINGLET &  &  &  &  &  &  &  &  & 0.33***\\
 CONCEPT &  &  &  &  &  &  &  &  & \\
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not consider other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our model does not consider that vocabulary acquisition could also be predicted by other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), as well as other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and simple nouns (``ball,'' ``dog'') are among the first words children across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
+For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
+It therefore seems reasonable to expect that bilingual children growing up in a bilingual household would be exposed to a similar set of referents and labels in each of their languages.
 Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.

### 2021-06-16 paper/paper_TE_bilingual_vocabulary_model_docx.Rmd
fixed typos/wordings

--- 2021-06-15 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
     address       : "Department of Psychology, 7141 Sherbrooke St.
-West, Montreal, QC, Canada, H2T1V2, or Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
+West, Montreal, QC, Canada, H2T1V2.
+Current address: Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
     email         : "rachelkytsui@gmail.com"
@@
 class             : "man"
-output            : papaja::apa6_pdf # change to apa6_docx for word document
+output            : papaja::apa6_docx # change to apa6_docx for word document
 citation_package  : biblatex
@@
 library(kableExtra)
+library(flextable)
 library(english)
@@
 
-One important consideration in thinking about bilingual children’s experience is whether they encounter and use their languages within the same or different contexts.
-This is known as the Complementarity Principle [@Grosjean_2016].
-For example, for school-aged bilinguals, school-related words are more likely to be known in the language of schooling rather than in the home language (Bialystok et al., 2010).
-If certain words are encountered in particular contexts where only one language is used, bilinguals may have fewer opportunities to learn translation equivalents for these words.
-However, we argue that the Complementarity Principle is unlikely to strongly impact bilingual word acquisition in infancy.
-Most words in children’s early vocabularies could be considered “home words”, which include words for social and daily routines (“hello”, “more”, “diaper”), common nouns (“doggie”), and everyday verbs (“walk”).
-Such words are likely to be encountered across contexts where children spend the majority of their time, such as at home and at childcare.
-Thus, growing up in a bilingual context from birth, bilingual children presumably encounter “home words” in both languages.
-Accordingly, we assume that for the most part, bilingual children’s opportunities for learning words in each of their languages will be proportional to their overall exposure to the language, and largely not subject to the Complementarity Principle.
-We further consider implications of this assumption in the discussion section.
-
-Finally, we must note that vocabulary acquisition is not solely tied to quantity of input, but is also predicted by a host of other factors such as children’s ability to segment words from the continuous stream of speech [e.g., @Brent_Siskind_2001; @Swingley_Humphrey_2018], children’s efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder & Fernald, 2013), cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), and family socioeconomic status (Fernald et al., 2013).
-Nonetheless, all else being equal, words that are encountered more frequently are acquired sooner than those encountered less frequently (Brent & Siskind, 2001; Goodman et al., 2008; Swingley & Humphrey, 2018).
-
 ### Word learnability as a function of developmental level
@@
 
-kable(box1,
-      format="latex", booktabs=TRUE, longtable = TRUE) %>%
-  column_spec(1, width = "450px") %>%
-  kable_styling(font_size = 9)
+#kable(box1,
+#      format="latex", booktabs=TRUE, longtable = TRUE) %>%
+#  column_spec(1, width = "450px") %>%
+#  kable_styling(font_size = 9)
+
+flextable(box1) %>%
+  width(width = 6.5)
 
@@
 
-landscape(
-  kable(table1,
-        format = "latex", booktabs = TRUE,
-        caption = "Table 1. Summary of the parameters in the Bilingual Vocabulary Model.") %>%
-    kableExtra::kable_styling(font_size = 7) %>%
-    pack_rows("Main Parameters", 1, 4) %>%
-    pack_rows("Derived Parameters", 5, 10) %>%
-    column_spec(2:4, width = "175px") %>%
-    footnote(general = "All vocabulary measures are constrained to be integers.")
-)
+#landscape(
+#  kable(table1,
+#        format = "latex", booktabs = TRUE,
+#        caption = "Table 1. Summary of the parameters in the Bilingual Vocabulary Model.") %>%
+#    kableExtra::kable_styling(font_size = 7) %>%
+#    pack_rows("Main Parameters", 1, 4) %>%
+#    pack_rows("Derived Parameters", 5, 10) %>%
+#    column_spec(2:4, width = "175px") %>%
+#    footnote(general = "All vocabulary measures are constrained to be integers.")
+#)
+
+flextable(table1) %>%
+  fontsize(part = "all", size = 8) %>% 
+  set_header_labels(Relationship.to.other.parameters = "Relationship to other parameters") %>%
+  set_caption("Table 1. Summary of the parameters in the Bilingual Vocabulary Model.") %>%
+  width(width = 2)
+
 ```
@@
 # print table
-#papaja::apa_table(mean_TE_binBALANCE,
-#                  caption = "Average number of translation equivalents (TE) produced by each balance group.",
-#                  col.names =c("","N","Mean","$SD$","range"),
-#                  align="c")
-
-kable(mean_TE_binBALANCE,
-      format="latex", booktabs=TRUE, align = "c",
-      caption = "Table 6. Average number of translation equivalents (TE) produced by each balance group.") %>%
-  column_spec(1, width = "150px") %>%
-  column_spec(2:4, width = "50px") %>%
-  column_spec(5, width = "75px") %>%
-  kable_styling(position = "center")
+papaja::apa_table(mean_TE_binBALANCE,
+                  caption = "Average number of translation equivalents (TE) produced by each balance group.",
+                  col.names =c("","N","Mean","$SD$","range"),
+                  align="c")
+
+#kable(mean_TE_binBALANCE,
+#      format="latex", booktabs=TRUE, align = "c",
+#      caption = "Table 6. Average number of translation equivalents (TE) produced by each balance group.") %>%
+#  column_spec(1, width = "150px") %>%
+#  column_spec(2:4, width = "50px") %>%
+#  column_spec(5, width = "75px") %>%
+#  kable_styling(position = "center")
 ```

### 2021-06-18 paper/paper_TE_bilingual_vocabulary_model.Rmd
fixed wordings

@@
 West, Montreal, QC, Canada, H2T1V2.
-Current address: Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
+Current affilitation: Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
     email         : "rachelkytsui@gmail.com"
@@
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn than singlets (Neutral Account).
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn as singlets (Neutral Account).
 To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level.
@@
 Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word “cat”).
-However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning including vocabulary size in each language and the developmental level of a child — a point that we will return to later in the introduction.
+However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals’ vocabulary learning, including vocabulary size in each language and the developmental level of a child — a point that we will return to later in the introduction.
 
@@
 
-### Neutral Account: Translation equivalents are similar to learn than singlets
+### Neutral Account: Translation equivalents are similar to learn as singlets
 The previous accounts rely on the idea that bilingual vocabulary development unfolds differently than monolingual development, as monolinguals encounter only singlets but bilinguals encounter both singlets and translation equivalents.
@@
 The Neutral Account implies that bilingual children’s two languages develop relatively independently.
-Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
+Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, and weakly if at all correlated across languages.
 For example, 30-month-old bilingual children’s processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman et al., 2010).
@@
 ### Vocabulary size in bilinguals’ two languages
-Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know.
-For example, a child with a less balanced vocabulary across the two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language.
-Conversely, it seems reasonable to expect that if a child knows a similar number of words in each language and thus has a more balanced vocabulary across the two languages, there would be more potential for some of those words to be translation equivalents.
-
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
@@
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
+
+Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know.
+For example, a child with a less balanced vocabulary across the two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language.
+Conversely, it seems reasonable to expect that if a child knows a similar number of words in each language and thus has a more balanced vocabulary across the two languages, there would be more potential for some of those words to be translation equivalents.
 
@@
 If 100/1000 (or 1/10) students are left-handed, and 200/1000 (or 1/5) students are psychology majors, then if these variables are independent we expect a proportion of 1/10 × 1/5 = 1/50 students to be left-handed psychology majors.
-To determine the number we expect to observe in the college, we multiple 1/50 × 1000 = 20 students.
+To determine the number we expect to observe in the college, we multiply 1/50 × 1000 = 20 students.
 When we compare this expected number to the actual number of left-handed students, there are three possible outcomes.
 First, we may observe many more than 20 left-handed psychology students at the college (say 100 students), which suggests that being left-handed increases the probability of majoring in psychology.
-Or, second, we may observe many fewer than 20 left-handed psychology students at the college (say 5 students), this suggests that being left-handed decreases the probability of majoring in psychology (in this example by a factor of 1/4).
+Or, second, we may observe many fewer than 20 left-handed psychology students at the college (say 5 students), which suggests that being left-handed decreases the probability of majoring in psychology (in this example by a factor of 1/4).
 Finally, if left-handedness and majoring in psychology are independent, we can predict the number of left-handed psychology students by multiplying the observed number of left-handed students (100) by the observed number of psychology students (200), and dividing by the total population of the college (1000), so for example 100×200/1000 = 20.
@@
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-Very young children would not be expected to know many of the “harder” words on the CDI, such as “lawn mower”, “sidewalk”, or “vitamins'', due to children’s immature cognitive machinery and conceptual development.
+Very young children would not be expected to know many of the “harder” words on the CDI, such as “lawn mower”, “sidewalk”, or “vitamins", due to children’s immature cognitive machinery and conceptual development.
 A more reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
@@
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
-We also compared predicted learning outcomes for when translation equivalents are harder to learn, or easier to learn, or similar to learn than singlets.
+We also compared predicted learning outcomes for when translation equivalents are harder, easier, or similar to learn as singlets.
 
 In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English [@Fenson_etal_2007] and Québec French (Trudeau et al., 1999).
-Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
+Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) to learning singlets.
 
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equal to learn as singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equally learnable as singlets).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets).
@@
 
-* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn than singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
 
@@
 Caregivers were instructed to check off the words produced by their child using either a CDI paper questionnaire or the same questionnaire administered on a tablet.
-Data from paper based questionnaires was double entered and checked by trained research assistants.
+Data from paper based questionnaires were double entered and checked by trained research assistants.
 
@@
 Children produced an average of `r vocabulary_average$mean[2]` translation equivalents (TE; SD = `r vocabulary_average$sd[2]`, range = `r vocabulary_average$min[2]` - `r vocabulary_average$max[2]`).
-The remainder of words were singlets: Children produced many more singlets in their dominant language (DOM-SINGLET; M = `r vocabulary_average$mean[3]`, SD = `r vocabulary_average$sd[3]`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r vocabulary_average$mean[4]`, SD = `r vocabulary_average$sd[4]`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
+The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = `r vocabulary_average$mean[3]`, SD = `r vocabulary_average$sd[3]`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r vocabulary_average$mean[4]`, SD = `r vocabulary_average$sd[4]`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
 On average, children’s concept vocabulary size was `r vocabulary_average$mean[1]` (CONCEPT; SD = `r vocabulary_average$sd[1]`, range = `r vocabulary_average$min[1]` - `r vocabulary_average$max[1]`).
@@
 This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, ($F$(`r aov_age_CONCEPT$df[1]`, `r aov_age_CONCEPT$df[2]`) = `r round(aov_age_CONCEPT$statistic[1],2)`, $p `r papaja::printp(aov_age_CONCEPT$p.value[1])`$).
-Older children of 28–33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[3],2)` ($p$s < .001), those middle children of 23–27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[2],2)`, and those younger children of 18–22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[1],2)` ($p$s < .001).
+Older children of 28–33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[3],2)` ($p$s < .001), middle children of 23–27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[2],2)`, and younger children of 18–22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[1],2)` ($p$s < .001).
 
@@
 Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
-To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate BIAS parameter.
+To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate the BIAS parameter.
 
@@
 The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-As for the number of learnable vocabulary (LEARNABLE), this was determined by the averaging of English and French productive CDI vocabulary at the 90th percentile at different ages which was obtained from Wordbank (Frank et al., 2016), and Table 7 lists the denominator at different ages.
+The number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank; Frank et al., 2016).
+Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
 For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since the 90th percentile information was available only up to 30 months.
+For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
 
@@
 Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
-To reproduce the Bilingual Vocabulary Model’s formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included in the model (denoted in R by using a colon rather than an asterisk between the interacting predictors).
+To reproduce the Bilingual Vocabulary Model’s formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included (denoted in R by using a colon rather than an asterisk between the interacting predictors).
 Therefore, our final linear regression model equation was:
 
-_ObservedTE ~ 0 + Dominant vocabulary:Non-dominant vocabulary/90 percentile of CDI items._
+_ObservedTE ~ 0 + Dominant vocabulary:Non-dominant vocabulary/90th percentile of CDI items._
 
@@
 
-_ObservedTE * 90 percentile of CDI items ~ 0 + Dominant vocabulary:Non-dominant vocabulary_)
-
-\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was consistent with the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+_ObservedTE * 90th percentile of CDI items ~ 0 + Dominant vocabulary:Non-dominant vocabulary_)
+
+\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was related to the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
 If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
@@
 # Discussion
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn than singlet words (i.e., the first label for a particular referent).
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn as singlet words (i.e., the first label for a particular referent).
 To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were available to be learned given the child’s age.
-The relative difficulty of learning translation equivalents relative to singlets was modeled via the bias parameter (BIAS), which indicated whether translation equivalent learning is consistent with the Avoidance (BIAS < 1), Preference (BIAS > 1), or Neutral Account (BIAS = 1).
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn given the child’s age.
+The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
 
@@
 
-Prediction Set 1 pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
+A first set of predictions pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
 This pattern is consistent with reports from previous research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
-Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
+Moreover, both the simulated and observed data showed that the children who produced more total words also produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
 This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
@@
 
-Prediction Set 2 pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
 We operationalized developmental level in terms of children’s age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
 Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
-Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
+Likewise, we confirmed the model-generated prediction that older children would produce more translation equivalents than younger children.
 This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David & Wei, 2008; Legacy et al., 2016).
@@
 
-Having validated our overall approach, Prediction 3 motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents learning are similar to learn than singlets), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
+Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents are similar to learn as singlets), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
@@
 ## Developmental change in translation equivalent learning
-The developmental change of bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
+The developmental change we observed in bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
+As revealed by previous studies, younger children and children with smaller vocabulary sizes (and thus less vocabulary knowledge) seem to have only a weak bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
 In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
-At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
-Yet, this explanation would not predict that younger bilinguals’ data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance account, before perhaps returning to the Neutral account once children realize that each referent should have a label in each language.
+In contrast, children with larger vocabulary sizes appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
+At first blush, strengthening of one-to-one mapping biases with age could explain why younger children appear to learn relatively more translation equivalents than older children.
+Yet, this explanation would not predict that younger bilinguals’ data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance Account (i.e.
+from no bias to a bias against many-to-one mappings), before perhaps returning to the Neutral Account once children realize that each referent should have a label in each language.
 Thus, changes in one-to-one mapping biases do not provide a complete explanation for our results.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not consider that vocabulary acquisition could also be predicted by other factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), as well as other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
-It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
-Such a model may better characterize and predict bilingual vocabulary development as a function of experience.
-Moreover, it would be important to extend our Bilingual Vocabulary Model to longitudinal data or data of a different bilingual population to investigate if it is possible to replicate the qualitative shift where bilingual children’s ability to learn translation equivalents appears to change across development.
+However, our model does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+It would be interesting for future studies to take into consideration the such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+Such a model may better characterize and predict individual differences in bilingual vocabulary development.
+Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data and/or data of a different bilingual.
 
@@
 Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
-However, following the Complementarity Principle [@Grosjean_2016], bilinguals may have different experiences in each of their languages.
+However, according to the Complementarity Principle [@Grosjean_2016], bilinguals may have different experiences in each of their languages.
 For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., “soap”, “bath”, “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
@@
 For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and common nouns (“ball”, “dog”) are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children growing up in a bilingual household would be exposed to a similar set of referents and labels in each of their languages.
+It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
 Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
@@
 
-Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the bias parameter (BIAS) and the parameter that accounts for how many words are potentially learnable at a particular age (LEARNABLE).
-Under the Bilingual Vocabulary Model, the learnability parameter and the bias parameter jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter changes by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also change by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
-Our model estimated the number of learnable words to be the number that children at the 90th percentile at a particular age produce.
+Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
+Under the Bilingual Vocabulary Model, these two parameters jointly and inversely predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter increases by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
+Our model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce.
 Small changes to this approach (e.g., taking the number of words children at the 95th percentile produce) would likely not drastically alter our results, nor change the qualitative shift that we observed in our data.

### 2021-06-18 paper/paper_TE_bilingual_vocabulary_model.tex
fixed wordings

@@
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn than singlets (Neutral Account).
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn as singlets (Neutral Account).
 To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals' vocabulary in each language and their developmental level.
@@
 Note that, in their study, expected translation equivalent knowledge was simulated based on the number of lexical items that overlapped between two randomly chosen English monolinguals (e.g., whether both monolinguals knew the word ``cat'').
-However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning including vocabulary size in each language and the developmental level of a child --- a point that we will return to later in the introduction.
+However, it is unclear whether this is an appropriate point of comparison for bilingual children as this approach may overlook variables that impact bilinguals' vocabulary learning, including vocabulary size in each language and the developmental level of a child --- a point that we will return to later in the introduction.
 
@@
 
-\hypertarget{neutral-account-translation-equivalents-are-similar-to-learn-than-singlets}{%
-\subsubsection{Neutral Account: Translation equivalents are similar to learn than singlets}\label{neutral-account-translation-equivalents-are-similar-to-learn-than-singlets}}
+\hypertarget{neutral-account-translation-equivalents-are-similar-to-learn-as-singlets}{%
+\subsubsection{Neutral Account: Translation equivalents are similar to learn as singlets}\label{neutral-account-translation-equivalents-are-similar-to-learn-as-singlets}}
 
@@
 The Neutral Account implies that bilingual children's two languages develop relatively independently.
-Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, with weakly if at all correlated across languages.
+Indeed, language and processing measures for bilingual children tend to be tightly correlated within a particular language, and weakly if at all correlated across languages.
 For example, 30-month-old bilingual children's processing efficiency in a particular language closely correlated with vocabulary size in that language, but was unrelated to vocabulary size in their other language (Marchman et al., 2010).
@@
 
-Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know.
-For example, a child with a less balanced vocabulary across the two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language.
-Conversely, it seems reasonable to expect that if a child knows a similar number of words in each language and thus has a more balanced vocabulary across the two languages, there would be more potential for some of those words to be translation equivalents.
-
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
@@
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-Very young children would not be expected to know many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins'\,', due to children's immature cognitive machinery and conceptual development.
+Very young children would not be expected to know many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins,'' due to children's immature cognitive machinery and conceptual development.
 A more reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
@@
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
-We also compared predicted learning outcomes for when translation equivalents are harder to learn, or easier to learn, or similar to learn than singlets.
+We also compared predicted learning outcomes for when translation equivalents are harder, easier, or similar to learn as singlets.
 
 In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1999).
-Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), whether translation equivalent learning is easier (Preference Account), or similar to learn than singlets (Neutral Account).
+Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) to learning singlets.
 
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50\% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equal to learn as singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equally learnable as singlets).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50\% more likely to be learned than singlets).
@@
 \item
-  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn than singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
 \end{itemize}
@@
 Children produced an average of 67.70 translation equivalents (TE; SD = 85.10, range = 1 - 409).
-The remainder of words were singlets: Children produced many more singlets in their dominant language (DOM-SINGLET; M = 138.40, SD = 124.40, range = 2 - 523) than in their non-dominant language (NONDOM-SINGLET; M = 21.20, SD = 20.10, range = 0 - 94), \emph{t}(228) = 13.89, \(p < .001\), \(d = 0.92\).
+The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = 138.40, SD = 124.40, range = 2 - 523) than in their non-dominant language (NONDOM-SINGLET; M = 21.20, SD = 20.10, range = 0 - 94), \emph{t}(228) = 13.89, \(p < .001\), \(d = 0.92\).
 On average, children's concept vocabulary size was 227.30 (CONCEPT; SD = 181.30, range = 4 - 695).
@@
 This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, (\(F\)(2, 226) = 90.86, \(p < .001\)).
-Older children of 28--33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of 414.60 (\(p\)s \textless{} .001), those middle children of 23--27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of 252.10, and those younger children of 18--22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of 119.90 (\(p\)s \textless{} .001).
+Older children of 28--33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of 414.60 (\(p\)s \textless{} .001), middle children of 23--27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of 252.10, and younger children of 18--22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of 119.90 (\(p\)s \textless{} .001).
 
@@
 Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
-To reproduce the Bilingual Vocabulary Model's formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included in the model (denoted in R by using a colon rather than an asterisk between the interacting predictors).
+To reproduce the Bilingual Vocabulary Model's formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included (denoted in R by using a colon rather than an asterisk between the interacting predictors).
 Therefore, our final linear regression model equation was:
 
-\emph{ObservedTE \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary/90 percentile of CDI items.}
+\emph{ObservedTE \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary/90th percentile of CDI items.}
 
@@
 
-\emph{ObservedTE * 90 percentile of CDI items \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary})
-
-\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was consistent with the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+\emph{ObservedTE * 90th percentile of CDI items \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary})
+
+\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was related to the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
 If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
@@
 
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn than singlet words (i.e., the first label for a particular referent).
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn as singlet words (i.e., the first label for a particular referent).
 To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were available to be learned given the child's age.
-The relative difficulty of learning translation equivalents relative to singlets was modeled via the bias parameter (BIAS), which indicated whether translation equivalent learning is consistent with the Avoidance (BIAS \textless{} 1), Preference (BIAS \textgreater{} 1), or Neutral Account (BIAS = 1).
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn given the child's age.
+The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
 
@@
 
-Prediction Set 1 pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
+A first set of predictions pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
 This pattern is consistent with reports from previous research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
-Additionally, both our simulated and observed data showed that the more words children knew in their dominant language, the more translation equivalents they produced.
+Moreover, both the simulated and observed data showed that the children who produced more total words also produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
 This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
@@
 
-Prediction Set 2 pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
 We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
 Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
-Likewise, Prediction 2b was confirmed by the observed data as older children produced more translation equivalents than younger children.
+Likewise, we confirmed the model-generated prediction that older children would produce more translation equivalents than younger children.
 This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy et al., 2016).
@@
 
-Having validated our overall approach, Prediction 3 motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents learning are similar to learn than singlets), with this model explaining 96\% of variance in the data.
+Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents are similar to learn as singlets), with this model explaining 96\% of variance in the data.
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
@@
 
-The developmental change of bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
-As revealed by previous studies, younger children and children with smaller vocabulary size and thus less vocabulary knowledge seem to not have a strong bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
+The developmental change we observed in bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
+As revealed by previous studies, younger children and children with smaller vocabulary sizes (and thus less vocabulary knowledge) seem to have only a weak bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
 In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
-In contrast, children with larger vocabulary size appear to become more certain about the one-to-one mapping relationships between referents and words (Lewis et al., 2020), while at the same time they also take better advantage of their bilingual exposure to accept that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
-At first blush, strengthening of one-to-one mapping biases over age could explain why younger children appear to learn relatively more translation equivalents than older children.
-Yet, this explanation would not predict that younger bilinguals' data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance account, before perhaps returning to the Neutral account once children realize that each referent should have a label in each language.
+In contrast, children with larger vocabulary sizes appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
+At first blush, strengthening of one-to-one mapping biases with age could explain why younger children appear to learn relatively more translation equivalents than older children.
+Yet, this explanation would not predict that younger bilinguals' data would follow the Preference Account as we observed, and might instead predict development from the Neutral to the Avoidance Account (i.e.~from no bias to a bias against many-to-one mappings), before perhaps returning to the Neutral Account once children realize that each referent should have a label in each language.
 Thus, changes in one-to-one mapping biases do not provide a complete explanation for our results.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not consider that vocabulary acquisition could also be predicted by other factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), as well as other qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
-It would be interesting for future studies to take into consideration the qualitative factors in a bilingual word learning model, including different amounts of input and the quality of that input.
-Such a model may better characterize and predict bilingual vocabulary development as a function of experience.
-Moreover, it would be important to extend our Bilingual Vocabulary Model to longitudinal data or data of a different bilingual population to investigate if it is possible to replicate the qualitative shift where bilingual children's ability to learn translation equivalents appears to change across development.
+However, our model does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+It would be interesting for future studies to take into consideration the such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+Such a model may better characterize and predict individual differences in bilingual vocabulary development.
+Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data and/or data of a different bilingual.
 
@@
 Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
-However, following the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
 For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
@@
 For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children growing up in a bilingual household would be exposed to a similar set of referents and labels in each of their languages.
+It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
 Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.

### 2021-07-02 data_dictionary.Rmd
created data dictionary

---
title: "Data dictionary: Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants"
author: "Rachel K.Y. Tsui, Ana Maria Gonzalez-Barrero, Esther Schott, & Krista Byers-Heinlein"
date: '`r format(Sys.time(), "%a %b %d %X %Y")`'
output:
  html_document:
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: 'hide'
    self_contained: true
  pdf_document:
    toc: yes
    toc_depth: 4
    latex_engine: xelatex
---

```{r setup, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  warning = TRUE, # show warnings during codebook generation
  message = TRUE, # show messages during codebook generation
  error = TRUE, # do not interrupt codebook generation in case of errors,
                # usually better for debugging
  echo = TRUE  # show R code
)
ggplot2::theme_set(ggplot2::theme_bw())
pander::panderOptions("table.split.table", Inf)
```

```{r library, message=FALSE, warning=FALSE}
# load libraries
library(codebook)
library(here)
library(dplyr)
library(tidyverse)
library(future)
library(labelled)
```

This is a data dictionary for the data set used in the paper "Are translation equivalents special? Evidence from simulations and empirical data from bilingual infants".

```{r message=FALSE, warning=FALSE}
#load dataset
keepers_ws_TE <- rio::import(here::here("data_keepers/keepers_ws_TE_final.csv"))
```

# Variables {.tabset}

## Overview
```{r}
codebook_items(keepers_ws_TE)
```

## baby_id & ID_testdate

* __baby_id__: Unique participant ID
* __ID_testdate__: Unique participant ID + date of test. Some participants made multiple visits; therefore, certain unique participant IDs have multiple rows of data.

The dataset has N = `r length(unique(keepers_ws_TE$baby_id))` participants and `r length(keepers_ws_TE$ID_testdate)` datapoints.

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  summarise(N_baby_id = n_distinct(baby_id),
            N_ID_testdate = n_distinct(ID_testdate))
```

## visit_num
* __visit_num__: Data collected 1st, 2nd or 3rd visit to the lab. Most children come to lab only once, some come back at a later age
* __multiple_visits__: Who came once or came multiple times. Values: Single visit = FALSE, multiple visits = TRUE

While the majority of participants made only 1 visit, some participants made multiple visits.

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  group_by(multiple_visits) %>%
  distinct(baby_id, .keep.all=T) %>% 
  count()
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  group_by(visit_num) %>% 
  distinct(ID_testdate) %>%
  mutate(total = n()) %>% 
  count()
```

## gender
* __gender__: categorical with 2 levels: F & M

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  distinct(baby_id, .keep_all = TRUE) %>% 
  count(gender) %>%
  mutate(percentage = round(n/sum(n)*100, 2)) 
```

## age
* __age_days__: age in days
* __age_continuous__: age in months (with decimal)
* __age_months_binned__: age in months (binned; without decimal)

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  summarize(mean_age_continuous = mean(age_continuous, na.rm = T),
            sd_age_continuous = sd(age_continuous, na.rm = T),
            min_age_continuous = min(age_continuous, na.rm = T),
            max_age_continuous = max(age_continuous, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  ggplot(aes(age_continuous)) +
  geom_histogram()
```

## years_education
* __years_education__: Maternal education in years
```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  summarize(mean_years_education = mean(years_education, na.rm = T),
            sd_years_education = sd(years_education, na.rm = T),
            min_years_education = min(years_education, na.rm = T),
            max_years_education = max(years_education, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  ggplot(aes(years_education)) +
  geom_histogram()
```

## lang_group
All the data points included in the analysis are from bilinguals.
```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  mutate(lang_group = as.factor(lang_group)) %>%
  count(lang_group) %>%
  mutate(percentage = round(n/sum(n)*100, 2)) 
```

## lang_exp
* __lang_exp_eng__: % of the time when a participant is exposed to __English__ over the course of life globally
* __lang_exp_fre__: % of the time when a participant is exposed to __French__ over the course of life globally
* __lang_exp_other__: % of the time when a participant is exposed to __a third language__ over the course of life globally

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  pivot_longer(c(lang_exp_eng, lang_exp_fre, lang_exp_other), names_to = "language", values_to = "lang_exp") %>%
  group_by(language) %>%
  summarize(mean_lang_exp = mean(lang_exp, na.rm = T),
            sd_lang_exp = sd(lang_exp, na.rm = T),
            min_lang_exp = min(lang_exp, na.rm = T),
            max_lang_exp = max(lang_exp, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  pivot_longer(c(lang_exp_eng, lang_exp_fre, lang_exp_other), names_to = "language", values_to = "lang_exp") %>%
  group_by(language) %>%
  summarize(mean_lang_exp = mean(lang_exp, na.rm = T)) %>% 
  mutate(language = recode(language, 
                           lang_exp_eng = "English", 
                           lang_exp_fre = "French",
                           lang_exp_other = "Other language")) %>%
  ggplot(aes(x = language, y = mean_lang_exp)) +
  geom_bar(stat="identity") +
  labs(x = "Language", 
       y = "Mean language exposure (%)")
```

## lang_dom
* __lang_dom__: Language dominance based on % language exposure (i.e., the language with a greater exposure is the dominant language)
* __lang_dom_vocab__: Language dominance based on vocabulary size (i.e., the language with a greater vocabulary size is the dominant language)

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  pivot_longer(c(lang_dom, lang_dom_vocab), names_to = "language_dominance", values_to = "language") %>%
  group_by(language_dominance, language) %>%
  summarize(n = n())
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  pivot_longer(c(lang_dom, lang_dom_vocab), names_to = "language_dominance", values_to = "language") %>%
  group_by(language_dominance, language) %>%
  summarize(n = n()) %>%
  mutate(language_dominance = recode(language_dominance, 
                           lang_dom = "Language dominance \n based on % language exposure", 
                           lang_dom_vocab = "Language dominance \n based on vocabulary size")) %>%
  ggplot(aes(x = language_dominance, y = n, fill = language)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x = "Types of language dominance", 
       y = "Number of participants")
```

For most children, the language in which they produced the most words was also the language that they heard most often, although this was not the case for some children. 
```{r message=FALSE, warning=FALSE}
## check to see how many children have consistent/inconsistent dominant language between vocabulary-defined and input-defined dominance
keepers_ws_TE %>%
  mutate(consistent_lang_dom = if_else(lang_dom != lang_dom_vocab, 0, 1)) %>% # 1= consistent, 0 = inconsistent
  summarise(n = n(),
            n_consistent = sum(consistent_lang_dom),
            percentage_consistent = n_consistent/n*100,
            n_inconsistent = n - n_consistent,
            percentage_inconsistent = n_inconsistent/n*100)
```

Thus, these two constructs were related, although not identical.
```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  ggplot(aes(x = balance_vocab, y = lang_nondom_input)) +
  stat_smooth(method = lm, se = F, color = "black") +
  geom_point(shape = 1) + 
  theme_light() + 
  labs(x = "Balance based on vocabulary (BALANCE)", 
       y = "Balance based on exposure") 
```

## cdi_filled
* __cdi_filled_by__: Who filled out CDI? Due to limitations with the platform we used to enter CDI data, this information is provided for each visits (e.g, if mother filled English CDI and father filled French, the answer is Mother and Father, without specifying which CDI was filled by whom)">Who filled out CDI? Due to limitations with the platform we used to enter CDI data, this information is provided for each visits (e.g, if mother filled English CDI and father filled French, the answer is Mother and Father, without specifying which CDI was filled by whom)
* __both_cdi_filled__: Are both CDIs filled? (Y for yes; N for N)
* __required_cdi_filled__: Is the required CDI filled? Required for bilinguals: CDI in English AND French. (Y for yes; N for N)
* __eng_cdi_filled__: Is the English CDI filled?
* __fre_cdi_filled__: Is the French CDI filled?
* __dom_cdi_filled__: Is the CDI in the dominant language (defined by exposure: lang_dom) filled?
* __cdi_available__: Is the CDI available for both languages? (bothFilled = both English and French filled)

### Who filled out the CDIs?
```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  # combine grandmother & other family member
  mutate(cdi_filled_by = replace(cdi_filled_by, cdi_filled_by == "Grandmother", "Other family member")) %>% 
  # count number of respondents
  count(cdi_filled_by) %>%
  mutate(percentage = round(n/sum(n)*100, 2)) 
```

### Are both English and French CDIs filled?
```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  pivot_longer(c(both_cdi_filled, required_cdi_filled, eng_cdi_filled, fre_cdi_filled, dom_cdi_filled, cdi_available),
               names_to = "variable", values_to = "response") %>%
  group_by(variable, response) %>%
  summarize(n = n()) %>%
  mutate(percentage = round(n/sum(n)*100, 2)) 
```

## vocabulary measures

### Types of vocabulary
* __total_words_eng__: Total number of words in English
* __total_words_fre__: Total number of words in French
* __word_vocab__: Total word vocabulary (= total_words_eng + total_words_fre)
* __concept_vocab__: Total concept vocabulary
* __number_of_te__: Total number of translation equivalents (TEs), i.e., words produced by one child in both English and French)"
* __eng_unique_words__: Total number of words in English CDI minus number of TEs. Number of words produced only in English (not in French)
* __fre_unique_words__: Total number of words in French CDI minus number of TEs. Number of words produced only in French (not in English)

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  pivot_longer(c(total_words_eng, total_words_fre, word_vocab, concept_vocab, number_of_te, eng_unique_words, fre_unique_words),
        names_to = "vocab_type", values_to = "vocab_score") %>%
  group_by(vocab_type) %>%
  summarize(mean = mean(vocab_score, na.rm = T),
            sd = sd(vocab_score, na.rm = T),
            min = min(vocab_score, na.rm = T),
            max = max(vocab_score, na.rm = T))
```

### Derived variables
Instead of coding total number of words produced in English/French, codes for total number of words produced in dominant/non-dominant language defined by vocabulary size (i.e., lang_dom_vocab: the language with a greater vocabulary size is the dominant language)

* __total_words_dom__: Total number of words in the dominant language
* __total_words_nondom__: Total number of words in the non-dominant language
* __total_singlet_dom__: Total number of words in the dominant language minus number of TEs. Number of words produced only in the dominant language.
* __total_singlet_nondom__: Total number of words in the non-dominant language minus number of TEs. Number of words produced only in the non-dominant language
* __singlet_vocab__: Total singlet vocabulary (total_singlet_dom + total_singlet_nondom)

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  pivot_longer(c(total_words_dom, total_words_nondom, total_singlet_dom, total_singlet_nondom, singlet_vocab),
        names_to = "vocab_type", values_to = "vocab_score") %>%
  group_by(vocab_type) %>%
  summarize(mean = mean(vocab_score, na.rm = T),
            sd = sd(vocab_score, na.rm = T),
            min = min(vocab_score, na.rm = T),
            max = max(vocab_score, na.rm = T))
```

## balance measures

### Defined by vocabulary size
* __balance_vocab__: **The main balance measure** we used in the current study is a *Vocabulary Balance*, which is determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages. The formula for calculating the vocabulary balance score is total_words_dom/(total_words_dom+total_words_nondom).

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  summarize(mean_balance_vocab = mean(balance_vocab, na.rm = T),
            sd_balance_vocab = sd(balance_vocab, na.rm = T),
            min_balance_vocab = min(balance_vocab, na.rm = T),
            max_balance_vocab = max(balance_vocab, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  ggplot(aes(balance_vocab)) +
  geom_histogram()
```

### Defined by exposure
Balance can also be considered in terms of input in each language. To make balance_vocab and balance_input comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance (lang_dom_vocab), rather than the language that children heard most and least often.

* __lang_dom_input__: % of the time when a participant is exposed to __the dominant language__ over the course of life globally
* __lang_nondom_input__: % of the time when a participant is exposed to __the non-dominant language__ over the course of life globally
* __balance_input__: The formula for calculating the input balance score is also NONDOM/(DOM+NONDOM).

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  summarize(mean_balance_input = mean(balance_vocab, na.rm = T),
            sd_balance_input = sd(balance_vocab, na.rm = T),
            min_balance_input = min(balance_vocab, na.rm = T),
            max_balance_input = max(balance_vocab, na.rm = T))
```

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>% 
  ggplot(aes(balance_input)) +
  geom_histogram()
```

## Wordbank percentile measures
* __age_months_percentile__: Instead of their actual age in months, this variable adjusts for the age range available in the Wordbank CDI-WS data. As the upper age limit of the CDI-WS is 30m, ages between 30m to 33m are all changed to 30m.
* __EngWS_90percentile__: The number of English words children can produce at the 90th percentile across ages obtained from Wordbank.
* __FrWS_90percentile__: The number of French words children can produce at the 90th percentile across ages obtained from Wordbank.

```{r message=FALSE, warning=FALSE}
keepers_ws_TE %>%
  distinct(age_months_percentile, .keep_all = TRUE) %>%
  pivot_longer(c(EngWS_90percentile, FrWS_90percentile), names_to = "language", values_to = "words_90percentile") %>%
  mutate(language = recode(language, 
                           EngWS_90percentile = "English",
                           FrWS_90percentile = "French")) %>%
  ggplot(aes(x = age_months_percentile, y = words_90percentile, fill = language)) +
  geom_bar(stat="identity", position=position_dodge()) +
  labs(x = "Age in months adjusted for the CDI-WS age range (18-30m)", 
       y = "Number of words produced at the 90th percentile \n (obtained from Wordbank)")
```

### 2021-12-06 01_data_wordbank.Rmd
updated paper

@@
 
-```{r setup, include=FALSE}
+```{r setup, include=FALSE, cache = TRUE}
 knitr::opts_chunk$set(echo = TRUE)
@@
 So we need to retrieve the percentile information from Wordbank (Frank et al., 2016).
-```{r load_wordbankr}
+```{r load_wordbankr, cache = TRUE}
 library(wordbankr)
@@
 ## English CDI-Words and sentences: Vocabulary size at 90th percentile
-```{r eng_percentile_info}
+```{r eng_percentile_info, cache = TRUE}
 english_ws_percentile <-  
@@
                       quantile = "standard") %>%
-  # keep only the 90% percentile
-  filter(quantile == 0.90) %>%
-  rename(EngWS_90percentile = production)
+  pivot_wider(names_from = quantile, values_from = production) %>%
+  rename(EngWS_90percentile = "0.9",
+         EngWS_75percentile = "0.75",
+         EngWS_50percentile = "0.5",
+         EngWS_25percentile = "0.25",
+         EngWS_10percentile = "0.1")
+
 ```
@@
 ## French CDI-Mots et énoncés: Vocabulary size at 90th percentile
-```{r fre_percentile_info}
+```{r fre_percentile_info, cache = TRUE}
 french_ws_percentile <-  
@@
                       quantile = "standard") %>%
-  # keep only the 90% percentile
-  filter(quantile == 0.90) %>%
-  rename(FrWS_90percentile = production)
+  pivot_wider(names_from = quantile, values_from = production) %>%
+  rename(FrWS_90percentile = "0.9",
+         FrWS_75percentile = "0.75",
+         FrWS_50percentile = "0.5",
+         FrWS_25percentile = "0.25",
+         FrWS_10percentile = "0.1")
+
 ```
@@
 Then we merge the two percentile dataframe by age.
-```{r merge_percentile}
+```{r merge_percentile, cache = TRUE}
 ws_percentile <- merge(english_ws_percentile, french_ws_percentile, 
@@
   # compute estimate total vocab by adding the two languages
-  mutate(Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
+  mutate(Average_10percentile = (EngWS_10percentile + FrWS_10percentile)/2,
+         Average_25percentile = (EngWS_25percentile + FrWS_25percentile)/2,
+         Average_50percentile = (EngWS_50percentile + FrWS_50percentile)/2,
+         Average_75percentile = (EngWS_75percentile + FrWS_75percentile)/2,
+         Average_90percentile = (EngWS_90percentile + FrWS_90percentile)/2) %>%
   # keep only necessary columns
-  select(age, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
+  select(-c(language.x, form.x, language.y, form.y)) %>%
   # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
   rename(age_months_percentile = age)
+
 ```

### 2021-12-06 02_clean_data.Rmd
updated paper

@@
 
-```{r load_data_raw}
+```{r load_data_raw, cache = TRUE}
 # raw archival data from Gonzalez-Barrero, Schott, & Byers-Heinlein’s (2020)
@@
 
-```{r}
+```{r cache = TRUE}
 # change which infants changed lang_group & create a "exclude.language_group_change" criterion for subsequent step
@@
 ## How many distinct infants and total number of data points are there?
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE %>% summarize(N_visits = n_distinct(ID_testdate),
@@
 ### Vocabulary score in total vocabulary in English/French, total word vocabulary, and total concept vocabulary
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE %>% filter(is.na(vocab_score)) %>% 
@@
 ### Unique words in English
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE %>% filter(is.na(eng_unique_words)) %>% 
@@
 ### Unique words in French
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE %>% filter(is.na(fre_unique_words)) %>% 
@@
 ### Number of translation equivalents
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE %>% filter(is.na(number_of_te)) %>% 
@@
 The keepers_data is saved in the folder "data_keepers"
-```{r}
+```{r cache = TRUE}
 write.csv(keepers_ws_TE, "data_keepers/keepers_ws_TE.csv", row.names = F)

### 2021-12-06 03_prepare_data.Rmd
updated paper

@@
 
-```{r load_data_raw}
+```{r load_data_raw, cache = TRUE}
 keepers_ws_TE <- read.csv(here::here("data_keepers/keepers_ws_TE.csv"))
@@
 # Pivot data from long to wide
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE_final <- keepers_ws_TE %>%
@@
 
-```{r}
+```{r cache = TRUE}
 keepers_ws_TE_final <- keepers_ws_TE_final %>%
@@
 
-```{r}
+```{r cache = TRUE}
 # read in the percentile data
@@
 The final keepers_data is saved in the folder "data_keepers"
-```{r}
+```{r cache = TRUE}
 write.csv(keepers_ws_TE_final, "data_keepers/keepers_ws_TE_final.csv", row.names = F)

### 2021-12-06 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated paper

@@
 This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
-To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
+To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language, and includes a bias parameter that varies the difficulty of learning translation equivalents according to each account.
 Study 2 tested model-derived predictions against vocabulary data from 200 French–English bilingual children aged 18–33 months.
@@
 
-```{r load_packages_settings, include = FALSE}
+```{r load_packages_settings, include = FALSE, cache = TRUE}
 library(rmarkdown)
@@
 
-```{r load_data}
+```{r load_data, cache = TRUE}
 # read in the final cleaned keepers_data
@@
 For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents [^1] (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics.
 For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak & Byers-Heinlein, 2019).
-That is, upon hearing the English word “apple”, the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent.
+That is, upon hearing the English word “apple,” the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
+In vocabulary acquisition, bilingual children must learn a first label for a referent  (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet).
 Is translation equivalent learning different from singlet learning?
@@
 
+[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word “hygge” refers to a type of cozy contentment and enjoyment of simplicity.
+Moreover, even when a word that could be considered a translation equivalent exists, adults’ concepts may not be congruent across languages (e.g., Malt & Majid, 2013).
+However, in early development, most vocabulary words are reasonably concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.
+
 ## Accounts of translation equivalent learning
@@
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level [@Markman_Wachtel_1988; @Markman_1992; @Markman_1994].
-For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug”, they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.
+This avoidance of translation equivalents — particularly for nouns — was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level [@Markman_Wachtel_1988; @Markman_1992; @Markman_1994].
+For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like “wug,” they assume that “wug” refers to the garlic press — the object unknown to them — rather than to the cup, the object for which they already know the word.
 
@@
 Indeed, experimental work has suggested bilingual experience in infancy might not support the development of one-to-one mapping biases such as mutual exclusivity, at least in early infancy.
-For example, when hearing a novel word like “nil”, monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010).
+For example, when hearing a novel word like “nil,” monolingual children aged 17–22 months looked towards a novel object rather than a familiar object, but bi- and multilingual children looked similarly to both objects (Byers-Heinlein & Werker, 2009; 2013; Houston-Price et al., 2010).
 A recent meta-analysis also indicated that bilingual children show mutual exclusivity to a weaker degree than monolinguals (Lewis et al., 2020).
@@
 At a minimum, word learning requires encoding and representing the relevant sounds of a word, creating a mental representation of its referent, and linking the two.
-When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple”, one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
+When a French–English bilingual child encounters the word _“pomme”_ after having learned “apple,” one part of that process has already occurred in that the referent is already represented; because part of the word learning task is already accomplished, translation equivalents might therefore be easier to learn than singlets (e.g., Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
 Moreover, research suggests that bilingual lexicons are not tightly encapsulated by language, but instead include cross-language mental links between words that are semantically related [e.g., @Floccia_etal_2020; @Jardak_Byers-Heinlein_2019; @Singh_2014].
@@
 The researchers found that the percentage of translation equivalents observed in English–Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence implied that singlets and translation equivalents are equally learnable.
+This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (_n_ = 27) in the context of a wide age range (8 months to 2.5 years).
 In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
@@
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
-That is, potential singlets might be “too hard” to be learned at a particular age.
+That is, many potential singlets could be “too hard” to be learned at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
+For example, an 14-month-old who has learned the French words _“ballon”_ and _“chien”_ would be capable of learning their translation equivalents “ball” and “dog”, as easier potential singlets like “cup”, but not later-acquired potential singlets like “tomorrow”, “who”, or “not”.
 Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
@@
 This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
-The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B | A) where P(B | A) = P(B) if A and B are independent (see Box 1 for a detailed example).
-The full model is shown in Figure 1.
+The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent[^2].
 In the next paragraphs, we define each of the model parameters in detail, and these are also summarized in Table 1.
 
-\newpage
-```{r box1, results="asis", tab.cap = NULL}
-box1 <- data.frame(title = "To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
-To determine independence, we must know the probability of being a psychology major, the probability of being left handed, and the probability of being both a psychology major and left-handed.
-The numerator of these probabilities will be the number of students who are left-handed, psychology majors, and both (respectively) and the denominator will be the total population of students we are observing.
+[^2]: To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
 Imagine a college of 1000 students.
@@
 Finally, if left-handedness and majoring in psychology are independent, we can predict the number of left-handed psychology students by multiplying the observed number of left-handed students (100) by the observed number of psychology students (200), and dividing by the total population of the college (1000), so for example 100×200/1000 = 20.
-Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.") %>%
-  rename("Box 1. Example of test for independence." = title)
-
-kable(box1,
-      format="latex", booktabs=TRUE, longtable = TRUE) %>%
-  column_spec(1, width = "450px") %>%
-  kable_styling(font_size = 9)
-
-```
-
-```{r fig1, fig.cap="Illustration of the Bilingual Vocabulary Model.", fig.align='center', out.width='90%'}
-
-knitr::include_graphics("./Figure1_illustration.jpg")
-```
+Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.
 
@@
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-Very young children would not be expected to know many of the “harder” words on the CDI, such as “lawn mower”, “sidewalk”, or “vitamins", due to children’s immature cognitive machinery and conceptual development.
-A more reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
+However, not all children will be capable of learning all of the CDI words.
+Typically, words that are more concrete, occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019), and very young children would not be expected to produce many of the “harder” words on the CDI, such as “lawn mower,” “sidewalk,” or “vitamins”.
+Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
+Since there are no CDI norms for French-English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
-Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
-Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable — if imperfect — estimate of the number of learnable words (LEARNABLE) that a child of Jamie’s age could potentially acquire in each language.
-Thus, we might expect that Jamie could potentially have learned up to 245 English words and 245 French words, although he has thus far only learned 50 in English and 20 in French.
+Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 240 words (retrieved from the Wordbank database version 0.3.1; Frank et al., 2016).
+Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that averaging between the English and French CDIs at the 90th percentile provides a reasonable — if imperfect — estimate of the number of learnable words (LEARNABLE) that a bilingual child of Jamie’s age could potentially acquire in each language.
+Thus, we might expect that Jamie could potentially have learned up to 240 words in English and 240 words in French, although he has thus far only learned 50 in English and 20 in French.
+Note that we model learnability in terms of the number of words expected to be learnable at a particular age, rather than modeling the learnability of individual words at a particular age — a simplifying choice that we will return to in the discussion section.
 
@@
 Adding this parameter, translation equivalents can be derived from TE = BIAS × (DOM × NONDOM)/LEARNABLE.
-For the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets; for the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets; for the the Neutral Account, BIAS is exactly 1 (i.e., the model is unbiased with respect to whether translation equivalents are more difficult or easier to acquire than singlets).
+For the Avoidance Account, BIAS will be less than 1, meaning that translation equivalentsare less easily learned than singlets; for the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets; for the the Neutral Account, BIAS is exactly 1 (i.e., the model is unbiased with respect to whether translation equivalents are more difficult or easier to acquire than singlets).
 Going back to the example of 18-month-old Jamie, we would set the denominator at 245 which is the number of potentially learnable words at 18 months.
@@
 
-Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.
-
-Balance of vocabulary (BALANCE) is the proportion of total words that children produce in each language.
-For convenience, balance is defined in reference to the non-dominant language with the formula NONDOM/(DOM+NONDOM), such that scores can range from 0.0 (completely unbalanced) to 0.5 (completely balanced).
-For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29.
-Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many of the words he produces are translation equivalents).
-
-Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
-Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child — that is, the total number of concepts that are lexicalized in either language.
-This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD).
-Finally, we can also calculate singlets that are produced in each language, that is words for which the child does not yet produce a translation equivalent.
-Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM).
-It is also possible to decompose children’s word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
-
-```{r table1, results = "asis"}
+```{r table1, results = "asis", cache = TRUE}
 
@@
 
-## Current research
-The current research aimed to better understand the nature of translation equivalent learning in bilingual children.
+Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.
+
+Balance of vocabulary (BALANCE) is the proportion of total words that children produce in each language.
+For convenience, balance is defined in reference to the non-dominant language with the formula NONDOM/(DOM+NONDOM), such that scores can range from 0.0 (completely unbalanced) to 0.5 (completely balanced).
+For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29.
+Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many of the words he produces are translation equivalents).
+
+Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
+Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child — that is, the total number of concepts that are lexicalized in either language.
+This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD).
+Finally, we can also calculate singlets that are produced in each language, that is words for which the child does not yet produce a translation equivalent.
+Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM).
+It is also possible to decompose children’s word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
+
+Finally, we note two simplifying assumptions made by our model.
+First, we assume that children have the same communicative need in each language: our model does not account for cases where bilinguals might systematically refer to a concept in only one language and not the other (e.g., Grosjean, 2016).
+Second, we assume that translation equivalents refer to the same underlying concept.
+Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt & Majid, 2013; Storm et al., 2015; White et al., 2020).
+However, we believe that these assumptions present a reasonable first approximation, especially in the case of young children, and return to these assumptions in the discussion section.
+
+## Current studies
+The current studies aimed to better understand the nature of translation equivalent learning in bilingual children.
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
@@
 # Study 1: Simulations
-Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction (see also Figure 1), which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction, which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
 Note that usually only three values are necessary to calculate all the other variables (see Table 1).
@@
 
-Three simulations were generated to explore expected patterns of translation equivalent learning under the Bilingual Vocabulary Model.
+We present three simulations generated by creating data points for hypothetical children by varying different parameters of the Bilingual Vocabulary Model, in order to explore expected patterns of translation equivalent learning.
+The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except for those that violated constraints in the model (e.g., NON-DOM could not be larger than DOM).
 In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NON-DOM), and word vocabulary (WORD).
@@
 
-```{r table2, results = "asis"}
+```{r table2, results = "asis", cache = TRUE}
 
@@
   Simulation = c("1", "2", "3"),
-  LEARNABLE = c("Constant at 600", "Varied at 300, 450, and 600", "Varied at 150, 300, 450, and 600"),
+  LEARNABLE = c("Constant at 600", "Varied at 100, 200, 300, 400, 500, 600", "Varied at 150, 300, 450, and 600"),
   DOM = c("Varied, ranging from 100 to LEARNABLE at an interval of 100", 
@@
   NONDOM = c("Varied, ranging from 0 to DOM at an interval of 10",
-             "Varied, ranging from 0 to DOM at an interval of 25",
+             "Varied, ranging from 0 to DOM at an interval of 20",
              "Varied, ranging from 0 to DOM at an interval of 25"),
@@
            "Varied at 0.5 (Avoidance Account), 1 (Neutral Account), and 1.5 (Preference Account)"),
-  n_data = c(216, 161, 166)
+  n_data = c(216, 301, 166)
 )
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-learning 26 month-old (i.e., the most verbal 26-month-old English-learner at the 90th percentile of vocabulary produces around 600 words as retrieved from the Wordbank database; Frank et al., 2016).
-We set BIAS to 1, meaning that in these examples translation equivalents are similarly easy to learn as singlets.
-
-We first illustrate with three hypothetical children.
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
+We set BIAS to 1, meaning that in these examples translation equivalents are equally easy to learn as singlets.
+
+We first illustrate with three simulated children.
 Infant Annie (small vocabulary, unbalanced exposure) produces 270 words in the dominant language and 30 words in her non-dominant language.
@@
 
-```{r table3, results="asis"}
+```{r table3, results="asis", cache = TRUE}
 
@@
 Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 2, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
 Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
@@
 * Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE). 
-* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 2).
-
-```{r simulation1, echo = FALSE}
+* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 1).
+
+```{r simulation1, echo = FALSE, cache = TRUE}
 # Simulation 1: simulating infants of the same developmental level with different word vocabularies and balances of exposure (LEARNABLE = 600)
@@
 
-```{r simulation1-plus_observed_data, echo = FALSE}
+```{r simulation1-plus_observed_data, echo = FALSE, cache = TRUE}
 Observed_plot_inputs_TE <- keepers_ws_TE %>% 
@@
 
-```{r fig2, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
+```{r fig1, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
 Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
-Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%", fig.pos = "H"}
+Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%", fig.pos = "H", cache = TRUE}
 
@@
 
-Translation equivalent knowledge was simulated across children at three developmental levels (the number of LEARNABLE words = 300, 450, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
-In total, data from 161 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
+Translation equivalent knowledge was simulated across children at four developmental levels (the number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
+In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
 Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
-We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children’s concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
-Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), to medium balanced (BALANCE = .20 - .35), to least balanced (BALANCE = .00 - .02).
-
-```{r simulation2, echo = FALSE}
-# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (LEARNABLE = 300, 450, or 600)
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
+Figure 2 plots the number of TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
+Across the three Panels (1A, 1B, and 1C),  simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+In Panel 1A, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
+Panels 1B and 1C show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
+Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
+Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
+
+* Prediction 2a: Have larger concept vocabularies (CONCEPT). 
+* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET.
+* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+
+```{r simulation2, echo = FALSE, cache = TRUE}
+# Simulation 2: simulating patterns of translation equivalent learning as a function of different developmental timepoints (LEARNABLE = 100, 200, 300, 400, 500, 600)
 
 ## generate the simulation data
-LEARNABLE <- seq(300, 600, by = 150) # set LEARNABLE to 300, 450, and 600
+LEARNABLE <- seq(100, 600, by = 100) # set LEARNABLE to 100, 200, 300, 400, 500, and 600
 
@@
   DOM_seq <- seq(100, LEARNABLE, by = 100) # sequence of D from 100 to V at an interval of 100
-  NONDOM_seq <- seq(0, LEARNABLE, by = 25) # sequence of N from 0 to V at an interval of 25
+  NONDOM_seq <- seq(0, LEARNABLE, by = 20) # sequence of N from 0 to V at an interval of 20
   DOM <- rep(DOM_seq, length(NONDOM_seq))
@@
          DOM_SINGLET = DOM-TE, # Unique words in dominant language
-         NONDOM_SINGLET = NONDOM-TE) %>% # Unique words in non-dominant language
+         NONDOM_SINGLET = NONDOM-TE, # Unique words in non-dominant language
+         CONCEPT = WORD - TE) %>% # concept vocabulary
   # Round to keep whole numbers only
@@
   # bin B into 3 groups
-  mutate(bin_BALANCE = case_when(BALANCE >= 0.36 ~ "Most Balanced",
-                                 BALANCE <= 0.20 ~ "Least Balanced",
-                                 TRUE ~ "Medium Balanced")) %>%
-  mutate(bin_BALANCE  = fct_rev(bin_BALANCE)) %>%
+  mutate(bin_BALANCE = cut(BALANCE, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
   arrange(DOM, NONDOM, bin_BALANCE)
@@
 simulation2 <- data_simulation2 %>%
-  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
+  pivot_longer(-c(LEARNABLE, WORD, CONCEPT, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
   filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("NONDOM_SINGLET", "DOM_SINGLET", "TE"))) %>%
-  mutate(LEARNABLE_label = case_when(LEARNABLE == 600 ~ "LEARNABLE = 600",
-                                     LEARNABLE == 450 ~ "LEARNABLE = 450",
-                                     LEARNABLE == 300 ~ "LEARNABLE = 300")) %>%
-  group_by(LEARNABLE, LEARNABLE_label, bin_BALANCE, vocab_type) %>%
-  summarise(n_words = mean(number)) %>%
-  ggplot(aes(x = bin_BALANCE, y = n_words, fill = vocab_type)) + 
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ LEARNABLE_label) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
-  scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
-  ylim(0, 600) +
+  mutate(vocab_type = factor(vocab_type, levels=c("TE", "DOM_SINGLET", "NONDOM_SINGLET")),
+         vocab_type = recode(vocab_type, 
+                             TE = "Panel 1A: \n Translation equivalents \n (TE)",
+                             DOM_SINGLET = "Panel 1B: \n Singlets in dominant language \n (DOM_SINGLET)",
+                             NONDOM_SINGLET = "Panel 1C: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+  ggplot(aes(x = LEARNABLE, y = number, color = bin_BALANCE, linetype = bin_BALANCE)) +
+  facet_grid(. ~ vocab_type) +
+  stat_smooth(method = lm, se = F) + 
+  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
   theme_minimal() +
@@
         plot.title.position = "plot",
-        text = element_text(size=10),
-        panel.spacing.x = unit(2, "lines")) + 
-  labs(x = "\nVocabulary balance (BALANCE)", 
+        text = element_text(size=12),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) + 
+  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
+  labs(x = "\nNumber of learnable words (LEARNABLE)", 
        y = "Number of concepts (CONCEPT)", 
-       title = "Panel A: Simulated data")
-```
-
-```{r simulation2-plus_observed_data, echo = FALSE}
-plot_stacked_byLEARNABLE <- keepers_ws_TE %>%
-  select(baby_id, age_months_percentile, balance_vocab, word_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
-  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months \n (LEARNABLE = 244.9 - 451.9)",
-                                      age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months \n (LEARNABLE = 491.8 - 604.1)",
-                                      TRUE ~ "28-33 months \n (LEARNABLE = 620.4 - 638.9)")) %>%
-  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
-                               balance_vocab <= 0.20 ~ "Least Balanced",
-                               TRUE ~ "Medium Balanced")) %>%
-  mutate(BalSubset = fct_rev(BalSubset)) %>%
-  arrange(desc(number_of_te)) %>%
+       color = "Vocabulary balance (BALANCE)",
+       linetype = "Vocabulary balance (BALANCE)",
+       title = "1) Simulated data") +
+  guides(color = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1)) 
+```
+
+```{r simulation2-plus_observed_data, echo = FALSE, cache = TRUE}
+Observed_plot_LEARNABLE <- keepers_ws_TE %>%
+  select(baby_id, age_months_percentile, Average_90percentile, balance_vocab, word_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
   mutate(TE = number_of_te) %>%
-  pivot_longer(-c(baby_id, age_months_percentile, balance_vocab, LEARNABLE_subset, BalSubset, word_vocab, TE), 
-               names_to = "Type", values_to = "NumberOfWords") %>%
-  mutate(Type = factor(Type, levels=c("total_singlet_nondom", "total_singlet_dom", "number_of_te"))) %>%
-  mutate(Type = recode(Type, 
-                       total_singlet_nondom = "Singlet in non-dominant language (NONDOM-SINGLET)", 
-                       total_singlet_dom = "Singlet in dominant language (DOM-SINGLET)", 
-                       number_of_te = "Translation equivalent (TE)")) %>%
-  group_by(LEARNABLE_subset, BalSubset, Type) %>%
-  summarise(n_words = mean(NumberOfWords)) %>%
-  ggplot(aes(x = BalSubset, y = n_words, fill = Type)) +
-  geom_bar(position="stack", stat="identity") +
-  facet_grid(. ~ LEARNABLE_subset) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
-  scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
-  ylim(0, 600) +
+  pivot_longer(c(total_singlet_nondom, total_singlet_dom, number_of_te), names_to = "vocab_type", values_to = "number_of_words") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("number_of_te", "total_singlet_dom", "total_singlet_nondom")),
+         vocab_type = recode(vocab_type, 
+                             number_of_te = "Panel 2A: \n Translation equivalents \n (TE)",
+                             total_singlet_dom = "Panel 2B: \n Singlets in dominant language \n (DOM_SINGLET)",
+                             total_singlet_nondom = "Panel 2C: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+  ggplot(aes(x = age_months_percentile, y = number_of_words, color = bin_BALANCE, linetype = bin_BALANCE)) +
+  facet_grid(. ~ vocab_type) +
+  stat_smooth(method = lm, se = F) + 
+  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
   theme_minimal() +
-  theme(axis.text.x = element_text(angle = 0) , 
-        legend.position = "bottom",
+  theme(legend.position = "bottom",
         plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
         plot.title.position = "plot",
-        text = element_text(size=10),
-        panel.spacing.x = unit(2, "lines")) + 
+        text = element_text(size=12),
+        panel.spacing.x = unit(2, "lines"),
+        panel.background = element_rect(fill = NA, color = "black")) + 
   guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
-  labs(x = "\nVocabulary balance (BALANCE)", 
+  ylim(0, 450) +
+  #xlim(200, 700) +
+  scale_x_continuous(breaks = seq(18, 30, by = 2)) +
+  labs(x = "\nAge (in months)", 
        y = "Number of concepts (CONCEPT)", 
-       title = "Panel B: Observed data")
-
-# combine Simulation2 visualization + observed data visualization
-plot_Simulation2plusData <- ggarrange(simulation2 + rremove("legend"), 
-                                      plot_stacked_byLEARNABLE,
+       color = "Vocabulary balance (BALANCE)",
+       linetype = "Vocabulary balance (BALANCE)",
+       title = "2) Observed data") +
+  guides(color = guide_legend(reverse = T, nrow = 1),
+         linetype = guide_legend(reverse = T, nrow = 1))
+
+plot_Simulation2plusData <- ggarrange(simulation2, 
+                                      Observed_plot_LEARNABLE,
                                       nrow = 2,
-                                      heights=c(0.8,1))
-```
-
-```{r fig3, fig.cap="Number of translation equivalents (TE) and singlets in dominant (DOM-SINGLET) and non-dominant language (NONDOM-SINGLET) across different developmental levels/ages, which sets the number of LEARNABLE words.
-Panel A represents the model simulation in Study 1, where developmental levels of simulated children are set at three values: LEARNABLE = 300, 450, and 600.
-Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right).
-Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.width="80%", fig.pos = "H"}
+                                      common.legend = TRUE, legend = "bottom")
+```
+
+```{r fig2, fig.cap="Number of translation equivalents (TE; Panel A), singlets in dominant (DOM-SINGLET; Panel B) and singlets in the non-dominant language (NONDOM-SINGLET; Panel C) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+Row 1 represents the simulated data in Study 1.
+Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.width="80%", fig.pos = "H", cache = TRUE}
 
@@
 ```
-
-In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
-Moreover, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
-Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Overall, we generated 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
-Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
-
-* Prediction 2a: Have larger concept vocabularies (CONCEPT). 
-* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
-* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET.
-* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
 
@@
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets).
-In Figure 4, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
+In Figure 3, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
 Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
@@
 
-```{r simulation3, echo = FALSE}
+```{r simulation3, echo = FALSE, cache = TRUE}
 # Simulation 3: Non-independence of dominant-language (DOM) and non-dominant language words (NONDOM)
@@
 
-```{r fig4, fig.cap="Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%", fig.pos = "H"}
+```{r fig3, fig.cap="Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%", fig.pos = "H", cache = TRUE}
 
@@
 ```
+
+## 1.4 Summary of model predictions
+Study 1 presented predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
+Prediction Set 1 illustrated that the more words children produced in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they would produce.
+Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) would also produce more translation equivalents.
+Prediction Set 2 described translation equivalent knowledge with relations to the number of learnable words constrained by a child’s developmental level.
+Older children were predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Moreover, the most balanced children produced more singlets in their non-dominant language, whereas the least balanced children produced more singlets in their dominant language.
+Finally, Prediction Set 3 quantitatively demonstrated the patterns of translation equivalent learning under the three different accounts: the Avoidance Account, the Preference Account, and the Neutral Account.
+Crucially, these predictions should not be thought of as fully independent from each other, but instead illustrate the interrelatedness of measures of bilingual vocabulary under the Bilingual Vocabulary Model.
+In Study 2, we used vocabulary data from bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
 
@@
 ## 2.1.1 Participants
-```{r participants, include = FALSE}
+```{r participants, include = FALSE, cache = TRUE}
 # how many distinct participants and data point we have
@@
 
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
+Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
 (2020).
+Data collection was conducted in Montréal, Québec, Canada.
+Montréal is a multicultural city where both English and French are widely used in society.
 Some children took part in more than one in-lab study (n = `r length(d_multiple$baby_id)`); thus, they contributed data at more than one time point.
@@
 Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 2)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant.
-Data collection was conducted in Montréal, Québec, Canada.
-Montréal is a multicultural city where both English and French are widely used in society.
 Children’s demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
@@
 &nbsp;
-```{r table4, results="asis"}
+```{r table4, results="asis", cache = TRUE}
 # Create dataframe of demographics
@@
 
-```{r cdi_filled_by, include = FALSE}
+```{r cdi_filled_by, include = FALSE, cache = TRUE}
 # who filled out the CDIs
@@
 
-```{r vocabulary_descriptives, include = FALSE}
+```{r vocabulary_descriptives, include = FALSE, cache = TRUE}
 # average vocabulary scores
@@
 
-```{r balance_descriptives, include = FALSE}
+```{r balance_descriptives, include = FALSE, cache = TRUE}
 # Average BALANCE score (defined by vocabulary balance)
@@
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, but for young bilinguals balance can also be considered in terms of input in each language.
-We therefore compare the vocabulary balance with the proportion of exposure bilingual children received in their non-dominant language.
-To make values comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance, rather than the language that children heard most and least often.
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Pearson et al., 1997; Hurtado et al., 2013).
+We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
 For most children, the language in which they produced the most words was also the language that they heard most often (`r dominance_consistency$n_consistent` children, `r round(dominance_consistency$percentage_consistent,2)`%), although this was not the case for some children (`r dominance_consistency$n_inconsistent` children, `r round(dominance_consistency$percentage_inconsistent,2)`%).
-The correlation between vocabulary-defined BALANCE and the raw percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 5).
-Thus, these two constructs were related, although not identical.
+The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
+The imperfect correlation between these two measures could result from measurement error alone (for a discussion, see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
 &nbsp;
-```{r fig5, fig.cap="Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.", echo=FALSE, dpi=600, fig.align='center', out.width="75%", fig.pos = "H"}
+```{r fig4, fig.cap="Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.", echo=FALSE, dpi=600, fig.align='center', out.width="75%", fig.pos = "H", cache = TRUE}
 
@@
   labs(x = "Balance based on vocabulary (BALANCE)", 
-       y = "Balance based on exposure") +
+       y = "Percent exposure to \n the non-dominant language") +
   theme(text = element_text(size=16))
@@
 Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson’s correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations).
-
-```{r testing_prediction1, include = FALSE}
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations[^3]).
+
+[^3]: Note that there was a negative correlation between balance and age in our sample.
+We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.
+
+```{r testing_prediction1, include = FALSE, cache = TRUE}
 
@@
 Prediction 1a was that children with more balanced vocabularies would produce more translation equivalents.
-As shown in Figure 2 Row 2, our vocabulary data confirmed the prediction, _r_(`r cor_TE_balance$parameter[1]`) = `r round(cor_TE_balance$estimate[1],2)`, $p `r papaja::printp(cor_TE_balance$p.value[1])`$, where children with the most balanced vocabulary produced the most translation equivalents.
+As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, _r_(`r cor_TE_balance$parameter[1]`) = `r round(cor_TE_balance$estimate[1],2)`, $p `r papaja::printp(cor_TE_balance$p.value[1])`$, where children with the most balanced vocabulary produced the most translation equivalents.
 We further tested this prediction by dividing children into 5 balance subset groups (0 < BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, _F_(`r aov_TE_binBALANCE$df[1]`, `r aov_TE_binBALANCE$df[2]`) = `r aov_TE_binBALANCE$statistic[1]`, $p = `r papaja::printp(aov_TE_binBALANCE$p.value[1])`$.
-The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the least translation equivalents.
+The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the fewest translation equivalents.
 Detailed descriptive statistics are reported in Table 6.
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): _r_(`r cor_TE_word$parameter[1]`) = `r round(cor_TE_word$estimate[1],2)`, $p `r papaja::printp(cor_TE_word$p.value[1])`$, and dominant-language vocabulary (DOM): _r_(`r cor_TE_dom$parameter[1]`) = `r round(cor_TE_dom$estimate[1],2)`, $p `r papaja::printp(cor_TE_dom$p.value[1])`$.
-Figure 2 Row 2 further illustrates these relationships observed in our dataset.
+Figure 1 Row 2 further illustrates these relationships observed in our dataset.
 
 Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
-As shown in Figure 2 Row 2, we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
-
-```{r table5, results = "asis"}
+As shown in Figure 1 Row 2 (Observed data), we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
+
+```{r table5, results = "asis", cache = TRUE}
 # correlation among variables in the observed data
@@
 
-```{r table6, results="asis"}
+```{r table6, results="asis", cache = TRUE}
 
@@
 
-```{r testing_prediction2, include = FALSE}
+```{r testing_prediction2, include = FALSE, cache = TRUE}
 
@@
 
-```{r table7, results = "asis"}
+```{r table7, results = "asis", cache = TRUE}
 # read in the percentile file
 ws_percentile <- read.csv(here::here("data_keepers/ws_percentile.csv")) %>%
+  # filter out age ranges that is smaller than 18 months
+  filter(age_months_percentile >= 18) %>%
   # as the upper age limit of the CDI-WS is 30m, create a new row for age 30-33m
@@
                                            TRUE ~ as.character(age_months_percentile))) %>%
+  # select only variables relevant for the 90th percentile
+  select(age_months_percentile, EngWS_90percentile, FrWS_90percentile, Average_90percentile) %>%
   # rename variables for presentation
@@
 
-```{r testing_prediction3, include = FALSE}
+```{r testing_prediction3, include = FALSE, cache = TRUE}
 
@@
 
-```{r fig6, fig.cap="The number of simulated and observed translation equivalents plotted against each other.
+```{r fig5, fig.cap="The number of simulated and observed translation equivalents plotted against each other.
 The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
-The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="60%", out.width="60%"}
+The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="60%", out.width="60%", cache = TRUE}
 
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
   geom_point(shape = 23) +
@@
 &nbsp;
-```{r fig7, fig.cap="The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
-The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%"}
+```{r fig6, fig.cap="The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
+The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}
 
@@
   xlim(0,80) +
-  geom_smooth(method = "lm", se = F) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
   geom_point(shape = 23) +
@@
        title = "More than 300 word vocabulary (WORD)") +
-  geom_smooth(method = "lm", se = F) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
   geom_point(shape = 23) +

### 2021-12-06 paper/paper_TE_bilingual_vocabulary_model.tex
updated paper

@@
 This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
-To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model --- the Bilingual Vocabulary Model --- which quantifies translation equivalent knowledge as a function of the probability of learning words in each language.
+To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model --- the Bilingual Vocabulary Model --- which quantifies translation equivalent knowledge as a function of the probability of learning words in each language, and includes a bias parameter that varies the difficulty of learning translation equivalents according to each account.
 Study 2 tested model-derived predictions against vocabulary data from 200 French--English bilingual children aged 18--33 months.
@@
 For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
-These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word ``hygge'' refers to a type of cozy contentment and enjoyment of simplicity.
+Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
+However, in early development, most vocabulary words are reasonably concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.} (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual's developing lexicon due to the strong overlap in their semantics.
@@
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent.
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet).
 Is translation equivalent learning different from singlet learning?
@@
 The phenomenon of missing translation equivalents led theorists to propose that young bilingual children do not differentiate their languages, and thus tend to learn only a single word for each referent.
-This avoidance of translation equivalents was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman, 1992, 1994; Markman \& Wachtel, 1988).
+This avoidance of translation equivalents --- particularly for nouns --- was thought to be due to word learning biases such as mutual exclusivity, whereby children assume that a referent is only associated with one word at the basic level (Markman, 1992, 1994; Markman \& Wachtel, 1988).
 For example, when monolingual children see a familiar object (e.g., a cup) next to a novel object (e.g., a garlic press) and hear a novel word like ``wug,'' they assume that ``wug'' refers to the garlic press --- the object unknown to them --- rather than to the cup, the object for which they already know the word.
@@
 The researchers found that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence implied that singlets and translation equivalents are equally learnable.
+This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (\emph{n} = 27) in the context of a wide age range (8 months to 2.5 years).
 In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
@@
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
-That is, potential singlets might be ``too hard'' to be learned at a particular age.
-Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
+That is, many potential singlets could be ``too hard'' to be learned at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
+For example, an 14-month-old who has learned the French words \emph{``ballon''} and \emph{``chien''} would be capable of learning their translation equivalents ``ball'' and ``dog,'' as easier potential singlets like ``cup,'' but not later-acquired potential singlets like ``tomorrow,'' ``who,'' or ``not.'' Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
 
@@
 This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
-The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B \textbar{} A) where P(B \textbar{} A) = P(B) if A and B are independent (see Box 1 for a detailed example).
-The full model is shown in Figure 1.
-In the next paragraphs, we define each of the model parameters in detail, and these are also summarized in Table 1.
-
-\newpage
-\begingroup\fontsize{9}{11}\selectfont
-
-\begin{longtable}{>{\raggedright\arraybackslash}p{450px}}
-\toprule
-Box 1. Example of test for independence.\\
-\midrule
-To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
-To determine independence, we must know the probability of being a psychology major, the probability of being left handed, and the probability of being both a psychology major and left-handed.
-The numerator of these probabilities will be the number of students who are left-handed, psychology majors, and both (respectively) and the denominator will be the total population of students we are observing.
+The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent\footnote{To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
 Imagine a college of 1000 students.
@@
 Finally, if left-handedness and majoring in psychology are independent, we can predict the number of left-handed psychology students by multiplying the observed number of left-handed students (100) by the observed number of psychology students (200), and dividing by the total population of the college (1000), so for example 100×200/1000 = 20.
-Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.\\
-\bottomrule
-\end{longtable}
-\endgroup{}
-
-\begin{figure}
-
-{\centering \includegraphics[width=0.9\linewidth]{./Figure1_illustration}
-
-}
-
-\caption{Illustration of the Bilingual Vocabulary Model.}\label{fig:fig1}
-\end{figure}
+Thus, comparing expected and observed numbers can inform us about the independence of the underlying phenomena.}.
+In the next paragraphs, we define each of the model parameters in detail, and these are also summarized in Table 1.
 
@@
 For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-Very young children would not be expected to know many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins,'' due to children's immature cognitive machinery and conceptual development.
-A more reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
+However, not all children will be capable of learning all of the CDI words.
+Typically, words that are more concrete, occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019), and very young children would not be expected to produce many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
+Since there are no CDI norms for French-English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
-Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 245 words (retrieved from the Wordbank database; Frank et al., 2016).
-Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that this provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a child of Jamie's age could potentially acquire in each language.
-Thus, we might expect that Jamie could potentially have learned up to 245 English words and 245 French words, although he has thus far only learned 50 in English and 20 in French.
+Monolingual children his age with the very largest productive vocabularies (those at the 90th percentile averaging between English and French norms) produce a total of 240 words (retrieved from the Wordbank database version 0.3.1; Frank et al., 2016).
+Although there is likely considerable individual variability as to the cognitive capacity even amongst children of the same age, we argue that averaging between the English and French CDIs at the 90th percentile provides a reasonable --- if imperfect --- estimate of the number of learnable words (LEARNABLE) that a bilingual child of Jamie's age could potentially acquire in each language.
+Thus, we might expect that Jamie could potentially have learned up to 240 words in English and 240 words in French, although he has thus far only learned 50 in English and 20 in French.
+Note that we model learnability in terms of the number of words expected to be learnable at a particular age, rather than modeling the learnability of individual words at a particular age --- a simplifying choice that we will return to in the discussion section.
 
@@
 Under the Neutral Account, we would expect Jamie to learn 1×(50×20/245) = 4.1 translation equivalents.
-
-Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.
-
-Balance of vocabulary (BALANCE) is the proportion of total words that children produce in each language.
-For convenience, balance is defined in reference to the non-dominant language with the formula NONDOM/(DOM+NONDOM), such that scores can range from 0.0 (completely unbalanced) to 0.5 (completely balanced).
-For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29.
-Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many of the words he produces are translation equivalents).
-
-Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
-Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child --- that is, the total number of concepts that are lexicalized in either language.
-This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD).
-Finally, we can also calculate singlets that are produced in each language, that is words for which the child does not yet produce a translation equivalent.
-Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM).
-It is also possible to decompose children's word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
 
@@
 
-\hypertarget{current-research}{%
-\subsection{Current research}\label{current-research}}
-
-The current research aimed to better understand the nature of translation equivalent learning in bilingual children.
+Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.
+
+Balance of vocabulary (BALANCE) is the proportion of total words that children produce in each language.
+For convenience, balance is defined in reference to the non-dominant language with the formula NONDOM/(DOM+NONDOM), such that scores can range from 0.0 (completely unbalanced) to 0.5 (completely balanced).
+For example, since 18-month-old Jamie produces 50 dominant vocabulary words and 20 non-dominant vocabulary words, he would have a balance score of 0.29.
+Note that this calculation does not take into account overlap in meaning across the two languages (i.e., how many of the words he produces are translation equivalents).
+
+Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
+Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child --- that is, the total number of concepts that are lexicalized in either language.
+This can be calculated by subtracting the number of translation equivalents (TE) from the word vocabulary (WORD).
+Finally, we can also calculate singlets that are produced in each language, that is words for which the child does not yet produce a translation equivalent.
+Singlets in the dominant language (DOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from dominant-language vocabulary (DOM); singlets in the non-dominant language (NONDOM-SINGLET) can be calculated by subtracting translation equivalents (TE) from non-dominant language vocabulary (NONDOM).
+It is also possible to decompose children's word vocabulary (WORD) into the sum of TE, DOM-SINGLET, and NONDOM-SINGLET.
+
+Finally, we note two simplifying assumptions made by our model.
+First, we assume that children have the same communicative need in each language: our model does not account for cases where bilinguals might systematically refer to a concept in only one language and not the other (e.g., Grosjean, 2016).
+Second, we assume that translation equivalents refer to the same underlying concept.
+Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt \& Majid, 2013; Storm et al., 2015; White et al., 2020).
+However, we believe that these assumptions present a reasonable first approximation, especially in the case of young children, and return to these assumptions in the discussion section.
+
+\hypertarget{current-studies}{%
+\subsection{Current studies}\label{current-studies}}
+
+The current studies aimed to better understand the nature of translation equivalent learning in bilingual children.
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
@@
 
-Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction (see also Figure 1), which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction, which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
 Note that usually only three values are necessary to calculate all the other variables (see Table 1).
@@
 1 & Constant at 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 10 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 216\\
-2 & Varied at 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 161\\
+2 & Varied at 100, 200, 300, 400, 500, 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 20 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Constant at 1 & 301\\
 3 & Varied at 150, 300, 450, and 600 & Varied, ranging from 100 to LEARNABLE at an interval of 100 & Varied, ranging from 0 to DOM at an interval of 25 & Calculated as WORD = DOM + NONDOM & Calculated as BALANCE = NONDOM / (DOM+NONDOM) & Varied at 0.5 (Avoidance Account), 1 (Neutral Account), and 1.5 (Preference Account) & 166\\
@@
 Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 2, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
 Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
@@
 \item
-  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 2).
+  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 1).
 \end{itemize}
 
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 3 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 25 rows containing missing values (geom_smooth).
+\end{verbatim}
+
 \begin{figure}[H]
 
-{\centering \includegraphics[width=1.1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig2-1}
+{\centering \includegraphics[width=1.1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig1-1}
 
@@
 Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
-Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig2}
+Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig1}
 \end{figure}
@@
 
-Translation equivalent knowledge was simulated across children at three developmental levels (the number of LEARNABLE words = 300, 450, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
-In total, data from 161 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
+Translation equivalent knowledge was simulated across children at four developmental levels (the number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
+In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
 Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
-We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages, so that simulated children's concept vocabulary (CONCEPT) could be decomposed as the sum of TE (translation equivalents), DOM-SINGLET, and NONDOM-SINGLET.
-Figure 3 plots this decomposition for simulated children of different developmental levels, with vocabulary ranging from most balanced (BALANCE = .35 - .50), to medium balanced (BALANCE = .20 - .35), to least balanced (BALANCE = .00 - .02).
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig3-1}
-
-}
-
-\caption{Number of translation equivalents (TE) and singlets in dominant (DOM-SINGLET) and non-dominant language (NONDOM-SINGLET) across different developmental levels/ages, which sets the number of LEARNABLE words.
-Panel A represents the model simulation in Study 1, where developmental levels of simulated children are set at three values: LEARNABLE = 300, 450, and 600.
-Panel B represents the observed vocabulary data in Study 2, where developmental level was divided into 3 subsets with children of 18-22 months (left), children of 23-27 months (middle), and children of 28-33 months (right).
-Proportion of balance (BALANCE) was divided into three groups, where the least balanced group had a range of .00 - .20 vocabulary balance, the medium balanced group had a range of .20 - .35, and the most balanced group had a range of .35 - .50.}\label{fig:fig3}
-\end{figure}
-
-In general, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
-Moreover, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
+Figure 2 plots the number of TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
+Across the three Panels (1A, 1B, and 1C), simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+In Panel 1A, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
 Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Overall, we generated 3 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
+Panels 1B and 1C show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
+Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
 Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
@@
 
+\begin{verbatim}
+## Warning: Removed 4 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 1 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig2-1}
+
+}
+
+\caption{Number of translation equivalents (TE; Panel A), singlets in dominant (DOM-SINGLET; Panel B) and singlets in the non-dominant language (NONDOM-SINGLET; Panel C) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+Row 1 represents the simulated data in Study 1.
+Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig2}
+\end{figure}
+
 \hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}{%
@@
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50\% more likely to be learned than singlets).
-In Figure 4, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
+In Figure 3, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
 Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
@@
 
-{\centering \includegraphics[width=1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig4-1}
+{\centering \includegraphics[width=1\linewidth,height=0.8\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig3-1}
 
@@
 
-\caption{Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.}\label{fig:fig4}
+\caption{Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.}\label{fig:fig3}
 \end{figure}
+
+\hypertarget{summary-of-model-predictions}{%
+\subsection{1.4 Summary of model predictions}\label{summary-of-model-predictions}}
+
+Study 1 presented predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
+Prediction Set 1 illustrated that the more words children produced in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they would produce.
+Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) would also produce more translation equivalents.
+Prediction Set 2 described translation equivalent knowledge with relations to the number of learnable words constrained by a child's developmental level.
+Older children were predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Moreover, the most balanced children produced more singlets in their non-dominant language, whereas the least balanced children produced more singlets in their dominant language.
+Finally, Prediction Set 3 quantitatively demonstrated the patterns of translation equivalent learning under the three different accounts: the Avoidance Account, the Preference Account, and the Neutral Account.
+Crucially, these predictions should not be thought of as fully independent from each other, but instead illustrate the interrelatedness of measures of bilingual vocabulary under the Bilingual Vocabulary Model.
+In Study 2, we used vocabulary data from bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
 
@@
 
-Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at the XYZ lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Data collection was conducted in Montréal, Québec, Canada.
+Montréal is a multicultural city where both English and French are widely used in society.
 Some children took part in more than one in-lab study (n = 28); thus, they contributed data at more than one time point.
@@
 Within the 229 data points, 59.83\% of children were English-dominant and 40.17\% were French-dominant.
-Data collection was conducted in Montréal, Québec, Canada.
-Montréal is a multicultural city where both English and French are widely used in society.
 Children's demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
@@
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, but for young bilinguals balance can also be considered in terms of input in each language.
-We therefore compare the vocabulary balance with the proportion of exposure bilingual children received in their non-dominant language.
-To make values comparable, the language designated as DOM and NONDOM was based on vocabulary-defined dominance, rather than the language that children heard most and least often.
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Pearson et al., 1997; Hurtado et al., 2013).
+We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
 For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79.04\%), although this was not the case for some children (48 children, 20.96\%).
-The correlation between vocabulary-defined BALANCE and the raw percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 5).
-Thus, these two constructs were related, although not identical.
+The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
+The imperfect correlation between these two measures could result from measurement error alone (for a discussion, see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 
-{\centering \includegraphics[width=0.75\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig5-1}
+{\centering \includegraphics[width=0.75\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig4-1}
 
@@
 
-\caption{Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.}\label{fig:fig5}
+\caption{Correlation between balance defined by vocabulary (BALANCE) and balance defined by exposure.}\label{fig:fig4}
 \end{figure}
@@
 Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson's correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations).
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between balance and age in our sample.
+We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.}).
 
 Prediction 1a was that children with more balanced vocabularies would produce more translation equivalents.
-As shown in Figure 2 Row 2, our vocabulary data confirmed the prediction, \emph{r}(227) = 0.25, \(p < .001\), where children with the most balanced vocabulary produced the most translation equivalents.
+As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, \emph{r}(227) = 0.25, \(p < .001\), where children with the most balanced vocabulary produced the most translation equivalents.
 We further tested this prediction by dividing children into 5 balance subset groups (0 \textless{} BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, \emph{F}(4, 224) = 3.61, \(p = .007\).
-The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the least translation equivalents.
+The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the fewest translation equivalents.
 Detailed descriptive statistics are reported in Table 6.
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): \emph{r}(227) = 0.90, \(p < .001\), and dominant-language vocabulary (DOM): \emph{r}(227) = 0.76, \(p < .001\).
-Figure 2 Row 2 further illustrates these relationships observed in our dataset.
+Figure 1 Row 2 further illustrates these relationships observed in our dataset.
 
 Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
-As shown in Figure 2 Row 2, we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
+As shown in Figure 1 Row 2 (Observed data), we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
 
@@
 \endhead
-Age (in month) & 0.95*** & -0.24*** & 0.65*** & 0.69*** & 0.45*** & 0.48*** & 0.65*** & 0.19** & 0.69***\\
-LEARNABLE &  & -0.23*** & 0.62*** & 0.65*** & 0.43*** & 0.44*** & 0.62*** & 0.21** & 0.66***\\
+Age (in month) & 0.96*** & -0.24*** & 0.65*** & 0.69*** & 0.45*** & 0.48*** & 0.65*** & 0.19** & 0.69***\\
+LEARNABLE &  & -0.23*** & 0.62*** & 0.66*** & 0.43*** & 0.45*** & 0.63*** & 0.21** & 0.67***\\
 BALANCE &  &  & -0.07 & -0.29*** & 0.35*** & 0.25*** & -0.58*** & 0.63*** & -0.21**\\
@@
 \midrule
-16 & 129.0 & 97.0 & 113.0\\
-17 & 201.5 & 161.1 & 181.3\\
-18 & 268.7 & 221.1 & 244.9\\
-19 & 330.5 & 277.0 & 303.8\\
-20 & 387.0 & 328.8 & 357.9\\
+18 & 259.4 & 220.1 & 239.7\\
+19 & 321.5 & 274.4 & 297.9\\
+20 & 378.3 & 325.1 & 351.7\\
+21 & 429.8 & 372.2 & 401.0\\
+22 & 476.1 & 415.7 & 445.9\\
 \addlinespace
-21 & 438.1 & 376.5 & 407.3\\
-22 & 483.8 & 420.0 & 451.9\\
-23 & 524.2 & 459.4 & 491.8\\
-24 & 559.3 & 494.7 & 527.0\\
-25 & 589.0 & 525.9 & 557.4\\
+23 & 517.0 & 455.6 & 486.3\\
+24 & 552.6 & 492.0 & 522.3\\
+25 & 583.3 & 524.8 & 554.0\\
+26 & 609.9 & 554.0 & 582.0\\
+27 & 633.0 & 579.6 & 606.3\\
 \addlinespace
-26 & 613.4 & 553.0 & 583.2\\
-27 & 632.4 & 575.9 & 604.1\\
-28 & 646.0 & 594.7 & 620.4\\
-29 & 654.3 & 609.4 & 631.9\\
-30 - 33 & 657.9 & 620.0 & 638.9\\
+28 & 652.5 & 601.7 & 627.1\\
+29 & 668.5 & 620.2 & 644.3\\
+30 - 33 & 680.9 & 635.1 & 658.0\\
 \bottomrule
@@
 Our model showed an excellent model fit of \(R^2\) = 0.96, indicating that our model explained 96\% of the variance in bilinguals' translation equivalent knowledge.
-The linear regression model estimated a BIAS coefficient of 1.02, \(p < .001\).
+The linear regression model estimated a BIAS coefficient of 1.04, \(p < .001\).
 This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words.
@@
 Cook's distance was estimated for our linear regression model listed above and identified two data points with a cook's distance over 0.4.
-After removing those two data points, the linear regression model returned a coefficient of 1.05, \(p < .001\), with \(R^2\) = 0.96.
+After removing those two data points, the linear regression model returned a coefficient of 1.07, \(p < .001\), with \(R^2\) = 0.96.
 As the model fit was similar to the model without eliminating the two outlier data points, we proceeded with the full data set keeping the two potential outlier data points.
@@
 
-{\centering \includegraphics[width=0.6\linewidth,height=0.6\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig6-1}
+{\centering \includegraphics[width=0.6\linewidth,height=0.6\textheight]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig5-1}
 
@@
 The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
-The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.}\label{fig:fig6}
+The diagonal dashed line represents the case where the bias parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal, and the solid blue line represents the model predictions.}\label{fig:fig5}
 \end{figure}
@@
 To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
-The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.02, \(p < .001\), whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.22, \(p < .001\).
-Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size (\(R^2\) = 0.97) than children with smaller vocabulary size (\(R^2\) = 0.88).
+The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.04, \(p < .001\), whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.21, \(p < .001\).
+Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size (\(R^2\) = 0.96) than children with smaller vocabulary size (\(R^2\) = 0.88).
 Overall, this analysis suggests that translation equivalent learning for children with larger vocabularies corresponds best to the Neutral Account, but translation equivalent learning for children with smaller vocabularies corresponds best to the Preference Account.
@@
 
-{\centering \includegraphics[width=1.2\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig7-1}
+{\centering \includegraphics[width=1.2\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig6-1}
 
@@
 \caption{The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
-The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.}\label{fig:fig7}
+The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.}\label{fig:fig6}
 \end{figure}

### 2021-12-06 simulations_old_versions/Analyze_TEs.Rmd
updated paper

@@
 ```{r}
-cdi_ws = readRDS(here::here("Data","output_analyzeSent.Rdata"))
+cdi_ws = readRDS(here::here("data_raw","output_analyzeSent.Rdata"))
 ```
@@
   pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("WORD", "DOM", "NONDOM"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             DOM = "Panel 1A: Dominant vocabulary \n (DOM)", 
-                             NONDOM = "Panel 1B: Non-dominant vocabulary \n (NONDOM)",
-                             WORD = "Panel 1C: Total vocabulary \n (WORD)")) %>%
+                             DOM = "Panel 1B: Dominant vocabulary",
+                             NONDOM = "Panel 1C: Non-dominant vocabulary",
+                             WORD = "Panel 1A: Total vocabulary")) %>%
   ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
   stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
-  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
+#  scale_color_brewer(palette="Accent")+
+  scale_color_manual(values=c("#762a83", "#abd9e9", "#a6d96a", "#fdae61", "#d53e4f")) + 
+  scale_linetype_manual(values=c("solid", "solid", "solid", "solid", "solid")) +
   theme_minimal() + 
@@
        linetype = "Vocabulary balance (BALANCE)",
-       title = "1) Simulated data") +
+       title = "Simulated data") +
   guides(color = guide_legend(reverse = T, nrow = 1),
@@
   pivot_longer(c(word_vocab, TotalNW_Dom, TotalNW_NonDom), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("TotalNW_Dom", "TotalNW_NonDom", "word_vocab"))) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "TotalNW_Dom", "TotalNW_NonDom"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             TotalNW_Dom = "Panel 2A: Dominant vocabulary \n (DOM)", 
-                             TotalNW_NonDom = "Panel 2B: Non-dominant vocabulary \n (NONDOM)",
-                             word_vocab = "Panel 2C: Total vocabulary \n (WORD)")) %>%
+                             TotalNW_Dom = "Panel 2B: Dominant vocabulary", 
+                             TotalNW_NonDom = "Panel 2C: Non-dominant vocabulary",
+                             word_vocab = "Panel 2A: Total vocabulary")) %>%
   mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) 
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
   ggplot(aes(x = number, y = number_of_te, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
   stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
-  scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
+  scale_color_manual(values=c("#762a83", "#abd9e9", "#a6d96a", "#fdae61", "#d53e4f")) + 
+  scale_linetype_manual(values=c("solid", "solid", "solid", "solid", "solid")) +
   theme_minimal() + 
@@
        linetype = "Vocabulary balance (BALANCE)",
-       title = "2) Observed data") +
+       title = "Observed data") +
   guides(colour = guide_legend(reverse = T, nrow = 1),
@@
 
-ggsave("Output/plot_Simulation1plusData.png", plot_Simulation1plusData,
+ggsave("Output/BU_plot_Simulation1plusData.png", plot_Simulation1plusData,
        width = 10, height = 10)
@@
   facet_grid(. ~ LEARNABLE_label) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) +
+  scale_fill_manual(values = c("#bdbdbd", "#737373", "#252525")) +
   scale_x_discrete(labels = function(bin_BALANCE) str_wrap(bin_BALANCE, width = 10)) + 
@@
         panel.spacing.x = unit(2, "lines")) + 
+  guides(fill=guide_legend(nrow=3,byrow=TRUE)) + 
   labs(x = "\nVocabulary balance (BALANCE)", 
@@
   facet_grid(. ~ LEARNABLE_subset) +
-  scale_fill_manual(values = c("#1f78b4", "#a6cee3", "#33a02c")) + 
+  scale_fill_manual(values = c("#bdbdbd", "#737373", "#252525")) +
   scale_x_discrete(labels = function(BalSubset) str_wrap(BalSubset, width = 10)) + 
@@
   geom_abline(linetype = 2, alpha = 0.5) +
-  theme_minimal() +
+  theme_minimal(base_size = 18) +
   theme(plot.title = element_text(hjust = 0.5))
@@
             ncol = 2, nrow = 1) %>%
-  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model"))
+  annotate_figure(bottom = text_grob("Number of TEs predicted by the Bilingual Vocabulary Model", size= 18))
 
-ggsave(here("Output/plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
+ggsave(here("Output/BU_plot_Simulation3_data_2WORDgroups.png"), TEprediction_percentile_facetByW300,
        height=5,width=10)

### 2021-12-06 Output/ws_percentile_plot.R

@@ -1,0 +1,31 @@
+library(here)
+library(tidyverse)
+library(tidylog)
+library(wordbankr)
+
+ws_percentile_plot <- ws_percentile %>%
+  pivot_longer(-age_months_percentile, names_to = "type", values_to = "number_of_words") %>%
+  mutate(type = recode(type,
+                       Average_90percentile = "Average",
+                       EngWS_90percentile = "English at the 90th percentile", 
+                       FrWS_90percentile = "French at the 90th percentile")) %>%
+  ggplot(aes(x= age_months_percentile, y = number_of_words, color = type, linetype = type)) +
+  theme_minimal(base_size = 18) +
+  geom_line(size = 1.5) +
+  scale_x_continuous(breaks=seq(16, 130, 1)) +
+  scale_color_manual(values=c("#984ea3", "#fb8072", "#80b1d3")) + 
+  scale_linetype_manual(values=c("solid", "dashed", "dashed")) +
+  labs(x = "Age (months)", 
+       y = "Number of productive vocabulary", 
+       color = "Language",
+       linetype = "Language") +
+  theme(legend.position="bottom") +
+  guides(colour = guide_legend(nrow = 3),
+         linetype = guide_legend(nrow = 3)) 
+
+ggsave("/Users/racheltsui/Documents/GitHub/TE_bilingual_vocabulary_model/Output/BU_ws_percentile_plot.png", 
+       ws_percentile_plot,
+       width = 6, height = 6)
+  
+
+  

### 2021-12-07 paper/paper_TE_bilingual_vocabulary_model.Rmd
Updated paper

@@
 
-[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word “hygge” refers to a type of cozy contentment and enjoyment of simplicity.
+[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word _“hygge”_ refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults’ concepts may not be congruent across languages (e.g., Malt & Majid, 2013).
@@
 In our data set, developmental level was approximated by children’s age.
-Figure 3 Panel B shows the concept vocabulary (CONCEPT) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
-To illustrate the acquisition of translation equivalents and singlets at different developmental levels, we divided children into three age groups: younger children of 18–22 months, middle children of 23–27 months, and older children of 28–33 months.
+Figure 2 Panel B shows the number of translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 
-# testing prediction 2a: To compare CONCEPT among the 3 LEARNABLE groups in the stacked bar chart
-
+# testing prediction 2a: Relationship between CONCEPT and Age
 ## correlation between age (in months) and CONCEPT
@@
 
-## one-way ANOVE to compare CONCEPT among the 3 LEARNABLE groups
-keepers_ws_TE_3LEARNABLEsubset <- keepers_ws_TE %>%
-  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months (V = 244.9 - 451.9)",
-                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months (V = 491.8 - 604.1)",
-                              TRUE ~ "28-33 months (V = 620.4 - 638.9)"))
-
-aov_age_CONCEPT <- tidy(aov(concept_vocab ~ LEARNABLE_subset, data = keepers_ws_TE_3LEARNABLEsubset))
-
-pairwise.t.test(keepers_ws_TE_3LEARNABLEsubset$concept_vocab, keepers_ws_TE_3LEARNABLEsubset$LEARNABLE_subset,
-                p.adjust.method = "BH")
-
-## mean CONCEPT by LEARNABLE subset
-mean_concept_LEARNABLEsubset <- keepers_ws_TE_3LEARNABLEsubset %>%
-  group_by(LEARNABLE_subset) %>%
-  summarise(mean_concept = mean(concept_vocab, na.rm = T),
-            sd_concept = sd(concept_vocab, na.rm = T),
-            min_concept = min(concept_vocab, na.rm = T),
-            max_concept = max(concept_vocab, na.rm = T)) %>%
-  mutate(across(is.numeric, ~ round(.,1)))
-
-# testing prediction 2b: To compare TE among the 3 LEARNABLE groups in the stacked bar chart
-
+# testing prediction 2b: Relationship between TE and Age
 ## correlation between age (in months) and TE
@@
 
-## one-way ANOVA to compare TE among LEARNABLE
-stacked_byLEARNABLE_wide <- keepers_ws_TE %>%
-  select(baby_id, age_months_percentile, balance_vocab, word_vocab, concept_vocab, 
-         number_of_te, total_singlet_dom, total_singlet_nondom) %>%
-  # Subset developmental level (i.e., Average_90percentile) by 3 levels 
-  mutate(LEARNABLE_subset = case_when(age_months_percentile <= 22 ~ "18-22 months",
-                              age_months_percentile > 22 & age_months_percentile <= 27 ~ "23-27 months",
-                              TRUE ~ "28-33 months")) %>%
-  # Subset balance level by 3 groups
-  mutate(BalSubset = case_when(balance_vocab > 0.35 ~ "Most Balanced",
-                               balance_vocab <= 0.20 ~ "Least Balanced",
-                               TRUE ~ "Medium Balanced"))
-
-aov_age_TE <- tidy(aov(number_of_te ~ LEARNABLE_subset, data = stacked_byLEARNABLE_wide))
-
-pairwise.t.test(stacked_byLEARNABLE_wide$number_of_te, stacked_byLEARNABLE_wide$LEARNABLE_subset,
-                 p.adjust.method = "BH")
-
-## mean TE by LEARNABLE subset
-mean_TE_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
-  group_by(LEARNABLE_subset) %>%
-  summarise(mean_TE = mean(number_of_te, na.rm=T), 
-            sd_TE = sd(number_of_te, na.rm=T),
-            min_TE = min(number_of_te, na.rm=T),
-            max_TE = max(number_of_te, na.rm=T)) %>%
-  mutate(across(is.numeric, ~ round(.,1)))
-
-# testing prediction 2c: To compare DOM-SINGLET among the 3 LEARNABLE groups in the stacked bar chart
-
+# testing prediction 2c: Relationship among DOM-SINGLET, Age, and BALANCE
 ## correlation between DOM-SINGLET and age (in month)
@@
 
-## one-way ANOVA to compare DOM-SINGLET among LEARNABLE
-aov_balance_DOMSING <- tidy(aov(total_singlet_dom ~ BalSubset, data = stacked_byLEARNABLE_wide))
-
-pairwise.t.test(stacked_byLEARNABLE_wide$total_singlet_dom, stacked_byLEARNABLE_wide$BalSubset,
-                 p.adjust.method = "BH")
-
-## mean DOM-SINGLET by LEARNABLE subset
-mean_DOMSINGLET_LEARNABLEsubset <- stacked_byLEARNABLE_wide %>%  
-  group_by(BalSubset) %>%
+## mean DOM-SINGLET by BALANCE subset
+mean_DOMSINGLET_BALANCEsubset <- keepers_ws_TE %>%
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  group_by(bin_BALANCE) %>%
   summarise(mean_DOM_SINGLET = mean(total_singlet_dom, na.rm=T), 
@@
             max_DOM_SINGLET = max(total_singlet_dom, na.rm=T)) %>%
-  mutate(across(is.numeric, ~ round(.,1)))
-
-# testing Prediction 2d: To compare NONDOM-SINGLET among the 3 LEARNABLE groups in the stacked barchart
-
+  mutate(across(is.numeric, ~ round(.,0)))
+
+# testing Prediction 2d: Relationship among NONDOM-SINGLET, Age, and BALANCE 
 ## correlation between NONDOM-SINGLET and age (in month)
@@
 
-## one-way ANOVA to compare NONDOM-SINGLET among LEARNABLE
-aov_balance_NONDOMSING <- tidy(aov(total_singlet_nondom ~ BalSubset, data = stacked_byLEARNABLE_wide))
-
-pairwise.t.test(stacked_byLEARNABLE_wide$total_singlet_nondom, stacked_byLEARNABLE_wide$BalSubset,
-                 p.adjust.method = "BH")
-
-## mean NONDOM-SINGLET by LEARNABLE subset
-mean_NONDOMSINGLET_LEARNABLEsubset <-stacked_byLEARNABLE_wide %>% 
-  group_by(BalSubset) %>%
+## mean DOM-SINGLET by BALANCE subset
+mean_NONDOMSINGLET_BALANCEsubset <- keepers_ws_TE %>%
+  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
+                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
+  group_by(bin_BALANCE) %>%
   summarise(mean_NONDOM_SINGLET = mean(total_singlet_nondom, na.rm=T), 
@@
             max_NONDOM_SINGLET = max(total_singlet_nondom, na.rm=T)) %>%
-  mutate(across(is.numeric, ~ round(.,1)))
+  mutate(across(is.numeric, ~ round(.,0)))
+
 ```
@@
 We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, _r_(`r cor_age_CONCEPT$parameter[1]`) = `r round(cor_age_CONCEPT$estimate[1],2)`, $p `r papaja::printp(cor_age_CONCEPT$p.value[1])`$, and therefore confirmed the prediction.
-This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, ($F$(`r aov_age_CONCEPT$df[1]`, `r aov_age_CONCEPT$df[2]`) = `r round(aov_age_CONCEPT$statistic[1],2)`, $p `r papaja::printp(aov_age_CONCEPT$p.value[1])`$).
-Older children of 28–33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[3],2)` ($p$s < .001), middle children of 23–27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[2],2)`, and younger children of 18–22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of `r round(mean_concept_LEARNABLEsubset$mean_concept[1],2)` ($p$s < .001).
 
 Prediction 2b was that older children would produce more translation equivalents than younger children.
-First, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, _r_(`r cor_age_TE$parameter[1]`) = `r round(cor_age_TE$estimate[1],2)`, $p `r papaja::printp(cor_age_TE$p.value[1])`$, and therefore confirmed the prediction.
-In a one-way ANOVA with age group as factor, we further found that groups differed in how many translation equivalents they produced ($F$(`r aov_age_TE$df[1]`, `r aov_age_TE$df[2]`) = `r round(aov_age_TE$statistic[1],2)`, $p `r papaja::printp(aov_age_TE$p.value[1])`$).
-Younger children of 18–22 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[1],2)` translation equivalents, middle children of 23–27 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[2],2)` translation equivalents, and older children of 28–33 months produced an average of `r round(mean_TE_LEARNABLEsubset$mean_TE[3],2)` translation equivalents ($p$s < .01).
+We observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, _r_(`r cor_age_TE$parameter[1]`) = `r round(cor_age_TE$estimate[1],2)`, $p `r papaja::printp(cor_age_TE$p.value[1])`$, and therefore confirmed the prediction.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_DOMSING$parameter[1]`) = `r round(cor_age_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_age_DOMSING$p.value[1])`$, and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), _r_(`r cor_BALANCE_DOMSING$parameter[1]`) = `r round(cor_BALANCE_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_DOMSING$p.value[1])`$.
-As shown in Figure 3 Panel B, children were divided into least balanced (range of balance: .00 - .20), medium balanced (range of balance: .20 - .35) and most balanced (range of balance: .35 - .50) groups (i.e., the same criteria as in Figure 5).
-In a one-way ANOVA with balance group as a between-subjects factor, we observed that the least balanced children produced the most singlets in their dominant language ($p$s < .001), with the least balanced, medium balanced, most balanced children producing respectively: `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[1],2)`, `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[2],2)`, and `r round(mean_DOMSINGLET_LEARNABLEsubset$mean_DOM_SINGLET[3],2)` words in their dominant language ($F$(`r aov_balance_DOMSING$df[1]`, `r aov_balance_DOMSING$df[2]`) = `r round(aov_balance_DOMSING$statistic[1],2)`, $p `r papaja::printp(aov_balance_DOMSING$p.value[1])`$).
+As illustrated in Figure 2 Panel 2B, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[1],0)`), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[5],0)`).
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_NONDOMSING$parameter[1]`) = `r round(cor_age_NONDOMSING$estimate[1],2)`, $p = `r papaja::printp(cor_age_NONDOMSING$p.value[1])`$, and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), _r_(`r cor_BALANCE_NONDOMSING$parameter[1]`) = `r round(cor_BALANCE_NONDOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_NONDOMSING$p.value[1])`$.
-In a one-way ANOVA with balance group as a between-subjects factor, we confirmed that children who differed in how balanced their vocabulary knowledge was also differed in how many singlets they produced in their non-dominant language ($F$(`r aov_balance_NONDOMSING$df[1]`, `r aov_balance_NONDOMSING$df[2]`) = `r round(aov_balance_NONDOMSING$statistic[1],2)`, $p `r papaja::printp(aov_balance_NONDOMSING$p.value[1])`$).
-As shown in Figure 3 Panel B, we observed that children produced very few singlets in their non-dominant language, although the most balanced children produced the most singlets in their non-dominant language (mean of the most balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[3],2)` > mean of the medium balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[2],2)` > mean of the most balanced children = `r round(mean_NONDOMSINGLET_LEARNABLEsubset$mean_NONDOM_SINGLET[1],2)`; $p$s < .001).
+As shown in Figure 2 Panel 2C, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[5],0)`) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[1],0)`).
 
@@
 Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
-To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate the BIAS parameter.
-
-First, we will walk through the parameters in this model.
-The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-The number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank; Frank et al., 2016).
-Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
-
-```{r table7, results = "asis", cache = TRUE}
+To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE, and we allowed the model to estimate the BIAS parameter.
+
+```{r percentile_info, include = FALSE, cache = TRUE}
 # read in the percentile file
@@
   mutate(across(where(is.numeric),round,1))
-
+```
+
+First, we will walk through the parameters in this model.
+The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
+Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
+Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
+For example, for an 18 month-old infant, the denominator was `r round(ws_percentile$"Average (LEARNABLE)"[1],1)` words which was calculated by averaging the `r round(ws_percentile$"Number of English words produced at 90th percentile"[1],1)` English words and `r round(ws_percentile$"Number of French words produced at 90th percentile"[1],1)` French words, based on what 18-month-old children would typically produce at the 90th percentile.
+For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
+
+```{r table7, results = "asis", cache = TRUE}
 kable(ws_percentile,
@@
   column_spec(2:4, width = "125px")
-
 ```
@@
   mutate(Predicted_TE_percentile = total_words_dom * total_words_nondom/Average_90percentile) %>%
-  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_90percentile, total_words_dom, total_words_nondom, 
+  select(baby_id, ID_testdate, Predicted_TE_percentile, number_of_te, Average_75percentile, Average_90percentile, total_words_dom, total_words_nondom, 
          word_vocab, age_months_binned)
 
-## linear regression model
+## linear regression model (90th percentile)
 test_prediction3_model <- keepers_ws_TE_simulation %>%
   lm(number_of_te*(Average_90percentile) ~ 0 + total_words_dom:total_words_nondom, data = .)
+
+## linear regression model (75th percentile for footnote 4)
+test_prediction3_model_75percentile <- keepers_ws_TE_simulation %>%
+  lm(number_of_te*(Average_75percentile) ~ 0 + total_words_dom:total_words_nondom, data = .)
 
@@
 The linear regression model estimated a BIAS coefficient of `r round(summary(test_prediction3_model)$coefficient[1],2)`, $p `r papaja::printp(summary(test_prediction3_model)$coefficient[4])`$.
-This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words.
-
-To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child’s expected translation equivalent knowledge (setting BIAS = 1), which is plotted against our observed data in Figure 6.
+This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^4].
+
+[^4]: We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
+Overall, model fit was slightly reduced compared to our main model, although fit was still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`, $p `r papaja::printp(summary(test_prediction3_model_75percentile)$coefficient[4])`$.
+The estimated BIAS parameter was lower at `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`.
+This change in the BIAS parameter is mathematically expected due to the reciprocal nature of LEARNABLE and BIAS, an issue we will return to in the discussion.
+
+To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child’s expected translation equivalent knowledge (setting BIAS = 1; also setting the intercept at 0 to replicate our linear regression model), which is plotted against our observed data in Figure 5.
 Expected and observed translation equivalents were closely aligned with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1), suggesting that the Neutral Account provides a parsimonious explanation for bilinguals’ translation equivalent knowledge.
@@
 
-Despite the good overall fit to the data, a close examination of Figure 6 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
+Despite the good overall fit to the data, a close examination of Figure 5 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
 Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
@@
 Three sets of model predictions were confirmed in our empirical dataset.
+In Table 8, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+
+```{r table8, results = "asis", cache = TRUE}
+
+table8 <- data.frame(
+  model_predictions = c(
+    "1A: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE).", 
+    "1B: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE). ",
+    "1C: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge",
+    "2A: Older children will have larger concept vocabularies (CONCEPT) than younger children.",
+    "2B: Older children will produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).",
+    "2C: Older children will produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least balanced vocabulary (BALANCE) will produce the most DOM-SINGLET.",
+    "2D: Older children will produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most balanced vocabulary (BALANCE) will produce the most NONDOM-SINGLET."
+    ),
+  previous_literature = c(
+    "Children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).",
+    "The number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+\n There is a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013)",
+    "Vocabulary size in the non-dominant language was positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).",
+    "Older children had larger concept vocabularies (Pearson et al., 1993).",
+    "Older bilingual children knew more translation equivalents than younger children  (David & Wei, 2008; Legacy et al., 2016)",
+    "Bilingual children learned words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011)",
+    ""
+  ),
+  observattion_in_our_data = c("Yes", "Yes", "Yes", "Yes", "Yes", "Yes", "")
+  ) %>%
+  rename("Model predictions" = model_predictions,
+         "Evidence in previous literature" = previous_literature,
+         "Did we observe the same pattern in our empirical data?" = observattion_in_our_data)
+
+landscape(
+  kable(table8,
+        align = "llc",
+        format = "latex", booktabs = TRUE,
+        caption = "Table 8. Summary of model predictions and evidence from previous literature and our empirical data.") %>%
+    column_spec(c(1), width = "200px") %>%
+    column_spec(c(2), width = "250px") %>%
+    column_spec(c(3), width = "150px") %>%
+    kableExtra::kable_styling(font_size = 7)
+)
+```
 
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
-For example, a caregiver may choose to say to their English–French bilingual child “Can you grab the livre?” if they know their child understands the French word “livre” but not the English equivalent “book”.
-This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English–French bilingual child “Can you grab the _livre_?” if they know their child understands the French word _“livre”_ but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
-It would be interesting for future studies to take into consideration the such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+However, our model does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
 Such a model may better characterize and predict individual differences in bilingual vocabulary development.
 Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data and/or data of a different bilingual.
+
+Note also that the age group that our paper currently looked at has been suggested to learn substantially more nouns before verbs or other word classes (i.e., noun bias; Gentner, 1982).
+It has been suggested that nouns that are frequently found in the input or have concrete meanings appeared to be the easiest for young children to grasp (Braginsky et al., 2019; Goodman et al., 2008).
+On the other hand, predicates including verbs, adjectives and adverbs appear to be more difficult to learn because their meaning is more abstract and is dependent on relationships among entities (Gertner, 2006).
+Due to these differences, different lexical categories seem to be learned by young children in different ways (Braginsky et al., 2019; Imai et al., 2005).
+Since meanings of nouns are more concrete, the boundaries to categorize objects are more stable.
+Yet, predicates have boundaries that are less clearly marked without knowledge of neighboring words (Imai et al., 2008; Saji et al., 2020).
+Future investigation may explore if bilingual vocabulary learning would differ among nouns and other lexical categories.
 
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a “bowl” in English but an _“assiette”_ [plate] in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
-Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) [@Bosch_Ramon-Casas_2014].
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storm et al., 2015; White et al., 2020).
+Therefore, it seems reasonable for our model to take the simplified view that early vocabulary may be treated similarly across languages, especially for the age group we were looking at.
+Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where vocabularies are kept language-specific and distinct across languages.
+
+Another limitation of our approach is that we modeled the number of words that might be learnable at a particular age, rather than modeling the learnability of individual words.
+For example, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English–French pair “banana” – _“banane”_) more easily than those that do not share similar phonological form (e.g., the English–French pair “dog” – _“chien”_) (Bosch & Ramon-Casas, 2014; Mitchell et al., 2021).
 Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
-While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
-Moreover, if they do prove to be important, such factors could be added to future iterations of the Bilingual Vocabulary Model.
-
-Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
-However, according to the Complementarity Principle [@Grosjean_2016], bilinguals may have different experiences in each of their languages.
-For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., “soap”, “bath”, “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
+There are many other individual differences between words that affect their learnability, such as frequency of occurrence in speech input, concreteness, babiness, word length, semantic category, etc.
+(Braginsky et al., 2019).
+It is likely that within the set of words that are learnable at a given age, some will be more easily learned than others, which our model does not take into account.
+However, the close correspondence between our model and data from bilingual children suggest that even if our model is a simplification, deviations from our assumptions might have a relatively small impact.
+On the other hand, individual item differences could be particularly consequential at younger ages, and the apparent shift we observed in development from the enhancement account to the avoidance account could be an artifact of this issue.
+Future iterations of the Bilingual Vocabulary Model could explicitly add such word-level factors into the model and test the importance of those factors in early bilingual vocabulary development.
+
+Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
+However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., “soap,” “bath,” “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
 At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (“mommy”, “daddy”), social routines (“hi”, “bye”, “yes”, “no”), and common nouns (“ball”, “dog”) are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
-Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquired across languages and cultures.
+It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
 Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
@@
 Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
-Under the Bilingual Vocabulary Model, these two parameters jointly and inversely predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter increases by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
-Our model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce.
-Small changes to this approach (e.g., taking the number of words children at the 95th percentile produce) would likely not drastically alter our results, nor change the qualitative shift that we observed in our data.
-Nonetheless, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.02, which supports the Neutral account.
+The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
+However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to .95.
+It is also possible that how to appropriately estimate the number of learnable words changes with development.
+The 90th percentile might be a good estimate of the number of potentially learnable words for older children, but the 75th percentile might be more realistic for younger children.
+It may be that the most cognitively advanced younger children are possibly more capable of lexicalizing more concepts than their peers; for example, better ability to update non-verbal predictions facilitates encoding new information and is found to correlate with larger vocabulary in infants of 12–24 months (e.g., Reuter et al., 2018).
+If the number of learnable words does change with development, the developmental change we observed could be related to how the number of potentially learnable words should be estimated, rather than a change in children’s bias to learn translation equivalents.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
 

### 2021-12-07 paper/paper_TE_bilingual_vocabulary_model.tex
Updated paper

@@
 For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
-These cross-language synonyms are known as translation equivalents\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word ``hygge'' refers to a type of cozy contentment and enjoyment of simplicity.
+These cross-language synonyms are known as translation equivalents\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
@@
 
-\begin{verbatim}
-## Warning: Removed 20 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 140 rows containing missing values (geom_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 20 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 140 rows containing missing values (geom_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 3 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 25 rows containing missing values (geom_smooth).
-\end{verbatim}
-
 \begin{figure}[H]
@@
 \end{itemize}
-
-\begin{verbatim}
-## Warning: Removed 4 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 1 rows containing missing values (geom_smooth).
-\end{verbatim}
 
@@
 In our data set, developmental level was approximated by children's age.
-Figure 3 Panel B shows the concept vocabulary (CONCEPT) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
-To illustrate the acquisition of translation equivalents and singlets at different developmental levels, we divided children into three age groups: younger children of 18--22 months, middle children of 23--27 months, and older children of 28--33 months.
+Figure 2 Panel B shows the number of translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, \emph{r}(227) = 0.69, \(p < .001\), and therefore confirmed the prediction.
-This pattern was further confirmed by a one-way ANOVA, where the three age groups significantly differed in the number of concept vocabulary they produced, (\(F\)(2, 226) = 90.86, \(p < .001\)).
-Older children of 28--33 months (i.e., at a later developmental level) produced the most with an average concept vocabulary of 414.60 (\(p\)s \textless{} .001), middle children of 23--27 months (i.e., at an intermediate developmental level) produced an average concept vocabulary of 252.10, and younger children of 18--22 months (i.e., at an earlier developmental level) produced the least with an average concept vocabulary of 119.90 (\(p\)s \textless{} .001).
 
 Prediction 2b was that older children would produce more translation equivalents than younger children.
-First, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, \emph{r}(227) = 0.48, \(p < .001\), and therefore confirmed the prediction.
-In a one-way ANOVA with age group as factor, we further found that groups differed in how many translation equivalents they produced (\(F\)(2, 226) = 31.74, \(p < .001\)).
-Younger children of 18--22 months produced an average of 33.70 translation equivalents, middle children of 23--27 months produced an average of 71.10 translation equivalents, and older children of 28--33 months produced an average of 131.40 translation equivalents (\(p\)s \textless{} .01).
+We observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, \emph{r}(227) = 0.48, \(p < .001\), and therefore confirmed the prediction.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.65, \(p < .001\), and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), \emph{r}(227) = -0.58, \(p < .001\).
-As shown in Figure 3 Panel B, children were divided into least balanced (range of balance: .00 - .20), medium balanced (range of balance: .20 - .35) and most balanced (range of balance: .35 - .50) groups (i.e., the same criteria as in Figure 5).
-In a one-way ANOVA with balance group as a between-subjects factor, we observed that the least balanced children produced the most singlets in their dominant language (\(p\)s \textless{} .001), with the least balanced, medium balanced, most balanced children producing respectively: 255.50, 141.50, and 71.90 words in their dominant language (\(F\)(2, 226) = 50.77, \(p < .001\)).
+As illustrated in Figure 2 Panel 2B, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = 294), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = 66).
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.19, \(p = .005\), and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), \emph{r}(227) = 0.63, \(p < .001\).
-In a one-way ANOVA with balance group as a between-subjects factor, we confirmed that children who differed in how balanced their vocabulary knowledge was also differed in how many singlets they produced in their non-dominant language (\(F\)(2, 226) = 61.89, \(p < .001\)).
-As shown in Figure 3 Panel B, we observed that children produced very few singlets in their non-dominant language, although the most balanced children produced the most singlets in their non-dominant language (mean of the most balanced children = 35.10 \textgreater{} mean of the medium balanced children = 15.50 \textgreater{} mean of the most balanced children = 5.70; \(p\)s \textless{} .001).
+As shown in Figure 2 Panel 2C, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = 39) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = 4).
 
@@
 Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
-To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE , and we allowed the model to estimate the BIAS parameter.
+To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE, and we allowed the model to estimate the BIAS parameter.
 
@@
 The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-The number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank; Frank et al., 2016).
+Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
 Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was 244.9 words which was calculated by averaging the 268.7 English words and 221.1 French words, based on what 18-month-old children would typically produce at the 90th percentile.
+For example, for an 18 month-old infant, the denominator was 239.70 words which was calculated by averaging the 259.40 English words and 220.10 French words, based on what 18-month-old children would typically produce at the 90th percentile.
 For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
@@
 The linear regression model estimated a BIAS coefficient of 1.04, \(p < .001\).
-This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words.
-
-To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child's expected translation equivalent knowledge (setting BIAS = 1), which is plotted against our observed data in Figure 6.
+This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
+Overall, model fit was slightly reduced compared to our main model, although fit was still good, \(R^2\) = 0.95, \(p < .001\).
+The estimated BIAS parameter was lower at 0.95.
+This change in the BIAS parameter is mathematically expected due to the reciprocal nature of LEARNABLE and BIAS, an issue we will return to in the discussion.}.
+
+To illustrate the close fit between the Neutral Account and our data, we used the Bilingual Vocabulary Model formula TE = 1 × (DOM×NONDOM/LEARNABLE) to estimate each child's expected translation equivalent knowledge (setting BIAS = 1; also setting the intercept at 0 to replicate our linear regression model), which is plotted against our observed data in Figure 5.
 Expected and observed translation equivalents were closely aligned with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1), suggesting that the Neutral Account provides a parsimonious explanation for bilinguals' translation equivalent knowledge.
@@
 Three sets of model predictions were confirmed in our empirical dataset.
+In Table 8, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:table8}Table 8. Summary of model predictions and evidence from previous literature and our empirical data.}
+\centering
+\fontsize{7}{9}\selectfont
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{200px}>{\raggedright\arraybackslash}p{250px}>{\centering\arraybackslash}p{150px}}
+\toprule
+Model predictions & Evidence in previous literature & Did we observe the same pattern in our empirical data?\\
+\midrule
+1A: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE).
+& Children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+& Yes\\
+1B: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE).
+& The number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
+ There is a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013) & Yes\\
+1C: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge & Vocabulary size in the non-dominant language was positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
+& Yes\\
+2A: Older children will have larger concept vocabularies (CONCEPT) than younger children. & Older children had larger concept vocabularies (Pearson et al., 1993). & Yes\\
+2B: Older children will produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+& Older bilingual children knew more translation equivalents than younger children  (David \& Wei, 2008; Legacy et al., 2016) & Yes\\
+\addlinespace
+2C: Older children will produce more dominant-language singlet words (DOM-SINGLET).
+Moreover, those with the least balanced vocabulary (BALANCE) will produce the most DOM-SINGLET.
+& Bilingual children learned words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011) & Yes\\
+2D: Older children will produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most balanced vocabulary (BALANCE) will produce the most NONDOM-SINGLET. &  & \\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
 
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
-For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the livre?'' if they know their child understands the French word ``livre'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the \emph{livre}?'' if they know their child understands the French word \emph{``livre''} but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict ,1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
-It would be interesting for future studies to take into consideration the such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
+However, our model does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
 Such a model may better characterize and predict individual differences in bilingual vocabulary development.
 Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data and/or data of a different bilingual.
+
+Note also that the age group that our paper currently looked at has been suggested to learn substantially more nouns before verbs or other word classes (i.e., noun bias; Gentner, 1982).
+It has been suggested that nouns that are frequently found in the input or have concrete meanings appeared to be the easiest for young children to grasp (Braginsky et al., 2019; Goodman et al., 2008).
+On the other hand, predicates including verbs, adjectives and adverbs appear to be more difficult to learn because their meaning is more abstract and is dependent on relationships among entities (Gertner, 2006).
+Due to these differences, different lexical categories seem to be learned by young children in different ways (Braginsky et al., 2019; Imai et al., 2005).
+Since meanings of nouns are more concrete, the boundaries to categorize objects are more stable.
+Yet, predicates have boundaries that are less clearly marked without knowledge of neighboring words (Imai et al., 2008; Saji et al., 2020).
+Future investigation may explore if bilingual vocabulary learning would differ among nouns and other lexical categories.
 
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a ``bowl'' in English but an \emph{``assiette''} {[}plate{]} in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, and little to no data from bilingual children.
-Second, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English--French pair ``banana'' -- \emph{``banane''}) more easily than those that do not share similar phonological form (e.g., the English--French pair ``dog'' -- \emph{``chien''}) (Bosch \& Ramon-Casas, 2014).
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storm et al., 2015; White et al., 2020).
+Therefore, it seems reasonable for our model to take the simplified view that early vocabulary may be treated similarly across languages, especially for the age group we were looking at.
+Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where vocabularies are kept language-specific and distinct across languages.
+
+Another limitation of our approach is that we modeled the number of words that might be learnable at a particular age, rather than modeling the learnability of individual words.
+For example, our model did not take into account that bilingual children appear to learn similar-sounding translation equivalents (i.e., cognates like the English--French pair ``banana'' -- \emph{``banane''}) more easily than those that do not share similar phonological form (e.g., the English--French pair ``dog'' -- \emph{``chien''}) (Bosch \& Ramon-Casas, 2014; Mitchell et al., 2021).
 Likewise, some bilingual children learn language pairs that share more cognates than others (e.g., Spanish and Italian share more phonologically similar translation equivalents than English and French; Schepens et al., 2013).
-While more research will be needed on how these factors impact bilingual vocabulary learning, the close correspondence between our model and data from bilingual children suggest that even if our assumptions are a simplification, deviations from these assumptions might have a relatively small impact.
-Moreover, if they do prove to be important, such factors could be added to future iterations of the Bilingual Vocabulary Model.
-
-Another assumption of our model was that bilingual children hear labels from both languages for the same set of referents.
+There are many other individual differences between words that affect their learnability, such as frequency of occurrence in speech input, concreteness, babiness, word length, semantic category, etc.
+(Braginsky et al., 2019).
+It is likely that within the set of words that are learnable at a given age, some will be more easily learned than others, which our model does not take into account.
+However, the close correspondence between our model and data from bilingual children suggest that even if our model is a simplification, deviations from our assumptions might have a relatively small impact.
+On the other hand, individual item differences could be particularly consequential at younger ages, and the apparent shift we observed in development from the enhancement account to the avoidance account could be an artifact of this issue.
+Future iterations of the Bilingual Vocabulary Model could explicitly add such word-level factors into the model and test the importance of those factors in early bilingual vocabulary development.
+
+Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
 However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
-For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
+For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
 At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
 For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to a similar set of referents and labels in each of their languages.
-Moreover, if indeed bilingual children tend to encounter different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
 Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
@@
 Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
-Under the Bilingual Vocabulary Model, these two parameters jointly and inversely predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter increases by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
-Our model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce.
-Small changes to this approach (e.g., taking the number of words children at the 95th percentile produce) would likely not drastically alter our results, nor change the qualitative shift that we observed in our data.
-Nonetheless, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.02, which supports the Neutral account.
+The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
+However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to .95.
+It is also possible that how to appropriately estimate the number of learnable words changes with development.
+The 90th percentile might be a good estimate of the number of potentially learnable words for older children, but the 75th percentile might be more realistic for younger children.
+It may be that the most cognitively advanced younger children are possibly more capable of lexicalizing more concepts than their peers; for example, better ability to update non-verbal predictions facilitates encoding new information and is found to correlate with larger vocabulary in infants of 12--24 months (e.g., Reuter et al., 2018).
+If the number of learnable words does change with development, the developmental change we observed could be related to how the number of potentially learnable words should be estimated, rather than a change in children's bias to learn translation equivalents.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
 

### 2021-12-10 01_data_wordbank.Rmd
updated paper

@@
   # rename age to prepare for the next step (i.e., merging with keepers_ws_TE
-  rename(age_months_percentile = age)
+  rename(age_months_percentile = age) %>%
+  # round to the nearest integer
+  mutate_if(is.numeric, round, digits = 0)
 

### 2021-12-10 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated paper

@@
 Moreover, even when a word that could be considered a translation equivalent exists, adults’ concepts may not be congruent across languages (e.g., Malt & Majid, 2013).
-However, in early development, most vocabulary words are reasonably concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
 Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.
@@
 We present three simulations generated by creating data points for hypothetical children by varying different parameters of the Bilingual Vocabulary Model, in order to explore expected patterns of translation equivalent learning.
-The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except for those that violated constraints in the model (e.g., NON-DOM could not be larger than DOM).
-In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NON-DOM), and word vocabulary (WORD).
-In the second simulation, we explored relationships between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE).
+The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except for those that violated constraints in the model (e.g., NONDOM could not be larger than DOM).
+In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), and word vocabulary (WORD).
+In the second simulation, we explored relations between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE).
 In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the bias parameter (BIAS) to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts.
@@
   # reorder columns
-  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)")
+  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)") %>%
+  # round to the nearest integer
+  mutate_if(is.numeric, round, digits = 0)
 
@@
       caption = "Table 3.
-Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.") %>%
+Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Vocabulary values are rounded to the nearest integer.") %>%
   pack_rows("Main Parameters", 1, 4) %>%
@@
 ## 1.3 Simulation 3: Bias towards or against translation equivalent learning compared to singlets
-In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
+In Simulations 1 and 2, we modeled cases in accordance with the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
 In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
@@
 For the Preference Account, BIAS will be greater than 1, meaning that TEs are more easily learned than singlets.
-On the other hand, for the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets.
+By contrast, for the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets.
 
@@
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
-Prediction Set 1 illustrated that the more words children produced in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they would produce.
-Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) would also produce more translation equivalents.
-Prediction Set 2 described translation equivalent knowledge with relations to the number of learnable words constrained by a child’s developmental level.
-Older children were predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
-Moreover, the most balanced children produced more singlets in their non-dominant language, whereas the least balanced children produced more singlets in their dominant language.
+Prediction Set 1 illustrated that the more words children produce in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they are predicted to produce.
+Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) are also expected to produce more translation equivalents.
+Prediction Set 2 described translation equivalent knowledge in relation to the number of learnable words constrained by a child’s developmental level.
+Older children are predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Moreover, the most balanced children are expected to produce more singlets in their non-dominant language, whereas the least balanced children are expected to produce more singlets in their dominant language.
 Finally, Prediction Set 3 quantitatively demonstrated the patterns of translation equivalent learning under the three different accounts: the Avoidance Account, the Preference Account, and the Neutral Account.
@@
 Following the approach in Study 1, children’s dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
-Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 2)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 2)`% were French-dominant.
+Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 1)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 1)`% were French-dominant.
 Children’s demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
@@
 Caregivers completed the original CDI English version [@Fenson_etal_2007] and its Québec French adaptation (Trudeau et al., 1999).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), and respondent not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
+We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), or else the respondent was not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
 Our analyses focused on the vocabulary checklist of this questionnaire, which includes different nouns, verbs, adjectives, and other words used by young children.
-There are 680 words in the English CDI version and 664 in the Québec French version.
+There are 680 words on the English CDI version and 664 in the Québec French version.
 
@@
 Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
-The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 2)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
+The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
 
@@
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
-For most children, the language in which they produced the most words was also the language that they heard most often (`r dominance_consistency$n_consistent` children, `r round(dominance_consistency$percentage_consistent,2)`%), although this was not the case for some children (`r dominance_consistency$n_inconsistent` children, `r round(dominance_consistency$percentage_inconsistent,2)`%).
+For most children, the language in which they produced the most words was also the language that they heard most often (`r dominance_consistency$n_consistent` children, `r round(dominance_consistency$percentage_consistent,1)`%), although this was not the case for a minority of children (`r dominance_consistency$n_inconsistent` children, `r round(dominance_consistency$percentage_inconsistent,1)`%).
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error alone (for a discussion, see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error alone (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 
-[^3]: Note that there was a negative correlation between balance and age in our sample.
+[^3]: Note that there was a negative correlation between BALANCE and age in our sample.
 We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.
@@
 
-# one-way ANOVA to compare TE among the 5 BALANCE groups
-
-keepers_ws_TE <- keepers_ws_TE %>% # divide children into 5 balance subset groups (0 < BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5)
-  mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
-                     labels = c("0.1", "0.2", "0.3", "0.4", "0.5")))
-
-aov_TE_binBALANCE <- tidy(keepers_ws_TE %>% # one-way ANOVA
-                            aov(number_of_te ~ bin_BALANCE, data = .))
-
 # correlation between TE and WORD
@@
 As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, _r_(`r cor_TE_balance$parameter[1]`) = `r round(cor_TE_balance$estimate[1],2)`, $p `r papaja::printp(cor_TE_balance$p.value[1])`$, where children with the most balanced vocabulary produced the most translation equivalents.
-We further tested this prediction by dividing children into 5 balance subset groups (0 < BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, _F_(`r aov_TE_binBALANCE$df[1]`, `r aov_TE_binBALANCE$df[2]`) = `r aov_TE_binBALANCE$statistic[1]`, $p = `r papaja::printp(aov_TE_binBALANCE$p.value[1])`$.
-The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the fewest translation equivalents.
-Detailed descriptive statistics are reported in Table 6.
 
@@
 
-```{r table6, results="asis", cache = TRUE}
-
-# TE descriptive statistics of the 5 balance group
-mean_TE_binBALANCE <- keepers_ws_TE %>% 
-  group_by(bin_BALANCE) %>%
-  summarise(n = n(),
-            mean_TE = mean(number_of_te, na.rm = T),
-            sd_TE = sd(number_of_te, na.rm = T),
-            min_TE = min(number_of_te, na.rm = T),
-            max_TE = max(number_of_te, na.rm = T)) %>%
-  mutate(across(!bin_BALANCE & !sd_TE, ~ round(.,0))) %>%
-  mutate_at(vars(sd_TE), ~ round(.,2)) %>%
-  # combine min and max to form range
-  unite("range", min_TE:max_TE, sep = " - ", remove = TRUE, na.rm = FALSE) %>%
-  # sort bin_BALANCE with descending order
-  arrange(desc(bin_BALANCE)) %>%
-  # rename each level of bin_BALANCE
-  mutate(bin_BALANCE = recode(bin_BALANCE,
-                              "0.5" = "0.4 < BALANCE \u2264 0.5",
-                              "0.4" = "0.3 < BALANCE \u2264 0.4",
-                              "0.3" = "0.2 < BALANCE \u2264 0.3",
-                              "0.2" = "0.1 < BALANCE \u2264 0.2",
-                              "0.1" = "0 < BALANCE \u2264 0.1")) %>%
-  # rename column
-  rename(" " = bin_BALANCE, "N" = n, "Mean" = mean_TE, "SD" = sd_TE)
-
-# print table
-#papaja::apa_table(mean_TE_binBALANCE,
-#                  caption = "Average number of translation equivalents (TE) produced by each balance group.",
-#                  col.names =c("","N","Mean","$SD$","range"),
-#                  align="c")
-
-kable(mean_TE_binBALANCE,
-      format="latex", booktabs=TRUE, align = "c",
-      caption = "Table 6. Average number of translation equivalents (TE) produced by each balance group.") %>%
-  column_spec(1, width = "150px") %>%
-  column_spec(2:4, width = "50px") %>%
-  column_spec(5, width = "75px") %>%
-  kable_styling(position = "center")
-```
-
 ## 2.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels
@@
          "Number of French words produced at 90th percentile" = FrWS_90percentile,
-         "Average (LEARNABLE)" = Average_90percentile) %>%
-  # round to 1 decimal place
-  mutate(across(where(is.numeric),round,1))
+         "Average (LEARNABLE)" = Average_90percentile) 
 ```
@@
 Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
-Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was `r round(ws_percentile$"Average (LEARNABLE)"[1],1)` words which was calculated by averaging the `r round(ws_percentile$"Number of English words produced at 90th percentile"[1],1)` English words and `r round(ws_percentile$"Number of French words produced at 90th percentile"[1],1)` French words, based on what 18-month-old children would typically produce at the 90th percentile.
+Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
+For example, for an 18 month-old infant, the denominator was `r ws_percentile$"Average (LEARNABLE)"[1]` words which was calculated by averaging the `r ws_percentile$"Number of English words produced at 90th percentile"[1]` English words and `r ws_percentile$"Number of French words produced at 90th percentile"[1]` French words, based on what 18-month-old children would typically produce at the 90th percentile.
 For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
 
-```{r table7, results = "asis", cache = TRUE}
+```{r table6, results = "asis", cache = TRUE}
 kable(ws_percentile,
+        digits = 0,
         format = "latex", booktabs = TRUE, align = "c",
-        caption = "Table 7.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.") %>%
+        caption = "Table 6.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.
+Values are rounded to the nearest integer.") %>%
   kableExtra::kable_styling(font_size = 10) %>%
@@
 
-\noindent (In R language, the model was entered as:
+\noindent (In R language, the model formula was rearranged and entered as:
 
@@
 The linear regression model estimated a BIAS coefficient of `r round(summary(test_prediction3_model)$coefficient[1],2)`, $p `r papaja::printp(summary(test_prediction3_model)$coefficient[4])`$.
-This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^4].
+This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^4].
 
@@
 Overall, model fit was slightly reduced compared to our main model, although fit was still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`, $p `r papaja::printp(summary(test_prediction3_model_75percentile)$coefficient[4])`$.
-The estimated BIAS parameter was lower at `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`.
+The estimated BIAS parameter was lower, at `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
 This change in the BIAS parameter is mathematically expected due to the reciprocal nature of LEARNABLE and BIAS, an issue we will return to in the discussion.
@@
 Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
-Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1) for children with more than 300 total vocabulary.
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.
 To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
@@
 Three sets of model predictions were confirmed in our empirical dataset.
-In Table 8, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
-
-```{r table8, results = "asis", cache = TRUE}
-
-table8 <- data.frame(
+In Table 7, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+
+```{r table7, results = "asis", cache = TRUE}
+
+table7 <- data.frame(
   model_predictions = c(
@@
 landscape(
-  kable(table8,
+  kable(table7,
         align = "llc",
         format = "latex", booktabs = TRUE,
-        caption = "Table 8. Summary of model predictions and evidence from previous literature and our empirical data.") %>%
+        caption = "Table 7. Summary of model predictions, similar evidence from previous literature, and whether our empirical data was consistent with the prediction.") %>%
     column_spec(c(1), width = "200px") %>%
@@
 
-A first set of predictions pertained to relationships between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
+A first set of predictions pertained to relations between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
 In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
@@
 In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
-In contrast, children with larger vocabulary sizes appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
+In contrast, children with larger vocabularies appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages [@Au_Glusman_1990; @Davidson_Tell_2005].
 At first blush, strengthening of one-to-one mapping biases with age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
-For example, a caregiver may choose to say to their English–French bilingual child “Can you grab the _livre_?” if they know their child understands the French word _“livre”_ but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English–French bilingual child “Can you give me the _livre_?” if they know their child understands the French word _“livre”_ but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
@@
 
+Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2013), and amount of language experience (e.g., Weisleder & Fernald, 2013).
+Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
+However, there could be much more variability in the developmental level of younger infants than of older infants.
+In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+
 ## Assumptions, limitations, and future directions
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
-However, our model does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our approach does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 Therefore, it seems reasonable for our model to take the simplified view that early vocabulary may be treated similarly across languages, especially for the age group we were looking at.
-Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where vocabularies are kept language-specific and distinct across languages.
+Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where concepts are less well aligned across languages.
 
@@
 
-Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
+We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
 Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
-Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.02, which supports the Neutral account.
+That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only `r ws_percentile$"Average (LEARNABLE)"[1]/2` words in each language are learnable for 18-month-olds, rather than `r ws_percentile$"Average (LEARNABLE)"[1]`), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of `r round(summary(percentile_model_less300w)$coefficient[1],2)` which supports the Preference account, we would estimate a parameter of `r round(summary(percentile_model_less300w)$coefficient[1]/2, 2)` which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around `r round(summary(test_prediction3_model)$coefficient[1],2)`, which supports the Neutral account.
 The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
-However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to .95.
-It is also possible that how to appropriately estimate the number of learnable words changes with development.
-The 90th percentile might be a good estimate of the number of potentially learnable words for older children, but the 75th percentile might be more realistic for younger children.
-It may be that the most cognitively advanced younger children are possibly more capable of lexicalizing more concepts than their peers; for example, better ability to update non-verbal predictions facilitates encoding new information and is found to correlate with larger vocabulary in infants of 12–24 months (e.g., Reuter et al., 2018).
-If the number of learnable words does change with development, the developmental change we observed could be related to how the number of potentially learnable words should be estimated, rather than a change in children’s bias to learn translation equivalents.
-Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is lower, although nevertheless still consistent with the Neutral Account.
+Future work might find an improved method for estimating the number of words that are potentially learnable for individual children.
 

### 2021-12-10 paper/paper_TE_bilingual_vocabulary_model.tex
updated paper

@@
 Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
-However, in early development, most vocabulary words are reasonably concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
 Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.} (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
@@
 \caption{\label{tab:table3}Table 3.
-Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.}
+Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Vocabulary values are rounded to the nearest integer.}
 \centering
@@
 \multicolumn{6}{l}{\textbf{Main Parameters}}\\
-\hspace{1em}BIAS & Bias parameter &  & 1.0 & 1.0 & 1.0\\
-\hspace{1em}LEARNABLE & Learnable words in each language &  & 600.0 & 600.0 & 600.0\\
-\hspace{1em}DOM & Words produced in the dominant language &  & 270.0 & 180.0 & 540.0\\
-\hspace{1em}NONDOM & Words produced in the non-dominant language &  & 30.0 & 120.0 & 360.0\\
+\hspace{1em}BIAS & Bias parameter &  & 1 & 1 & 1\\
+\hspace{1em}LEARNABLE & Learnable words in each language &  & 600 & 600 & 600\\
+\hspace{1em}DOM & Words produced in the dominant language &  & 270 & 180 & 540\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language &  & 30 & 120 & 360\\
 \addlinespace[0.3em]
 \multicolumn{6}{l}{\textbf{Derived Parameters}}\\
-\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & DOM + NONDOM & 300.0 & 300.0 & 900.0\\
-\hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0.1 & 0.4 & 0.4\\
-\hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 13.5 & 36.0 & 324.0\\
-\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286.5 & 264.0 & 576.0\\
-\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256.5 & 144.0 & 216.0\\
-\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16.5 & 84.0 & 36.0\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & DOM + NONDOM & 300 & 300 & 900\\
+\hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0 & 0 & 0\\
+\hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 14 & 36 & 324\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286 & 264 & 576\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256 & 144 & 216\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16 & 84 & 36\\
 \bottomrule
@@
 
-In Simulations 1 and 2, we modeled cases in accordance to the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
+In Simulations 1 and 2, we modeled cases in accordance with the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
 In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
@@
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
-Prediction Set 1 illustrated that the more words children produced in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they would produce.
-Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) would also produce more translation equivalents.
-Prediction Set 2 described translation equivalent knowledge with relations to the number of learnable words constrained by a child's developmental level.
-Older children were predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
-Moreover, the most balanced children produced more singlets in their non-dominant language, whereas the least balanced children produced more singlets in their dominant language.
+Prediction Set 1 illustrated that the more words children produce in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they are predicted to produce.
+Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) are also expected to produce more translation equivalents.
+Prediction Set 2 described translation equivalent knowledge in relation to the number of learnable words constrained by a child's developmental level.
+Older children are predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Moreover, the most balanced children are expected to produce more singlets in their non-dominant language, whereas the least balanced children are expected to produce more singlets in their dominant language.
 Finally, Prediction Set 3 quantitatively demonstrated the patterns of translation equivalent learning under the three different accounts: the Avoidance Account, the Preference Account, and the Neutral Account.
@@
 Following the approach in Study 1, children's dominant language was deemed to be the language in which the child produced a greater number of words; vocabulary balance was then determined based on the proportion of words produced in the non-dominant language relative to the total words produced across both languages using the same formula as in Study 1: NONDOM/(DOM+NONDOM).
-Within the 229 data points, 59.83\% of children were English-dominant and 40.17\% were French-dominant.
+Within the 229 data points, 59.80\% of children were English-dominant and 40.20\% were French-dominant.
 Children's demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
@@
 Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1999).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), and respondent not indicated (24\%).
+We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), or else the respondent was not indicated (24\%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
 Our analyses focused on the vocabulary checklist of this questionnaire, which includes different nouns, verbs, adjectives, and other words used by young children.
-There are 680 words in the English CDI version and 664 in the Québec French version.
+There are 680 words on the English CDI version and 664 in the Québec French version.
 
@@
 Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
-The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.17\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).
+The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.20\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).
 
@@
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
-For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79.04\%), although this was not the case for some children (48 children, 20.96\%).
+For most children, the language in which they produced the most words was also the language that they heard most often (181 children, 79\%), although this was not the case for a minority of children (48 children, 21\%).
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error alone (for a discussion, see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error alone (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson's correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between balance and age in our sample.
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between BALANCE and age in our sample.
 We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.}).
@@
 As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, \emph{r}(227) = 0.25, \(p < .001\), where children with the most balanced vocabulary produced the most translation equivalents.
-We further tested this prediction by dividing children into 5 balance subset groups (0 \textless{} BALANCE ≤ 0.1, 0.2, 0.3, 0.4, and 0.5), and a one-way ANOVA revealed a significant effect of BALANCE, \emph{F}(4, 224) = 3.61, \(p = .007\).
-The children with a BALANCE score of 0.5 (i.e., with more balanced vocabulary) produced the most translation equivalents, whereas children with a BALANCE score of 0.1 (i.e., with less balanced vocabulary) produced the fewest translation equivalents.
-Detailed descriptive statistics are reported in Table 6.
 
@@
 
-\begin{table}
-
-\caption{\label{tab:table6}Table 6. Average number of translation equivalents (TE) produced by each balance group.}
-\centering
-\begin{tabular}[t]{>{\centering\arraybackslash}p{150px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{50px}>{\centering\arraybackslash}p{75px}}
-\toprule
-  & N & Mean & SD & range\\
-\midrule
-0.4 < BALANCE ≤ 0.5 & 69 & 92 & 114.44 & 4 - 409\\
-0.3 < BALANCE ≤ 0.4 & 58 & 68 & 89.30 & 1 - 376\\
-0.2 < BALANCE ≤ 0.3 & 52 & 64 & 57.13 & 5 - 221\\
-0.1 < BALANCE ≤ 0.2 & 32 & 47 & 35.17 & 7 - 137\\
-0 < BALANCE ≤ 0.1 & 18 & 18 & 11.57 & 2 - 52\\
-\bottomrule
-\end{tabular}
-\end{table}
-
 \hypertarget{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}{%
@@
 Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
-Table 7 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was 239.70 words which was calculated by averaging the 259.40 English words and 220.10 French words, based on what 18-month-old children would typically produce at the 90th percentile.
+Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
+For example, for an 18 month-old infant, the denominator was 240 words which was calculated by averaging the 259 English words and 220 French words, based on what 18-month-old children would typically produce at the 90th percentile.
 For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
@@
 
-\caption{\label{tab:table7}Table 7.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.}
+\caption{\label{tab:table6}Table 6.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.
+Values are rounded to the nearest integer.}
 \centering
@@
 \midrule
-18 & 259.4 & 220.1 & 239.7\\
-19 & 321.5 & 274.4 & 297.9\\
-20 & 378.3 & 325.1 & 351.7\\
-21 & 429.8 & 372.2 & 401.0\\
-22 & 476.1 & 415.7 & 445.9\\
+18 & 259 & 220 & 240\\
+19 & 321 & 274 & 298\\
+20 & 378 & 325 & 352\\
+21 & 430 & 372 & 401\\
+22 & 476 & 416 & 446\\
 \addlinespace
-23 & 517.0 & 455.6 & 486.3\\
-24 & 552.6 & 492.0 & 522.3\\
-25 & 583.3 & 524.8 & 554.0\\
-26 & 609.9 & 554.0 & 582.0\\
-27 & 633.0 & 579.6 & 606.3\\
+23 & 517 & 456 & 486\\
+24 & 553 & 492 & 522\\
+25 & 583 & 525 & 554\\
+26 & 610 & 554 & 582\\
+27 & 633 & 580 & 606\\
 \addlinespace
-28 & 652.5 & 601.7 & 627.1\\
-29 & 668.5 & 620.2 & 644.3\\
-30 - 33 & 680.9 & 635.1 & 658.0\\
+28 & 653 & 602 & 627\\
+29 & 668 & 620 & 644\\
+30 - 33 & 681 & 635 & 658\\
 \bottomrule
@@
 The linear regression model estimated a BIAS coefficient of 1.04, \(p < .001\).
-This value is extremely close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
+This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
 Overall, model fit was slightly reduced compared to our main model, although fit was still good, \(R^2\) = 0.95, \(p < .001\).
-The estimated BIAS parameter was lower at 0.95.
+The estimated BIAS parameter was lower, at 0.95, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
 This change in the BIAS parameter is mathematically expected due to the reciprocal nature of LEARNABLE and BIAS, an issue we will return to in the discussion.}.
@@
 Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
-Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account of the Bilingual Vocabulary Model (i.e., BIAS = 1) for children with more than 300 total vocabulary.
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.
 To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
@@
 Three sets of model predictions were confirmed in our empirical dataset.
-In Table 8, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+In Table 7, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
 
@@
 
-\caption{\label{tab:table8}Table 8. Summary of model predictions and evidence from previous literature and our empirical data.}
+\caption{\label{tab:table7}Table 7. Summary of model predictions, similar evidence from previous literature, and whether our empirical data was consistent with the prediction.}
 \centering
@@
 In other words, children with less experience in word learning may be more inclined to accept multiple words for the same referent (Halberda, 2003; Merriman et al., 1989).
-In contrast, children with larger vocabulary sizes appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
+In contrast, children with larger vocabularies appear to become more certain about one-to-one mapping relationships between referents and words (Lewis et al., 2020), while simultaneously using their bilingual experience to understand that referents can have different words between languages (Au \& Glusman, 1990; Davidson \& Tell, 2005).
 At first blush, strengthening of one-to-one mapping biases with age could explain why younger children appear to learn relatively more translation equivalents than older children.
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
-For example, a caregiver may choose to say to their English--French bilingual child ``Can you grab the \emph{livre}?'' if they know their child understands the French word \emph{``livre''} but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English--French bilingual child ``Can you give me the \emph{livre}?'' if they know their child understands the French word \emph{``livre''} but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
@@
 
+Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2013), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
+Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
+However, there could be much more variability in the developmental level of younger infants than of older infants.
+In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+
 \hypertarget{assumptions-limitations-and-future-directions}{%
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
-However, our model does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our approach does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 
-Finally, we must note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
+We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
 Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 122 words in each language are learnable for 18-month-olds, rather than 244), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.22 which supports the Preference account, we would estimate a parameter of 1.11 which is closer to the Neutral Account).
-Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.02, which supports the Neutral account.
+That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 120 words in each language are learnable for 18-month-olds, rather than 240), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.21 which supports the Preference account, we would estimate a parameter of 1.10 which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.04, which supports the Neutral account.
 The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
-However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to .95.
-It is also possible that how to appropriately estimate the number of learnable words changes with development.
-The 90th percentile might be a good estimate of the number of potentially learnable words for older children, but the 75th percentile might be more realistic for younger children.
-It may be that the most cognitively advanced younger children are possibly more capable of lexicalizing more concepts than their peers; for example, better ability to update non-verbal predictions facilitates encoding new information and is found to correlate with larger vocabulary in infants of 12--24 months (e.g., Reuter et al., 2018).
-If the number of learnable words does change with development, the developmental change we observed could be related to how the number of potentially learnable words should be estimated, rather than a change in children's bias to learn translation equivalents.
-Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to 0.95, which is lower, although nevertheless still consistent with the Neutral Account.
+Future work might find an improved method for estimating the number of words that are potentially learnable for individual children.
 

### 2021-12-13 paper/paper_TE_bilingual_vocabulary_model.Rmd
Updated paper

@@
 
-* Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE).
-* Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE). 
-* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 1).
+* Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE) (Figure 1, Panel 1A).
+* Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE) (Figure 1, Panels 1A and 1B). 
+* Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE).
+Unlike for WORD and DOM this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
 
@@
   pivot_longer(c(WORD, DOM, NONDOM), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("DOM", "NONDOM", "WORD"))) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("WORD", "DOM", "NONDOM"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             DOM = "Panel 1A: \n Dominant vocabulary \n (DOM)", 
-                             NONDOM = "Panel 1B: \n Non-dominant vocabulary \n (NONDOM)",
-                             WORD = "Panel 1C: \n Total vocabulary \n (WORD)")) %>%
+                             WORD = "Panel 1A: \n Total vocabulary \n (WORD)",
+                             DOM = "Panel 1B: \n Dominant vocabulary \n (DOM)", 
+                             NONDOM = "Panel 1C: \n Non-dominant vocabulary \n (NONDOM)")) %>%
   ggplot(aes(x = number, y = TE, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
   pivot_longer(c(word_vocab, total_words_dom, total_words_nondom), names_to = "vocab_type", values_to = "number") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("total_words_dom", "total_words_nondom", "word_vocab"))) %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("word_vocab", "total_words_dom", "total_words_nondom"))) %>%
   mutate(vocab_type = recode(vocab_type, 
-                             total_words_dom = "Panel 2A: \n Dominant vocabulary \n (DOM)", 
-                             total_words_nondom = "Panel 2B:\n Non-dominant vocabulary \n (NONDOM)",
-                             word_vocab = "Panel 2C: \n Total vocabulary \n (WORD)")) %>%
+                             word_vocab = "Panel 2A: \n Total vocabulary \n (WORD)",
+                             total_words_dom = "Panel 2B: \n Dominant vocabulary \n (DOM)", 
+                             total_words_nondom = "Panel 2C:\n Non-dominant vocabulary \n (NONDOM)")) %>%
   mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
@@
 
-```{r fig1, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
+```{r fig1, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to word vocabulary (WORD; Panel A), dominant vocabulary size (DOM; Panel B), and non-dominant vocabulary size (NONDOM; Panel C).
 Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
@@
 We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
-Figure 2 plots the number of TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
-Across the three Panels (1A, 1B, and 1C),  simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
-In Panel 1A, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
+As shown in Panel 1A, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+In Panel 1B, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
 Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Panels 1B and 1C show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
+Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
 Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
@@
 
-* Prediction 2a: Have larger concept vocabularies (CONCEPT). 
-* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
-* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET.
-* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+* Prediction 2a: Have larger concept vocabularies (CONCEPT). (Figure 2, Panel 1A)
+* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE). (Figure 2, Panel 1B)
+* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET. (Figure 2, Panel 1C)
+* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET. (Figure 2, Panel 1D)
 
@@
 simulation2 <- data_simulation2 %>%
-  pivot_longer(-c(LEARNABLE, WORD, CONCEPT, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
+  pivot_longer(-c(LEARNABLE, WORD, BALANCE, bin_BALANCE), names_to = "vocab_type", values_to = "number") %>%
   filter(vocab_type %notin% c("DOM", "NONDOM")) %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("TE", "DOM_SINGLET", "NONDOM_SINGLET")),
+  mutate(vocab_type = factor(vocab_type, levels=c("CONCEPT", "TE", "DOM_SINGLET", "NONDOM_SINGLET")),
          vocab_type = recode(vocab_type, 
-                             TE = "Panel 1A: \n Translation equivalents \n (TE)",
-                             DOM_SINGLET = "Panel 1B: \n Singlets in dominant language \n (DOM_SINGLET)",
-                             NONDOM_SINGLET = "Panel 1C: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             CONCEPT = "Panel 1A: \n Concept vocabulary \n (CONCEPT)",
+                             TE = "Panel 1B: \n Translation equivalents \n (TE)",
+                             DOM_SINGLET = "Panel 1C: \n Singlets in dominant language \n (DOM_SINGLET)",
+                             NONDOM_SINGLET = "Panel 1D: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
   ggplot(aes(x = LEARNABLE, y = number, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
   theme(legend.position = "bottom",
-        plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
+        plot.title = element_text(face = 'bold', size = 10, hjust = 0, vjust = 2),
         plot.title.position = "plot",
-        text = element_text(size=12),
+        text = element_text(size=10),
         panel.spacing.x = unit(2, "lines"),
@@
 Observed_plot_LEARNABLE <- keepers_ws_TE %>%
-  select(baby_id, age_months_percentile, Average_90percentile, balance_vocab, word_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
+  select(baby_id, age_months_percentile, Average_90percentile, balance_vocab, 
+         word_vocab, concept_vocab, number_of_te, total_singlet_dom, total_singlet_nondom) %>%
   mutate(bin_BALANCE = cut(balance_vocab, breaks = c(0, 0.1, 0.2, 0.3, 0.4, 0.5), include.lowest=T,
                      labels = c("0.1", "0.2", "0.3", "0.4", "0.5"))) %>%
-  mutate(TE = number_of_te) %>%
-  pivot_longer(c(total_singlet_nondom, total_singlet_dom, number_of_te), names_to = "vocab_type", values_to = "number_of_words") %>%
-  mutate(vocab_type = factor(vocab_type, levels=c("number_of_te", "total_singlet_dom", "total_singlet_nondom")),
+  pivot_longer(c(total_singlet_nondom, total_singlet_dom, number_of_te, concept_vocab), names_to = "vocab_type", values_to = "number_of_words") %>%
+  mutate(vocab_type = factor(vocab_type, levels=c("concept_vocab", "number_of_te", "total_singlet_dom", "total_singlet_nondom")),
          vocab_type = recode(vocab_type, 
-                             number_of_te = "Panel 2A: \n Translation equivalents \n (TE)",
-                             total_singlet_dom = "Panel 2B: \n Singlets in dominant language \n (DOM_SINGLET)",
-                             total_singlet_nondom = "Panel 2C: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             concept_vocab = "Panel 2A: \n Concept vocabulary \n (CONCEPT)",
+                             number_of_te = "Panel 2B: \n Translation equivalents \n (TE)",
+                             total_singlet_dom = "Panel 2C: \n Singlets in dominant language \n (DOM_SINGLET)",
+                             total_singlet_nondom = "Panel 2D: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
   ggplot(aes(x = age_months_percentile, y = number_of_words, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
   ylim(0, 450) +
-  #xlim(200, 700) +
   scale_x_continuous(breaks = seq(18, 30, by = 2)) +
@@
 
-```{r fig2, fig.cap="Number of translation equivalents (TE; Panel A), singlets in dominant (DOM-SINGLET; Panel B) and singlets in the non-dominant language (NONDOM-SINGLET; Panel C) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+```{r fig2, fig.cap="Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
 Row 1 represents the simulated data in Study 1.
-Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.width="80%", fig.pos = "H", cache = TRUE}
+Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.width="100%", fig.pos = "H", cache = TRUE}
 
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): _r_(`r cor_TE_word$parameter[1]`) = `r round(cor_TE_word$estimate[1],2)`, $p `r papaja::printp(cor_TE_word$p.value[1])`$, and dominant-language vocabulary (DOM): _r_(`r cor_TE_dom$parameter[1]`) = `r round(cor_TE_dom$estimate[1],2)`, $p `r papaja::printp(cor_TE_dom$p.value[1])`$.
-Figure 1 Row 2 further illustrates these relationships observed in our dataset.
+Figure 1 Panels 2A and 2B further illustrates these relationships observed in our dataset.
 
 Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
-As shown in Figure 1 Row 2 (Observed data), we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
+As shown in Figure 1 Panel 2C, we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
 
@@
 In our data set, developmental level was approximated by children’s age.
-Figure 2 Panel B shows the number of translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 Prediction 2a was that older children (i.e., those at a later developmental level) would have larger concept vocabularies than younger children (i.e., those at an earlier developmental level).
-We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, _r_(`r cor_age_CONCEPT$parameter[1]`) = `r round(cor_age_CONCEPT$estimate[1],2)`, $p `r papaja::printp(cor_age_CONCEPT$p.value[1])`$, and therefore confirmed the prediction.
+As shown in Figure 2 Panel 2A, we observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, _r_(`r cor_age_CONCEPT$parameter[1]`) = `r round(cor_age_CONCEPT$estimate[1],2)`, $p `r papaja::printp(cor_age_CONCEPT$p.value[1])`$, and therefore confirmed the prediction.
 
 Prediction 2b was that older children would produce more translation equivalents than younger children.
-We observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, _r_(`r cor_age_TE$parameter[1]`) = `r round(cor_age_TE$estimate[1],2)`, $p `r papaja::printp(cor_age_TE$p.value[1])`$, and therefore confirmed the prediction.
+As illustrated in Figure 2 Panel 2B, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, _r_(`r cor_age_TE$parameter[1]`) = `r round(cor_age_TE$estimate[1],2)`, $p `r papaja::printp(cor_age_TE$p.value[1])`$, and therefore confirmed the prediction.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_DOMSING$parameter[1]`) = `r round(cor_age_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_age_DOMSING$p.value[1])`$, and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), _r_(`r cor_BALANCE_DOMSING$parameter[1]`) = `r round(cor_BALANCE_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_DOMSING$p.value[1])`$.
-As illustrated in Figure 2 Panel 2B, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[1],0)`), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[5],0)`).
+As illustrated in Figure 2 Panel 2C, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[1],0)`), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[5],0)`).
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_NONDOMSING$parameter[1]`) = `r round(cor_age_NONDOMSING$estimate[1],2)`, $p = `r papaja::printp(cor_age_NONDOMSING$p.value[1])`$, and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), _r_(`r cor_BALANCE_NONDOMSING$parameter[1]`) = `r round(cor_BALANCE_NONDOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_NONDOMSING$p.value[1])`$.
-As shown in Figure 2 Panel 2C, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[5],0)`) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[1],0)`).
+As shown in Figure 2 Panel 2D, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[5],0)`) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[1],0)`).
 
@@
 Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
-Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2013), and amount of language experience (e.g., Weisleder & Fernald, 2013).
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder & Fernald, 2013).
 Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
@@
 
-Note also that the age group that our paper currently looked at has been suggested to learn substantially more nouns before verbs or other word classes (i.e., noun bias; Gentner, 1982).
-It has been suggested that nouns that are frequently found in the input or have concrete meanings appeared to be the easiest for young children to grasp (Braginsky et al., 2019; Goodman et al., 2008).
-On the other hand, predicates including verbs, adjectives and adverbs appear to be more difficult to learn because their meaning is more abstract and is dependent on relationships among entities (Gertner, 2006).
-Due to these differences, different lexical categories seem to be learned by young children in different ways (Braginsky et al., 2019; Imai et al., 2005).
-Since meanings of nouns are more concrete, the boundaries to categorize objects are more stable.
-Yet, predicates have boundaries that are less clearly marked without knowledge of neighboring words (Imai et al., 2008; Saji et al., 2020).
-Future investigation may explore if bilingual vocabulary learning would differ among nouns and other lexical categories.
+Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
+However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., “soap,” “bath,” “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
+For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquired across languages and cultures.
+It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
 
@@
 
-Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
-However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
-For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., “soap,” “bath,” “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
-Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
-Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
-
 We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
@@
 
+Finally, the bilingual children in our sample were learning one particular language pair: French and English.
+In both French and English, children tend to initially learn more nouns than verbs or other word classes (i.e., noun bias; Gentner, 1982), whereas children learning other languages such as Mandarin show more equal learning of nouns and verbs (Tardif, 1996).
+Moreover, French and English are also typologically similar, for example they share the same basic subject-verb-object word order.
+Similarities across bilinguals’ two languages might make it more likely that children will acquire translation equivalents.
+Future research could extend our approach to bilinguals learning different language pairs, and could also examine translation equivalent learning separately for different word classes.
+
 ## Conclusions

### 2021-12-13 paper/paper_TE_bilingual_vocabulary_model.tex
Updated paper

@@
 \item
-  Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE).
+  Prediction 1a: Children with more balanced vocabularies (BALANCE) will produce more translation equivalents (TE) (Figure 1, Panel 1A).
 \item
-  Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE).
+  Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE) (Figure 1, Panels 1A and 1B).
 \item
-  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (see panel 1B of Figure 1).
+  Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE).
+Unlike for WORD and DOM this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
 \end{itemize}
@@
 
-\caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to dominant vocabulary size (DOM; Panel A), non-dominant vocabulary size (NONDOM; Panel B), and word vocabulary (WORD; Panel C).
+\caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to word vocabulary (WORD; Panel A), dominant vocabulary size (DOM; Panel B), and non-dominant vocabulary size (NONDOM; Panel C).
 Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
@@
 We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
-Figure 2 plots the number of TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
-Across the three Panels (1A, 1B, and 1C), simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
-In Panel 1A, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
+Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
+As shown in Panel 1A, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
+In Panel 1B, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
 Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Panels 1B and 1C show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
+Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
 Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
@@
 \item
-  Prediction 2a: Have larger concept vocabularies (CONCEPT).
+  Prediction 2a: Have larger concept vocabularies (CONCEPT). (Figure 2, Panel 1A)
 \item
-  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE).
+  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE). (Figure 2, Panel 1B)
 \item
-  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least \emph{balanced} vocabulary (BALANCE) will produce the most DOM-SINGLET.
+  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least \emph{balanced} vocabulary (BALANCE) will produce the most DOM-SINGLET. (Figure 2, Panel 1C)
 \item
-  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most \emph{balanced} vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET).
+Moreover, those with the most \emph{balanced} vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
+(Figure 2, Panel 1D)
 \end{itemize}
@@
 
-{\centering \includegraphics[width=0.8\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig2-1}
+{\centering \includegraphics[width=1\linewidth]{paper_TE_bilingual_vocabulary_model_files/figure-latex/fig2-1}
 
@@
 
-\caption{Number of translation equivalents (TE; Panel A), singlets in dominant (DOM-SINGLET; Panel B) and singlets in the non-dominant language (NONDOM-SINGLET; Panel C) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+\caption{Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
 Row 1 represents the simulated data in Study 1.
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): \emph{r}(227) = 0.90, \(p < .001\), and dominant-language vocabulary (DOM): \emph{r}(227) = 0.76, \(p < .001\).
-Figure 1 Row 2 further illustrates these relationships observed in our dataset.
+Figure 1 Panels 2A and 2B further illustrates these relationships observed in our dataset.
 
 Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
-As shown in Figure 1 Row 2 (Observed data), we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
+As shown in Figure 1 Panel 2C, we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
 
@@
 In our data set, developmental level was approximated by children's age.
-Figure 2 Panel B shows the number of translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
 
 Prediction 2a was that older children (i.e., those at a later developmental level) would have larger concept vocabularies than younger children (i.e., those at an earlier developmental level).
-We observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, \emph{r}(227) = 0.69, \(p < .001\), and therefore confirmed the prediction.
+As shown in Figure 2 Panel 2A, we observed a positive correlation between age (used as a proxy for developmental level, which determines LEARNABLE) and concept vocabulary (CONCEPT) in our dataset, \emph{r}(227) = 0.69, \(p < .001\), and therefore confirmed the prediction.
 
 Prediction 2b was that older children would produce more translation equivalents than younger children.
-We observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, \emph{r}(227) = 0.48, \(p < .001\), and therefore confirmed the prediction.
+As illustrated in Figure 2 Panel 2B, we observed a positive correlation between age (our proxy for LEARNABLE) and number of translation equivalents in our dataset, \emph{r}(227) = 0.48, \(p < .001\), and therefore confirmed the prediction.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.65, \(p < .001\), and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), \emph{r}(227) = -0.58, \(p < .001\).
-As illustrated in Figure 2 Panel 2B, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = 294), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = 66).
+As illustrated in Figure 2 Panel 2C, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = 294), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = 66).
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.19, \(p = .005\), and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), \emph{r}(227) = 0.63, \(p < .001\).
-As shown in Figure 2 Panel 2C, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = 39) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = 4).
+As shown in Figure 2 Panel 2D, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = 39) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = 4).
 
@@
 Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
-Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2013), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
 Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
@@
 
-Note also that the age group that our paper currently looked at has been suggested to learn substantially more nouns before verbs or other word classes (i.e., noun bias; Gentner, 1982).
-It has been suggested that nouns that are frequently found in the input or have concrete meanings appeared to be the easiest for young children to grasp (Braginsky et al., 2019; Goodman et al., 2008).
-On the other hand, predicates including verbs, adjectives and adverbs appear to be more difficult to learn because their meaning is more abstract and is dependent on relationships among entities (Gertner, 2006).
-Due to these differences, different lexical categories seem to be learned by young children in different ways (Braginsky et al., 2019; Imai et al., 2005).
-Since meanings of nouns are more concrete, the boundaries to categorize objects are more stable.
-Yet, predicates have boundaries that are less clearly marked without knowledge of neighboring words (Imai et al., 2008; Saji et al., 2020).
-Future investigation may explore if bilingual vocabulary learning would differ among nouns and other lexical categories.
+Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
+However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
+For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
+For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
+It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
 
@@
 
-Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
-However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
-For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
-Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
-Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
-
 We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
@@
 Future work might find an improved method for estimating the number of words that are potentially learnable for individual children.
+
+Finally, the bilingual children in our sample were learning one particular language pair: French and English.
+In both French and English, children tend to initially learn more nouns than verbs or other word classes (i.e., noun bias; Gentner, 1982), whereas children learning other languages such as Mandarin show more equal learning of nouns and verbs (Tardif, 1996).
+Moreover, French and English are also typologically similar, for example they share the same basic subject-verb-object word order.
+Similarities across bilinguals' two languages might make it more likely that children will acquire translation equivalents.
+Future research could extend our approach to bilinguals learning different language pairs, and could also examine translation equivalent learning separately for different word classes.
 

### 2021-12-16 paper/paper_TE_bilingual_vocabulary_model.Rmd
Updated paper

@@
 At smaller vocabulary sizes, data matched the Preference Account, while at larger vocabulary sizes they matched the Neutral Account.
-Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.
+Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and potentially reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.
 
@@
 
+We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
+BIAS was once again constant at 1.
+Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
+Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
+In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
+Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
+Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
+
 ```{r table3, results="asis", cache = TRUE}
@@
 ```
-
-We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
-BIAS was once again constant at 1.
-Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
-Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
-Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
-In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
 
@@
 ```{r fig1, fig.cap="Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to word vocabulary (WORD; Panel A), dominant vocabulary size (DOM; Panel B), and non-dominant vocabulary size (NONDOM; Panel C).
-Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
-Row 2 represents the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%", fig.pos = "H", cache = TRUE}
+Row 1 plots the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
+Row 2 plots the observed vocabulary data in Study 2.", echo=FALSE, dpi=600, fig.align='center', fig.height=7, out.height="80%", out.width="110%", fig.pos = "H", cache = TRUE}
 
@@
 
-Translation equivalent knowledge was simulated across children at four developmental levels (the number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
+Translation equivalent knowledge was simulated across children at six developmental levels (the number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
 In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
@@
 Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages.
+Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages, and that more balanced children would produce more singlets in their non-dominant language, but fewer singlets in their dominant language than less balanced children.
 Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
@@
                              TE = "Panel 1B: \n Translation equivalents \n (TE)",
-                             DOM_SINGLET = "Panel 1C: \n Singlets in dominant language \n (DOM_SINGLET)",
-                             NONDOM_SINGLET = "Panel 1D: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             DOM_SINGLET = "Panel 1C: Singlets in \n dominant language \n (DOM_SINGLET)",
+                             NONDOM_SINGLET = "Panel 1D: Singlets in \n non-dominant language \n (NONDOM_SINGLET)")) %>%
   ggplot(aes(x = LEARNABLE, y = number, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
                              number_of_te = "Panel 2B: \n Translation equivalents \n (TE)",
-                             total_singlet_dom = "Panel 2C: \n Singlets in dominant language \n (DOM_SINGLET)",
-                             total_singlet_nondom = "Panel 2D: \n Singlets in non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             total_singlet_dom = "Panel 2C: Singlets in \n dominant language \n (DOM_SINGLET)",
+                             total_singlet_nondom = "Panel 2D: Singlets in \n non-dominant language \n (NONDOM_SINGLET)")) %>%
   ggplot(aes(x = age_months_percentile, y = number_of_words, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
 In Figure 3, we illustrate the three different scenarios of simulated translation equivalent (TE) knowledge.
-Again, we continue to observe a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
+Again, we observed a pattern consistent with prediction 1a where, in all cases, simulated children with more balanced vocabularies (BALANCE) produced more translation equivalents (TE).
 Thus, overall relationships between BALANCE and TE remained similar across the Avoidance, Preference, and Neutral Accounts.
@@
 ## 2.1 Method
-Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certification Number 10000439) and informed consent was obtained from the children’s parents.
+Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certificate Number 10000439) and informed consent was obtained from the children’s parents.
 
@@
 
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),2)` - `r round(max(keepers_ws_TE$age_continuous),2)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
+Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),1)` - `r round(max(keepers_ws_TE$age_continuous),1)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
 (2020).
@@
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
-A full list of translation equivalents is available at [https://osf.io/2t5kw/].
+A full list of translation equivalents is available at https://osf.io/2t5kw/.
 
@@
 Data analyses were conducted using R (Version 4.0.2, 2020).
-Analysis scripts and the data set used in the present study are available at [https://osf.io/2t5kw/].
+Analysis scripts and the data set used in the present study are available at https://osf.io/2t5kw/.
 We first present descriptive measures of vocabulary, and then tests of the three sets of predictions generated in Study 1.
@@
 
-On average, bilinguals in the sample had a mean word vocabulary size (WORD) of `r vocabulary_average$mean[7]` (SD = `r vocabulary_average$sd[7]`), with a wide range of `r vocabulary_average$min[7]` - `r vocabulary_average$max[7]` words.
-As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = `r vocabulary_average$mean[5]`, SD = `r vocabulary_average$sd[5]`, range = `r vocabulary_average$min[5]` - `r vocabulary_average$max[5]`) than in their non-dominant language (NONDOM; M = `r vocabulary_average$mean[6]`, SD = `r vocabulary_average$sd[6]`, range = `r vocabulary_average$min[6]` - `r vocabulary_average$max[6]`), _t_(`r t_test_DOM_NONDOM$parameter[1]`) = `r t_test_DOM_NONDOM$statistic[1]`, $p `r papaja::printp(t_test_DOM_NONDOM$p.value[1])`$, $d = `r round(d_DOM_NONDOM, 2)`$.
-
-Children produced an average of `r vocabulary_average$mean[2]` translation equivalents (TE; SD = `r vocabulary_average$sd[2]`, range = `r vocabulary_average$min[2]` - `r vocabulary_average$max[2]`).
-The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = `r vocabulary_average$mean[3]`, SD = `r vocabulary_average$sd[3]`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r vocabulary_average$mean[4]`, SD = `r vocabulary_average$sd[4]`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
-On average, children’s concept vocabulary size was `r vocabulary_average$mean[1]` (CONCEPT; SD = `r vocabulary_average$sd[1]`, range = `r vocabulary_average$min[1]` - `r vocabulary_average$max[1]`).
+On average, bilinguals in the sample had a mean word vocabulary size (WORD) of `r round(vocabulary_average$mean[7],1)` (SD = `r round(vocabulary_average$sd[7],1)`), with a wide range of `r vocabulary_average$min[7]` - `r vocabulary_average$max[7]` words.
+As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = `r round(vocabulary_average$mean[5],1)`, SD = `r round(vocabulary_average$sd[5],1)`, range = `r vocabulary_average$min[5]` - `r vocabulary_average$max[5]`) than in their non-dominant language (NONDOM; M = `r round(vocabulary_average$mean[6],1)`, SD = `r round(vocabulary_average$sd[6],1)`, range = `r vocabulary_average$min[6]` - `r vocabulary_average$max[6]`), _t_(`r t_test_DOM_NONDOM$parameter[1]`) = `r t_test_DOM_NONDOM$statistic[1]`, $p `r papaja::printp(t_test_DOM_NONDOM$p.value[1])`$, $d = `r round(d_DOM_NONDOM, 2)`$.
+
+Children produced an average of `r round(vocabulary_average$mean[2],1)` translation equivalents (TE; SD = `r round(vocabulary_average$sd[2],1)`, range = `r vocabulary_average$min[2]` - `r vocabulary_average$max[2]`).
+The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = `r round(vocabulary_average$mean[3],1)`, SD = `r round(vocabulary_average$sd[3],1)`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r round(vocabulary_average$mean[4],1)`, SD = `r round(vocabulary_average$sd[4],1)`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
+On average, children’s concept vocabulary size was `r round(vocabulary_average$mean[1],1)` words (CONCEPT; SD = `r round(vocabulary_average$sd[1],1)`, range = `r vocabulary_average$min[1]` - `r vocabulary_average$max[1]`).
 
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error alone (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 In our data set, developmental level was approximated by children’s age.
-Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (our proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_DOMSING$parameter[1]`) = `r round(cor_age_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_age_DOMSING$p.value[1])`$, and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), _r_(`r cor_BALANCE_DOMSING$parameter[1]`) = `r round(cor_BALANCE_DOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_DOMSING$p.value[1])`$.
-As illustrated in Figure 2 Panel 2C, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[1],0)`), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = `r round(mean_DOMSINGLET_BALANCEsubset$mean_DOM_SINGLET[5],0)`).
+This is illustrated in Figure 2 Panel 2C.
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), _r_(`r cor_age_NONDOMSING$parameter[1]`) = `r round(cor_age_NONDOMSING$estimate[1],2)`, $p = `r papaja::printp(cor_age_NONDOMSING$p.value[1])`$, and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), _r_(`r cor_BALANCE_NONDOMSING$parameter[1]`) = `r round(cor_BALANCE_NONDOMSING$estimate[1],2)`, $p `r papaja::printp(cor_BALANCE_NONDOMSING$p.value[1])`$.
-As shown in Figure 2 Panel 2D, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[5],0)`) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = `r round(mean_NONDOMSINGLET_BALANCEsubset$mean_NONDOM_SINGLET[1],0)`).
+This is illustrated in Figure 2 Panel 2D.
 
@@
 For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
+Finally, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
 
@@
         caption = "Table 6.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
 Values are rounded to the nearest integer.") %>%
@@
 
-Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
 To reproduce the Bilingual Vocabulary Model’s formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included (denoted in R by using a colon rather than an asterisk between the interacting predictors).
-Therefore, our final linear regression model equation was:
+Therefore, our final linear regression model equation was[^4]:
 
@@
 
-\noindent (In R language, the model formula was rearranged and entered as:
-
-_ObservedTE * 90th percentile of CDI items ~ 0 + Dominant vocabulary:Non-dominant vocabulary_)
-
-\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was related to the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+[^4]: Note that for the purposes of running the model in R, the formula was rearranged and entered as _ObservedTE * 90th percentile of CDI items ~ 0 + Dominant vocabulary:Non-dominant vocabulary_.
+
+With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model indicates how the BIAS parameter is related to the empirical vocabulary data, which in turn indicates whether bilingual children were biased towards or against learning translation equivalents.
 If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
-Otherwise, a coefficient less than 1 represents a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 represents a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
+Otherwise, a coefficient less than 1 indicates a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 indicates a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
 
@@
 Our model showed an excellent model fit of $R^2$ = `r round(summary(test_prediction3_model)$r.squared,2)`, indicating that our model explained `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of the variance in bilinguals’ translation equivalent knowledge.
-The linear regression model estimated a BIAS coefficient of `r round(summary(test_prediction3_model)$coefficient[1],2)`, $p `r papaja::printp(summary(test_prediction3_model)$coefficient[4])`$.
-This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^4].
-
-[^4]: We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
-Overall, model fit was slightly reduced compared to our main model, although fit was still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`, $p `r papaja::printp(summary(test_prediction3_model_75percentile)$coefficient[4])`$.
+BIAS coefficient was estimated at `r round(summary(test_prediction3_model)$coefficient[1],2)`.
+This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^5].
+
+[^5]: We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
+Overall, model fit was slightly reduced compared to our main model, although fit was still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`.
 The estimated BIAS parameter was lower, at `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
@@
 Cook’s distance was estimated for our linear regression model listed above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a cook’s distance over 0.4.
-After removing those `r words(length(number_of_outliers$ID_testdate))` data points, the linear regression model returned a coefficient of `r round(summary(model_remove_outliers)$coefficient[1],2)`, $p `r papaja::printp(summary(model_remove_outliers)$coefficient[4])`$, with $R^2$ = `r round(summary(model_remove_outliers)$r.squared,2)`.
+After removing those `r words(length(number_of_outliers$ID_testdate))` data points, the linear regression model returned a coefficient of `r round(summary(model_remove_outliers)$coefficient[1],2)`, with $R^2$ = `r round(summary(model_remove_outliers)$r.squared,2)`.
 As the model fit was similar to the model without eliminating the `r words(length(number_of_outliers$ID_testdate))` outlier data points, we proceeded with the full data set keeping the `r words(length(number_of_outliers$ID_testdate))` potential outlier data points.
@@
 To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
-The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_more300w)$coefficient[1],2)`, $p `r papaja::printp(summary(percentile_model_more300w)$coefficient[4])`$, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_less300w)$coefficient[1],2)`, $p `r papaja::printp(summary(percentile_model_less300w)$coefficient[4])`$.
+The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_more300w)$coefficient[1],2)`, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_less300w)$coefficient[1],2)`.
 Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size ($R^2$ = `r round(summary(percentile_model_more300w)$r.squared,2)`) than children with smaller vocabulary size ($R^2$ = `r round(summary(percentile_model_less300w)$r.squared,2)`).
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
-For example, a caregiver may choose to say to their English–French bilingual child “Can you give me the _livre_?” if they know their child understands the French word _“livre”_ but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English–French bilingual child “Can you give me the _livre_?” if they know their child understands the French word “_livre_” but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
 Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
-Overall, more research will be needed to understand why translation equivalents appear to be over-represented in younger bilinguals’ vocabularies.
-
-Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
+
+Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to changes in input patterns or word learning biases.
 Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder & Fernald, 2013).
@@
 Such a model may better characterize and predict individual differences in bilingual vocabulary development.
-Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data and/or data of a different bilingual.
+Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data.
 
@@
 We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
-Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only `r ws_percentile$"Average (LEARNABLE)"[1]/2` words in each language are learnable for 18-month-olds, rather than `r ws_percentile$"Average (LEARNABLE)"[1]`), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of `r round(summary(percentile_model_less300w)$coefficient[1],2)` which supports the Preference account, we would estimate a parameter of `r round(summary(percentile_model_less300w)$coefficient[1]/2, 2)` which is closer to the Neutral Account).
+Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will produce based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only `r ws_percentile$"Average (LEARNABLE)"[1]/2` words in each language are learnable for 18-month-olds, rather than `r ws_percentile$"Average (LEARNABLE)"[1]`), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of `r round(summary(percentile_model_less300w)$coefficient[1],2)` which supports the Preference account, we would estimate a parameter of `r round(summary(percentile_model_less300w)$coefficient[1]/2, 2)` which is closer to the Neutral Account).
 Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around `r round(summary(test_prediction3_model)$coefficient[1],2)`, which supports the Neutral account.
@@
 However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is lower, although nevertheless still consistent with the Neutral Account.
-Future work might find an improved method for estimating the number of words that are potentially learnable for individual children.
+It will be important for future work to more fully explore the learnability parameter, for example, finding a method for estimating the number of words that are potentially learnable for individual children.
 
@@
 
+# Acknowledgments
+\noindent We are grateful to all the families who participated in this research.
+This work was funded by a grant from the Natural Sciences and Engineering Research Council of Canada (NSERC 2011-402470, 2018-04390), a National Institutes of Health grant (NIH R03HD079779), and a Fonds de Recherche du Québec-Société et Culture grant (FQRSC 2012-NP-145009) to Byers-Heinlein, by a FQRSC (2018-B3-205717) postdoctoral fellowship to Gonzalez-Barrero, and by a FQRSC graduate fellowship and Concordia University graduate fellowship to Schott.
+Byers-Heinlen holds the Concordia University Research Chair in Bilingualism and Open Science.
+This work was presented at the 2021 Boston University Conference on Language Development.
+We thank the members of the Concordia Infant Research Lab for their comments on earlier versions of this paper, and Roman Feiman for thoughtful discussion that led to new insights.
+
+&nbsp;
+&nbsp;
+
+# Author Contributions
+\noindent __Rachel Ka-Ying Tsui:__ Conceptualization, Formal analysis, Visualization, Writing - Original draft preparation, Writing - Review & Editing, Project administration.
+**Ana Maria Gonzalez-Barrero:** Conceptualization, Data Curation, Writing - Review & Editing.
+**Esther Schott:** Conceptualization, Data Curation, Writing - Review & Editing.
+**Krista Byers-Heinlein:** Conceptualization, Writing - Original draft preparation, Writing - Review & Editing,  Supervision, Project administration.
+
+\newpage
+
 # References

### 2021-12-16 paper/paper_TE_bilingual_vocabulary_model.tex
Updated paper

@@
 At smaller vocabulary sizes, data matched the Preference Account, while at larger vocabulary sizes they matched the Neutral Account.
-Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.
+Our findings show that patterns of translation equivalent learning emerge predictably from the word learning process, and potentially reveal a qualitative shift in translation equivalent learning as bilingual children develop and learn more words.
 }
@@
 
+We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
+BIAS was once again constant at 1.
+Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
+Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
+In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
+Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
+Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
+
 \begin{landscape}\begin{table}
@@
 
-We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
-BIAS was once again constant at 1.
-Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
-Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
-Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
-In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
-
 \begin{itemize}
@@
 
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 3 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 25 rows containing missing values (geom_smooth).
+\end{verbatim}
+
 \begin{figure}[H]
@@
 \caption{Number of translation equivalents (TE) across different levels of vocabulary balance (BALANCE) in relation to word vocabulary (WORD; Panel A), dominant vocabulary size (DOM; Panel B), and non-dominant vocabulary size (NONDOM; Panel C).
-Row 1 represents the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
-Row 2 represents the observed vocabulary data in Study 2.}\label{fig:fig1}
+Row 1 plots the simulated data in Study 1 while holding the number of learnable words (LEARNABLE) constant at 600 and BIAS constant at 1.
+Row 2 plots the observed vocabulary data in Study 2.}\label{fig:fig1}
 \end{figure}
@@
 
+\begin{verbatim}
+## Warning: Removed 39 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 1 rows containing missing values (geom_smooth).
+\end{verbatim}
+
 \begin{figure}[H]
@@
 
-Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certification Number 10000439) and informed consent was obtained from the children's parents.
+Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certificate Number 10000439) and informed consent was obtained from the children's parents.
 
@@
 
-Archival data from 200 bilingual children acquiring English and French (age range: 18.38 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Archival data from 200 bilingual children acquiring English and French (age range: 18.40 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
 Data collection was conducted in Montréal, Québec, Canada.
@@
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
-A full list of translation equivalents is available at {[}\url{https://osf.io/2t5kw/}{]}.
+A full list of translation equivalents is available at \url{https://osf.io/2t5kw/}.
 
@@
 Data analyses were conducted using R (Version 4.0.2, 2020).
-Analysis scripts and the data set used in the present study are available at {[}\url{https://osf.io/2t5kw/}{]}.
+Analysis scripts and the data set used in the present study are available at \url{https://osf.io/2t5kw/}.
 We first present descriptive measures of vocabulary, and then tests of the three sets of predictions generated in Study 1.
@@
 The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = 138.40, SD = 124.40, range = 2 - 523) than in their non-dominant language (NONDOM-SINGLET; M = 21.20, SD = 20.10, range = 0 - 94), \emph{t}(228) = 13.89, \(p < .001\), \(d = 0.92\).
-On average, children's concept vocabulary size was 227.30 (CONCEPT; SD = 181.30, range = 4 - 695).
+On average, children's concept vocabulary size was 227.30 words (CONCEPT; SD = 181.30, range = 4 - 695).
 
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error alone (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 In our data set, developmental level was approximated by children's age.
-Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (a proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (our proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 This pattern was confirmed by the results from our dataset, with a positive correlation between dominant-language singlets (DOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.65, \(p < .001\), and a negative correlation between BALANCE and dominant-language singlets (DOM-SINGLET), \emph{r}(227) = -0.58, \(p < .001\).
-As illustrated in Figure 2 Panel 2C, children with the least balanced vocabulary of 0.1 produced the most singlets in their dominant language (mean = 294), whereas those with the most balanced vocabulary of 0.5 produced the fewest (mean = 66).
+This is illustrated in Figure 2 Panel 2C.
 
@@
 This pattern was also observed in our dataset, with a positive correlation between the number of non-dominant singlets (NONDOM-SINGLET) and age (which determined LEARNABLE), \emph{r}(227) = 0.19, \(p = .005\), and a positive correlation between BALANCE and the number of non-dominant singlets (NONDOM-SINGLET), \emph{r}(227) = 0.63, \(p < .001\).
-As shown in Figure 2 Panel 2D, we observed that children produced very few singlets in their non-dominant language, although the most balanced children with BALANCE of 0.5 produced the most singlets in their non-dominant language (mean = 39) and the least balanced children with BALANCE of 0.1 produced the fewest (mean = 4).
+This is illustrated in Figure 2 Panel 2D.
 
@@
 \caption{\label{tab:table6}Table 6.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages, and the average between the two which serves as the denominator in our computation model.
+The number of total English and French productive CDI vocabulary at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
 Values are rounded to the nearest integer.}
@@
 
-Furthermore, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
 To reproduce the Bilingual Vocabulary Model's formula TE = DOM×NONDOM/LEARNABLE, an interaction between dominant and non-dominant vocabulary was entered in the model, but main effects were not included (denoted in R by using a colon rather than an asterisk between the interacting predictors).
-Therefore, our final linear regression model equation was:
+Therefore, our final linear regression model equation was\footnote{Note that for the purposes of running the model in R, the formula was rearranged and entered as \emph{ObservedTE * 90th percentile of CDI items \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary}.}:
 
@@
 
-\noindent (In R language, the model formula was rearranged and entered as:
-
-\emph{ObservedTE * 90th percentile of CDI items \textasciitilde{} 0 + Dominant vocabulary:Non-dominant vocabulary})
-
-\noindent With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model would indicate how the BIAS parameter was related to the empirical vocabulary data, which would then indicate whether bilingual children were biased towards or against learning translation equivalents.
+With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model indicates how the BIAS parameter is related to the empirical vocabulary data, which in turn indicates whether bilingual children were biased towards or against learning translation equivalents.
 If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
-Otherwise, a coefficient less than 1 represents a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 represents a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
+Otherwise, a coefficient less than 1 indicates a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 indicates a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
 
 Our model showed an excellent model fit of \(R^2\) = 0.96, indicating that our model explained 96\% of the variance in bilinguals' translation equivalent knowledge.
-The linear regression model estimated a BIAS coefficient of 1.04, \(p < .001\).
+BIAS coefficient was estimated at 1.04.
 This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
-Overall, model fit was slightly reduced compared to our main model, although fit was still good, \(R^2\) = 0.95, \(p < .001\).
+Overall, model fit was slightly reduced compared to our main model, although fit was still good, \(R^2\) = 0.95.
 The estimated BIAS parameter was lower, at 0.95, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
@@
 Cook's distance was estimated for our linear regression model listed above and identified two data points with a cook's distance over 0.4.
-After removing those two data points, the linear regression model returned a coefficient of 1.07, \(p < .001\), with \(R^2\) = 0.96.
+After removing those two data points, the linear regression model returned a coefficient of 1.07, with \(R^2\) = 0.96.
 As the model fit was similar to the model without eliminating the two outlier data points, we proceeded with the full data set keeping the two potential outlier data points.
@@
 To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
-The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.04, \(p < .001\), whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.21, \(p < .001\).
+The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.04, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.21.
 Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size (\(R^2\) = 0.96) than children with smaller vocabulary size (\(R^2\) = 0.88).
@@
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
-For example, a caregiver may choose to say to their English--French bilingual child ``Can you give me the \emph{livre}?'' if they know their child understands the French word \emph{``livre''} but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+For example, a caregiver may choose to say to their English--French bilingual child ``Can you give me the \emph{livre}?'' if they know their child understands the French word ``\emph{livre}'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
 However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
 Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
-Overall, more research will be needed to understand why translation equivalents appear to be over-represented in younger bilinguals' vocabularies.
-
-Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to word learning biases.
+
+Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to changes in input patterns or word learning biases.
 Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
@@
 We must also note the reciprocal relationship in the Bilingual Vocabulary Model between the parameter that indicates whether or not children are biased to learn translation equivalents and the parameter that accounts for how many words are potentially learnable at a particular age.
-Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will learn based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decrease by a factor of two (e.g., whereby only 120 words in each language are learnable for 18-month-olds, rather than 240), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.21 which supports the Preference account, we would estimate a parameter of 1.10 which is closer to the Neutral Account).
+Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will produce based on the number of words that they know in each of their languages.
+That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only 120 words in each language are learnable for 18-month-olds, rather than 240), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.21 which supports the Preference account, we would estimate a parameter of 1.10 which is closer to the Neutral Account).
 Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.04, which supports the Neutral account.
@@
 However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to 0.95, which is lower, although nevertheless still consistent with the Neutral Account.
-Future work might find an improved method for estimating the number of words that are potentially learnable for individual children.
+It will be important for future work to more fully explore the learnability parameter, for example, finding a method for estimating the number of words that are potentially learnable for individual children.
 
@@
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
+
+\newpage
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+\noindent We are grateful to all the families who participated in this research.
+This work was funded by a grant from the Natural Sciences and Engineering Research Council of Canada (NSERC 2011-402470, 2018-04390), a National Institutes of Health grant (NIH R03HD079779), and a Fonds de Recherche du Québec-Société et Culture grant (FQRSC 2012-NP-145009) to Byers-Heinlein, by a FQRSC (2018-B3-205717) postdoctoral fellowship to Gonzalez-Barrero, and by a FQRSC graduate fellowship and Concordia University graduate fellowship to Schott.
+Byers-Heinlen holds the Concordia University Research Chair in Bilingualism and Open Science.
+This work was presented at the 2021 Boston University Conference on Language Development.
+We thank the members of the Concordia Infant Research Lab for their comments on earlier versions of this paper, and Roman Feiman for thoughtful discussion that led to new insights.
+
+~
+~
+
+\hypertarget{author-contributions}{%
+\section{Author Contributions}\label{author-contributions}}
+
+\noindent \textbf{Rachel Ka-Ying Tsui:} Conceptualization, Formal analysis, Visualization, Writing - Original draft preparation, Writing - Review \& Editing, Project administration.
+\textbf{Ana Maria Gonzalez-Barrero:} Conceptualization, Data Curation, Writing - Review \& Editing.
+\textbf{Esther Schott:} Conceptualization, Data Curation, Writing - Review \& Editing.
+\textbf{Krista Byers-Heinlein:} Conceptualization, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
 

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated paper

@@
 For instance, to refer to the same crisp red-skinned fruit, an English–French bilingual child must use the word “apple” when speaking English, and the word _“pomme”_ when speaking French.
-These cross-language synonyms are known as translation equivalents [^1] (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children’s first words (e.g., David & Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual’s developing lexicon due to the strong overlap in their semantics.
@@
 That is, upon hearing the English word “apple,” the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent  (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet).
+In vocabulary acquisition, bilingual children must learn a first label for a referent  (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)[^1].
 Is translation equivalent learning different from singlet learning?
@@
 If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) — which is the possible number of words that could overlap across both languages — to predict the number of translation equivalents.
-We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral account.
+We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral Account.
 Adding this parameter, translation equivalents can be derived from TE = BIAS × (DOM × NONDOM)/LEARNABLE.
@@
 
-A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets.
 We operationalized developmental level in terms of children’s age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
@@
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
-Modeling their data separately, we found evidence for the Preference account: younger children at around `r round(keepers_ws_TE %>% filter(word_vocab < 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months appeared to learn translation equivalents more easily than singlets, whereas older children at around `r round(keepers_ws_TE %>% filter(word_vocab > 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months learned translation equivalents similarly to singlets.
+Modeling their data separately, we found evidence for the Preference Account: younger children at around `r round(keepers_ws_TE %>% filter(word_vocab < 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months appeared to learn translation equivalents more easily than singlets, whereas older children at around `r round(keepers_ws_TE %>% filter(word_vocab > 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months learned translation equivalents similarly to singlets.
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
@@
 
-Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
+Another possible explanation for our findings is that the nature of bilingual input changes as children become more advanced word learners.
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
@@
 It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
-Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance account (e.g., fewer than expected translation equivalents), which is not what we observed.
+Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance Account (e.g., fewer than expected translation equivalents), which is not what we observed.
 Nonetheless, future studies of bilingual corpora could directly address whether early translation equivalent learning might be impacted by the Complementarity Principle.
 
-Another limitation of our model is that it takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
+Another consideration is that our model is that it takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
 However, the reality of bilingual experience might be more complex.
@@
 There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storm et al., 2015; White et al., 2020).
-Therefore, it seems reasonable for our model to take the simplified view that early vocabulary may be treated similarly across languages, especially for the age group we were looking at.
+Therefore, it seems reasonable to assume that early vocabulary may be represented similarly across languages, especially for the age group we were looking at.
 Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where concepts are less well aligned across languages.
@@
 However, the close correspondence between our model and data from bilingual children suggest that even if our model is a simplification, deviations from our assumptions might have a relatively small impact.
-On the other hand, individual item differences could be particularly consequential at younger ages, and the apparent shift we observed in development from the enhancement account to the avoidance account could be an artifact of this issue.
+On the other hand, individual item differences could be particularly consequential at younger ages, and the apparent shift we observed in development from the Preference Account to the Neutral Account could be an artifact of this issue.
 Future iterations of the Bilingual Vocabulary Model could explicitly add such word-level factors into the model and test the importance of those factors in early bilingual vocabulary development.
@@
 Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will produce based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only `r ws_percentile$"Average (LEARNABLE)"[1]/2` words in each language are learnable for 18-month-olds, rather than `r ws_percentile$"Average (LEARNABLE)"[1]`), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of `r round(summary(percentile_model_less300w)$coefficient[1],2)` which supports the Preference account, we would estimate a parameter of `r round(summary(percentile_model_less300w)$coefficient[1]/2, 2)` which is closer to the Neutral Account).
-Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around `r round(summary(test_prediction3_model)$coefficient[1],2)`, which supports the Neutral account.
+That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only `r ws_percentile$"Average (LEARNABLE)"[1]/2` words in each language are learnable for 18-month-olds, rather than `r ws_percentile$"Average (LEARNABLE)"[1]`), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of `r round(summary(percentile_model_less300w)$coefficient[1],2)` which supports the Preference Account, we would estimate a parameter of `r round(summary(percentile_model_less300w)$coefficient[1]/2, 2)` which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around `r round(summary(test_prediction3_model)$coefficient[1],2)`, which supports the Neutral Account.
 The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
 However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is lower, although nevertheless still consistent with the Neutral Account.
-It will be important for future work to more fully explore the learnability parameter, for example, finding a method for estimating the number of words that are potentially learnable for individual children.
+It will be important for future work to more fully explore the learnability parameter, for example, by estimating the number of words that are potentially learnable for individual children while taking into account factors beyond chronological age.
 
@@
 Results using archival data from a large number of young French–English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
-Moreover, our model parsimoniously explained previously disparate observations about bilingual children’s translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Moreover, our model parsimoniously explains previously disparate observations about bilingual children’s translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
@@
 # Author Contributions
-\noindent __Rachel Ka-Ying Tsui:__ Conceptualization, Formal analysis, Visualization, Writing - Original draft preparation, Writing - Review & Editing, Project administration.
+\noindent __Rachel Ka-Ying Tsui:__ Conceptualization, Data Curation, Formal analysis, Investigation, Methodology, Visualization, Writing - Original draft preparation, Writing - Review & Editing, Project administration.
 **Ana Maria Gonzalez-Barrero:** Conceptualization, Data Curation, Writing - Review & Editing.
 **Esther Schott:** Conceptualization, Data Curation, Writing - Review & Editing.
-**Krista Byers-Heinlein:** Conceptualization, Writing - Original draft preparation, Writing - Review & Editing,  Supervision, Project administration.
+**Krista Byers-Heinlein:** Conceptualization, Funding Acquisition, Methodology, Resources, Supervision, Writing - Original draft preparation, Writing - Review & Editing, Supervision, Project administration.
 

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.tex
updated paper

@@
 For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
-These cross-language synonyms are known as translation equivalents\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
-Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
-However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
-Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.} (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
+These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
 Translation equivalents are thought to hold a special status in a bilingual's developing lexicon due to the strong overlap in their semantics.
@@
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet).
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
+Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
+However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.}.
 Is translation equivalent learning different from singlet learning?
@@
 If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) --- which is the possible number of words that could overlap across both languages --- to predict the number of translation equivalents.
-We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral account.
+We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral Account.
 Adding this parameter, translation equivalents can be derived from TE = BIAS × (DOM × NONDOM)/LEARNABLE.
@@
 \end{itemize}
-
-\begin{verbatim}
-## Warning: Removed 39 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 1 rows containing missing values (geom_smooth).
-\end{verbatim}
 
@@
 
-A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets (i.e., words without a translation equivalent).
+A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets.
 We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
@@
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
-Modeling their data separately, we found evidence for the Preference account: younger children at around 22 months appeared to learn translation equivalents more easily than singlets, whereas older children at around 28 months learned translation equivalents similarly to singlets.
+Modeling their data separately, we found evidence for the Preference Account: younger children at around 22 months appeared to learn translation equivalents more easily than singlets, whereas older children at around 28 months learned translation equivalents similarly to singlets.
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
@@
 
-Another possible explanation is that the nature of bilingual input changes as children become more advanced word learners.
+Another possible explanation for our findings is that the nature of bilingual input changes as children become more advanced word learners.
 Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
@@
 Under the Bilingual Vocabulary Model, these two parameters jointly predict the number of translation equivalents that a child will produce based on the number of words that they know in each of their languages.
-That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only 120 words in each language are learnable for 18-month-olds, rather than 240), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.21 which supports the Preference account, we would estimate a parameter of 1.10 which is closer to the Neutral Account).
-Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.04, which supports the Neutral account.
+That is, if the assumed learnability parameter decreases by a factor of two (e.g., whereby only 120 words in each language are learnable for 18-month-olds, rather than 240), then estimates of the bias parameter will also decrease by a factor of two (i.e., rather than a parameter of 2.21 which supports the Preference Account, we would estimate a parameter of 1.10 which is closer to the Neutral Account).
+Our main model estimated the number of learnable words to be that which children at the 90th percentile at a particular age produce, and this approach resulted in a bias parameter around 1.04, which supports the Neutral Account.
 The 90th percentile was chosen prior to analysis, following the reasoning that the largest vocabulary size observed at a particular age would be a reasonable estimate of how many words were potentially learnable at that age.
 However, we also explored the effects of setting the number of learnable words at the 75th percentile, and as mathematically expected, the estimated bias parameter decreased to 0.95, which is lower, although nevertheless still consistent with the Neutral Account.
-It will be important for future work to more fully explore the learnability parameter, for example, finding a method for estimating the number of words that are potentially learnable for individual children.
+It will be important for future work to more fully explore the learnability parameter, for example, by estimating the number of words that are potentially learnable for individual children while taking into account factors beyond chronological age.
 
@@
 Results using archival data from a large number of young French--English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
-Moreover, our model parsimoniously explained previously disparate observations about bilingual children's translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Moreover, our model parsimoniously explains previously disparate observations about bilingual children's translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
@@
 
-\noindent \textbf{Rachel Ka-Ying Tsui:} Conceptualization, Formal analysis, Visualization, Writing - Original draft preparation, Writing - Review \& Editing, Project administration.
+\noindent \textbf{Rachel Ka-Ying Tsui:} Conceptualization, Data Curation, Formal analysis, Investigation, Methodology, Visualization, Writing - Original draft preparation, Writing - Review \& Editing, Project administration.
 \textbf{Ana Maria Gonzalez-Barrero:} Conceptualization, Data Curation, Writing - Review \& Editing.
 \textbf{Esther Schott:} Conceptualization, Data Curation, Writing - Review \& Editing.
-\textbf{Krista Byers-Heinlein:} Conceptualization, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
+\textbf{Krista Byers-Heinlein:} Conceptualization, Funding Acquisition, Methodology, Resources, Supervision, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
 

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.Rmd
Updated paper

@@
 Adding this parameter, translation equivalents can be derived from TE = BIAS × (DOM × NONDOM)/LEARNABLE.
-For the Avoidance Account, BIAS will be less than 1, meaning that translation equivalentsare less easily learned than singlets; for the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets; for the the Neutral Account, BIAS is exactly 1 (i.e., the model is unbiased with respect to whether translation equivalents are more difficult or easier to acquire than singlets).
-Going back to the example of 18-month-old Jamie, we would set the denominator at 245 which is the number of potentially learnable words at 18 months.
-If translation equivalents are half as easy to learn as singlets (following the Avoidance Account), we would expect Jamie to produce .5×(50×20/245) = 2.0 translation equivalents.
-Conversely, if translation equivalents are twice as easy to learn as singlets (following the Preference Account), we would expect Jamie to produce 2×(50×20/245) = 8.2 translation equivalents.
-Under the Neutral Account, we would expect Jamie to learn 1×(50×20/245) = 4.1 translation equivalents.
+For the Avoidance Account, BIAS will be less than 1, meaning that translation equivalents are less easily learned than singlets; for the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets; for the Neutral Account, BIAS is exactly 1 (i.e., the model is unbiased with respect to whether translation equivalents are more difficult or easier to acquire than singlets).
+Going back to the example of 18-month-old Jamie, we would set the denominator at 240 which is the number of potentially learnable words at 18 months.
+If translation equivalents are half as easy to learn as singlets (following the Avoidance Account), we would expect Jamie to produce .5×(50×20/240) = 2.1 translation equivalents.
+Conversely, if translation equivalents are twice as easy to learn as singlets (following the Preference Account), we would expect Jamie to produce 2×(50×20/240) = 8.3 translation equivalents.
+Under the Neutral Account, we would expect Jamie to learn 1×(50×20/240) = 4.2 translation equivalents.
 
@@
                       "Bias parameter (BIAS)",
-                      "Total number of data points generated"),
+                      "Number of simulated children"),
         align = "clllllll",
@@
       caption = "Table 3.
-Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
 Vocabulary values are rounded to the nearest integer.") %>%
@@
 
-* Prediction 2a: Have larger concept vocabularies (CONCEPT). (Figure 2, Panel 1A)
-* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE). (Figure 2, Panel 1B)
-* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET. (Figure 2, Panel 1C)
-* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET. (Figure 2, Panel 1D)
+* Prediction 2a: Have larger concept vocabularies (CONCEPT) (Figure 2, Panel 1A).
+* Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE) (Figure 2, Panel 1B).
+* Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least _balanced_ vocabulary (BALANCE) will produce the most DOM-SINGLET (Figure 2, Panel 1C).
+* Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most _balanced_ vocabulary (BALANCE) will produce the most NONDOM-SINGLET (Figure 2, Panel 1D).
 
@@
 Our analyses focused on the vocabulary checklist of this questionnaire, which includes different nouns, verbs, adjectives, and other words used by young children.
-There are 680 words on the English CDI version and 664 in the Québec French version.
+There are 680 words on the English CDI version and 664 on the Québec French version.
 
@@
 As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, _r_(`r cor_TE_balance$parameter[1]`) = `r round(cor_TE_balance$estimate[1],2)`, $p `r papaja::printp(cor_TE_balance$p.value[1])`$, where children with the most balanced vocabulary produced the most translation equivalents.
-
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): _r_(`r cor_TE_word$parameter[1]`) = `r round(cor_TE_word$estimate[1],2)`, $p `r papaja::printp(cor_TE_word$p.value[1])`$, and dominant-language vocabulary (DOM): _r_(`r cor_TE_dom$parameter[1]`) = `r round(cor_TE_dom$estimate[1],2)`, $p `r papaja::printp(cor_TE_dom$p.value[1])`$.
@@
 Despite the good overall fit to the data, a close examination of Figure 5 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
-Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
+Figure 6 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
 Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.
@@
 
-Another assumption of our model was that bilingual children use labels from both languages for the same set of referents.
+One assumption of our model was that bilingual children use labels from both languages for the same set of referents.
 However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.tex
Updated paper

@@
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
-For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
+For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word ``pomme'' when speaking French.
 These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
@@
 For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
-That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
+That is, upon hearing the English word ``apple,'' the corresponding French word ``pomme'' is more easily activated in bilinguals' minds.
 In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
@@
 The researchers found that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (\emph{n} = 27) in the context of a wide age range (8 months to 2.5 years).
+This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (n = 27) in the context of a wide age range (8 months to 2.5 years).
 In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
@@
 \toprule
-Simulation & Learnable words (LEARNABLE) & Words in dominant Language (DOM) & Words in non-dominant language (NONDOM) & Word vocabulary (WORD) & Balance of vocabulary (BALANCE) & Bias parameter (BIAS) & Total number of data points generated\\
+Simulation & Learnable words (LEARNABLE) & Words in dominant Language (DOM) & Words in non-dominant language (NONDOM) & Word vocabulary (WORD) & Balance of vocabulary (BALANCE) & Bias parameter (BIAS) & Number of simulated children\\
 \midrule
@@
 \caption{\label{tab:table3}Table 3.
-Examples for Simulation 1 of three hypothetical children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
 Vocabulary values are rounded to the nearest integer.}
@@
 \item
-  Prediction 2a: Have larger concept vocabularies (CONCEPT). (Figure 2, Panel 1A)
+  Prediction 2a: Have larger concept vocabularies (CONCEPT) (Figure 2, Panel 1A).
 \item
-  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE). (Figure 2, Panel 1B)
+  Prediction 2b: Produce more translation equivalents (TE), regardless of vocabulary balance (BALANCE) (Figure 2, Panel 1B).
 \item
-  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least \emph{balanced} vocabulary (BALANCE) will produce the most DOM-SINGLET. (Figure 2, Panel 1C)
+  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least balanced vocabulary (BALANCE) will produce the most DOM-SINGLET (Figure 2, Panel 1C).
 \item
-  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET).
-Moreover, those with the most \emph{balanced} vocabulary (BALANCE) will produce the most NONDOM-SINGLET.
-(Figure 2, Panel 1D)
+  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most balanced vocabulary (BALANCE) will produce the most NONDOM-SINGLET (Figure 2, Panel 1D).
 \end{itemize}
@@
 As shown in Figure 1 Row 2 (Observed data), our vocabulary data confirmed the prediction, \emph{r}(227) = 0.25, \(p < .001\), where children with the most balanced vocabulary produced the most translation equivalents.
-
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): \emph{r}(227) = 0.90, \(p < .001\), and dominant-language vocabulary (DOM): \emph{r}(227) = 0.76, \(p < .001\).
@@
 Despite the good overall fit to the data, a close examination of Figure 5 suggested that the model might less closely fit the data of children with smaller vocabulary sizes.
-Figure 7 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
+Figure 6 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
 Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated references

@@
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Pearson et al., 1997; Hurtado et al., 2013).
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997).
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.

### 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.tex
updated references

@@
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
-For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word ``pomme'' when speaking French.
+For instance, to refer to the same crisp red-skinned fruit, an English--French bilingual child must use the word ``apple'' when speaking English, and the word \emph{``pomme''} when speaking French.
 These cross-language synonyms are known as translation equivalents (also called doublets; Umbel et al., 1992), and are observed amongst bilingual children's first words (e.g., David \& Wei, 2008; De Houwer et al., 2006; Pearson et al., 1995).
@@
 For example, studies with bilingual toddlers show that the associative semantic properties of a word in one language facilitate the activation of its translation equivalent (e.g., Bilson et al., 2015; Floccia et al., 2020; Jardak \& Byers-Heinlein, 2019).
-That is, upon hearing the English word ``apple,'' the corresponding French word ``pomme'' is more easily activated in bilinguals' minds.
+That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
 In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
@@
 The researchers found that the percentage of translation equivalents observed in English--Spanish bilingual children was similar to the by-chance percentage of translation equivalents between randomly-paired children.
-This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (n = 27) in the context of a wide age range (8 months to 2.5 years).
+This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (\emph{n} = 27) in the context of a wide age range (8 months to 2.5 years).
 In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
@@
 \item
-  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least balanced vocabulary (BALANCE) will produce the most DOM-SINGLET (Figure 2, Panel 1C).
+  Prediction 2c: Produce more dominant-language singlet words (DOM-SINGLET). Moreover, those with the least \emph{balanced} vocabulary (BALANCE) will produce the most DOM-SINGLET (Figure 2, Panel 1C).
 \item
-  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET). Moreover, those with the most balanced vocabulary (BALANCE) will produce the most NONDOM-SINGLET (Figure 2, Panel 1D).
+  Prediction 2d: Produce more non-dominant-language singlets (NONDOM-SINGLET).
+Moreover, those with the most \emph{balanced} vocabulary (BALANCE) will produce the most NONDOM-SINGLET (Figure 2, Panel 1D).
 \end{itemize}
+
+\begin{verbatim}
+## Warning: Removed 39 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 1 rows containing missing values (geom_smooth).
+\end{verbatim}
 
@@
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Pearson et al., 1997; Hurtado et al., 2013).
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997).
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
@@
 
-\leavevmode\hypertarget{ref-Bialystok_etal_2010}{}%
-Bialystok, E., Luk, G., Peets, K.
-F., \& Yang, S.
-(2010).
-Receptive vocabulary differences in monolingual and bilingual children.
-\emph{Bilingualism (Cambridge, England)}, \emph{13}(4), 525--531.
-\url{https://doi.org/10.1017/S1366728909990423}
-
 \leavevmode\hypertarget{ref-Bilson_etal_2015}{}%
@@
 
+\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2021}{}%
+Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (2021). \emph{Six solutions for more reliable infant research}. Preprint. Retrieved from \url{https://psyarxiv.com/u37fy}
+
 \leavevmode\hypertarget{ref-Byers-Heinlein_etal_2020}{}%
@@
 \url{https://doi.org/10.1016/j.cognition.2013.05.010}
+
+\leavevmode\hypertarget{ref-Cristia_etal_2014}{}%
+Cristia, A., Seidl, A., Junge, C., Soderstrom, M., \& Hagoort, P.
+(2014).
+Predicting individual variation in language from infant speech perception measures.
+\emph{Child Development}, \emph{85}(4), 1330--1345.
+\url{https://doi.org/10.1111/cdev.12193}
 
@@
 
+\leavevmode\hypertarget{ref-Gentner_1982}{}%
+Gentner, D.
+(1982).
+Why nouns are learned before verbs: Linguistic relativity versus natural partitioning.
+In S.
+A.
+Kucza (Ed.), \emph{Language development: Language, thought and culture} (pp.
+301--334).
+Hillsdale, NJ: Lawrence Erlbaum Associates.
+
 \leavevmode\hypertarget{ref-Gonzalez-Barrero_2020}{}%
@@
 Preprint.
-https://doi.org/\url{https://psyarxiv.com/x7s4u/}
+Retrieved from \url{https://psyarxiv.com/x7s4u/}
 
@@
 \url{https://doi.org/10.1016/j.cognition.2020.104191}
+
+\leavevmode\hypertarget{ref-Malt_Majid_2013}{}%
+Malt, B. C., \& Majid, A. (2013). How thought is mapped into words. \emph{Wiley Interdisciplinary Reviews: Cognitive Science}, \emph{4}(6), 583--597. \url{https://doi.org/10.1002/wcs.1251}
 
@@
 
+\leavevmode\hypertarget{ref-Mitchell_etal_2021}{}%
+Mitchell, L., Tsui, R.
+K.-Y., \& Byers-Heinlein, K.
+(2021).
+\emph{Bilingual infants learn cognates more rapidly than non-cognates}.
+Poster presented at the 46th Annual Boston University Conference on Language Development, Boston, United States.
+Retrieved from \url{https://www.bu.edu/bucld/files/2021/09/BUCLD46_Abstracts_at-a-glance.pdf}
+
 \leavevmode\hypertarget{ref-Montanari_2010}{}%
@@
 \leavevmode\hypertarget{ref-Pearson_etal_1997}{}%
-Pearson, B. Z., Fernández, S. C., Lewedeg, V., \& Oller, D. K. (1997). The relation of input factors to lexical learning by bilingual infants. \emph{Applied Psycholinguistics}, \emph{18}(1), 41--58.
+Pearson, B.
+Z., Fernández, S.
+C., Lewedeg, V., \& Oller, D.
+K.
+(1997).
+The relation of input factors to lexical learning by bilingual infants.
+\emph{Applied Psycholinguistics}, \emph{18}(1), 41--58.
+\url{https://doi.org/10.1017/S0142716400009863}
 
@@
 
+\leavevmode\hypertarget{ref-Reuter_etal_2018}{}%
+Reuter, T., Emberson, L., Romberg, A., \& Lew-Williams, C.
+(2018).
+Individual differences in nonverbal prediction and vocabulary size in infancy.
+\emph{Cognition}, \emph{176}, 215--219.
+\url{https://doi.org/10.1016/j.cognition.2018.03.006}
+
 \leavevmode\hypertarget{ref-Rowe_2012}{}%
@@
 Singh, L. (2014). One world, two languages: Cross-language semantic priming in bilingual toddlers. \emph{Child Development}, \emph{85}(2), 755--766. \url{https://doi.org/10.1111/cdev.12133}
+
+\leavevmode\hypertarget{ref-Storms_etal_2015}{}%
+Storms, G., Ameel, E., \& Malt, B.
+C.
+(2015).
+Development of cross-language lexical influence.
+\emph{International Journal of Bilingual Education and Bilingualism}, \emph{18}(5), 529--547.
+\url{https://doi.org/10.1080/13670050.2015.1027142}
 
@@
 
+\leavevmode\hypertarget{ref-Unsworth_2016}{}%
+Unsworth, S.
+(2016).
+Quantity and quality of language input in bilingual language development.
+In E.
+Nicoladis \& S.
+Montanari (Eds.), \emph{Bilingualism across the lifespan} (pp.
+103--122).
+Berlin: De Gruyter Mouton.
+
 \leavevmode\hypertarget{ref-Volterra_Taeschner_1978}{}%
-Volterra, V., \& Taeschner, T. (1978). The acquisition and development of language by bilingual children. \emph{Journal of Child Language}, \emph{5}, 311--326.
+Volterra, V., \& Taeschner, T.
+(1978).
+The acquisition and development of language by bilingual children.
+\emph{Journal of Child Language}, \emph{5}, 311--326.
+\url{https://doi.org/10.1017/S0305000900007492}
 
 \leavevmode\hypertarget{ref-Weisleder_Fernald_2013}{}%
-Weisleder, A., \& Fernald, A. (2013). Talking to children matters: Early language experience strengthens processing and builds vocabulary. \emph{Psychological Science}, \emph{24}(11), 2143--2152.
+Weisleder, A., \& Fernald, A.
+(2013).
+Talking to children matters: Early language experience strengthens processing and builds vocabulary.
+\emph{Psychological Science}, \emph{24}(11), 2143--2152.
+\url{https://doi.org/10.1177/0956797613488145}
+
+\leavevmode\hypertarget{ref-White_etal_2020}{}%
+White, A., Malt, B.
+C., Verheyen, S., \& Storms, G.
+(2020).
+Language-specificity in monolingual and bilingual later lexical development.
+\emph{Language Learning and Development}, \emph{16}(4), 309--329.
+\url{https://doi.org/10.1080/15475441.2020.1769632}
 
 \leavevmode\hypertarget{ref-Yu_Smith_2012}{}%
-Yu, C., \& Smith, L. (2012). Embodied attention and word learning by toddlers. \emph{Cognition}, \emph{125}(2), 244--262.
+Yu, C., \& Smith, L. (2012). Embodied attention and word learning by toddlers. \emph{Cognition}, \emph{125}(2), 244--262. \url{https://doi.org/10.1016/j.cognition.2012.06.016}
 

### 2021-12-22 paper/knit to word/paper_TE_bilingual_vocabulary_model_doc.Rmd
updated paper

--- 2021-12-17 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
 West, Montreal, QC, Canada, H2T1V2.
-Current affilitation: Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
+Rachel Ka-Ying Tsui is now at Laboratory for Language Development, RIKEN Center for Brain Science, 2-1 Hirosawa, Wako-shi, Saitama, Japan, 351-0198"
     email         : "rachelkytsui@gmail.com"
@@
   The acquisition of translation equivalents is often considered a special component of bilingual children’s vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
-This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
+This study examined three contrasting accounts for bilingual children’s acquisition of translation equivalents relative to singlets (i.e., words that are first labels for a referent): the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents and singlets are learned similarly.
 To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model — the Bilingual Vocabulary Model — which quantifies translation equivalent knowledge as a function of the probability of learning words in each language, and includes a bias parameter that varies the difficulty of learning translation equivalents according to each account.
@@
 That is, upon hearing the English word “apple,” the corresponding French word _“pomme”_ is more easily activated in bilinguals’ minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)[^1].
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a “singlet”; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent, in a different language, learned following a singlet)[^1].
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn as singlets (Neutral Account).
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents and singlets are learned similarly (Neutral Account).
 To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals’ vocabulary in each language and their developmental level.
@@
 
-[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word _“hygge”_ refers to a type of cozy contentment and enjoyment of simplicity.
+[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word “hygge”, which refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults’ concepts may not be congruent across languages (e.g., Malt & Majid, 2013).
-However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storms et al., 2015; White et al., 2020).
 Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.
@@
 When encountering a potential singlet, mutual exclusivity would be equally useful for bilinguals as it is for monolinguals, supporting them in associating an unlabeled referent with a novel word.
-However, a strong form of mutual exclusivity might prevent bilinguals from associating a translation equivalent word with its referent, given that in this case the referent is already associated with another word (albeit in the other language).
+However, a strong form of mutual exclusivity might prevent bilinguals from associating a translation equivalent with its referent, given that in this case the referent is already associated with another word (albeit in the other language).
 Thus, mutual exclusivity could prevent bilinguals from acquiring translation equivalents, leading to an abundance of singlets in their vocabularies.
@@
 
-### Neutral Account: Translation equivalents are similar to learn as singlets
+### Neutral Account: Translation equivalents and singlets are learned similarly
 The previous accounts rely on the idea that bilingual vocabulary development unfolds differently than monolingual development, as monolinguals encounter only singlets but bilinguals encounter both singlets and translation equivalents.
@@
 Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson & Fernandez, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2014).
-Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
+Bilingual children with greater exposure to a particular language tend to process that language faster, and in turn learn more words in that language.
 
@@
 This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (_n_ = 27) in the context of a wide age range (8 months to 2.5 years).
-In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
+In sum, the Neutral Account predicts that, for bilingual infants, learning translation equivalents is similar to learning singlets.
 
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes & Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011).
-Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
+In general, more language exposure leads to a larger vocabulary size (e.g., Barnes & Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011).
+Bilingual children usually know more words in the language to which they have greater exposure (i.e; dominant language) relative to the language to which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place & Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 Because translation equivalents are words from different languages that refer to the same concept, the number of words a bilingual knows in each of their languages will necessarily constrain the number of translation equivalent pairs they could possibly know.
-For example, a child with a less balanced vocabulary across the two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language.
+For example, a child with a less balanced vocabulary across their two languages might only say 5 words in one language but many more words in the other language; this means that the child could only produce a maximum of 5 translation equivalents, regardless of how many words they know in their other language.
 Conversely, it seems reasonable to expect that if a child knows a similar number of words in each language and thus has a more balanced vocabulary across the two languages, there would be more potential for some of those words to be translation equivalents.
@@
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
-That is, many potential singlets could be “too hard” to be learned at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
-For example, an 14-month-old who has learned the French words _“ballon”_ and _“chien”_ would be capable of learning their translation equivalents “ball” and “dog”, as easier potential singlets like “cup”, but not later-acquired potential singlets like “tomorrow”, “who”, or “not”.
+That is, many potential singlets could be “too hard” to learn at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
+For example, an 14-month-old who has learned the French words _“ballon”_ and _“chien”_ would be capable of learning their translation equivalents “ball” and “dog”, as well as easier potential singlets like “cup”, but not later-acquired potential singlets like "vitamins", “tomorrow”, “who”, or “not”.
 Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
@@
 We formalize learning a translation equivalent pair as the joint probability of learning each of the words in the pair.
-This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
+This provides a straightforward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
 The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent[^2].
@@
 However, not all children will be capable of learning all of the CDI words.
-Typically, words that are more concrete, occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019), and very young children would not be expected to produce many of the “harder” words on the CDI, such as “lawn mower,” “sidewalk,” or “vitamins”.
+Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019).
+That is, very young children would not be expected to produce many of the advanced words on the CDI, such as “lawn mower,” “sidewalk,” or “vitamins”.
 Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
-Since there are no CDI norms for French-English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
+Since there are no CDI norms for French–English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
@@
 Conversely, if translation equivalents are twice as easy to learn as singlets (following the Preference Account), we would expect Jamie to produce 2×(50×20/240) = 8.3 translation equivalents.
-Under the Neutral Account, we would expect Jamie to learn 1×(50×20/240) = 4.2 translation equivalents.
+Under the Neutral Account, we would expect Jamie to produce 1×(50×20/240) = 4.2 translation equivalents.
 
@@
   Variable = c("LEARNABLE", "DOM", "NONDOM", "BIAS", "BALANCE", "WORD", "TE", "CONCEPT", "DOM-SINGLET", "NONDOM-SINGLET"),
-  Definition = c("Number of learnable words in each language, given the child’s developmental level",
+  Definition = c("Number of words expected to be learnable in each language, given the child’s developmental level",
                  "Words produced in the dominant language",
@@
     column_spec(2:4, width = "175px") %>%
-    footnote(general = "All vocabulary measures are constrained to be integers.")
+    footnote(general = "All vocabulary measures are constrained to be integers.",
+             threeparttable = T)
 )
@@
 Second, we assume that translation equivalents refer to the same underlying concept.
-Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt & Majid, 2013; Storm et al., 2015; White et al., 2020).
+Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt & Majid, 2013; Storms et al., 2015; White et al., 2020).
 However, we believe that these assumptions present a reasonable first approximation, especially in the case of young children, and return to these assumptions in the discussion section.
@@
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
-We also compared predicted learning outcomes for when translation equivalents are harder, easier, or similar to learn as singlets.
+We also compared predicted learning outcomes for when translation equivalent learning is harder than, easier than, or similar to singlet learning.
 
 In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French–English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English [@Fenson_etal_2007] and Québec French (Trudeau et al., 1999).
-Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) to learning singlets.
+Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalent learning is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlets.
 
@@
 We present three simulations generated by creating data points for hypothetical children by varying different parameters of the Bilingual Vocabulary Model, in order to explore expected patterns of translation equivalent learning.
-The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except for those that violated constraints in the model (e.g., NONDOM could not be larger than DOM).
+The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except those that violated constraints in the model (e.g., NONDOM could not be larger than DOM).
 In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), and word vocabulary (WORD).
 In the second simulation, we explored relations between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE).
-In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the bias parameter (BIAS) to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts.
+In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the BIAS parameter to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts.
 A summary of the parameter values used in each simulation is provided in Table 2.
@@
     column_spec(8, width = "45px") %>%
-    kable_styling(font_size = 8)
+    kable_styling(font_size = 8) %>%
+    footnote(general = "The number of children simulated is derived from the total number of allowed combinations of the parameters (i.e., excluding combinations that violate parameter constraints; for example NONDOM could not be larger than DOM).",
+             threeparttable = T)
 )
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
-We set BIAS to 1, meaning that in these examples translation equivalents are equally easy to learn as singlets.
-
-We first illustrate with three simulated children.
-Infant Annie (small vocabulary, unbalanced exposure) produces 270 words in the dominant language and 30 words in her non-dominant language.
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English–French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
+We set BIAS to 1, meaning that in these examples translation equivalents and singlets are equally easy to learn.
+
+Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
 She has a word vocabulary of 300, and a balance score of .10 (10% of her words are in the non-dominant language).
 Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents.
-Infant Bernie (small vocabulary, balanced exposure) produces 180 dominant-language words, and 120 non-dominant language words.
+Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
 Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40% of his words are in the non-dominant language).
 Based on our formula, we expect Bernie to produce 36 translation equivalents.
-Comparing Annie and Bernie, two children who produce the same word vocabulary (i.e., WORD is held constant), the child with more balanced language vocabulary (Bernie) is expected to produce more translation equivalents.
-Like Bernie, infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM) for a total of 900 words (WORD), and thus BALANCE = .40.
+Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
+Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
 Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
@@
 We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
-BIAS was once again constant at 1.
+BIAS remained constant at 1.
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
@@
 Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as the number of DOM (dominant language words) and WORD (word vocabulary) increased, TE also increased regardless of BALANCE.
+Moreover, Panels 1A and 1C show that, as DOM (dominant language words) and WORD (word vocabulary) increased, TE (translation equivalents) also increased regardless of BALANCE.
 Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
@@
       caption = "Table 3.
-Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
 Vocabulary values are rounded to the nearest integer.") %>%
@@
         legend.position = "bottom",
+        legend.key.width=unit(1.1,"cm"),
         plot.title = element_text(face = 'bold', size = 12, hjust = 0, vjust = 2),
@@
         legend.position = "bottom",
+        legend.key.width=unit(1.1,"cm"),
         panel.spacing.x = unit(2, "lines"),
@@
 
-Translation equivalent knowledge was simulated across children at six developmental levels (the number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
+Translation equivalent knowledge was simulated across children at six developmental levels (number of LEARNABLE words = 100, 200, 300, 400, 500, 600), in conjunction with a wide range of values for words in the dominant language (DOM) and the non-dominant language (NONDOM).
 In total, data from 301 simulated children were generated (see Table 2 for a summary of the parameters used in this simulation).
@@
 Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages, and that more balanced children would produce more singlets in their non-dominant language, but fewer singlets in their dominant language than less balanced children.
+Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages, and that more balanced children produced more singlets in their non-dominant language, but fewer singlets in their dominant language than less balanced children.
 Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
@@
   theme(legend.position = "bottom",
+        legend.key.width=unit(1.1,"cm"),
         plot.title = element_text(face = 'bold', size = 10, hjust = 0, vjust = 2),
@@
   labs(x = "\nNumber of learnable words (LEARNABLE)", 
-       y = "Number of concepts (CONCEPT)", 
+       y = "Number produced", 
        color = "Vocabulary balance (BALANCE)",
@@
   labs(x = "\nAge (in months)", 
-       y = "Number of concepts (CONCEPT)", 
+       y = "Number produced", 
        color = "Vocabulary balance (BALANCE)",
@@
 
-```{r fig2, fig.cap="Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+```{r fig2, fig.cap="Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which set the number of LEARNABLE words.
 Row 1 represents the simulated data in Study 1.
@@
 
-Translation equivalent (TE) knowledge was first simulated across different developmental levels (as indicated by number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
+Translation equivalent (TE) knowledge was first simulated across different developmental levels (number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
 Again, BALANCE and word vocabulary (WORD) were calculated based on the values of DOM and NONDOM.
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equally learnable as singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs and singlets are equally learnable).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50% more likely to be learned than singlets).
@@
 
-* Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+* Prediction 3: Whether learning translation equivalents is harder than, easier than, or similar to learning singlets, the slope of translation equivalent learning will change as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents and singlets are similar to learn (i.e., Neutral Account).
 
@@
   theme(axis.text.x = element_text(angle = 90),
-        legend.position = "bottom") +
-  theme(plot.title = element_text(hjust = 0.5),
+        legend.position = "bottom",
+        legend.key.width=unit(1.1,"cm"),
+        plot.title = element_text(hjust = 0.5),
         text = element_text(size=11),
@@
 
-```{r fig3, fig.cap="Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%", fig.pos = "H", cache = TRUE}
+```{r fig3, fig.cap="Different scenarios of expected translation equivalent learning (TE) as a function of WORD vocabulary, under scenarios where learning TEs is harder than (BIAS < 1), easier than (BIAS > 1), or similar to (BIAS = 1) learning singlets.", echo=FALSE, dpi=600, fig.align='center', fig.height=3.5, out.height="80%", out.width="100%", fig.pos = "H", cache = TRUE}
 
@@
 ## 1.4 Summary of model predictions
-Study 1 presented predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Study 1 generated predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
-Prediction Set 1 illustrated that the more words children produce in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they are predicted to produce.
+Prediction Set 1 illustrated that the more words children produce in total word vocabulary, in their dominant language, or in their non-dominant language, the more translation equivalents they are predicted to produce.
 Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) are also expected to produce more translation equivalents.
 Prediction Set 2 described translation equivalent knowledge in relation to the number of learnable words constrained by a child’s developmental level.
-Older children are predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Older children are predicted to have a larger concept vocabulary (i.e., to have lexicalized more concepts), including translation equivalents and singlets in both dominant and non-dominant languages, than younger children.
 Moreover, the most balanced children are expected to produce more singlets in their non-dominant language, whereas the least balanced children are expected to produce more singlets in their dominant language.
@@
 Crucially, these predictions should not be thought of as fully independent from each other, but instead illustrate the interrelatedness of measures of bilingual vocabulary under the Bilingual Vocabulary Model.
-In Study 2, we used vocabulary data from bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
+In Study 2, we used vocabulary data from real bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
 
 # Study 2: Empirical data
-In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
+In Study 1, we used simulations based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
 In Study 2, we tested these predictions using archival vocabulary data from 200 French–English bilingual children aged 18 to 33 months.
@@
   group_by(gender) %>% 
-  distinct(baby_id, .keep.all=T) 
+  distinct(baby_id, .keep.all = T) 
   # %>% count()
@@
 
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),1)` - `r round(max(keepers_ws_TE$age_continuous),1)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
+Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range = `r round(min(keepers_ws_TE$age_continuous),1)` – `r round(max(keepers_ws_TE$age_continuous),1)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
 (2020).
@@
 Caregivers completed the original CDI English version [@Fenson_etal_2007] and its Québec French adaptation (Trudeau et al., 1999).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), others (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), or else the respondent was not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
+We asked the caregiver more familiar with each language to complete the respective CDI form.
+The forms were mainly filled out by mothers (`r round(d_cdi_filled_by$percentage[2],0)`%), fathers (`r round(d_cdi_filled_by$percentage[1],0)`%), both parents (`r round(d_cdi_filled_by$percentage[3],0)`%), other caregivers (< `r round(d_cdi_filled_by$percentage[4],0)`%; e.g., grandmother), or else the respondent was not indicated (`r round(d_cdi_filled_by$percentage[5],0)`%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
@@
 In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
-Words that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were also considered translation equivalents.
+Words with the same meaning that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were considered translation equivalents.
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
@@
 
-After determining the dominant language of a child based on the vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
+After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
 Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).
@@
 On average, bilinguals in the sample had a mean word vocabulary size (WORD) of `r round(vocabulary_average$mean[7],1)` (SD = `r round(vocabulary_average$sd[7],1)`), with a wide range of `r vocabulary_average$min[7]` - `r vocabulary_average$max[7]` words.
-As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = `r round(vocabulary_average$mean[5],1)`, SD = `r round(vocabulary_average$sd[5],1)`, range = `r vocabulary_average$min[5]` - `r vocabulary_average$max[5]`) than in their non-dominant language (NONDOM; M = `r round(vocabulary_average$mean[6],1)`, SD = `r round(vocabulary_average$sd[6],1)`, range = `r vocabulary_average$min[6]` - `r vocabulary_average$max[6]`), _t_(`r t_test_DOM_NONDOM$parameter[1]`) = `r t_test_DOM_NONDOM$statistic[1]`, $p `r papaja::printp(t_test_DOM_NONDOM$p.value[1])`$, $d = `r round(d_DOM_NONDOM, 2)`$.
-
-Children produced an average of `r round(vocabulary_average$mean[2],1)` translation equivalents (TE; SD = `r round(vocabulary_average$sd[2],1)`, range = `r vocabulary_average$min[2]` - `r vocabulary_average$max[2]`).
-The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = `r round(vocabulary_average$mean[3],1)`, SD = `r round(vocabulary_average$sd[3],1)`, range = `r vocabulary_average$min[3]` - `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r round(vocabulary_average$mean[4],1)`, SD = `r round(vocabulary_average$sd[4],1)`, range = `r vocabulary_average$min[4]` - `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
-On average, children’s concept vocabulary size was `r round(vocabulary_average$mean[1],1)` words (CONCEPT; SD = `r round(vocabulary_average$sd[1],1)`, range = `r vocabulary_average$min[1]` - `r vocabulary_average$max[1]`).
+As expected by the way language dominance was defined, children produced more words in their dominant language (DOM; M = `r round(vocabulary_average$mean[5],1)`, SD = `r round(vocabulary_average$sd[5],1)`, range = `r vocabulary_average$min[5]` – `r vocabulary_average$max[5]`) than in their non-dominant language (NONDOM; M = `r round(vocabulary_average$mean[6],1)`, SD = `r round(vocabulary_average$sd[6],1)`, range = `r vocabulary_average$min[6]` – `r vocabulary_average$max[6]`), _t_(`r t_test_DOM_NONDOM$parameter[1]`) = `r t_test_DOM_NONDOM$statistic[1]`, $p `r papaja::printp(t_test_DOM_NONDOM$p.value[1])`$, $d = `r round(d_DOM_NONDOM, 2)`$.
+
+Children produced an average of `r round(vocabulary_average$mean[2],1)` translation equivalents (TE; SD = `r round(vocabulary_average$sd[2],1)`, range = `r vocabulary_average$min[2]` – `r vocabulary_average$max[2]`).
+The remainder of words were singlets, and children produced many more singlets in their dominant language (DOM-SINGLET; M = `r round(vocabulary_average$mean[3],1)`, SD = `r round(vocabulary_average$sd[3],1)`, range = `r vocabulary_average$min[3]` – `r vocabulary_average$max[3]`) than in their non-dominant language (NONDOM-SINGLET; M = `r round(vocabulary_average$mean[4],1)`, SD = `r round(vocabulary_average$sd[4],1)`, range = `r vocabulary_average$min[4]` – `r vocabulary_average$max[4]`), _t_(`r t_test_SINGLET$parameter[1]`) = `r t_test_SINGLET$statistic[1]`, $p `r papaja::printp(t_test_SINGLET$p.value[1])`$, $d = `r round(d_SINGLET, 2)`$.
+On average, children’s concept vocabulary size was `r round(vocabulary_average$mean[1],1)` words (CONCEPT; SD = `r round(vocabulary_average$sd[1],1)`, range = `r vocabulary_average$min[1]` – `r vocabulary_average$max[1]`).
 
@@
 Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
-The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
+The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` – `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` – `r round(balance_average_group$max[2], 2)`).
 
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., in press, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson’s correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations[^3]).
-
-[^3]: Note that there was a negative correlation between BALANCE and age in our sample.
-We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction Set 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations[^3]).
+
+[^3]: Note that there was a negative correlation between BALANCE and age in our sample, such that older children had vocabularies that were less equal across their languages.
+This is not predicted by our model, and we can only speculate as to why this pattern was observed.
+For example, older children may spend more time in childcare where mostly one language is used, therefore leading to a less balanced vocabulary learning.
+Another possibility is that older children increasingly elicit speech in their dominant language, which in turn amplifies vocabulary imbalance.
+Testing these possibilities is beyond the scope of the current paper.
 
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): _r_(`r cor_TE_word$parameter[1]`) = `r round(cor_TE_word$estimate[1],2)`, $p `r papaja::printp(cor_TE_word$p.value[1])`$, and dominant-language vocabulary (DOM): _r_(`r cor_TE_dom$parameter[1]`) = `r round(cor_TE_dom$estimate[1],2)`, $p `r papaja::printp(cor_TE_dom$p.value[1])`$.
-Figure 1 Panels 2A and 2B further illustrates these relationships observed in our dataset.
-
-Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
+Figure 1 Panels 2A and 2B further illustrate these relationships observed in our dataset.
+
+Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE) — specifically, that this relationship would be nearly perfect.
 As shown in Figure 1 Panel 2C, we observed that these two variables were indeed nearly perfectly correlated, _r_(`r cor_TE_nondom$parameter[1]`) = `r round(cor_TE_nondom$estimate[1],2)`, $p `r papaja::printp(cor_TE_nondom$p.value[1])`$.
@@
 In our data set, developmental level was approximated by children’s age.
-Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (our proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET), and singlets in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of age (our proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 ## 2.2.4 Testing Prediction Set 3: Rate of translation equivalent learning
-Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
+Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, and the Preference Account.
 To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE, and we allowed the model to estimate the BIAS parameter.
@@
 First, we will walk through the parameters in this model.
-The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
+Dominant vocabulary size (DOM) and non-dominant vocabulary size (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
+Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary norms at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
 Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was `r ws_percentile$"Average (LEARNABLE)"[1]` words which was calculated by averaging the `r ws_percentile$"Number of English words produced at 90th percentile"[1]` English words and `r ws_percentile$"Number of French words produced at 90th percentile"[1]` French words, based on what 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
-Finally, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
-
+For example, for an 18-month-old infant, the denominator was `r ws_percentile$"Average (LEARNABLE)"[1]` words, which was calculated by averaging the `r ws_percentile$"Number of English words produced at 90th percentile"[1]` English words and `r ws_percentile$"Number of French words produced at 90th percentile"[1]` French words that 18-month-old children would typically produce at the 90th percentile.
+For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since 90th percentile information was available only up to 30 months.
+Finally, the intercept of the linear regression model was set at 0, since no translation equivalents are expected to be produced if a child does not have any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
+
+&nbsp;
 ```{r table6, results = "asis", cache = TRUE}
@@
         caption = "Table 6.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
+The number of total English and French productive CDI vocabulary words at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
 Values are rounded to the nearest integer.") %>%
@@
   column_spec(1, width = "75px") %>%
-  column_spec(2:4, width = "125px")
+  column_spec(2:4, width = "125px") %>%
+  footnote(general = "90th percentile information was only available up to age 30 months from WordBank, so we used that value for all children in our dataset aged 30 months and older.",
+           threeparttable = T)
+
 ```
@@
 
-With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model indicates how the BIAS parameter is related to the empirical vocabulary data, which in turn indicates whether bilingual children were biased towards or against learning translation equivalents.
-If the coefficient is close to 1, then there is no bias and translation equivalents are learned equally to other words (i.e., the Neutral Account).
+With the observed number of translation equivalents as the dependent variable, the regression coefficient estimated by the model indicates how the BIAS parameter relates to the empirical vocabulary data, which in turn indicates whether bilingual children were biased towards or against learning translation equivalents.
+If the coefficient is close to 1, then there is no bias and translation equivalents and singlets are learned similarly (i.e., the Neutral Account).
 Otherwise, a coefficient less than 1 indicates a bias against learning translation equivalents where translation equivalents are less easily learned (i.e., the Avoidance Account), and a coefficient greater than 1 indicates a bias towards learning translation equivalents where translation equivalents are more easily learned (i.e., the Preference Account).
@@
 Our model showed an excellent model fit of $R^2$ = `r round(summary(test_prediction3_model)$r.squared,2)`, indicating that our model explained `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of the variance in bilinguals’ translation equivalent knowledge.
-BIAS coefficient was estimated at `r round(summary(test_prediction3_model)$coefficient[1],2)`.
-This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words[^5].
+The BIAS coefficient was estimated at `r round(summary(test_prediction3_model)$coefficient[1],2)`.
+This value is close to 1, suggesting that our data are consistent with the Neutral account, whereby translation equivalents are learned equivalently to other words[^5].
 
 [^5]: We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
-Overall, model fit was slightly reduced compared to our main model, although fit was still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`.
+Overall, model fit was slightly reduced compared to our main model, although still good, $R^2$ = `r round(summary(test_prediction3_model_75percentile)$r.squared,2)`.
 The estimated BIAS parameter was lower, at `r round(summary(test_prediction3_model_75percentile)$coefficient[1],2)`, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
@@
 Note that visual inspection suggested that there could be some possible outliers.
-Cook’s distance was estimated for our linear regression model listed above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a cook’s distance over 0.4.
+Cook’s distance was estimated for our linear regression model described above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a cook’s distance over 0.4.
 After removing those `r words(length(number_of_outliers$ID_testdate))` data points, the linear regression model returned a coefficient of `r round(summary(model_remove_outliers)$coefficient[1],2)`, with $R^2$ = `r round(summary(model_remove_outliers)$r.squared,2)`.
@@
 Figure 6 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
-Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.
-To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
-The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_more300w)$coefficient[1],2)`, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_less300w)$coefficient[1],2)`.
-Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size ($R^2$ = `r round(summary(percentile_model_more300w)$r.squared,2)`) than children with smaller vocabulary size ($R^2$ = `r round(summary(percentile_model_less300w)$r.squared,2)`).
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 word vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS > 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 word vocabulary.
+To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 word vocabulary and for those with more than 300 word vocabulary.
+The model for those with a larger word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_more300w)$coefficient[1],2)`, whereas the model for those with a smaller word vocabulary (WORD) returned a coefficient of BIAS = `r round(summary(percentile_model_less300w)$coefficient[1],2)`.
+Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary sizes ($R^2$ = `r round(summary(percentile_model_more300w)$r.squared,2)`) than children with smaller vocabulary sizes ($R^2$ = `r round(summary(percentile_model_less300w)$r.squared,2)`).
 Overall, this analysis suggests that translation equivalent learning for children with larger vocabularies corresponds best to the Neutral Account, but translation equivalent learning for children with smaller vocabularies corresponds best to the Preference Account.
@@
 &nbsp;
-```{r fig6, fig.cap="The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
+```{r fig6, fig.cap="The number of observed translation equivalents as a function of the number of expected translation equivalents under the Bilingual Vocabulary Model (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
 The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.", echo=FALSE, dpi=600, fig.align='center', fig.height=4, out.width="120%", cache = TRUE}
@@
 
+```{r include=FALSE}
+# Log (as suggested by a reviewer to log-tranform the predicted and observed TEs in testing Prediction 3; yet, we observed the same patterns in the log-transformed model as in the previous models.
+So we kept the previous models)
+
+## Overall model
+test_prediction3_model_log <- keepers_ws_TE_simulation %>%
+  mutate(predicted_TE_percentile = (total_words_dom*total_words_nondom)/Average_90percentile,
+         log_number_of_te = log(number_of_te),
+         log_predicted_TE_percentile = log(predicted_TE_percentile)) %>%
+  lm(log_number_of_te ~ 0 + log_predicted_TE_percentile, data = .)
+
+summary(test_prediction3_model_log)
+
+## Those with > 300 word vocab
+older_test_prediction3_model_log <- keepers_ws_TE_simulation %>%
+  filter(word_vocab > 300) %>%
+  mutate(predicted_TE_percentile = (total_words_dom*total_words_nondom)/Average_90percentile,
+         log_number_of_te = log(number_of_te),
+         log_predicted_TE_percentile = log(predicted_TE_percentile)) %>%
+  lm(log_number_of_te ~ 0 + log_predicted_TE_percentile, data = .)
+
+summary(older_test_prediction3_model_log)
+
+## Those with < 300 word vocab
+younger_test_prediction3_model_log <- keepers_ws_TE_simulation %>%
+  filter(word_vocab < 300) %>%
+  mutate(predicted_TE_percentile = (total_words_dom*total_words_nondom)/Average_90percentile,
+         log_number_of_te = log(number_of_te),
+         log_predicted_TE_percentile = log(predicted_TE_percentile)) %>%
+  lm(log_number_of_te ~ 0 + log_predicted_TE_percentile, data = .)
+
+summary(younger_test_prediction3_model_log)
+
+## Plot
+log <- keepers_ws_TE_simulation %>%
+  mutate(predicted_TE_percentile = (total_words_dom*total_words_nondom)/Average_90percentile,
+         log_number_of_te = log(number_of_te),
+         log_Predicted_TE_percentile = log(predicted_TE_percentile)) %>%
+#  pivot_longer(log_Predicted_TE_75percentile:log_Predicted_TE_90percentile, names_to = "Percentile", values_to = "log_Predicted_TE") %>%
+  ggplot(aes(x = log_Predicted_TE_percentile, y = log_number_of_te)) +
+#  facet_grid(. ~ Percentile) +
+  labs(x = "Log-transformed number of TEs \n predicted by the Bilingual Vocabulary Model", y = "Log-transformed observed total number of TEs") +
+  theme(plot.title = element_text(hjust = 0.5)) +
+  #geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
+  geom_point(shape = 23) +
+  geom_abline(slope = 1, linetype = 2, alpha = 0.5) +
+  xlim(-6,6) +
+  ylim(-6,6) +
+  theme_minimal() + 
+  theme(text = element_text(size=14))
+
+ggsave(here("Output/log_prediction.png"), log,
+       height=5,width=5)
+
+```
+
 # Discussion
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn as singlet words (i.e., the first label for a particular referent).
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether learning translation equivalents is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words (i.e., the first label for a particular referent).
 To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn given the child’s age.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn, given a child’s age.
 The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
@@
 ## Confirmation of model predictions
-In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18-33 months.
-Three sets of model predictions were confirmed in our empirical dataset.
-In Table 7, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18–33 months.
+All three sets of model predictions were confirmed in our empirical dataset.
+In Table 7, we summarize the model predictions along with the evidence from our empirical dataset and other findings in the literature.
 
@@
     "The number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
-\n There is a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013)",
+\n There is a positive correlation between bilingual children’s dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013)",
     "Vocabulary size in the non-dominant language was positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).",
@@
 Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children’s dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
@@
 Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents are similar to learn as singlets), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
-Modeling their data separately, we found evidence for the Preference Account: younger children at around `r round(keepers_ws_TE %>% filter(word_vocab < 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months appeared to learn translation equivalents more easily than singlets, whereas older children at around `r round(keepers_ws_TE %>% filter(word_vocab > 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months learned translation equivalents similarly to singlets.
+Modeling their data separately, we found evidence for the Preference Account: younger children at around `r round(keepers_ws_TE %>% filter(word_vocab < 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months old appeared to learn translation equivalents more easily than singlets, whereas older children at around `r round(keepers_ws_TE %>% filter(word_vocab > 300) %>% summarise(mean_age = mean(age_continuous, na.rm = T)),0)` months old learned translation equivalents similarly to singlets.
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
@@
 However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
-For example, a French–English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., “soap,” “bath,” “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquired across languages and cultures.
+For example, a French–English bilingual child who always spends bathtime with an English-speaking caregiver might encounter and use bath words primarily in English (e.g., “soap,” “bath,” “bubbles”), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produce (e.g., Braginsky et al., 2019; Tardif et al., 2008).
+For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquire across languages and cultures.
 It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
@@
 
-Another consideration is that our model is that it takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
+Another consideration is that our model takes a somewhat simplified view of translation equivalents, assuming that children encounter the same conceptual categories in each of their languages and are exposed to the corresponding words.
 However, the reality of bilingual experience might be more complex.
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a “bowl” in English but an _“assiette”_ [plate] in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storm et al., 2015; White et al., 2020).
-Therefore, it seems reasonable to assume that early vocabulary may be represented similarly across languages, especially for the age group we were looking at.
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storms et al., 2015; White et al., 2020).
+Therefore, it seems reasonable to assume that early vocabulary may be represented similarly across languages, especially for the age group we looked at in this study.
 Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where concepts are less well aligned across languages.

### 2021-12-22 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated paper

@@
 However, not all children will be capable of learning all of the CDI words.
-Typically, words that are more concrete, occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019), and very young children would not be expected to produce many of the “harder” words on the CDI, such as “lawn mower,” “sidewalk,” or “vitamins”.
+Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019).
+That is, very young children would not be expected to produce many of the advanced words on the CDI, such as “lawn mower,” “sidewalk,” or “vitamins”.
 Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
-Since there are no CDI norms for French-English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
+Since there are no CDI norms for French—English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
@@
 
-Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),1)` - `r round(max(keepers_ws_TE$age_continuous),1)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
+Archival data from `r length(unique(keepers_ws_TE$baby_id))` bilingual children acquiring English and French (age range: `r round(min(keepers_ws_TE$age_continuous),1)` – `r round(max(keepers_ws_TE$age_continuous),1)` months; `r length(d_gender$gender[d_gender$gender == "F"])` girls and `r length(d_gender$gender[d_gender$gender == "M"])` boys) who participated in prior studies at the Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.
 (2020).
@@
 In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
-Words that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were also considered translation equivalents.
+Words with the same meaning that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were also considered translation equivalents.
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.

### 2021-12-22 paper/paper_TE_bilingual_vocabulary_model.tex
updated paper

@@
 The acquisition of translation equivalents is often considered a special component of bilingual children's vocabulary development, as bilinguals have to learn words that share the same meaning across their two languages.
-This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to words that are first labels for a referent: the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents are similar to learn.
+This study examined three contrasting accounts for bilingual children's acquisition of translation equivalents relative to singlets (i.e., words that are first labels for a referent): the Avoidance Account whereby translation equivalents are harder to learn, the Preference Account whereby translation equivalents are easier to learn, and the Neutral Account whereby translation equivalents and singlets are learned similarly.
 To adjudicate between these accounts, Study 1 explored patterns of translation equivalent learning under a novel computational model --- the Bilingual Vocabulary Model --- which quantifies translation equivalent knowledge as a function of the probability of learning words in each language, and includes a bias parameter that varies the difficulty of learning translation equivalents according to each account.
@@
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''} refers to a type of cozy contentment and enjoyment of simplicity.
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent, in a different language, learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word ``hygge,'' which refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
-However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storm et al., 2020; White et al., 2020).
+However, in early development, vocabulary words tend to be concrete, and evidence suggests that concepts are largely shared across languages (Storms et al., 2015; White et al., 2020).
 Thus, we make the assumption that most early words have a translation equivalent and that these are fairly congruent for the age range of focus in the current paper.}.
 Is translation equivalent learning different from singlet learning?
-The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents are similar to learn as singlets (Neutral Account).
+The current paper contrasts three competing accounts: 1) translation equivalents are harder to learn than singlets (Avoidance Account), 2) translation equivalents are easier to learn than singlets (Preference Account), and 3) translation equivalents and singlets are learned similarly (Neutral Account).
 To adjudicate between these accounts, we introduce the Bilingual Vocabulary Model, which provides a computational account of vocabulary learning, with parameters including bilinguals' vocabulary in each language and their developmental level.
@@
 
-\hypertarget{neutral-account-translation-equivalents-are-similar-to-learn-as-singlets}{%
-\subsubsection{Neutral Account: Translation equivalents are similar to learn as singlets}\label{neutral-account-translation-equivalents-are-similar-to-learn-as-singlets}}
+\hypertarget{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}{%
+\subsubsection{Neutral Account: Translation equivalents and singlets are learned similarly}\label{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}}
 
@@
 Due to differences in the amount of language exposure, bilingual children seldom show equal vocabulary growth in both of their languages (e.g., Pearson \& Fernandez, 1994; Pearson et al., 1997), and the amount of exposure to a particular language has been reported to modulate the within-language association between language processing ability and vocabulary size (Hurtado et al., 2014).
-Bilingual children with greater exposure to a particular language tended to process that language faster, and in turn learned more words in that language.
+Bilingual children with greater exposure to a particular language tend to process that language faster, and in turn learn more words in that language.
 
@@
 This evidence implied that singlets and translation equivalents are equally learnable, although the paper was limited by its somewhat small sample (\emph{n} = 27) in the context of a wide age range (8 months to 2.5 years).
-In sum, the Neutral Account predicts that translation equivalents are similar for bilingual children to learn as singlets.
+In sum, the Neutral Account predicts that, for bilingual infants, learning translation equivalents is similar to learning singlets.
 
@@
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
-In general, more language exposure leads to larger vocabulary size (e.g., Barnes \& Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011).
-Bilingual children usually know more words in the language in which they have greater exposure (i.e; dominant language) relative to the language in which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
+In general, more language exposure leads to a larger vocabulary size (e.g., Barnes \& Garcia, 2013; Boyce et al., 2013; Hurtado et al., 2014; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011).
+Bilingual children usually know more words in the language to which they have greater exposure (i.e; dominant language) relative to the language to which they have less exposure (i.e., non-dominant language; Pearson et al., 1997; Place \& Hoff, 2011).
 This is because the more often a bilingual hears a language, the more opportunities there will be for learning new words in that language.
@@
 If this is the case, then children might be more likely to learn translation equivalents than singlets, simply because translation equivalents are more likely to be learnable at their stage of development.
-That is, many potential singlets could be ``too hard'' to be learned at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
-For example, an 14-month-old who has learned the French words \emph{``ballon''} and \emph{``chien''} would be capable of learning their translation equivalents ``ball'' and ``dog,'' as easier potential singlets like ``cup,'' but not later-acquired potential singlets like ``tomorrow,'' ``who,'' or ``not.'' Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
+That is, many potential singlets could be ``too hard'' to learn at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
+For example, an 14-month-old who has learned the French words \emph{``ballon''} and \emph{``chien''} would be capable of learning their translation equivalents ``ball'' and ``dog,'' as well as easier potential singlets like ``cup,'' but not later-acquired potential singlets like ``vitamins,'' ``tomorrow,'' ``who,'' or ``not.'' Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
 
@@
 We formalize learning a translation equivalent pair as the joint probability of learning each of the words in the pair.
-This provides a straight-forward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
+This provides a straightforward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
 The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent\footnote{To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
@@
 However, not all children will be capable of learning all of the CDI words.
-Typically, words that are more concrete, occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019), and very young children would not be expected to produce many of the ``harder'' words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
-Since there are no CDI norms for French-English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
+Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019).
+That is, very young children would not be expected to produce many of the advanced words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
+Since there are no CDI norms for French---English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
@@
 \fontsize{7}{9}\selectfont
+\begin{threeparttable}
 \begin{tabular}[t]{l>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}>{\raggedright\arraybackslash}p{175px}}
@@
 \multicolumn{4}{l}{\textbf{Main Parameters}}\\
-\hspace{1em}LEARNABLE & Number of learnable words in each language, given the child’s developmental level & Varies by age.
+\hspace{1em}LEARNABLE & Number of words expected to be learnable in each language, given the child’s developmental level & Varies by age.
 No greater than the number of words on CDI.
@@
 \bottomrule
-\multicolumn{4}{l}{\rule{0pt}{1em}\textit{Note: }}\\
-\multicolumn{4}{l}{\rule{0pt}{1em}All vocabulary measures are constrained to be integers.}\\
 \end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item All vocabulary measures are constrained to be integers.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 Second, we assume that translation equivalents refer to the same underlying concept.
-Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt \& Majid, 2013; Storm et al., 2015; White et al., 2020).
+Thus, our model does not consider the nuanced and language-specific conceptual representations that bilingual adults might ultimately develop (Malt \& Majid, 2013; Storms et al., 2015; White et al., 2020).
 However, we believe that these assumptions present a reasonable first approximation, especially in the case of young children, and return to these assumptions in the discussion section.
@@
 Study 1 simulated the expected patterns of translation equivalent learning under the Bilingual Vocabulary Model proposed in the introduction, with reference to the proportion of words learned in the dominant and non-dominant language and the number of words that are learnable at various developmental levels.
-We also compared predicted learning outcomes for when translation equivalents are harder, easier, or similar to learn as singlets.
+We also compared predicted learning outcomes for when translation equivalent learning is harder than, easier than, or similar to singlet learning.
 
 In Study 2, we examined real-world translation equivalent development in light of the predictions from the Bilingual Vocabulary Model, using archival data from 200 French--English bilingual children aged 18 to 33 months, whose vocabularies and translation equivalent knowledge were measured by parent report using the MacArthur-Bates CDI: Words and Sentences form in English (Fenson et al., 2007) and Québec French (Trudeau et al., 1999).
-Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalents learning is harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) to learning singlets.
+Together, the Bilingual Vocabulary Model and real-world data allowed us to examine contrasting hypotheses about translation equivalents: whether translation equivalent learning is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlets.
 
@@
 \fontsize{8}{10}\selectfont
+\begin{threeparttable}
 \begin{tabular}[t]{>{\centering\arraybackslash}p{45px}>{\raggedright\arraybackslash}p{45px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{75px}>{\raggedright\arraybackslash}p{45px}}
@@
 \end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item The number of children simulated is derived from the total number of allowed combinations of the parameters (i.e., excluding combinations that violate parameter constraints; for example NONDOM could not be larger than DOM).
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English-French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
-We set BIAS to 1, meaning that in these examples translation equivalents are equally easy to learn as singlets.
-
-We first illustrate with three simulated children.
-Infant Annie (small vocabulary, unbalanced exposure) produces 270 words in the dominant language and 30 words in her non-dominant language.
+For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English--French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
+We set BIAS to 1, meaning that in these examples translation equivalents and singlets are equally easy to learn.
+
+Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
 She has a word vocabulary of 300, and a balance score of .10 (10\% of her words are in the non-dominant language).
 Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents.
-Infant Bernie (small vocabulary, balanced exposure) produces 180 dominant-language words, and 120 non-dominant language words.
+Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
 Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40\% of his words are in the non-dominant language).
 Based on our formula, we expect Bernie to produce 36 translation equivalents.
-Comparing Annie and Bernie, two children who produce the same word vocabulary (i.e., WORD is held constant), the child with more balanced language vocabulary (Bernie) is expected to produce more translation equivalents.
-Like Bernie, infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM) for a total of 900 words (WORD), and thus BALANCE = .40.
+Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
+Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
 Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
@@
 \caption{\label{tab:table3}Table 3.
-Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balance (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
 Vocabulary values are rounded to the nearest integer.}
@@
 
-\begin{verbatim}
-## Warning: Removed 39 rows containing non-finite values (stat_smooth).
-\end{verbatim}
-
-\begin{verbatim}
-## Warning: Removed 1 rows containing missing values (geom_smooth).
-\end{verbatim}
-
 \begin{figure}[H]
@@
 
-\caption{Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which sets the number of LEARNABLE words.
+\caption{Number of concept vocabulary (CONCEPT; Panel A), translation equivalents (TE; Panel B), singlets in dominant (DOM-SINGLET; Panel C) and singlets in the non-dominant language (NONDOM-SINGLET; Panel D) across different levels of vocabulary balance (BALANCE) in relation to different developmental levels/ages, which set the number of LEARNABLE words.
 Row 1 represents the simulated data in Study 1.
@@
 To illustrate the Avoidance Account, BIAS was set at .5 (i.e., TEs are 50\% less likely to be learned than singlets).
-To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs are equally learnable as singlets).
+To illustrate the Neutral Account, BIAS was set at 1 (i.e., TEs and singlets are equally learnable).
 Finally, to illustrate the Preference Account, BIAS was set at 1.5 (i.e., TE are 50\% more likely to be learned than singlets).
@@
 \item
-  Prediction 3: Whether translation equivalents are harder to learn, easier to learn, or similar to learn as singlets will change the slope of translation equivalent learning as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents are similar to learn as singlets (i.e., Neutral Account).
+  Prediction 3: Whether learning translation equivalents is harder than, easier than, or similar to learning singlets, the slope of translation equivalent learning will change as a function of word vocabulary (WORD), with a shallower slope if TEs are less easily learned (i.e., Avoidance Account), and a steeper slope if TEs are more easily learned (i.e., Preference Account) compared to where translation equivalents and singlets are similar to learn (i.e., Neutral Account).
 \end{itemize}
@@
 
-\caption{Different scenarios of expected translation equivalents learning (TE) as a function of WORD vocabulary, under scenarios where TEs are harder to learn (BIAS < 1), easier to learn (BIAS > 1), or similar to learn (BIAS = 1) as singlets.}\label{fig:fig3}
+\caption{Different scenarios of expected translation equivalent learning (TE) as a function of WORD vocabulary, under scenarios where learning TEs is harder than (BIAS < 1), easier than (BIAS > 1), or similar to (BIAS = 1) learning singlets.}\label{fig:fig3}
 \end{figure}
@@
 
-Study 1 presented predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+Study 1 generated predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
-Prediction Set 1 illustrated that the more words children produce in total vocabulary, their dominant language, or the non-dominant language, the more translation equivalents they are predicted to produce.
+Prediction Set 1 illustrated that the more words children produce in total word vocabulary, in their dominant language, or in their non-dominant language, the more translation equivalents they are predicted to produce.
 Moreover, children with more balanced vocabularies (i.e., producing a similar number of words in each of their languages) are also expected to produce more translation equivalents.
 Prediction Set 2 described translation equivalent knowledge in relation to the number of learnable words constrained by a child's developmental level.
-Older children are predicted to produce more concept vocabularies including translation equivalents and singlets in both dominant and non-dominant languages than younger children.
+Older children are predicted to have a larger concept vocabulary (i.e., to have lexicalized more concepts), including translation equivalents and singlets in both dominant and non-dominant languages, than younger children.
 Moreover, the most balanced children are expected to produce more singlets in their non-dominant language, whereas the least balanced children are expected to produce more singlets in their dominant language.
@@
 Crucially, these predictions should not be thought of as fully independent from each other, but instead illustrate the interrelatedness of measures of bilingual vocabulary under the Bilingual Vocabulary Model.
-In Study 2, we used vocabulary data from bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
+In Study 2, we used vocabulary data from real bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
 
@@
 
-In Study 1, we used a simulation based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
+In Study 1, we used simulations based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
 In Study 2, we tested these predictions using archival vocabulary data from 200 French--English bilingual children aged 18 to 33 months.
@@
 
-Archival data from 200 bilingual children acquiring English and French (age range: 18.40 - 33.50 months; 94 girls and 106 boys) who participated in prior studies at Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
+Archival data from 200 bilingual children acquiring English and French (age range: 18.40 -- 33.50 months; 94 girls and 106 boys) who participated in prior studies at the Concordia Infant Research Lab were included in the present study, drawn from the same set of participants as Gonzalez-Barrero et al.~(2020).
 Data collection was conducted in Montréal, Québec, Canada.
@@
 Caregivers completed the original CDI English version (Fenson et al., 2007) and its Québec French adaptation (Trudeau et al., 1999).
-We asked the caregiver more familiar with each language to complete the respective CDI form, and the forms are mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), others (\textless{} 1\%; e.g., grandmother), or else the respondent was not indicated (24\%).
+We asked the caregiver more familiar with each language to complete the respective CDI form.
+The forms were mainly filled out by mothers (64\%), fathers (7\%), both parents (4\%), other caregivers (\textless{} 1\%; e.g., grandmother), or else the respondent was not indicated (24\%).
 In some cases different caregivers filled out each form, while in other cases the same caregiver filled out both forms.
@@
 In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
-Words that had similar phonetic realizations (e.g., English ``alligator'' and French \emph{``alligator''}) were also considered translation equivalents.
+Words with the same meaning that had similar phonetic realizations (e.g., English ``alligator'' and French \emph{``alligator''}) were also considered translation equivalents.
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., in press, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson's correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between BALANCE and age in our sample.
-We believe that this could be related to patterns of language exposure at child care (i.e., attending childcare in one language could have made exposure less balanced), which is not related to our model per se.}).
+Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction Set 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between BALANCE and age in our sample, such that older children had vocabularies that were less equal across their languages.
+This is not predicted by our model, and we can only speculate as to why this pattern was observed.
+For example, older children may spend more time in childcare where mostly one language is used, therefore leading to a less balanced vocabulary learning.
+Another possibility is that older children increasingly elicit speech in their dominant language, which in turn amplifies vocabulary imbalance.
+Testing these possibilities is beyond the scope of the current paper.}).
 
@@
 Prediction 1b was that children with larger word vocabularies and larger dominant-language vocabularies would produce more translation equivalents, and the results from our dataset confirmed this prediction, for word vocabulary (WORD): \emph{r}(227) = 0.90, \(p < .001\), and dominant-language vocabulary (DOM): \emph{r}(227) = 0.76, \(p < .001\).
-Figure 1 Panels 2A and 2B further illustrates these relationships observed in our dataset.
-
-Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE), specifically that this relationship would be nearly perfect.
+Figure 1 Panels 2A and 2B further illustrate these relationships observed in our dataset.
+
+Prediction 1c was that children who produce more words in the non-dominant language (NONDOM) would produce more translation equivalents (TE) --- specifically, that this relationship would be nearly perfect.
 As shown in Figure 1 Panel 2C, we observed that these two variables were indeed nearly perfectly correlated, \emph{r}(227) = 0.99, \(p < .001\).
@@
 In our data set, developmental level was approximated by children's age.
-Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET) and in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of different ages (our proxy for developmental level), used to estimate the number of LEARNABLE words.
+Figure 2 Panel B shows the number of concept vocabulary (CONCEPT), translation equivalents (TE), singlets in the dominant language (DOM-SINGLET), and singlets in the non-dominant language (NONDOM-SINGLET) of the bilingual children as a function of age (our proxy for developmental level), used to estimate the number of LEARNABLE words.
 
@@
 
-Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, or the Preference Account.
+Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, and the Preference Account.
 To directly test the correspondence of our data with these different accounts, we built a linear regression model predicting the observed number of translation equivalents from the Bilingual Vocabulary Model using the formula TE = DOM×NONDOM/LEARNABLE, and we allowed the model to estimate the BIAS parameter.
@@
 First, we will walk through the parameters in this model.
-The size of dominant vocabulary (DOM) and size of non-dominant vocabulary (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
-Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
+Dominant vocabulary size (DOM) and non-dominant vocabulary size (NONDOM) were taken to be the number of words produced by individual children observed in the vocabulary data.
+Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary norms at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
 Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
-For example, for an 18 month-old infant, the denominator was 240 words which was calculated by averaging the 259 English words and 220 French words, based on what 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 to 33 months in our dataset, the 90th percentile of 30-month-old children was used since 90th percentile information was available only up to 30 months.
-Finally, the intercept of the linear regression model was set at 0 since no translation equivalents are expected to be produced if a child does not know any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
+For example, for an 18-month-old infant, the denominator was 240 words, which was calculated by averaging the 259 English words and 220 French words that 18-month-old children would typically produce at the 90th percentile.
+For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since 90th percentile information was available only up to 30 months.
+Finally, the intercept of the linear regression model was set at 0, since no translation equivalents are expected to be produced if a child does not have any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
+
+~
 
@@
 \caption{\label{tab:table6}Table 6.
-The number of total English and French productive CDI vocabulary at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
+The number of total English and French productive CDI vocabulary words at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
 Values are rounded to the nearest integer.}
@@
 \fontsize{10}{12}\selectfont
+\begin{threeparttable}
 \begin{tabular}[t]{>{\centering\arraybackslash}p{75px}>{\centering\arraybackslash}p{125px}>{\centering\arraybackslash}p{125px}>{\centering\arraybackslash}p{125px}}
@@
 \end{tabular}
+\begin{tablenotes}
+\item \textit{Note: } 
+\item 90th percentile information was only available up to age 30 months from WordBank, so we used that value for all children in our dataset aged 30 months and older.
+\end{tablenotes}
+\end{threeparttable}
 \end{table}
@@
 Our model showed an excellent model fit of \(R^2\) = 0.96, indicating that our model explained 96\% of the variance in bilinguals' translation equivalent knowledge.
-BIAS coefficient was estimated at 1.04.
-This value is close to 1, suggesting that our data are consistent with the account whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
-Overall, model fit was slightly reduced compared to our main model, although fit was still good, \(R^2\) = 0.95.
+The BIAS coefficient was estimated at 1.04.
+This value is close to 1, suggesting that our data are consistent with the Neutral account, whereby translation equivalents are learned equivalently to other words\footnote{We also explored how these results would vary if we set the number of LEARNABLE words at the 75th percentile, rather than at the 90th percentile.
+Overall, model fit was slightly reduced compared to our main model, although still good, \(R^2\) = 0.95.
 The estimated BIAS parameter was lower, at 0.95, which is nonetheless still quite close to 1, and thus consistent with the Neutral Account.
@@
 Note that visual inspection suggested that there could be some possible outliers.
-Cook's distance was estimated for our linear regression model listed above and identified two data points with a cook's distance over 0.4.
+Cook's distance was estimated for our linear regression model described above and identified two data points with a cook's distance over 0.4.
 After removing those two data points, the linear regression model returned a coefficient of 1.07, with \(R^2\) = 0.96.
@@
 Figure 6 displays the model fit separately for children with a word vocabulary (WORD) less than 300 words and those with a word vocabulary (WORD) of 300 or greater.
-Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 total vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 total vocabulary.
-To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 total vocabulary and for those with more than 300 total vocabulary.
-The model for those with larger total word vocabulary (WORD) returned a coefficient of BIAS = 1.04, whereas the model for those with less than 300 total word vocabulary (WORD) returned a coefficient of BIAS = 2.21.
-Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary size (\(R^2\) = 0.96) than children with smaller vocabulary size (\(R^2\) = 0.88).
+Based on visual inspection, the slope of translation equivalent learning appeared steeper for children with less than 300 word vocabulary, suggesting that translation equivalents are more easily learned (i.e., BIAS \textgreater{} 1); whereas the slope of translation equivalent learning appeared to align with the Neutral Account (i.e., BIAS = 1) for children with more than 300 word vocabulary.
+To further explore this pattern, we ran the same linear regression twice, separately for children with less than 300 word vocabulary and for those with more than 300 word vocabulary.
+The model for those with a larger word vocabulary (WORD) returned a coefficient of BIAS = 1.04, whereas the model for those with a smaller word vocabulary (WORD) returned a coefficient of BIAS = 2.21.
+Both models fit well, although a somewhat better fit was obtained for children with larger vocabulary sizes (\(R^2\) = 0.96) than children with smaller vocabulary sizes (\(R^2\) = 0.88).
 Overall, this analysis suggests that translation equivalent learning for children with larger vocabularies corresponds best to the Neutral Account, but translation equivalent learning for children with smaller vocabularies corresponds best to the Preference Account.
@@
 
-\caption{The number of observed translation equivalents as a function of number of expected translation equivalents under the Bilingual Vocabulary Model other (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
+\caption{The number of observed translation equivalents as a function of the number of expected translation equivalents under the Bilingual Vocabulary Model (represented by the blue solid line), plotted separately for children with fewer than 300 word vocabulary (left panel) and for those with more than 300 word vocabulary (right panel).
 The dashed diagonal line represents the case where the parameter equals 1 (BIAS = 1) such that the predicted and observed number of TEs are equal.}\label{fig:fig6}
@@
 
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether translation equivalents are harder (Avoidance Account), easier (Preference Account), or similar (Neutral Account) for bilingual children to learn as singlet words (i.e., the first label for a particular referent).
+The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether learning translation equivalents is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words (i.e., the first label for a particular referent).
 To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn given the child's age.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn, given a child's age.
 The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
@@
 
-In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18-33 months.
-Three sets of model predictions were confirmed in our empirical dataset.
-In Table 7, we summarize the model predictions along with the evidence in our empirical dataset and findings in the literature.
+In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18--33 months.
+All three sets of model predictions were confirmed in our empirical dataset.
+In Table 7, we summarize the model predictions along with the evidence from our empirical dataset and other findings in the literature.
 
@@
 & The number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
- There is a positive correlation between bilingual children’s size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013) & Yes\\
+ There is a positive correlation between bilingual children’s dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013) & Yes\\
 1C: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE); but unlike for WORD and DOM this does not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge & Vocabulary size in the non-dominant language was positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
@@
 Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children's size of dominant language vocabulary and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
+This pattern is consistent with previous research reporting a positive correlation between bilingual children's dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
 Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
@@
 Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents are similar to learn as singlets), with this model explaining 96\% of variance in the data.
+The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining 96\% of variance in the data.
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
-Modeling their data separately, we found evidence for the Preference Account: younger children at around 22 months appeared to learn translation equivalents more easily than singlets, whereas older children at around 28 months learned translation equivalents similarly to singlets.
+Modeling their data separately, we found evidence for the Preference Account: younger children at around 22 months old appeared to learn translation equivalents more easily than singlets, whereas older children at around 28 months old learned translation equivalents similarly to singlets.
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
@@
 However, according to the Complementarity Principle (Grosjean, 2016), bilinguals may have different experiences in each of their languages.
-For example, a French--English bilingual child who always spends bathtime with an English-speaking parent might encounter and use bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
-At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produced (e.g., Braginsky et al., 2019; Tardif et al., 2008).
-For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquired across languages and cultures.
+For example, a French--English bilingual child who always spends bathtime with an English-speaking caregiver might encounter and use bath words primarily in English (e.g., ``soap,'' ``bath,'' ``bubbles''), therefore having less opportunity to acquire their translation equivalents in French.
+At the same time, cross-linguistic data has provided evidence of a high degree of commonality in the first words children produce (e.g., Braginsky et al., 2019; Tardif et al., 2008).
+For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquire across languages and cultures.
 It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
@@
 Still other languages categorize objects differently within conceptual categories (e.g., a shallow dish might be called a ``bowl'' in English but an \emph{``assiette''} {[}plate{]} in French).
-There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storm et al., 2015; White et al., 2020).
-Therefore, it seems reasonable to assume that early vocabulary may be represented similarly across languages, especially for the age group we were looking at.
+There is mixed evidence for whether bilingual adults maintain separate (Jared et al., 2012) versus integrated (Ameel et al., 2009) conceptual representations across their two languages, while evidence for young bilingual children suggests that conceptual representations are mostly shared between languages (Storms et al., 2015; White et al., 2020).
+Therefore, it seems reasonable to assume that early vocabulary may be represented similarly across languages, especially for the age group we looked at in this study.
 Nonetheless, future iterations of the Bilingual Vocabulary Model may consider capturing the cases where concepts are less well aligned across languages.
@@
 (2013).
-Vocabulary growth and composition in monolingual and bilingual basque infants and toddlers.
+Vocabulary growth and composition in monolingual and bilingual {B}asque infants and toddlers.
 \emph{International Journal of Bilingualism}, \emph{17}(3), 357--374.
@@
 (2013).
-An examination of language input and vocabulary development of young latino dual language learners living in poverty.
+An examination of language input and vocabulary development of young {L}atino dual language learners living in poverty.
 \emph{First Language}, \emph{33}(6), 572--593.
@@
 \leavevmode\hypertarget{ref-Byers-Heinlein_etal_2021}{}%
-Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (2021). \emph{Six solutions for more reliable infant research}. Preprint. Retrieved from \url{https://psyarxiv.com/u37fy}
+Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (in press). Six solutions for more reliable infant research. \emph{Infant and Child Development}. \url{https://doi.org/10.31234/osf.io/u37fy}
 
@@
 (2008).
-Individual differences in the lexical development of french--english bilingual children.
+Individual differences in the lexical development of {F}rench--{E}nglish bilingual children.
 \emph{International Journal of Bilingual Education and Bilingualism}, \emph{11}(5), 598--618.
@@
 Early understanding of two words for the same thing: A CDI study of lexical comprehension in infant bilinguals.
-\emph{International Journal of Bilingual}, \emph{10}(3), 331--347.
+\emph{International Journal of Bilingualism}, \emph{10}(3), 331--347.
 \url{https://doi.org/10.1177/13670069060100030401}
@@
 (2007).
-\emph{MacArthur-bates communicative development inventories (CDIs)} (2nd ed.).
+\emph{MacArthur-{B}ates {C}ommunicative {D}evelopment {I}nventories (CDIs)} (2nd ed.).
 Baltimore, MD: Brookes Publishing.
@@
 (2013).
-SES differences in language processing skill and vocabulary are evident at 18 monthss.
+SES differences in language processing skill and vocabulary are evident at 18 months.
 \emph{Developmental Science}, \emph{16}, 234--248.
@@
 (2014).
-Relative language exposure, processing efficiency and vocabulary in spanish--english bilingual toddler.
+Relative language exposure, processing efficiency and vocabulary in {S}panish--{E}nglish bilingual toddler.
 \emph{Bilingualism: Language and Cognition}, \emph{17}(1), 189--202.
@@
 Dog or chien?
-Translation equivalents in the receptive and expressive vocabularies of young french--english bilinguals.
+Translation equivalents in the receptive and expressive vocabularies of young {F}rench--{E}nglish bilinguals.
 \emph{Journal of Child Language}, \emph{44}(4), 881--904.
@@
 (2010).
-How vocabulary size in two languages relates to efficiency in spoken word recognition by young spanish-english bilinguals.
+How vocabulary size in two languages relates to efficiency in spoken word recognition by young {S}panish-{E}nglish bilinguals.
 \emph{Journal of Child Language}, \emph{37}(4), 817--840.
@@
 Gunnar \& M.
-Maratsos (Eds.), \emph{The minnesota symposia on child psychology} (Vol.
+Maratsos (Eds.), \emph{The {M}innesota {S}ymposia on {C}hild {P}sychology} (Vol.
 25, pp.
@@
 (2018).
-Quantitative linguistic predictors of infants' learning of specific english words.
+Quantitative linguistic predictors of infants' learning of specific {E}nglish words.
 \emph{Child Development}, \emph{89}(4), 1247--1267.
@@
 (1996).
-Nouns are not always learned before verbs: Evidence from mandarin speakers' early vocabularies.
+Nouns are not always learned before verbs: Evidence from {M}andarin speakers' early vocabularies.
 \emph{Developmental Psychology}, \emph{32}(3), 492--504.
@@
 (1999).
-Une adaptation en français québecois du MacArthur communicative development inventory {[}a quebec french adaptation of the MacArthur communicative development inventory{]}.
+Une adaptation en français québecois du MacArthur {C}ommunicative {D}evelopment {I}nventory {[}a {Q}uebec {F}rench adaptation of the MacArthur {C}ommunicative {D}evelopment {I}nventory{]}.
 \emph{Revue d'orthophonie Et d'audiologie}, \emph{23}, 31--73.

### 2022-02-18 paper/paper_TE_bilingual_vocabulary_model.Rmd
updated paper

@@
 
-[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word “hygge”, which refers to a type of cozy contentment and enjoyment of simplicity.
+[^1]: We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word _“hygge”_, which refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults’ concepts may not be congruent across languages (e.g., Malt & Majid, 2013).
@@
 That is, many potential singlets could be “too hard” to learn at a particular age, while by definition potential translational equivalents refer to concepts that children are already able to lexicalize.
-For example, an 14-month-old who has learned the French words _“ballon”_ and _“chien”_ would be capable of learning their translation equivalents “ball” and “dog”, as well as easier potential singlets like “cup”, but not later-acquired potential singlets like "vitamins", “tomorrow”, “who”, or “not”.
+For example, an 14-month-old who has learned the French words _“ballon”_ and _“chien”_ would be capable of learning their translation equivalents “ball” and “dog”, as well as easier potential singlets like “cup”, but not later-acquired potential singlets like “vitamins”, “tomorrow”, “who”, or “not”.
 Thus, a seeming overabundance of translation equivalents might be a product of developmental constraints on word learning, rather than due to semantic facilitation.
@@
 
-[^2]: To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
+[^2]: To test the independence of two events from the same population, as an example, we might ask whether psychology majors are more likely to be left-handed.
 Imagine a college of 1000 students.
@@
 
-The model takes four main parameters: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS > 1) or against (BIAS < 1) learning translation equivalents.
+The model takes four main parameters as predictors: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS > 1) or against (BIAS < 1) learning translation equivalents.
 The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language.
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI.
-For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-However, not all children will be capable of learning all of the CDI words.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600–700 depending on the language of the adaptation.
+Yet, not all children will be capable of learning all of the CDI words.
 Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019).
@@
 Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child’s developmental level, which could be approximated by their age.
-Since there are no CDI norms for French—English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
+Since there are no CDI norms for French–English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
@@
 Using the mathematical concept of independence, we can then quantify the number of translation equivalents (TE) expected given children’s vocabulary sizes in the dominant (DOM) and non-dominant (NONDOM) languages, as well as the number of potentially learnable words (LEARNABLE).
-If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) — which is the possible number of words that could overlap across both languages — to predict the number of translation equivalents.
+If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total number of learnable words in one language (LEARNABLE) — which is the possible number of words that could overlap across both languages — to predict the number of translation equivalents.
 We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral Account.
@@
 
-Finally, based on the main parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters.
+Finally, based on the four main predictive parameters, we can calculate additional, commonly-reported descriptors of bilingual vocabulary, which we detail below and describe as derived parameters (see also Table 1).
 
@@
 
-Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
+Word vocabulary (WORD; sometimes called total vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
 Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child — that is, the total number of concepts that are lexicalized in either language.
@@
 Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction, which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
-Note that usually only three values are necessary to calculate all the other variables (see Table 1).
-Most commonly, we can calculate other variables based on the total number of learnable words (LEARNABLE) together with either the words known in each language (DOM and NONDOM) or word vocabulary plus balance (WORD and BALANCE) which allow us to compute DOM and NONDOM.
-It is also possible to calculate other variables based on the total number of learnable words (LEARNABLE) with balance and words known in either language (BALANCE and DOM or BALANCE and NONDOM).
-
-We present three simulations generated by creating data points for hypothetical children by varying different parameters of the Bilingual Vocabulary Model, in order to explore expected patterns of translation equivalent learning.
+Simulations were conducted by holding constant or varying the main parameters across a set of possible values, and computing expected translation equivalents and derived parameters from these values.
 The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except those that violated constraints in the model (e.g., NONDOM could not be larger than DOM).
-In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), and word vocabulary (WORD).
-In the second simulation, we explored relations between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE).
-In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the BIAS parameter to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts.
+
+For each simulation, we generated predictions for expected patterns in bilingual children’s vocabulaires.
+Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David & Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016).
+However, others have examined derived parameters such as BALANCE (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), WORD (Legacy et al., 2016; Montanari, 2010), CONCEPT (Pearson et al., 1993), DOM-SINGLET and NONDOM-SINGLET (De Houwer et al., 2006; Pearson et al., 1993; 1995).
+To facilitate comparison with the extant literature where main parameters are not always reported, we generated predictions expressed both in terms of main parameters and derived parameters.
+This provides alternate ways of testing the model, depending on which parameters are available in the dataset.
+Note that predictions from the main parameters and those from derived parameters do not provide independent tests of our model for the same dataset, as they can be algebraically calculated from one another.
+As an analogy, a child’s age is not independent from their birth date and today’s date, yet we might make reference to their age in some situations and their birth date in others.
+Similarly, elucidating relations among both main and derived parameters provides a more complete description of expected patterns in bilingual vocabulary development.
+
+In the first simulation, we examined relations between translation equivalents (TE), dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), word vocabulary (WORD), and vocabulary balance (BALANCE), while keeping the number of learnable words (LEARNABLE) constant.
+In the second simulation, we varied the number of learnable words (LEARNABLE), and examined expected number of translation equivalents (TE), singlets in each language (DOM-SINGLET and NONDOM-SINGLET), and concept vocabulary (CONCEPT).
+In both of the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the BIAS parameter to examine expected patterns of translation equivalent (TE) learning as a function of word vocabulary (WORD) under the Avoidance, Preference, and Neutral Accounts.
 A summary of the parameter values used in each simulation is provided in Table 2.
@@
 ## 1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
-In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English–French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
-We set BIAS to 1, meaning that in these examples translation equivalents and singlets are equally easy to learn.
-
-Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
-She has a word vocabulary of 300, and a balance score of .10 (10% of her words are in the non-dominant language).
-Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents.
-Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
-Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40% of his words are in the non-dominant language).
-Based on our formula, we expect Bernie to produce 36 translation equivalents.
-Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
-Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
-Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
-Infants Bernie and Charlie illustrate that for two children equal in BALANCE, the child with larger word vocabulary (WORD) is expected to produce more translation equivalents (TE).
-Other vocabulary metrics are calculated for each hypothetical child as described in Table 3.
-
-We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
-BIAS remained constant at 1.
+In Simulation 1, we examined patterns of translation equivalent learning where simulated children are at the same developmental level and thus have the same number of potentially LEARNABLE words, but their vocabulary size in each language (DOM and NONDOM) varied.
+For readers less familiar with this type of approach, we provide expanded examples of three simulated infants in the Appendix.
+For convenience, we set LEARNABLE constant at 600, which roughly corresponds to what is expected for an English–French 27-month-old bilingual child.
+For example, a 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and a 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages (Frank et al., 2016).
+BIAS was set constant at 1, meaning that in these simulations translation equivalents and singlets are equally easy to learn.
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
 Based on these values, we derived simulated children’s word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+
+In Figure 1, we plotted TE knowledge as a function of WORD, DOM, and NONDOM at different levels of BALANCE.
 Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as DOM (dominant language words) and WORD (word vocabulary) increased, TE (translation equivalents) also increased regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+Moreover, Panels 1A and 1B show that, as  WORD (word vocabulary) and DOM (dominant language words) increased, TE (translation equivalents) also increased regardless of BALANCE.
+Interestingly, Panel 1C shows that NONDOM and TE were extremely tightly coupled.
 In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
-
-```{r table3, results="asis", cache = TRUE}
-
-# create columns that contain information to be added to table3 in the next step
-## Definition column: the definition of each variable in the table
-Definition <- c("Bias parameter", "Learnable words in each language", "Words produced in the dominant language", "Words produced in the non-dominant language",
-                "Word vocabulary (or total vocabulary size)", "Vocabulary balance", "Translation equivalents produced", 
-                "Concept vocabulary (or total conceptual vocabulary size)", "Singlets in dominant language", "Singlets in non-dominant language")
-## Calculation column: the formula used to calculate the derived parameters
-Calculation <- c("", "", "", "", 
-                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM × NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
-
-table3 <- data.frame(
-    # define the main parameters
-    hypotheical_child = c("Infant Anne (small vocabulary, unbalanced)",
-                        "Infant Bernie (small vocabulary, balanced)",
-                        "Infant Charlie (large vocabulary, balanced)"),
-    BIAS = 1, 
-    LEARNABLE = 600, 
-    DOM = c(270, 180, 540),
-    NONDOM = c(30, 120, 360)
-  ) %>%
-  # calculate the derived parameters from the main parameters
-  mutate(WORD = DOM + NONDOM,
-         BALANCE = NONDOM/(DOM+NONDOM),
-         TE = DOM*NONDOM/LEARNABLE,
-         CONCEPT = WORD - TE,
-         "DOM-SINGLET" = DOM - TE, 
-         "NONDOM-SINGLET" = NONDOM - TE) %>%
-  # transpose the data set for presentation
-  pivot_longer(-("hypotheical_child"), names_to = "Variable", values_to = "value") %>%
-  pivot_wider(names_from = "hypotheical_child", values_from = "value") %>%
-  # append other information for presentation
-  mutate(Definition = Definition, 
-         Calculation = Calculation) %>%
-  # reorder columns
-  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)") %>%
-  # round to the nearest integer
-  mutate_if(is.numeric, round, digits = 0)
-
-landscape(
-  kable(table3,
-      align = "llllll",
-      format="latex", booktabs = TRUE,
-      caption = "Table 3.
-Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
-Vocabulary values are rounded to the nearest integer.") %>%
-  pack_rows("Main Parameters", 1, 4) %>%
-  pack_rows("Derived Parameters", 5, 10) %>%
-  column_spec(c(1), width = "100px") %>%
-  column_spec(c(2:3), width = "140px") %>%
-  column_spec(c(4:6), width = "60px") %>%
-  kable_styling(font_size = 8) 
-)
-```
 
@@
 * Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE).
-Unlike for WORD and DOM this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
+Unlike for WORD and DOM, this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
 
@@
 ## 1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
-In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE held constant at 600).
+In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE was held constant at 600).
 As laid out in the introduction, under the Bilingual Vocabulary Model, the learnability of different words changes with a child’s developmental level, where LEARNABLE increases as a child grows older.
@@
 Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
-As shown in Panel 1A, simulated children at a later developmental level had larger concept vocabularies (CONCEPT).
-In Panel 1B, we continued to observe a pattern reported in prediction 1a, whereby simulated children with more balanced vocabularies produced more translation equivalents (TE).
-Moreover, regardless of balance, simulated children at later developmental levels (i.e., older children with more potentially LEARNABLE words) acquired more translation equivalents (TE).
-Panels 1C and 1D show that simulated children at later developmental levels also produced more singlets in the dominant and non-dominant languages, and that more balanced children produced more singlets in their non-dominant language, but fewer singlets in their dominant language than less balanced children.
 Overall, we generated 4 additional predictions (Prediction Set 2) made by the Bilingual Vocabulary Model.
-Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will
+Compared to children at an earlier developmental level (i.e., younger infants with fewer potentially learnable words), children at a later developmental level (i.e., older infants with more potentially learnable words) will:
 
@@
                              TE = "Panel 1B: \n Translation equivalents \n (TE)",
-                             DOM_SINGLET = "Panel 1C: Singlets in \n dominant language \n (DOM_SINGLET)",
-                             NONDOM_SINGLET = "Panel 1D: Singlets in \n non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             DOM_SINGLET = "Panel 1C: Singlets in \n dominant language \n (DOM-SINGLET)",
+                             NONDOM_SINGLET = "Panel 1D: Singlets in \n non-dominant language \n (NONDOM-SINGLET)")) %>%
   ggplot(aes(x = LEARNABLE, y = number, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
                              number_of_te = "Panel 2B: \n Translation equivalents \n (TE)",
-                             total_singlet_dom = "Panel 2C: Singlets in \n dominant language \n (DOM_SINGLET)",
-                             total_singlet_nondom = "Panel 2D: Singlets in \n non-dominant language \n (NONDOM_SINGLET)")) %>%
+                             total_singlet_dom = "Panel 2C: Singlets in \n dominant language \n (DOM-SINGLET)",
+                             total_singlet_nondom = "Panel 2D: Singlets in \n non-dominant language \n (NONDOM-SINGLET)")) %>%
   ggplot(aes(x = age_months_percentile, y = number_of_words, color = bin_BALANCE, linetype = bin_BALANCE)) +
@@
 
-## 1.3 Simulation 3: Bias towards or against translation equivalent learning compared to singlets
-In Simulations 1 and 2, we modeled cases in accordance with the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
-In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
+## 1.3 Simulation 3: Bias towards or against translation equivalent learning
+In Simulations 1 and 2, we modeled cases with the bias parameter held constant at 1 and translation equivalent (TE) was calculated as DOM×NONDOM/LEARNABLE.
+Thus, these simulations were in accordance with the Neutral Account where dominant-language and non-dominant language vocabularies were learned independently.
+In our final simulation, we examined cases where dominant-language and non-dominant language vocabularies were not independent, corresponding to the Avoidance Account and the Preference Account.
 Mathematically, this requires varying the BIAS parameter.
-For the Preference Account, BIAS will be greater than 1, meaning that TEs are more easily learned than singlets.
-By contrast, for the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets.
+For the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets.
+By contrast, for the Avoidance Account, BIAS will be less than 1, meaning that translation equivalents are less easily learned than singlets.
 
 Translation equivalent (TE) knowledge was first simulated across different developmental levels (number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
-Again, BALANCE and word vocabulary (WORD) were calculated based on the values of DOM and NONDOM.
+Again, BALANCE and word vocabulary (WORD) were derived based on the values of DOM and NONDOM.
 The final simulated data set contained 166 data points (see Table 2 for a summary of the parameters used).
@@
 Study 1 generated predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+The generated predictions were expressed in terms of the model’s four main predictive parameters as well as a function of derived parameters to illustrate expected patterns in bilingual children’s vocabularies.
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
+
 Prediction Set 1 illustrated that the more words children produce in total word vocabulary, in their dominant language, or in their non-dominant language, the more translation equivalents they are predicted to produce.
@@
 Older children are predicted to have a larger concept vocabulary (i.e., to have lexicalized more concepts), including translation equivalents and singlets in both dominant and non-dominant languages, than younger children.
-Moreover, the most balanced children are expected to produce more singlets in their non-dominant language, whereas the least balanced children are expected to produce more singlets in their dominant language.
+Moreover, the most balanced children are expected to produce more singlets in their non-dominant language than the least balanced children, whereas the least balanced children are expected to produce more singlets in their dominant language than the most balanced children.
 Finally, Prediction Set 3 quantitatively demonstrated the patterns of translation equivalent learning under the three different accounts: the Avoidance Account, the Preference Account, and the Neutral Account.
-Crucially, these predictions should not be thought of as fully independent from each other, but instead illustrate the interrelatedness of measures of bilingual vocabulary under the Bilingual Vocabulary Model.
-In Study 2, we used vocabulary data from real bilingual infants to investigate whether these same patterns of interrelatedness would be observed, thus validating our model.
+In Study 2, we used vocabulary data from real bilingual infants to investigate our predictions, thus validating our model.
 
@@
 Participants were recruited through government birth lists, online ads, daycares, and infant-parent group activities (e.g., children’s library activities).
-Inclusion criteria were the following: full term-pregnancy ( i.e., > 36 weeks of gestation), normal birth weight (> 2500 grams), and absence of major medical conditions (i.e., meningitis).
+Inclusion criteria were the following: full term-pregnancy ( i.e., > 36 weeks of gestation), normal birth weight (> 2500 grams), and absence of major medical conditions (e.g., meningitis).
 Only children who had complete data in both CDI forms (i.e., English and French) were retained for analysis.
@@
 Within the `r length(keepers_ws_TE$ID_testdate)` data points, `r round(d_lang_dom$percent_EngDom, 1)`% of children were English-dominant and `r round(d_lang_dom$percent_FrDom, 1)`% were French-dominant.
-Children’s demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
+Children’s demographic characteristics including age, maternal education, and language exposure, are presented in Table 3.
 
 &nbsp;
-```{r table4, results="asis", cache = TRUE}
+```{r table3, results="asis", cache = TRUE}
 # Create dataframe of demographics
@@
 papaja::apa_table(d_participants, format.args = list(digits = 1),
-                  caption = "Table 4. Demographic characteristics of participants (data points = 229).",
+                  caption = "Table 3. Demographic characteristics of participants (data points = 229).",
                   col.names =c("","Mean","$SD$","Range"),
@@
 In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
-Words with the same meaning that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were also considered translation equivalents.
+Words with the same meaning that had similar phonetic realizations (e.g., English “alligator” and French _“alligator”_) were considered translation equivalents.
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
@@
 The LEQ and MAPLE provide a global language exposure estimate based on the number of hours the child is exposed to each language within all contexts (e.g., home, daycare, etc.).
-Children’s average global exposure to each language is described in Table 4.
+Children’s average global exposure to each language is described in Table 3.
 
 ## 2.1.3 Procedure
-Caregivers were asked to fill out the CDI questionnaires as part of their child’s participation in experimental studies on language development, speech perception, and word learning.
+Caregivers were asked to fill out the CDI questionnaires as part of their child’s participation in experimental studies on language and cognitive development.
 Caregivers were instructed to check off the words produced by their child using either a CDI paper questionnaire or the same questionnaire administered on a tablet.
@@
 On average, bilingual children in our sample had a balance score BALANCE of `r balance_average$mean[1]` (SD = `r balance_average$sd[1]`), ranging from `r balance_average$min[1]` to `r balance_average$max[1]`.
-Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
+SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
 The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., in press, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2022, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 ## 2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures
-Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson’s correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction Set 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations[^3]).
+Prediction Set 1 pertained to the pairwise relationships between translation equivalents (TE), word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), and vocabulary balance (BALANCE), which we examined through Pearson’s correlations.
+Overall, the univariate statistics showed strong correspondence with the relationships expressed in Prediction Set 1 under the Bilingual Vocabulary Model (see Table 4 for a full table of pairwise correlations[^3]).
 
@@
 
-```{r table5, results = "asis", cache = TRUE}
+```{r table4, results = "asis", cache = TRUE}
 # correlation among variables in the observed data
@@
 papaja::apa_table(correlation_table,
-                  caption = "Table 5. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).",
+                  caption = "Table 4. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).",
                   note = "*** p < .001, ** p < .01, * p < .05.",
@@
 Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary norms at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
-Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
+Table 5 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
 For example, for an 18-month-old infant, the denominator was `r ws_percentile$"Average (LEARNABLE)"[1]` words, which was calculated by averaging the `r ws_percentile$"Number of English words produced at 90th percentile"[1]` English words and `r ws_percentile$"Number of French words produced at 90th percentile"[1]` French words that 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since 90th percentile information was available only up to 30 months.
+For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since vocabulary norms were only available up to 30 months.
 Finally, the intercept of the linear regression model was set at 0, since no translation equivalents are expected to be produced if a child does not have any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
@@
 &nbsp;
-```{r table6, results = "asis", cache = TRUE}
+```{r table5, results = "asis", cache = TRUE}
 kable(ws_percentile,
@@
         format = "latex", booktabs = TRUE, align = "c",
-        caption = "Table 6.
+        caption = "Table 5.
 The number of total English and French productive CDI vocabulary words at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
@@
 
-```{r fig5, fig.cap="The number of simulated and observed translation equivalents plotted against each other.
+```{r fig5, fig.cap="The number of observed and predicted translation equivalents plotted against each other.
 The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
@@
 # Discussion
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether learning translation equivalents is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words (i.e., the first label for a particular referent).
-To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn, given a child’s age.
-The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
-
-## Confirmation of model predictions
-In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18–33 months.
-All three sets of model predictions were confirmed in our empirical dataset.
-In Table 7, we summarize the model predictions along with the evidence from our empirical dataset and other findings in the literature.
-
-```{r table7, results = "asis", cache = TRUE}
+The aim of the current study was to better understand translation equivalent learning in bilingual children.
+To do so, we developed a simple model, the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets (i.e., the first label for a particular referent) were possible for an individual child to learn.
+The nature of translation equivalent learning was modeled via a parameter that indicated whether translation equivalent learning is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words.
+
+In Study 1, we generated specific predictions from the model via a series of simulations, and in Study 2 we tested these predictions against CDI production data from 200 bilingual children aged 18–33 months.
+There were two main findings: 1) The predictions from the Bilingual Vocabulary Model were born out in our data, thus showing that our simple quantitative model can explain numerous patterns in young bilinguals’ word learning; and 2) younger children’s learning of translation equivalents was more consistent with the Preference Account, while older children’s learning was more consistent with the Neutral account.
+In what follows, we elaborate on each of these findings.
+
+## Understanding patterns in bilingual children’s vocabulary development
+
+Translation equivalent learning has been of interest in studies of bilingual children’s development from the very earliest descriptive work (Ronjat, 2013), to more recent quantitative research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+Many previous studies have examined univariate relations between translation equivalent knowledge and predictors such as age or vocabulary balance (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), but such observations lacked a unifying framework.
+The current paper developed and validated such a framework — the Bilingual Vocabulary Model — which specifies the relationship amongst multiple parameters simultaneously.
+Translation equivalent production is predicted from four parameters: vocabulary size in the dominant language, vocabulary size in the non-dominant language, the number of potentially learnable words in each language given a child’s developmental level, and a bias parameter indicating whether translation equivalent learning is harder than, easier than, or similar to singlet learning.
+Translation equivalent production can also be reexpressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
+
+We validated the Bilingual Vocabulary Model by generating a set of predictions in Study 1 which were then tested on our empirical dataset in Study 2.
+First, we predicted that children would produce more translation equivalents if they have more balanced vocabularies (i.e., producing a similar number of words in each language), produce more words (whether in total, in their dominant language, or in their non-dominant language), and if they are at a more advanced developmental level (i.e., older).
+Specifically, our model predicted that vocabulary size in the non-dominant language would have a near-perfect relationship with translation equivalent production.
+Second, we predicted that children would produce more singlets in their dominant language if they had less balanced vocabularies, and more singlets in their non-dominant language if they had more balanced vocabularies.
+Each of these predictions was confirmed in our empirical data, and the patterns found were consistent with numerous previous reports in the literature (Boyce et al., 2013; David & Wei, 2008; Hoff et al., 2012; Legacy et al., 2016; Marchman et al., 2010; Montanari, 2010; Pearson et al., 1993; 1995; 1997; Place & Hoff, 2011; Poulin-Dubois et al., 2013).
+We summarize our model’s predictions and related empirical literature in Table 6.
+This work therefore advances the field by providing a unifying model that explains patterns in bilingual children’s vocabulary development.
+
+```{r table6, results = "asis", cache = TRUE}
 
@@
         format = "latex", booktabs = TRUE,
-        caption = "Table 7. Summary of model predictions, similar evidence from previous literature, and whether our empirical data was consistent with the prediction.") %>%
+        caption = "Table 6. Summary of model predictions, related findings from previous literature, and whether our empirical data were consistent with the prediction.") %>%
     column_spec(c(1), width = "200px") %>%
@@
 
-A first set of predictions pertained to relations between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
-In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words also produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
-Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children’s dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
-Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
-
-A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets.
-We operationalized developmental level in terms of children’s age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
-Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
-Likewise, we confirmed the model-generated prediction that older children would produce more translation equivalents than younger children.
-This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David & Wei, 2008; Legacy et al., 2016).
-Predictions 2c and 2d were also confirmed by our vocabulary data.
-While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place & Hoff, 2011).
-Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
-
-Overall, we observed a strong correspondence between the data simulated under the Bilingual Vocabulary Model and our observed data.
-Moreover, our model predicted numerous disparate patterns that have been previously reported in the literature.
-
-Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
+## The status of translation equivalents in bilingual vocabulary development 
+Applying the Bilingual Vocabulary Model to a large set of empirical vocabulary data from bilingual children allowed us to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+Contrary to early theories of bilingual development, in our data children did not seem to avoid producing a word if they knew its translation equivalent in the other language (Avoidance Account; Imedadze, 1967; Swain & Wesche, 1975; Volterra & Taeschner, 1978); the number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining `r round(summary(test_prediction3_model)$r.squared,2) * 100`% of variance in the data.
+This overall finding is in line with previous reports that bilinguals’ vocabularies in each language develop relatively independently (Hurtado et al., 2014; Marchman et al., 2010).
+
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
@@
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
-This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
-The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
-Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
-The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children’s own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
-
-## Developmental change in translation equivalent learning
-The developmental change we observed in bilingual children’s ability to learn translation equivalents could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
+
+We hypothesize that this pattern could stem from changes in children’s word learning abilities across development.
+Previous studies have suggested that strong semantic overlap helps bilingual children learn translation equivalents more easily than singlets (Bilson et al., 2015).
+However, it is possible that the boost children receive from this semantic facilitation changes as bilingual children develop.
+Younger bilinguals, who are relative word learning novices, might strongly benefit from translation equivalents, especially if establishing a conceptual representation is a particularly hard part of the word learning process.
+For translation equivalent words, children have already formed a conceptual representation and only need to learn a second label for a particular referent (Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
+By contrast, older bilinguals are relative word learning experts.
+It may be that establishing a conceptual representation is somewhat easy, and thus any facilitation effects of translation equivalents might be relatively small.
+Such developmental changes would be consistent with work showing that children’s word learning abilities develop substantially across the first years of life (Bergelson, 2021; Hirsh-Pasek & Golinkoff, 2000; Nazzi & Bertoncini, 2003; Tsui et al., 2019).
+
+An additional possibility is that the nature of bilingual input changes as children become more advanced word learners.
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
+For example, a caregiver may choose to say to their English–French bilingual child “Can you give me the _livre_?” if they know their child understands the French word _“livre”_ but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
+Thus, changes in bilingual input may not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
+
+Another explanation for the developmental pattern we observed is that the way in which age and developmental level are related changes across infancy.
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder & Fernald, 2013).
+Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
+However, there could be much more variability in the developmental level of younger infants than of older infants.
+In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+Longitudinal data might be useful for taking these individual differences into account.
+
+Finally, we must consider whether the pattern we observed could be related to changes in children’s use of one-to-one mapping biases such as mutual exclusivity.
 As revealed by previous studies, younger children and children with smaller vocabulary sizes (and thus less vocabulary knowledge) seem to have only a weak bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
@@
 
-Another possible explanation for our findings is that the nature of bilingual input changes as children become more advanced word learners.
-Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child’s vocabulary (Kremin et al., 2021; Nicoladis & Secco, 2000).
-For example, a caregiver may choose to say to their English–French bilingual child “Can you give me the _livre_?” if they know their child understands the French word “_livre_” but not the English equivalent “book.” This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
-However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
-Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
-
-Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to changes in input patterns or word learning biases.
-Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder & Fernald, 2013).
-Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
-However, there could be much more variability in the developmental level of younger infants than of older infants.
-In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
-Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
-
 ## Assumptions, limitations, and future directions
@@
 To do so, our model parameters included the number of words produced in each language, as well as children’s developmental level.
+This represents an important methodological advance compared to previous approaches, which used data from randomly-paired monolinguals or the vocabularies of two different bilinguals as a reference point (e.g., Bilson et al., 2015; Pearson et al., 1995).
 However, our approach does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
@@
 Such a model may better characterize and predict individual differences in bilingual vocabulary development.
-Moreover, it would be important to further test our Bilingual Vocabulary Model using longitudinal data.
 
@@
 For example, words for important people (“mommy,” “daddy”), social routines (“hi,” “bye,” “yes,” “no”), and common nouns (“ball,” “dog”) are among the first words children acquire across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+It therefore seems reasonable to expect that young bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
 Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance Account (e.g., fewer than expected translation equivalents), which is not what we observed.
@@
 We proposed the Bilingual Vocabulary Model, which provides a quantitative way to test these accounts, by modeling translation equivalent learning in relation to vocabulary size in each language and the number of potentially learnable words, which is constrained by children’s developmental level.
-Results using archival data from a large number of young French–English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
-Moreover, our model parsimoniously explains previously disparate observations about bilingual children’s translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Results using archival data from a large number of young French–English bilingual children showed that our model parsimoniously explains previously disparate observations about bilingual children’s translation equivalent learning.
+Overall, our model was a good fit to the Neutral Account suggesting that translation equivalent learning may be neither easier nor harder for bilingual children to learn than singlets, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
+We hypothesize that such a developmental change reflects changes in bilingual children’s word learning abilities, where bilingual children initially gain substantial benefit from the semantic overlap in translation equivalents, and subsequently benefit less from this type of facilitation as they grow older and become more proficient at word learning.
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
@@
 
+# Appendix
+To more clearly illustrate the relationships between different variables in the Bilingual Vocabulary Model, we provide a more expanded and detailed illustration of our simulation approach.
+Here, we simulated three hypothetical children Annie, Bernie and Charlie who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
+As we did in the main text, we again set LEARNABLE = 600 and BIAS = 1 in this example.
+Other vocabulary metrics for each of the three hypothetical children are described in Table A1.
+
+Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
+She has a word vocabulary of 300, and a balance score of .10 (10% of her words are in the non-dominant language).
+Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table A1, Annie is expected to produce 13.5 translation equivalents.
+
+Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
+Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40% of his words are in the non-dominant language).
+Based on our formula, we expect Bernie to produce 36 translation equivalents.
+Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
+This corresponds to Prediction 1a in the main text.
+
+Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
+Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
+Infants Bernie and Charlie illustrate that for two children equal in BALANCE, the child with larger word vocabulary (WORD) is expected to produce more translation equivalents (TE).
+This corresponds to Prediction 1b in the main text.
+
+```{r tableA1, results="asis", cache = TRUE}
+
+# create columns that contain information to be added to table3 in the next step
+## Definition column: the definition of each variable in the table
+Definition <- c("Bias parameter", "Learnable words in each language", "Words produced in the dominant language", "Words produced in the non-dominant language",
+                "Word vocabulary (or total vocabulary size)", "Vocabulary balance", "Translation equivalents produced", 
+                "Concept vocabulary (or total conceptual vocabulary size)", "Singlets in dominant language", "Singlets in non-dominant language")
+## Calculation column: the formula used to calculate the derived parameters
+Calculation <- c("", "", "", "", 
+                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM × NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
+
+table3 <- data.frame(
+    # define the main parameters
+    hypotheical_child = c("Infant Anne (small vocabulary, unbalanced)",
+                        "Infant Bernie (small vocabulary, balanced)",
+                        "Infant Charlie (large vocabulary, balanced)"),
+    BIAS = 1, 
+    LEARNABLE = 600, 
+    DOM = c(270, 180, 540),
+    NONDOM = c(30, 120, 360)
+  ) %>%
+  # calculate the derived parameters from the main parameters
+  mutate(WORD = DOM + NONDOM,
+         BALANCE = NONDOM/(DOM+NONDOM),
+         TE = DOM*NONDOM/LEARNABLE,
+         CONCEPT = WORD - TE,
+         "DOM-SINGLET" = DOM - TE, 
+         "NONDOM-SINGLET" = NONDOM - TE) %>%
+  # transpose the data set for presentation
+  pivot_longer(-("hypotheical_child"), names_to = "Variable", values_to = "value") %>%
+  pivot_wider(names_from = "hypotheical_child", values_from = "value") %>%
+  # append other information for presentation
+  mutate(Definition = Definition, 
+         Calculation = Calculation) %>%
+  # reorder columns
+  relocate(c(Definition, Calculation), .before = "Infant Anne (small vocabulary, unbalanced)") %>%
+  # round to the nearest integer
+  mutate_if(is.numeric, round, digits = 0)
+
+landscape(
+  kable(table3,
+      align = "llllll",
+      format="latex", booktabs = TRUE,
+      caption = "Table A1.
+Expanded examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Vocabulary values are rounded to the nearest integer.") %>%
+  pack_rows("Main Parameters", 1, 4) %>%
+  pack_rows("Derived Parameters", 5, 10) %>%
+  column_spec(c(1), width = "100px") %>%
+  column_spec(c(2:3), width = "140px") %>%
+  column_spec(c(4:6), width = "60px") %>%
+  kable_styling(font_size = 8) 
+)
+```
+
+\newpage
+
 # References

### 2022-02-18 paper/paper_TE_bilingual_vocabulary_model.tex
updated paper

@@
 That is, upon hearing the English word ``apple,'' the corresponding French word \emph{``pomme''} is more easily activated in bilinguals' minds.
-In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent, in a different language, learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word ``hygge,'' which refers to a type of cozy contentment and enjoyment of simplicity.
+In vocabulary acquisition, bilingual children must learn a first label for a referent (a ``singlet''; Umbel et al., 1992) before they can learn its translation equivalent (i.e., a second label for a referent, in a different language, learned following a singlet)\footnote{We note that in some cases a singlet may not have a close translation equivalent, for example the recently popular Danish word \emph{``hygge''}, which refers to a type of cozy contentment and enjoyment of simplicity.
 Moreover, even when a word that could be considered a translation equivalent exists, adults' concepts may not be congruent across languages (e.g., Malt \& Majid, 2013).
@@
 This provides a straightforward empirical test of different theoretical accounts of translation equivalent learning, by asking whether or not the probability of knowing a word is independent of knowing its translation equivalent.
-The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent\footnote{To test the independence of two events from the same population, as an example, we might ask whether Psychology majors are more likely to be left-handed.
+The logic is similar to that of the familiar chi-squared test for independence, where the independence of two events from the same population is tested as the probability of their intersection computed by multiplying the probability of each individual event: P(A and B) = P(A) × P(B) if A and B are independent\footnote{To test the independence of two events from the same population, as an example, we might ask whether psychology majors are more likely to be left-handed.
 Imagine a college of 1000 students.
@@
 
-The model takes four main parameters: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS \textgreater{} 1) or against (BIAS \textless{} 1) learning translation equivalents.
+The model takes four main parameters as predictors: the number of words produced in the dominant language (DOM), the number of words produced in the non-dominant language (NONDOM), vocabulary size of potentially learnable words in each language (LEARNABLE), and a bias parameter (BIAS) which indicates whether the model is biased towards (BIAS \textgreater{} 1) or against (BIAS \textless{} 1) learning translation equivalents.
 The language in which a child knows more words is the dominant language, whereas the one in which a child knows fewer words is the non-dominant language.
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI.
-For convenience, consider the effect of setting LEARNABLE to 600, as a round number (the actual number of CDI items is usually slightly higher than 600, depending on the language of the adaptation).
-However, not all children will be capable of learning all of the CDI words.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600--700 depending on the language of the adaptation.
+Yet, not all children will be capable of learning all of the CDI words.
 Typically, words that are more concrete occur more frequently in the input, and words that are associated with babies tend to be acquired earlier than other words (Braginsky et al., 2019).
 That is, very young children would not be expected to produce many of the advanced words on the CDI, such as ``lawn mower,'' ``sidewalk,'' or ``vitamins.'' Thus, a reasonable solution might be to determine how many CDI words are potentially learnable given the child's developmental level, which could be approximated by their age.
-Since there are no CDI norms for French---English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
+Since there are no CDI norms for French--English bilingual children, we will take the average English and French CDI norms to approximate how many words children might lexicalize in either language.
 For example, imagine that Jamie who is 18 months old produces 50 English words and 20 French words, thus a total of 70 words.
@@
 Using the mathematical concept of independence, we can then quantify the number of translation equivalents (TE) expected given children's vocabulary sizes in the dominant (DOM) and non-dominant (NONDOM) languages, as well as the number of potentially learnable words (LEARNABLE).
-If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total population of learnable words in one language (LEARNABLE) --- which is the possible number of words that could overlap across both languages --- to predict the number of translation equivalents.
+If dominant-language and non-dominant-language words are learned independently from each other, we multiply DOM × NONDOM (the number of words known in the dominant and non-dominant language respectively), and divide by the total number of learnable words in one language (LEARNABLE) --- which is the possible number of words that could overlap across both languages --- to predict the number of translation equivalents.
 We further introduce the bias parameter (BIAS), which allows us to examine whether translation equivalent learning is best described by the Avoidance, Preference, or Neutral Account.
@@
 
-Word vocabulary (WORD; sometimes called total productive vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
+Word vocabulary (WORD; sometimes called total vocabulary) is the total number of words that a child produces across the two languages, calculated as the sum of the dominant vocabulary (DOM) and non-dominant vocabulary (NONDOM).
 Concept vocabulary (CONCEPT; sometimes called total conceptual vocabulary) is the number of concepts that are lexicalized by the child --- that is, the total number of concepts that are lexicalized in either language.
@@
 Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction, which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
-Note that usually only three values are necessary to calculate all the other variables (see Table 1).
-Most commonly, we can calculate other variables based on the total number of learnable words (LEARNABLE) together with either the words known in each language (DOM and NONDOM) or word vocabulary plus balance (WORD and BALANCE) which allow us to compute DOM and NONDOM.
-It is also possible to calculate other variables based on the total number of learnable words (LEARNABLE) with balance and words known in either language (BALANCE and DOM or BALANCE and NONDOM).
-
-We present three simulations generated by creating data points for hypothetical children by varying different parameters of the Bilingual Vocabulary Model, in order to explore expected patterns of translation equivalent learning.
+Simulations were conducted by holding constant or varying the main parameters across a set of possible values, and computing expected translation equivalents and derived parameters from these values.
 The number of children in each simulation was fully determined by the number of levels varied for each parameter: each combination of parameters was included, except those that violated constraints in the model (e.g., NONDOM could not be larger than DOM).
-In the first simulation, we examined how translation equivalent learning relates to vocabulary balance (BALANCE), as well as different metrics of vocabulary size, including dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), and word vocabulary (WORD).
-In the second simulation, we explored relations between translation equivalents (TE), balance (BALANCE), and learnable words (LEARNABLE).
-In the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the BIAS parameter to compare translation equivalent learning under the Avoidance, Preference, and Neutral Accounts.
+
+For each simulation, we generated predictions for expected patterns in bilingual children's vocabulaires.
+Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David \& Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016).
+However, others have examined derived parameters such as BALANCE (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), WORD (Legacy et al., 2016; Montanari, 2010), CONCEPT (Pearson et al., 1993), DOM-SINGLET and NONDOM-SINGLET (De Houwer et al., 2006; Pearson et al., 1993; 1995).
+To facilitate comparison with the extant literature where main parameters are not always reported, we generated predictions expressed both in terms of main parameters and derived parameters.
+This provides alternate ways of testing the model, depending on which parameters are available in the dataset.
+Note that predictions from the main parameters and those from derived parameters do not provide independent tests of our model for the same dataset, as they can be algebraically calculated from one another.
+As an analogy, a child's age is not independent from their birth date and today's date, yet we might make reference to their age in some situations and their birth date in others.
+Similarly, elucidating relations among both main and derived parameters provides a more complete description of expected patterns in bilingual vocabulary development.
+
+In the first simulation, we examined relations between translation equivalents (TE), dominant-language vocabulary (DOM), non-dominant language vocabulary (NONDOM), word vocabulary (WORD), and vocabulary balance (BALANCE), while keeping the number of learnable words (LEARNABLE) constant.
+In the second simulation, we varied the number of learnable words (LEARNABLE), and examined expected number of translation equivalents (TE), singlets in each language (DOM-SINGLET and NONDOM-SINGLET), and concept vocabulary (CONCEPT).
+In both of the first two simulations, the BIAS parameter was held constant at 1 (Neutral Account); in the third simulation, we varied the BIAS parameter to examine expected patterns of translation equivalent (TE) learning as a function of word vocabulary (WORD) under the Avoidance, Preference, and Neutral Accounts.
 A summary of the parameter values used in each simulation is provided in Table 2.
@@
 
-In Simulation 1, we first illustrate the relationships between different variables in the model by simulating three hypothetical children who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
-For convenience, we set LEARNABLE = 600 in this example, which roughly corresponds to what is expected for an English--French 27-month-old bilingual child (i.e., the most verbal 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and the most verbal 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages; Frank et al., 2016).
-We set BIAS to 1, meaning that in these examples translation equivalents and singlets are equally easy to learn.
-
-Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
-She has a word vocabulary of 300, and a balance score of .10 (10\% of her words are in the non-dominant language).
-Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table 3, Annie is expected to produce 13.5 translation equivalents.
-Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
-Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40\% of his words are in the non-dominant language).
-Based on our formula, we expect Bernie to produce 36 translation equivalents.
-Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
-Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
-Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
-Infants Bernie and Charlie illustrate that for two children equal in BALANCE, the child with larger word vocabulary (WORD) is expected to produce more translation equivalents (TE).
-Other vocabulary metrics are calculated for each hypothetical child as described in Table 3.
-
-We then broadened this simulation to the more general case and examined patterns of translation equivalent learning, where simulated children had the capacity to learn 600 words (LEARNABLE held constant at 600), and their vocabulary size in each language (DOM and NONDOM) varied.
-BIAS remained constant at 1.
+In Simulation 1, we examined patterns of translation equivalent learning where simulated children are at the same developmental level and thus have the same number of potentially LEARNABLE words, but their vocabulary size in each language (DOM and NONDOM) varied.
+For readers less familiar with this type of approach, we provide expanded examples of three simulated infants in the Appendix.
+For convenience, we set LEARNABLE constant at 600, which roughly corresponds to what is expected for an English--French 27-month-old bilingual child.
+For example, a 27-month-old English-learner at the 90th percentile of vocabulary produces around 633 words and a 27-month-old French-learner produces around 580 words as retrieved from the Wordbank database, giving us an average of 606 words across the two languages (Frank et al., 2016).
+BIAS was set constant at 1, meaning that in these simulations translation equivalents and singlets are equally easy to learn.
 Data from a total of 216 simulated children were generated (see Table 2 for a summary of the parameter values used in this simulation).
 Based on these values, we derived simulated children's word vocabulary (WORD, calculated as DOM+NONDOM) and their vocabulary balance (BALANCE, calculated as NONDOM/(DOM+NONDOM)).
-In Figure 1, we plotted TE knowledge as a function of DOM, NONDOM, and WORD at different levels of BALANCE.
+
+In Figure 1, we plotted TE knowledge as a function of WORD, DOM, and NONDOM at different levels of BALANCE.
 Across all three Panels (1A, 1B, and 1C), simulated children with the most balanced vocabulary consistently produced more translation equivalents than other children.
-Moreover, Panels 1A and 1C show that, as DOM (dominant language words) and WORD (word vocabulary) increased, TE (translation equivalents) also increased regardless of BALANCE.
-Interestingly, Panel 1B shows that NONDOM and TE were extremely tightly coupled.
+Moreover, Panels 1A and 1B show that, as WORD (word vocabulary) and DOM (dominant language words) increased, TE (translation equivalents) also increased regardless of BALANCE.
+Interestingly, Panel 1C shows that NONDOM and TE were extremely tightly coupled.
 In sum, we observed three important patterns, which served as Prediction Set 1 from the Bilingual Vocabulary Model for Study 2:
-
-\begin{landscape}\begin{table}
-
-\caption{\label{tab:table3}Table 3.
-Examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
-Vocabulary values are rounded to the nearest integer.}
-\centering
-\fontsize{8}{10}\selectfont
-\begin{tabular}[t]{>{\raggedright\arraybackslash}p{100px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}}
-\toprule
-Variable & Definition & Calculation & Infant Anne (small vocabulary, unbalanced) & Infant Bernie (small vocabulary, balanced) & Infant Charlie (large vocabulary, balanced)\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Main Parameters}}\\
-\hspace{1em}BIAS & Bias parameter &  & 1 & 1 & 1\\
-\hspace{1em}LEARNABLE & Learnable words in each language &  & 600 & 600 & 600\\
-\hspace{1em}DOM & Words produced in the dominant language &  & 270 & 180 & 540\\
-\hspace{1em}NONDOM & Words produced in the non-dominant language &  & 30 & 120 & 360\\
-\addlinespace[0.3em]
-\multicolumn{6}{l}{\textbf{Derived Parameters}}\\
-\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & DOM + NONDOM & 300 & 300 & 900\\
-\hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0 & 0 & 0\\
-\hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 14 & 36 & 324\\
-\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286 & 264 & 576\\
-\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256 & 144 & 216\\
-\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16 & 84 & 36\\
-\bottomrule
-\end{tabular}
-\end{table}
-\end{landscape}
 
@@
 \item
-  Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE) (Figure 1, Panels 1A and 1B).
+  Prediction 1b: Children who produce more total words (WORD) or more dominant-language words (DOM) will produce more translation equivalents (TE) (Figure 1, Panels 1A and 1B).\\
 \item
   Prediction 1c: Children who produce more non-dominant language words (NONDOM) will produce more translation equivalents (TE).
-Unlike for WORD and DOM this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
+Unlike for WORD and DOM, this will not interact with BALANCE; instead, non-dominant vocabulary size will be an almost perfect predictor of translation equivalent knowledge (Figure 1, Panel 1C).
 \end{itemize}
+
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 20 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 140 rows containing missing values (geom_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 3 rows containing non-finite values (stat_smooth).
+\end{verbatim}
+
+\begin{verbatim}
+## Warning: Removed 25 rows containing missing values (geom_smooth).
+\end{verbatim}
 
@@
 
-In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE held constant at 600).
+In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE was held constant at 600).
 As laid out in the introduction, under the Bilingual Vocabulary Model, the learnability of different words changes with a child's developmental level, where LEARNABLE increases as a child grows older.
@@
 
-\hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}{%
-\subsection{1.3 Simulation 3: Bias towards or against translation equivalent learning compared to singlets}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning-compared-to-singlets}}
-
-In Simulations 1 and 2, we modeled cases in accordance with the Neutral Account where dominant-language and non-dominant language words were learned independently, such that the bias parameter (BIAS) was exactly 1 when we calculated TE as DOM×NONDOM/LEARNABLE.
-In our final simulation, we examined cases where dominant-language and non-dominant language words were not independent, corresponding to the Avoidance Account and the Preference Account.
+\hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning}{%
+\subsection{1.3 Simulation 3: Bias towards or against translation equivalent learning}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning}}
+
+In Simulations 1 and 2, we modeled cases with the bias parameter held constant at 1 and translation equivalent (TE) was calculated as DOM×NONDOM/LEARNABLE.
+Thus, these simulations were in accordance with the Neutral Account where dominant-language and non-dominant language vocabularies were learned independently.
+In our final simulation, we examined cases where dominant-language and non-dominant language vocabularies were not independent, corresponding to the Avoidance Account and the Preference Account.
 Mathematically, this requires varying the BIAS parameter.
-For the Preference Account, BIAS will be greater than 1, meaning that TEs are more easily learned than singlets.
-By contrast, for the Avoidance Account, BIAS will be less than 1, meaning that TEs are less easily learned than singlets.
+For the Preference Account, BIAS will be greater than 1, meaning that translation equivalents are more easily learned than singlets.
+By contrast, for the Avoidance Account, BIAS will be less than 1, meaning that translation equivalents are less easily learned than singlets.
 
 Translation equivalent (TE) knowledge was first simulated across different developmental levels (number of LEARNABLE words = 150, 300, 450, 600), in conjunction with a wide range of values for DOM and NONDOM.
-Again, BALANCE and word vocabulary (WORD) were calculated based on the values of DOM and NONDOM.
+Again, BALANCE and word vocabulary (WORD) were derived based on the values of DOM and NONDOM.
 The final simulated data set contained 166 data points (see Table 2 for a summary of the parameters used).
@@
 Study 1 generated predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
+The generated predictions were expressed in terms of the model's four main predictive parameters as well as a function of derived parameters to illustrate expected patterns in bilingual children's vocabularies.
 Assuming that translation equivalents were learned similarly to singlets (i.e., the Neutral Account), three sets of predictions were made.
+
 Prediction Set 1 illustrated that the more words children produce in total word vocabulary, in their dominant language, or in their non-dominant language, the more translation equivalents they are predicted to produce.
@@
 Participants were recruited through government birth lists, online ads, daycares, and infant-parent group activities (e.g., children's library activities).
-Inclusion criteria were the following: full term-pregnancy ( i.e., \textgreater{} 36 weeks of gestation), normal birth weight (\textgreater{} 2500 grams), and absence of major medical conditions (i.e., meningitis).
+Inclusion criteria were the following: full term-pregnancy ( i.e., \textgreater{} 36 weeks of gestation), normal birth weight (\textgreater{} 2500 grams), and absence of major medical conditions (e.g., meningitis).
 Only children who had complete data in both CDI forms (i.e., English and French) were retained for analysis.
@@
 Within the 229 data points, 59.80\% of children were English-dominant and 40.20\% were French-dominant.
-Children's demographic characteristics including age, maternal education, and language exposure, are presented in Table 4.
+Children's demographic characteristics including age, maternal education, and language exposure, are presented in Table 3.
 
@@
 
-\caption{\label{tab:table4}Table 4. Demographic characteristics of participants (data points = 229).}
+\caption{\label{tab:table3}Table 3. Demographic characteristics of participants (data points = 229).}
 
@@
 In cases of disagreement, a discussion of the likely uses of the word in question by children (rather than potential adult uses of the word) was conducted and then a decision was made.
-Words with the same meaning that had similar phonetic realizations (e.g., English ``alligator'' and French \emph{``alligator''}) were also considered translation equivalents.
+Words with the same meaning that had similar phonetic realizations (e.g., English ``alligator'' and French \emph{``alligator''}) were considered translation equivalents.
 Most of the items on both vocabulary checklists had an equivalent word in the other language, which resulted in a total of 611 translation equivalents.
@@
 The LEQ and MAPLE provide a global language exposure estimate based on the number of hours the child is exposed to each language within all contexts (e.g., home, daycare, etc.).
-Children's average global exposure to each language is described in Table 4.
+Children's average global exposure to each language is described in Table 3.
 
@@
 
-Caregivers were asked to fill out the CDI questionnaires as part of their child's participation in experimental studies on language development, speech perception, and word learning.
+Caregivers were asked to fill out the CDI questionnaires as part of their child's participation in experimental studies on language and cognitive development.
 Caregivers were instructed to check off the words produced by their child using either a CDI paper questionnaire or the same questionnaire administered on a tablet.
@@
 On average, bilingual children in our sample had a balance score BALANCE of 0.31 (SD = 0.13), ranging from 0.02 to 0.50.
-Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
+SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
 The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.20\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., in press, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2022, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 
-Prediction Set 1 pertained to the pairwise relationships between word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), vocabulary balance (BALANCE), and translation equivalents (TE), which we examined through Pearson's correlations.
-Overall, the univariate statistics showed strong correspondence with the relationships predicted by Prediction Set 1 under the Bilingual Vocabulary Model (see Table 5 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between BALANCE and age in our sample, such that older children had vocabularies that were less equal across their languages.
+Prediction Set 1 pertained to the pairwise relationships between translation equivalents (TE), word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), and vocabulary balance (BALANCE), which we examined through Pearson's correlations.
+Overall, the univariate statistics showed strong correspondence with the relationships expressed in Prediction Set 1 under the Bilingual Vocabulary Model (see Table 4 for a full table of pairwise correlations\footnote{Note that there was a negative correlation between BALANCE and age in our sample, such that older children had vocabularies that were less equal across their languages.
 This is not predicted by our model, and we can only speculate as to why this pattern was observed.
@@
 \begin{longtable}{llllllllll}\noalign{\getlongtablewidth\global\LTcapwidth=\longtablewidth}
-\caption{\label{tab:table5}Table 5. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).}\\
+\caption{\label{tab:table4}Table 4. Pairwise correlations among variables (corrected for multiple comparisons using Benjamini and Yekutieli [2001]).}\\
 \toprule
@@
 \endfirsthead
-\caption*{\normalfont{Table \ref{tab:table5} continued}}\\
+\caption*{\normalfont{Table \ref{tab:table4} continued}}\\
 \toprule
@@
 Again, the number of learnable words (LEARNABLE) was determined by averaging English and French productive CDI vocabulary norms at the 90th percentile at different ages (percentile information obtained from Wordbank version 0.3.1; Frank et al., 2016).
-Table 6 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
+Table 5 lists the denominators derived from averaging these percentiles, or LEARNABLE, at different ages.
 For example, for an 18-month-old infant, the denominator was 240 words, which was calculated by averaging the 259 English words and 220 French words that 18-month-old children would typically produce at the 90th percentile.
-For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since 90th percentile information was available only up to 30 months.
+For children who were between 31 and 33 months old in our dataset, the 90th percentile for 30-month-old children was used, since vocabulary norms were only available up to 30 months.
 Finally, the intercept of the linear regression model was set at 0, since no translation equivalents are expected to be produced if a child does not have any dominant or non-dominant vocabulary (i.e., when the predictor variables are 0).
@@
 
-\caption{\label{tab:table6}Table 6.
+\caption{\label{tab:table5}Table 5.
 The number of total English and French productive CDI vocabulary words at the 90th percentile at different ages according to WordBank, and the average between the two which serves as the denominator in our computation model.
@@
 
-\caption{The number of simulated and observed translation equivalents plotted against each other.
+\caption{The number of observed and predicted translation equivalents plotted against each other.
 The dots represent the value of a child tested on the CDI, with their observed number of TEs and the expected number of TEs based on our model.
@@
 
-The aim of the current study was to better understand translation equivalent learning in bilingual children, specifically investigating whether learning translation equivalents is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words (i.e., the first label for a particular referent).
-To test these accounts, we developed the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
-The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets were possible to learn, given a child's age.
-The relative difficulty of learning translation equivalents compared to singlets was modeled via a parameter that indicated whether translation equivalent learning was consistent with the Avoidance, Preference, or Neutral Account.
-
-\hypertarget{confirmation-of-model-predictions}{%
-\subsection{Confirmation of model predictions}\label{confirmation-of-model-predictions}}
-
-In Study 1, we simulated vocabulary and translation equivalent knowledge based on the Bilingual Vocabulary Model, and in Study 2 we tested three sets of model-generated predictions using archival CDI data from 200 bilingual children aged 18--33 months.
-All three sets of model predictions were confirmed in our empirical dataset.
-In Table 7, we summarize the model predictions along with the evidence from our empirical dataset and other findings in the literature.
+The aim of the current study was to better understand translation equivalent learning in bilingual children.
+To do so, we developed a simple model, the Bilingual Vocabulary Model, which quantifies the number of translation equivalents that children produce as a product of words they know in their dominant and non-dominant language, divided by the number of words that are learnable at their developmental level.
+The inclusion of a learnability parameter was a unique aspect of our approach, and was crucial to quantifying how many translation equivalents versus singlets (i.e., the first label for a particular referent) were possible for an individual child to learn.
+The nature of translation equivalent learning was modeled via a parameter that indicated whether translation equivalent learning is harder than (Avoidance Account), easier than (Preference Account), or similar to (Neutral Account) learning singlet words.
+
+In Study 1, we generated specific predictions from the model via a series of simulations, and in Study 2 we tested these predictions against CDI production data from 200 bilingual children aged 18--33 months.
+There were two main findings: 1) The predictions from the Bilingual Vocabulary Model were born out in our data, thus showing that our simple quantitative model can explain numerous patterns in young bilinguals' word learning; and 2) younger children's learning of translation equivalents was more consistent with the Preference Account, while older children's learning was more consistent with the Neutral account.
+In what follows, we elaborate on each of these findings.
+
+\hypertarget{understanding-patterns-in-bilingual-childrens-vocabulary-development}{%
+\subsection{Understanding patterns in bilingual children's vocabulary development}\label{understanding-patterns-in-bilingual-childrens-vocabulary-development}}
+
+Translation equivalent learning has been of interest in studies of bilingual children's development from the very earliest descriptive work (Ronjat, 2013), to more recent quantitative research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+Many previous studies have examined univariate relations between translation equivalent knowledge and predictors such as age or vocabulary balance (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), but such observations lacked a unifying framework.
+The current paper developed and validated such a framework --- the Bilingual Vocabulary Model --- which specifies the relationship amongst multiple parameters simultaneously.
+Translation equivalent production is predicted from four parameters: vocabulary size in the dominant language, vocabulary size in the non-dominant language, the number of potentially learnable words in each language given a child's developmental level, and a bias parameter indicating whether translation equivalent learning is harder than, easier than, or similar to singlet learning.
+Translation equivalent production can also be reexpressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
+
+We validated the Bilingual Vocabulary Model by generating a set of predictions in Study 1 which were then tested on our empirical dataset in Study 2.
+First, we predicted that children would produce more translation equivalents if they have more balanced vocabularies (i.e., producing a similar number of words in each language), produce more words (whether in total, in their dominant language, or in their non-dominant language), and if they are at a more advanced developmental level (i.e., older).
+Specifically, our model predicted that vocabulary size in the non-dominant language would have a near-perfect relationship with translation equivalent production.
+Second, we predicted that children would produce more singlets in their dominant language if they had less balanced vocabularies, and more singlets in their non-dominant language if they had more balanced vocabularies.
+Each of these predictions was confirmed in our empirical data, and the patterns found were consistent with numerous previous reports in the literature (Boyce et al., 2013; David \& Wei, 2008; Hoff et al., 2012; Legacy et al., 2016; Marchman et al., 2010; Montanari, 2010; Pearson et al., 1993; 1995; 1997; Place \& Hoff, 2011; Poulin-Dubois et al., 2013).
+We summarize our model's predictions and related empirical literature in Table 6.
+This work therefore advances the field by providing a unifying model that explains patterns in bilingual children's vocabulary development.
 
@@
 
-\caption{\label{tab:table7}Table 7. Summary of model predictions, similar evidence from previous literature, and whether our empirical data was consistent with the prediction.}
+\caption{\label{tab:table6}Table 6. Summary of model predictions, related findings from previous literature, and whether our empirical data were consistent with the prediction.}
 \centering
@@
 
-A first set of predictions pertained to relations between translation equivalent knowledge, vocabulary balance, and vocabulary size in the dominant and non-dominant languages.
-In both the simulated and observed data, children with more balanced vocabularies (i.e., those who produced a similar number of words in each of their languages) produced more translation equivalents.
-This pattern is consistent with reports from previous research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
-Moreover, both the simulated and observed data showed that the children who produced more total words also produced more translation equivalents, which is in line with previous research showing that the number of translation equivalents a bilingual child knows increases along with their total vocabulary size (Legacy et al., 2016; Montanari, 2010).
-Additionally, both our simulated and observed data showed that the more words children produced in their dominant language, the more translation equivalents they also produced.
-This pattern is consistent with previous research reporting a positive correlation between bilingual children's dominant language vocabulary size and the proportion of translation equivalents (Legacy et al., 2016; Poulin-Dubois et al., 2013).
-Finally, both our simulated and observed data showed that the more words children produced in their non-dominant language, the more translation equivalents they produced.
-A similar pattern has been reported by Legacy and colleagues (2016), where vocabulary size in the non-dominant language positively correlated with the proportion of translation equivalents known by the child (Legacy et al., 2016).
-
-A second set of predictions pertained to the relationship between the number of potentially learnable words for a child (constrained by their developmental level) and the production of translation equivalents and singlets.
-We operationalized developmental level in terms of children's age, and set the number of learnable words at the number produced by children at the 90th percentile for that age (averaged across French and English).
-Both simulated and observed data showed older children had larger concept vocabularies, a pattern consistent with reports from previous literature (Pearson et al., 1993).
-Likewise, we confirmed the model-generated prediction that older children would produce more translation equivalents than younger children.
-This pattern is consistent with the literature that bilingual children learn more translation equivalents as they grow older (David \& Wei, 2008; Legacy et al., 2016).
-Predictions 2c and 2d were also confirmed by our vocabulary data.
-While children produced more singlets in both the dominant and non-dominant languages with age, the least balanced children produced the most singlets in their dominant language and the most balanced children produced the most singlets in their non-dominant language.
-These patterns are also in line with the notion that bilingual children learn words in proportion to their relative exposure to each language (e.g., Boyce et al., 2013; Hoff et al., 2012; Marchman et al., 2010; Pearson et al., 1997; Place \& Hoff, 2011).
-Therefore, within the number of words that are potentially learnable at a particular developmental level, bilingual children with less balanced language exposure have more opportunities to learn more words in their dominant language than their non-dominant language, whereas bilingual children with more balanced language exposure have more equal opportunities to learn words in each of their language.
-
-Overall, we observed a strong correspondence between the data simulated under the Bilingual Vocabulary Model and our observed data.
-Moreover, our model predicted numerous disparate patterns that have been previously reported in the literature.
-
-Having validated our overall approach, our third prediction motivated using the Bilingual Vocabulary Model to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
-The number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining 96\% of variance in the data.
+\hypertarget{the-status-of-translation-equivalents-in-bilingual-vocabulary-development}{%
+\subsection{The status of translation equivalents in bilingual vocabulary development}\label{the-status-of-translation-equivalents-in-bilingual-vocabulary-development}}
+
+Applying the Bilingual Vocabulary Model to a large set of empirical vocabulary data from bilingual children allowed us to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
+Contrary to early theories of bilingual development, in our data children did not seem to avoid producing a word if they knew its translation equivalent in the other language (Avoidance Account; Imedadze, 1967; Swain \& Wesche, 1975; Volterra \& Taeschner, 1978); the number of translation equivalents children produced was a very close fit to the Neutral Account (i.e., translation equivalents and singlets are learned similarly), with this model explaining 96\% of variance in the data.
+This overall finding is in line with previous reports that bilinguals' vocabularies in each language develop relatively independently (Hurtado et al., 2014; Marchman et al., 2010).
+
 However, there was some indication that the Neutral Account provided a poorer fit for children with smaller vocabulary sizes.
@@
 This could indicate a qualitative shift in word learning that occurs as bilingual children develop and learn more words, from the Preference Account to the Neutral Account.
-This pattern of a qualitative shift contradicts previous evidence proposing that bilingual children between the ages of 6 months and 7 years learn translation equivalents more easily than singlets (Bilson et al., 2015).
-The discrepancy could potentially be explained by the difference in how expected patterns of translation equivalent learning were simulated in each study.
-Previous approaches simulated bilingual language learning using data from randomly-paired monolinguals or lexicons of two different bilinguals as a reference point for the Neutral Account (e.g., Bilson et al., 2015; Pearson et al., 1995).
-The Bilingual Vocabulary Model represents a significant theoretical and methodological advance, as it does not make reference to randomly-paired children, and instead uses children's own dominant and non-dominant vocabulary size, together with their developmental level, to gauge how many translation equivalents they are expected to learn.
-
-\hypertarget{developmental-change-in-translation-equivalent-learning}{%
-\subsection{Developmental change in translation equivalent learning}\label{developmental-change-in-translation-equivalent-learning}}
-
-The developmental change we observed in bilingual children's ability to learn translation equivalents could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
+
+We hypothesize that this pattern could stem from changes in children's word learning abilities across development.
+Previous studies have suggested that strong semantic overlap helps bilingual children learn translation equivalents more easily than singlets (Bilson et al., 2015).
+However, it is possible that the boost children receive from this semantic facilitation changes as bilingual children develop.
+Younger bilinguals, who are relative word learning novices, might strongly benefit from translation equivalents, especially if establishing a conceptual representation is a particularly hard part of the word learning process.
+For translation equivalent words, children have already formed a conceptual representation and only need to learn a second label for a particular referent (Montanari, 2010; Poulin-Dubois et al., 2013; 2018).
+By contrast, older bilinguals are relative word learning experts.
+It may be that establishing a conceptual representation is somewhat easy, and thus any facilitation effects of translation equivalents might be relatively small.
+Such developmental changes would be consistent with work showing that children's word learning abilities develop substantially across the first years of life (Bergelson, 2021; Hirsh-Pasek \& Golinkoff, 2000; Nazzi \& Bertoncini, 2003; Tsui et al., 2019).
+
+An additional possibility is that the nature of bilingual input changes as children become more advanced word learners.
+Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
+For example, a caregiver may choose to say to their English--French bilingual child ``Can you give me the \emph{livre}?'' if they know their child understands the French word \emph{``livre''} but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
+However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
+Thus, changes in bilingual input may not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
+
+Another explanation for the developmental pattern we observed is that the way in which age and developmental level are related changes across infancy.
+Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
+Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
+However, there could be much more variability in the developmental level of younger infants than of older infants.
+In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
+Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
+Longitudinal data might be useful for taking these individual differences into account.
+
+Finally, we must consider whether the pattern we observed could be related to changes in children's use of one-to-one mapping biases such as mutual exclusivity.
 As revealed by previous studies, younger children and children with smaller vocabulary sizes (and thus less vocabulary knowledge) seem to have only a weak bias for a one-to-one mapping between words and referents (Halberda, 2003; Lewis et al., 2020; Merriman et al., 1989).
@@
 
-Another possible explanation for our findings is that the nature of bilingual input changes as children become more advanced word learners.
-Some recent research has suggested that bilingual parents sometimes code-switch to use a word that they know to be in their child's vocabulary (Kremin et al., 2021; Nicoladis \& Secco, 2000).
-For example, a caregiver may choose to say to their English--French bilingual child ``Can you give me the \emph{livre}?'' if they know their child understands the French word ``\emph{livre}'' but not the English equivalent ``book.'' This may provide fewer opportunities for children to learn translation equivalents, since they would be less exposed to the unfamiliar translation equivalents.
-However, this observation would predict that young bilinguals would know fewer translation equivalents as a proportion of their vocabularies than older bilinguals, which was opposite to what we observed.
-Thus, changes in bilingual input also do not provide an adequate explanation for our results of a qualitative change in translation equivalent learning.
-
-Finally, the developmental pattern we observed could be due to differences in how age and developmental level are related across infancy, rather than to changes in input patterns or word learning biases.
-Indeed, even controlling for age, vocabulary size is linked to multiple factors, including cognitive abilities (e.g., Reuter et al., 2018), speech perception skills (e.g., Cristia et al., 2014), and amount of language experience (e.g., Weisleder \& Fernald, 2013).
-Our model used age as a proxy for developmental level, assuming that all children of a particular age could potentially learn the same number of words as children at the 90th vocabulary percentile.
-However, there could be much more variability in the developmental level of younger infants than of older infants.
-In this case, the 90th percentile might be a reasonable estimate for the number of learnable words for most older infants, but an overestimate of the number of learnable words for all but the most cognitively advanced younger infants.
-Given the importance of the learnability parameter in our models, future research will be needed to more precisely quantify the number of words that are learnable by particular children at particular ages.
-
 \hypertarget{assumptions-limitations-and-future-directions}{%
@@
 To do so, our model parameters included the number of words produced in each language, as well as children's developmental level.
+This represents an important methodological advance compared to previous approaches, which used data from randomly-paired monolinguals or the vocabularies of two different bilinguals as a reference point (e.g., Bilson et al., 2015; Pearson et al., 1995).
 However, our approach does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
@@
 For example, words for important people (``mommy,'' ``daddy''), social routines (``hi,'' ``bye,'' ``yes,'' ``no''), and common nouns (``ball,'' ``dog'') are among the first words children acquire across languages and cultures.
-It therefore seems reasonable to expect that bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
+It therefore seems reasonable to expect that young bilingual children would be exposed to and produce words for a similar set of referents in each of their languages.
 Moreover, if indeed bilingual children tend to use different words in different linguistic contexts, we would have expected our data to be consistent with the Avoidance Account (e.g., fewer than expected translation equivalents), which is not what we observed.
@@
 We proposed the Bilingual Vocabulary Model, which provides a quantitative way to test these accounts, by modeling translation equivalent learning in relation to vocabulary size in each language and the number of potentially learnable words, which is constrained by children's developmental level.
-Results using archival data from a large number of young French--English bilingual children showed that our model was a good fit to the Neutral Account, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
-Moreover, our model parsimoniously explains previously disparate observations about bilingual children's translation equivalent learning, for example that the number of translation equivalents children produce is tightly linked to their vocabulary size in their non-dominant language, and thus all else equal children with more balanced vocabularies will produce more translation equivalents.
+Results using archival data from a large number of young French--English bilingual children showed that our model parsimoniously explains previously disparate observations about bilingual children's translation equivalent learning.
+Overall, our model was a good fit to the Neutral Account suggesting that translation equivalent learning may be neither easier nor harder for bilingual children to learn than singlets, although younger children may show a preference for translation equivalent learning in line with the Preference Account.
+We hypothesize that such a developmental change reflects changes in bilingual children's word learning abilities, where bilingual children initially gain substantial benefit from the semantic overlap in translation equivalents, and subsequently benefit less from this type of facilitation as they grow older and become more proficient at word learning.
 Future studies with data from other populations of bilinguals will be important to more fully test the Bilingual Vocabulary Model.
@@
 \textbf{Krista Byers-Heinlein:} Conceptualization, Funding Acquisition, Methodology, Resources, Supervision, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
+
+\newpage
+
+\hypertarget{appendix}{%
+\section{Appendix}\label{appendix}}
+
+To more clearly illustrate the relationships between different variables in the Bilingual Vocabulary Model, we provide a more expanded and detailed illustration of our simulation approach.
+Here, we simulated three hypothetical children Annie, Bernie and Charlie who are at the same developmental level and thus have the same number of potentially learnable words (LEARNABLE), but with different word vocabularies (WORD) and BALANCE.
+As we did in the main text, we again set LEARNABLE = 600 and BIAS = 1 in this example.
+Other vocabulary metrics for each of the three hypothetical children are described in Table A1.
+
+Our first simulated infant, Annie (small vocabulary, unbalanced exposure), produces 270 words in the dominant language and 30 words in her non-dominant language.
+She has a word vocabulary of 300, and a balance score of .10 (10\% of her words are in the non-dominant language).
+Based on the formula TE = DOM×NONDOM/LEARNABLE (we drop BIAS from the formula since it is 1 here) and as seen in Table A1, Annie is expected to produce 13.5 translation equivalents.
+
+Our second infant, Bernie (small vocabulary, balanced exposure), produces 180 dominant-language words and 120 non-dominant language words.
+Like Annie, he has a word vocabulary of 300, but he has a higher balance score of .40 (40\% of his words are in the non-dominant language).
+Based on our formula, we expect Bernie to produce 36 translation equivalents.
+Comparing Annie and Bernie, two children who produce the same total number of word vocabulary (i.e., WORD is held constant), the child with the more balanced bilingual vocabulary (Bernie) is expected to produce more translation equivalents.
+This corresponds to Prediction 1a in the main text.
+
+Like Bernie, our third infant Charlie also has a balanced vocabulary, but has a larger word vocabulary (WORD), producing 540 words in the dominant language (DOM) and 360 in the non-dominant language (NONDOM), for a total of 900 words (WORD), and thus BALANCE = .40.
+Based on our formula for Simulation 1, we expect Charlie to produce 324 translation equivalents (TE).
+Infants Bernie and Charlie illustrate that for two children equal in BALANCE, the child with larger word vocabulary (WORD) is expected to produce more translation equivalents (TE).
+This corresponds to Prediction 1b in the main text.
+
+\begin{landscape}\begin{table}
+
+\caption{\label{tab:tableA1}Table A1.
+Expanded examples for Simulation 1 of three simulated children with different hypothetical word vocabularies (WORD) and vocabulary balances (BALANCE), where the number of learnable words (LEARNABLE) = 600 and BIAS = 1.
+Vocabulary values are rounded to the nearest integer.}
+\centering
+\fontsize{8}{10}\selectfont
+\begin{tabular}[t]{>{\raggedright\arraybackslash}p{100px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{140px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}>{\raggedright\arraybackslash}p{60px}}
+\toprule
+Variable & Definition & Calculation & Infant Anne (small vocabulary, unbalanced) & Infant Bernie (small vocabulary, balanced) & Infant Charlie (large vocabulary, balanced)\\
+\midrule
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Main Parameters}}\\
+\hspace{1em}BIAS & Bias parameter &  & 1 & 1 & 1\\
+\hspace{1em}LEARNABLE & Learnable words in each language &  & 600 & 600 & 600\\
+\hspace{1em}DOM & Words produced in the dominant language &  & 270 & 180 & 540\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language &  & 30 & 120 & 360\\
+\addlinespace[0.3em]
+\multicolumn{6}{l}{\textbf{Derived Parameters}}\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & DOM + NONDOM & 300 & 300 & 900\\
+\hspace{1em}BALANCE & Vocabulary balance & NONDOM / (DOM + NONDOM) & 0 & 0 & 0\\
+\hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 14 & 36 & 324\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286 & 264 & 576\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256 & 144 & 216\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16 & 84 & 36\\
+\bottomrule
+\end{tabular}
+\end{table}
+\end{landscape}
 
@@
 
-\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2021}{}%
-Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (in press). Six solutions for more reliable infant research. \emph{Infant and Child Development}. \url{https://doi.org/10.31234/osf.io/u37fy}
+\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2022}{}%
+Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (2022). Six solutions for more reliable infant research. \emph{Infant and Child Development}. \url{https://doi.org/10.1002/icd.2296}
 
@@
 Halberda, J. (2003). The development of a word-learning strategy. \emph{Cognition}, \emph{87}(1), B23--B34. \url{https://doi.org/10.1016/S0010-0277(02)00186-5}
+
+\leavevmode\hypertarget{ref-Hirsh-Pasek_etal_2000}{}%
+Hirsh-Pasek, K., Golinkoff, R.
+M., \& Hollich, G.
+(2000).
+An emergentist coalition model for word learning: Mapping words to objects is a product of the interaction of multiple cues.
+In R.
+M.
+Golinkoff, K.
+Hirsh-Pasek, L.
+Bloom, L.
+B.
+Smith, A.
+L.
+Woodward, N.
+Akhtar, \ldots{} G.
+Hollich (Eds.), \emph{Becoming a word learner: A debate on lexical acquisition} (pp.
+136--164).
+Oxford: Oxford University Press.
 
@@
 
+\leavevmode\hypertarget{ref-Nazzi_Bertoncini_2003}{}%
+Nazzi, T., \& Bertoncini, J.
+(2003).
+Before and after the vocabulary spurt: Two modes of word acquisition?
+\emph{Developmental Science}, \emph{6}(2), 136--142.
+\url{https://doi.org/10.1111/1467-7687.00263}
+
 \leavevmode\hypertarget{ref-Nicoladis_Secco_2000}{}%
@@
 
+\leavevmode\hypertarget{ref-Tsui_etal_2019}{}%
+Tsui, A.
+S.
+M., Byers-Heinlein, K., \& Fennell, C.
+(2019).
+Associative word learning in infancy: A meta-analysis of the switch task.
+\emph{Developmental Psychology}, \emph{55}(5), 934--950.
+\url{https://doi.org/10.1037/dev0000699}
+
 \leavevmode\hypertarget{ref-Umbel_etal_1992}{}%

### 2022-02-22 paper/knit to word/paper_TE_bilingual_vocabulary_model_doc220221.Rmd
doc version updated

--- 2022-02-18 paper/paper_TE_bilingual_vocabulary_model.Rmd
@@
 On average, bilingual children in our sample had a balance score BALANCE of `r balance_average$mean[1]` (SD = `r balance_average$sd[1]`), ranging from `r balance_average$min[1]` to `r balance_average$max[1]`.
-SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
+Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
 The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).

### 2022-03-16 paper/knit to word/paper_TE_bilingual_vocabulary_model_220316.Rmd
fixed typos

--- 2022-02-22 paper/knit to word/paper_TE_bilingual_vocabulary_model_doc220221.Rmd
@@
   - \DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
+  - \DeclareUnicodeCharacter{2212}{\ensuremath{−}}
 
@@
 class             : "man"
-output            : papaja::apa6_word # change to apa6_docx for word document
+output            : papaja::apa6_docx # change to apa6_docx for word document
 citation_package  : biblatex
@@
 
+# 1 Introduction
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
@@
 
-## Accounts of translation equivalent learning
-
-### Avoidance Account: Translation equivalents are harder to learn than singlets
+## 1.1 Accounts of translation equivalent learning
+
+### 1.1.1 Avoidance Account: Translation equivalents are harder to learn than singlets
 Early theories of bilingual development claimed that translation equivalents are conspicuously missing from bilingual children’s early vocabularies [e.g., @Imedadze_1967; @Swain_Wesche_1975; @Volterra_Taeschner_1978].
@@
 
-### Preference Account: Translation equivalents are easier to learn than singlets
+### 1.1.2 Preference Account: Translation equivalents are easier to learn than singlets
 Contrary to the Avoidance Account, the Preference Account posits that translation equivalents are easier to learn than singlets.
@@
 This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English “dog” and French _“chien”_) and words that had other cross-language relations (e.g., “cat” and _“chien”_).
-The authors then simulated bilingual vocabulaires by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals.
+The authors then simulated bilingual vocabularies by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals.
 Comparison with actual bilingual children’s vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
@@
 
-### Neutral Account: Translation equivalents and singlets are learned similarly
+### 1.1.3 Neutral Account: Translation equivalents and singlets are learned similarly
 The previous accounts rely on the idea that bilingual vocabulary development unfolds differently than monolingual development, as monolinguals encounter only singlets but bilinguals encounter both singlets and translation equivalents.
@@
 
-## Contributors to translation equivalent knowledge
+## 1.2 Contributors to translation equivalent knowledge
 The previous section discussed three theoretical accounts concerning the relative learnability of translation equivalents.
@@
 
-### Vocabulary size in bilinguals’ two languages
+### 1.1.1 Vocabulary size in bilinguals’ two languages
 Balance between the two vocabulary sizes is a function of the number of words bilingual children produce in each language, which tends to be tightly linked to their exposure to each language.
@@
 
-### Word learnability as a function of developmental level
+### 1.1.2 Word learnability as a function of developmental level
 An often overlooked factor that could contribute to bilingual children’s learning of translation equivalents is related to the changes in the learnability of different words over time based on children’s developmental level.
@@
 
-## The Bilingual Vocabulary Model
+## 1.3 The Bilingual Vocabulary Model
 Taking into account the contributions of language exposure and developmental level to bilingual children’s vocabulary acquisition, we put forward the Bilingual Vocabulary Model.
@@
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600–700 depending on the language of the adaptation.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development Inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600–700 depending on the language of the adaptation.
 Yet, not all children will be capable of learning all of the CDI words.
@@
   "Relationship to other parameters" = c("Maximum number that could be learned in DOM or NONDOM",
-                                         "DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM",
-                                         "NONDOM = BALANCE×WORD; NONDOM = WORD - DOM",
+                                         "DOM = (1 – BALANCE)×WORD; DOM = WORD – NONDOM",
+                                         "NONDOM = BALANCE×WORD; NONDOM = WORD – DOM",
                                          "",
                                          "BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)",
-                                         "WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)",
+                                         "WORD = DOM+NONDOM; WORD = DOM/(1 – BALANCE); WORD = NONDOM/(BALANCE)",
                                          "TE = BIAS × DOM×NONDOM/LEARNABLE",
-                                         "CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET",
-                                         "DOM-SINGLET = DOM - TE",
-                                         "NONDOM_SINGLET = NONCOM - TE")
+                                         "CONCEPT = WORD – TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET",
+                                         "DOM-SINGLET = DOM – TE",
+                                         "NONDOM-SINGLET = NONCOM – TE")
 )
@@
 
-## Current studies
+## 1.4 Current studies
 The current studies aimed to better understand the nature of translation equivalent learning in bilingual children.
@@
 
-# Study 1: Simulations
+# 2 Study 1: Simulations
 Study 1 provides a computational implementation of the Bilingual Vocabulary Model outlined in the introduction, which we use to simulate different scenarios to examine the effect of vocabulary sizes and developmental variables on translation equivalent learning.
@@
 
-For each simulation, we generated predictions for expected patterns in bilingual children’s vocabulaires.
+For each simulation, we generated predictions for expected patterns in bilingual children’s vocabularies.
 Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David & Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016).
@@
 
-## 1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
+## 2.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary
 In Simulation 1, we examined patterns of translation equivalent learning where simulated children are at the same developmental level and thus have the same number of potentially LEARNABLE words, but their vocabulary size in each language (DOM and NONDOM) varied.
@@
   stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_color_manual(values=c("#FFCE00", "#F89301", "#F76385", "#954698", "#3F3A81")) + 
   scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
@@
   stat_smooth(method = lm, se = F) +
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_color_manual(values=c("#FFCE00", "#F89301", "#F76385", "#954698", "#3F3A81")) + 
   scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) + 
@@
 
-## 1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
+## 2.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels
 In our previous simulation, we assumed that each simulated child was at the same developmental level and had the capacity to learn up to 600 words in each language (i.e., LEARNABLE was held constant at 600).
@@
 Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
-We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM – TE and NONDOM-SINGLET = NONDOM – TE.
 Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
@@
   stat_smooth(method = lm, se = F) + 
-  scale_color_manual(values=c("#bdbdbd", "#737373", "#525252", "#252525", "#000000")) + 
+  scale_color_manual(values=c("#FFCE00", "#F89301", "#F76385", "#954698", "#3F3A81")) + 
   scale_linetype_manual(values=c("longdash", "dotdash", "dashed", "twodash", "solid")) +
@@
 
-## 1.3 Simulation 3: Bias towards or against translation equivalent learning
+## 2.3 Simulation 3: Bias towards or against translation equivalent learning
 In Simulations 1 and 2, we modeled cases with the bias parameter held constant at 1 and translation equivalent (TE) was calculated as DOM×NONDOM/LEARNABLE.
@@
 
-## 1.4 Summary of model predictions
+## 2.4 Summary of model predictions
 Study 1 generated predictions under the Bilingual Vocabulary Model with regards to the effect of vocabulary sizes and developmental variables on translation equivalent learning.
@@
 
-# Study 2: Empirical data
+# 3 Study 2: Empirical data
 In Study 1, we used simulations based on the Bilingual Vocabulary Model to generate several predictions about the relationship between translation equivalent knowledge and other vocabulary variables.
@@
 
-## 2.1 Method
+## 3.1 Method
 Ethics approval was obtained by the Human Research Ethics Board of Concordia University (Certificate Number 10000439) and informed consent was obtained from the children’s parents.
 
-## 2.1.1 Participants
+### 3.1.1 Participants
 ```{r participants, include = FALSE, cache = TRUE}
@@
 
-## 2.1.2 Measures
+## 3.1.2 Measures
 
@@
 
-### MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)
+### 3.1.2.1 MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)
 Bilingual children’s expressive vocabulary was measured by the Words and Sentences form of the MacArthur-Bates CDI.
@@
 
-After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
-Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).
-
-### Language Exposure Questionnaire using the MAPLE approach
+After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM – TE and NONDOM – TE as in Study 1).
+Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD – TE as in Study 1).
+
+### 3.1.2.2 Language Exposure Questionnaire using the MAPLE approach
 Children’s language exposure was measured using the Language Exposure Questionnaire (LEQ; Bosch & Sebastián-Gallés, 2001) and the Multilingual Approach to Parent Language Estimates (MAPLE; Byers-Heinlein et al., 2020).
@@
 
-## 2.1.3 Procedure
+## 3.1.3 Procedure
 Caregivers were asked to fill out the CDI questionnaires as part of their child’s participation in experimental studies on language and cognitive development.
@@
 
-## 2.2 Results
+## 3.2 Results
 Data analyses were conducted using R (Version 4.0.2, 2020).
@@
 
-## 2.2.1 Descriptive measures of vocabulary
+## 3.2.1 Descriptive measures of vocabulary
 
@@
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997).
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Hurtado et al., 2014; Pearson et al., 1997).
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, _r_(`r correlation_balance$parameter[1]`) = `r round(correlation_balance$estimate[1], 2)`, $p `r papaja::printp(correlation_balance$p.value[1])`$ (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2022, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 
-## 2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures
+## 3.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures
 Prediction Set 1 pertained to the pairwise relationships between translation equivalents (TE), word vocabulary (WORD), dominant (DOM) and non-dominant vocabulary (NONDOM), and vocabulary balance (BALANCE), which we examined through Pearson’s correlations.
@@
 
-## 2.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels
+## 3.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels
 Prediction Set 2 pertained to expected patterns of acquisition of translation equivalents and singlets for children of different developmental levels.
@@
 
-## 2.2.4 Testing Prediction Set 3: Rate of translation equivalent learning
+## 3.2.4 Testing Prediction Set 3: Rate of translation equivalent learning
 Prediction Set 3 pertained to the overall nature of translation equivalent learning, describing expected patterns of translation equivalent learning under the Neutral Account, the Avoidance Account, and the Preference Account.
@@
 Note that visual inspection suggested that there could be some possible outliers.
-Cook’s distance was estimated for our linear regression model described above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a cook’s distance over 0.4.
+Cook’s distance was estimated for our linear regression model described above and identified `r words(length(number_of_outliers$ID_testdate))` data points with a Cook’s distance over 0.4.
 After removing those `r words(length(number_of_outliers$ID_testdate))` data points, the linear regression model returned a coefficient of `r round(summary(model_remove_outliers)$coefficient[1],2)`, with $R^2$ = `r round(summary(model_remove_outliers)$r.squared,2)`.
@@
   theme(plot.title = element_text(hjust = 0.5)) +
-  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x, color = "#3236a6", alpha = 0.75, size = 2) +
+  geom_point(shape = 23, color = "#3236a6", alpha = 0.50) +
+  geom_abline(linetype = 2) +
   xlim(-10,450) +
@@
   xlim(0,80) +
-  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x, color = "#3236a6", alpha = 0.75) +
+  geom_point(shape = 23, color = "#3236a6", alpha = 0.50) +
+  geom_abline(linetype = 2) +
   theme_minimal() +
@@
        title = "More than 300 word vocabulary (WORD)") +
-  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x) +
-  geom_point(shape = 23) +
-  geom_abline(linetype = 2, alpha = 0.5) +
+  geom_smooth(method = "lm", se = F, formula = y ~ 0 + x, color = "#3236a6", alpha = 0.75) +
+  geom_point(shape = 23, color = "#3236a6", alpha = 0.50) +
+  geom_abline(linetype = 2) +
   theme_minimal() +
@@
 
-# Discussion
+# 4 Discussion
 The aim of the current study was to better understand translation equivalent learning in bilingual children.
@@
 
-## Understanding patterns in bilingual children’s vocabulary development
-
-Translation equivalent learning has been of interest in studies of bilingual children’s development from the very earliest descriptive work (Ronjat, 2013), to more recent quantitative research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+## 4.1 Understanding patterns in bilingual children’s vocabulary development
+
+Translation equivalent learning has been of interest in studies of bilingual children’s development from the very earliest descriptive work (Ronjat, 1913), to more recent quantitative research (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
 Many previous studies have examined univariate relations between translation equivalent knowledge and predictors such as age or vocabulary balance (David & Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), but such observations lacked a unifying framework.
@@
 Translation equivalent production is predicted from four parameters: vocabulary size in the dominant language, vocabulary size in the non-dominant language, the number of potentially learnable words in each language given a child’s developmental level, and a bias parameter indicating whether translation equivalent learning is harder than, easier than, or similar to singlet learning.
-Translation equivalent production can also be reexpressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
+Translation equivalent production can also be re-expressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
 
@@
 
-## The status of translation equivalents in bilingual vocabulary development 
+## 4.2 The status of translation equivalents in bilingual vocabulary development 
 Applying the Bilingual Vocabulary Model to a large set of empirical vocabulary data from bilingual children allowed us to quantitatively test three conceptual accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
@@
 It may be that establishing a conceptual representation is somewhat easy, and thus any facilitation effects of translation equivalents might be relatively small.
-Such developmental changes would be consistent with work showing that children’s word learning abilities develop substantially across the first years of life (Bergelson, 2021; Hirsh-Pasek & Golinkoff, 2000; Nazzi & Bertoncini, 2003; Tsui et al., 2019).
+Such developmental changes would be consistent with work showing that children’s word learning abilities develop substantially across the first years of life (Bergelson, 2020; Hirsh-Pasek & Golinkoff, 2000; Nazzi & Bertoncini, 2003; Tsui et al., 2019).
 
@@
 
-## Assumptions, limitations, and future directions
+## 4.3 Assumptions, limitations, and future directions
 Our Bilingual Vocabulary Model presented an integrated computational account of translation equivalent learning, focusing on the joint probability of learning the word for a concept in each language.
@@
 This represents an important methodological advance compared to previous approaches, which used data from randomly-paired monolinguals or the vocabularies of two different bilinguals as a reference point (e.g., Bilson et al., 2015; Pearson et al., 1995).
-However, our approach does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our approach does not directly model individual difference factors such as children’s ability to segment words from the continuous stream of speech (e.g., Brent & Siskind, 2001; Swingley & Humphrey, 2018), children’s efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder & Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents’ interaction with their children (e.g., Blewitt et al., 2009; Yu & Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 
-## Conclusions
-In sum, the acquisition of translation equivalents has been considered a special component in bilingual children’ vocabulary development.
+# 5 Conclusions
+In sum, the acquisition of translation equivalents has been considered a special component in bilingual children’s vocabulary development.
 Previous research has put forward three diverging accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
@@
 This work was funded by a grant from the Natural Sciences and Engineering Research Council of Canada (NSERC 2011-402470, 2018-04390), a National Institutes of Health grant (NIH R03HD079779), and a Fonds de Recherche du Québec-Société et Culture grant (FQRSC 2012-NP-145009) to Byers-Heinlein, by a FQRSC (2018-B3-205717) postdoctoral fellowship to Gonzalez-Barrero, and by a FQRSC graduate fellowship and Concordia University graduate fellowship to Schott.
-Byers-Heinlen holds the Concordia University Research Chair in Bilingualism and Open Science.
+Byers-Heinlein holds the Concordia University Research Chair in Bilingualism and Open Science.
 This work was presented at the 2021 Boston University Conference on Language Development.
@@
 **Esther Schott:** Conceptualization, Data Curation, Writing - Review & Editing.
-**Krista Byers-Heinlein:** Conceptualization, Funding Acquisition, Methodology, Resources, Supervision, Writing - Original draft preparation, Writing - Review & Editing, Supervision, Project administration.
+**Krista Byers-Heinlein:** Conceptualization, Funding Acquisition, Methodology, Resources, Writing - Original draft preparation, Writing - Review & Editing, Supervision, Project administration.
 
@@
 Calculation <- c("", "", "", "", 
-                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM × NONDOM / LEARNABLE", "WORD - TE", "DOM - TE", "NONDOM - TE")
+                 "DOM + NONDOM", "NONDOM / (DOM + NONDOM)", "DOM × NONDOM / LEARNABLE", "WORD – TE", "DOM – TE", "NONDOM – TE")
 

### 2022-03-16 paper/paper_TE_bilingual_vocabulary_model.Rmd
fixed typos

@@
 On average, bilingual children in our sample had a balance score BALANCE of `r balance_average$mean[1]` (SD = `r balance_average$sd[1]`), ranging from `r balance_average$min[1]` to `r balance_average$max[1]`.
-SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
+Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, _t_(`r round(t_test_balance$parameter[1],2)`) = `r round(t_test_balance$statistic[1],2)`, $p = `r papaja::printp(t_test_balance$p.value[1])`$, $d = `r round(d_balance, 2)`$.
 The `r round(d_lang_dom$percent_EngDom, 1)`% of children who were English-dominant had an average BALANCE of `r round(balance_average_group$mean[1], 2)` (SD = `r round(balance_average_group$sd[1], 2)`, range = `r round(balance_average_group$min[1], 2)` - `r round(balance_average_group$max[1], 2)`) whereas the remaining `r round(d_lang_dom$percent_FrDom, 1)`% who were French-dominant had an average BALANCE of `r round(balance_average_group$mean[2], 2)` (SD = `r round(balance_average_group$sd[2], 2)`, range = `r round(balance_average_group$min[2], 2)` - `r round(balance_average_group$max[2], 2)`).
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997).
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children’s relative input in each language (e.g., David & Wei, 2008; Hurtado et al., 2014; Pearson et al., 1997).
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.

### 2022-03-16 paper/paper_TE_bilingual_vocabulary_model.tex
fixed typos

@@
 \DeclareUnicodeCharacter{2264}{\ensuremath{\le}}
+\DeclareUnicodeCharacter{2212}{\ensuremath{−}}
 \ifxetex
@@
 
+\hypertarget{introduction}{%
+\section{1 Introduction}\label{introduction}}
+
 Bilingual children must learn words that take a different form in each of their languages, but share the same or highly similar meanings.
@@
 \hypertarget{accounts-of-translation-equivalent-learning}{%
-\subsection{Accounts of translation equivalent learning}\label{accounts-of-translation-equivalent-learning}}
+\subsection{1.1 Accounts of translation equivalent learning}\label{accounts-of-translation-equivalent-learning}}
 
 \hypertarget{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}{%
-\subsubsection{Avoidance Account: Translation equivalents are harder to learn than singlets}\label{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}}
+\subsubsection{1.1.1 Avoidance Account: Translation equivalents are harder to learn than singlets}\label{avoidance-account-translation-equivalents-are-harder-to-learn-than-singlets}}
 
@@
 \hypertarget{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}{%
-\subsubsection{Preference Account: Translation equivalents are easier to learn than singlets}\label{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}}
+\subsubsection{1.1.2 Preference Account: Translation equivalents are easier to learn than singlets}\label{preference-account-translation-equivalents-are-easier-to-learn-than-singlets}}
 
@@
 This effect applied not only to words within the same language, but also to words across languages including translation equivalents (e.g., English ``dog'' and French \emph{``chien''}) and words that had other cross-language relations (e.g., ``cat'' and \emph{``chien''}).
-The authors then simulated bilingual vocabulaires by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals.
+The authors then simulated bilingual vocabularies by modeling bilingual lexicons as combinations of two independent vocabulary-size-matched monolinguals.
 Comparison with actual bilingual children's vocabulary data revealed that bilingual children acquired more translation equivalents than predicted by the simulation.
@@
 \hypertarget{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}{%
-\subsubsection{Neutral Account: Translation equivalents and singlets are learned similarly}\label{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}}
+\subsubsection{1.1.3 Neutral Account: Translation equivalents and singlets are learned similarly}\label{neutral-account-translation-equivalents-and-singlets-are-learned-similarly}}
 
@@
 \hypertarget{contributors-to-translation-equivalent-knowledge}{%
-\subsection{Contributors to translation equivalent knowledge}\label{contributors-to-translation-equivalent-knowledge}}
+\subsection{1.2 Contributors to translation equivalent knowledge}\label{contributors-to-translation-equivalent-knowledge}}
 
@@
 \hypertarget{vocabulary-size-in-bilinguals-two-languages}{%
-\subsubsection{Vocabulary size in bilinguals' two languages}\label{vocabulary-size-in-bilinguals-two-languages}}
+\subsubsection{1.1.1 Vocabulary size in bilinguals' two languages}\label{vocabulary-size-in-bilinguals-two-languages}}
 
@@
 \hypertarget{word-learnability-as-a-function-of-developmental-level}{%
-\subsubsection{Word learnability as a function of developmental level}\label{word-learnability-as-a-function-of-developmental-level}}
+\subsubsection{1.1.2 Word learnability as a function of developmental level}\label{word-learnability-as-a-function-of-developmental-level}}
 
@@
 \hypertarget{the-bilingual-vocabulary-model}{%
-\subsection{The Bilingual Vocabulary Model}\label{the-bilingual-vocabulary-model}}
+\subsection{1.3 The Bilingual Vocabulary Model}\label{the-bilingual-vocabulary-model}}
 
@@
 Next, we turn to the LEARNABLE parameter (i.e., the number of potentially learnable words).
-If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600--700 depending on the language of the adaptation.
+If DOM and NONDOM are measured with an instrument such as the MacArthur-Bates Communicative Development Inventories (CDI; Fenson et al., 2007), one option would be to set LEARNABLE to be the total number of items on the CDI, which is typically 600--700 depending on the language of the adaptation.
 Yet, not all children will be capable of learning all of the CDI words.
@@
 & Maximum number that could be learned in DOM or NONDOM\\
-\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (children always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1-BALANCE)×WORD; DOM = WORD - NONDOM\\
-\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (children always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD - DOM\\
+\hspace{1em}DOM & Words produced in the dominant language & DOM ≥ NONDOM (children always produce more words in dominant than non-dominant language); DOM ≤ LEARNABLE & DOM = (1 – BALANCE)×WORD; DOM = WORD – NONDOM\\
+\hspace{1em}NONDOM & Words produced in the non-dominant language & NONDOM ≤ DOM (children always produce fewer words in non-dominant than dominant language); NONDOM ≤ LEARNABLE & NONDOM = BALANCE×WORD; NONDOM = WORD – DOM\\
 \hspace{1em}BIAS & Bias parameter & BIAS < 1 implies the Avoidance Account; BIAS > 1 implies the Preference Account; BIAS = 1 implies the Neutral Account & \\
@@
 \hspace{1em}BALANCE & Balance (relative proportion of words produced in the non-dominant language to the total words produced in both languages) & 0 ≤ BALANCE ≤ .50 (greater values indicate children producing a more similar number of words in their two languages) & BALANCE = NONDOM/WORD; BALANCE = NONDOM/(DOM+NONDOM)\\
-\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1-BALANCE); WORD = NONDOM/(BALANCE)\\
+\hspace{1em}WORD & Word vocabulary (or total vocabulary size) & W ≤ 2×LEARNABLE (maximum word vocabulary is knowing each word in both languages) & WORD = DOM+NONDOM; WORD = DOM/(1 – BALANCE); WORD = NONDOM/(BALANCE)\\
 \hspace{1em}TE & Translation equivalents produced &  & TE = BIAS × DOM×NONDOM/LEARNABLE\\
-\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD - TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
-\hspace{1em}DOM-SINGLET & Singlets in dominant language &  & DOM-SINGLET = DOM - TE\\
-\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language &  & NONDOM\_SINGLET = NONCOM - TE\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) &  & CONCEPT = WORD – TE; CONCEPT = TE + DOM-SINGLET + NONDOM-SINGLET\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language &  & DOM-SINGLET = DOM – TE\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language &  & NONDOM-SINGLET = NONCOM – TE\\
 \bottomrule
@@
 \hypertarget{current-studies}{%
-\subsection{Current studies}\label{current-studies}}
+\subsection{1.4 Current studies}\label{current-studies}}
 
@@
 \hypertarget{study-1-simulations}{%
-\section{Study 1: Simulations}\label{study-1-simulations}}
+\section{2 Study 1: Simulations}\label{study-1-simulations}}
 
@@
 
-For each simulation, we generated predictions for expected patterns in bilingual children's vocabulaires.
+For each simulation, we generated predictions for expected patterns in bilingual children's vocabularies.
 Several previous studies in the literature have examined relations between translation equivalent knowledge and our main parameters including age or developmental level (which determines LEARNABLE; David \& Wei, 2008; Legacy et al., 2016), DOM (Legacy et al., 2016; Poulin-Dubois et al., 2013), and NON-DOM (Legacy et al., 2016).
@@
 \hypertarget{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}{%
-\subsection{1.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}
+\subsection{2.1 Simulation 1: Children of the same developmental level with different word vocabularies and balances of vocabulary}\label{simulation-1-children-of-the-same-developmental-level-with-different-word-vocabularies-and-balances-of-vocabulary}}
 
@@
 \hypertarget{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}{%
-\subsection{1.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}
+\subsection{2.2 Simulation 2: Acquisition of translation equivalents and singlets at different developmental levels}\label{simulation-2-acquisition-of-translation-equivalents-and-singlets-at-different-developmental-levels}}
 
@@
 Again, balance (BALANCE) was calculated based on the values of DOM and NONDOM.
-We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM - TE and NONDOM-SINGLET = NONDOM - TE.
+We also calculated the number of singlet words in the dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by the formula DOM-SINGLET = DOM -- TE and NONDOM-SINGLET = NONDOM -- TE.
 Figure 2 plots the number of CONCEPT, TE, DOM-SINGLET, and NONDOM-SINGLET at different BALANCE levels across different levels of LEARNABLE.
@@
 \hypertarget{simulation-3-bias-towards-or-against-translation-equivalent-learning}{%
-\subsection{1.3 Simulation 3: Bias towards or against translation equivalent learning}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning}}
+\subsection{2.3 Simulation 3: Bias towards or against translation equivalent learning}\label{simulation-3-bias-towards-or-against-translation-equivalent-learning}}
 
@@
 \hypertarget{summary-of-model-predictions}{%
-\subsection{1.4 Summary of model predictions}\label{summary-of-model-predictions}}
+\subsection{2.4 Summary of model predictions}\label{summary-of-model-predictions}}
 
@@
 \hypertarget{study-2-empirical-data}{%
-\section{Study 2: Empirical data}\label{study-2-empirical-data}}
+\section{3 Study 2: Empirical data}\label{study-2-empirical-data}}
 
@@
 \hypertarget{method}{%
-\subsection{2.1 Method}\label{method}}
+\subsection{3.1 Method}\label{method}}
 
@@
 \hypertarget{participants}{%
-\subsection{2.1.1 Participants}\label{participants}}
+\subsubsection{3.1.1 Participants}\label{participants}}
 
@@
 \hypertarget{measures}{%
-\subsection{2.1.2 Measures}\label{measures}}
+\subsection{3.1.2 Measures}\label{measures}}
 
 \hypertarget{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}{%
-\subsubsection{MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)}\label{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}}
+\subsubsection{3.1.2.1 MacArthur-Bates Communicative Development Inventories: Words and Sentences (CDI)}\label{macarthur-bates-communicative-development-inventories-words-and-sentences-cdi}}
 
@@
 
-After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM - TE and NONDOM - TE as in Study 1).
-Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD - TE as in Study 1).
+After determining the dominant language of a child based on relative vocabulary size, we then computed the number of singlets that children knew in their dominant (DOM-SINGLET) and non-dominant (NONDOM-SINGLET) languages by deducting the number of translation equivalents produced from the total number of words produced in each language (i.e., DOM -- TE and NONDOM -- TE as in Study 1).
+Concept vocabulary (CONCEPT) was computed based on the number of concepts for which a child produced a word, calculated by subtracting the number of translation equivalents from word vocabulary (i.e., WORD -- TE as in Study 1).
 
 \hypertarget{language-exposure-questionnaire-using-the-maple-approach}{%
-\subsubsection{Language Exposure Questionnaire using the MAPLE approach}\label{language-exposure-questionnaire-using-the-maple-approach}}
+\subsubsection{3.1.2.2 Language Exposure Questionnaire using the MAPLE approach}\label{language-exposure-questionnaire-using-the-maple-approach}}
 
@@
 \hypertarget{procedure}{%
-\subsection{2.1.3 Procedure}\label{procedure}}
+\subsection{3.1.3 Procedure}\label{procedure}}
 
@@
 \hypertarget{results}{%
-\subsection{2.2 Results}\label{results}}
+\subsection{3.2 Results}\label{results}}
 
@@
 \hypertarget{descriptive-measures-of-vocabulary}{%
-\subsection{2.2.1 Descriptive measures of vocabulary}\label{descriptive-measures-of-vocabulary}}
+\subsection{3.2.1 Descriptive measures of vocabulary}\label{descriptive-measures-of-vocabulary}}
 
@@
 On average, bilingual children in our sample had a balance score BALANCE of 0.31 (SD = 0.13), ranging from 0.02 to 0.50.
-SSimilar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
+Similar vocabulary balance was found between the children who were English-dominant and those who were French-dominant, \emph{t}(200.43) = 0.57, \(p = .566\), \(d = 0.08\).
 The 59.80\% of children who were English-dominant had an average BALANCE of 0.31 (SD = 0.13, range = 0.02 - 0.50) whereas the remaining 40.20\% who were French-dominant had an average BALANCE of 0.30 (SD = 0.12, range = 0.05 - 0.50).
 
-Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Hurtado et al., 2013; Pearson et al., 1997).
+Note that in this paper, we defined BALANCE in terms of relative vocabulary in each language, which numerous studies have found is related to children's relative input in each language (e.g., David \& Wei, 2008; Hurtado et al., 2014; Pearson et al., 1997).
 We therefore investigated the relationship between vocabulary balance and the percent of exposure bilingual children received in their non-dominant language.
@@
 The correlation between BALANCE and the percentage of exposure to the non-dominant language was moderate, \emph{r}(227) = 0.45, \(p < .001\) (see also Figure 4).
-The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2022, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
+The imperfect correlation between these two measures could result from measurement error (see Byers-Heinlein et al., 2021, for a discussion of attenuation of correlation due to measurement error in developmental research) or could also be because vocabulary size in each language is determined by factors beyond parent-reported proportion of input in each language, for example the quality of that input (for a discussion of input quality for bilinguals, see Unsworth, 2016).
 
@@
 \hypertarget{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}{%
-\subsection{2.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures}\label{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}}
+\subsection{3.2.2 Testing Prediction Set 1: Univariate relationships between translation equivalents and different vocabulary measures}\label{testing-prediction-set-1-univariate-relationships-between-translation-equivalents-and-different-vocabulary-measures}}
 
@@
 \hypertarget{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}{%
-\subsection{2.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels}\label{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}}
+\subsection{3.2.3 Testing Prediction Set 2: The vocabulary composition of bilingual children at different developmental levels}\label{testing-prediction-set-2-the-vocabulary-composition-of-bilingual-children-at-different-developmental-levels}}
 
@@
 \hypertarget{testing-prediction-set-3-rate-of-translation-equivalent-learning}{%
-\subsection{2.2.4 Testing Prediction Set 3: Rate of translation equivalent learning}\label{testing-prediction-set-3-rate-of-translation-equivalent-learning}}
+\subsection{3.2.4 Testing Prediction Set 3: Rate of translation equivalent learning}\label{testing-prediction-set-3-rate-of-translation-equivalent-learning}}
 
@@
 Note that visual inspection suggested that there could be some possible outliers.
-Cook's distance was estimated for our linear regression model described above and identified two data points with a cook's distance over 0.4.
+Cook's distance was estimated for our linear regression model described above and identified two data points with a Cook's distance over 0.4.
 After removing those two data points, the linear regression model returned a coefficient of 1.07, with \(R^2\) = 0.96.
@@
 \hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
+\section{4 Discussion}\label{discussion}}
 
@@
 \hypertarget{understanding-patterns-in-bilingual-childrens-vocabulary-development}{%
-\subsection{Understanding patterns in bilingual children's vocabulary development}\label{understanding-patterns-in-bilingual-childrens-vocabulary-development}}
-
-Translation equivalent learning has been of interest in studies of bilingual children's development from the very earliest descriptive work (Ronjat, 2013), to more recent quantitative research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
+\subsection{4.1 Understanding patterns in bilingual children's vocabulary development}\label{understanding-patterns-in-bilingual-childrens-vocabulary-development}}
+
+Translation equivalent learning has been of interest in studies of bilingual children's development from the very earliest descriptive work (Ronjat, 1913), to more recent quantitative research (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997).
 Many previous studies have examined univariate relations between translation equivalent knowledge and predictors such as age or vocabulary balance (David \& Wei, 2008; Legacy et al., 2016; Montanari, 2010; Pearson et al., 1995; 1997), but such observations lacked a unifying framework.
@@
 Translation equivalent production is predicted from four parameters: vocabulary size in the dominant language, vocabulary size in the non-dominant language, the number of potentially learnable words in each language given a child's developmental level, and a bias parameter indicating whether translation equivalent learning is harder than, easier than, or similar to singlet learning.
-Translation equivalent production can also be reexpressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
+Translation equivalent production can also be re-expressed via other derived variables such as vocabulary balance, word vocabulary, concept vocabulary, and singlet vocabulary (i.e., words for which the child does not yet produce a translation equivalent), which can be calculated algebraically from the original parameters (e.g., word vocabulary = dominant vocabulary + non-dominant vocabulary).
 
@@
 \hypertarget{the-status-of-translation-equivalents-in-bilingual-vocabulary-development}{%
-\subsection{The status of translation equivalents in bilingual vocabulary development}\label{the-status-of-translation-equivalents-in-bilingual-vocabulary-development}}
+\subsection{4.2 The status of translation equivalents in bilingual vocabulary development}\label{the-status-of-translation-equivalents-in-bilingual-vocabulary-development}}
 
@@
 It may be that establishing a conceptual representation is somewhat easy, and thus any facilitation effects of translation equivalents might be relatively small.
-Such developmental changes would be consistent with work showing that children's word learning abilities develop substantially across the first years of life (Bergelson, 2021; Hirsh-Pasek \& Golinkoff, 2000; Nazzi \& Bertoncini, 2003; Tsui et al., 2019).
+Such developmental changes would be consistent with work showing that children's word learning abilities develop substantially across the first years of life (Bergelson, 2020; Hirsh-Pasek \& Golinkoff, 2000; Nazzi \& Bertoncini, 2003; Tsui et al., 2019).
 
@@
 \hypertarget{assumptions-limitations-and-future-directions}{%
-\subsection{Assumptions, limitations, and future directions}\label{assumptions-limitations-and-future-directions}}
+\subsection{4.3 Assumptions, limitations, and future directions}\label{assumptions-limitations-and-future-directions}}
 
@@
 This represents an important methodological advance compared to previous approaches, which used data from randomly-paired monolinguals or the vocabularies of two different bilinguals as a reference point (e.g., Bilson et al., 2015; Pearson et al., 1995).
-However, our approach does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2013; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
+However, our approach does not directly model individual difference factors such as children's ability to segment words from the continuous stream of speech (e.g., Brent \& Siskind, 2001; Swingley \& Humphrey, 2018), children's efficiency of processing words they hear (e.g., Hurtado et al., 2014; Weisleder \& Fernald, 2013), and cognitive development and perceptual bias (e.g., Benedict, 1979; Goodman et al., 2008), nor does it consider qualitative factors including family socioeconomic status (e.g., Fernald et al., 2013; Hoff, 2003), parents' interaction with their children (e.g., Blewitt et al., 2009; Yu \& Smith, 2012), and the quality of parental language input over time (e.g., Raneri et al., 2020; Rowe, 2012).
 It would be interesting for future studies to take into consideration such factors in a bilingual word learning model, including different amounts of input and the quality of that input.
@@
 \hypertarget{conclusions}{%
-\subsection{Conclusions}\label{conclusions}}
-
-In sum, the acquisition of translation equivalents has been considered a special component in bilingual children' vocabulary development.
+\section{5 Conclusions}\label{conclusions}}
+
+In sum, the acquisition of translation equivalents has been considered a special component in bilingual children's vocabulary development.
 Previous research has put forward three diverging accounts of translation equivalent learning: the Avoidance Account, the Preference Account, and the Neutral Account.
@@
 \textbf{Esther Schott:} Conceptualization, Data Curation, Writing - Review \& Editing.
-\textbf{Krista Byers-Heinlein:} Conceptualization, Funding Acquisition, Methodology, Resources, Supervision, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
+\textbf{Krista Byers-Heinlein:} Conceptualization, Funding Acquisition, Methodology, Resources, Writing - Original draft preparation, Writing - Review \& Editing, Supervision, Project administration.
 
@@
 \hspace{1em}TE & Translation equivalents produced & DOM × NONDOM / LEARNABLE & 14 & 36 & 324\\
-\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD - TE & 286 & 264 & 576\\
-\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM - TE & 256 & 144 & 216\\
-\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM - TE & 16 & 84 & 36\\
+\hspace{1em}CONCEPT & Concept vocabulary (or total conceptual vocabulary size) & WORD – TE & 286 & 264 & 576\\
+\hspace{1em}DOM-SINGLET & Singlets in dominant language & DOM – TE & 256 & 144 & 216\\
+\hspace{1em}NONDOM-SINGLET & Singlets in non-dominant language & NONDOM – TE & 16 & 84 & 36\\
 \bottomrule
@@
 
-\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2022}{}%
-Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (2022). Six solutions for more reliable infant research. \emph{Infant and Child Development}. \url{https://doi.org/10.1002/icd.2296}
+\leavevmode\hypertarget{ref-Byers-Heinlein_etal_2021}{}%
+Byers-Heinlein, K., Bergmann, C., \& Savalei, V. (2021). Six solutions for more reliable infant research. \emph{Infant and Child Development}, e2296. \url{https://doi.org/10.1002/icd.2296}
 
@@
 
+\leavevmode\hypertarget{ref-Ronjat_1913}{}%
+Ronjat, J. (1913). \emph{Le développement du langage observé chez un enfant bilingue}. Paris: Champion.
+
 \leavevmode\hypertarget{ref-Rowe_2012}{}%
@@
 (1999).
-Une adaptation en français québecois du MacArthur {C}ommunicative {D}evelopment {I}nventory {[}a {Q}uebec {F}rench adaptation of the MacArthur {C}ommunicative {D}evelopment {I}nventory{]}.
-\emph{Revue d'orthophonie Et d'audiologie}, \emph{23}, 31--73.
+Une adaptation en français {Q}uébécois du {M}acArthur {C}ommunicative {D}evelopment {I}nventory.
+\emph{La Revue d'orthophonie Et d'audiologie}, \emph{23}, 61--73.
 
